### 2014-01-22 presentations/popl-2013/jgross-student-talk.tex
Add some presentations

\documentclass[utf8x]{beamer}

\usepackage{beamerthemesplit}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{hyperref}
\usepackage{pgf,tikz}
\usetikzlibrary{arrows}
\usepackage[all]{xy}

\makeatletter
\AtBeginDocument{% example environment modified from the example environment in lshort.sty, using the verbatim package
  \newwrite\examplesx@out
  \newenvironment{examplesx}{%
    \begingroup% Lets Keep the Changes Local
      \@bsphack
      \immediate\openout \examplesx@out \jobname.exa
      \let\do\@makeother\dospecials\catcode`\^^M\active
      \def\verbatim@processline{%
        \immediate\write\examplesx@out{\the\verbatim@line}}%
      \verbatim@start
    }{%
    \immediate\closeout\examplesx@out\@esphack\endgroup%
    \noindent\makebox[\textwidth][l]{%
      \begin{minipage}[c]{0.45\textwidth}%
        \small\verbatiminput{\jobname.exa}
      \end{minipage}%
      \hspace*{0.1\textwidth}%
      \framebox{%
        \begin{minipage}{0.45\textwidth}%
          \small\input{\jobname.exa}%
        \end{minipage}
      }%
    }\vspace*{\parskip}%
  }
}
\makeatother

\pdfinfo {
  /Title (Database Management on top of Category Theory in Coq)
  /Subject (POPL)
  /Author (Jason Gross)
  /Keywords (POPL, Coq, Category Theory, Databases)
}

\begin{document}

\title[Database Management on top of Category Theory in Coq]{Building Database Management on top of Category Theory in Coq}
\date{POPL 2013}
\author{Jason Gross --- \href{mailto:jgross@mit.edu}{jgross@mit.edu}}

\begin{frame}
  \titlepage
  This document is available at \url{http://web.mit.edu/jgross/Public/POPL/\jobname.pdf}.

  My category theory library is available at \url{https://bitbucket.org/JasonGross/catdb}.
\end{frame}

\section*{Outline}
  \begin{frame}{Outline}
    \tableofcontents[pausesections]
  \end{frame}

\section{Introduction --- Databases and Category Theory}    
  \subsection{Categories}
    \begin{frame}{Categories}
      A category is:
      \begin{itemize}
        \item
          a collection of objects, \pause together with
        \item
          arrows between those objects, \pause together with
        \item
          a composition law for the arrows satisfying coherence conditions:
          \begin{itemize}
            \item existence of identity
            \item associativity
          \end{itemize} \pause
      \end{itemize}
      \begin{center}
        \fbox{
          $$
          \xymatrix{
            a \ar@/^{0.5pc}/[rr]^{f} \ar@/_{0.5pc}/[d]_{g} \ar@/^{0.5pc}/[d]^{h} && b \ar@/^{0.5pc}/[ll]^{i} \ar@/^{1.5pc}/[dll]^{j} \\
            c
          }
          $$
        }
      \end{center}
    \end{frame}

  \subsection{Relational Databases}
    \begin{frame}{Relational Databases}
      A database schema for a relational database can be modeled as a
      \begin{itemize}
        \item collection of \textbf{tables}, \pause together with 
        \item a collection of \textbf{attributes} or column-labels for each table, \pause together with
        \item integrity constraints \pause
      \end{itemize}
      \begin{center}
        \fbox{
          $$
          \xymatrix{
            *\txt{professor} \ar@/^{1pc}/[rr]^{\text{department}} \ar@/_{1pc}/[dd]_{\text{first name}} \ar@/^{1pc}/[dd]^{\text{last name}} && *\txt{department} \ar@/^{1pc}/[ll]^{\text{department head}} \ar@/^{2pc}/[ddll]^{\text{name}} \\ \\
            *\txt{string} &&
          }
          $$
          }
      \end{center}
    \end{frame}

  \subsection{Relational Database Schema = Category}
    \begin{frame}{Relational Database Schema = Category}
      \begin{center}
        \only<1,3>{
          \fbox{
          $$
          \xymatrix{
            a \ar@/^{0.5pc}/[rr]^{f} \ar@/_{0.5pc}/[d]_{g} \ar@/^{0.5pc}/[d]^{h} && b \ar@/^{0.5pc}/[ll]^{i} \ar@/^{1.5pc}/[dll]^{j} \\
            c
          }
          $$
          }
        }
        \only<3>{\raisebox{-3.5ex}[\height][\depth]{$\cong$}}
        \only<2,3>{
          \fbox{\only<3>{\small}
          $$
          \xymatrix{
            *\txt{professor} \ar@/^{1pc}/[rr]^{\text{department}} \ar@/_{1pc}/[dd]_{\text{f.~name}} \ar@/^{1pc}/[dd]^{\text{l.~name}} && *\txt{department} \ar@/^{1pc}/[ll]^{\text{department head}} \ar@/^{2pc}/[ddll]^{\text{name}} \\ \\
            *\txt{string} &&
          }
          $$
          }
        }
        \only<3>{\vspace{4ex} \\ The diagrams are ``the same''.}
      \end{center}
    \end{frame}

  \subsection{Usefulness}
    \begin{frame}{Usefulness of Categorical Databases}
      \begin{itemize}
        \item
          Built in notion of path equivalence (multiple equivalent paths of foreign keys can be a pain in typical database management). \pause
        \item
          Provides a rigorous language for data migration between databases (another hard task in standard database management).
      \end{itemize}
    \end{frame}

\section{Category Theory in Coq}

  \begin{frame}{Category Theory in Coq}
    \begin{itemize}
      \item
        Many people learn a proof assistant by coding up category theory. \pause
      \item
        Category theory is relatively simple to code up. \pause
        \begin{itemize}
          \item
            Standard rigorous formulation of concepts exists in the literature. \pause
          \item
            It's rare to get caught up in minute details of proofs. \pause
          \item
            If you can define something categorically, it's probably interesting.
        \end{itemize}
    \end{itemize}
  \end{frame}

  \subsection{Universe Levels}
    \begin{frame}
      \frametitle{Universe Levels \only<1-7>{(Russel's Paradox)}}
      \begin{itemize}
        \item<1-7>
          Consider, na\"ively, the set of all sets. \only<2->{Does it contain itself?} \only<3-7>{
          \begin{itemize}
            \item
              It's a set, and it contains all sets, so it must.
          \end{itemize}
        }
        \item<4-7>
          Consider the set of all sets that do not contain themselves.  Does it contain itself?
          \only<5-7>{
          \begin{itemize}
            \item<5->
              If it contains itself, then it is not a set that doesn't contain itself, and so it cannot be a member of itself; contradiction.  Thus it cannot contain itself. \pause
            \item<6->
              If it does not contain itself, then it is a set that does not contain itself, and thus must be a member of itself; contradiction.  Thus it cannot fail to contain itself.
          \end{itemize}
        }
        \item<7->
          This is the \textbf{paradox of na\"ive set theory}.
        \item<8->
          Solution: universe levels
          \begin{itemize}
            \item<9->
              \texttt{Set} or \texttt{Type}(0) is the collection of all sets, \texttt{Type}(1) is the collection of all \texttt{Type}(0)s, \ldots, \texttt{Type}($i + 1$) is the collection of all \text{Type}($i$)s
            \item<9->
              The \textbf{universe level} of an object of type \texttt{Type}($i$) is $i$
          \end{itemize}
        \item<10->
          In some cases, Coq can infer the universe level of an inductive type from the universe levels of its parameters; when this happens, the inductive type is polymorphic over universe levels.
        \item<11->
          It's useful to talk about ``a category whose objects are of type \texttt{T}'' rather than just ``a category''.
      \end{itemize}
    \end{frame}

  \subsection{Limits and Colimits}
    \begin{frame}{Limits and Colimits}
      \begin{itemize}
        \item
          Categorical \textbf{limits} are like \textbf{Cartesian products}, subject to constraints about equality of components \pause
        \item
          Categorical \textbf{colimits} are like \textbf{disjoint unions}, modulo equivalence relations
      \end{itemize}
    \end{frame}

    \begin{frame}{Coq Category}
      \begin{itemize}
        \item Coq has all limits
          \begin{itemize}
            \item
              Product types provide products (function types, e.g, \texttt{forall~a~:~A,~f~a} is the product $\prod_{a \in A} f(a)$)
            \item
              Sigma types provide constraints about equality of components (e.g., \texttt{\{ f : A $\to$ B | f a = f b \}})
          \end{itemize} \pause
        \item Coq has some colimits
          \begin{itemize}
            \item
              Sigma types provide disjoint unions (e.g., \texttt{\{ j : J \& f j \}} is the disjoint union $\bigsqcup_{j \in J} f(j)$)
            \item
              Quotients are \ldots\space hard
          \end{itemize}
      \end{itemize}
    \end{frame}

    \begin{frame}{Quotients}
      \begin{itemize}
        \item Quotients can be defined via axioms \pause
        \begin{itemize}
          \item
            \texttt{proof\_irrelevance}; \texttt{(A $\leftrightarrow$ B) $\to$ A = B} for propositions; either decidable existence, or a way of turning proofs of existence into objects (\texttt{constructive\_indefinite\_description : (exists~x,~P~x)~$\to$~\{~x~|~P~x~\}})
          \item
            Not computational
        \end{itemize} \pause
        \item Quotients can be defined via setoids \pause
        \begin{itemize}
          \item
            All objects carry around extra information of what the equivalence relation is
          \item
            This is somewhat clunky
          \item
            Not first-class quotients
        \end{itemize}
      \end{itemize}
    \end{frame}

    \begin{frame}{Limits and Colimits (High-Level Summary)}
      \begin{itemize}
        \item
          There are two categorical constructions (limits and colimits) that are ``dual'' \pause
        \item
          Coq's type-system fully implements only one of these (limits) \pause
        \item
          It's harder to define colimits inside of Coq than limits, in general, even for the ones that Coq does support
      \end{itemize}
    \end{frame}

  \begin{frame}[Thank You]
    \fontsize{128}{0}\selectfont\begin{center}Thank You!\end{center}
  \end{frame}
\end{document}

### 2014-01-22 presentations/popl-2013/minute-madness.tex
Add some presentations

--- 2014-01-22 presentations/popl-2013/jgross-student-talk.tex
@@
 \pdfinfo {
-  /Title (Database Management on top of Category Theory in Coq)
+  /Title (POPL --- Minute Madness)
   /Subject (POPL)
   /Author (Jason Gross)
-  /Keywords (POPL, Coq, Category Theory, Databases)
+  /Keywords (POPL, Coq, PLMW, Minute Madness, Category Theory)
 }
@@
 
-\title[Database Management on top of Category Theory in Coq]{Building Database Management on top of Category Theory in Coq}
+\title{POPL --- Minute Madness}
 \date{POPL 2013}
@@
 
-\begin{frame}
-  \titlepage
-  This document is available at \url{http://web.mit.edu/jgross/Public/POPL/\jobname.pdf}.
-
-  My category theory library is available at \url{https://bitbucket.org/JasonGross/catdb}.
-\end{frame}
-
-\section*{Outline}
-  \begin{frame}{Outline}
-    \tableofcontents[pausesections]
-  \end{frame}
-
-\section{Introduction --- Databases and Category Theory}    
-  \subsection{Categories}
-    \begin{frame}{Categories}
-      A category is:
-      \begin{itemize}
-        \item
-          a collection of objects, \pause together with
-        \item
-          arrows between those objects, \pause together with
-        \item
-          a composition law for the arrows satisfying coherence conditions:
-          \begin{itemize}
-            \item existence of identity
-            \item associativity
-          \end{itemize} \pause
-      \end{itemize}
+\begin{frame}{Database Management on top of Category Theory in Coq \\ Category of Relational Schemas $\cong$ Category of Categories}
   \begin{center}
@@
         $$
-        }
-      \end{center}
-    \end{frame}
-
-  \subsection{Relational Databases}
-    \begin{frame}{Relational Databases}
-      A database schema for a relational database can be modeled as a
-      \begin{itemize}
-        \item collection of \textbf{tables}, \pause together with 
-        \item a collection of \textbf{attributes} or column-labels for each table, \pause together with
-        \item integrity constraints \pause
-      \end{itemize}
-      \begin{center}
-        \fbox{
-          $$
-          \xymatrix{
-            *\txt{professor} \ar@/^{1pc}/[rr]^{\text{department}} \ar@/_{1pc}/[dd]_{\text{first name}} \ar@/^{1pc}/[dd]^{\text{last name}} && *\txt{department} \ar@/^{1pc}/[ll]^{\text{department head}} \ar@/^{2pc}/[ddll]^{\text{name}} \\ \\
-            *\txt{string} &&
-          }
-          $$
-          }
-      \end{center}
-    \end{frame}
-
-  \subsection{Relational Database Schema = Category}
-    \begin{frame}{Relational Database Schema = Category}
-      \begin{center}
-        \only<1,3>{
-          \fbox{
-          $$
-          \xymatrix{
-            a \ar@/^{0.5pc}/[rr]^{f} \ar@/_{0.5pc}/[d]_{g} \ar@/^{0.5pc}/[d]^{h} && b \ar@/^{0.5pc}/[ll]^{i} \ar@/^{1.5pc}/[dll]^{j} \\
-            c
-          }
-          $$
-          }
-        }
-        \only<3>{\raisebox{-3.5ex}[\height][\depth]{$\cong$}}
-        \only<2,3>{
-          \fbox{\only<3>{\small}
+      } \raisebox{-3.5ex}[\height][\depth]{$\cong$}
+    \fbox{\small
       $$
@@
       }
-        }
-        \only<3>{\vspace{4ex} \\ The diagrams are ``the same''.}
+
+    \vspace{4ex} \\ The diagrams are ``the same''.
   \end{center}
 \end{frame}
-
-  \subsection{Usefulness}
-    \begin{frame}{Usefulness of Categorical Databases}
-      \begin{itemize}
-        \item
-          Built in notion of path equivalence (multiple equivalent paths of foreign keys can be a pain in typical database management). \pause
-        \item
-          Provides a rigorous language for data migration between databases (another hard task in standard database management).
-      \end{itemize}
-    \end{frame}
-
-\section{Category Theory in Coq}
-
-  \begin{frame}{Category Theory in Coq}
-    \begin{itemize}
-      \item
-        Many people learn a proof assistant by coding up category theory. \pause
-      \item
-        Category theory is relatively simple to code up. \pause
-        \begin{itemize}
-          \item
-            Standard rigorous formulation of concepts exists in the literature. \pause
-          \item
-            It's rare to get caught up in minute details of proofs. \pause
-          \item
-            If you can define something categorically, it's probably interesting.
-        \end{itemize}
-    \end{itemize}
-  \end{frame}
-
-  \subsection{Universe Levels}
-    \begin{frame}
-      \frametitle{Universe Levels \only<1-7>{(Russel's Paradox)}}
-      \begin{itemize}
-        \item<1-7>
-          Consider, na\"ively, the set of all sets. \only<2->{Does it contain itself?} \only<3-7>{
-          \begin{itemize}
-            \item
-              It's a set, and it contains all sets, so it must.
-          \end{itemize}
-        }
-        \item<4-7>
-          Consider the set of all sets that do not contain themselves.  Does it contain itself?
-          \only<5-7>{
-          \begin{itemize}
-            \item<5->
-              If it contains itself, then it is not a set that doesn't contain itself, and so it cannot be a member of itself; contradiction.  Thus it cannot contain itself. \pause
-            \item<6->
-              If it does not contain itself, then it is a set that does not contain itself, and thus must be a member of itself; contradiction.  Thus it cannot fail to contain itself.
-          \end{itemize}
-        }
-        \item<7->
-          This is the \textbf{paradox of na\"ive set theory}.
-        \item<8->
-          Solution: universe levels
-          \begin{itemize}
-            \item<9->
-              \texttt{Set} or \texttt{Type}(0) is the collection of all sets, \texttt{Type}(1) is the collection of all \texttt{Type}(0)s, \ldots, \texttt{Type}($i + 1$) is the collection of all \text{Type}($i$)s
-            \item<9->
-              The \textbf{universe level} of an object of type \texttt{Type}($i$) is $i$
-          \end{itemize}
-        \item<10->
-          In some cases, Coq can infer the universe level of an inductive type from the universe levels of its parameters; when this happens, the inductive type is polymorphic over universe levels.
-        \item<11->
-          It's useful to talk about ``a category whose objects are of type \texttt{T}'' rather than just ``a category''.
-      \end{itemize}
-    \end{frame}
-
-  \subsection{Limits and Colimits}
-    \begin{frame}{Limits and Colimits}
-      \begin{itemize}
-        \item
-          Categorical \textbf{limits} are like \textbf{Cartesian products}, subject to constraints about equality of components \pause
-        \item
-          Categorical \textbf{colimits} are like \textbf{disjoint unions}, modulo equivalence relations
-      \end{itemize}
-    \end{frame}
-
-    \begin{frame}{Coq Category}
-      \begin{itemize}
-        \item Coq has all limits
-          \begin{itemize}
-            \item
-              Product types provide products (function types, e.g, \texttt{forall~a~:~A,~f~a} is the product $\prod_{a \in A} f(a)$)
-            \item
-              Sigma types provide constraints about equality of components (e.g., \texttt{\{ f : A $\to$ B | f a = f b \}})
-          \end{itemize} \pause
-        \item Coq has some colimits
-          \begin{itemize}
-            \item
-              Sigma types provide disjoint unions (e.g., \texttt{\{ j : J \& f j \}} is the disjoint union $\bigsqcup_{j \in J} f(j)$)
-            \item
-              Quotients are \ldots\space hard
-          \end{itemize}
-      \end{itemize}
-    \end{frame}
-
-    \begin{frame}{Quotients}
-      \begin{itemize}
-        \item Quotients can be defined via axioms \pause
-        \begin{itemize}
-          \item
-            \texttt{proof\_irrelevance}; \texttt{(A $\leftrightarrow$ B) $\to$ A = B} for propositions; either decidable existence, or a way of turning proofs of existence into objects (\texttt{constructive\_indefinite\_description : (exists~x,~P~x)~$\to$~\{~x~|~P~x~\}})
-          \item
-            Not computational
-        \end{itemize} \pause
-        \item Quotients can be defined via setoids \pause
-        \begin{itemize}
-          \item
-            All objects carry around extra information of what the equivalence relation is
-          \item
-            This is somewhat clunky
-          \item
-            Not first-class quotients
-        \end{itemize}
-      \end{itemize}
-    \end{frame}
-
-    \begin{frame}{Limits and Colimits (High-Level Summary)}
-      \begin{itemize}
-        \item
-          There are two categorical constructions (limits and colimits) that are ``dual'' \pause
-        \item
-          Coq's type-system fully implements only one of these (limits) \pause
-        \item
-          It's harder to define colimits inside of Coq than limits, in general, even for the ones that Coq does support
-      \end{itemize}
-    \end{frame}
-
-  \begin{frame}[Thank You]
-    \fontsize{128}{0}\selectfont\begin{center}Thank You!\end{center}
-  \end{frame}
 \end{document}

### 2014-04-12 presentations/coq-workshop-2014/coq-workshop-proposal-tactics-in-terms.tex
Add coq workshop proposal

\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage[utf8]{inputenc}
% copy-pastable ligatures
\input{glyphtounicode}
\pdfgentounicode=1
\usepackage[margin=1in]{geometry}
\usepackage{verbatim}
\usepackage{comment}
\usepackage{xcolor}
\usepackage{fancyhdr}
\title{Presentation of Three Neat Tricks in Coq 8.5}
\author{Jason Gross}
\date{ITP 2014: 6th Coq Workshop}
\usepackage[pdftex,
            pdfauthor={Jason Gross},
            pdftitle={Presentation of Three Neat Tricks in Coq 8.5},
            pdfsubject={ITP 2014: 6th Coq Workshop},
            pdfkeywords={Coq 8.5, Ltac, Gallina, binders, notations, tactics, recursion, recursive tactics, typeclasses, universe polymorphism, universes}]{hyperref}
\chead{Presentation of Three Neat Tricks in Coq 8.5 \\ Jason Gross \\ ITP 2014: 6th Coq Workshop}
\thispagestyle{fancy}
% Keywords:
% Coq 8.5
% Ltac
% Gallina
% binders
% notations
% tactics
% recursion
% recursive tactics
% typeclasses
% universe polymorphism
% universes

\setcounter{secnumdepth}{0}
\setlength{\headheight}{35pt}

\makeatletter
\begingroup
  \catcode`{=11% { is a letter
  \catcode`}=11% } is a letter
  \catcode`<=1% < acts like {
  \catcode`>=2% > acts like }
  \gdef\@makeopenbrace<\catcode`{>
  \gdef\@makeclosebrace<\catcode`}>
  \global\let\mytextbraceleft={
  \global\let\mytextbraceright=}
\endgroup

\newcommand{\processcommands}[1]{{%
  \catcode`\\=0\relax
  \@makeopenbrace=1\relax
  \@makeclosebrace=2\relax
  \def\{{\texttt{\mytextbraceleft}}
  \def\}{\texttt{\mytextbraceright}}
  \scantokens{#1}%
}}
\newcommand{\processcommandsinverbatimline}{\expandafter\processcommands\expandafter{\the\verbatim@line}}
\newenvironment{coqcode}{\begingroup
  \vspace{0.5\baselineskip}
  \let\trivlist\relax
  \let\endtrivlist\relax
  \let\item\relax
  \setlength{\parsep}{0pt}%
  \setlength{\parskip}{0pt}%
  \setlength{\topsep}{0pt}%
  \setlength{\@topsepadd}{0pt}%
  \setlength{\partopsep}{0pt}%
  \setlength{\@topsep}{0pt}%
  \let\old@@par\@@par
  \let\@@par\relax
  \let\old@vskip\vskip
  \let\vskip\relax
%  \setlength{\parfillskip}{0pt}%
  \verbatim
  \let\@@par\old@@par
  \let\vskip\old@vskip
  \let\verbatim@processline=\processcommandsinverbatimline
}{\endverbatim\endgroup\vspace{0.5\baselineskip}}
\makeatother

\DeclareUnicodeCharacter{2218}{\ensuremath{\circ}}
\DeclareUnicodeCharacter{2192}{\ensuremath{\to}}
\DeclareUnicodeCharacter{2200}{\ensuremath{\forall}}

% copy some colors from coqdoc
\definecolor{varpurple}{rgb}{0.4,0,0.4}
\definecolor{constrmaroon}{rgb}{0.6,0,0}
\definecolor{defgreen}{rgb}{0,0.4,0}
\definecolor{indblue}{rgb}{0,0,0.8}
\definecolor{kwred}{rgb}{0.8,0.1,0.1}

% macro for typesetting keywords
\newcommand{\coqdockw}[1]{\texttt {\textcolor{kwred}{#1}}}

% macro for typesetting variable identifiers
\newcommand{\coqdocvar}[1]{\textcolor{varpurple}{#1}}

% macro for typesetting constant identifiers
\newcommand{\coqdoccst}[1]{\texttt{\textcolor{defgreen}{#1}}}%{\textsf{#1}}

% macro for typesetting module identifiers
\newcommand{\coqdocmod}[1]{\textsc{\textsf{#1}}}

\newcommand{\coqd}[1]{\texttt {\textcolor{defgreen}{#1}}}
% macro for typesetting module constant identifiers (e.g. Parameters in
% module types)
\newcommand{\coqdocax}[1]{\texttt {\textcolor{indblue}{#1}}}

% macro for typesetting inductive type identifiers
\newcommand{\coqdocind}[1]{\texttt{\textcolor{indblue}{#1}}}%{\textbf{\textsf{#1}}}

% macro for typesetting constructor identifiers
\newcommand{\coqdocconstr}[1]{\texttt {\textcolor{constrmaroon}{#1}}}

% macro for typesetting tactic identifiers
\newcommand{\coqdoctac}[1]{\texttt{#1}}

% These are the real macros used by coqdoc, their typesetting is 
% based on the above macros by default.

\newcommand{\coqdoclibrary}[1]{\coqdoccst{#1}}
\newcommand{\coqdocinductive}[1]{\coqdocind{#1}}
\newcommand{\coqdocdefinition}[1]{\coqdoccst{#1}}
\newcommand{\coqdocvariable}[1]{\coqdocvar{#1}}
\newcommand{\coqdocconstructor}[1]{\coqdocconstr{#1}}
\newcommand{\coqdoclemma}[1]{\coqdoccst{#1}}
\newcommand{\coqdocclass}[1]{\coqdocind{#1}}
\newcommand{\coqdocinstance}[1]{\coqdoccst{#1}}
\newcommand{\coqdocmethod}[1]{\coqdoccst{#1}}
\newcommand{\coqdocabbreviation}[1]{\coqdoccst{#1}}
\newcommand{\coqdocrecord}[1]{\coqdocind{#1}}
\newcommand{\coqdocprojection}[1]{\coqdoccst{#1}}
\newcommand{\coqdocnotation}[1]{\coqdockw{#1}}
\newcommand{\coqdocsection}[1]{\coqdoccst{#1}}
\newcommand{\coqdocaxiom}[1]{\coqdocax{#1}}
\newcommand{\coqdocmodule}[1]{\coqdocmod{#1}}

\begin{document}
\begin{abstract}
Coq 8.5 has a number of new features.  It has more powerful universe polymorphism support.  It allows tactics to be run at interpretation to construct other terms.  The ability to switch from Gallina to Ltac in arbitrary locations nicely complements the \texttt{constr:}~notation permitting the switch from Ltac to Gallina in tactics, and opens up many new possibilities.  I propose to present three tricks involving these new features: tactics in terms allows the construction of tactics that recurse under binders; tactics in terms together with typeclasses allows overloading notations based on the type of their arguments; and there is a way to talk about universe levels explicitly, helped along by tactics in terms.
\end{abstract}
\subsection{Tactics in Terms}
Coq 8.5 features tactics in terms using the syntax \texttt{\$(\ldots)\$} to denote a tactic script, which evaluates to the proof tree it builds in proving a goal.  For example, we can simplify terms generated by tactics before saving them.
\begin{coqcode}
\coqdockw{Definition} \coqdocdefinition{I_am_one} : \coqdocinductive{nat} := \$(\coqdockw{let} \coqdocvar{x} := \coqdockw{constr}:(2 \coqdocnotation{-} 1) \coqdoctac{in} \coqdoctac{exact} \coqdocvar{x})\$.
\coqdockw{Print} \coqdocdefinition{I_am_one}.
(* I_am_one = 2 - 1 : nat *)
\coqdockw{Definition} \coqdocdefinition{I_am_exactly_one} : \coqdocvar{nat} := \$(\coqdockw{let} \coqdocvar{x} := \coqdockw{constr}:(2 - 1) \coqdoctac{in}
                                       \coqdockw{let} \coqdocvar{y} := (\coqdoctac{eval} \coqdoctac{compute} \coqdoctac{in} \coqdocvar{x}) \coqdoctac{in}
                                       \coqdoctac{exact} \coqdocvar{y})\$.
\coqdockw{Print} \coqdocdefinition{I_am_exactly_one}.
(* I_am_exactly_one = 1 : nat *)
\end{coqcode}
\subsection{Recursive Tactics}
The ability to go back and forth between Gallina and Ltac at will allows tactics that recurse under binders, by using Gallina to add binders and Ltac to recurse. Here is an example that returns the left projection of an \texttt{and} under binders:
\begin{coqcode}
\coqdockw{Ltac} \coqdocvar{ret_and_left} \coqdocvar{f} :=
  \coqdockw{let} \coqdocvar{tac} := \coqdocvar{ret_and_left} \coqdoctac{in}
  \coqdockw{let} \coqdocvar{T} := \coqdocvar{type} \coqdocvar{of} \coqdocvar{f} \coqdoctac{in}
  \coqdockw{lazymatch} \coqdoctac{eval} \coqdoctac{hnf} \coqdoctac{in} \coqdocvar{T} \coqdockw{with}
    \ensuremath{|} ?\coqdocvar{a} \coqdocnotation{\ensuremath{\land}} ?\coqdocvar{b} \ensuremath{\Rightarrow} \coqdoctac{exact} (\coqdoclemma{proj1} \coqdocvar{f})
    \ensuremath{|} ?\coqdocvar{T'} \coqdocnotation{\ensuremath{\rightarrow}} \coqdocvar{_}
      \ensuremath{\Rightarrow} \coqdockw{let} \coqdocvar{ret} := \coqdockw{constr}:(\coqdockw{\ensuremath{\lambda}} \coqdocvar{x'} : \coqdocvar{T'} \ensuremath{\Rightarrow} \coqdockw{let} \coqdocvar{fx} := \coqdocvar{f} \coqdocvar{x'} \coqdoctac{in} \$(\coqdocvar{tac} \coqdocvar{fx})\$) \coqdoctac{in}
         \coqdockw{let} \coqdocvar{ret'} := (\coqdoctac{eval} \coqdoctac{cbv} \coqdockw{zeta} \coqdoctac{in} \coqdocvar{ret})  \coqdoctac{in}
         \coqdoctac{exact} \coqdocvar{ret'}
    \coqdockw{end}.
\coqdockw{Goal} \coqdockw{\ensuremath{\forall}} \coqdocvar{A} \coqdocvar{B} : \coqdockw{Prop}, (\coqdocvar{A} \coqdocnotation{\ensuremath{\rightarrow}} \coqdocvar{A} \coqdocnotation{\ensuremath{\rightarrow}} \coqdocvar{A} \coqdocnotation{\ensuremath{\land}} \coqdocvar{B}) \coqdocnotation{\ensuremath{\rightarrow}} \coqdocinductive{True}.
  \coqdoctac{intros} \coqdocvar{A} \coqdocvar{B} \coqdocvar{H}.
  \coqdoctac{pose} \$(\coqdocvar{ret_and_left} \coqdocvar{H})\$.
  (* \ensuremath{\lambda} {x'} {x'0} : {A} \ensuremath{\Rightarrow} {proj1} ({H} {x'} {x'0}) : {A} {\ensuremath{\rightarrow}} {A} {\ensuremath{\rightarrow}} {A} *)
\end{coqcode}
This code works around some bugs in the current implementation, such as the fact that the tactic currently being defined is not available inside of \texttt{\$(\ldots)\$}, and binders added within \texttt{constr:}~are not renamed when going under binders.

\subsection{Overloading Notations on Argument Types}
Coq provides a way of overloading notations via scopes; together with binding scopes to types, this allows reusing the same notation to mean similar but different things when you are expecting terms of different resulting types.  Using typeclasses, we can extend this overloading to a dependence on the type of the arguments, though this breaks rewriting.  Using tactics in terms (in notations), we can fix rewriting.  Due to limitations in the current trunk version of Coq, this breaks using evars in such notations.\footnote{\url{https://coq.inria.fr/bugs/show\_bug.cgi?id=3278}}  This is useful, for example, if you want to define both horizontal and vertical composition using the same symbol, which both return the same type of morphism, but have different argument types.  Here is a very simple example:

\newcommand{\Ts}[1]{T\ensuremath{_#1}}
\newcommand{\ts}[1]{t\ensuremath{_#1}}
\begin{coqcode}
\coqdockw{Parameters} \coqdocaxiom{\Ts1} \coqdocaxiom{\Ts2} \coqdocaxiom{\Ts3} : \coqdockw{Type}.
\coqdockw{Parameter} \coqdocaxiom{F} : \coqdocaxiom{\Ts1} \coqdocnotation{\ensuremath{\rightarrow}} \coqdocaxiom{\Ts3}.
\coqdockw{Parameter} \coqdocaxiom{G} : \coqdocaxiom{\Ts2} \coqdocnotation{\ensuremath{\rightarrow}} \coqdocaxiom{\Ts3}.
\coqdockw{Parameters} (\coqdocaxiom{\ts1} : \coqdocaxiom{\Ts1}) (\coqdocaxiom{\ts2} : \coqdocaxiom{\Ts2}).
\vspace{0em}
\coqdockw{Class} \coqdocrecord{MyNotation} \{\coqdocvar{A} \coqdocvar{R}\} (\coqdocvar{a} : \coqdocvar{A}) (\coqdocvar{r} : \coqdocvar{R}) := \{\}.

\coqdockw{Definition} \coqdocdefinition{mynotation} \coqdocvar{A} \coqdocvar{R} \coqdocvar{a} \coqdocvar{r} `\{@\coqdocclass{MyNotation} \coqdocvar{A} \coqdocvar{R} \coqdocvar{a} \coqdocvar{r}\} := \coqdocvar{r}.

\coqdockw{Instance} \coqdocinstance{MyF} \coqdocvar{x} : \coqdocclass{MyNotation} \coqdocvar{x} (\coqdocaxiom{F} \coqdocvar{x}) \ensuremath{|} 10.
\coqdockw{Instance} \coqdocinstance{MyG} \coqdocvar{x} : \coqdocclass{MyNotation} \coqdocvar{x} (\coqdocaxiom{G} \coqdocvar{x}) \ensuremath{|} 100.
\vspace{0em}
\coqdockw{Notation} "\% x"
    := (\$(\coqdockw{let} \coqdocvar{ret} := \coqdockw{constr}:(@\coqdocdefinition{mynotation} \coqdocvar{_} \coqdocvar{_} \coqdocvar{x} \coqdocvar{y} \coqdocvar{_}) \coqdoctac{in}
          \coqdockw{let} \coqdocvar{ret'} := (\coqdoctac{eval} \coqdoctac{cbv} \coqdockw{beta} \coqdockw{delta} [\coqdocdefinition{mynotation}] \coqdoctac{in} \coqdocvar{ret}) \coqdoctac{in}
          \coqdoctac{exact} \coqdocvar{ret'})\$)
  (\coqdockw{at} \coqdockw{level} 40, \coqdockw{only} \coqdockw{parsing}).

\coqdockw{Check} (\coqdocnotation{\%} \coqdocaxiom{\ts1}). (* F \ts1 : \Ts3 *)
\coqdockw{Check} (\coqdocnotation{\%} \coqdocaxiom{\ts2}). (* G \ts2 : \Ts3 *)
\end{coqcode}

If we wanted to display these notations as well, we could add extra notations for each constructor, for example,
\begin{coqcode}
\coqdockw{Notation} "\% x" := (\coqdocaxiom{F} \coqdocvar{x}).
\coqdockw{Notation} "\% x" := (\coqdocaxiom{G} \coqdocvar{x}).
\end{coqcode}
\noindent above the general notation above.

\begin{comment}
\newcommand{\Ts}[1]{T\ensuremath{_#1}}
\newcommand{\ts}[1]{t\ensuremath{_#1}}
\begin{coqcode}
\coqdockw{Parameters} \coqdocaxiom{\Ts1} \coqdocaxiom{\Ts2} : \coqdockw{Type}.
\coqdockw{Parameter} \coqdocvar{Compose11} : \coqdocaxiom{\Ts1} \coqdocnotation{\ensuremath{\rightarrow}} \coqdocaxiom{\Ts1} \coqdocnotation{\ensuremath{\rightarrow}} \coqdocaxiom{\Ts1}.
\coqdockw{Parameter} \coqdocvar{Compose12} : \coqdocaxiom{\Ts1} \coqdocnotation{\ensuremath{\rightarrow}} \coqdocaxiom{\Ts2} \coqdocnotation{\ensuremath{\rightarrow}} \coqdocaxiom{\Ts1}.
\coqdockw{Parameter} \coqdocvar{Compose21} : \coqdocaxiom{\Ts2} \coqdocnotation{\ensuremath{\rightarrow}} \coqdocaxiom{\Ts1} \coqdocnotation{\ensuremath{\rightarrow}} \coqdocaxiom{\Ts1}.
\coqdockw{Parameter} \coqdocvar{Compose22} : \coqdocaxiom{\Ts2} \coqdocnotation{\ensuremath{\rightarrow}} \coqdocaxiom{\Ts2} \coqdocnotation{\ensuremath{\rightarrow}} \coqdocaxiom{\Ts2}.
\coqdockw{Parameters} (\coqdocaxiom{\ts1} : \coqdocaxiom{\Ts1}) (\coqdocaxiom{\ts2} : \coqdocaxiom{\Ts2}).
\vspace{0em}
\coqdockw{Class} \coqdocvar{ComposeTo} \{\coqdocvar{A} \coqdocvar{B} \coqdocvar{C}\} (\coqdocvar{a} : \coqdocvar{A}) (\coqdocvar{b} : \coqdocvar{B}) (\coqdocvar{c} : \coqdocvar{C}) := \{\}.

\coqdockw{Definition} \coqdocvar{composition} \coqdocvar{A} \coqdocvar{B} \coqdocvar{C} \coqdocvar{a} \coqdocvar{b} \coqdocvar{c} `\{@\coqdocvar{ComposeTo} \coqdocvar{A} \coqdocvar{B} \coqdocvar{C} \coqdocvar{a} \coqdocvar{b} \coqdocvar{c}\} := \coqdocvar{c}.

\coqdockw{Instance} \coqdocvar{ComposeTo11} \coqdocvar{x} \coqdocvar{y} : \coqdocvar{ComposeTo} \coqdocvar{x} \coqdocvar{y} (\coqdocvar{Compose11} \coqdocvar{x} \coqdocvar{y}) \ensuremath{|} 10.

\coqdockw{Instance} \coqdocvar{ComposeTo12} \coqdocvar{x} \coqdocvar{y} : \coqdocvar{ComposeTo} \coqdocvar{x} \coqdocvar{y} (\coqdocvar{Compose12} \coqdocvar{x} \coqdocvar{y}) \ensuremath{|} 100.

\coqdockw{Instance} \coqdocvar{ComposeTo21} \coqdocvar{x} \coqdocvar{y} : \coqdocvar{ComposeTo} \coqdocvar{x} \coqdocvar{y} (\coqdocvar{Compose21} \coqdocvar{x} \coqdocvar{y}) \ensuremath{|} 100.

\coqdockw{Instance} \coqdocvar{ComposeTo22} \coqdocvar{x} \coqdocvar{y} : \coqdocvar{ComposeTo} \coqdocvar{x} \coqdocvar{y} (\coqdocvar{Compose22} \coqdocvar{x} \coqdocvar{y}) \ensuremath{|} 1000.
\vspace{0em}
\coqdockw{Notation} "x ∘ y"
    := (\$(\coqdockw{let} \coqdocvar{ret} := \coqdockw{constr}:(@\coqdocvar{composition} \coqdocvar{_} \coqdocvar{_} \coqdocvar{_} \coqdocvar{x} \coqdocvar{y} \coqdocvar{_} \coqdocvar{_}) \coqdoctac{in}
          \coqdockw{let} \coqdocvar{ret'} := (\coqdoctac{eval} \coqdoctac{cbv} \coqdockw{beta} \coqdockw{delta} [\coqdocvar{composition}] \coqdoctac{in} \coqdocvar{ret}) \coqdoctac{in}
          \coqdoctac{exact} \coqdocvar{ret'})\$)
  (\coqdockw{at} \coqdockw{level} 40, \coqdockw{left} \coqdockw{associativity}, \coqdockw{only} \coqdockw{parsing}).

\coqdockw{Check} (\coqdocaxiom{\ts1} \coqdocnotation{∘} \coqdocaxiom{\ts1}). (* Compose11 \ts1 \ts1 : \Ts1 *)
\coqdockw{Check} (\coqdocaxiom{\ts1} \coqdocnotation{∘} \coqdocaxiom{\ts2}). (* Compose12 \ts1 \ts2 : \Ts1 *)
\coqdockw{Check} (\coqdocaxiom{\ts2} \coqdocnotation{∘} \coqdocaxiom{\ts1}). (* Compose21 \ts1 \ts1 : \Ts1 *)
\coqdockw{Check} (\coqdocaxiom{\ts2} \coqdocnotation{∘} \coqdocaxiom{\ts2}). (* Compose22 \ts2 \ts2 : \Ts2 *)
\end{coqcode}
\end{comment}

\subsection{Explicit Universe Levels}
The final trick involves the new, more powerful universe polymorphism algorithm that Matthieu Sozeau has recently implemented.  Although Coq still uses typical ambiguity and gives no explicit way to talk about universes, it is possible to define a \texttt{Lift} function which takes a type and returns the same type in a higher universe.  It is possible to use this to prove, for example, than if one universe satisfies functional extensionality, then so do all lower universes; that theorem is useful for theorems that end up needing multiple instances of functional extensionality at various universe levels to avoid universe inconsistencies.  Here is the definition of \texttt{Lift}, and an associated lift for equality.  (The lift for equality only works correctly if \texttt{eq} is polymorphic.)
\begin{coqcode}
\coqdockw{Set} \coqdockw{Universe} \coqdockw{Polymorphism}.
\coqdockw{Inductive} \coqdocinductive{eq} \{\coqdocvar{A}\} (\coqdocvar{x} : \coqdocvar{A}) : \coqdocvar{A} \coqdocnotation{\ensuremath{\rightarrow}} \coqdockw{Set} := \coqdocconstructor{eq_refl} : \coqdocinductive{eq} \coqdocvar{x} \coqdocvar{x}.
\coqdockw{Notation} "x = y :> A" := (@\coqdocinductive{eq} \coqdocvar{A} \coqdocvar{x} \coqdocvar{y}) : \coqdocvar{type_scope}.
\coqdockw{Notation} "x = y" := (\coqdocvar{x} \coqdocnotation{=} \coqdocvar{y} \coqdocnotation{:>} \coqdocvar{_}) : \coqdocvar{type_scope}.
\coqdockw{Definition} \coqdocdefinition{Lift}
 : \$(\coqdockw{let} \coqdocvar{U1} := \coqdockw{constr}:(\coqdockw{Type}) \coqdoctac{in}
     \coqdockw{let} \coqdocvar{U0} := \coqdockw{constr}:(\coqdockw{Type} : \coqdocvar{U1}) \coqdoctac{in}
     \coqdockw{let} \coqdocvar{U0'} := (\coqdoctac{eval} \coqdoctac{simpl} \coqdoctac{in} \coqdocvar{U0}) \coqdoctac{in}
     \coqdoctac{exact} (\coqdocvar{U0'} \coqdocnotation{\ensuremath{\rightarrow}} \coqdocvar{U1}))\$
    := \coqdockw{\ensuremath{\lambda}} \coqdocvar{A} \ensuremath{\Rightarrow} \coqdocvar{A}.
\coqdockw{Definition} \coqdocdefinition{lift_eq} \{\coqdocvar{T} \coqdocvar{x} \coqdocvar{y}\} (\coqdocvar{p} : \coqdocvar{x} \coqdocnotation{=} \coqdocvar{y} \coqdocnotation{:>} \coqdocvar{T}) : \coqdocvar{x} \coqdocnotation{=} \coqdocvar{y} \coqdocnotation{:>} \coqdocvar{Lift} \coqdocvar{T}
    := \coqdockw{match} \coqdocvar{p} \coqdockw{with} \coqdocconstructor{eq_refl} \coqdocvar{_} \ensuremath{\Rightarrow} \coqdocconstructor{eq_refl} \coqdocvar{_} \coqdockw{end}.
\end{coqcode}
\end{document}

### 2020-07-30 papers/source/fiat-crypto/pldi-2018/response.md
Add a bunch of paper sources

Thanks for the wealth of suggestions on improving the paper's presentation.  We'll be glad to consider reordering material or adding new information.  For now, we'd like to point out how a few of the requests for expansion can be addressed with existing material in the paper, possibly in the context of some brief new clarifications.

- Review A asks for more detailed comparison with other approaches.  First, on the side of performance of related code, competing approaches involve writing code by hand, and they generally copy quite faithfully from the best known implementations (e.g., OpenSSL), so our comparison with best-known versions is a strong proxy for the performance elements of HACL*, Jasmin, and Vale.  On the side of developer effort, we hope the message came across that, once a new algorithm is implemented in our framework, it becomes almost free to recompile for a new prime modulus or a new hardware-architecture class, while as far as we know, no high-performance past project had implemented/verified more than one configuration of that kind per algorithm.  Furthermore, we don't know of any prior high-performance verification of NISTP256 or Montgomery reduction, which are used by the vast majority of TLS connections today.
- Review B asks how developers use the tool and what comes out in the end.  We should perhaps have been clearer that our story involves at least three layers: the framework, generic implementations of algorithms, and instantiations of those algorithms to new parameters.  Targeting a new algorithm involves new coding and proving work (often nontrivial).  However, compiling an existing algorithm with new parameters really does only require the level of effort seen in the "Input" part of Figure 1 (no proving, etc.).  Much of the PLDI audience probably takes that sort of thing for granted, but as far as we know, all past approaches have required days of work to "recompile" an existing algorithm for new parameters.  And, yes, the "Output" part of Figure 1 is representative of what our tool produces.  It can be copied-and-pasted into a project, compiled and linked as a library, etc.
- Review C expresses skepticism that crypto-library authors will use this tool.  We're not sure what to say beyond our original (blinded) description of how one of the most used crypto libraries has already adopted our framework (and more has happened since paper submission).  It seems there is a nontrivial segment of the crypto world where stronger security guarantees can outweigh the last 10% of performance.
- Review D says that we failed to discuss side-channel resistance, though we believe the issue was treated fully (but perhaps the presentation deserves rearranging).  In addition to several other mentions throughout the paper, see the paragraph starting on line 951.  Essentially, our target language only exposes instructions that preserve constant time.
- Review D asks how the compiler chooses a representation.  The main part of our framework is designed to take the representation as a parameter, and it is hard to predict analytically which choice will perform best.  Thus, the experiments of Figure 3 involve heuristic generation of several representations per configuration (prime modulus and hardware target), compiling with each choice, and keeping the best-performing version.  Currently we only bother generating 2 representations per configuration, but it is easy to add new heuristics to the mix, with no new formal-methods work.
- Review E asks who performs CPS translation and makes important design choices.  Currently the programmer implementing a new generic algorithm codes and proves a CPS version manually.  We would like to apply CPS translation automatically, though our initial efforts were flummoxed by the dependent types used by our direct-style algorithms.  Design choices like data-representation changes for points are also made manually by the programmer implementing a new algorithm.  From this question and parts of other reviews, we see that one point we failed to make clearly is that nontrivial manual work goes into implementing a new verified, generic algorithm, while we then provide a good automation story for instantiating that algorithm with new parameters.

More responses past the 700-word limit
======================================

We see two main use cases for the artifact produced in this work.
First it can be used in a plug-and-play fashion to generate modular arithmetic code using the already-implemented strategies.
This use case is rather narrow but very expedient when it is applicable.
More importantly, we think there is a lot going for implementing new optimized algorithms in the style we propose.
In particular, the code reuse made possible by our library (relying on a predictable specializing compiler) allows new operations (e.g. polynomial multiplication) and tricks (e.g. Karatsuba's multiplication, interleaved carrying and reduction) to be added with less effort than using any other methodology we are aware of.
For example, we implemented (and proved correct) Goldilocks Karatsuba multiplication at a conference lunch break when the author of the technique inquired whether our framework could support it.

Review D
--------

We wholly agree that going from the C code our pipeline generates to speed-record-setting assembly code requires a whole new set of techniques in addition to those described in this paper.
In our opinion, our output language is a natural interface between domain-specific and microarchitecture-specific optimizations, so that the C-to-assembly translation would not need to understand modular arithmetic or limbed representations to produce the best assembly we know to wish for.
As the first reviewer points out, a simple translation validation should be sufficient to connect the assembly code that includes low-level optimizations to the output of our pipeline it was generated from.
In fact, as seen in libsecp256k-c64, it is even possible to improve on our performance within the C language by munging source code (at the same abstraction layer as our generated code) into a form that happens to persuade a particular compiler to produce the desired register allocation and instruction schedule.

Asymmetric cryptography implementations (ours and others') use very simple CPU instructions and cannot benefit from the AES/SHA implementations found on some popular CPUs.

We view producing straight-line code as a strength rather than a limitation.
Any conventional compiler (e.g., gcc, ocamlc) could turn code roughly at the same abstraction layer as our templates into machine code that uses loops and branches, but so far it has taken significant amounts of (duplicative) human effort to produce specialized constant-time code.
We do not generate loops or complicated control-flow logic (e.g., as seen in elliptic-curve scalar multiplication routines that use precomputed tables) because writing (and verifying) the code with loops in the first place is pleasant and reliable when compared to doing the same for specialized field-arithmetic implementations.
Our library already includes verified code with iteration and conditionals, and we have integrated it (with proofs) with our generated field arithmetic.
There is just nothing novel about it.

Review E
--------

The question "In Section 2.3, are we assuming $p < 2^{64}$?" may have been caused by our typo of $\mathbb N_p$ in place of plain $\mathbb N$ on line 333.  Sorry!

On guarantees throughout our pipline: each run of the entire pipeline, from a specification like "a * b mod m" to C code, generates a proof certificate for functional correctness.
For the early phases, the proof is generated anew using tactics on each execution.
The flattening and word-size-selection transformations (the "certified compiler") are proven correct once and for all, and this proof is appropriately linked into the end-to-end proof.

The hand-modified version in section 6.2 is proven equivalent to the generated code it replaces before word-size selection using the `ring` tactic from Coq's standard library.
It is indeed ad-hoc but has an end-to-end Coq-checkable correctness proof against the same specification as the rest of the library, and the proof is very automated, in the tradition of translation validation.
We also intended that manual transformation as an experiment to zero in on the nature of our performance gap with other code; we are not suggesting manual code tweaking as part of the intended usage of a future, improved version of our framework (and it is easily skipped now if being within 10% of best-known performance is acceptable).

### 2020-07-30 papers/source/fiat-crypto/popl-2018/response.md
Add a bunch of paper sources

We're very glad to have reviews from 3 experts!
The comments point out some perspective issues that we're eager to address by revising the paper.
We'll mention those before returning to answer some more specific questions, after the 600-word limit.

Most importanly, we want to emphasize what we see as the main novelty of the paper: encapsulation of intricate machine arithmetic and its proofs in libraries, so that not only is there no significant manual proof effort for each new prime modulus; we don't even need to do something like call an SMT solver to prove each new suite of code.  Rather, we have proved the whole family of algorithms correct once and for all, in a way that still lets us generate pretty fast, idiomatic low-level code.

Next, there's a perspective summed up well in Reviewer C's comment that other recent work seems "much ahead [of us] in terms of the state-of-the-art of verified cryptography."
However, none of that work presents higher-level abstractions like we called out in the prior response paragraph.
We tried to sketch out our bigger-picture plan in the paper, where we do plan to connect to proofs of cryptographic security, of constant-time execution, and with a trusted base that ends in assembly language.
Those connections have not been made yet, but we felt that our main novelty (described in the prior response paragraph) is valuable and interesting enough on its own.

We appreciate the pointers to 3 papers (on HACL*, Jasmin, and Vale) that weren't published as of the POPL deadline.
As far as we know, none of those demonstrate almost-complete code or proof reuse across prime moduli for machine arithmetic, as we do.
(We should point out that the mainstream assembly-generating scripts that the Vale paper mentions are operating at a much lower level of abstraction than for our work, not coming close to abstracting out a prime modulus.)
However, as the reviewers point out, those other projects include complementary parts of a verification stack, so we plan to look into ways to connect to these systems, rather than building our own alternatives.
For instance, our pipeline is a perfect match with Jasmin: the final terms that we produce are effectively in a subset of the Jasmin input language.
We aren't aware of any duplication of functionality between our respective tools: every one of our phases is needed to produce code low-level enough for Jasmin or Vale to accept as input.

It seems worth reiterating that generation of C code is just an element of our current experimental setup: our formal development actually bottoms out in a simple straightline-code language (with a subset of the features of Jasmin), and we do not intend to use C in any way in the extensions we sketch as important future work.

Finally, multiple reviewers commented on the amount of space we spend in the paper on reviewing material well-known to experts on ECC.
We tried to negotiate the balance between keeping those experts interested and keeping a general POPL audience engaged; we will think about tweaking the balance based on the reviewers' remarks.
However, we do think that our presentation of key strategies often includes ideas that are "obvious in retrospect," but that were not known in this form to crypto experts.
Formulating the ideas in this way was crucial to making our Coq proofs tractable.
One example is the question we consider in Section 4.3, on flexible representation of multi-digit numbers.

# Further detailed responses (past word limit)

## How to prove equivalence between our initial generated version of a routine and another version where we have done some optimizations manually?

With straightline low-level code like is involved here, we just need to evaluate the two versions as functional programs and use Coq's `ring` tactic to prove equivalence of results.
In other words, naive symbolic execution suffices.

## Avoiding timing side channels

By design, our output AST type can *only* represent constant-time code; there are no branches, loops, or divisions.
(We do assume that the processor actually implements these instructions in a constant-time fashion, but checking that must be done experimentally as long as we do not have access to a precise specification or implementation that captures this timing behavior.)
We only mention timing analysis as part of future work that would extend our derivation pipeline to higher-level cryptographic code.

## Where is the generated code in the supplementary material?

The generated code is in `src/Specific/NISTP256/AMD64/` and `src/Specific/X25519/C64/`; to see output in our AST format, look at the files ending in `.log`.
To see C, run `make c` to generate `.c` files.

## fsatz: can we deal with 1/(1/x)?

Yes, and it is not a special case for the algorithm.
`src/Algebra/Field_test.v` contains several more complicated examples; adding `Lemma inv_inv x (H:x <> 0) : 1/(1/x) = x. fsatz. Qed.` passes.

## Effort to choose good carrying strategies

Selecting a carry chain is a complex problem; choices depend heavily on the target architecture, and there is not really a good way to evaluate them other than empirical testing.
Some general heuristics exist but may not provide optimal choices.
However, we could easily integrate any heuristic into our build process, with no effect on proofs.

## Meaning of "Our initial implementation, misguided by types, ..."

We are referring to a classic sort of mistake when programming with dependent types: calling it a day when a routine type-checks, even before it has been proven correct.
We were lulled into a false sense of security by the width-tracking dependent types of bitvectors, in a tricky loop with a surprising design space that the crypto literature hadn't spelled out.

## Extending our results to cover signatures and cyphers, not just lower-level ECC operations

Actually, we have composed the operations into more complex schemes (ed25519) in Gallina, but we haven't formalized extracting C function calls because we eventually plan to cut out C and output straight to assembly.
Still, we used our generated field-arithmetic code as subroutines in complete Gallina (Coq) programs for signing and signature verification, which we proved equivalent to their simple specifications.
So carrying out this exercise ensures that we chose the right formal interface for the code that makes up the main novelty of our project.
(We mention this aspect in the paragraph with line 397.)

## New work that needs to be done for each new curve

For each new *point format*, it is necessary to write a new correctness proof, although the proof would almost certainly be a single invocation of `fsatz`.
The library already includes code and equivalence proofs for affine, XYZT, and Niels variants of Edwards coordiantes; affine, Jacobian, and Projective Weierstrass coordinates; and affine and XZ Montgomery coordinates.
When reusing an existing point format for a new curve or a new hardware architecture for an existing curve, the new implementation work is roughly just in the configuration snippets shown near the beginning of our Experimental Results section.

## Why not implement and benchmark more operations than just for 25519 and P256?

P256 and X25519 make up the vast majority of ECC usage on the Internet.
For example TLS ECDHE was measured to be 97% P256 in 2015[1]; since then X25519 has been deployed, reaching around 30% of connections established by Google Chrome.
We find it much more valuable to increase the depth of our verification (we already use higher-level specs than in any other work we know of, but there is still room to extend towards assembly) and performance of the code (fixing up the low-level inefficiencies pointed out by the reviewers and ourselves) than work on less commonly used curves.
If we wanted to further demonstrate the flexibility of the framework, it would be more meaningful to create different implementations of X25519 and P256, because that is where the performance engineering research is focused.
We could, of course, use the existing library to generate code for other curves (e.g. P384), but since we (and most libraries we know) would not prioritize their performance, the benchmarks would be less revealing.
Once we have a functioning pipeline to assembly code, we might seek to imitate the x25519-amd64-64 implementation that is fastest in our benchmarks but awkward to represent as C code.
However, since we intentionally chose our implementation strategies for X25519 and P256 to be as different as possible, every technique we know for implementing these curves (or any curves at a similar security level) would be an interpolation between the tricks used in our saturated-Montgomery-multiplication P256 code and unsaturated-Solinas-reduction X25519 code (e.g. x25519-amd64-64 uses saturated arithmetic with Solinas reduction).
New optimizations like specialized Karatsuba multiplication used for Goldilocks primes (e.g. X448) would need to be added and proven once as high-level optimizations, but again, we see expanding the work in other ways as more practically valuable at this moment.

[1]: "TLS in the wild: An Internet-wide analysis of TLS-based protocols for electronic communication" https://www.net.in.tum.de/fileadmin/bibtex/publications/papers/2016-tls-ndss.pdf

## Explanation of local-variable bitwidth inference

Sorry for the possibly cryptic description "abstract interpretation with intervals"!
We do think that (with the exception of bitwise operators and their sometimes-not-obvious conservative rules) we follow the abstract-interpretation literature pretty closely, even avoiding the complexities of code with interesting control flow, since we only analyze straightline code.
Here's another try at a summary: the analysis starts with some lower and upper bound for each variable, and applies sound (but not necessarily complete) operation-specific rules to derive bounds for further variables defined in the program, computing each's bound as a simple function of bounds for its operands.
Each variable is assigned the fastest integer type that encompasses its full computed interval.
This transformation and analysis is proven to preserve the output value for all code.

### 2020-07-30 papers/source/fiat-crypto/fiat-crypto.tex
Add a bunch of paper sources

\documentclass[conference,letterpaper]{IEEEtran}
%\overfullrule=1mm
\bibliographystyle{IEEEtran}

\usepackage[utf8]{inputenc}
\usepackage{newunicodechar}
\newunicodechar{λ}{$\lambda$}
\newunicodechar{∀}{$\forall$}
\newunicodechar{⌊}{$\lfloor$}
\newunicodechar{⌋}{$\rfloor$}

\usepackage{cite}
\usepackage{etex}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{url}
\usepackage{xcolor}
\usepackage{ifthen}
\usepackage{textcomp}
\usepackage{makecell}
\renewcommand{\cellalign}{cl}
\usepackage{balance}

%% Some recommended packages.
\usepackage{listings}
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

% packages for the carry-chain diagram
\usepackage{wrapfig}
\usepackage[all, frame]{xy}
\usepackage{mathtools}
\include{digitboxed}

\usepackage{tikz}
\usepackage{tkz-graph}
\usetikzlibrary{shapes.misc,positioning,arrows,snakes,backgrounds}
\usepackage{bm}

\usepackage{pgfplots}
\usepackage{xtab}

\newcommand{\kw}[1]{\mathsf{#1}}
\newcommand{\id}[1]{\bm{\mathsf{#1}}}
\newcommand{\nat}[0]{\mathbb N}
\newcommand{\intZ}[0]{\mathbb Z}
\newcommand{\floor}[1]{{\left \lfloor #1 \right \rfloor}}
\newcommand{\bunderline}[1]{\underline{#1\mkern-2mu}\mkern2mu }

% ~ godbolt.org coloring
\newcommand{\docolor}[2]{{\setlength{\fboxsep}{0pt}%
  \ifthenelse{\equal{#1}{A}}{\colorbox{green!15}{#2}}{%
  \ifthenelse{\equal{#1}{B}}{\colorbox{blue!15}{#2}}{%
  \ifthenelse{\equal{#1}{C}}{\colorbox{yellow!15}{#2}}{%
  \ifthenelse{\equal{#1}{D}}{\colorbox{red!15}{#2}}{%
    #1#2%
  }}}}%
}}%
{
\catcode`\!\active
\global\def\enableverbatimcolors{%
  \catcode`\!\active
  \def!##1[[##2]]{\docolor{##1}{##2}}%
}
}

\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO: #1}}}

\hyphenation{Comp-Cert}

\begin{document}

\title{Simple High-Level Code For Cryptographic Arithmetic -- With Proofs, Without Compromises}

\author{%
 \IEEEauthorblockN{Andres Erbsen \quad Jade Philipoom \quad Jason Gross \quad Robert Sloan \quad Adam Chlipala}
 \IEEEauthorblockA{%
   MIT CSAIL, \\
   Cambridge, MA, USA \\
   \{\texttt{andreser}, \texttt{jadep}, \texttt{jgross}\}\texttt{@mit.edu}, \texttt{rob.sloan@alum.mit.edu}, \texttt{adamc@csail.mit.edu} %TODO(@csail.mit.edu for all?)
 }
}

\maketitle

\begin{abstract}
  We introduce a new approach for implementing cryptographic arithmetic in short high-level code with machine-checked proofs of functional correctness.
  We further demonstrate that simple partial evaluation is sufficient to transform such initial code into the fastest-known C code,
  breaking the decades-old pattern that the only fast implementations are those whose instruction-level steps were written out by hand.

  These techniques were used to build an elliptic-curve library that achieves competitive performance for 80 prime fields and multiple CPU architectures, showing that implementation and proof effort scales with the number and complexity of conceptually different algorithms, not their use cases.
  As one outcome, we present the first verified high-performance implementation of P-256, the most widely used elliptic curve.
  Implementations from our library were included in BoringSSL to replace existing specialized code, for inclusion in several large deployments for Chrome, Android, and CloudFlare.
\end{abstract}

\IEEEpeerreviewmaketitle

\section{Motivation}

The predominant practice today is to implement cryptographic primitives at the lowest possible level of abstraction: usually flat C code, ideally assembly language if engineering time is plentiful.
Low-level implementation has been the only way to achieve good performance, ensure that execution time does not leak secret information, and enable the code to be called from any programming language.
Yet placing the burden of every detail on the programmer also leaves a lot to be desired.
The demand for implementation experts' time is high, resulting in an ever-growing backlog of algorithms to be reimplemented in specialized code ``as soon as possible.''
Worse, bugs are also common enough that it is often not economical to go beyond fixing the behavior for problematic inputs and understand the root cause and impact of the defect.

There is hope that these issues can eventually be overcome by redoubling community efforts to produce high-quality crypto code.
However, even the most renowned implementors are not infallible, and if a typo among thousands of lines of monotonous code can only be recognized as incorrect by top experts,\footnote{We refer to the \texttt{ed25519-amd64-64-24k} bug that is discussed along with other instructive real-world implementation defects in Appendix \ref{bugsapp}.} it is reasonable to expect diminishing returns for achieving correctness through manual review.
A number of tools for computer-aided verification have been proposed~\cite{verif25519,gfverif,ECCstar,hacl,vale,jasmin}, enabling developers to eliminate incorrect-output bugs conclusively.
Three solid implementations of the X25519 Diffie-Hellman function were verified.
In the beautiful world where X25519 was the only arithmetic-based crypto primitive we need, now would be the time to declare victory and go home.
Yet most of the Internet still uses P-256, and the current proposals for post-quantum cryptosystems are far from Curve25519's combination of performance and simplicity.

A closer look into the development and verification process (with an eye towards replicating it for another function) reveals a discouraging amount of duplicated effort.
First, expert time is spent on mechanically writing out the instruction-level steps required to perform the desired arithmetic, drawing from experience and folklore.
Then, expert time is spent to optimize the code by applying a carefully chosen combination of simple transformations.
Finally, expert time is spent on annotating the now-munged code with assertions relating values of run-time variables to specification variables, until the SMT solver (used as a black box) is convinced that each assertion is implied by the previous ones.
As the number of steps required to multiply two numbers is quadratic in the size of those numbers, effort grows accordingly: for example ``Verifying Curve25519 Software''~\cite{verif25519} reports that multiplication modulo $2^{255}-19$ required 27 times the number of assertions used for $2^{127}-1$.
It is important to note that both examples use \emph{the same} simple modular multiplication algorithm, and the same correctness argument should apply, just with different parameters.
This is just one example of how manual repetition of algorithmic steps inside primitive cryptographic arithmetic implementations makes them notoriously difficult to write, review, or prove correct.

Hence we suggest a rather different way of implementing cryptographic primitives:
to implement general algorithms in the style that best illustrates their correctness arguments
and then derive parameter-specific low-level code through partial evaluation.
For example, the algorithm used for multiplication modulo $2^{127}-1$ and $2^{255}-19$ is first encoded as a purely functional program operating on lists of digit-weight pairs: it generates a list of partial products, shifts down the terms with weight greater than the modulus, and scales them accordingly.
It is then instantiated with lists of concrete weights and length, for example ten limbs with weights $2^{\lceil 25.5 i\rceil}$, and partially evaluated so that no run-time manipulation of lists and weights remains, resulting in straight-line code very similar to what an expert would have written.
As the optimization pipeline is proven correct once and for all, correctness of all code generated from a high-level algorithm can be proven directly on its high-level implementation.
A nice side benefit of this approach is that the same high-level algorithm and proof can be used for multiple target configurations with very little effort, enabling the use of fast implementations in cases that were previously not worth the effort.
Finally, the generated code is compiled with a conventional C compiler.
As current compiler support for different CPU features varies wildly in optimization quality, this step can limit the performance of our code (especially when the best compilation would make careful use of carry flags or vector instructions).
We aim for the fastest code in terms of ALU functions, but we leave instruction selection, scheduling, and register allocation out of scope.

In practice, we use the fast arithmetic implementations reported on here in conjunction with a library of verified cryptographic algorithms, the rest of which deviates much less from common implementation practice.
All formal reasoning is done in the Coq proof assistant, and the overall trusted computing base also includes a simple pretty-printer and the C language toolchain.
The development, called \textbf{Fiat Cryptography}, is available under the MIT license at:
\begin{center}
  \texttt{\url{https://github.com/mit-plv/fiat-crypto}}
\end{center}

The approach we advocate for is already used in widely deployed code.
Most applications relying on the implementations described in this paper are using them through BoringSSL, Google's OpenSSL-derived crypto library that includes our Curve25519 and P-256 implementations.
A high-profile user is the Chrome Web browser, so today about half of HTTPS connections opened by Web browsers worldwide use our fast verified code (Chrome versions since 65 have about 60\% market share~\cite{browserMarketShare}, and 90\% of connections use X25519 or P-256~\cite{davidenTLSECC}).
BoringSSL is also used in Android and on a majority of Google's servers, so a Google search from Chrome would likely be relying on our code on both sides of the connection.
Maintainers appreciate the reduced need to worry about bugs.

\begin{figure}
\footnotesize
\enableverbatimcolors
\textbf{Input:}
\begin{verbatim}
Definition wwmm_step A B k S ret :=
  divmod_r_cps A (λ '(A, T1),
  !A[[@scmul_cps r _ T1 B _ (λ aB,]]
  !B[[@add_cps r _ S aB _ (λ S,]]
  divmod_r_cps S (λ '(_, s),
  mul_split_cps' r s k (λ '(q, _),
  !C[[@scmul_cps r _ q M _ (λ qM,]]
  !D[[@add_longer_cps r _ S qM _ (λ S,]]
  divmod_r_cps S (λ '(S, _),
  @drop_high_cps (S R_numlimbs) S _ (λ S,
  ret (A, S) ))))))))).
\end{verbatim}
\begin{verbatim}
Fixpoint wwmm_loop A B k len_A ret :=
  match len_A with
  | O => ret
  | S len_A' => λ '(A, S),
    wwmm_step A B k S (wwmm_loop A B k len_A' ret)
  end.
\end{verbatim}
\textbf{Output:}
\begin{verbatim}
void wwmm_p256(u64 out[4], u64 x[4], u64 y[4]) {
  !A[[u64 x17, x18 = mulx_u64(x[0], y[0]);]]
  !A[[u64 x20, x21 = mulx_u64(x[0], y[1]);]]
  !A[[u64 x23, x24 = mulx_u64(x[0], y[2]);]]
  !A[[u64 x26, x27 = mulx_u64(x[0], y[3]);]]
  !B[[u64 x29, u8 x30 = addcarryx_u64(0x0, x18, x20);]]
  !B[[u64 x32, u8 x33 = addcarryx_u64(x30, x21, x23);]]
  !B[[u64 x35, u8 x36 = addcarryx_u64(x33, x24, x26);]]
  !B[[u64 x38, u8 _ = addcarryx_u64(0x0, x36, x27);]]
  !C[[u64 x41, x42 = mulx_u64(x17, 0xffffffffffffffff);]]
  !C[[u64 x44, x45 = mulx_u64(x17, 0xffffffff);]]
  !C[[u64 x47, x48 = mulx_u64(x17, 0xffffffff00000001);]]
  !D[[u64 x50, u8 x51 = addcarryx_u64(0x0, x42, x44);]]
\end{verbatim}
  \vspace{-1.8mm}
\verb|  // |\textbf{\texttt{100 more lines...}}
  \caption{\label{top-level-example} Word-by-word Montgomery Multiplication, output specialized to P-256 ($2^{256} - 2^{224} + 2^{192} + 2^{96} - 1$)}
\end{figure}

\section{\label{overview}Overview}

We will explain how we write algorithm implementations in a high-level way that keeps the structural principles as clear and readable as possible, which facilitates proving functional correctness.
Our explanation builds up to showing how other security constraints can be satisfied without sacrificing the appeal of our new strategy.
Our output code follows industry-standard ``constant-time'' coding practices and is only bested in throughput by platform-specific assembly-language implementations (and only by a modest margin).
This section will give a tour of the essential ingredients of our methodology, and the next section will explain in detail how to implement field arithmetic for implementation-friendly primes like $2^{255}-19$.
Building on this intuition (and code!), Section \ref{saturated} culminates in a specializable implementation of the word-by-word Montgomery reduction algorithm for multiplication modulo any prime.
Section \ref{bounds} will explain the details of the compiler that turns the described implementations into the kind of code that implementation experts write by hand for each arithmetic primitive.
We report performance measurements in Section \ref{experiments} and discuss related work and future directions in Sections \ref{related} and \ref{conclusion}.

Fig. \ref{top-level-example} gives a concrete example of what our framework provides.
The algorithm in the top half of the figure is a transcription of the word-by-word Montgomery reduction algorithm, following Gueron's Algorithm 4~\cite{montexp2011}.
As is required by the current implementation of our optimization pipeline, this parameter-agnostic code is written in \emph{continuation-passing style} where functions do not return normally but instead take as arguments callbacks to invoke when answers become available.
The notation \mbox{\small\texttt{λ '(x, y), ...}} encodes an anonymous function with arguments \texttt{\small x} and \texttt{\small y}.

The bottom half of the figure shows the result of specializing the general algorithm to P-256, the elliptic curve most commonly used in TLS and unfortunately not supported by any of the verified libraries mentioned earlier.
Note that some operations influence the generated code by creating explicit instruction sequences (highlighted in matching colors), while other operations instead control which input expressions should be fed into others.
The number of low-level lines per high-level line varies between curves and target hardware architectures.
The generated code is accompanied by a Coq proof of correctness stated in terms of the straightline-code language semantics and the Coq standard-library definitions of integer arithmetic.

\subsection{High-Level Arithmetic Library Designed For Specialization}\label{ex-basesys}

At the most basic level, all cryptographic arithmetic algorithms come down to implementing large mathematical objects (integers or polynomials of several hundred bits) in terms of arithmetic provided by the hardware platform (for example integers modulo $2^{64}$).
To start out with, let us consider the example of addition, with the simplifying precondition that no digits exceed half of the hardware-supported maximum.

\noindent {\small \begin{align*}
  &\kw{type} \; \id{num} = \id{list} \; \nat \\
  &\id{add} : \id{num} \to \id{num} \to \id{num} \\
  &\id{add} \; (a :: as) \; (b :: bs) = \kw{let} \; x = a + b \; \kw{in} \; x :: \id{add} \; as \; bs \\
  &\id{add} \; as \; [] = as \\
  &\id{add} \; [] \; bs = bs
\end{align*}}%

\noindent To give the specification of this function acting on a base-$2^{64}$ little-endian representation, we define an abstraction function evaluating each digit list $\ell$ into a single number.
\noindent {\small \begin{align*}
  &\floor{\_} : \id{num} \to \nat \\
  &\floor{a::as} = a + 2^{64}\floor{as}\\
  &\floor{[]} = 0
\end{align*}}%
To prove correctness, we show (by naive induction on $a$) that evaluating the result of $\id{add}$ behaves as expected:
$$\forall a, b. \; \floor{\id{add} \; a \; b} = \floor{a} + \floor{b}$$

\noindent
Using the standard list type in Coq, the theorem becomes
{
\begin{verbatim}
Lemma eval_add : ∀a b, ⌊add a b⌋ = ⌊a⌋+⌊b⌋.
 induction a,b;cbn; rewrite ?IHa;nia. Qed.
\end{verbatim}
}

The above proof is only possible because we are intentionally avoiding details of the underlying machine arithmetic --
if the list of digits were defined to contain $64$-bit words instead of $\nat$, repeated application of $\id{add}$ would eventually result in integer overflow.
This simplification is an instance of the pattern of anticipating low-level optimizations in writing high-level code: we do expect to avoid overflow, and our choice of a digit representation is motivated precisely by that aim.
It is just that the proofs of overflow-freedom will be injected in a later stage of our pipeline, as long as earlier stages like our current one are implemented correctly.
There is good reason for postponing this reasoning: generally we care about the \emph{context} of higher-level code calling our arithmetic primitives.
In an arbitrary context, an $\id{add}$ implemented using $64$-bit words would need to propagate carries from each word to the next, causing an unnecessary slowdown when called with inputs known to be small.

\subsection{Partial Evaluation For Specific Parameters}\label{partial}

It is impossible to achieve competitive performance with arithmetic code that manipulates dynamically allocated lists at runtime.
The fastest code will implement, for example, a single numeric addition with straightline code that keeps as much state as possible in registers.
Expert implementers today write that straightline code manually, applying various rules of thumb.
Our alternative is to use \emph{partial evaluation} in Coq to generate all such specialized routines, beginning with a single library of high-level functional implementations that generalize the patterns lurking behind hand-written implementations favored today.

Consider the case where we know statically that each number we add will have 3 digits.
A particular addition in our top-level algorithm may have the form $\id{add} \; [a_1, a_2, a_3] \; [b_1, b_2, b_3]$, where the $a_i$s and $b_i$s are unknown program inputs.
While we cannot make compile-time simplifications based on the values of the digits, we \emph{can} reduce away all the overhead of dynamic allocation of lists.
We use Coq's term-reduction machinery, which allows us to choose $\lambda$-calculus-style reduction rules to apply until reaching a normal form.
Here is what happens with our example, when we ask Coq to leave $\kw{let}$ and $+$ unreduced but apply other rules.

\noindent {\small \begin{eqnarray*}
  \id{add} \; [a_1, a_2, a_3] \; [b_1, b_2, b_3] &\Downarrow& \kw{let} \; n_1 = a_1 + b_1 \; \kw{in} \; n_1 :: \\
  && \kw{let} \; n_2 = a_2 + b_2 \; \kw{in} \; n_2 :: \\
  && \kw{let} \; n_3 = a_3 + b_3 \; \kw{in} \; n_3 :: []
\end{eqnarray*}}%

\noindent We have made progress: no run-time case analysis on lists remains.
Unfortunately, $\kw{let}$ expressions are intermixed with list constructions, leading to code that looks rather different than assembly.
To persuade Coq's built-in term reduction to do what we want, we first translate arithmetic operations to \emph{continuation-passing style}.
Concretely, we can rewrite $\id{add}$.

\noindent {\small \begin{align*}
  &\id{add'} : \forall \alpha. \; \id{num} \to \id{num} \to (\id{num} \to \alpha) \to \alpha \\
  &\id{add'} \; (a :: as) \; (b :: bs) \; k =\\
  &\hspace{.1in} \kw{let} \; n = a + b \; \kw{in} \\
  &\hspace{.1in} \id{add'} \; as \; bs \; (\lambda \ell. \; k \; (n :: \ell)) \\
  &\id{add'} \; as \; [] \; k = k \; as \\
  &\id{add'} \; [] \; bs \; k = k \; bs
\end{align*}}%

\noindent Reduction turns this function into straightline code.

\noindent {\small \begin{eqnarray*}
  \id{add'} \; [a_1, a_2, a_3] \; [b_1, b_2, b_3] \; (\lambda \ell. \; \ell) &\Downarrow& \kw{let} \; n_1 = a_1 + b_1 \; \kw{in} \\
  && \kw{let} \; n_2 = a_2 + b_2 \; \kw{in} \\
  && \kw{let} \; n_3 = a_3 + b_3 \; \kw{in} \\
  && [n_1, n_2, n_3]
\end{eqnarray*}}%

\noindent When $\id{add'}$ is applied to a particular continuation, we can also reduce away the result list.
Chaining together sequences of function calls leads to idiomatic and efficient straightline code, based just on Coq's normal term reduction, preserving sharing of $\kw{let}$-bound variables.
This level of inlining is common for the inner loops of crypto primitives, and it will also simplify the static analysis described in the next subsection.

\begin{figure*}
  \begin{tabular}{|p{4cm}|l|l|l|l|}
\hline
    prime & architecture & \# limbs & base & representation (distributing large $x$ into $x_0...x_n$)\\
\hline
    $2^{256} - 2^{224} + 2^{192} + 2^{96} - 1$ (P-256) & 64-bit & 4 & $2^{64}$ & $x = x_0 + 2^{64}x_1 + 2^{128}x_2 + 2^{192}x_3$ \\
\hline
  $2^{255}-19$ (Curve25519) & 64-bit & 5 & $2^{51}$ & $x = x_0 + 2^{51}x_1 + 2^{102}x_2 + 2^{153}x_3 + 2^{204}x_4$ \\
\hline
  $2^{255}-19$ (Curve25519) & 32-bit & 10 & $2^{25.5}$ & $x = x_0 + 2^{26}x_1 + 2^{51}x_2 + 2^{77}x_3 + ... + 2^{204}x_8 + 2^{230}x_9$ \\
\hline
  $2^{448} - 2^{224} - 1$ (p448) & 64-bit & 8 & $2^{56}$ & $x = x_0 + 2^{56}x_1 + 2^{112}x_2 + ... + 2^{392}x_7$ \\
\hline
    $2^{127} - 1$ & 64-bit & 3 & $2^{42.5}$ & $x = x_0 + 2^{43}x_1 + 2^{85}x_2$ \\
\hline
\end{tabular}
  \caption{\label{fig:representationexamples}Examples of big-integer representations for different primes and integer widths}
\end{figure*}

\subsection{Word-Size Inference}\label{ex-bounds}

Up to this point, we have derived code that looks almost exactly like the C code we want to produce.
The code is structured to avoid overflows when run with fixed-precision integers, but so far it is only proven correct for natural numbers.
The final major step is to infer a range of possible values for each variable, allowing us to assign each one a register or stack-allocated variable of the appropriate bit width.

The bounds-inference pass works by standard abstract interpretation with intervals.
As inputs, we require lower and upper bounds for the integer values of all arguments of a function.
These bounds are then pushed through all operations to infer bounds for temporary variables.
Each temporary is assigned the smallest bit width that can accommodate its full interval.

As an artificial example, assume the input bounds $a_1, a_2, a_3, b_1 \in [0, 2^{31}]$; $b_2, b_3 \in [0, 2^{30}]$.
The analysis concludes $n_1 \in [0, 2^{32}]$; $n_2, n_3 \in [0, 2^{30} + 2^{31}]$.
The first temporary is just barely too big to fit in a 32-bit register, while the second two will fit just fine.
Therefore, assuming the available temporary sizes are 32-bit and 64-bit, we can transform the code with precise size annotations.

\noindent {\small \begin{align*}
  &\kw{let} \; n_1 : \nat_{2^{64}} = a_1 + b_1 \; \kw{in} \\
  &\kw{let} \; n_2 : \nat_{2^{32}} = a_2 + b_2 \; \kw{in} \\
  &\kw{let} \; n_3 : \nat_{2^{32}} = a_3 + b_3 \; \kw{in} \\
  &[n_1, n_2, n_3]
\end{align*}}%

\noindent Note how we may infer different temporary widths based on different bounds for arguments.
As a result, the same primitive inlined within different larger procedures may get different bounds inferred.
World-champion code for real algorithms takes advantage of this opportunity.

This phase of our pipeline is systematic enough that we chose to implement it as a certified compiler.
That is, we define a type of abstract syntax trees (ASTs) for the sorts of programs that earlier phases produce, we reify those programs into our AST type, and we run compiler passes written in Coq's Gallina functional programming language.
Each pass is proved correct once and for all, as Section \ref{bounds} explains in more detail.

\subsection{Compilation To Constant-Time Machine Code}\label{ex-asm}

What results is straightline code very similar to that written by hand by experts, represented as ASTs in a simple language with arithmetic and bitwise operators.
Our correctness proofs connect this AST to specifications in terms of integer arithmetic, such as the one for $\id{add}$ above.
All operations provided in our lowest-level AST are implemented with input-independent execution time in many commodity compilers and processors, and if so, our generated code is trivially free of timing leaks.
Each function is pretty-printed as C code and compiled with a normal C compiler, ready to be benchmarked or included in a library.
We are well aware that top implementation experts can translate C to assembly better than the compilers, and we do not try to compete with them: while better instruction scheduling and register allocation for arithmetic-heavy code would definitely be valuable, it is outside the scope of this project.
Nevertheless, we are excited to report that our library generates the fastest-known C code for all operations we have benchmarked (Section \ref{experiments}).

\begin{figure*}
\begin{alignat*}{5}
    s \times t & = 1 \times s_0t_0 && + 2^{43} \times s_0t_1 && + 2^{85} \times s_0t_2 \\
    & && + 2^{43} \times s_1t_0 && + 2^{86} \times s_1t_1 && + 2^{128} \times s_1t_2 \\
    & && && + 2^{85} \times s_2t_0 && + 2^{128} \times s_2t_1 && + 2^{170} \times s_2t_2 \\
    & = s_0t_0 && + 2^{43}(s_0t_1 + s_1t_0) && + 2^{85}(s_0t_2 + 2s_1t_1 + s_2t_0) && + 2^{127}(2s_1t_2 + 2s_2t_1) && + 2^{170}\times s_2t_2
\end{alignat*}
\caption{\label{distribute}Distributing terms for multiplication mod $2^{127}-1$}
\end{figure*}

\section{Arithmetic Template Library}\label{basesystems}

Recall Section \ref{ex-basesys}'s toy $\id{add}$ function for little-endian big integers.
We will now describe our full-scale library, starting with the core \emph{unsaturated} arithmetic subset, the foundation for all big-integer arithmetic in our development.
For those who prefer to read code, we suggest \texttt{src/Demo.v} in the framework's source code, which contains a succinct standalone development of the unsaturated-arithmetic library up to and including prime-shape-aware modular reduction.
The concrete examples derived in this section are within established implementation practice, and an expert would be able to reproduce them given an informal description of the strategy.
Our contribution is to encode this general wisdom in concrete algorithms and provide these with correctness certificates without sacrificing the programmer's sanity.

\subsection{Multi-Limbed Arithmetic}\label{unsaturated}

Cryptographic modular arithmetic implementations distribute very large numbers across smaller ``limbs'' of 32- or 64-bit integers.
Fig. \ref{fig:representationexamples} shows a small sample of fast representations for different primes.
Notice that many of these implementations use bases other than $2^{32}$ or $2^{64}$, leaving bits unused in each limb: these are called \emph{unsaturated} representations.
Conversely, the ones using all available bits are called \emph{saturated}.

Another interesting feature shown in the examples is that the exponents of some
bases, such as the original 32-bit Curve25519 representation~\cite{curve25519}, are not whole numbers.
In the actual representation, this choice corresponds to an alternating pattern, so ``base $2^{25.5}$'' uses 26 bits in the first limb, 25 in the second, 26 in the third, and so on.
Because of the alternation, these are called \emph{mixed-radix} bases, as opposed to \emph{uniform-radix} ones.

These unorthodox big integers are fast primarily because of a specific modular-reduction trick, which is most efficient when the number of bits in the prime corresponds to a limb boundary.
For instance, reduction modulo $2^{255}-19$ is fastest when the bitwidths of a prefix of limbs sum to 255.
Every unsaturated representation in our examples is designed to fulfill this criterion.

\subsection{Partial Modular Reduction}
\label{solinasreduce}

Suppose we are working modulo a $k$-bit prime $m$.
Multiplying two $k$-bit numbers can produce up to $2k$ bits, potentially much larger than $m$.
However, if we only care about what the result is mod $m$, we can perform a \emph{partial modular reduction} to reduce the upper bound while preserving modular equivalence.
(The reduction is ``partial'' because the result is not guaranteed to be the minimal residue.)

The most popular choices of primes in elliptic-curve cryptography are of the form $m = 2^k - c_l 2^{t_l} - \ldots - c_0 2^{t_0}$, encompassing what have been called ``generalized Mersenne primes,'' ``Solinas primes,'' ``Crandall primes,'' ``pseudo-Mersenne primes,'' and ``Mersenne primes.''
Although any number could be expressed this way, and the algorithms we describe would still apply, choices of $m$ with relatively few terms ($l \ll k$) and small $c_i$ facilitate fast arithmetic.

Set $s=2^k$ and $c = c_l 2^{t_l} + \ldots + c_0 2^{t_0}$, so $m = s - c$.
To reduce $x$ mod $m$, first \emph{split} $x$ by finding $a$ and $b$ such that $x = a + sb$.
Then, a simple derivation yields a division-free procedure for partial modular reduction.
% TODO(jgross): fix mod alignment?
\begin{align*}
x \bmod m &= (a + sb) \bmod (s - c) \\
                                             % &= (a(s-c+c) + b) \bmod (s - c) \\
         &= (a + (s-c)b + cb) \bmod (s - c) \\
         &= (a + cb) \bmod m
\end{align*}

The choice of $a$ and $b$ does not further affect the correctness of this formula, but it does influence how much the input is reduced: picking $a=x$ and $b=0$ would make this formula a no-op.
One might pick $a = x \bmod s$, although the formula does not require it.
Here is where careful choices of big-integer representation help: having $s$ be at a limb boundary allows for a good split without any computation!

Our Coq proof of this trick is reproduced here:
{%\footnotesize
\begin{verbatim}
Lemma reduction_rule a b s c (_: s-c <> 0)
 : (a + s * b) mod (s - c) =
   (a + c * b) mod (s - c).
Proof.
 replace (a+s*b) with ((a+c*b)+b*(s-c)).
 rewrite add_mod,mod_mult,add_0_r,mod_mod.
 all: auto; nsatz.
Qed.
\end{verbatim}}

\subsection{Example: Multiplication Modulo $2^{127}-1$}

Before describing the general modular-reduction algorithm implemented in our library, we will walk through multiplication specialized to just one modulus and representation.
To simplify matters a bit, we use the (relatively) small modulus $2^{127} - 1$.
Say we want to multiply 2 numbers $s$ and $t$ in its field, with those inputs broken up as $s = s_0 + 2^{43} s_1 + 2^{85} s_2$ and $t = t_0 + 2^{43} t_1 + 2^{85} t_2$.
Distributing multiplication repeatedly over addition gives us the answer form shown in Fig. \ref{distribute}.

We format the first intermediate term suggestively: down each column, the powers of two are very close together, differing by at most one.
Therefore, it is easy to add down the columns to form our final answer, split conveniently into digits with integral bit widths.
%The number of primitive addition and multiplication operations is reasonable (and the primitive multiplications can be implemented as self-additions, which are often faster in hardware).

At this point we have a double-wide answer for multiplication, and we need to do modular reduction to shrink it down to single-wide.
For our example, the last two digits can be rearranged so that the modular-reduction rule applies:
% N.B. We use \phantom{\vphantom{.}=\vphantom{.}} to make sure that ``='' is aligned with space on either side; the space only shows up when LaTeX registers a character before/after the character, so we fool it into thinking that there is a character there.
\begin{align*}
    &\phantom{\vphantom{.}=\vphantom{.}} 2^{127}(2s_1t_2 + 2s_2t_1) + 2^{170}s_2t_2 & \pmod{2^{127}-1} \\
    &= 2^{127}((2s_1t_2 + 2s_2t_1) + 2^{43}s_2t_2) & \pmod{2^{127}-1} \\
    &= 1((2s_1t_2 + 2s_2t_1) + 2^{43}s_2t_2) & \pmod{2^{127}-1}
\end{align*}

As a result, we can merge the second-last digit into the first and merge the last digit into the second, leading to this final formula specifying the limbs of a single-width answer:
\begin{align*}
  (s_0t_0 + 2s_1t_2 + 2s_2t_1) &+ 2^{43}(s_0t_1 + s_1t_0 + s_2t_2) \\
  &+ 2^{85}(s_0t_2 + 2s_1t_1 + s_2t_0)
\end{align*}

\subsection{Associational Representation}
% The Central Insight to Simplify Coding and Verifying Generic Arithmetic

As is evident by now, the most efficient code makes use of sophisticated and specific big-number representations, but many concrete implementations operate on the same set of underlying principles.
Our strategy of writing high-level templates allows us to capture the underlying principles in a fully general way, without committing to a run-time representation.
Abstracting away the implementation-specific details, like the exact number of limbs or whether the base system is mixed- or uniform-radix, enables simple high-level proofs of all necessary properties.

A key insight that allows us to simplify the arithmetic proofs is to use multiple different representations for big integers, as long as changes between these representations can be simplified away once the prime shape is known.
We have two main representations of big integers in the core library: \emph{associational} and \emph{positional}.
A big integer in associational form is represented as a list of pairs; the first element in each pair is a \emph{weight}, known at compile time, and the second is a runtime value.
The decimal number 95 might be encoded as \texttt{[(16, 5); (1, 15)]}, or equivalently as \texttt{[(1, 5); (10, 9)]} or \texttt{[(1, 1); (1, 6); (44, 2)]}.
As long as the sum of products is correct, we do not care about the order, whether the weights are powers of each other, or whether there are multiple pairs of equal weight.

In associational format, arithmetic operations are extremely easy to reason about.
Addition is just concatenation of the lists.
Schoolbook multiplication (see Fig. \ref{assocmul}) is also simple: $(a_1\cdot x_1 + \ldots)(b_1\cdot y_1 + \ldots) = (a_1b_1\cdot x_1y_1 + \ldots)$, where $a_1b_1$ is the new compile-time-known weight.
Because of the flexibility allowed by associational representation, we do not have to worry about whether the new compile-time weight produced is actually in the base.
If we are working with the base $2^{25.5}$ in associational representation, and we multiply two pairs that both have weight $2^{26}$, we can just keep the $2^{52}$ rather than the base's $2^{51}$ weight.

\begin{figure}
  %\footnotesize
  \begin{verbatim}
Definition mul (p q:list(Z*Z)) : list(Z*Z)
:= flat_map (λt,
    map (λt',(fst t*fst t', snd t*snd t'))
   q) p.
	\end{verbatim}
  \begin{verbatim}
Lemma eval_map_mul (a x:Z) (p:list (Z*Z))
  : eval (map (λt, (a*fst t, x*snd t)) p)
  = a*x * eval p.
Proof. induction p; push; nsatz.  Qed.
	\end{verbatim}
  \begin{verbatim}
Lemma eval_mul p q
  : eval (mul p q) = eval p * eval q.
Proof.
induction p; cbv [mul]; push; nsatz. Qed.
\end{verbatim}
  \caption{\label{assocmul}Definition and correctness proof of multiplication}
\end{figure}

Positional representation, on the other hand, uses lists of runtime values with weight functions.
There is only one runtime value for each weight, and order matters.
This format is closer to how a number would actually be represented at runtime.
Separating associational from positional representations lets us separate reasoning about arithmetic from reasoning about making weights line up.
Especially in a system that must account for mixed-radix bases, the latter step is nontrivial and is better not mixed with other reasoning.

\begin{figure}
  \begin{verbatim}
Fixpoint place (t:Z*Z) (i:nat) : nat*Z :=
 if dec (fst t mod weight i = 0) then
  let c:=fst t / weight i in (i, c*snd t)
 else match i with
      | S i' => place t i'
      | O => (O, fst t*snd t) end.
	\end{verbatim}
  \begin{verbatim}
Definition from_associational n p :=
  List.fold_right (λt,
      let wx := place t (pred n) in
      add_to_nth (fst wx) (snd wx) )
    (zeros n) p.
	\end{verbatim}
  \begin{verbatim}
Lemma place_in_range (t:Z*Z) (n:nat)
: (fst (place t n) < S n)%nat.
Proof.
 cbv [place]; induction n; break_match;
 autorewrite with cancel_pair; omega. Qed.
	\end{verbatim}
  \begin{verbatim}
Lemma weight_place t i
: weight (fst(place t i)) * snd(place t i)
  = fst t * snd t.
Proof. (* ... 4 lines elided ... *) Qed.
	\end{verbatim}
  \begin{verbatim}
Lemma eval_from_associational {n} p
: eval (from_associational (S n) p)
  = Associational.eval p.
Proof.
 cbv [from_associational]; induction p;
 push;
 try pose proof place_in_range a n;
 try omega; try nsatz.                Qed.
 \end{verbatim}
  \caption{\label{fromassoc}Coq definition and selected correctness proofs of conversion from associational to positional representation}
\end{figure}

Fig. \ref{fromassoc} shows the code and proofs for conversion from associational to positional format, including the fixing of misaligned weights.
The helper function \texttt{place} takes a pair $t = (w, x)$ (compile-time weight and runtime value) and recursively searches for the highest-index output of the weight function that $w$ is a multiple of.
After finding it, \texttt{place} returns an index $i$ and the runtime value to be added at that index: ($w / weight_i \cdot x$).
Then, \texttt{from\_associational} simply adds each limb at the correct index.

\begin{figure}
%\footnotesize
\begin{verbatim}
Definition split (s:Z) (p:list (Z*Z))
  : list(Z*Z)*list(Z*Z) :=
 let hl :=
   partition (λt, fst t mod s =? 0) p in
 (snd hl,
  map (λt, (fst t / s, snd t)) (fst hl)).
\end{verbatim}
\begin{verbatim}
Definition reduce(s:Z)(c:list _)(p:list _)
  : list(Z*Z) :=
 let lh := split s p in
 fst lh ++ mul c (snd lh).
\end{verbatim}
\begin{verbatim}
Lemma eval_split s p (s_nz:s<>0) :
    eval (fst (split s p))
     + s * eval (snd (split s p))
  = eval p.
Proof. (* ... (7 lines elided) *) Qed.
\end{verbatim}
\begin{verbatim}
Lemma eval_reduce s c p
       (s_nz:s<>0)(m_nz:s-eval c<>0) :
    eval (reduce s c p) mod (s - eval c)
  = eval p mod (s - eval c).
Proof. cbv [reduce]; push.
 rewrite<-reduction_rule,eval_split; auto.
Qed.
\end{verbatim}
\caption{\label{splitreduce}Coq code for prime-shape-sensitive modular reduction}
%TODO: if we need to cut a figure, I propose this one
\end{figure}

The modular-reduction trick we explained in Section \ref{solinasreduce} can also be succinctly represented in associational format (Fig. \ref{splitreduce}).
First we define \texttt{split}, which partitions the list, separating out pairs with compile-time weights that are multiples of the number at which we are splitting (\texttt{s}, in the code).
Then it divides all the weights that are multiples of \texttt{s} by \texttt{s} and returns both lists.
The \texttt{reduce} function simply multiplies the ``high'' list by \texttt{c} and adds the result to the ``low'' list.
Using the \texttt{mul} we defined earlier recreates the folklore modular-reduction pattern for prime shapes in which \texttt{c} is itself multi-limb.

% The original motivation behind saturated arithmetic is that the sum (or difference, or product) of two $s$-bit values no longer fits in a $s$-bit register: instead, the low and high half of the output are returned separately.
% However, the correspnding implementation techniques are not locked into operating on radix $w_i = 2^{si}$, they are also useful (for non-performance-critical operations) on the mixed-radix basesystems designed for efficient unsaturated arithmetic.
% Furthermore, the high level flows for saturated addition and subtraction do not require the inputs to be non-redundant, but nevertheless produce non-redundant output.
% This allows for a generic way of putting unsaturated representations of field elements into a canonical form, which can then be converted into a saturated representation with simple arithmetic and bitwise operations.
% In particular: it is sufficient to proceed in three steps.
% First, our code ensures that the limb $x_n$ with the highest weight $w_n$ is represented canonically by replacing it with two limbs $x_n \bmod w_n$ and $\lfloor x_n / w_n \rfloor$ and applying the modular reduction procedure described above to eliminate the limb at $w_{n+1}$.
% Assuming $x_i < w_{i+2}/W_i$, this establishes that the overall value is $\le 2w_n$ (we do not actually prove this in general, we run bounds analysis and then prove the specific inequality by \texttt{omega}).
% Second, our code subtracts the modulus $m$, producing a carry bit.
% Third, our code adds the modulus, producing a carry bit if and only if the previous subtraction did not produce one.
% The output limbs of the saturated operation that did not produce a carry bit (and thus represents the original value modulo $m$).
% We learned this technique from Mike Hamburg (todo:cite Goldilocks code); we would not surprised if it has been presented before.
%
% We wrote generic implementations of these algorithms, proved them correct, and generated C code for basesystems used in cryptography.
% However, even leaving aside verification, we would like to call the reader's attention to the fact that the above descriptions of the algorithms, unlike any implementation of them we have seen, do not include arithmetic on indices of limbs.

\subsection{Carrying}

In unsaturated representations, it is not necessary to carry immediately after every addition.
For example, with 51-bit limbs on a 64-bit architecture, it would take 13 additions to risk overflow.
Choosing which limbs to carry and when is part of the design and is critical for keeping the limb values bounded.
Generic operations are easily parameterized on carry strategies, for example ``after each multiplication carry from limb 4 to limb 5, then from limb 0 to limb 1, then from limb 5 to limb 6,'' etc.
The library includes a conservative default.

\subsection{Saturated Arithmetic}\label{saturated}

The code described so far was designed with unsaturated representations in mind.
However, unsaturated-arithmetic performance degrades rapidly when used with moduli that it is not a good match for, so older curves such as P-256 need to be implemented using saturated arithmetic.
Basic arithmetic on saturated digits is complicated by the fact that addition and multiplication outputs may not fit in single words, necessitating the use of two-output addition and multiplication instructions.
While the partial products in Fig. \ref{assocmul} also need to be stored in pairs of registers, the halves are never separated, avoiding the need to reason about their representation.
On the other hand, saturated-arithmetic algorithms can be seen as treating carry bits (and the low and high halves of a multiplication) as separate limbs.
This conceptual framework allows us to reuse most of our unsaturated-associational-arithmetic procedures on saturated representations while still generating the expected low-level code that differs significantly from that of unsaturated arithmetic.

\begin{figure}
  %\footnotesize
\begin{verbatim}
Definition from_associational
    n (p:list (list Z)) :=
 List.fold_right (λt,
  let p := place t (pred n) in
  cons_to_nth (fst p) (snd p)) (nils n) p.
\end{verbatim}
  \caption{\label{columnsfromassoc}Conversion from associational to columns}
\end{figure}

First, we write another version of \texttt{mul} that looks very similar to the unsaturated one but uses a two-output multiplication instruction to compute the partial product of $(a,x)$ and $(b,y)$ as \texttt{let xy := mul x y in [(a*b, fst xy); (a*b*bound, snd xy)]} instead of \texttt{[(a*b, x*y)]}.
%TODO: is the meaning of bound clear?

Second, rather than using the unsaturated \texttt{from\_associational}, we first convert to an intermediate representation we call \emph{columns}, named after the conventional layout of partial products during pencil-and-paper multiplication.
While associational representations are lists of pairs of integers (\texttt{list (Z * Z)}) and positional ones are lists of integers (\texttt{list Z}), columns representations are lists of lists of integers (\texttt{list (list Z)}).
Columns representations are very similar to positional ones, except that the values at each position have not yet been added together.
Starting from the columns representation, we can add terms together in a specific order so that carry bits are used right after they are produced, enabling them to be stored in flag registers in assembly code.
The code to convert from associational to columns is very similar to associational-to-positional conversion (Fig. \ref{columnsfromassoc}).

With these changes in place, much of the code shown before for associational works on positional representations as well.
For instance, the partial modular-reduction implementation still applies unchanged; using the new \texttt{from\_associational} is sufficient.
The takeaway here is that even completely changing the underlying hardware instructions we used for basic arithmetic did not require redoing all the work from unsaturated representations.

Our most substantial use of saturated arithmetic was for \emph{Montgomery modular reduction}, shown in full earlier in Fig. \ref{top-level-example}.
For moduli $m$ without convenient special structure, there is not much room to optimize the computation of $ab \bmod m$.
Instead, it pays off to keep all intermediate values $a$ in a scaled form $aR \bmod m$ for some parameter $R$ and instead optimize the computation of $(aR)(bR)R^{-1} \bmod m = (ab)R \bmod m$.
We will use underlines to denote the (unique) Montgomery forms of numbers: $\bunderline{a} = aR \bmod m$.

\begin{figure}
  %\footnotesize
\begin{verbatim}
Lemma S2_mod
: eval (sat_add S1 (scmul q m)) mod m
  = eval S1 mod m.
Proof. push; zsimplify; auto. Qed.
\end{verbatim}
  \caption{\label{wwmod}Proof that adjusting $S$ to be $0 \bmod m$ preserves the correct value $\bmod\, m$}
\end{figure}

First, consider the simple case where $S = \bunderline{a}\bunderline{b}$ is divisible by $R$: we can simply divide it by $R$, ignoring the modulus $m$ completely.
To make division by $R$ fast, we use a minimal saturated representation for $S$ and pick $R$ to match the weight of some limb.
This way, dividing by $R$ needs no code at all: simply forget about the less significant limbs!

Of course, $S$ is extremely unlikely to be divisible by $R$.
However, since we only care about the answer $\bmod\, m$, we can add a multiple of $m$ to adjust $S$ to be $0 \bmod R$.
In particular, we want to add $qm$ such that $qm \bmod R = -(S \bmod R)$.
Multiplying both sides by $m^{-1} \bmod R$ gives $q = (-m^{-1} \bmod R)(S \bmod R) \bmod R$, which is very cheap to compute: the first factor is a compile-time constant, and $S \bmod R$ is simply the limbs of $S$ below $R$.
The Coq proof of this step of our word-by-word implementation is in Fig. $\ref{wwmod}$.

\begin{figure}
  % TODO: get rid of this example?
  \setlength{\abovedisplayskip}{0pt}%
  \setlength{\belowdisplayskip}{0pt}%
  Given \mbox{\small $\texttt{eval}\ S < \texttt{eval}\ m + \texttt{eval}\ B$};\\
  To show: \par% must be \par not \\ for line spacing, see https://tex.stackexchange.com/a/142950/2066
  \small
  $\texttt{eval}\ (\texttt{div\_r}\ (\texttt{add}\ (\texttt{add}\ S\ (\texttt{scmul}\ a\ B))\ (\texttt{scmul}\ q\ m)))$
  \begin{align*}
    & < \texttt{eval}\ m + \texttt{eval}\ B \\
    (\texttt{eval}\ S + a*\texttt{eval}\ B + q*\texttt{eval}\ m)/r
    & < \texttt{eval}\ m + \texttt{eval}\ B \\
(\texttt{eval}\ m * r + r * \texttt{eval}\ B - 1)/r
    & < \texttt{eval}\ m + \texttt{eval}\ B \\
\texttt{eval}\ m + \texttt{eval}\ B - 1
    & < \texttt{eval}\ m + \texttt{eval}\ B
  \end{align*}
  \caption{\label{wwbound} Intermediate goals of the proof that the word-by-word Montgomery reduction state value remains bounded by a constant, generated by single-stepping our proof script}
\end{figure}

This procedure is further optimized based on the observation that it only acts on the higher limbs of $S$ additively.
For $n$-limb $\bunderline{a}$, instead of first computing $\bunderline{a}\bunderline{b}$ and then dividing it by $R=r^n$, it is fine to divide eagerly by $r$ after multiplying $\bunderline{b}$ with a single limb of $\bunderline{a}$ -- adding the partial products with the more significant limbs of $\bunderline{a}$ commutes with adding multiples of $m$.
Even better, adding a multiple of $m$ and a multiple of $\bunderline{b}$ can only increase the length of $S$ by one limb, balancing out the loss of one limb when dividing $S$ by $r$ and allowing for an implementation that loops through limbs of $\bunderline{a}$ while operating in-place on $S$.
The proof of the last fact is delicate -- for example, it is not true that one of the additions always produces a newly nonzero value in the most significant limb and the other never does.
In our library, this subtlety is handled by a 15-line/100-word proof script; the high-level steps, shown in Fig. \ref{wwbound}, admittedly fail to capture the amount of thought that was required to find the invariant that makes the bounds line up exactly right.

%% \subsection{The Story So Far}

%% Let us take stock of how the final packaging of this section's results fits into the overall story of the paper.
%% We have \emph{generic} implementations of key modular-arithmetic algorithms coded as functional programs in Coq.
%% Each operation is proved correct once and for all.
%% Moreover, the operations' input and output formats have been chosen to enforce a clear separation between compile-time (e.g., weight of a digit) and runtime (e.g., a digit itself) values, though nothing formal has been proved about that separation.
%% Instead, in the context of the top-level code for a primitive, where the compile-time values are apparent, we apply \emph{partial evaluation} as explained in Section \ref{partial}.
%% The result is, for each instance of an arithmetic operation, relatively simple first-order code.
%% That code is the input to our final compiler phases, described next.

\section{Certified Bounds Inference}\label{bounds}

Recall from Section \ref{partial} how we use partial evaluation to specialize the functions from the last section to particular parameters.
The resulting code is elementary enough that it becomes more practical to apply relatively well-understood ideas from \emph{certified compilers}.
That is, as sketched in Section \ref{ex-bounds}, we can define an explicit type of program abstract syntax trees (ASTs), write compiler passes over it as Coq functional programs, and prove those passes correct.
%once and for all. The contrast is with earlier transformation phases, which work using Coq's proof-generating tactics.

\subsection{Abstract Syntax Trees}

The results of partial evaluation fit, with minor massaging, into this intermediate language:
$$\begin{array}{rrcl}
  \textrm{Base types} & b \\
  \textrm{Types} & \tau &::=& b \mid \kw{unit} \mid \tau \times \tau \\
  \textrm{Variables} & x \\
  \textrm{Operators} & o \\
  \textrm{Expressions} & e &::=& x \mid o(e) \mid () \mid (e, e) \\
  &&& \mid \kw{let} \; (x_1, \ldots, x_n) = e \; \kw{in} \; e
\end{array}$$

Types are trees of pair-type operators $\times$ where the leaves are one-element $\kw{unit}$ types and base types $b$, the latter of which come from a domain that is a parameter to our compiler.
It will be instantiated differently for different target hardware architectures, which may have different primitive integer types.
When we reach the certified compiler's part of the pipeline, we have converted earlier uses of lists into tuples, so we can optimize away any overhead of such value packaging.

Another language parameter is the set of available primitive operators $o$, each of which takes a single argument, which is often a tuple of base-type values.
Our $\kw{let}$ construct bakes in destructuring of tuples, in fact using typing to ensure that all tuple structure is deconstructed fully, with variables bound only to the base values at a tuple's leaves.
Our deep embedding of this language in Coq uses dependent types to enforce that constraint, along with usual properties like lack of dangling variable references and type agreement between operators and their arguments.

Several of the key compiler passes are polymorphic in the choices of base types and operators, but bounds inference is specialized to a set of operators.
We assume that each of the following is available for each type of machine integers (e.g., 32-bit vs. 64-bit).
$$\begin{array}{@{}l@{}}
  \textrm{Integer literals: } n \\
  \textrm{Unary arithmetic operators: } -e \\
  \textrm{Binary arithmetic operators: } e_1 + e_2, e_1 - e_2, e_1 \times e_2 \\
  \textrm{Bitwise operators: } e_1 \ll e_2, e_1 \gg e_2, e_1 \; \& \; e_2, e_1 \mid e_2 \\
  \textrm{Conditionals: } \mathsf{if} \; e_1 \neq 0 \; \mathsf{then} \; e_2 \; \mathsf{else} \; e_3 \\
  \textrm{Carrying: } \mathsf{addWithCarry}(e_1, e_2, c), \mathsf{carryOfAdd}(e_1, e_2, c) \\
  \textrm{Borrowing: } \mathsf{subWithBorrow}(c, e_1, e_2), \mathsf{borrowOfSub}(c, e_1, e_2) \\
  \textrm{Two-output multiplication: } \mathsf{mul2}(e_1, e_2)
%| IdWithAlt T1 T2 Tout : op (Tbase T1 * Tbase T2) (Tbase Tout)
\end{array}$$

We explain only the last three categories, since the earlier ones are familiar from C programming.
%TODO: agl drew a \rfloor symbol here, idk why
To chain together multiword additions, as discussed in the prior section, we need to save overflow bits (i.e., carry flags) from earlier additions, to use as inputs into later additions.
The $\mathsf{addWithCarry}$ operation implements this three-input form, while $\mathsf{carryOfAdd}$ extracts the new carry flag resulting from such an addition.
Analogous operators support subtraction with \emph{borrowing}, again in the grade-school-arithmetic sense.
Finally, we have $\mathsf{mul2}$ to multiply two numbers to produce a two-number result, since multiplication at the largest available word size may produce outputs too large to fit in that word size.

All operators correspond directly to common assembly instructions.
Thus the final outputs of compilation look very much like assembly programs, just with unlimited supplies of temporary variables, rather than registers.
$$\begin{array}{rrcl@{}}
  \textrm{Operands} & O &::=& x \mid n \\
  \textrm{Expressions} & e &::=& \kw{let} \; (x_1, \ldots, x_n) = o(O, \ldots, O) \; \kw{in} \; e \\
  &&& \mid (O, \ldots, O)
\end{array}$$

We no longer work with first-class tuples.
Instead, programs are sequences of primitive operations, applied to constants and variables, binding their (perhaps multiple) results to new variables.
A function body, represented in this type, ends in the function's (perhaps multiple) return values.

Such functions are easily pretty-printed as C code, which is how we compile them for our experiments.
Note also that the language enforces the \emph{constant time} security property by construction: the running time of an expression leaks no information about the values of the free variables.
We do hope, in follow-on work, to prove that a particular compiler preserves the constant-time property in generated assembly.
For now, we presume that popular compilers like GCC and Clang do preserve constant time, modulo the usual cycle of occasional bugs and bug fixes.
(One additional source-level restriction is important, forcing conditional expressions to be those supported by native processor instructions like conditional move.)

\subsection{Phases of Certified Compilation}

%% \begin{figure*}
%%   \centering
%%   \tikzstyle{rf} = [rectangle, draw=black, inner sep=3pt]
%%   \tikzstyle{rrf} = [rounded rectangle, draw=black, inner sep=3pt]
%%   \tikzstyle{line}=[-]
%%   \tikzstyle{arrow}=[<-]
%%   \begin{tikzpicture}
%%     \node[anchor=south] (root) at (0, 0) [rf] {
%%       Gallina
%%     };
%%     \node[anchor=south] (n1) at (1.9, 0) [rf] {
%%       AST
%%     };
%%     \node[anchor=south] (n2) at (4.15, 0) [rf] {
%%       flat
%%     };
%%     \node[anchor=south] (n3) at (7.9, 0) [rf] {
%%       flat
%%     };
%%     \node[anchor=south] (n4) at (10.00, 0) [rf] {
%%       flat
%%     };
%%     \node[anchor=south] (n5) at (11.8, 0) [rf] {
%%       flat
%%     };

%%     \node[anchor=east] (n1exp) at (1.5, 0.5) {reify};
%%     \node[anchor=east] (n2exp) at (3.75, 0.5) {linearize};
%%     \node[anchor=east] (n3exp) at (7.5, 0.5) {const.-fold \& inline};
%%     \node[anchor=east] (n4exp) at (9.5, 0.5) {bounds};
%%     \node[anchor=east] (n5exp) at (11.4, 0.5) {$\eta$-exp.};

%%     \draw [arrow] (n1) to (root);
%%     \draw [arrow] (n2) to (n1);
%%     \draw [arrow] (n3) to (n2);
%%     \draw [arrow] (n4) to (n3);
%%     \draw [arrow] (n5) to (n4);
%%   \end{tikzpicture}
%%   \caption{Phases of certified compilation}
%%   \label{fig-certified}
%% \end{figure*}

To begin the certified-compilation phase of our pipeline, we need to \emph{reify} native Coq programs as terms of this AST type.
To illustrate the transformations we perform on ASTs, we walk through what the compiler does to an example program:

\noindent {\small \begin{align*}
  &\kw{let} \; (x_1, x_2, x_3) = x \; \kw{in} \\
  &\kw{let} \; (y_1, y_2) = ((\kw{let} \; z = x_2 \times 1 \times x_3 \; \kw{in} \; z + 0), x_2) \; \kw{in} \\
  &y_1 \times y_2 \times x_1
\end{align*}}%

\noindent The first phase is \emph{linearize}, which cancels out all intermediate uses of tuples and immediate $\kw{let}$-bound variables and moves all $\kw{let}$s to the top level.
%Optionally, this phase can also create temporaries for all subexpressions of multiple-operation terms, putting the output in A-normal form (ANF)~\cite{ANF}.

\noindent {\small \begin{align*}
  &\kw{let} \; (x_1, x_2, x_3) = x \; \kw{in} \\
  &\kw{let} \; z = x_2 \times 1 \times x_3 \; \kw{in} \\
  &\kw{let} \; y_1 = z + 0 \; \kw{in} \\
  &y_1 \times x_2 \times x_1
\end{align*}}%

\noindent Next is \emph{constant folding}, which applies simple arithmetic identities and inlines constants and variable aliases.

\noindent {\small \begin{align*}
  &\kw{let} \; (x_1, x_2, x_3) = x \; \kw{in} \\
  &\kw{let} \; z = x_2 \times x_3 \; \kw{in} \\
  &z \times x_2 \times x_1
\end{align*}}%

\noindent At this point we run the core phase, \emph{bounds inference}, the one least like the phases of standard C compilers.
The phase is parameterized over a list of available fixed-precision base types with their ranges; for our example, assume the hardware supports bit sizes 8, 16, 32, and 64.
Intervals for program inputs, like $x$ in our running example, are given as additional inputs to the algorithm.
Let us take them to be as follows: $x_1 \in [0, 2^8], x_2 \in [0, 2^{13}], x_3 \in [0, 2^{18}]$.
The output of the algorithm has annotated each variable definition and arithmetic operator with a finite type.

\noindent {\small \begin{align*}
  &\kw{let} \; (x_1 : \nat_{2^{16}}, \; x_2 : \nat_{2^{16}}, \; x_3 : \nat_{2^{32}}) = x \; \kw{in} \\
  &\kw{let} \; z : \nat_{2^{32}} = x_2 \times_{\nat_{2^{32}}} x_3 \; \kw{in} \\
  &z \times_{\nat_{2^{64}}} x_2 \times_{\nat_{2^{64}}} x_1
\end{align*}}%

\noindent Our biggest proof challenge here was in the interval rules for bitwise operators applied to negative numbers, a subject mostly missing from Coq's standard library.
%% Bitwise operators are essential for full-scale crypto primitives, and these operators pose an additional challenge in our full implementation.
%% In particular, Coq computes bitwise operations on negative integers using an infinite-width two's-complement representation.
%% Much of our time in the bounds-analysis proof was spent on appropriate bounds for negative as well as nonnegative inputs, since Coq's standard library contains few theorems about applying bitwise operators to negative numbers (but they can show up in crypto algorithms).

%The final compiler phase is \emph{$\eta$-expansion}.
%Because our $\kw{let}$ construct bakes in the decomposition of tuples, execution of the previous passes will block on any $\kw{let}$ binding for an opaque operator that returns a tuple (e.g., two-output multiplication), because Coq cannot ``peek underneath'' the operator to determine the rest of the syntax tree.
%The $\eta$-expansion phase destructures all binders, passing each reconstructed tuple to the rest of the syntax tree, allowing reduction of the previous phases to proceed.

\subsection{Important Design Choices}

%\paragraph{Encoding term structure in Coq.}

Most phases of the compiler use a term encoding called parametric higher-order abstract syntax (PHOAS)~\cite{PHOAS}.
Briefly, this encoding uses variables of the metalanguage (Coq's Gallina) to encode variables of the object language, to avoid most kinds of bookkeeping about variable environments; and for the most part we found that it lived up to that promise.
However, we needed to convert to a first-order representation (de Bruijn indices) and back for the bounds-inference phase, essentially because it calls for a forward analysis followed by a backward transformation: calculate intervals for variables, then rewrite the program bottom-up with precise base types for all variables.
We could not find a way with PHOAS to write a recursive function that returns both bounds information and a new term, taking better than quadratic time, while it was trivial to do with first-order terms.
We also found that the established style of term well-formedness judgment for PHOAS was not well-designed for large, automatically generated terms like ours: proving well-formedness would frequently take unreasonably long, as the proof terms are quadratic in the size of the syntax tree.
The fix was to switch well-formedness from an inductive definition into an executable recursive function that returns a linear number of simple constraints in propositional logic.
% where atomic formulas are equalities between variables, which admit more compact proofs.

\subsection{Extensibility with Nonobvious Algebraic Identities}

Classic abstract interpretation with intervals works surprisingly well for most of the cryptographic code we have generated.
However, a few spans of code call for some algebra to establish the tightest bounds.
We considered extending our abstract interpreter with general algebraic simplification, perhaps based on E-graphs as in SMT solvers~\cite{simplify}, but in the end we settled on an approach that is both simpler and more accommodating of unusual reasoning patterns.
The best-known among our motivating examples is Karatsuba's multiplication, which expresses a multi-digit multiplication in a way with fewer single-word multiplications (relatively slow) but more additions (relatively fast), where bounds checking should recognize algebraic equivalence to a simple multiplication.
Naive analysis would treat $(a + b)(c+d) - ac$ equivalently to $(a+b)(c+d) - xy$, where $x$ and $y$ are known to be within the same ranges as $a$ and $c$ but are not necessarily equal to them. The latter expression can underflow.
We add a new primitive operator $\mathsf{equivalently}$ such that, within the high-level functional program, we can write the above expression as $\mathsf{equivalently}\left((a+b)(c+d)-ac, ad+bc+bd\right)$.
The latter argument is used for bounds analysis.
The soundness theorem of bounds analysis requires, as a side condition, equality for all pairs of arguments to $\mathsf{equivalently}$.
These side conditions are solved automatically during compilation using Coq tactics, in this case \texttt{ring} algebra.
Compilation then replaces any $\mathsf{equivalently}\left(e_1, e_2\right)$ with just $e_1$.

\section{Experimental Results}\label{experiments}

The first purpose of this section is to confirm that implementing optimized algorithms in high-level code and then separately specializing to concrete parameters actually achieves the expected performance.
Given the previous sections, this conclusion should not be surprising: as code generation is extremely predictable, it is fair to think of the high-level implementations as simple templates for low-level code.
Thus, it would be possible write new high-level code to mimic every low-level implementation.
Such a strategy would, of course, be incredibly unsatisfying, although perhaps still worthwhile as a means to reduce effort required for correctness proofs.
Thus, we wish additionally to demonstrate that the two simple generic-arithmetic implementations discussed in Sections \ref{solinasreduce} and \ref{saturated} are sufficient to achieve good performance in all cases.
Furthermore, we do our best to call out differences that enable more specialized implementations to be faster, speculating on how we might modify our framework in the future based on those differences.

\subsection{X25519 Scalar Multiplication}

We measure the number of CPU cycles different implementations take to multiply a secret scalar and a public Curve25519 point (represented by the $x$ coordinate in Montgomery coordinates).
Despite the end-to-end task posed for this benchmark, we believe the differences between implementations we compare against lie in the field-arithmetic implementation --
all use the same scalar-multiplication algorithm and two almost-identical variants of the Montgomery-curve $x$-coordinate differential-addition formulas.

The benchmarks were run on an Intel Broadwell i7-5600U processor in a kernel module
%\footnote{Thanks to Jason Donenfeld for teaching us how and setting up benchmarks for popular Curve25519 implementations! %TODO:rephrase? Moved to acknowledgments
% TODO: should we incldue a link to his run of the benchmarks confirming a similar ranking on Skylake?
% https://www.mail-archive.com/curves@moderncrypto.org/msg00953.html
%}
with interrupts, power management, Hyper Threading, and Turbo Boost features disabled.
Each measurement represents the average of a batch of 15,000 consecutive trials, with time measured using the \texttt{RDTSC} instruction and converted to CPU cycles by multiplying by the ratio of CPU and timestamp-counter clock speeds.
C code was compiled using \texttt{gcc} 7.3 with \texttt{-O3 -march=native -mtune=native -fwrapv} (we also tried \texttt{clang} 5.0, but it produced $\sim$10\% slower code for the implementations here).

% gcc 7.3 always fastest
% NAME, TSC * 2.494/2.6, usec, raw TSC
\begin{center}%
  \newcommand{\makecpu}[2][100]{{\textcolor{black!#1}{\rule{\dimexpr6em * #2 / 546000\relax}{1ex}}}}%
  \newcommand{\makethiscpu}[1]{\makecpu[50]{#1}}%
  \newcommand{\makeothercpu}[1]{\makecpu[100]{#1}}%
  \begin{tabular}[]{lrl}
  Implementation & CPU cycles & \\
\hline
% \texttt{OLHFR}\cite{oliveira_sac2017}, asm       & \texttt{117877} & \makeothercpu{117877} \\ % 122887 % TODO this is newer usermode number
\texttt{OLHFR}, asm       & \texttt{121444} & \makeothercpu{126606} \\ % % 122887 % TODO this is kbench9000 number for older version
\texttt{amd64-64}, asm       & \texttt{151586} & \makeothercpu{151586} \\ % 158029
\textit{this work B}, 64-bit & \texttt{152195} & \makethiscpu{152195} \\ % 158664
\texttt{sandy2x}, asm        & \texttt{154313} & \makeothercpu{154313} \\ % 160872
\texttt{hacl-star}, 64-bit   & \texttt{154982} & \makeothercpu{154982} \\ % 161570
\texttt{donna64}, 64-bit C   & \texttt{168502} & \makeothercpu{168502} \\ % 175664
\textit{this work A}, 64-bit & \texttt{174637} & \makethiscpu{174637} \\ % 182060
\textit{this work, 32-bit  } & \texttt{310585} & \makethiscpu{310585} \\ % 323786
\texttt{donna32}, 32-bit C   & \texttt{529812} & \makeothercpu{529812} \\ % 552331
\hline
  \end{tabular}

%%   \pgfplotsset{
%%     ylabel right/.style={
%%       after end axis/.append code={
%%         \node [rotate=90, anchor=north] at (rel axis cs:1.1,0.5) {#1};
%%       }
%%     }
%%   }%
%%   \begin{tikzpicture}
%%     \begin{axis}[
%% %        xlabel=Implementation,
%%         ylabel=CPU cycles,
%%         ylabel right=\textmu{}s at 2.6GHz,
%%         yticklabel style={
%%           rotate=30,
%%           anchor=east,
%%         %  /pgf/number format/fixed,
%%         %  /pgf/number format/precision=6
%%         },
%%         scaled y ticks=false,
%%  %       ybar,
%%         ymin=0,
%%         symbolic x coords={%
%%           {\texttt{amd64-64}, asm      },
%%           {\textit{this work B}, 64-bit},
%%           {\texttt{sandy2x}, asm       },
%%           {\texttt{hacl-star}, 64-bit  },
%%           {\texttt{donna64}, 64-bit C  },
%%           {\textit{this work A}, 64-bit},
%%           {\textit{this work, 32-bit  }},
%%           {\texttt{donna32}, 32-bit C  }%
%%         },
%%  %       xtick={%
%%  %         {\texttt{amd64-64}, asm      },
%%  %         {\textit{this work B}, 64-bit},
%%  %         {\texttt{sandy2x}, asm       },
%%  %         {\texttt{hacl-star}, 64-bit  },
%%  %         {\texttt{donna64}, 64-bit C  },
%%  %         {\textit{this work A}, 64-bit},
%%  %         {\textit{this work, 32-bit  }},
%%  %         {\texttt{donna32}, 32-bit C  }%
%%         %       },
%%         ytick pos=left,
%%         yticklabel={$\pgfmathprintnumber\tick$},
%%         extra y ticks = {0,130000,260000,390000,540000},
%%         extra y tick labels = {,50,100,150,200},
%%         extra y tick style={
%%           ytick pos=right,
%%           yticklabel pos=right,
%%           yticklabel style={rotate=-30,anchor=west}
%%         },
%% %        extra y tick label={$\pgfmathprintnumber\tick$},
%%         x tick label style={rotate=80,anchor=east},
%%         xtick=data,
%%         only marks,
%%       ]
%%       \addplot+[ycomb] coordinates {
%%         ({\texttt{amd64-64}, asm      }%
%%         ,151586)
%%         ({\textit{this work B}, 64-bit}%
%%         ,152195)
%%         ({\texttt{sandy2x}, asm       }%
%%         ,154313)
%%         ({\texttt{hacl-star}, 64-bit  }%
%%         ,154982)
%%         ({\texttt{donna64}, 64-bit C  }%
%%         ,168502)
%%         ({\textit{this work A}, 64-bit}%
%%         ,174637)
%%         ({\textit{this work, 32-bit  }}%
%%         ,310585)
%%         ({\texttt{donna32}, 32-bit C  }%
%%         ,529812)
%%       };
%%       \addplot+[ycomb] coordinates {
%%         ({\textit{this work B}, 64-bit}%
%%         ,152195)
%%         ({\textit{this work A}, 64-bit}%
%%         ,174637)
%%         ({\textit{this work, 32-bit  }}%
%%         ,310585)
%%       };
%%     \end{axis}
%%   \end{tikzpicture}
\end{center}

In order, we compare against \texttt{OLHFR}, the non-precomputation-based implementation from \cite{oliveira_sac2017}; \texttt{amd64-64} and \texttt{sandy2x}, the fastest assembly implementations from SUPERCOP~\cite{SUPERCOP} that use scalar and vector instructions respectively; the verified X25519 implementation from the HACL$^*$ project~\cite{hacl}; and the best-known high-performance C implementation \texttt{curve25519-donna}, in both 64-bit and 32-bit variants.
We measured cycle counts for all comparison implementations from a checkout of kbench9000~\cite{kbench} at commit \texttt{59063a52}.
The field arithmetic in both \texttt{amd64-64} and \texttt{hacl-star} has been verified using SMT solvers~\cite{verif25519,ECCstar}.
Of course, care should be taken when extrapolating these results to other compilers and CPUs: each assembly implementation is likely to be fastest on the microarchitecture it was optimized for, and performance of C code additionally varies with the quality of compiler optimizations for the microarchitecture in question.

We report on our code generated using the standard representations for both 32-bit and 64-bit, though we are primarily interested in the latter, since we benchmark on a 64-bit processor.
We actually report on two of our 64-bit variants, both of which have correctness proofs of the same strength.
Variant \emph{A} is the output of the toolchain described in Section \ref{basesystems}.
Variant \emph{B} contains an additional implementation grouping multiplications by the modular-reduction coefficient $c=19$ with one of the inputs of the multiplication instead of the output ($19\times (a\times b) \longrightarrow (19\times a)\times b$).
This algebra pays off because $19\times a$ can be computed using a 64-bit multiplication, whereas $19\times (ab)$ would require a significantly more expensive 128-bit multiplication.
We currently implement this trick for just $2^{255}-19$ as an ad-hoc code replacement after partial evaluation but before word-size inference, using the Coq \texttt{ring} algebra tactic to prove that the code change does not change the output.
We are planning on incorporating a more general version of this optimization into the main pipeline, after a planned refactoring to add automatic translation to continuation-passing style (see Section \ref{conclusion}).

The results of a similar benchmark on an earlier version of our pipeline were good enough to convince the maintainers of the BoringSSL library to adopt our methodology, resulting in this Curve25519 code being shipped in Chrome 64 and used by default for TLS connection establishment in other Google products and services.
Previously, BoringSSL included the \texttt{amd64-64} assembly code and a 32-bit C implementation as a fallback, which was the first to be replaced with our generated code.
Then, the idea was raised of taking advantage of lookup tables to optimize certain point ECC multiplications.
While the BoringSSL developers had not previously found it worthwhile to modify 64-bit assembly code and review the changes, they made use of our code-generation pipeline (without even consulting us, the tool authors) and installed a new 64-bit C version.
The new code (our generated code linked with manually written code using lookup tables) was more than twice as fast as the old version and was easily chosen for adoption, enabling the retirement of \texttt{amd64-64} from BoringSSL.

% https://boringssl-review.googlesource.com/c/boringssl/+/24805
% Use 51-bit limbs from fiat-crypto in 64-bit.
%
% Our 64-bit performance was much lower than it could have been, since we
% weren't using the 64-bit multipliers. Fortunately, fiat-crypto is
% awesome, so this is just a matter of synthesizing new code and
% integration work.
%
% Functions without the signature fiat-crypto curly braces were written by
% hand and warrant more review. (It's just redistributing some bits.)
%
%
% https://boringssl-review.googlesource.com/c/boringssl/+/25524
% Now that we have 64-bit C code, courtesy of fiat-crypto, the tradeoff
% for carrying the assembly changes:
%
% Assembly:
% Did 16000 Curve25519 base-point multiplication operations in 1059932us (15095.3 ops/sec)
% Did 16000 Curve25519 arbitrary point multiplication operations in 1060023us (15094.0 ops/sec)
%
% fiat64:
% Did 39000 Curve25519 base-point multiplication operations in 1004712us (38817.1 ops/sec)
% Did 14000 Curve25519 arbitrary point multiplication operations in 1006827us (13905.1 ops/sec)
%
% The assembly is still about 9% faster than fiat64, but fiat64 gets to
% use the Ed25519 tables for the base point multiplication, so overall it
% is actually faster to disable the assembly:
%
% >>> 1/(1/15094.0 + 1/15095.3)
% 7547.324986004976
% >>> 1/(1/38817.1 + 1/13905.1)
% 10237.73016319501
%
% (At the cost of touching a 30kB table.)
%
% The assembly implementation is no longer pulling its weight. Remove it
% and use the fiat code in all build configurations.

\subsection{P-256 Mixed Addition}

Next, we benchmark our Montgomery modular arithmetic as used for in-place point addition on the P-256 elliptic curve with one precomputed input (Jacobian += affine).
A scalar-multiplication algorithm using precomputed tables would use some number of these additions depending on the table size.
The assembly-language implementation \texttt{nistz256} was reported on by Gueron and Krasnov~\cite{nistz256} and included in OpenSSL; we also measure its newer counterpart that makes use of the ADX instruction-set extension.
% Our choice to use Montgomery reduction was inspired by the impressive performance of that implementation.
The 64-bit C code we benchmark is also from OpenSSL and uses unsaturated-style modular reduction, carefully adding a couple of multiples of the prime each time before performing a reduction step with a negative coefficient to avoid underflow.
These P-256 implementations here are unverified.
The measurement methodology is the same as for our X25519 benchmarks, except that we did not manage to get \texttt{nistz256} running in a kernel module and report userspace measurements instead.
% We also passed \texttt{-ipo} to \texttt{icc}.

\begin{center}\begin{tabular}[]{lrrrr}
% tilde means measured in userspace, everything else kbench9000
% 2.494*TSC
  Implementation        & fastest       & \texttt{clang}& \texttt{gcc}& \texttt{icc} \\
\hline
  \texttt{nistz256} +ADX &  \verb|~550|  &  &  &  \\
  \texttt{nistz256} AMD64&  \verb|~650|  &  &  &  \\
  \textit{this work A}   &  \verb|1143|  & \texttt{1811}& \texttt{1828}& \texttt{1143}\\
  OpenSSL, 64-bit C      &  \verb|1151|  & \texttt{1151}& \texttt{2079}& \texttt{1404}\\
  \textit{this work B}   &  \verb|1343|  & \texttt{1343}& \texttt{2784}& \texttt{1521}\\
\hline
\end{tabular}\end{center}

 \begin{figure*}
 \begin{tikzpicture}
 	\begin{axis}[
 		height=3.4cm,
 		title style={font=\small},
 		title=64-Bit Field Arithmetic Benchmarks,
 		width=\textwidth,
 		legend pos= north west,
% 		xtick distance=64,
 		legend style={font=\tiny},
 		label style={font=\footnotesize},
 		xlabel style={at={(0.5,0.1)}, anchor=north},
 		legend columns=2,
 		tick label style={font=\footnotesize},
 		grid=major,
 		ymin=0,
         xmin=90,
         xmax=560,
 		xlabel=log2(prime),
 		ylabel=Time (seconds)]
		\addplot[color=red,mark=*,mark size=2pt] coordinates {
			(127.0, 0.26)
			(129.0, 0.38)
			(130.0, 0.39)
			(137.0, 0.37)
			(140.0, 0.38)
			(141.0, 0.39)
			(150.0, 0.38)
			(150.0, 0.39)
			(152.0, 0.38)
			(158.0, 0.37)
			(165.0, 0.38)
			(166.0, 0.39)
			(171.0, 0.38)
			(174.0, 0.38)
			(174.0, 0.38)
			(189.0, 0.39)
			(190.0, 0.38)
			(191.0, 0.36)
			(192.0, 0.31)
			(194.0, 0.5)
			(196.0, 0.5)
			(198.0, 0.51)
			(204.37503943134692, 0.51)
			(206.0, 0.5)
			(212.0, 0.49)
			(213.0, 0.49)
			(216.0, 0.51)
			(221.0, 0.51)
			(222.0, 0.53)
			(224.0, 0.5)
			(226.0, 0.5)
			(230.0, 0.54)
			(235.0, 0.5)
			(243.0, 0.5)
			(251.0, 0.5)
			(253.98877343250717, 0.5)
			(255.0, 0.48)
			(255.0, 0.52)
			(256.0, 0.38)
			(255.99999999966408, 0.38)
			(256.0, 0.38)
			(256.0, 0.37)
			(255.9980614856364, 0.39)
			(266.0, 0.66)
			(285.0, 0.73)
			(291.0, 0.68)
			(321.0, 0.8)
			(322.0, 0.83)
			(336.0, 0.8)
			(336.0, 0.86)
			(338.0, 0.8)
			(369.0, 0.79)
			(379.0, 0.79)
			(382.0, 0.92)
			(383.0, 0.75)
			(383.0, 0.75)
			(383.0, 0.76)
			(384.0, 0.64)
			(384.0, 0.67)
			(383.9998899269044, 0.63)
			(383.467605550083, 0.62)
			(389.0, 0.97)
			(401.0, 0.97)
			(413.0, 0.99)
			(414.0, 0.98)
			(416.0, 0.98)
			(444.0, 0.96)
			(448.0, 0.79)
			(450.0, 1.22)
			(452.0, 1.24)
			(468.0, 1.23)
			(480.0, 1.18)
			(488.0, 1.2)
			(489.0, 1.17)
			(495.0, 1.17)
			(509.97423531735535, 1.2)
			(511.0, 1.13)
			(511.0, 1.12)
			(511.9891505409899, 0.99)
			(512.0, 0.95)
			(521.0, 1.4)
		};
		\addlegendentry{GMP mpn\_sec API}

		\addplot[color=red,mark=o,mark size=2pt] coordinates {
			(127.0, 0.15)
			(129.0, 0.27)
			(130.0, 0.28)
			(137.0, 0.27)
			(140.0, 0.27)
			(141.0, 0.27)
			(150.0, 0.3)
			(150.0, 0.29)
			(152.0, 0.27)
			(158.0, 0.27)
			(165.0, 0.27)
			(166.0, 0.27)
			(171.0, 0.27)
			(174.0, 0.28)
			(174.0, 0.27)
			(189.0, 0.27)
			(190.0, 0.27)
			(191.0, 0.26)
			(192.0, 0.24)
			(194.0, 0.34)
			(196.0, 0.34)
			(198.0, 0.34)
			(204.37503943134692, 0.34)
			(206.0, 0.34)
			(212.0, 0.35)
			(213.0, 0.37)
			(216.0, 0.35)
			(221.0, 0.36)
			(222.0, 0.35)
			(224.0, 0.35)
			(226.0, 0.35)
			(230.0, 0.35)
			(235.0, 0.34)
			(243.0, 0.34)
			(251.0, 0.35)
			(253.98877343250717, 0.35)
			(255.0, 0.35)
			(255.0, 0.34)
			(256.0, 0.34)
			(255.99999999966408, 0.33)
			(256.0, 0.34)
			(256.0, 0.34)
			(255.9980614856364, 0.34)
			(266.0, 0.45)
			(285.0, 0.43)
			(291.0, 0.42)
			(321.0, 0.54)
			(322.0, 0.53)
			(336.0, 0.53)
			(336.0, 0.53)
			(338.0, 0.54)
			(369.0, 0.52)
			(379.0, 0.55)
			(382.0, 0.57)
			(383.0, 0.5)
			(383.0, 0.51)
			(383.0, 0.51)
			(384.0, 0.47)
			(384.0, 0.48)
			(383.9998899269044, 0.46)
			(383.467605550083, 0.46)
			(389.0, 0.6)
			(401.0, 0.61)
			(413.0, 0.62)
			(414.0, 0.6)
			(416.0, 0.6)
			(444.0, 0.6)
			(448.0, 0.52)
			(450.0, 0.74)
			(452.0, 0.71)
			(468.0, 0.71)
			(480.0, 0.71)
			(488.0, 0.7)
			(489.0, 0.69)
			(495.0, 0.69)
			(509.97423531735535, 0.7)
			(511.0, 0.66)
			(511.0, 0.66)
			(511.9891505409899, 0.62)
			(512.0, 0.62)
			(521.0, 0.81)
		};
		\addlegendentry{GMP mpn API}

		\addplot[color=blue,mark=triangle*,mark size=2pt] coordinates {
			(127.0, 0.03)
			(129.0, 0.03)
			(130.0, 0.03)
			(137.0, 0.03)
			(140.0, 0.03)
			(141.0, 0.03)
			(150.0, 0.03)
			(150.0, 0.03)
			(152.0, 0.03)
			(158.0, 0.03)
			(165.0, 0.03)
			(166.0, 0.03)
			(171.0, 0.03)
			(174.0, 0.03)
			(174.0, 0.03)
			(189.0, 0.04)
			(190.0, 0.04)
			(191.0, 0.04)
			(192.0, 0.05)
			(194.0, 0.04)
			(196.0, 0.04)
			(198.0, 0.04)
			(204.37503943134692, 0.14)
			(206.0, 0.04)
			(212.0, 0.05)
			(213.0, 0.04)
			(216.0, 0.04)
			(221.0, 0.05)
			(222.0, 0.05)
			(224.0, 0.13)
			(226.0, 0.04)
			(230.0, 0.05)
			(235.0, 0.06)
			(243.0, 0.06)
			(251.0, 0.06)
			(253.98877343250717, 0.12)
			(255.0, 0.06)
			(255.0, 0.06)
			(256.0, 0.06)
			(255.99999999966408, 0.11)
			(256.0, 0.1)
			(256.0, 0.13)
			(255.9980614856364, 0.11)
			(266.0, 0.06)
			(285.0, 0.06)
			(291.0, 0.07)
			(321.0, 0.1)
			(322.0, 0.07)
			(336.0, 0.1)
			(336.0, 0.09)
			(338.0, 0.1)
			(369.0, 0.13)
			(379.0, 0.12)
			(382.0, 0.13)
			(383.0, 0.13)
			(383.0, 0.13)
			(383.0, 0.13)
			(384.0, 0.25)
			(384.0, 0.13)
			(383.9998899269044, 0.23)
			(383.467605550083, 0.23)
			(389.0, 0.13)
			(401.0, 0.14)
			(413.0, 0.16)
			(414.0, 0.15)
			(416.0, 0.14)
			(444.0, 0.17)
			(448.0, 0.12)
			(450.0, 0.13)
			(452.0, 0.16)
			(468.0, 0.16)
			(480.0, 0.13)
			(488.0, 0.19)
			(489.0, 0.2)
			(495.0, 0.19)
			(511.0, 0.25)
			(511.0, 0.25)
			(512.0, 0.24)
			(521.0, 0.18)
		};
		\addlegendentry{this work}

	\end{axis}
\end{tikzpicture}
 \begin{tikzpicture}
 \begin{axis}[
 ymin=0,
 ymax=15.5,
 xmin=90,
 xmax=560,
 ticks=none,
 height=3.4cm,
 width=\textwidth,
 ytick={0,1.67,3.33,5,...,15},
 %label style={font=\footnotesize},
 %xlabel style={at={(0.5,0.1)}, anchor=north},
 %tick label style={font=\footnotesize},
 grid=major,
 ]
 \end{axis}
 	\begin{axis}[
        height=3.4cm,
 		title style={font=\small},
 		title=32-Bit Field Arithmetic Benchmarks,
 		width=\textwidth,
 		legend pos= north west,
% 		xtick distance=64,
 		legend style={font=\tiny},
 		label style={font=\footnotesize},
 		xlabel style={at={(0.5,0.1)}, anchor=north},
 		legend columns=2,
 		tick label style={font=\footnotesize},
 		grid=major,
 		ymin=0,
        ymax=15.5,
        xmin=90,
        xmax=560,
 		xlabel=log2(prime),
 		ylabel=Time (seconds)]
		\addplot[color=red,mark=*,mark size=2pt] coordinates {
			(127.0, 2.86)
			(129.0, 3.38)
			(130.0, 3.56)
			(137.0, 3.41)
			(140.0, 3.43)
			(141.0, 3.43)
			(150.0, 3.56)
			(150.0, 3.38)
			(152.0, 3.4)
			(158.0, 3.4)
			(165.0, 4.02)
			(166.0, 4.02)
			(171.0, 4.04)
			(174.0, 4.03)
			(174.0, 4.02)
			(189.0, 4.05)
			(190.0, 4.1)
			(191.0, 4.03)
			(192.0, 3.56)
			(194.0, 4.66)
			(196.0, 4.64)
			(198.0, 4.72)
			(204.37503943134692, 4.66)
			(206.0, 4.62)
			(212.0, 4.68)
			(213.0, 4.68)
			(216.0, 4.67)
			(221.0, 4.68)
			(222.0, 4.72)
			(224.0, 4.13)
			(226.0, 5.25)
			(230.0, 5.29)
			(235.0, 5.31)
			(243.0, 5.29)
			(251.0, 5.3)
			(253.98877343250717, 5.26)
			(255.0, 5.25)
			(255.0, 5.27)
			(256.0, 4.71)
			(255.99999999966408, 4.7)
			(256.0, 4.72)
			(256.0, 4.77)
			(255.9980614856364, 4.78)
			(266.0, 6.1)
			(285.0, 6.13)
			(291.0, 6.94)
			(321.0, 7.6)
			(322.0, 7.66)
			(336.0, 7.64)
			(336.0, 7.58)
			(338.0, 7.66)
			(369.0, 8.41)
			(379.0, 8.44)
			(382.0, 8.41)
			(383.0, 8.44)
			(383.0, 8.47)
			(383.0, 8.45)
			(384.0, 7.62)
			(384.0, 7.62)
			(383.9998899269044, 7.64)
			(383.467605550083, 7.66)
			(389.0, 9.41)
			(401.0, 9.35)
			(413.0, 9.48)
			(414.0, 9.4)
			(416.0, 8.54)
			(444.0, 10.31)
			(448.0, 9.57)
			(450.0, 11.37)
			(452.0, 11.33)
			(468.0, 11.37)
			(480.0, 10.47)
			(488.0, 12.23)
			(489.0, 12.26)
			(495.0, 12.2)
			(509.97423531735535, 12.17)
			(511.0, 12.21)
			(511.0, 12.23)
			(511.9891505409899, 11.26)
			(512.0, 11.23)
			(521.0, 13.3)
		};
		\addlegendentry{GMP mpn\_sec API}

		\addplot[color=red,mark=o,mark size=2pt] coordinates {
			(127.0, 3.23)
			(129.0, 3.77)
			(130.0, 3.79)
			(137.0, 3.78)
			(140.0, 3.77)
			(141.0, 3.81)
			(150.0, 3.79)
			(150.0, 3.8)
			(152.0, 3.82)
			(158.0, 3.77)
			(165.0, 4.45)
			(166.0, 4.43)
			(171.0, 4.51)
			(174.0, 4.39)
			(174.0, 4.4)
			(189.0, 4.4)
			(190.0, 4.42)
			(191.0, 4.4)
			(192.0, 4.23)
			(194.0, 4.94)
			(196.0, 4.94)
			(198.0, 4.97)
			(204.37503943134692, 5.03)
			(206.0, 4.91)
			(212.0, 4.91)
			(213.0, 4.94)
			(216.0, 4.92)
			(221.0, 4.92)
			(222.0, 4.87)
			(224.0, 4.85)
			(226.0, 5.65)
			(230.0, 5.71)
			(235.0, 5.69)
			(243.0, 5.62)
			(251.0, 5.65)
			(253.98877343250717, 5.7)
			(255.0, 5.7)
			(255.0, 5.71)
			(256.0, 5.49)
			(255.99999999966408, 5.46)
			(256.0, 5.45)
			(256.0, 5.48)
			(255.9980614856364, 5.46)
			(266.0, 6.32)
			(285.0, 6.34)
			(291.0, 6.98)
			(321.0, 7.66)
			(322.0, 7.74)
			(336.0, 7.74)
			(336.0, 7.69)
			(338.0, 7.67)
			(369.0, 9.03)
			(379.0, 9.25)
			(382.0, 9.04)
			(383.0, 9.11)
			(383.0, 9.13)
			(383.0, 9.11)
			(384.0, 8.8)
			(384.0, 8.82)
			(383.9998899269044, 8.94)
			(383.467605550083, 8.84)
			(389.0, 9.93)
			(401.0, 9.92)
			(413.0, 9.93)
			(414.0, 9.86)
			(416.0, 9.67)
			(444.0, 10.89)
			(448.0, 10.51)
			(450.0, 11.63)
			(452.0, 11.63)
			(468.0, 11.63)
			(480.0, 11.33)
			(488.0, 12.92)
			(489.0, 12.81)
			(495.0, 13.1)
			(509.97423531735535, 12.9)
			(511.0, 13.07)
			(511.0, 12.9)
			(511.9891505409899, 12.58)
			(512.0, 12.55)
			(521.0, 13.91)
		};
		\addlegendentry{GMP mpn API}

		\addplot[color=blue,mark=triangle*,mark size=2pt] coordinates {
			(127.0, 0.3)
			(129.0, 0.35)
			(130.0, 0.44)
			(137.0, 0.48)
			(140.0, 0.51)
			(141.0, 0.51)
			(150.0, 0.42)
			(150.0, 0.49)
			(152.0, 0.5)
			(158.0, 0.52)
			(165.0, 0.59)
			(166.0, 0.61)
			(171.0, 0.57)
			(174.0, 0.58)
			(174.0, 0.61)
			(189.0, 0.7)
			(190.0, 0.71)
			(191.0, 0.66)
			(192.0, 2.41)
			(194.0, 0.75)
			(196.0, 0.77)
			(198.0, 0.76)
			(206.0, 0.76)
			(212.0, 0.86)
			(213.0, 0.7)
			(216.0, 0.7)
			(221.0, 0.8)
			(222.0, 0.87)
			(226.0, 0.87)
			(230.0, 0.83)
			(235.0, 0.9)
			(243.0, 0.86)
			(251.0, 1.12)
			(253.98877343250717, 3.97)
			(255.0, 1.01)
			(255.0, 1.43)
			(256.0, 1.2)
			(256.0, 1.65)
			(266.0, 1.01)
			(285.0, 1.13)
			(291.0, 1.33)
			(321.0, 1.72)
			(322.0, 1.37)
			(336.0, 1.67)
			(336.0, 1.59)
			(338.0, 1.7)
			(369.0, 2.44)
			(379.0, 2.47)
			(382.0, 2.66)
			(383.0, 2.63)
			(383.0, 2.6)
			(383.0, 3.58)
			(384.0, 3.95)
			(389.0, 2.89)
			(401.0, 2.85)
			(413.0, 3.53)
			(414.0, 3.72)
			(416.0, 2.48)
			(444.0, 3.7)
			(448.0, 3.18)
			(452.0, 3.23)
			(468.0, 3.2)
			(480.0, 3.58)
			(488.0, 7.99)
			(489.0, 7.7)
			(495.0, 6.07)
			(511.0, 9.73)
			(521.0, 3.9)
		};
		\addlegendentry{this work}

	\end{axis}
\end{tikzpicture}
\caption{\label{graphs}Performance comparison of our generated C code vs. handwritten using \texttt{libgmp}}
\end{figure*}

Saturated arithmetic is a known weak point of current compilers, resulting in implementors either opting for alternative arithmetic strategies or switching to assembly language.
Our programs are not immune to these issues: when we first ran our P-256 code, it produced incorrect output because \texttt{gcc} 7.1.1 had generated incorrect code%
\footnote{\url{https://gcc.gnu.org/bugzilla/show_bug.cgi?id=81300}, \url{https://gcc.gnu.org/bugzilla/show_bug.cgi?id=81294}}%
; \texttt{clang} 4.0 exited with a mere segmentation fault.%
\footnote{\url{https://bugs.llvm.org/show_bug.cgi?id=24943}}
Even in later compiler versions where these issues have stopped appearing, the code generated for saturated arithmetic varies a lot between compilers and is obviously suboptimal:
for example, there are ample redundant moves of carry flags, perhaps because the compilers do not consider flag registers during register allocation.
The same pattern is also present for X25519, although less pronounced: the two fastest assembly implementations listed earlier use a 4-limb saturated representation, but the speedup over 5-limb unsaturated assembly is smaller than the slowdown incurred in C code due to heavy use of carry flags.
Furthermore, expressing the same computation using intrinsics such as \verb|_mulx_u64| (variant A in the table) or using \texttt{uint128} and a bit shift (variant B) can produce a large performance difference, in different directions on different compilers.
% As arithmetic operations in our C source line up very closely with those in the \texttt{nistz256} assembly code, and as the difference between \texttt{icc} and \texttt{gcc} is bigger than the difference between our code and the fastest assembly code, we attribute most of the performance difference to low-level compilation.

The BoringSSL team had a positive enough experience with adopting our framework for Curve25519 that they decided to use our generated code to replace their P-256 implementation as well.
First, they replaced their handwritten 64-bit C implementation.
Second, while they had never bothered to write a specialized 32-bit P-256 implementation before, they also generated one with our framework.
\texttt{nistz256} remains as an option for use in server contexts where performance is critical and where patches can be applied quickly when new bugs are found.
The latter is not a purely theoretical concern -- Appendix \ref{bugsapp} contains an incomplete list of issues discovered in previous \texttt{nistz256} versions.

The two curves thus generated with our framework for BoringSSL together account for over 99\% of ECDH connections.
Chrome version 65 is the first release to use our P-256 code.

% https://boringssl-review.googlesource.com/c/boringssl/+/23244

% maybe write a section answering openssl mailing lists skepticism about
% low-level compilers, "what to do about the compiler?", basically saying it is
% critical infrastructure no matter what?

\subsection{Benchmarking All the Primes}\label{instantiate}

Even though the primes supported by the modular-reduction algorithms we implemented vastly outnumber those actually used for cryptography, we could not resist the temptation to do a breadth-oriented benchmark for plausible cryptographic parameters.
The only other implementation of constant-time modular arithmetic for arbitrary modulus sizes we found is the \verb|mpn_sec| section of the GNU Multiple Precision Arithmetic Library\footnote{\url{https://gmplib.org/}}.
We also include the non-constant-time \texttt{mpn} interface for comparison.
The list of primes to test was generated by scraping the archives of the ECC discussion list \texttt{curves@moderncrypto.org}, finding all supported textual representations of prime numbers.

One measurement in Fig. \ref{graphs} corresponds to 1000 sequential computations of a 256-bit Montgomery ladder, aiming to simulate cryptographic use accurately while still constraining the source of performance variation to field arithmetic.
A single source file was used for all \verb|mpn_sec|-based implementations; the modulus was specified as a preprocessor macro to allow compiler optimizations to try their best to specialize GMP code.
Our arithmetic functions were instantiated using standard heuristics for picking the compile-time parameters based on the shorthand textual representation of the prime, picking a champion configuration out of all cases where correctness proof succeeds, with two unsaturated and one Montgomery-based candidates.
The 64-bit trials were run on an x86 Intel Haswell processor and 32-bit trials were run on an ARMv7-A Qualcomm Krait (LG Nexus 4).

We see a significant performance advantage for our code, even compared to the GMP version that ``cheats'' by leaking secrets through timing.
Speedups range between 1.25X and 10X.
Appendix \ref{data} includes the full details, with Tables \ref{table64} and \ref{table32} recording all experimental data points, including for an additional comparison implementation in C++.
In our current experiments, compilation of saturated arithmetic in Coq times out for a handful of larger primes.
There is no inherent reason for this slowness, but our work-in-progress reimplementation of this pipeline (Section \ref{conclusion}) is designed to be executable using the Haskell extraction feature of Coq,
aiming to sidestep such performance unpredictability conclusively.

\section{Related Work}\label{related}

% I think it might make sense to organize/split this section based on whether some project or plan goes up or down in terms of abstraction level.
% First, verification projects:

% I think verif25519 is the predominant prior work, and we should give adaquate exposure unless political reasons demand otherwise.
% From the perspective of our high-level implementations, there is very little difference between the abstraction levels of verif25519 (qhasm) and HACl-star (common subset of F* with mutable state and C).
% HACL-star specifications are much nicer, but only about halfway to ours
% HACL-star proofs are slightly nicer, but only questionably so.
% HACL-star covers more code, but the additional higher-level code it covers acts as its own specification during verification.

% The second most interesting piece of prior work is gfverif, because it is fully automatic, but it is not published and is alpha-alpha quality according to the authors.
% of course, it still requires writing all the code by hand
% gröbner basis solvers are just really good for addition and multiplication, which make up the vast majority of unsaturated implementations.
% it is unclear how it could be extended to do anything beyond unsatruated modular arithmetic

% Scala LMS and other templating systems should be mentioned here

% Our main comparison to them is that our code is actually a natural expression of the algorithm
% because of this, verification is much less painful
% because of this, genericity (to the extent supported by the algorithm) is freely available

% Separately, we might think of extending our library to deal with assembly as opposde to C.
% there are some efforts that already do pretty well on this

% Vale and Jasmin deserve the current mentions

% verifiedBignum, compcert, cakeml

% a simple success criteria for a compiler: if you take handwritten assembly that is line-by-line expressible as compiler input and translate it to the input language, the compiler should not generate code slower than that assembly. This would help our P-256 code greatly.

% supporting loops is not a todo-item, we already have proofs about loops and we don't want the current C extraction to deal with them anyway

\subsection{Verified Elliptic-Curve Cryptography}

Several past projects have produced verified ECC implementations.
We first summarize them and then give a unified comparison against our new work.

Chen et al.~\cite{verif25519} verified an assembly implementation of Curve25519, using a mix of automatic SAT solving and manual Coq proof for remaining goals.
A later project by Bernstein and Schwabe~\cite{gfverif}, described as an ``alpha test,'' explores an alternative workflow using the Sage computer-algebra system.
This line of work benefits from the chance for direct analysis of common C and assembly programs for unsaturated arithmetic, so that obviously established standards of performance can be maintained.

Vale~\cite{vale} supports compile-time metaprogramming of assembly code, with a cleaner syntax to accomplish the same tasks done via Perl scripts in OpenSSL.
There is a superficial similarity to the flexible code generation used in our own work.
However, Vale and OpenSSL use comparatively shallow metaprogramming, essentially just doing macro substitution, simple compile-time offset arithmetic, and loop unrolling.
Vale has not been used to write code parameterized on a prime modulus (and OpenSSL includes no such code).
%Vale connects to Dafny in a proof-assistant-inspired style that keeps Vale itself out of the trusted code base.
A verified static analysis checks that assembly code does not leak secrets, including through timing channels.

HACL$^*$~\cite{hacl} is a cryptographic library implemented and verified in the F$^*$ programming language, providing all the functionality needed to run TLS 1.3 with the latest primitives.
Primitives are implemented in the Low$^*$ imperative subset of F$^*$~\cite{kremlin}, which supports automatic semantics-preserving translation to C.
As a result, while taking advantage of F$^*$'s high-level features for specification, HACL$^*$ performs comparably to leading C libraries.
Additionally, abstract types are used for secret data to avoid side-channel leaks.

Jasmin~\cite{jasmin} is a low-level language that wraps assembly-style straightline code with C-style control flow.
It has a Coq-verified compiler to 64-bit x86 assembly (with other targets planned), along with support for verification of memory safety and absence of information leaks, via reductions to Dafny.
A Dafny reduction for functional-correctness proof exists but has not yet been used in a significant case study.

We compare against these past projects in a few main claims.

\emph{Advantage of our work:} \textbf{first verified high-performance implementation of P-256}.
Though this curve is the most widely used in TLS today, no prior projects had demonstrated performance-competitive versions with correctness proofs.
A predecessor system to HACL$^*$~\cite{ECCstar} verified more curves, including P-256, but in high-level F$^*$ code, incurring performance overhead above 100X.

\emph{Advantage of our work:} \textbf{code and proof reuse across algorithms and parameters}.
A more fundamental appeal of our approach is that it finally enables the high-performance-cryptography domain to benefit from pleasant software-engineering practices of abstraction and code reuse.
The code becomes easier to understand and maintain (and therefore to prove), and we also gain the ability to compile fast versions of new algorithm variants automatically, with correctness proofs.
Considering the cited past work, it is perhaps worth emphasizing how big of a shift it is to do implementation and proving at a parameterized level: as far as we are aware, no past project on high-performance verified ECC has ever \emph{verified implementations of multiple elliptic curves} or \emph{verified multiple implementations of one curve, targeted at different hardware characteristics}.
With our framework, all of those variants can be generated automatically, at low developer cost; and indeed the BoringSSL team took advantage of that flexibility.

\emph{Advantage of our work:} \textbf{small trusted code base}.
Every past project we mentioned includes either an SMT solver or a computer-algebra system in the trusted code base.
Furthermore, each past project also trusts some program-verification tool that processes program syntax and outputs symbolic proof obligations.
We trust only the standard Coq theorem prover, which has a proof-checking kernel significantly smaller than the dependencies just mentioned, and the (rather short, whiteboard-level) statements of the formal claims that we make (plus, for now, the C compiler; see below).

\emph{Disadvantage of our work:} \textbf{stopping at C rather than assembly}.
Several past projects apply to assembly code rather than C code.
As a result, they can achieve higher performance and remove the C compiler from the trusted base.
Our generated code is already rather close to assembly, differing only in the need for some register allocation and instruction scheduling.
In fact, it seems worth pointing out that every phase of our code-generation pipeline is necessary to get code \emph{low-level enough} to be accepted as input by Jasmin or Vale.
We have to admit to nontrivial frustration with popular C compilers in the crypto domain, keeping us from using them to bridge this gap with low performance cost.
We wish compiler authors would follow the maxim that, if a programmer takes handwritten assembly and translates it line-by-line to the compiler's input language, the compiler should not generate code slower than the original assembly.
Unfortunately, no popular compiler seems to obey that maxim today, so we are looking into writing our own (with correctness proof) or integrating with one of the projects mentioned above.

\emph{Disadvantage of our work:} \textbf{constant-time guarantee mechanized only for straightline code}.
Our development also includes higher-level cryptographic code that calls the primitives described here, with little innovation beyond comprehensive functional-correctness proofs.
That code is written following ``constant-time'' coding rules and thus should be free of timing side channels barring human error.
However, we do not yet do certified compilation or static analysis for timing leaks on the higher-level code.
To this end, it could be advantageous to take inspiration from, or adopt, one of the tools mentioned above.

\subsection{Other Related Work}

Performance-oriented synthesis of domain-specific code (without proofs of correctness) has previously been done using explicit templates (e.g. Template Haskell~\cite{TemplateHaskell}) and more sophisticated multistage programming (e.g. Lightweight Modular Staging (LMS)~\cite{LMS}).
More specialized frameworks along these lines include FFTW~\cite{FFTW} and Spiral~\cite{SPIRAL}.
Out of these, our strategy is most similar to the one from LMS, differing mainly in the choice of using existing (proof-generating) Coq facilities for partial evaluation and rewriting rather than implementing them ourselves.
While verified compilers (e.g., CakeML~\cite{CakeML}, CompCert~\cite{Compcert}) and translation validators~\cite{NeculaTransVal} are useful for creating soundly optimized versions of a reference program, we are not aware of any that could cope with abstraction-collapsing synthesis as done in this work or LMS.

Myreen and Curello verified a general-purpose big-integer library~\cite{verifiedBignum}.
The code uses a single uniform base system, does not include specialized modular-reduction optimizations, and does not run in constant time.
However, their verification extends down to AMD64 assembly using verified decompilation.
%The proof effort is roughly similar to ours (6227 lines of HOL).

Verification of cryptographic protocols (e.g., FCF~\cite{FCF}, CertiCrypt~\cite{CertiCrypt}) is complementary to this work: given a good formal specification of a protocol, it can be shown separately that an implementation corresponds to the protocol (as we do for elliptic-curve primitives used in TLS) and that the protocol is secure (out of scope for this paper).
The work by Beringer et al.~\cite{HMAC} is a good example of this pattern, composing a protocol-security proof, a correctness proof for its C-level implementation, and a correctness proof for the C compiler.

\section{Discussion}\label{conclusion}

We would like to remark on the aspects of elliptic-curve-cryptography implementation that made this approach work as well as it did, to aid future application in other contexts.
The most general (and perhaps the most important) takeaway is that effort put into structuring code in the most instructive manner possible pays off double during verification, enough to justify the development of new tooling to make that code run fast.
In cases where generalizing an algorithm makes its operation and invariants more apparent, we think it simply makes sense to prove correctness for the general version and use partial evaluation to derive the desired code, even if a specialized implementation has already been written.

On the other hand, we do not believe that the phase distinction between arithmetic and bounds analysis is fundamental, although it is convenient.
When viewing uniform positional representations as polynomials, this distinction corresponds to finding a bound on the degree of the polynomial using partial evaluation and finding bounds on its coefficients using certified compilation, but cryptographic implementation literature (e.g.~\cite{curve25519}) has treated degree and coefficient ranges rather similarly.
We are aware of one example that would be simpler if the high-level algorithms operated on tuples $(l, x, h)$ s.t. $l \le x \le h$ instead of arbitrary-precision integers $x$.
The BoringSSL P-256 implementation that was replaced with our code used prime-shape-specific modular reduction with negative $c_i$ and avoided limb underflow by adding the modulus to a field element before any addition where the input limb ranges did not already rule out underflow.
A certified compiler that needs to preserve the exact value of each limb cannot use this strategy, but it could be encoded in the framework so far by figuring out when to add the modulus for balance in some other way in the high-level algorithm.

The bounds-analysis pipeline described in this paper only works on straight-line code, greatly simplifying the implementation.
It is geared towards being used on the bodies of inner loops where naively picking the largest available integer type for each variable would be costly, but temporaries are numerous enough that picking sizes by hand would be tedious and error-prone.
The analysis could be generalized to loops and recursion by iterating until convergence, but it is unclear whether that is called for in any application.

While we do appreciate being able to support a number of plausibly useful elliptic curves with the same (parameterized) code and proof,
we believe that engineering benefits, especially simplicity of code and proof, make this approach worth considering even when targeting a single configuration.

We find it arbitrary and unfortunate that arithmetic algorithms need to be written in continuation-passing style.
As CPS was only necessary for using Coq's term reduction for partial evaluation, writing a certified partial evaluator would be sufficient to avoid it.
Preliminary experiments have given promising results, but we have not finalized the interface yet.
We also expect that replacing the controlled term reduction with a certified compilation pass would significantly improve the compilation time.
%\todo{say something about having to remove dependent types because we don't know how to reify them?}

We would like to shrink our trusted base by connecting to a certified compiler targeting assembly.
However, existing compilers are not smart enough at mapping the data flow between instructions onto time and space available on the hardware.
Thus another fruitful future-work area is studying those optimizations, principally combined register allocation and instruction scheduling, even independently of proof.
Alternatively, hand-optimized assembly code could be checked to implement the same machine-word-level computation as our generated code using an equality-checking engine like those found in modern SMT solvers (e.g., \cite{simplify}), allowing verification of optimizations that are out-of-reach for compilers, for example effective use of vector instructions.

\section*{Acknowledgments}

This work was supported in part by a Google Research Award and National Science Foundation grants CCF-1253229, CCF-1512611, and CCF-1521584.
We benefited greatly from a fruitful collaboration with Google involving David Benjamin, Thai Duong, Adam Langley, Dominic Rizzo, and Marius Schilder.
Robert Sloan contributed to this project as a student at MIT, before joining Google.
We thank Jason Donenfeld for teaching us how to benchmark arithmetic code with Linux kernel modules, as well as for setting up benchmarks for popular Curve25519 implementations.
For comments on drafts of the paper, we thank Daniel J. Bernstein, Tej Chajed, Istvan Chung, Karl Samuel Gruetter, Ivan Kuraj, Adam Langley, Derek Leung, Devin Neal, Rahul Sridhar, Peng Wang, Ray Wang, and Daniel Ziegler.

\balance
\bibliography{fiat-crypto}

\newpage

\appendices

\section{\label{bugsapp}Study of Bugs in Similar Crypto Code}
\begin{figure*}
  \begin{tabular}{llll}
    Reference/\emph{Category}                                                           & Specification               & Implementation              & Defect       \\
 \hline \emph{Carrying}\\ \hline
 \href{https://github.com/golang/go/issues/13515}{go\#13515}               & Modular exponentiation      & \texttt{uintptr}-sized Montgomery form, Go & carry handling\\
 \href{https://tweetnacl.cr.yp.to/tweetnacl-20131229.pdf}{\cite[p. 2]{TweetNaCl}\textsuperscript{\ref{tweetnacl-correction}}} & F25519 mul, square          & \texttt{ed25519-amd64-64-24k} & carry handling\\
 \href{https://github.com/openssl/openssl/commit/29851264f11ccc70c6c0140d7e3d8d93ef5c9b11}{openssl\#ef5c9b11} & Modular exponentiation & 64-bit Montgomery form, AMD64 & carry handling\\
 \href{https://git.openssl.org/gitweb/?p=openssl.git;a=commitdiff;h=4b8736a22e758c371bc2f8b3534dc0c274acf42c;ds=sidebyside}{openssl\#74acf42c} & Poly1305 & multiple implementations & carry handling \\
 \href{https://git.lysator.liu.se/nettle/nettle/commit/c71d2c9d20eeebb985e3872e4550137209e3ce4d}{nettle\#09e3ce4d} & secp-256r1 modular reduction & & carry handling\\
 \href{https://www.openssl.org/news/secadv/20170126.txt}{CVE-2017-3732} & $x^2 \bmod m$ & Montgomery form, AMD64 assembly & \href{https://boringssl.googlesource.com/boringssl/+/d103616db14ca9587f074efaf9f09a48b8ca80cb%5E%21/}{carry, exploitable}\\
 \href{https://rt.openssl.org/Ticket/Display.html?id=1593&user=guest&pass=guest}{openssl\#1593} & P384 modular reduction & & \href{https://eprint.iacr.org/2011/633.pdf}{carry, exploitable}\\
    \href{https://web.archive.org/web/20160305001036/http://blog.skylable.com/2014/05/tweetnacl-carrybit-bug/}{tweetnacl-U32} & irrelevant & bit-twiddly C & assumed 32-bit \texttt{long}\\
    \hline \emph{Canonicalization} \\ \hline
 \href{https://github.com/agl/curve25519-donna/commit/2647eeba59fb628914c79ce691df794a8edc799f}{donna\#8edc799f} & GF($2^{255}-19$) internal to wire &  32-bit pseudo-Mersenne, C & non-canonical\\
 \href{https://git.openssl.org/gitweb/?p=openssl.git;a=commitdiff;h=b62b2454fadfccaf5e055a1810d72174c2633b8f;ds=sidebyside}{openssl\#c2633b8f} & a + b mod p256 &  Montgomery form, AMD64 assembly & \href{https://mta.openssl.org/pipermail/openssl-dev/2016-August/008179.html}{non-canonical}\\
 \href{http://seb.dbzteam.org/blog/2014/04/28/tweetnacl_arithmetic_bug.html}{tweetnacl-m15} & GF($2^{255}-19$) freeze & bit-twiddly C & bounds? typo?\\
    \hline \emph{Misc. number system} \\ \hline
 \href{https://rt.openssl.org/Ticket/Display.html?id=3607}{openssl\#3607}  & P-256 field element squaring & 64-bit Montgomery form, AMD64 & limb overflow\\
 \href{https://git.openssl.org/gitweb/?p=openssl.git;a=commitdiff;h=dc3c5067cd90f3f2159e5d53c57b92730c687d7e;ds=sidebyside}{openssl\#0c687d7e} & Poly1305 & 32-bit pseudo-Mersenne, x86 and ARM & bad truncation\\
 \href{https://github.com/openssl/openssl/commit/a7a44ba55cb4f884c6bc9ceac90072dea38e66d0}{CVE-2014-3570} & Bignum squaring & asm &  limb overflow\\
 \href{https://github.com/golang/crypto/commit/84e98f45760e87786b7f24603b8166a6fa09811d}{go\#fa09811d} & poly1305 reduction & AMD64 asm, missing subtraction of 3 & found quickly\\
 \href{https://git.openssl.org/gitweb/?p=openssl.git;a=commitdiff;h=bbe9769ba66ab2512678a87b0d9b266ba970db05;ds=sidebyside}{openssl\#a970db05} & Poly1305 &  Lazy reduction in x86 asm & lost bit 59\\
 \href{https://git.openssl.org/gitweb/?p=openssl.git;a=commitdiff;h=1ea8ae5090f557fea2e5b4d5758b10566825d74b;ds=sidebyside}{openssl\#6825d74b} & Poly1305 & AVX2 addition and reduction & bounds?\\
 \href{https://github.com/bitcoin-core/secp256k1/commit/5de4c5dffd22aa4510a5c97d0ad4a9c2eed71d85}{bitcoin\#eed71d85} & ECDSA-secp256k1 x*B & mixed addition Jacobian+Affine & missing case\\
    \hline \emph{Elliptic Curves} \\ \hline
 \href{https://ed25519.cr.yp.to/python/ed25519.py}{ed25519.py} & Ed25519 & accepts signatures other impls reject & missing $h \bmod l$\\
 \href{http://hg.openjdk.java.net/jdk8u/jdk8u/jdk/rev/d99101781d7e}{openjdk\#01781d7e} & EC scalarmult & mixed addition Jacobian+Affine & \href{https://twitter.com/asanso/status/887691580018176000}{missing case}\\
 \href{https://blogs.adobe.com/security/2017/03/critical-vulnerability-uncovered-in-json-encryption.html}{jose-adobe} & ECDH-ES & 5 different libraries affected & not on curve\\
 \href{http://euklid.org/pdf/ECC_Invalid_Curve.pdf}{invalid-curve} & NIST ECDH & irrelevant & not on curve\\
 \href{https://github.com/google/end-to-end/issues/340}{end-to-end\#340} & Curve25519 library & twisted Edwards coordinates & $(0, 1) = \infty$\\
 \href{https://git.openssl.org/gitweb/?p=openssl.git;a=commitdiff;h=e3057a57caf4274ea1fb074518e4714059dfcabf;ds=sidebyside}{openssl\#59dfcabf} & Weier. affine \texttt{<->} Jacobian &  Montgomery form, AMD64 and C & $\infty$ confusion\\
    \hline \emph{Crypto Primitives} \\ \hline
 \href{http://www.dest-unreach.org/socat/contrib/socat-secadv7.html}{socat\#7} & DH in Z*p & irrelevant & non-prime $p$\\
 \href{https://web.archive.org/web/20071010042708/http://www.imc.org/ietf-openpgp/mail-archive/msg14307.html}{CVE-2006-4339} & RSA-PKCS-1 sig. verification & irrelevant & padding check\\
 \href{https://getmonero.org/2017/05/17/disclosure-of-a-major-bug-in-cryptonote-based-currencies.html}{CryptoNote}
    & Anti-double-spending tag &  additive curve25519 curve point & missed order$(P) \neq l$\\
  \end{tabular}
  \caption{Survey of bugs in algebra-based cryptography implementations}\label{bugs}
\end{figure*}

Highly optimized handwritten cryptographic arithmetic implementations have an uninspiring history of serious bugs, even when written and audited by respected experts.
To get a sense of the details, we surveyed project bug trackers and other Internet sources, stopping after finding 26 bugs (each hyperlinked to its bug report) in implementation of nontrivial cryptography-specific optimizations.
Fig. \ref{bugs} summarizes our findings, in terms of 5 emergent categories.
The first three categories have to do with representing large integers using multiple machine-word-sized integers, with custom positional number systems.
Carrying is fairly analogous to the same concept in grade-school arithmetic, and canonicalization is a part of converting back from a custom representation into a standard one.
Elliptic curve formulas are part of high-level algebraic optimizations, above the level of operations on single large integers.
Crypto primitives bring it all together to provide functionality like digital signatures.

Here is a sampling of root causes behind these bugs.
\begin{itemize}
\item Mechanical errors:
One of the two bugs uncovered in OpenSSL issue 3607 was summarized by its author as ``Got math wrong :-('', which we think refers to a pencil-and-paper execution of numerical range analysis.
    The discussion was concluded when the patched version was found to be ``good for \textasciitilde{}6B random tests'' and the reviewer saw that ``there aren't any low-hanging bugs left.''
In \texttt{ed25519-amd64-64-24k}, one of 16,184 repetitive (and handwritten) lines should have been \texttt{r2 += 0 + carry} instead of \texttt{r1 += 0 + carry} \cite[p. 2]{TweetNaCl}.
\item Confusion over intermediate specifications:
OpenSSL bug 1953 was traced back to confusion between the postconditions of exact division with remainder and an operation that produces a $q$ and $r$ s.t. $x = qm + r$ but does not guarantee that $r$ is the smallest possible.
    The probability of a random test triggering this bug was bounded to $10\cdot{}2^{-29}$~\cite{OpenSSLBignumExploit}.
\item Mathematical misconceptions:
  The
    \href{https://getmonero.org/2017/05/17/disclosure-of-a-major-bug-in-cryptonote-based-currencies.html}{CryptoNote
    doublespending bug} arose from use of an algorithm on a composite-order
    elliptic curve when it is only applicable in a prime-order group.
\end{itemize}

\section{\label{data}Full Results of Many-Primes Experiments}

Tables \ref{table64} and \ref{table32} contain the full results of our performance experiments on many primes.
Recall the basic experimental setup:
\begin{itemize}
\item Scrape all prime numbers mentioned in the archives of the ECC mailing list \texttt{curves@moderncrypto.org}.  Crucially, we record not just the numeric values of the primes but also the \emph{ways in which they are expressed} in terms of additions and subtractions of powers of 2 and small multiples thereof.
\item We run a small Python script (shorter than 300 lines) to inspect the shapes of these prime formulas, using simple heuristics to choose the parameters to our Coq library: not just a prime modulus of arithmetic but also how to divide a big integer into digits and which sequence of carry operations to perform in modular reduction.  Actually, the script generates multiple variants, by considering 64-bit vs. 32-bit hardware architectures and by considering the Montgomery and Solinas arithmetic strategies.  The main operation defined in each case is a Montgomery ladder step.
\item We run our Coq pipeline on every variant, culminating in C code pretty-printed for each one.
\item 64-bit configurations are compiled and run on an x86 Linux desktop machine, while 32-bit configurations are compiled and run on an ARM Android mobile device.  We save the running time of each variation.
\item We also compile and run fixed C and C++ implementations using \texttt{libgmp}.
\end{itemize}

The three comparison implementations are:
\begin{itemize}
\item GMP C constant-time, the best comparison with the goals of the code we generate, since running time is required to be independent of integer inputs
\item GMP C variable time, taking advantage of additional optimizations that leak input values through timing
\item GMP C++, the only one of the comparison implementations that does not include manual memory management
\end{itemize}

All three comparison programs are conventional in that they are fixed C or C++ programs, where the prime modulus is set as a preprocessor macro.
It is up to the C compiler and \texttt{libgmp} to take advantage of properties of each modulus.
The final column in each table shows how much better our specialized generation does.
We take the ratio of variable-time C GMP (the fastest GMP code) to whichever of our generated variants is faster.

Some columns in the tables contain dashes in place of numbers of seconds needed for one trial.
Those spots indicate configurations where our Coq compilation currently times out or exhausts available memory.
Considering that Coq is not designed as a platform for executing an optimizing compiler, we are pleased that we get as many successful compilations as we do!
However, we continue working on optimizations to our implementation, to push up the size of prime whose code we can compile quickly.
The timing bottleneck is generally in reification, where repeated inefficient manipulation of terms and contexts by Ltac incurs significant overhead.
The memory bottleneck generally shows up at \texttt{Qed}-time.
Note also that some configurations are expected to fail to build, for instance when applying the Solinas strategy to so-called ``Montgomery-friendly'' primes like $2^{256} - 88 \cdot 2^{240} - 1$, where implementation experts would never choose Solinas.

Among successful compilations, time ranges between tens of seconds and levels best run overnight.

% \section{\label{weighing}Weighing Our Code Base}
%
% %Lacking an established method for quantifying how difficult some piece of code is to maintain, we will resort to a mix of presenting solid (but rather uninformative) numbers and providing anecdotes about our experience developing this library.
%
% % While we started working on the first version of this library two years ago, real progress was made rather sproadically.
% % We implemented new features as we figured out how to do them nicely, making no attempt to work continuously.
% % Even so, the vast majority of the work we did has been replaced with alternatives that in hindsight seem obviously superior.
% % For these reasons, we find it rather meaningless to provide any numeric estimate of developer time -- numbers from 1 person-year to 8 person years could constructed using various thresholds for counting activity.
%
% % We also do not see our library as a feat in large-scale implementation or a collection of programming pearls.
% % Out of the 38000 lines of our code relevant to this paper, around 15000 would probably be best migrated to the Coq standard library.
% % For example, this includes numerous simple lemmas about the basic operations on natural numbers and integers along with tactic code for manipulating standard library objects.
% % Maybe a half of these were additions were created for this project, with the rest being salvaged from other projects.
% % Maybe a quarter of our code (in each file) is braindead boilerplate, working around deficencies in Coq's metaprogramming facilities.
% % To maintain compatability with tactic-based code generation, we cannot use the Coq module system for parametrization and need to import definitions from other files one-by-one, specifiying their parameters again each time.
% % Similarly, to preserve sanity while writing tactics that operate on algebraic code, we wrote a 2000-line algebraic hierarchy library that mirror's Coq's standard library in most aspects but doesn't require all uses of (for example) an operation that happens to form a ring to be marked as such ahead of time.
% % Every definition in our basesytem library is duplicated in CPS and non-CPS form, and the two are proven equivalent -- while avoiding this isn't trivial, a certified CPS-compiler for the relecant fragment of Gallina would do the trick.
% % Some of this is already improving: every time Coq developers fix a bug in the \texttt{nsatz} tactic, one of our elliptic curve proofs drops from several hundred lines to one line.
% % This list could go on; measuring this library by weight is meaningless again.
%
% We can also give a short quantitative summary of our code base, as a proxy for work required to develop and maintain it.
% Just the code for unsaturated arithmetic, described in Section \ref{basesystems}, requires 160 lines (each 80 characters or less) of code and proof, which can then trivially be used to generate all unsaturated reduction examples in this paper.
% However, trying to synthesize a chained carry operation using this code would result in exponential blow-up due to loss of sharing.
% Rewriting functions in continuation-passing style to force sharing of subexpressions (as demonstrated in Section \ref{partial}), the same library grows to around 1000 lines.
% The extensions for saturated arithmetic add 800 lines, and all proofs and code we needed to add for Montgomery reduction total 1500 lines.
% For comparison, the \texttt{amd64-51} implementation of X25519 contains 1900
% lines of assembly code (with P-256 implementations being substantially longer),
% and the translator from assembly to SMT-solver formulas used by \cite{verif25519} is 8800 lines of OCaml.
% The elliptic-curves library is rather thin: a total of 1300 lines of code for 3 curve shapes and 8 point formats, including all equivalence proofs.
% Certified-compiler-phase implementations are rather verbose and sometimes include more lines for parameters than actual code: the largest one is bounds inference (about 2500 lines code+proof), adding up to a total around 15,000 lines.
% With another 15,000 lines of utility lemmas and tactics that seem reasonable candidates to move into Coq's standard library, the total development adds up to 38,000 lines.

\pagebreak
\footnotetext{\label{tweetnacl-correction}An earlier version of this paper (prepublication) incorrectly referred to this implementation as NaCl ed25519. In fact, the bug was discovered as the code underwent NaCl's auditing process and never made it into a release.}
\ %
\clearpage
\nobalance

\bottomcaption{\label{table64}Full 64-bit benchmark data. Our code tried both Solinas and Montgomery implementations for each prime, and we test against three GMP-based implementations: one that is constant-time (gmpsec), one that is variable time (gmpvar), and GMP's C++ API. Our code is constant-time, so gmpsec is the best comparison; however, even with that constraint removed from GMP and not us, we compare favorably to gmpvar.}

\tablehead{%
  \hline
  & \multicolumn{2}{c|}{\textbf{Our Code}} & \multicolumn{3}{c|}{\textbf{GMP Code}} & \\
  \cline{2-6}
  \textbf{Prime} & \textbf{Sol.} & \textbf{Mont.} & \textbf{const time} & \textbf{var time} & \textbf{C++} & \textbf{Speed\allowbreak-up} \\ \hline}
\footnotesize
\centering
\begin{xtabular}{|l|p{0.475cm}|p{0.75cm}@{}|p{0.7cm}@{}|p{0.6cm}@{}|p{0.6cm}@{}|p{0.75cm}@{}|@{}}
	$2^{127}-1$ & 0.03 & 0.04 & 0.26 & 0.15 & 0.67 & 5.0 \\
	$2^{129}-25$ & 0.03 & 0.07 & 0.38 & 0.27 & 0.8 & 9.0 \\
	$2^{130}-5$ & 0.03 & 0.09 & 0.39 & 0.28 & 0.79 & 9.33 \\
	$2^{137}-13$ & 0.03 & 0.08 & 0.37 & 0.27 & 0.8 & 9.0 \\
	$2^{140}-27$ & 0.03 & 0.08 & 0.38 & 0.27 & 0.8 & 9.0 \\
	$2^{141}-9$ & 0.03 & 0.08 & 0.39 & 0.27 & 0.83 & 9.0 \\
	$2^{150}-3$ & 0.03 & 0.08 & 0.38 & 0.3 & 0.8 & 10.0 \\
	$2^{150}-5$ & 0.03 & 0.08 & 0.39 & 0.29 & 0.84 & 9.67 \\
	$2^{152}-17$ & 0.03 & 0.08 & 0.38 & 0.27 & 0.82 & 9.0 \\
	$2^{158}-15$ & 0.03 & 0.08 & 0.37 & 0.27 & 0.76 & 9.0 \\
	$2^{165}-25$ & 0.03 & 0.08 & 0.38 & 0.27 & 0.78 & 9.0 \\
	$2^{166}-5$ & 0.03 & 0.08 & 0.39 & 0.27 & 0.79 & 9.0 \\
	$2^{171}-19$ & 0.03 & 0.08 & 0.38 & 0.27 & 0.79 & 9.0 \\
	$2^{174}-17$ & 0.03 & 0.08 & 0.38 & 0.28 & 0.78 & 9.33 \\
	$2^{174}-3$ & 0.03 & 0.08 & 0.38 & 0.27 & 0.78 & 9.0 \\
	$2^{189}-25$ & 0.04 & 0.08 & 0.39 & 0.27 & 0.8 & 6.75 \\
	$2^{190}-11$ & 0.04 & 0.08 & 0.38 & 0.27 & 0.78 & 6.75 \\
	$2^{191}-19$ & 0.04 & 0.09 & 0.36 & 0.26 & 0.78 & 6.5 \\
	$2^{192}-2^{64}-1$ & 0.05 & 0.07 & 0.31 & 0.24 & 0.79 & 4.8 \\
	$2^{194}-33$ & 0.04 & 0.12 & 0.5 & 0.34 & 0.93 & 8.5 \\
	$2^{196}-15$ & 0.04 & 0.12 & 0.5 & 0.34 & 0.89 & 8.5 \\
	$2^{198}-17$ & 0.04 & 0.12 & 0.51 & 0.34 & 0.87 & 8.5 \\
	$2^{205}-45\cdot 2^{198}-1$ & - & 0.14 & 0.51 & 0.34 & 0.87 & 2.43 \\
	$2^{206}-5$ & 0.04 & 0.14 & 0.5 & 0.34 & 0.84 & 8.5 \\
	$2^{212}-29$ & 0.05 & 0.12 & 0.49 & 0.35 & 0.87 & 7.0 \\
	$2^{213}-3$ & 0.04 & 0.13 & 0.49 & 0.37 & 0.88 & 9.25 \\
	$2^{216}-2^{108}-1$ & 0.04 & 0.12 & 0.51 & 0.35 & 0.88 & 8.75 \\
	$2^{221}-3$ & 0.05 & 0.15 & 0.51 & 0.36 & 0.89 & 7.2 \\
	$2^{222}-117$ & 0.05 & 0.12 & 0.53 & 0.35 & 0.91 & 7.0 \\
	$2^{224}-2^{96}+1$ & - & 0.13 & 0.5 & 0.35 & 0.88 & 2.69 \\
	$2^{226}-5$ & 0.04 & 0.13 & 0.5 & 0.35 & 0.92 & 8.75 \\
	$2^{230}-27$ & 0.05 & 0.13 & 0.54 & 0.35 & 0.91 & 7.0 \\
	$2^{235}-15$ & 0.06 & 0.13 & 0.5 & 0.34 & 0.89 & 5.67 \\
	$2^{243}-9$ & 0.06 & 0.13 & 0.5 & 0.34 & 0.89 & 5.67 \\
	$2^{251}-9$ & 0.06 & 0.13 & 0.5 & 0.35 & 0.94 & 5.83 \\
	$2^{254}-127\cdot 2^{240}-1$ & - & 0.12 & 0.5 & 0.35 & 0.92 & 2.92 \\
	$2^{255}-19$ & 0.06 & 0.13 & 0.48 & 0.35 & 0.9 & 5.83 \\
	$2^{255}-765$ & 0.06 & 0.13 & 0.52 & 0.34 & 0.9 & 5.67 \\
	$2^{256}-189$ & 0.06 & 0.14 & 0.38 & 0.34 & 0.87 & 5.67 \\
	\makecell{$2^{256}-2^{224}$ \\ \quad $+2^{192}+2^{96}-1$} & - & 0.11 & 0.38 & 0.33 & 0.84 & 3.0 \\
	$2^{256}-2^{32}-977$ & 0.1 & 0.12 & 0.38 & 0.34 & 0.87 & 3.4 \\
	$2^{256}-4294968273$ & 0.14 & 0.13 & 0.37 & 0.34 & 0.86 & 2.62 \\
	$2^{256}-88\cdot 2^{240}-1$ & - & 0.11 & 0.39 & 0.34 & 0.88 & 3.09 \\
	$2^{266}-3$ & 0.06 & 0.18 & 0.66 & 0.45 & 1.13 & 7.5 \\
	$2^{285}-9$ & 0.06 & 0.18 & 0.73 & 0.43 & 0.97 & 7.17 \\
	$2^{291}-19$ & 0.07 & 0.18 & 0.68 & 0.42 & 1.0 & 6.0 \\
	$2^{321}-9$ & 0.1 & 0.26 & 0.8 & 0.54 & 1.18 & 5.4 \\
	$2^{322}-2^{161}-1$ & 0.07 & 0.27 & 0.83 & 0.53 & 1.15 & 7.57 \\
	$2^{336}-17$ & 0.1 & 0.27 & 0.8 & 0.53 & 1.11 & 5.3 \\
	$2^{336}-3$ & 0.09 & 0.27 & 0.86 & 0.53 & 1.08 & 5.89 \\
	$2^{338}-15$ & 0.1 & 0.25 & 0.8 & 0.54 & 1.06 & 5.4 \\
	$2^{369}-25$ & 0.13 & 0.26 & 0.79 & 0.52 & 1.1 & 4.0 \\
	$2^{379}-19$ & 0.12 & 0.26 & 0.79 & 0.55 & 1.07 & 4.58 \\
	$2^{382}-105$ & 0.13 & 0.25 & 0.92 & 0.57 & 1.11 & 4.38 \\
	$2^{383}-187$ & 0.13 & 0.28 & 0.75 & 0.5 & 1.05 & 3.85 \\
	$2^{383}-31$ & 0.13 & 0.26 & 0.75 & 0.51 & 1.05 & 3.92 \\
	$2^{383}-421$ & 0.13 & 0.25 & 0.76 & 0.51 & 1.06 & 3.92 \\
	\makecell{$2^{384}-2^{128}$ \\ \quad $-2^{96}+2^{32}-1$} & - & 0.25 & 0.64 & 0.47 & 0.98 & 1.88 \\
	$2^{384}-317$ & 0.13 & 0.26 & 0.67 & 0.48 & 1.0 & 3.69 \\
	$2^{384}-5\cdot 2^{368}-1$ & - & 0.23 & 0.63 & 0.46 & 0.99 & 2.0 \\
	$2^{384}-79\cdot 2^{376}-1$ & - & 0.23 & 0.62 & 0.46 & 0.99 & 2.0 \\
	$2^{389}-21$ & 0.13 & - & 0.97 & 0.6 & 1.22 & 4.62 \\
	$2^{401}-31$ & 0.14 & - & 0.97 & 0.61 & 1.17 & 4.36 \\
	$2^{413}-21$ & 0.16 & - & 0.99 & 0.62 & 1.22 & 3.88 \\
	$2^{414}-17$ & 0.15 & - & 0.98 & 0.6 & 1.21 & 4.0 \\
	$2^{416}-2^{208}-1$ & 0.14 & - & 0.98 & 0.6 & 1.16 & 4.29 \\
	$2^{444}-17$ & 0.17 & - & 0.96 & 0.6 & 1.2 & 3.53 \\
	$2^{448}-2^{224}-1$ & 0.12 & - & 0.79 & 0.52 & 1.06 & 4.33 \\
	$2^{450}-2^{225}-1$ & 0.13 & - & 1.22 & 0.74 & 1.34 & 5.69 \\
	$2^{452}-3$ & 0.16 & - & 1.24 & 0.71 & 1.32 & 4.44 \\
	$2^{468}-17$ & 0.16 & - & 1.23 & 0.71 & 1.29 & 4.44 \\
	$2^{480}-2^{240}-1$ & 0.13 & - & 1.18 & 0.71 & 1.28 & 5.46 \\
	$2^{488}-17$ & 0.19 & - & 1.2 & 0.7 & 1.28 & 3.68 \\
	$2^{489}-21$ & 0.2 & - & 1.17 & 0.69 & 1.27 & 3.45 \\
	$2^{495}-31$ & 0.19 & - & 1.17 & 0.69 & 1.3 & 3.63 \\
	$2^{510}-290\cdot 2^{496}-1$ & - & - & 1.2 & 0.7 & 1.28 & - \\
	$2^{511}-187$ & 0.25 & - & 1.13 & 0.66 & 1.21 & 2.64 \\
	$2^{511}-481$ & 0.25 & - & 1.12 & 0.66 & 1.24 & 2.64 \\
	$2^{512}-491\cdot 2^{496}-1$ & - & - & 0.99 & 0.62 & 1.15 & - \\
	$2^{512}-569$ & 0.24 & - & 0.95 & 0.62 & 1.14 & 2.58 \\
	$2^{521}-1$ & 0.18 & - & 1.4 & 0.81 & 1.44 & 4.5 \\
\hline
\end{xtabular}

\clearpage

\bottomcaption{\label{table32}Full 32-bit benchmark data. Many of the 32-bit Montgomery implementations exceeded the one-hour timeout for proofs, because 32-bit code involves approximately four times as many operations. The C++ GMP program was not benchmarked on 32-bit.}

\tablehead{%
  \hline
  & \multicolumn{2}{c|}{\textbf{Our Code}} & \multicolumn{2}{c|}{\textbf{GMP Code}} & \\
  \cline{2-5}
  \textbf{Prime} & \textbf{Sol.} & \textbf{Mont.} & \textbf{const time} & \textbf{var time} & \textbf{Speed\allowbreak-up} \\ \hline}
\footnotesize
\centering
\begin{xtabular}{|l|p{0.6cm}@{}|p{0.75cm}@{}|p{0.75cm}@{}|p{0.7cm}@{}|p{0.75cm}@{}|}
	$2^{127}-1$ & 0.3 & 1.19 & 2.86 & 3.23 & 9.53 \\
	$2^{129}-25$ & 0.35 & 1.7 & 3.38 & 3.77 & 9.66 \\
	$2^{130}-5$ & 0.44 & 1.87 & 3.56 & 3.79 & 8.09 \\
	$2^{137}-13$ & 0.48 & 2.06 & 3.41 & 3.78 & 7.1 \\
	$2^{140}-27$ & 0.51 & 1.98 & 3.43 & 3.77 & 6.73 \\
	$2^{141}-9$ & 0.51 & 2.0 & 3.43 & 3.81 & 6.73 \\
	$2^{150}-3$ & 0.42 & 2.0 & 3.56 & 3.79 & 8.48 \\
	$2^{150}-5$ & 0.49 & 1.99 & 3.38 & 3.8 & 6.9 \\
	$2^{152}-17$ & 0.5 & 1.96 & 3.4 & 3.82 & 6.8 \\
	$2^{158}-15$ & 0.52 & 2.04 & 3.4 & 3.77 & 6.54 \\
	$2^{165}-25$ & 0.59 & 2.46 & 4.02 & 4.45 & 6.81 \\
	$2^{166}-5$ & 0.61 & 2.43 & 4.02 & 4.43 & 6.59 \\
	$2^{171}-19$ & 0.57 & 2.68 & 4.04 & 4.51 & 7.09 \\
	$2^{174}-17$ & 0.58 & 2.63 & 4.03 & 4.39 & 6.95 \\
	$2^{174}-3$ & 0.61 & 2.62 & 4.02 & 4.4 & 6.59 \\
	$2^{189}-25$ & 0.7 & 2.65 & 4.05 & 4.4 & 5.79 \\
	$2^{190}-11$ & 0.71 & 2.64 & 4.1 & 4.42 & 5.77 \\
	$2^{191}-19$ & 0.66 & 2.69 & 4.03 & 4.4 & 6.11 \\
	$2^{192}-2^{64}-1$ & - & 2.41 & 3.56 & 4.23 & 1.48 \\
	$2^{194}-33$ & 0.75 & - & 4.66 & 4.94 & 6.21 \\
	$2^{196}-15$ & 0.77 & - & 4.64 & 4.94 & 6.03 \\
	$2^{198}-17$ & 0.76 & - & 4.72 & 4.97 & 6.21 \\
	$2^{205}-45\cdot 2^{198}-1$ & - & - & 4.66 & 5.03 & - \\
	$2^{206}-5$ & 0.76 & - & 4.62 & 4.91 & 6.08 \\
	$2^{212}-29$ & 0.86 & - & 4.68 & 4.91 & 5.44 \\
	$2^{213}-3$ & 0.7 & - & 4.68 & 4.94 & 6.69 \\
	$2^{216}-2^{108}-1$ & 0.7 & - & 4.67 & 4.92 & 6.67 \\
	$2^{221}-3$ & 0.8 & - & 4.68 & 4.92 & 5.85 \\
	$2^{222}-117$ & 0.87 & - & 4.72 & 4.87 & 5.43 \\
	$2^{224}-2^{96}+1$ & - & - & 4.13 & 4.85 & - \\
	$2^{226}-5$ & 0.87 & - & 5.25 & 5.65 & 6.03 \\
	$2^{230}-27$ & 0.83 & - & 5.29 & 5.71 & 6.37 \\
	$2^{235}-15$ & 0.9 & - & 5.31 & 5.69 & 5.9 \\
	$2^{243}-9$ & 0.86 & - & 5.29 & 5.62 & 6.15 \\
	$2^{251}-9$ & 1.12 & - & 5.3 & 5.65 & 4.73 \\
	$2^{254}-127\cdot 2^{240}-1$ & - & 3.97 & 5.26 & 5.7 & 1.32 \\
	$2^{255}-19$ & 1.01 & - & 5.25 & 5.7 & 5.2 \\
	$2^{255}-765$ & 1.43 & - & 5.27 & 5.71 & 3.69 \\
	$2^{256}-189$ & 1.2 & - & 4.71 & 5.49 & 3.93 \\
	$2^{256}-2^{224}+2^{192}+2^{96}-1$ & - & - & 4.7 & 5.46 & - \\
	$2^{256}-2^{32}-977$ & 1.65 & - & 4.72 & 5.45 & 2.86 \\
	$2^{256}-4294968273$ & - & - & 4.77 & 5.48 & - \\
	$2^{256}-88\cdot 2^{240}-1$ & - & - & 4.78 & 5.46 & - \\
	$2^{266}-3$ & 1.01 & - & 6.1 & 6.32 & 6.04 \\
	$2^{285}-9$ & 1.13 & - & 6.13 & 6.34 & 5.42 \\
	$2^{291}-19$ & 1.33 & - & 6.94 & 6.98 & 5.22 \\
	$2^{321}-9$ & 1.72 & - & 7.6 & 7.66 & 4.42 \\
	$2^{322}-2^{161}-1$ & 1.37 & - & 7.66 & 7.74 & 5.59 \\
	$2^{336}-17$ & 1.67 & - & 7.64 & 7.74 & 4.57 \\
	$2^{336}-3$ & 1.59 & - & 7.58 & 7.69 & 4.77 \\
	$2^{338}-15$ & 1.7 & - & 7.66 & 7.67 & 4.51 \\
	$2^{369}-25$ & 2.44 & - & 8.41 & 9.03 & 3.45 \\
	$2^{379}-19$ & 2.47 & - & 8.44 & 9.25 & 3.42 \\
	$2^{382}-105$ & 2.66 & - & 8.41 & 9.04 & 3.16 \\
	$2^{383}-187$ & 2.63 & - & 8.44 & 9.11 & 3.21 \\
	$2^{383}-31$ & 2.6 & - & 8.47 & 9.13 & 3.26 \\
	$2^{383}-421$ & 3.58 & - & 8.45 & 9.11 & 2.36 \\
	$2^{384}-2^{128}-2^{96}+2^{32}-1$ & - & - & 7.62 & 8.8 & - \\
	$2^{384}-317$ & 3.95 & - & 7.62 & 8.82 & 1.93 \\
	$2^{384}-5\cdot 2^{368}-1$ & - & - & 7.64 & 8.94 & - \\
	$2^{384}-79\cdot 2^{376}-1$ & - & - & 7.66 & 8.84 & - \\
	$2^{389}-21$ & 2.89 & - & 9.41 & 9.93 & 3.26 \\
	$2^{401}-31$ & 2.85 & - & 9.35 & 9.92 & 3.28 \\
	$2^{413}-21$ & 3.53 & - & 9.48 & 9.93 & 2.69 \\
	$2^{414}-17$ & 3.72 & - & 9.4 & 9.86 & 2.53 \\
	$2^{416}-2^{208}-1$ & 2.48 & - & 8.54 & 9.67 & 3.44 \\
	$2^{444}-17$ & 3.7 & - & 10.31 & 10.89 & 2.79 \\
	$2^{448}-2^{224}-1$ & 3.18 & - & 9.57 & 10.51 & 3.01 \\
	$2^{450}-2^{225}-1$ & - & - & 11.37 & 11.63 & - \\
	$2^{452}-3$ & 3.23 & - & 11.33 & 11.63 & 3.51 \\
	$2^{468}-17$ & 3.2 & - & 11.37 & 11.63 & 3.55 \\
	$2^{480}-2^{240}-1$ & 3.58 & - & 10.47 & 11.33 & 2.92 \\
	$2^{488}-17$ & 7.99 & - & 12.23 & 12.92 & 1.53 \\
	$2^{489}-21$ & 7.7 & - & 12.26 & 12.81 & 1.59 \\
	$2^{495}-31$ & 6.07 & - & 12.2 & 13.1 & 2.01 \\
	$2^{510}-290\cdot 2^{496}-1$ & - & - & 12.17 & 12.9 & - \\
	$2^{511}-187$ & 9.73 & - & 12.21 & 13.07 & 1.25 \\
	$2^{511}-481$ & - & - & 12.23 & 12.9 & - \\
	$2^{512}-491\cdot 2^{496}-1$ & - & - & 11.26 & 12.58 & - \\
	$2^{512}-569$ & - & - & 11.23 & 12.55 & - \\
	$2^{521}-1$ & 3.9 & - & 13.3 & 13.91 & 3.41 \\
\hline
\end{xtabular}
\end{document}

### 2020-07-30 papers/source/rewriting/rewriting.tex
Add a bunch of paper sources

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\RequirePackage{luatex85}
\PassOptionsToPackage{dvipsnames}{xcolor}
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2021}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
%\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
\usepackage{wrapfig}
\setlength{\intextsep}{0pt} % don't add space above nor below
\usepackage{subfig}
\newcommand{\subfigureautorefname}{\figureautorefname}
%% http://ctan.org/pkg/subcaption

\usepackage{hyperref}
\usepackage{minted}
\usemintedstyle{bw}
\usepackage{amsthm}
%\usepackage{amssymb}
\usepackage[all]{xy}
\usepackage{stmaryrd}
\usepackage[utf8]{inputenc}
\usepackage{textgreek}
\usepackage{newunicodechar}
\usepackage{usebib}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{bcprules}
\usepackage{multicol}
\usepackage{apptools}
\usepackage{adjustbox}
\usepgfplotslibrary{units}
\usepgfplotslibrary{external}
\tikzexternalize
\newunicodechar{→}{\ensuremath{\to}}
\newunicodechar{λ}{\ensuremath{\lambda}}
\newunicodechar{∀}{\ensuremath{\forall}}

\usepackage{etoolbox,xpatch}
% https://tex.stackexchange.com/a/343506/2066 for avoiding red boxes around minted errors
\makeatletter
\AtBeginEnvironment{minted}{\dontdofcolorbox}
\def\dontdofcolorbox{\renewcommand\fcolorbox[4][]{##4}}
\xpatchcmd{\inputminted}{\minted@fvset}{\minted@fvset\dontdofcolorbox}{}{}
\xpatchcmd{\mintinline}{\minted@fvset}{\minted@fvset\dontdofcolorbox}{}{} % see https://tex.stackexchange.com/a/401250/
\makeatother

% https://tex.stackexchange.com/a/183682/2066
\makeatletter

% define a macro \Autoref to allow multiple references to be passed to \autoref
\newcommand\Autoref[1]{\@first@ref#1,@}
\def\@throw@dot#1.#2@{#1}% discard everything after the dot
\def\@set@refname#1{%    % set \@refname to autoefname+s using \getrefbykeydefault
    \edef\@tmp{\getrefbykeydefault{#1}{anchor}{}}%
    \xdef\@tmp{\expandafter\@throw@dot\@tmp.@}%
    \ltx@IfUndefined{\@tmp autorefnameplural}%
         {\def\@refname{\@nameuse{\@tmp autorefname}s}}%
         {\def\@refname{\@nameuse{\@tmp autorefnameplural}}}%
}
\def\@first@ref#1,#2{%
  \ifx#2@\autoref{#1}\let\@nextref\@gobble% only one ref, revert to normal \autoref
  \else%
    \@set@refname{#1}%  set \@refname to autoref name
    \@refname~\ref{#1}% add autoefname and first reference
    \let\@nextref\@next@ref% push processing to \@next@ref
  \fi%
  \@nextref#2%
}
\def\@next@ref#1,#2{%
   \ifx#2@ and~\ref{#1}\let\@nextref\@gobble% at end: print and+\ref and stop
   \else, \ref{#1}% print  ,+\ref and continue
   \fi%
   \@nextref#2%
}

\makeatother

\newbibfield{author}
\newbibfield{title}
\bibinput{rewriting-lower}
\newcommand{\citetitle}[1]{\emph{\usebibentry{#1}{title}}\nocite{#1}}
\newcommand{\todo}[1]{\textbf{TODO: #1}}
\newcommand{\coqbug}[1]{\href{https://github.com/coq/coq/issues/#1}{\##1}}
\newcommand{\Rtac}{\ensuremath{\mathcal{R}_\text{\emph{tac}}}}
\newcommand{\Ltac}{\ensuremath{\mathcal{L}_\text{\emph{tac}}}}
\newcommand{\defeq}{=}
\makeatletter
\newcommand{\letin}[1][{\ensuremath{\cdots}}{\ensuremath{\cdots}}]{%
  \texttt{let }\@firstoftwo#1\texttt{ in }\@secondoftwo#1
}
\newcommand{\einput}[1]{\@@input #1 \space}
\newcommand{\beginTikzpictureStamped}[2][]{%
  {%
    \everyeof{\noexpand}% IDK why \noexpand is the magic one, but I got it from http://mirrors.ibiblio.org/CTAN/macros/latex/contrib/oberdiek/catchfile.pdf
    \long\xdef\@tikzstamp{#2}%
  }%
  \def\@dobegintikzpicture{\begin{tikzpicture}[#1]}%
  \expandafter\@dobegintikzpicture\expandafter\def\expandafter\tikzstamp\expandafter{\@tikzstamp}%
}
\newcommand{\TikzpictureStamped}[3][]{\beginTikzpictureStamped[#1]{#2}#3\end{tikzpicture}}%
\makeatother
\newcommand{\asserteq}[2]{\ifthenelse{\equal{#1}{#2}}{}{\GenericError{}{Not equal: \detokenize{#1} != \detokenize{#2}}{}{}}}

\makeatletter
\AtAppendix{%
  \newwrite\@appendixpagefile
  \immediate\openout\@appendixpagefile=\jobname.appendix.startpage
  \immediate\write\@appendixpagefile{\detokenize{\def\appendixstartpage}{\thepage}}
}
\newcommand{\supplement}{%
  \clearpage
  \newwrite\@supplementpagefile
  \immediate\openout\@supplementpagefile=\jobname.supplement.startpage
  \immediate\write\@supplementpagefile{\detokenize{\def\supplementstartpage}{\thepage}}
}%
\makeatother

\begin{document}

%% Title information
\title{A Framework for Building Verified Partial Evaluators}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
%\subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{Jason Gross}
%\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{0000-0002-9427-4891}             %% \orcid is optional
\affiliation{
  %\position{Position1}
  \department{CSAIL}              %% \department is recommended
  \institution{MIT}            %% \institution is required
  %\streetaddress{Street1 Address1}
  \city{Cambridge}
  \state{MA}
  \postcode{02139}
  \country{USA}                    %% \country is recommended
}
\email{jgross@mit.edu}          %% \email is recommended

%% Author with two affiliations and emails.
\author{Andres Erbsen}
%\authornote{with author2 note}          %% \authornote is optional;
%% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
    %\position{Position2a}
    \department{CSAIL}             %% \department is recommended
    \institution{MIT}           %% \institution is required
    %\streetaddress{Street2a Address2a}
    \city{Cambridge}
    \state{MA}
    \postcode{02139}
    \country{USA}                    %% \country is recommended
}
\email{andreser@mit.edu}         %% \email is recommended

%% Author with two affiliations and emails.
\author{Adam Chlipala}
%\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  %\position{Position2a}
  \department{CSAIL}             %% \department is recommended
  \institution{MIT}           %% \institution is required
  %\streetaddress{Street2a Address2a}
  \city{Cambridge}
  \state{MA}
  \postcode{02139}
  \country{USA}                    %% \country is recommended
}
\email{adamc@csail.mit.edu}         %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  \emph{Partial evaluation} is a classic technique for generating lean, customized code from libraries that start with more bells and whistles.
  It is also an attractive approach to creation of \emph{formally verified} systems, where theorems can be proved about libraries, yielding correctness of all specializations ``for free.''
  However, it can be challenging to make library specialization both performant (at compile time and runtime) and trustworthy.
  We present a new approach, prototyped in the Coq proof assistant, which supports specialization at the speed of native-code execution, without adding to the trusted code base.
  Our extensible engine, which combines the traditional concepts of tailored term reduction and automatic rewriting from hint databases, is also of interest to replace these ingredients in proof assistants' proof checkers and tactic engines, at the same time as it supports extraction to standalone compilers from library parameters to specialized code.
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
%% \begin{CCSXML}
%% <ccs2012>
%% <concept>
%% <concept_id>10011007.10011006.10011008</concept_id>
%% <concept_desc>Software and its engineering~General programming languages</concept_desc>
%% <concept_significance>500</concept_significance>
%% </concept>
%% <concept>
%% <concept_id>10003456.10003457.10003521.10003525</concept_id>
%% <concept_desc>Social and professional topics~History of programming languages</concept_desc>
%% <concept_significance>300</concept_significance>
%% </concept>
%% </ccs2012>
%% \end{CCSXML}

%% \ccsdesc[500]{Software and its engineering~General programming languages}
%% \ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
%% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction} \label{sec:intro}

Mechanized proof is gaining in importance for development of critical software infrastructure.
Oft-cited examples include the CompCert verified C compiler~\cite{Compcert} and the seL4 verified operating-system microkernel~\cite{seL4SOSP09}.
Here we have very flexible systems that are ready to adapt to varieties of workloads, be they C source programs for CompCert or application binaries for seL4.
For a verified operating system, such adaptation takes place at \emph{runtime}, when we launch the application.
However, some important bits of software infrastructure commonly do adaptation at \emph{compile time}, such that the fully general infrastructure software is not even installed in a deployed system.

Of course, compilers are a natural example of that pattern, as we would not expect CompCert itself to be installed on an embedded system whose application code was compiled with it.
The problem is that writing a compiler is rather labor-intensive, with its crafting of syntax-tree types for source, target, and intermediate languages, its fine-tuning of code for transformation passes that manipulate syntax trees explicitly, and so on.
An appealing alternative is \emph{partial evaluation}~\cite{PartialEvaluation}, which relies on reusable compiler facilities to specialize library code to parameters, with no need to write that library code in terms of syntax-tree manipulations.
Cutting-edge tools in this tradition even make it possible to use high-level functional languages to generate performance-competitive low-level code, as in Scala's Lightweight Modular Staging~\cite{LMS}.

It is natural to try to port this approach to construction of systems with mechanized proofs.
On one hand, the typed functional languages in popular proof assistants' logics make excellent hosts for flexible libraries, which can often be specialized through means as simple as partial application of curried functions.
Term-reduction systems built into the proof assistants can then generate the lean residual programs.
On the other hand, it is surprisingly difficult to realize the last sentence with good performance.
The challenge is that we are not just implementing algorithms; we also want a proof to be checked by a small proof checker, and there is tension in designing such a checker, as fancier reduction strategies grow the trusted code base.
It would seem like an abandonment of the spirit of proof assistants to bake in a reduction strategy per library, yet effective partial evaluation tends to be rather fine-tuned in this way.
Performance tuning matters when generated code is thousands of lines long.

In this paper, we present an approach to verified partial evaluation in proof assistants, which requires no changes to proof checkers.
To make the relevance concrete, we use the example of Fiat Cryptography~\cite{FiatCryptoSP19}, a Coq library that generates code for big-integer modular arithmetic at the heart of elliptic-curve-cryptography algorithms.
This domain-specific compiler has been adopted, for instance, in the Chrome Web browser, such that about half of all HTTPS connections from browsers are now initiated using code generated (with proof) by Fiat Cryptography.
However, Fiat Cryptography was only used successfully to build C code for the two most widely used curves (P-256 and Curve25519).
Their method of partial evaluation timed out trying to compile code for the third most widely used curve (P-384).
Additionally, to achieve acceptable reduction performance, the library code had to be written manually in continuation-passing style.
We will demonstrate a new Coq library that corrects both weaknesses, while maintaining the generality afforded by allowing rewrite rules to be mixed with partial evaluation.

\subsection{A Motivating Example} \label{sec:motivating-example} \label{sec:explain-ident.eagerly} \label{sec:explain-eval-rect} \label{sec:explain-'}

We are interested in partial-evaluation examples that mix higher-order functions, inductive datatypes, and arithmetic simplification.
For instance, consider the following Coq code.

\begin{minted}[fontsize=\small]{coq}
Definition prefixSums (ls:list nat) : list nat :=
  let ls' := combine ls (seq 0 (length ls)) in
  let ls'' := map (λ p, fst p * snd p) ls' in
  let '(_, ls''') := fold_left (λ '(acc, ls''') n,
    let acc' := acc + n in (acc', acc' :: ls''')) ls'' (0, []) in
  ls'''.
\end{minted}
%\todo{Do we want to mention dlet?  We only get the specified output if we use dlet rather than let...}
% Adam's reaction: no, it will confuse readers too much to mention dlet this early.

This function first computes list \mintinline{coq}{ls'} that pairs each element of input list \mintinline{coq}{ls} with its position, so, for instance, list $[a; b; c]$ becomes $[(a, 0); (b, 1); (c, 2)]$.
Then we map over the list of pairs, multiplying the components at each position.
Finally, we traverse that list, building up a list of all prefix sums.

We would like to specialize this function to particular list lengths.
That is, we know in advance how many list elements we will pass in, but we do not know the values of those elements.
For a given length, we can construct a schematic list with one free variable per element.
For example, to specialize to length four, we can apply the function to list \mintinline{coq}{[a; b; c; d]}, and we expect this output:
\begin{minted}[fontsize=\small]{coq}
let acc := b + c * 2 in
let acc' := acc + d * 3 in
  [acc'; acc; b; 0]
\end{minted}

Notice how subterm sharing via \mintinline{coq}{let}s is important.
As list length grows, we avoid quadratic blowup in term size through sharing.
Also notice how we simplified the first two multiplications with $a \cdot 0 = 0$ and $b \cdot 1 = b$ (each of which requires explicit proof in Coq), using other arithmetic identities to avoid introducing new variables for the first two prefix sums of \mintinline{coq}{ls''}, as they are themselves constants or variables, after simplification.

To set up our partial evaluator, we prove the algebraic laws that it should use for simplification, starting with basic arithmetic identities.
\begin{minted}[fontsize=\small]{coq}
Lemma zero_plus : forall n, 0 + n = n.       Lemma times_zero : forall n, n * 0 = 0.
Lemma plus_zero : forall n, n + 0 = n.       Lemma times_one  : forall n, n * 1 = n.
\end{minted}

Next, we prove a law for each list-related function, connecting it to the primitive-recursion combinator for some inductive type (natural numbers or lists, as appropriate).
We use a special apostrophe marker to indicate a quantified variable that may only match with \emph{compile-time constants}.
We also use a further marker \mintinline{coq}{ident.eagerly} to ask the reducer to simplify a case of primitive recursion by complete traversal of the designated argument's constructor tree.
%\todo{Hey, it sure would be nice to generate all of these to-eliminator lemmas automatically, too!  Next step for the plugin?}
%\todo{Reply from Jason: This is actually quite tricky, and I want to claim that it is significantly beyond the scope of the rewriter.  If you have a constant which is already defined in terms of the eliminator, it's quite easy for me to write a tactic that makes the one-step unfolding lemma.  And it's also quite easy to throw \mintinline{coq}{ident.eagerly} in front.  However, if the constant is not already defined in terms of the eliminator constant, and is instead a fixpoint, then you need something that is basically as powerful as Lean's elaborator to figure out the right recursion principle.}
\begin{minted}[fontsize=\small]{coq}
Lemma eval_map A B (f : A -> B) l
: map f l = ident.eagerly list_rect _ _ [] (λ x _ l', f x :: l') l.
Lemma eval_fold_left A B (f : A -> B -> A) l a
: fold_left f l a = ident.eagerly list_rect _ _ (λ a, a) (λ x _ r a, r (f a x)) l a.
Lemma eval_combine A B (la : list A) (lb : list B)
: combine la lb =
list_rect _ (λ _, []) (λ x _ r lb, list_case (λ _, _) [] (λ y ys, (x,y)::r ys) lb) la lb.
Lemma eval_length A (ls : list A)
: length ls = list_rect _ 0 (λ _ _ n, S n) ls.
\end{minted}
%\todo{Explain the Proper lemma and that eagerly is incompatible with do\_again.}
% I think would be too much too soon to talk about do\_again here.

With all the lemmas available, we can package them up into a rewriter, which triggers generation of a specialized rewrite procedure and its soundness proof.
Our Coq plugin introduces a new command \mintinline{coq}{Make} for building rewriters
\begin{minted}[fontsize=\small]{coq}
Make rewriter := Rewriter For (zero_plus, plus_zero, times_zero, times_one, eval_map,
  eval_fold_left, do_again eval_length, do_again eval_combine,
  eval_rect nat, eval_rect list, eval_rect prod) (with delta) (with extra idents (seq)).
\end{minted}
Most inputs to \mintinline{coq}{Rewriter For} list quantified equalities to use for left-to-right rewriting.
However, we also use options \mintinline{coq}{do_again}, to request that some rules trigger an extra bottom-up pass after being used for rewriting; \mintinline{coq}{eval_rect}, to queue up eager evaluation of a call to a primitive-recursion combinator on a known recursive argument; \mintinline{coq}{with delta}, to request evaluation of all monomorphic operations on concrete inputs; and \mintinline{coq}{with extra idents}, to inform the engine of further permitted identifiers that do not appear directly in any of the rewrite rules.

Our plugin also provides new tactics like \mintinline{coq}{Rewrite_rhs_for}, which applies a rewriter to the righthand side of an equality goal.
That last tactic is just what we need to synthesize a specialized \mintinline{coq}{prefixSums} for list length four, along with a proof of its equivalence to the original function.
\begin{minted}[fontsize=\small]{coq}
Definition prefixSums4 :
 {f : nat → nat → nat → nat → list nat | ∀ a b c d, f a b c d = prefixSums [a;b;c;d]}
  := ltac:(eexists; Rewrite_rhs_for rewriter; reflexivity).
\end{minted}

\subsection{Concerns of Trusted-Code-Base Size} \label{sec:trusted-code-base-size}

Crafting a reduction strategy is challenging enough in a standalone tool.
A large part of the difficulty in a proof assistant is reducing in a way that leaves a proof trail that can be checked efficiently by a small kernel.
Most proof assistants present user-friendly surface tactic languages that generate proof traces in terms of more-elementary tactic steps.
The trusted proof checker only needs to know about the elementary steps, and there is pressure to be sure that these steps are indeed elementary, not requiring excessive amounts of kernel code.
However, hardcoding a new reduction strategy in the kernel can bring dramatic performance improvements.
Generating thousands of lines of code with partial evaluation would be intractable if we were outputting sequences of primitive rewrite steps justifying every little term manipulation, so we must take advantage of the time-honored feature of type-theoretic proof assistants that reductions included in the definitional equality need not be requested explicitly.
We discuss the performance issues in more detail in \autoref{sec:setoid-rewrite-bottlenecks}.

Which kernel-level reductions \emph{does} Coq support today?
Currently, the trusted code base knows about four different kinds of reduction: left-to-right conversion, right-to-left conversion, a virtual machine (VM) written in C based on the OCaml compiler, and a compiler to native code.
Furthermore, the first two are parameterized on an arbitrary user-specified ordering of which constants to unfold when, in addition to internal heuristics about what to do when the user has not specified an unfolding order for given constants.
Recently, native support for 63-bit integers~\cite{coq-pr-ints} and IEEE 754-2008 binary64 floats~\cite{coq-pr-floats} have been added to the VM and native machines.
A recent pull request proposes adding support for native arrays~\cite{coq-pr-arrays}.

% Not sure if we want to include the following. -Adam
%% Some users of Coq are concerned by this increase in the TCB.
%% Out of the 34\,000 lines of code in the Coq kernel (about 5\,000 of them comments), about 4\,250 lines are the native compiler\footnote{In files matching \texttt{kernel/*native*}.}, about 6\,650 additional lines are dedicated to the VM\footnote{In files matching \texttt{kernel/*vm*}, \texttt{kernel/*byte*}, \texttt{kernel/cemitcodes*}, \texttt{kernel/clambda*}, \texttt{kernel/csymtable*}, \texttt{kernel/vconv*}, and \texttt{kernel/genOpcodeFiles*}.}, and an additional 3\,100 lines are devoted to conversion and reduction\footnote{In \texttt{kernel/reduction*}, \texttt{kernel/cClosure*}, \texttt{kernel/primred*}, and \texttt{kernel/conv\_oracle*}.}.

To summarize, there has been quite a lot of ``complexity creep'' in the Coq trusted base, to support efficient reduction, and yet realistic partial evaluation has \emph{still} been rather challenging.
Even the additional three reduction mechanisms outside Coq's kernel (\texttt{cbn}, \texttt{simpl}, \texttt{cbv}) are not at first glance sufficient for verified partial evaluation.

%% Definition foo \defeq
%%      (Pipeline.BoundsPipeline
%%         true None [64; 128]
%%         ltac:(let r \defeq Reify (fun f g => mulmod (weight 51 1) (2^255) [(1,19)] 5 f g) in
%%               exact r)
%%                (Some (repeat (@None _) 5), ((Some (repeat (@None _) 5), tt)))
%%                ZRange.type.base.option.None).
%%
%% Time Compute foo.
%%
%% Definition bar \defeq
%%      (Pipeline.BoundsPipeline
%%         true None [64; 128]
%%         ltac:(let r \defeq Reify (fun f g => mulmod (weight 51 2) (2^255) [(1,19)] 10 f g) in
%%               exact r)
%%                (Some (repeat (@None _) 10), ((Some (repeat (@None _) 10), tt)))
%%                ZRange.type.base.option.None).
%%
%% Time Compute bar.

\subsection{Our Solution} \label{sec:our-solution}

\citet{Aehlig} presented a very relevant solution to a related problem, using \emph{normalization by evaluation (NbE)}~\cite{NbE} to bootstrap reduction of open terms on top of full reduction, as built into a proof assistant.
However, it was simultaneously true that they expanded the proof-assistant trusted code base in ways specific to their technique, and that they did not report any experiments actually using the tool for partial evaluation (just traditional full reduction), potentially hiding performance-scaling challenges or other practical issues.
We have adapted their approach in a new Coq library embodying \textbf{the first partial-evaluation approach to satisfy the following criteria}.

\begin{itemize}
\item It integrates with a general-purpose, foundational proof assistant, \textbf{without growing the trusted base}.
\item For a wide variety of initial functional programs, it provides \textbf{fast} partial evaluation with reasonable memory use.
\item It allows reduction that \textbf{mixes} \emph{rules of the definitional equality} with \emph{equalities proven explicitly as theorems}.
\item It \textbf{preserves sharing} of common subterms.
\item It also allows \textbf{extraction of standalone partial evaluators}.
\end{itemize}

Our contributions include answers to a number of challenges that arise in scaling NbE-based partial evaluation in a proof assistant.
First, we rework the approach of \citet{Aehlig} to function \emph{without extending a proof assistant's trusted code base}, which, among other challenges, requires us to prove termination of reduction and encode pattern matching explicitly (leading us to adopt the performance-tuned approach of \citet{maranget2008compiling}).

Second, using partial evaluation to generate residual terms thousands of lines long raises \emph{new scaling challenges}:
\begin{itemize}
\item Output terms may contain so \emph{many nested variable binders} that we expect it to be performance-prohibitive to perform bookkeeping operations on first-order-encoded terms (e.g., with de Bruijn indices, as is done in \Rtac{} by \citet{rtac}).  For instance, while the reported performance experiments of \citet{Aehlig} generate only closed terms with no binders, Fiat Cryptography may generate a single routine (e.g., multiplication for curve P-384) with nearly a thousand nested binders.
%\todo{Consider being more specific here about expected cost of first-order representation, maybe using some experiment.}
%  \todo{Note from Jason: We haven't actually checked that the bookkeeping is performance-prohibitive.  I suspect there will be a cost, but that if when we made fiat-crypto there was an existing system that used de Bruijn indices and did this, (along with built-in let-lifting, etc), I suspect we might have been a bit more sad, but not prohibitively more sad.  I think we should be much more tentative with this bullet point.}
\item Naive representation of terms without proper \emph{sharing of common subterms} can lead to fatal term-size blow-up.  Fiat Cryptography's arithmetic routines rely on significant sharing of this kind.
\item Unconditional rewrite rules are in general insufficient, and we need \emph{rules with side conditions}.  For instance, in Fiat Cryptography, some rules for simplifying modular arithmetic depend on proofs that operations in subterms do not overflow.
\item However, it is also not reasonable to expect a general engine to discharge all side conditions on the spot.  We need integration with \emph{abstract interpretation} that can analyze whole programs to support reduction.
\end{itemize}

Briefly, our respective solutions to these problems are the \emph{parametric higher-order abstract syntax (PHOAS)}~\cite{PhoasICFP08} term encoding, a \emph{let-lifting} transformation threaded throughout reduction, extension of rewrite rules with executable Boolean side conditions, and a design pattern that uses decorator function calls to include analysis results in a program.

Finally, we carry out the \emph{first large-scale performance-scaling evaluation} of partial evaluation in a proof assistant, covering all elliptic curves from the published Fiat Cryptography experiments, along with microbenchmarks.

This paper proceeds through explanations of the trust stories behind our approach and earlier ones (\autoref{sec:trust}), the core structure of our engine (\autoref{sec:structure}), the additional scaling challenges we faced (\autoref{sec:scaling}), performance experiments (\autoref{sec:evaluation}), and related work (\autoref{sec:related}) and conclusions.
Our implementation is included as an anonymous supplement.

\section{Trust, Reduction, and Rewriting} \label{sec:trust}

Since much of the narrative behind our design process depends on tradeoffs between performance and trustworthiness, we start by reviewing the general situation in proof assistants.

Across a variety of proof assistants, simplification of functional programs is a workhorse operation.
Proof assistants like Coq that are based on type theory typically build in \emph{definitional equality} relations, identifying terms up to reductions like $\beta$-reduction and unfolding of named identifiers.
What looks like a single ``obvious'' step in an on-paper equational proof may require many of these reductions, so it is handy to have built-in support for checking a claimed reduction.
\autoref{subfig:change} diagrams how such steps work in a system like Coq, where the system implementation is divided between a trusted \emph{kernel}, for checking \emph{proof terms} in a minimal language, and additional untrusted support, like a \emph{tactic} engine evaluating a language of higher-level proof steps, in the process generating proof terms out of simpler building blocks.
It is standard to include a primitive proof step that validates any reduction compatible with the definitional equality, as the latter is decidable.
The figure shows a tactic that simplifies a goal using that facility.

In proof goals containing free variables, executing subterms can get stuck before reaching normal forms.
However, we can often achieve further simplification by using equational rules that we prove explicitly, rather than just relying on the rules built into the definitional equality and its decidable equivalence checker.
Coq's \texttt{autorewrite} tactic, as diagrammed in \autoref{subfig:autorewrite}, is a good example: it takes in a database of quantified equalities and applies them repeatedly to rewrite in a goal.
It is important that Coq's kernel does not trust the \texttt{autorewrite} tactic.
Instead, the tactic must output a proof term that, in some sense, is the moral equivalent of a line-by-line equational proof.
It can be challenging to keep these proof terms small enough, as naive rewrite-by-rewrite versions repeatedly copy large parts of proof goals, justifying a rewrite like $C[e_1] = C[e_2]$ for some context $C$ given a proof of $e_1 = e_2$, with the full value of $C$ replicated in the proof term for that single rewrite.
Overcoming these challenges while retaining decidability of proof checking is tricky, since we may use \texttt{autorewrite} with rule sets that do not always lead to terminating reduction.
Coq includes more experimental alternatives like \texttt{rewrite\_strat}, which use bottom-up construction of multi-rewrite proofs, with sharing of common contexts.
Still, as \autoref{sec:evaluation} will show, these methods that generate substantial proof terms are at significant performance disadvantages.
We also experimented with the corresponding tactics in the Lean proof assistant, with similarly dissapointing results (\autoref{sec:lean}).

\begin{figure} \centering
  \subfloat[Reduction via the definitional equality]{%
    \includegraphics[width=0.48\textwidth]{trust1}
    \label{subfig:change}} \hfill %\\
  \subfloat[Rewriting via proved rules]{%
    \includegraphics[width=0.48\textwidth]{trust2}
    \label{subfig:autorewrite}} \\
  \subfloat[Approach of \citet{Aehlig}]{%
    \includegraphics[width=0.48\textwidth]{trust3}
    \label{subfig:aehlig}} \hfill %\\
  \subfloat[Our approach]{%
    \includegraphics[width=0.48\textwidth]{trust4}
    \label{subfig:perfect}}

  \caption{Different approaches to reduction and rewriting}
  \label{fig:trust}
\end{figure}

Now we summarize how \citet{Aehlig} provide flexible and fast interleaving of standard $\lambda$-calculus reduction and use of proved equalities (the next section will go into more detail).
\autoref{subfig:aehlig} demonstrates a workflow based on \emph{a deep embedding of a core ML-like language}.
That is, within the logic of the proof assistant (Isabelle/HOL, in their case), a type of syntax trees for ML programs is defined, with an associated operational semantics.
The basic strategy is, for a particular set of rewrite rules and a particular term to simplify, to \emph{generate a (deeply embedded) ML program that, if it terminates, produces a syntax tree for the simplified term}.
Their tactic uses \emph{reification} to create ML versions of rule sets and terms.
They also wrote a reduction function in ML and proved it sound once and for all, against the ML operational semantics.
Combining that proof with proofs generated by reification, we conclude that an application of the reduction function to the reified rules and term is indeed an ML term that generates correct answers.
The tactic then ``throws the ML term over the wall,'' using a general code-generation framework for Isabelle/HOL~\cite{CodeGen}.
Trusted code compiles the ML code into the concrete syntax of a mainstream ML language, Standard ML in their case, and compiles it with an off-the-shelf compiler.
The output of that compiled program is then passed back over to the tactic, in terms of an axiomatic assertion that the ML semantics really yields that answer.

As \citet{Aehlig} argue, their use of external compilation and evaluation of ML code adds no real complexity on top of that required by the proof assistant -- after all, the proof assistant itself must be compiled and executed somehow.
However, the perceived increase of trusted code base is not spurious:
it is one thing to trust that the toolchain and execution environment used by the proof assistant and the partial evaluator are well-behaved,
and another to rely on two descriptions of ML (one deeply embedded in the proof assistant and another implied by the compiler) to agree on every detail of the semantics.
Furthermore, there still is new trusted code to translate from the deeply embedded ML subset into the concrete syntax of the full-scale ML language.
% Aehlig et al. defined an ML language quite specialized to their use case, building in exactly the one algebraic datatype they need to represent NbE results, and it is unsatisfying to add such a specialized translator to the trusted base.
% However, we could imagine using a very general deep embedding of the proof assistant in itself, as with Template-Coq~\cite{TemplateCoq}.
% In that case, the expansion of the trusted code base would be even more worrying, to include an interpretation function from a complete syntactic embedding of the language within itself.
% Any inconsistency between the mechanized operational semantics (which would not be small, for a full-scale practical logic) and the actual implementation could break logical soundness.
The vast majority of proof-assistant developments today rely on no such embeddings with associated mechanized semantics, so need we really add one to a proof-checking kernel to support efficient partial evaluation?

Our answer, diagrammed in \autoref{subfig:perfect}, shows a different way.
We still reify terms and rules into a deeply embedded language.
However, \emph{the reduction engine is implemented directly in the logic}, rather than as a deeply embedded syntax tree of an ML program.
As a result, the kernel's own reduction engine is prepared to execute our reduction engine for us -- using an operation that would be included in a type-theoretic proof assistant in any case, with no special support for a language deep embedding.
We also stage the process for performance reasons.
First, the \mintinline{coq}{Make} command creates a rewriter out of a list of rewrite rules, by specializing a generic partial-evaluation engine, which has a generic proof that applies to any set of proved rewrite rules.
We perform partial evaluation on the specialized partial evaluator, using Coq's normal reduction mechanisms, under the theory that we can afford to pay performance costs at this stage because we only need to create new rewriters relatively infrequently.
Then individual rewritings involve reifying terms, asking the kernel to execute the specialized evaluator on them, and simplifying an application of an interpretation function to the result (this last step must be done using Coq's normal reduction, and it is the bottleneck for outputs with enormous numbers of nested binders as discussed in section \ref{sec:micro}).

We would like to emphasize that, while we prototyped our implementation in Coq in particular, the trade-off space that we navigate seems fundamental, so that it should be the case both that our approach can be adapted to other proof assistants and that this case study may inform proof-assistant design.
The general game here is to stock the trusted proof-checking kernel with as few primitive rules as we can get away with, while still providing enough flexibility and performance.
Every proof assistant we are aware of has a small functional language at its core, and we argue that is quite natural to include a primitive for efficient full reduction of programs.
Our empirical result is that such a primitive can form the basis for bootstrapping other kinds of efficient reduction, perhaps suggesting that a future Coq version could fruitfully shrink its kernel by eliminating other built-in reduction strategies.

\subsection{Our Approach in Nine Steps} \label{sec:nine-steps}

Here is a bit more detail on the steps that go into applying our Coq plugin, many of which we expand on in the following sections.
In order to build a precomputed rewriter with the \mintinline{coq}{Make} command, the following actions are performed:
\begin{enumerate}
\item
  The given lemma statements are scraped for which named functions and types the rewriter package will support.
\item
  Inductive types enumerating all available primitive types and functions are emitted.
\item
  Tactics generate all of the necessary definitions and prove all of the necessary lemmas for dealing with this particular set of inductive codes.
  Definitions include operations like Boolean equality on type codes and lemmas like ``all representable primitive types have decidable equality.''
\item
  The statements of rewrite rules are reified and soundness and syntactic-well-formedness lemmas are proven about each of them.
  Each instance of the former involves wrapping the user-provided proof with the right adapter to apply to the reified version.
\item
  The definitions needed to perform reification and rewriting and the lemmas needed to prove correctness are assembled into a single package that can be passed by name to the rewriting tactic.
\end{enumerate}

When we want to rewrite with a rewriter package in a goal, the following steps are performed:
\begin{enumerate}
\item
  We rearrange the goal into a single logical formula: all free-variable quantification in the proof context is replaced by changing the equality goal into an equality between two functions (taking the free variables as inputs).
  %\todo{Pedantic Note From Jason: it's not actually an equality between two functions, it's an \mintinline{coq}{equiv} between two functions where \mintinline{coq}{equiv} is a custom relation we define indexed over type codes that is equality up to function extensionality}
\item
  We reify the side of the goal we want to simplify, using the inductive codes in the specified package.  That side of the goal is then replaced with a call to a denotation function on the reified version.
\item
  We use a theorem stating that rewriting preserves denotations of well-formed terms to replace the denotation subterm with the denotation of the rewriter applied to the same reified term.
  We use Coq's built-in full reduction (\texttt{vm\_compute}) to reduce the application of the rewriter to the reified term.
\item
  Finally, we run \texttt{cbv} (a standard call-by-value reducer) to simplify away the invocation of the denotation function on the concrete syntax tree from rewriting.
  %\todo{Is it worth mentioning that we'd instead use \texttt{vm\_compute} if it supported whitelists (and also said whitelists were recorded in the casts)?}
\end{enumerate}

\section{The Structure of a Rewriter} \label{sec:structure}

We now simultaneously review the approach of \citet{Aehlig} and introduce some notable differences in our own approach, noting similarities to the reflective rewriter of \citet{rtac} where applicable.

First, let us describe the language of terms we support rewriting in.
Note that, while we support rewriting in full-scale Coq proofs, where the metalanguage is dependently typed, the object language of our rewriter is nearly simply typed, with limited support for calling polymorphic functions.
However, we still support identifiers whose definitions use dependent types, since our reducer does not need to look into definitions.
\begin{align*}
  e ::={}& \phantom{\mid} \texttt{App }e_1\texttt{ }e_2 \mid \texttt{Let }v \defeq e_1\texttt{ In }e_2 %\\
  %&
  \mid \texttt{Abs }(\lambda v.\,e) \mid \texttt{Var }v \mid \texttt{Ident }i
\end{align*}
The \texttt{Ident} case is for identifiers, which are described by an enumeration specific to a use of our library.
For example, the identifiers might be codes for $+$, $\cdot$, and literal constants.
We write $\llbracket e \rrbracket$ for a standard denotational semantics.

\subsection{Pattern-Matching Compilation and Evaluation} \label{sec:pattern-matching-compilation-and-evaluation}

\citet{Aehlig} feed a specific set of user-provided rewrite rules to their engine by generating code for an ML function, which takes in deeply embedded term syntax (actually \emph{doubly} deeply embedded, within the syntax of the deeply embedded ML!) and uses ML pattern matching to decide which rule to apply at the top level.
Thus, they delegate efficient implementation of pattern matching to the underlying ML implementation.
As we instead build our rewriter in Coq's logic, we have no such option to defer to ML.
Indeed, Coq's logic only includes primitive pattern-matching constructs to match one constructor at a time.
%The frontend desugars nested pattern matching into trees of matches, but we do not have access to that desugaring within the language.
%It is arguably a good design decision to keep Coq working this way, since putting efficient nested pattern matching into the kernel would grow the trusted base substantially.

We could follow a naive strategy of repeatedly matching each subterm against a pattern for every rewrite rule, as in the rewriter of \citet{rtac}, but in that case we do a lot of duplicate work when rewrite rules use overlapping function symbols.
Instead, we adopted the approach of \citet{maranget2008compiling}, who describes compilation of pattern matches in OCaml to decision trees that eliminate needless repeated work (for example, decomposing an expression into $x + y + z$ only once even if two different rules match on that pattern).
We have not yet implemented any of the optimizations described therein for finding \emph{minimal} decision trees.

There are three steps to turn a set of rewrite rules into a functional program that takes in an expression and reduces according to the rules.
The first step is pattern-matching compilation: we must compile the lefthand sides of the rewrite rules to a decision tree that describes how and in what order to decompose the expression, as well as describing which rewrite rules to try at which steps of decomposition.
Because the decision tree is merely a decomposition hint, we require no proofs about it to ensure soundness of our rewriter.
The second step is decision-tree evaluation, during which we decompose the expression as per the decision tree, selecting which rewrite rules to attempt.
The only correctness lemma needed for this stage is that any result it returns is equivalent to picking some rewrite rule and rewriting with it.
The third and final step is to actually rewrite with the chosen rule.
Here the correctness condition is that we must not change the semantics of the expression.
Said another way, any rewrite-rule replacement expression must match the semantics of the rewrite-rule pattern.

While pattern matching begins with comparing one pattern against one expression, Maranget's approach works with intermediate goals that check multiple patterns against multiple expressions.
A decision tree describes how to match a vector (or list) of patterns against a vector of expressions.
It is built from these constructors:
\begin{itemize}
  \item \texttt{TryLeaf k onfailure}: Try the $k^\text{th}$ rewrite rule; if it fails, keep going with \texttt{onfailure}.
  \item \texttt{Failure}: Abort; nothing left to try.
  \item \texttt{Switch icases app\_case default}:
    With the first element of the vector, match on its kind; if it is an identifier matching something in \texttt{icases}, which is a list of pairs of identifiers and decision trees, remove the first element of the vector and run that decision tree; if it is an application and \texttt{app\_case} is not \texttt{None}, try the \texttt{app\_case} decision tree, replacing the first element of each vector with the two elements of the function and the argument it is applied to; otherwise, do not modify the vectors and use the \texttt{default} decision tree.
  \item \texttt{Swap i cont}: Swap the first element of the vector with the $i^\texttt{th}$ element (0-indexed) and keep going with \texttt{cont}.
\end{itemize}

Consider the encoding of two simple example rewrite rules, where we follow Coq's \Ltac{} language in prefacing pattern variables with question marks.
\begin{align*}
  ?n + 0 & \to n %\\
  &
  \texttt{fst}_{\mathbb{Z},\mathbb{Z}}(?x, ?y) & \to x
\end{align*}
We embed them in an AST type for patterns, which largely follows our ASTs for expressions.
\begin{verbatim}
0. App (App (Ident +) Wildcard) (Ident (Literal 0))
1. App (Ident fst) (App (App (Ident pair) Wildcard) Wildcard)
\end{verbatim}
The decision tree produced is \label{sec:compiled-pattern}
\[\resizebox{230px}{!}{\xymatrix@R-1pc{
  *++[o][F-]\txt{} \ar[d]_{\txt{App}} \\
  *++[o][F-]\txt{} \ar[r]^-{\txt{App}} \ar@/_1.5pc/[dr]_{\txt{\texttt{fst}}} & *++[o][F-]\txt{} \ar[r]^-{+} & *++[o][F-]\txt{Swap 0$\leftrightarrow$1} \ar[r] & *++[o][F-]\txt{} \ar[rr]^-{\txt{\texttt{Literal~0}}} && *++[o][F-]\txt{TryLeaf 0} \\
  & *++[o][F-]\txt{} \ar[r]_-{\txt{App}} & *++[o][F-]\txt{} \ar[r]_-{\txt{App}} & *++[o][F-]\txt{} \ar[rr]_-{\txt{\texttt{pair}}} && *++[o][F-]\txt{TryLeaf 1}
}}\]
\noindent where every non-swap node implicitly has a ``default'' case arrow to \texttt{Failure} and circles represent \texttt{Switch} nodes.

We implement, in Coq's logic, an evaluator for these trees against terms.
Note that we use Coq's normal partial evaluation to turn our general decision-tree evaluator into a specialized matcher to get reasonable efficiency.
Although this partial evaluation of our partial evaluator is subject to the same performance challenges we highlighted in the introduction, it only has to be done once for each set of rewrite rules, and we are targeting cases where the time of per-goal reduction dominates this time of meta-compilation.

For our running example of two rules, specializing gives us this match expression.
\begin{minted}[fontsize=\small]{coq}
match e with
| App f y => match f with
  | Ident fst => match y with
    | App (App (Ident pair) x) y => x | _ => e end
  | App (Ident +) x => match y with
    | Ident (Literal 0) => x | _ => e end | _ => e end | _ => e end.
\end{minted}

\subsection{Adding Higher-Order Features} \label{sec:thunk-eval-subst-term}

Fast rewriting at the top level of a term is the key ingredient for supporting customized algebraic simplification.
However, not only do we want to rewrite throughout the structure of a term, but we also want to integrate with simplification of higher-order terms, in a way where we can prove to Coq that our syntax-simplification function always terminates.
Normalization by evaluation (NbE)~\cite{NbE} is an elegant technique for adding the latter aspect, in a way where we avoid needing to implement our own $\lambda$-term reducer or prove it terminating.

To orient expectations: we would like to enable the following reduction
\begin{align*}
  (\lambda f\ x\ y.\, f\ x\ y)\ (+)\ z\ 0 & \leadsto z
\end{align*}
\noindent using the rewrite rule
\begin{align*}
  ?n + 0 & \to n
\end{align*}

\citet{Aehlig} also use NbE, and we begin by reviewing its most classic variant, for performing full $\beta$-reduction in a simply typed term in a guaranteed-terminating way.
The simply typed $\lambda$-calculus syntax we use is:
\begin{align*}
  t & ::= t \to t ~|~ b
  & e & ::= \lambda v.\, e ~|~ e~e ~|~ v ~|~ c
\end{align*}
\noindent with $v$ for variables, $c$ for constants, and $b$ for base types.

We can now define normalization by evaluation.
First, we choose a ``semantic'' representation for each syntactic type, which serves as the result type of an intermediate interpreter.
\begin{align*}
  \text{NbE}_t(t_1 \to t_2) & \defeq \text{NbE}_t(t_1) \to \text{NbE}_t(t_2) \\
  \text{NbE}_t(b) & \defeq \texttt{expr}(b)
\end{align*}
Function types are handled as in a simple denotational semantics, while base types receive the perhaps-counterintuitive treatment that the result of ``executing'' one is a syntactic expression of the same type.
We write $\texttt{expr}(b)$ for the metalanguage type of object-language syntax trees of type $b$, relying on a type family $\texttt{expr}$.

\begin{figure}%[b]
\begin{align*}
  \text{reify}_t & : \text{NbE}_t(t) \to \text{expr}(t) \\
  \text{reify}_{t_1 \to t_2}(f) & \defeq \lambda v.\,\text{reify}_{t_2}(f(\text{reflect}_{t_1}(v))) \\
  \text{reify}_{b}(f) & \defeq f \\ \noalign{\vskip7pt}
  \text{reflect}_t & : \text{expr}(t) \to \text{NbE}_t(t) \\
  \text{reflect}_{t_1\to t_2}(e) & \defeq \lambda x.\,\text{reflect}_{t_2}(e(\text{reify}_{t_1}(x)) \\
  \text{reflect}_{b}(e) & \defeq e \\ \noalign{\vskip7pt}
  \text{reduce} & : \text{expr}(t) \to \text{NbE}_t(t) \\
  \text{reduce}(\lambda v. \; e) & \defeq \lambda x. \; \text{reduce}([x/v]e) \\
  \text{reduce}(e_1~e_2) & \defeq \left(\text{reduce}(e_1)\right)(\text{reduce}(e_2)) \\
  \text{reduce}(x) & \defeq x \\
  \text{reduce}(c) & \defeq \text{reflect}(c) \\ \noalign{\vskip7pt}
  \text{NbE} & : \text{expr}(t) \to \text{expr}(t) \\
  \text{NbE}(e) & \defeq \text{reify}(\text{reduce}(e))
\end{align*}
\caption{\label{fig:nbe}Implementation of normalization by evaluation}
\end{figure}

Now the core of NbE, shown in \autoref{fig:nbe}, is a pair of dual functions reify and reflect, for converting back and forth between syntax and semantics of the object language, defined by primitive recursion on type syntax.
We split out analysis of term syntax in a separate function reduce, defined by primitive recursion on term syntax, when usually this functionality would be mixed in with reflect.
The reason for this choice will become clear when we extend NbE to handle our full problem domain.

% These definitions apply some of the usual corner-cutting that we see in on-paper descriptions of $\lambda$-term transformations.
We write $v$ for object-language variables and $x$ for metalanguage (Coq) variables, and we overload $\lambda$ notation using the metavariable kind to signal whether we are building a host $\lambda$ or a $\lambda$ syntax tree for the embedded language.
The crucial first clause for reduce replaces object-language variable $v$ with fresh metalanguage variable $x$, and then we are somehow tracking that all free variables in an argument to reduce must have been replaced with metalanguage variables by the time we reach them.
We reveal in \autoref{sec:PHOAS} the encoding decisions that make all the above legitimate, but first let us see how to integrate use of the rewriting operation from the previous section.
To fuse NbE with rewriting, we only modify the constant case of \texttt{reduce}.
First, we bind our specialized decision-tree engine under the name rewrite-head.
Recall that this function only tries to apply rewrite rules at the top level of its input.

In the constant case, we still reflect the constant, but underneath the binders introduced by full $\eta$-expansion, we perform one instance of rewriting.
In other words, we change this one function-definition clause:
\begin{align*}
  \text{reflect}_{b}(e) & \defeq \text{rewrite-head}(e)
\end{align*}

It is important to note that a constant of function type will be $\eta$-expanded only once for each syntactic occurrence in the starting term, though the expanded function is effectively a thunk, waiting to perform rewriting again each time it is called.
From first principles, it is not clear why such a strategy terminates on all possible input terms, though we work up to convincing Coq of that fact.

The details so far are essentially the same as in the approach of \citet{Aehlig}.
Recall that their rewriter was implemented in a deeply embedded ML, while ours is implemented in Coq's logic, which enforces termination of all functions.
Aehlig et al. did not prove termination, which indeed does not hold for their rewriter in general, which works with untyped terms, not to mention the possibility of rule-specific ML functions that diverge themselves.
In contrast, we need to convince Coq up-front that our interleaved $\lambda$-term normalization and algebraic simplification always terminate.
Additionally, we need to prove that our rewriter preserves denotations of terms, which can easily devolve into tedious binder bookkeeping, depending on encoding.

The next section introduces the techniques we use to avoid explicit termination proof or binder bookkeeping, in the context of a more general analysis of scaling challenges.

\section{Scaling Challenges} \label{sec:scaling}

\citet{Aehlig} only evaluated their implementation against closed programs.
What happens when we try to apply the approach to partial-evaluation problems that should generate thousands of lines of low-level code?

\subsection{Variable Environments Will Be Large} \label{sec:PHOAS}
We should think carefully about representation of ASTs, since many primitive operations on variables will run in the course of a single partial evaluation.
For instance, \citet{Aehlig} reported a significant performance improvement changing variable nodes from using strings to using de Bruijn indices~\cite{debruijn1972}.
However, de Bruijn indices and other first-order representations remain painful to work with.
We often need to fix up indices in a term being substituted in a new context.
Even looking up a variable in an environment tends to incur linear time overhead, thanks to traversal of a list.
Perhaps we can do better with some kind of balanced-tree data structure, but there is a fundamental performance gap versus the arrays that can be used in imperative implementations.
Unfortunately, it is difficult to integrate arrays soundly in a logic.
Also, even ignoring performance overheads, tedious binder bookkeeping complicates proofs.

Our strategy is to use a variable encoding that pushes all first-order bookkeeping off on Coq's kernel, which is itself performance-tuned with some crucial pieces of imperative code.
Parametric higher-order abstract syntax (PHOAS)~\cite{PhoasICFP08} is a dependently typed encoding of syntax where binders are managed by the enclosing type system.
It allows for relatively easy implementation and proof for NbE, so we adopted it for our framework.

Here is the actual inductive definition of term syntax for our object language, PHOAS-style.
The characteristic oddity is that the core syntax type \texttt{expr} is parameterized on a dependent type family for representing variables.
However, the final representation type \texttt{Expr} uses first-class polymorphism over choices of variable type, bootstrapping on the metalanguage's parametricity to ensure that a syntax tree is agnostic to variable type.
\begin{minted}[fontsize=\small]{coq}
  Inductive type := arrow (s d : type) | base (b : base_type).
  Infix "→" := arrow.
  Inductive expr (var : type -> Type) : type -> Type :=
  | Var {t} (v : var t) : expr var t
  | Abs {s d} (f : var s -> expr var d) : expr var (s → d)
  | App {s d} (f : expr var (s → d)) (x : expr var s) : expr var d
  | Const {t} (c : const t) : expr var t
  Definition Expr (t : type) : Type := forall var, expr var t.
\end{minted}

A good example of encoding adequacy is assigning a simple denotational semantics.
First, a simple recursive function assigns meanings to types.
\begin{minted}[fontsize=\small]{coq}
Fixpoint denoteT (t : type) : Type
  := match t with
     | arrow s d => denoteT s -> denoteT d
     | base b    => denote_base_type b
     end.
\end{minted}

Next we see the convenience of being able to \emph{use} an expression by choosing how it should represent variables.
Specifically, it is natural to choose \emph{the type-denotation function itself} as the variable representation.
Especially note how this choice makes rigorous the convention we followed in the prior section (e.g., in the suspicious function-abstraction clause of function reduce), where a recursive function enforces that values have always been substituted for variables early enough.
\begin{minted}[fontsize=\small]{coq}
Fixpoint denoteE {t} (e : expr denoteT t) : denoteT t
  := match e with
     | Var v     => v
     | Abs f     => λ x, denoteE (f x)
     | App f x   => (denoteE f) (denoteE x)
     | Ident c   => denoteI c
     end.
Definition DenoteE {t} (E : Expr t) : denoteT t
  := denoteE (E denoteT).
\end{minted}

It is now easy to follow the same script in making our rewriting-enabled NbE fully formal.
Note especially the first clause of \texttt{reduce}, where we avoid variable substitution precisely because we have chosen to represent variables with normalized semantic values.
The subtlety there is that base-type semantic values are themselves expression syntax trees, which depend on a nested choice of variable representation, which we retain as a parameter throughout these recursive functions.
The final definition $\lambda$-quantifies over that choice.
\begin{minted}[fontsize=\small]{coq}
Fixpoint nbeT var (t : type) : Type
  := match t with
     | arrow s d => nbeT var s -> nbeT var d
     | base b    => expr var b
     end.
Fixpoint reify {var t} : nbeT var t -> expr var t
  := match t with
     | arrow s d => λ f, Abs (λ x, reify (f (reflect (Var x))))
     | base b    => λ e, e
     end
with reflect {var t} : expr var t -> nbeT var t
  := match t with
     | arrow s d => λ e, λ x, reflect (App e (reify x))
     | base b    => rewrite_head
     end.
Fixpoint reduce {var t}
  (e : expr (nbeT var) t) : nbeT var t
  := match e with
     | Abs e     => λ x, reduce (e (Var x))
     | App e1 e2 => (reduce e1) (reduce e2)
     | Var x     => x
     | Ident c   => reflect (Ident c)
     end.
Definition Rewrite {t} (E : Expr t) : Expr t
  := λ var, reify (reduce (E (nbeT var t))).
\end{minted}

One subtlety hidden above in implicit arguments is in the final clause of \texttt{reduce}, where the two applications of the \texttt{Ident} constructor use different variable representations.
With all those details hashed out, we can prove a pleasingly simple correctness theorem, with a lemma for each main definition, with inductive structure mirroring recursive structure of the definition, also appealing to correctness of last section's pattern-compilation operations.
$$\forall t, E : \mintinline{coq}{Expr t}. \; \llbracket \mintinline{coq}{Rewrite}(E) \rrbracket = \llbracket E \rrbracket$$

Even before getting to the correctness theorem, we needed to convince Coq that the function terminates.
While for \citet{Aehlig}, a termination proof would have been a whole separate enterprise, it turns out that PHOAS and NbE line up so well that Coq accepts the above code with no additional termination proof.
As a result, the Coq kernel is ready to run our \mintinline{coq}{Rewrite} procedure during checking.

To understand how we now apply the soundness theorem in a tactic, it is important to note how the Coq kernel builds in reduction strategies.
These strategies have, to an extent, been tuned to work well to show equivalence between a simple denotational-semantics application and the semantic value it produces.
In contrast, it is rather difficult to code up one reduction strategy that works well for all partial-evaluation tasks.
Therefore, we should restrict ourselves to (1) running full reduction in the style of functional-language interpreters and (2) running normal reduction on ``known-good'' goals like correctness of evaluation of a denotational semantics on a concrete input.

Operationally, then, we apply our tactic in a goal containing a term $e$ that we want to partially evaluate.
In standard proof-by-reflection style, we \emph{reify} $e$ into some $E$ where $\llbracket E \rrbracket = e$, replacing $e$ accordingly, asking Coq's kernel to validate the equivalence via standard reduction.
Now we use the \mintinline{coq}{Rewrite} correctness theorem to replace $\llbracket E \rrbracket$ with $\llbracket \mintinline{coq}{Rewrite}(E) \rrbracket$.
Next we may ask the Coq kernel to simplify $\mintinline{coq}{Rewrite}(E)$ by \emph{full reduction via compilation to native code}, since we carefully designed $\mintinline{coq}{Rewrite}(E)$ and its dependencies to produce closed syntax trees, so that reduction will not get stuck pattern-matching on free variables.
Finally, where $E'$ is the result of that reduction, we simplify $\llbracket E' \rrbracket$ with standard reduction, producing a normal-looking Coq term.

%% The payoffs from fully satisfying Coq's type checker are:
%% \begin{enumerate}
%% \item We know that this procedure always terminates, and Coq's kernel is therefore willing to run the procedure for us implicitly during proof checking.  In a sense, we have bootstrapped this reduction strategy into the conversion rule of the type theory.
%% \item In that setting, all bookkeeping about variable binding and environments is handled by the kernel, whose implementation in OCaml allows certain efficient implementation strategies not available to us in the logic.
%% \end{enumerate}

\subsection{Subterm Sharing is Crucial} \label{sec:under-lets}

For some large-scale partial-evaluation problems, it is important to represent output programs with sharing of common subterms.
Redundantly inlining shared subterms can lead to exponential increase in space requirements.
Consider the Fiat Cryptography~\cite{FiatCryptoSP19} example of generating a 64-bit implementation of field arithmetic for the P-256 elliptic curve.
The library has been converted manually to continuation-passing style, allowing proper generation of \mintinline{coq}{let} binders, whose variables are often mentioned multiple times.
We ran their code generator (actually just a subset of its functionality, but optimized by us a bit further, as explained in \autoref{sec:macro}) on the P-256 example and found it took about 15 seconds to finish.
Then we modified reduction to inline \mintinline{coq}{let} binders instead of preserving them, at which point the reduction job terminated with an out-of-memory error, on a machine with 64 GB of RAM.
(The successful run uses under 2 GB.)

We see a tension here between performance and niceness of library implementation.
The Fiat Cryptography authors found it necessary to CPS-convert their code to coax Coq into adequate reduction performance.
Then all of their correctness theorems were complicated by reasoning about continuations.
It feels like a slippery slope on the path to implementing a domain-specific compiler, rather than taking advantage of the pleasing simplicity of partial evaluation on natural functional programs.
Our reduction engine takes shared-subterm preservation seriously while applying to libraries in direct style.

Our approach is \mintinline{coq}{let}-lifting: we lift \mintinline{coq}{let}s to top level, so that applications of functions to \mintinline{coq}{let}s are available for rewriting.
For example, we can perform the rewriting
\begin{align*}
  \texttt{map}\ (\lambda x.\, y+x)\ (\letin[{z:=e}{[0;1;2;z;z+1]}]) \\
  \leadsto \; \letin[{z:=e}{[y;y+1;y+2;y+z;y+(z+1)]}]
\end{align*}
using the rules
\begin{align*}
  \texttt{map}\ {?f}\ [] & \to []
  &
  \texttt{map}\ {?f}\ ({?x} :: {?xs}) & \to f\ x :: \texttt{map}\ f\ xs
  & {?n} + 0 & \to n %\\
\end{align*}

Our approach is to define a telescope-style type family called \mintinline{coq}{UnderLets}:
\begin{minted}[fontsize=\small]{coq}
Inductive UnderLets {var} (T : Type) :=
| Base (v : T)
| UnderLet {A}(e : @expr var A)(f : var A -> UnderLets T).
\end{minted}
A value of type \mintinline{coq}{UnderLets T} is a series of \texttt{let} binders (where each expression \mintinline{coq}{e} may mention earlier-bound variables) ending in a value of type \mintinline{coq}{T}.
It is easy to build various ``smart constructors'' working with this type, for instance to construct a function application by lifting the \texttt{let}s of both function and argument to a common top level.

Such constructors are used to implement an NbE strategy that outputs \mintinline{coq}{UnderLets} telescopes.
Recall that the NbE type interpretation mapped base types to expression syntax trees.
We now parameterize that type interpretation by a Boolean declaring whether we want to introduce telescopes.

\begin{minted}[fontsize=\small]{coq}
Fixpoint nbeT' {var} (with_lets : bool) (t : type)
  := match t with
     | base t => if with_lets then @UnderLets var (@expr var t) else @expr var t
     | arrow s d => nbeT' false s -> nbeT' true d
     end.
Definition nbeT := nbeT' false.
Definition nbeT_with_lets := nbeT' true.
\end{minted}

%%  - Here are some examples:
%%    - `value Z := UnderLets (expr Z)`
%%    - `value (Z -> Z) := expr Z -> UnderLets (expr Z)`
%%    - `value (Z -> Z -> Z) := expr Z -> expr Z -> UnderLets (expr Z)`
%%    - `value ((Z -> Z) -> Z) := (expr Z -> UnderLets (expr Z)) -> UnderLets (expr Z)`

There are cases where naive preservation of \texttt{let} binders blocks later rewrites from triggering and leads to suboptimal performance, so we include some heuristics.
For instance, when the expression being bound is a constant, we always inline.
When the expression being bound is a series of list ``cons'' operations, we introduce a name for each individual list element, since such a list might be traversed multiple times in different ways.

\subsection{Rules Need Side Conditions} \label{sec:side-conditions}

Many useful algebraic simplifications require side conditions.
One simple case is supporting \emph{nonlinear} patterns, where a pattern variable appears multiple times.
We can encode nonlinearity on top of linear patterns via side conditions.
\begin{align*}
  {?n_1} + {?m} - {?n_2} & \to m\text{\ \ if\ \ }n_1 = n_2
\end{align*}

The trouble is how to support predictable solving of side conditions during partial evaluation, where we may be rewriting in open terms.
We decided to sidestep this problem by allowing side conditions only as executable Boolean functions, to be applied only to variables that are confirmed as \emph{compile-time constants}, unlike \citet{rtac} who support general unification variables.
We added a variant of pattern variable that only matches constants.
Semantically, this variable style has no additional meaning, and in fact we implement it as a special identity function that should be called in the right places within Coq lemma statements.
Rather, use of this identity function triggers the right behavior in our tactic code that reifies lemma statements.
We introduce a notation where a prefixed apostrophe signals a call to the ``constants only'' function.

Our reification inspects the hypotheses of lemma statements, using type classes to find decidable realizations of the predicates that are used, synthesizing one Boolean expression of our deeply embedded term language, standing for a decision procedure for the hypotheses.
The \mintinline{coq}{Make} command fails if any such expression contains pattern variables not marked as constants.
Therefore, matching of rules can safely run side conditions, knowing that Coq's full-reduction engine can determine their truth efficiently.

\subsection{Side Conditions Need Abstract Interpretation} \label{sec:abs-int}

With our limitation that side conditions are decided by executable Boolean procedures, we cannot yet handle directly some of the rewrites needed for realistic partial evaluation.
For instance, Fiat Cryptography reduces high-level functional to low-level code that only uses integer types available on the target hardware.
The starting library code works with arbitrary-precision integers, while the generated low-level code should be careful to avoid unintended integer overflow.
As a result, the setup may be too naive for our running example rule ${?n} + 0 \to n$.
When we get to reducing fixed-precision-integer terms, we must be legalistic:
\begin{align*}
  \texttt{add\_with\_carry}_{64}({?n}, 0) & \to (0, n)\text{\ \ if\ \ }0 \le n < 2^{64}
\end{align*}

We developed a design pattern to handle this kind of rule.

First, we introduce a family of functions $\texttt{clip}_{l,u}$, each of which forces its integer argument to respect lower bound $l$ and upper bound $u$.
Partial evaluation is proved with respect to unknown realizations of these functions, only requiring that $\texttt{clip}_{l, u}(n) = n$ when $l \leq n < u$.
Now, before we begin partial evaluation, we can run a verified abstract interpreter to find conservative bounds for each program variable.
When bounds $l$ and $u$ are found for variable $x$, it is sound to replace $x$ with $\texttt{clip}_{l,u}(x)$.
Therefore, at the end of this phase, we assume all variable occurrences have been rewritten in this manner to record their proved bounds.

Second, we proceed with our example rule refactored:
\begin{align*}
  \texttt{add\_with\_carry}_{64}(\texttt{clip}_{'{?l},'{?u}}({?n}), 0) & \to (0, \texttt{clip}_{l,u}(n)) %\\
  %&
  \text{\ \ if\ \ }u < 2^{64}
\end{align*}
If the abstract interpreter did its job, then all lower and upper bounds are constants, and we can execute side conditions straightforwardly during pattern matching.

\subsection{Limitations and Preprocessing} \label{sec:implementation-and-usage}

%\todo{It's not clear to Jason what in this section belongs here, and what isn't interesting (yet?) to readers}
%\todo{I (Jason) have no idea what the story for this section should be}
%\todo{Insofar as text is actually kept in this section, make sure any duplication between here and the out-of-paper appendix is deliberate}
%\todo{If this section is moved or removed, be sure to restructure \autoref{sec:code-from-implementation-and-usage} accordingly}

We now note some details of the rewriting framework that were previously glossed over, which are useful for using the code or implementing something similar, but which do not add fundamental capabilities to the approach.
Although the rewriting framework does not support dependently typed constants, we can automatically preprocess uses of eliminators like \mintinline{coq}{nat_rect} and \mintinline{coq}{list_rect} into non-dependent versions.
The tactic that does this preprocessing is extensible via \Ltac{}'s reassignment feature.
Since pattern-matching compilation mixed with NbE requires knowing how many arguments a constant can be applied to, internally we must use a version of the recursion principle whose type arguments do not contain arrows; current preprocessing can handle recursion principles with either no arrows or one arrow in the motive.

Recall from \autoref{sec:explain-eval-rect} that \mintinline{coq}{eval_rect} is a definition provided by our framework for eagerly evaluating recursion associated with certain types.
It functions by triggering typeclass resolution for the lemmas reducing the recursion principle associated to the given type.
We provide instances for \texttt{nat}, \texttt{prod}, \texttt{list}, \texttt{option}, and \texttt{bool}.
Users may add more instances if they desire.

Recall again from \autoref{sec:explain-ident.eagerly} that we use \mintinline{coq}{ident.eagerly} to ask the reducer to simplify a case of primitive recursion by complete traversal of the designated argument's constructor tree.
Our current version only allows a limited, hard-coded set of eliminators with \mintinline{coq}{ident.eagerly} (\texttt{nat\_rect} on return types with either zero or one arrows, \texttt{list\_rect} on return types with either zero or one arrows, and \texttt{List.nth\_default}), but nothing in principle prevents automatic generation of the necessary code.

Note that \mintinline{coq}{Let_In} is the constant we use for writing \letin{} expressions that do not reduce under $\zeta$ (Coq's reduction rule for \mintinline{coq}{let}-inlining).
Throughout most of this paper, anywhere that \letin{} appears, we have actually used \mintinline{coq}{Let_In} in the code.
It would alternatively be possible to extend the reification preprocessor to automatically convert \letin{} to \mintinline{coq}{Let_In}, but this strategy may cause problems when converting the interpretation of the reified term with the pre-reified term, as Coq's conversion does not allow fine-tuning of when to inline or unfold \mintinline{coq}{let}s.

\section{Evaluation}\label{sec:evaluation}

% TODO: Presumably ``attached to this submission as an anonymized supplement'' will change for the final submission
Our implementation, attached to this submission as an anonymized supplement with a roadmap in \autoref{sec:CodeSupplement-more}, includes a mix of Coq code for the proved core of rewriting, tactic code for setting up proper use of that core, and OCaml plugin code for the manipulations beyond the current capabilities of the tactic language.
We report here on experiments to isolate performance benefits for rewriting under binders and reducing higher-order structure.

\subsection{Microbenchmarks} \label{sec:micro}

\def\NoBindersSubfloatNval{3}%
\def\NoBindersSubfloatXRow{\thisrowno{0}*(2^(\nval+1)-1)}%

We start with microbenchmarks focusing attention on particular aspects of reduction and rewriting, with \autoref{sec:additionalMicro} going into more detail.

\subsubsection{Rewriting Without Binders} \label{sec:micro:Plus0Tree}

Consider the code defined by the expression $\text{tree}_{n,m}(v)$ in \autoref{fig:micro:Plus0Tree:code}.
%\todo{make sure the figure spacing is okay}
%RSolve[{t[0]==m, t[n]==2*t[n-1] + m},t[n],n]
%t[n]=m(2^{n+1}-1)
We want to remove all of the${}+0$s.
There are $\Theta(m \cdot 2^n)$ such rewriting locations.
We can start from this expression directly, in which case reification alone takes as much time as Coq's \texttt{rewrite}.
\begin{wrapfigure}[7]{r}{0pt}
{\small $\begin{aligned}
\text{iter}_m(v) & = v + \underbrace{0 + 0 + \cdots + 0}_m \\
\text{tree}_{0,m}(v) &= \text{iter}_m(v + v) \\
\text{tree}_{n+1,m}(v) &= \text{iter}_m(\text{tree}_{n,m}(v) + \text{tree}_{n,m}(v))
\end{aligned}$}%
\caption{\label{fig:micro:Plus0Tree:code}Expressions computing initial code}
\end{wrapfigure}
As the reification method was not especially optimized, and there exist fast reification methods~\cite{ReificationITP18}, we instead start from a call to a recursive function that generates such an expression.

\begin{figure*}[b]
  \newsavebox{\NestedBindersSubfloat}%
  \sbox{\NestedBindersSubfloat}{%
    \adjustbox{valign=t}{\resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
      \einput{perf-UnderLetsPlus0-cbv;rewrite-strat(bottomup).txt}
      \einput{perf-UnderLetsPlus0-cbv;rewrite-strat(topdown).txt}
      \einput{perf-UnderLetsPlus0-cbv;setoid-rewrite.txt}
      \einput{perf-UnderLetsPlus0-Rewrite-for.txt}
      \einput{perf-UnderLetsPlus0-rewriting.txt}
    }
      \pgfplotsset{every axis legend/.append style={
          at={(0.5,-0.2)},
          anchor=north}}
      \begin{axis}[xlabel=\# of let binders,ylabel=time (s),ymax=65]
        \addplot[mark=o,color=red] table{perf-UnderLetsPlus0-cbv;rewrite-strat(bottomup).txt};
        \addplot[mark=triangle,color=red] table{perf-UnderLetsPlus0-cbv;rewrite-strat(topdown).txt};
        \addplot[mark=square,color=red] table{perf-UnderLetsPlus0-cbv;setoid-rewrite.txt};
        \addplot[mark=+,color=blue] table{perf-UnderLetsPlus0-Rewrite-for.txt};
        \addplot[mark=x,color=ForestGreen] table{perf-UnderLetsPlus0-rewriting.txt};
        \legend{rewrite\_strat bottomup,rewrite\_strat topdown,setoid\_rewrite,{Our approach including reification, cbv, etc.},Our approach (only rewriting)}
      \end{axis}
    \end{tikzpicture}}}}%
  \newsavebox{\BindersAndRecursiveFunctionsSubfloat}%
  \sbox{\BindersAndRecursiveFunctionsSubfloat}{%
    \adjustbox{valign=t}{\resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
        \einput{perf-LiftLetsMap-rewrite-strat(bottomup,bottomup).txt}
        \einput{perf-LiftLetsMap-rewrite-strat(topdown,bottomup).txt}
        \einput{perf-LiftLetsMap-setoid-rewrite.txt}
        \einput{perf-LiftLetsMap-Rewrite-for.txt}
        \einput{perf-LiftLetsMap-rewriting.txt}
        \einput{perf-LiftLetsMap-cps+vm-compute.txt}
    }
      \pgfplotsset{every axis legend/.append style={
          at={(0.5,-0.2)},
          anchor=north}}
      \begin{axis}[xlabel=$n\cdot m$,ylabel=time (s),scaled x ticks=false,ymax=27]
        \addplot[mark=o,color=red] table{perf-LiftLetsMap-rewrite-strat(bottomup,bottomup).txt};
        \addplot[mark=triangle,color=red] table{perf-LiftLetsMap-rewrite-strat(topdown,bottomup).txt};
        \addplot[mark=square,color=red] table{perf-LiftLetsMap-setoid-rewrite.txt};
        \addplot[mark=+,color=blue] table{perf-LiftLetsMap-Rewrite-for.txt};
        \addplot[mark=x,color=ForestGreen] table{perf-LiftLetsMap-rewriting.txt};
        \addplot[mark=*,color=purple] table{perf-LiftLetsMap-cps+vm-compute.txt};
        \legend{rewrite\_strat bottomup,rewrite\_strat topdown,repeat setoid\_rewrite,{Our approach including reification, cbv, etc.},Our approach (only rewriting),cps+vm\_compute}
      \end{axis}
    \end{tikzpicture}}}}%
  \newsavebox{\NoBindersSubfloat}%
  \sbox{\NoBindersSubfloat}{%
    \edef\nval{\NoBindersSubfloatNval}%
    \adjustbox{valign=t}{\resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
        \einput{perf-Plus0Tree-only-n-\nval-cbv;rewrite-strat(bottomup).txt}
        \einput{perf-Plus0Tree-only-n-\nval-cbv;setoid-rewrite.txt}
        \einput{perf-Plus0Tree-only-n-\nval-cbv;rewrite-strat(topdown).txt}
        \einput{perf-Plus0Tree-only-n-\nval-cbv;rewrite!.txt}
        \einput{perf-Plus0Tree-only-n-\nval-Rewrite-for.txt}
        \einput{perf-Plus0Tree-only-n-\nval-rewriting.txt}
        \nval
    }
      \pgfplotsset{every axis legend/.append style={
          at={(0.5,-0.2)},
          anchor=north}}
      % since n = 1, we have 2^n=twice as many rewrite locations as the value on the x axis, so we need to double things
      %,xticklabel={\pgfkeys{/pgf/fpu}\pgfmathparse{2^\nval*\tick}$\mathsf{\pgfmathprintnumber{\pgfmathresult}}$}
      \begin{axis}[xlabel=\# of rewrite locations,scaled x ticks=false,ylabel=time (s),ymax=7]% ymax=10]
        \addplot[mark=square,color=red] table[x expr={\NoBindersSubfloatXRow}]{perf-Plus0Tree-only-n-\nval-cbv;rewrite-strat(bottomup).txt};
        \addplot[mark=*,color=red] table[x expr={\NoBindersSubfloatXRow}]{perf-Plus0Tree-only-n-\nval-cbv;setoid-rewrite.txt};
        \addplot[mark=triangle,color=red] table[x expr={\NoBindersSubfloatXRow}]{perf-Plus0Tree-only-n-\nval-cbv;rewrite-strat(topdown).txt};
        \addplot[mark=o,color=red] table[x expr={\NoBindersSubfloatXRow}]{perf-Plus0Tree-only-n-\nval-cbv;rewrite!.txt};
        \addplot[mark=+,color=blue] table[x expr={\NoBindersSubfloatXRow}]{perf-Plus0Tree-only-n-\nval-Rewrite-for.txt};
        \addplot[mark=x,color=ForestGreen] table[x expr={\NoBindersSubfloatXRow}]{perf-Plus0Tree-only-n-\nval-rewriting.txt};
        \legend{rewrite\_strat bottomup,setoid\_rewrite,rewrite\_strat topdown,rewrite!,{Our approach including reification, cbv, etc.},Our approach (only rewriting)}
      \end{axis}
    \end{tikzpicture}}}}%
  \newsavebox{\FiatCryptoSubfloat}%
  \sbox{\FiatCryptoSubfloat}{%
    \adjustbox{valign=t}{\resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
      \einput{perf-new-vm-over-old-vm.txt.md5}
      \einput{perf-old-vm-over-old-vm.txt.md5}
      \einput{perf-new-extraction-over-old-vm.txt.md5}
    }
      \pgfplotsset{every axis legend/.append style={
          at={(0.5,-0.2)},
          anchor=north}}
      \begin{axis}[xlabel=prime,ylabel=slowdown factor over old approach,xmode=log, ymode=log,log basis x={2}]
        \addplot[mark=+] table{perf-new-vm-over-old-vm.txt};
        \addplot[mark=o] table{perf-old-vm-over-old-vm.txt};
        \addplot[mark=*] table{perf-new-extraction-over-old-vm.txt};
        \legend{Our approach w/ Coq's VM,Old approach (handwritten-CPS+VM),Our approach w/ extracted OCaml}
      \end{axis}
    \end{tikzpicture}}}}%
  \newcommand{\vphantomSubfloatOne}{\vphantom{\usebox{\NestedBindersSubfloat}}\vphantom{\usebox{\NoBindersSubfloat}}}%
  \newcommand{\vphantomSubfloatTwo}{\vphantom{\usebox{\BindersAndRecursiveFunctionsSubfloat}}\vphantom{\usebox{\FiatCryptoSubfloat}}}%
  \subfloat[No binders]{\usebox{\NoBindersSubfloat}\vphantomSubfloatOne\label{fig:timing-Plus0Tree}}%
  \quad
  \subfloat[Nested binders]{\usebox{\NestedBindersSubfloat}\vphantomSubfloatOne\label{fig:timing-UnderLetsPlus0}}%
  \quad
  \subfloat[Binders and recursive functions]{\usebox{\BindersAndRecursiveFunctionsSubfloat}\vphantomSubfloatTwo\label{fig:timing-LiftLetsMap}}%
  \quad
  \subfloat[Fiat Cryptography]{\usebox{\FiatCryptoSubfloat}\vphantomSubfloatTwo\label{fig:scaling}}

  \caption{Timing of different partial-evaluation implementations} \label{fig:multi-timing}
\end{figure*}

%RSolve[{t[0]==m, t[n]==2*t[n-1] + m},t[n],n]
%t[n]=m(2^{n+1}-1)
\autoref{fig:timing-Plus0Tree} shows the results for $n = \NoBindersSubfloatNval$ as we scale $m$.
The comparison points are Coq's \texttt{rewrite!}, \texttt{setoid\_rewrite}, and \texttt{rewrite\_strat}.
The first two perform one rewrite at a time, taking minimal advantage of commonalities across them and thus generating quite large, redundant proof terms.
The third makes top-down or bottom-up passes with combined generation of proof terms.
For our own approach, we list both the total time and the time taken for core execution of a verified rewrite engine, without counting reification (converting goals to ASTs) or its inverse (interpreting results back to normal-looking goals).

The comparison here is very favorable for our approach so long as $m > 2$.
The competing tactics spike upward toward timeouts at just around a thousand rewrite locations, while our engine is still under two seconds for examples with tens of thousands of rewrite locations.
When $m < 2$, Coq's \texttt{rewrite!} tactic does a little bit better than our engine, corresponding roughly to the overhead incurred by our term representation (which, for example, stores the types at every application node) when most of the term is in fact unchanged by rewriting.
See \autoref{sec:additionalPlots:Plus0Tree}\footnote{Like several forward references in this section, this one goes to an appendix included within the main submission page limit, to avoid interrupting the flow in presenting the most important results.} for more detailed plots.

\subsubsection{Rewriting Under Binders} \label{sec:micro:UnderLetsPlus0}

\begin{wrapfigure}[7]{r}{0pt}
{\small $\begin{aligned}
  & \letin[{v_1 := v_0 + v_0 + 0}{}] \\ \noalign{\vskip-7pt}
  & \vdots \\ \noalign{\vskip-3pt}
  & \letin[{v_n := v_{n-1} + v_{n-1} + 0}{}] \\
  & v_n + v_n + 0
\end{aligned}$}%
%  & \letin[{v_2 := v_1 + v_1 + 0}{}] \\
\caption{\label{fig:micro:UnderLetsPlus0:code}Initial code}
\end{wrapfigure}

Consider now the code in \autoref{fig:micro:UnderLetsPlus0:code}, which is a version of the code above where redundant expressions are shared via \mintinline{coq}{let} bindings.

\autoref{fig:timing-UnderLetsPlus0} shows the results.
The comparison here is again very favorable for our approach.
The competing tactics spike upward toward timeouts at just a few hundred generated binders, while our engine is only taking about 10 seconds for examples with 5,000 nested binders.

\subsubsection{Performance Bottlenecks of Proof-Producing Rewriting} \label{sec:micro:setoid-rewrite-bottlenecks-lite}

Although we have made our comparison against the built-in tactics \mintinline{coq}{setoid_rewrite} and \mintinline{coq}{rewrite_strat}, by analyzing the performance in detail, we can argue that these performance bottlenecks are likely to hold for any proof assistant designed like Coq.
Detailed debugging reveals five performance bottlenecks in the existing rewriting tactics, which we discuss in \autoref{sec:setoid-rewrite-bottlenecks}.\footnote{Also included within the submission page limit, though of interest mostly to proof-assistant-implementation experts.}

\subsubsection{Binders and Recursive Functions} \label{sec:micro:LiftLetsMap}

\begin{wrapfigure}[11]{r}{0pt}
{\small %\allowdisplaybreaks
$\begin{aligned}
  \text{map\_dbl}(\ell) & \defeq \begin{cases} [] & \text{if }\ell = [] \\
      \letin[{y := h + h}{}] & \text{if }\ell = h::t \\
      y :: \text{map\_dbl}(t) &
      \end{cases} \\
  \text{make}(n, m, v) & \defeq \begin{cases} [\underbrace{v, \ldots, v}_n] & \text{if }m = 0 \\
      \text{map\_dbl}(\text{make}(n, m-1, v)) & \text{if }m > 0
      \end{cases} \\
  \text{example}_{n, m} & \defeq \forall v,\ \text{make}(n, m, v) = []
\end{aligned}$}%
\caption{\label{fig:micro:LiftLetsMap:code}Initial code for binders and recursive functions}
\end{wrapfigure}

The next experiment uses the code in \autoref{fig:micro:LiftLetsMap:code}.
%Note that the ${}=[]$ at the end is just a dummy placeholder to allow us to check reduction and rewriting of this term.
Note that the \letin{} binding blocks further reduction of map\_dbl when we iterate it $m$ times in make, and so we need to take care to preserve sharing when reducing here.

\autoref{fig:timing-LiftLetsMap} compares performance between our approach, \texttt{repeat setoid\_rewrite}, and two variants of \texttt{rewrite\_strat}.
Additionally, we consider another option, which was adopted by Fiat Cryptography at a larger scale: rewrite our functions to improve reduction behavior.
Specifically, both functions are rewritten in continuation-passing style, which makes them harder to read and reason about but allows standard VM-based reduction to achieve good performance.
The figure shows that \texttt{rewrite\_strat} variants are essentially unusable for this example, with \texttt{setoid\_rewrite} performing only marginally better, while our approach applied to the original, more readable definitions loses ground steadily to VM-based reduction on CPSed code.
On the largest terms ($n \cdot m > 20,000$), the gap is 6s vs.\ 0.1s of compilation time, which should often be acceptable in return for simplified coding and proofs, plus the ability to mix proved rewrite rules with built-in reductions.
Note that about 99\% of the difference between the full time of our method and just the rewriting is spent in the final \texttt{cbv} at the end, used to denote our output term from refied syntax.
We blame this performance on the unfortunate fact that reduction in Coq is quadratic in the number of nested binders present; see Coq bug \coqbug{11151}.
%traded for simplified coding and proofs, plus the ability to mix proved rewrite rules with built-in reductions.
See \autoref{sec:LiftLetsMap-more} for more on this microbenchmark.

\subsubsection{Full Reduction} \label{sec:micro:Eratosthenes}

\begin{wrapfigure}[18]{r}{0pt}
  \resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
      \einput{perf-SieveOfEratosthenes-simpl.txt}
      \einput{perf-SieveOfEratosthenes-cbn.txt}
      \einput{perf-SieveOfEratosthenes-lazy.txt}
      \einput{perf-SieveOfEratosthenes-Rewrite-for.txt}
      \einput{perf-SieveOfEratosthenes-rewriting.txt}
      \einput{perf-SieveOfEratosthenes-cbv.txt}
      \einput{perf-SieveOfEratosthenes-vm-compute.txt}
      \einput{perf-SieveOfEratosthenes-native(2)(real).txt}
  }
    \pgfplotsset{every axis legend/.append style={
        at={(0.5,-0.2)},
        anchor=north}}
    \begin{axis}[xlabel=upper bound,ylabel=time (s),ymax=65]
      \addplot[mark=square,color=red] table{perf-SieveOfEratosthenes-simpl.txt};
      \addplot[mark=triangle,color=red] table{perf-SieveOfEratosthenes-cbn.txt};
      \addplot[mark=square*,color=red] table{perf-SieveOfEratosthenes-lazy.txt};
      \addplot[mark=+,color=blue] table{perf-SieveOfEratosthenes-Rewrite-for.txt};
      \addplot[mark=x,color=ForestGreen] table{perf-SieveOfEratosthenes-rewriting.txt};
      \addplot[mark=triangle*,color=orange] table{perf-SieveOfEratosthenes-cbv.txt};
      \addplot[mark=*,color=purple] table{perf-SieveOfEratosthenes-vm-compute.txt};
      \addplot[mark=-,color=red] table{perf-SieveOfEratosthenes-native(2)(real).txt};
      \legend{simpl,cbn,lazy,{Our approach including reification, cbv, etc.},Our approach (only rewriting),cbv,vm\_compute,native\_compute}
    \end{axis}
  \end{tikzpicture}}
  \caption{\label{fig:timing-SieveOfEratosthenes}Full evaluation, Sieve of Eratosthenes}
\end{wrapfigure}

The final experiment involves full reduction in computing the Sieve of Eratosthenes, taking inspiration on benchmark choice from \citet{Aehlig}.
We find in \autoref{fig:timing-SieveOfEratosthenes} that we are slower than \texttt{vm\_compute}, \mintinline{coq}{native_compute}, and \texttt{cbv}, but faster than \mintinline{coq}{lazy}, and of course much faster than \texttt{simpl} and \texttt{cbn}, which are quite slow.

\subsection{Macrobenchmark: Fiat Cryptography} \label{sec:macro}

Finally, we consider an experiment (described in more detail in \autoref{sec:additionalMacro}) replicating the generation of performance-competitive finite-field-arithmetic code for all popular elliptic curves by \citet{FiatCryptoSP19}.
In all cases, we generate essentially the same code as they did, so we only measure performance of the code-generation process.
We stage partial evaluation with three different reduction engines (i.e., three \mintinline{coq}{Make} invocations), respectively applying 85, 56, and 44 rewrite rules (with only 2 rules shared across engines), taking total time of about 5 minutes to generate all three engines.
These engines support 95 distinct function symbols.

\autoref{fig:scaling} graphs running time of three different partial-evaluation methods for Fiat Cryptography, as the prime modulus of arithmetic scales up.
Times are normalized to the performance of the original method, which relied entirely on standard Coq reduction.
Actually, in the course of running this experiment, we found a way to improve the old approach for a fairer comparison.
It had relied on Coq's configurable \texttt{cbv} tactic to perform reduction with selected rules of the definitional equality, which the Fiat Cryptography developers had applied to blacklist identifiers that should be left for compile-time execution.
By instead hiding those identifiers behind opaque module-signature ascription, we were able to run Coq's more-optimized virtual-machine-based reducer.

As the figure shows, our approach running partial evaluation inside Coq's kernel begins with about a 10$\times$ performance disadvantage vs.\ the original method.
With log scale on both axes, we see that this disadvantage narrows to become nearly negligible for the largest primes, of around 500 bits.
(We used the same set of prime moduli as in the experiments run by \citet{FiatCryptoSP19}, which were chosen based on searching the archives of an elliptic-curves mailing list for all prime numbers.)
It makes sense that execution inside Coq leaves our new approach at a disadvantage, as we are essentially running an interpreter (our normalizer) within an interpreter (Coq's kernel), while the old approach ran just the latter directly.
Also recall that the old approach required rewriting Fiat Cryptography's library of arithmetic functions in continuation-passing style, enduring this complexity in library correctness proofs, while our new approach applies to a direct-style library.
Finally, the old approach included a custom reflection-based arithmetic simplifier for term syntax, run after traditional reduction, whereas now we are able to apply a generic engine that combines both, without requiring more than proving traditional rewrite rules.

The figure also confirms clear performance advantage of running reduction in code extracted to OCaml, which is possible because our plugin produces verified code in Coq's functional language.
By the time we reach middle-of-the-pack prime size around 300 bits, the extracted version is running about 10$\times$ as quickly as the baseline.

\section{Related Work}\label{sec:related}

We have already discussed the work of \citet{Aehlig}, which introduced the basic structure that our engine shares, but which required a substantially larger trusted code base, did not tackle certain challenges in scaling to large partial-evaluation problems, and did not report any performance experiments in partial evaluation.

We have also mentioned \Rtac{}~\cite{rtac}, which implements an experimental reflective version of \texttt{rewrite\_strat} supporting arbitrary setoid relations, unification variables, and arbitrary semi-decidable side conditions solvable by other reflective tactics, using de Bruijn indexing to manage binders.
We were unfortunately unable to get the rewriter to work with Coq 8.10 and were also not able to determine from the paper how to repurpose the rewriter to handle our benchmarks.

Our implementation builds on fast full reduction in Coq's kernel, via a virtual machine~\cite{vmcompute} or compilation to native code~\cite{nativecompute}.
Especially the latter is similar in adopting an NbE style for full reduction, simplifying even under $\lambda$s, on top of a more traditional implementation of OCaml that never executes preemptively under $\lambda$s.
Neither approach unifies support for rewriting with proved rules, and partial evaluation only applies in very limited cases, where functions that should not be evaluated at compile time must have properly opaque definitions that the evaluator will not consult.
Neither implementation involved a machine-checked proof suitable to bootstrap on top of reduction support in a kernel providing simpler reduction.

% A more limited form of code generation for cryptography was already widespread through libraries like OpenSSL, which specifically uses Perl scripts to generate assembly code.
% The Vale tool suite~\cite{Vale} formalizes these practices with a more principled language and associated verification tools.
% However, code generation done in this style is significantly simpler than what we treat here, amounting mostly to loop unrolling, macro substitution, and computation of compile-time constants.
% Also, Vale involves a significantly larger trusted code base than with our approach, with no reduction to some kernel proof checker, instead placing trust in language-specific tooling and an SMT solver.

A variety of forms of pragmatic partial evaluation have been demonstrated, with Lightweight Modular Staging~\cite{LMS} in Scala as one of the best-known current examples.
A kind of type-based overloading for staging annotations is used to smooth the rough edges in writing code that manipulates syntax trees.
The LMS-Verify system~\cite{LMSVerify} can be used for formal verification of generated code after-the-fact.
Typically LMS-Verify has been used with relatively shallow properties (though potentially applied to larger and more sophisticated code bases than we tackle), not scaling to the kinds of functional-correctness properties that concern us here, justifying investment in verified partial evaluators.

\section{Future Work}

There are a number of natural extensions to our engine.
For instance, we do not yet allow pattern variables marked as ``constants only'' to apply to container datatypes; we limit the mixing of higher-order and polymorphic types, as well as limiting use of first-class polymorphism; we do not support rewriting with equalities of non-fully-applied functions; we only support decidable predicates as rule side conditions, and the predicates may only mention pattern variables restricted to matching constants; we have hardcoded support for a small set of container types and their eliminators; we support rewriting with equality and no other relations (e.g., subset inclusion); and we require decidable equality for all types mentioned in rules.
It may be helpful to design an engine that lifts some or all of these limitations, building on the basic structure that we present here.

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

%% Bibliography
\clearpage
\nocite{*}
\bibliography{rewriting}
\clearpage

\appendix

\section{Performance Bottlenecks of Proof-Producing Rewriting} \label{sec:setoid-rewrite-bottlenecks}

Although we have made our performance comparison against the built-in Coq tactics \mintinline{coq}{setoid_rewrite} and \mintinline{coq}{rewrite_strat}, by analyzing the performance in detail, we can argue that these performance bottlenecks are likely to hold for any proof assistant designed like Coq.
Detailed debugging reveals five performance bottlenecks in the existing rewriting tactics.
%(This section goes into detail that readers not interested in proof-assistant minutiae may want to skip, turning ahead to \autoref{sec:micro:LiftLetsMap}.)

\subsection{Bad performance scaling in sizes of existential-variable contexts}

We found that even when there are no occurrences fully matching the rule, \mintinline{coq}{setoid_rewrite} can still be \emph{cubic} in the number of binders (or, more accurately, quadratic in the number of binders with an additional multiplicative linear factor of the number of head-symbol matches).
Rewriting without any successful matches takes nearly as much time as \mintinline{coq}{setoid_rewrite} in this microbenchmark; by the time we are looking at goals with 400 binders, the difference is less than 5\%.

We posit that this overhead comes from \mintinline{coq}{setoid_rewrite} looking for head-symbol matches and then creating evars (existential variables) to instantiate the arguments of the lemmas for each head-symbol-match location; hence even if there are no matches of the rule as a whole, there may still be head-symbol matches.
Since Coq uses a locally nameless representation~\cite{LocallyNameless} for its terms, evar contexts are necessarily representated as \emph{named} contexts.
Representing a substitution between named contexts takes linear space, even when the substitution is trivial, and hence each evar incurs overhead linear in the number of binders above it.
Furthermore, fresh-name generation in Coq is quadratic in the size of the context, and since evar-context creation uses fresh-name generation, the additional multiplicative factor likely comes from fresh-name generation.
(Note, though, that this pattern suggests that the true performance is quartic rather than merely cubic.  However, doing a linear regression on a $\log$-$\log$ of the data suggests that the performance is genuinely cubic rather than quartic.)
%\todo{Can we cite an Coq issue that would de-anonymize us? 12524 (Andres says ``I would not cite, just claim coqdev acceptance if appropriate''}

Note that this overhead is inherent to the use of a locally nameless term representation.
To fix it, Coq would likely have to represent identity evar contexts using a compact representation, which is only naturally available for de Bruijn representations.
Any rewriting system that uses unification variables with a locally nameless (or named) context will incur at least quadratic overhead on this benchmark.

Note that \mintinline{coq}{rewrite_strat} uses exactly the same rewriting engine as \mintinline{coq}{setoid_rewrite}, just with a different strategy.
We found that \mintinline{coq}{setoid_rewrite} and \mintinline{coq}{rewrite_strat} have identical performance when there are no matches and generate identical proof terms when there are matches.
Hence we can conclude that the difference in performance between \mintinline{coq}{rewrite_strat} and \mintinline{coq}{setoid_rewrite} is entirely due to an increased number of failed rewrite attempts.

\subsection{Proof-term size}

Setting aside the performance bottleneck in constructing the matches in the first place, we can ask the question: how much cost is associated to the proof terms?
One way to ask this question in Coq is to see how long it takes to run \mintinline{coq}{Qed}.
While \mintinline{coq}{Qed}-time is asymptotically better, it is still quadratic in the number of binders.
This outcome is unsurprising, because the proof-term size is quadratic in the number of binders.
On this microbenchmark, we found that \mintinline{coq}{Qed}-time hits one second at about 250 binders, and using the best-fit quadratic line suggests that it would hit 10 seconds at about 800 binders and 100 seconds at about 2\,500 binders.
While this may be reasonable for the microbenchmarks, which only contain as many rewrite occurrences as there are binders, it would become unwieldy to try to build and typecheck such a proof with a rule for every primitive reduction step, which would be required if we want to avoid manually CPS-converting the code in Fiat Cryptography.

The quadratic factor in the proof term comes because we repeat subterms of the goal linearly in the number of rewrites.
For example, if we want to rewrite \mintinline{coq}{f (f x)} into \mintinline{coq}{g (g x)} by the equation \mintinline{coq}{∀ x, f x = g x}, then we will first rewrite \mintinline{coq}{f x} into \mintinline{coq}{g x}, and then rewrite \mintinline{coq}{f (g x)} into \mintinline{coq}{g (g x)}.
Note that \mintinline{coq}{g x} occurs three times (and will continue to occur in every subsequent step).

% We put this figure here so that the layout is better.
\begin{figure*}
\newcommand{\PlusZeroTreeMValSubfloat}[1]{%
  \def\mval{#1}%
  \subfloat[No binders ($m=\mval$)]{%
  \resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
    \einput{perf-Plus0Tree-only-m-\mval-cbv;rewrite-strat(bottomup).txt}
    \einput{perf-Plus0Tree-only-m-\mval-cbv;setoid-rewrite.txt}
    \einput{perf-Plus0Tree-only-m-\mval-cbv;rewrite-strat(topdown).txt}
    \einput{perf-Plus0Tree-only-m-\mval-cbv;rewrite!.txt}
    \einput{perf-Plus0Tree-only-m-\mval-Rewrite-for.txt}
    \einput{perf-Plus0Tree-only-m-\mval-rewriting.txt}
    \mval
  }
    \pgfplotsset{every axis legend/.append style={
        at={(0.5,-0.2)},
        anchor=north}}
    \begin{axis}[xlabel=$n$,ylabel=time (s),ymode=log,ymin=0.002,ymax=110]
      \addplot[mark=square,color=red] table{perf-Plus0Tree-only-m-\mval-cbv;rewrite-strat(bottomup).txt};
      \addplot[mark=*,color=red] table{perf-Plus0Tree-only-m-\mval-cbv;setoid-rewrite.txt};
      \addplot[mark=triangle,color=red] table{perf-Plus0Tree-only-m-\mval-cbv;rewrite-strat(topdown).txt};
      \addplot[mark=o,color=red] table{perf-Plus0Tree-only-m-\mval-cbv;rewrite!.txt};
      \addplot[mark=+,color=blue] table{perf-Plus0Tree-only-m-\mval-Rewrite-for.txt};
      \addplot[mark=x,color=ForestGreen] table{perf-Plus0Tree-only-m-\mval-rewriting.txt};
      \legend{rewrite\_strat bottomup,setoid\_rewrite,rewrite\_strat topdown,rewrite!,{Our approach including reification, cbv, etc.},Our approach (only rewriting)}
    \end{axis}
  \end{tikzpicture}}%
  \label{fig:timing-Plus0Tree-m=\mval}}%
}%
\PlusZeroTreeMValSubfloat{1}\qquad
\PlusZeroTreeMValSubfloat{2}\qquad
\PlusZeroTreeMValSubfloat{3}
  \caption{Timing of different partial-evaluation implementations for code with no binders for fixed $m$.  Note that we have a logarithmic time scale, because term size is proportional to $2^n$.} \label{fig:timing-Plus0Tree-fixed-m}
\end{figure*}

\subsection{Poor subterm sharing}

How easy is it to share subterms and create a linearly sized proof?
While it is relatively straightforward to share subterms using \mintinline{coq}{let} binders when the rewrite locations are not under any binders, it is not at all obvious how to share subterms when the terms occur under different binders.
Hence any rewriting algorithm that does not find a way to share subterms across different contexts will incur a quadratic factor in proof-building and proof-checking time, and we expect this factor will be significant enough to make applications to projects as large as Fiat Crypto infeasible.

\subsection{Overhead from the \mintinline{coq}{let} typing rule}

Suppose we had a proof-producing rewriting algorithm that shared subterms even under binders.
Would it be enough?
It turns out that even when the proof size is linear in the number of binders, the cost to typecheck it in Coq is still quadratic!
The reason is that when checking that \texttt{f : T} in a context \mintinline{coq}{x := v}, to check that \mintinline{coq}{let x := v in f} has type \texttt{T} (assuming that \mintinline{coq}{x} does not occur in \texttt{T}), Coq will substitute \mintinline{coq}{v} for \mintinline{coq}{x} in \texttt{T}.
So if a proof term has $n$ \mintinline{coq}{let} binders (e.g., used for sharing subterms), Coq will perform $n$ substitutions on the type of the proof term, even if none of the \mintinline{coq}{let}-binders are used.
If the number of \mintinline{coq}{let}-binders is linear in the size of the type, there is quadratic overhead in proof-checking time, even when the proof-term size is linear.

We performed a microbenchmark on a rewriting goal with no binders (because there is an obvious algorithm for sharing subterms in that case) and found that the proof-checking time reached about one second at about 2\,000 binders and reached 10 seconds at about 7\,000 binders.
While these results might seem good enough for Fiat Cryptography, we expect that there are hundreds of thousands of primitive reduction/rewriting steps even when there are only a few hundred binders in the output term, and we would need \mintinline{coq}{let}-binders for each of them.
Furthermore, we expect that getting such an algorithm correct would be quite tricky.

% We put this figure here so that the layout is better.
\begin{figure*}
\newcommand{\PlusZeroTreeNValSubfloat}[1]{%
  \def\nval{#1}%
  \subfloat[No binders ($n=\nval$)]{%
  \resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
      \einput{perf-Plus0Tree-only-n-\nval-cbv;rewrite-strat(bottomup).txt}
      \einput{perf-Plus0Tree-only-n-\nval-cbv;setoid-rewrite.txt}
      \einput{perf-Plus0Tree-only-n-\nval-cbv;rewrite-strat(topdown).txt}
      \einput{perf-Plus0Tree-only-n-\nval-cbv;rewrite!.txt}
      \einput{perf-Plus0Tree-only-n-\nval-Rewrite-for.txt}
      \einput{perf-Plus0Tree-only-n-\nval-rewriting.txt}
      \nval
  }
    \pgfplotsset{every axis legend/.append style={
        at={(0.5,-0.2)},
        anchor=north}}
    \begin{axis}[xlabel=$m$,ylabel=time (s),ymax=10]
      \addplot[mark=square,color=red] table{perf-Plus0Tree-only-n-\nval-cbv;rewrite-strat(bottomup).txt};
      \addplot[mark=*,color=red] table{perf-Plus0Tree-only-n-\nval-cbv;setoid-rewrite.txt};
      \addplot[mark=triangle,color=red] table{perf-Plus0Tree-only-n-\nval-cbv;rewrite-strat(topdown).txt};
      \addplot[mark=o,color=red] table{perf-Plus0Tree-only-n-\nval-cbv;rewrite!.txt};
      \addplot[mark=+,color=blue] table{perf-Plus0Tree-only-n-\nval-Rewrite-for.txt};
      \addplot[mark=x,color=ForestGreen] table{perf-Plus0Tree-only-n-\nval-rewriting.txt};
      \legend{rewrite\_strat bottomup,setoid\_rewrite,rewrite\_strat topdown,rewrite!,{Our approach including reification, cbv, etc.},Our approach (only rewriting)}
    \end{axis}
  \end{tikzpicture}}%
  \label{fig:timing-Plus0Tree-n=\nval}}%
}%
\PlusZeroTreeNValSubfloat{1}\qquad
\PlusZeroTreeNValSubfloat{2}\qquad
\PlusZeroTreeNValSubfloat{3}\qquad
%\PlusZeroTreeNValSubfloat{4}\qquad
%\PlusZeroTreeNValSubfloat{5}\qquad
%\PlusZeroTreeNValSubfloat{6}\qquad
%\PlusZeroTreeNValSubfloat{7}\qquad
%\PlusZeroTreeNValSubfloat{8}\qquad
\PlusZeroTreeNValSubfloat{9}
  \caption{Timing of different partial-evaluation implementations for code with no binders for fixed $n$ (1, 2, 3, and then we jump to 9)} \label{fig:timing-Plus0Tree-fixed-n}
\end{figure*}

% \begin{figure}
%   \begin{tikzpicture}[only marks]
%   \pgfplotsset{every axis legend/.append style={
%       at={(0.5,-0.2)},
%       anchor=north}}
%     \begin{axis}[xlabel=$\text{term size} = (m+1)\cdot 2^{n+1}-m$,ylabel=time (s)]
%       \addplot[mark=+,color=blue] table{perf-Plus0Tree-Rewrite-for.txt};
%       \addplot[mark=x,color=ForestGreen] table{perf-Plus0Tree-rewriting.txt};
%       \addplot[mark=o,color=red] table{perf-Plus0Tree-cbv;rewrite!.txt};
%       \addplot[mark=*,color=red] table{perf-Plus0Tree-cbv;setoid-rewrite.txt};
%       \addplot[mark=square,color=red] table{perf-Plus0Tree-cbv;rewrite-strat(bottomup).txt};
%       \addplot[mark=triangle,color=red] table{perf-Plus0Tree-cbv;rewrite-strat(topdown).txt};
%       \legend{Our approach including reification,Our approach (only rewriting),rewrite!,setoid\_rewrite,rewrite\_strat bottomup,rewrite\_strat topdown}
%     \end{axis}
%   \end{tikzpicture}
%   \caption{\label{fig:timing-Plus0Tree}Timing of different partial-evaluation implementations for code with no binders}
% \end{figure}

Fixing this quadratic bottleneck would, as far as we can tell, require deep changes in how Coq is implemented; it would either require reworking all of Coq to operate on some efficient representation of delayed substitutions paired with unsubstituted terms, or else it would require changing the typing rules of the type theory itself to remove this substitution from the typing rule for \mintinline{coq}{let}.
Note that there is a similar issue that crops up for function application and abstraction.

\subsection{Inherent advantages of reflection}

Finally, even if this quadratic bottleneck were fixed, \citet{Aehlig} reported a $10\times$--$100\times$ speed-up over the \emph{simp} tactic in Isabelle, which performs all of the intermediate rewriting steps via the kernel API.
Their results suggest that even if all of the super-linear bottlenecks were fixed---no small undertaking---rewriting and partial evaluation via reflection might still be orders of magnitude faster than any proof-term-generating tactic.

\section{Additional Benchmarking Plots} \label{sec:additionalPlots}

\subsection{Rewriting Without Binders} \label{sec:additionalPlots:Plus0Tree}

The code in \autoref{fig:micro:Plus0Tree:code} in \autoref{sec:micro:Plus0Tree} is parameterized on both $n$, the height of the tree, and $m$, the number of rewriting occurrences per node.
The plot in \autoref{fig:timing-Plus0Tree} displays only the case of $n=\NoBindersSubfloatNval$.
The plots in \autoref{fig:timing-Plus0Tree-fixed-m} display how performance scales as a factor of $n$ for fixed $m$, and the plots in \autoref{fig:timing-Plus0Tree-fixed-n} display how performance scales as a factor of $m$ for fixed $n$.
Note the logarithmic scaling on the time axis in the plots in \autoref{fig:timing-Plus0Tree-fixed-m}, as term size is proportional to $m\cdot 2^n$.

We can see from these graphs and the ones in \autoref{fig:timing-Plus0Tree-fixed-n} that
(a) we incur constant overhead over most of the other methods, which dominates on small examples;
(b) when the term is quite large and there are few opportunities for rewriting relative to the term size (i.e., $m \le 2$), we are worse than \mintinline{coq}{rewrite !Z.add_0_r} but still better than the other methods; and
(c) when there are many opportunities for rewriting relative to the term size ($m > 2$), we thoroughly dominate the other methods.

\clearpage

\subsection{Additional Information on the Fiat Cryptography Benchmark} \label{sec:additionalPlots:FiatCrypto} \label{sec:additionalMacro}

\begin{figure*}[b]
  \newcommand{\MacroSubfloat}[3]{%
    \subfloat[Timing of different partial-evaluation implementations for Fiat Cryptography as prime modulus grows (only #2 #3)]{%
  \resizebox{0.45\textwidth}{!}{\beginTikzpictureStamped[only marks]{
      \einput{perf-new-vm-times--only-#1-#3.txt}
      \einput{perf-old-vm-times--only-#1-#3.txt}
      \einput{perf-new-extraction-times--only-#1-#3.txt}
  }
    \pgfplotsset{every axis legend/.append style={
        at={(0.5,-0.2)},
        anchor=north}}
    \begin{axis}[xlabel=prime,ylabel=time (s),xmode=log, ymode=log,log basis x={2}]
      \addplot[mark=+] table{perf-new-vm-times--only-#1-#3.txt};
      \addplot[mark=o] table{perf-old-vm-times--only-#1-#3.txt};
      \addplot[mark=*] table{perf-new-extraction-times--only-#1-#3.txt};
      \legend{Our approach w/ Coq's VM,Old approach (handwritten-CPS+VM),Our approach w/ extracted OCaml}
    \end{axis}
  \end{tikzpicture}}%
  \label{fig:timing--only-#1-#3}}%
  }%
  \MacroSubfloat{UnsaturatedSolinas}{unsaturated Solinas}{x32}\qquad
  \MacroSubfloat{UnsaturatedSolinas}{unsaturated Solinas}{x64}\\
  \MacroSubfloat{WordByWordMontgomery}{word-by-word Montgomery}{x32}\qquad
  \MacroSubfloat{WordByWordMontgomery}{word-by-word Montgomery}{x64}%
  \caption{\label{fig:timing-macro-various}Timing of different partial-evaluation implementations for Fiat Cryptography {vs.} prime modulus}
\end{figure*}

It may also be useful to see performance results with absolute times, rather than normalized execution ratios {vs.} the original Fiat Cryptography implementation.
Furthermore, the benchmarks fit into four quite different groupings: elements of the cross product of two algorithms (unsaturated Solinas and word-by-word Montgomery) and bitwidths of target architectures (32-bit or 64-bit).
Here we provide absolute-time graphs by grouping in \autoref{fig:timing-macro-various}.

\clearpage
\supplement

\section{Additional Information on Microbenchmarks} \label{sec:additionalMicro}

We performed all benchmarks on a 3.5 GHz Intel Haswell running Linux and Coq 8.10.0.
We name the subsections here with the names that show up in the code supplement.

\subsection{UnderLetsPlus0} \label{sec:UnderLetsPlus0-more}

We provide more detail on the ``nested binders'' microbenchmark of \autoref{sec:micro:UnderLetsPlus0} displayed in \autoref{fig:timing-UnderLetsPlus0}.

Recall that we are removing all of the${}+0$s from
{\small \begin{align*}
  & \letin[{v_1 := v_0 + v_0 + 0}{}] \\
%  & \letin[{v_2 := v_1 + v_1 + 0}{}] \\
  & \vdots \\
  & \letin[{v_n := v_{n-1} + v_{n-1} + 0}{}] \\
  & v_n + v_n + 0
\end{align*}}%

The code used to define this microbenchmark is
\begin{minted}[fontsize=\small]{coq}
Definition make_lets_def (n:nat) (v acc : Z) :=
 @nat_rect (fun _ => Z * Z -> Z)
   (fun '(v, acc) => acc + acc + v)
   (fun _ rec '(v, acc) =>
     dlet acc := acc + acc + v in rec (v, acc))
   n
   (v, acc).
\end{minted}
We note some details of the rewriting framework that were glossed over in the main body of the paper, which are useful for using the code:
Although the rewriting framework does not support dependently typed constants, we can automatically preprocess uses of eliminators like \mintinline{coq}{nat_rect} and \mintinline{coq}{list_rect} into non-dependent versions.
The tactic that does this preprocessing is extensible via \Ltac{}'s reassignment feature.
Since pattern-matching compilation mixed with NbE requires knowing how many arguments a constant can be applied to, we must internally use a version of the recursion principle whose type arguments do not contain arrows; current preprocessing can handle recursion principles with either no arrows or one arrow in the motive.
Even though we will eventually plug in 0 for $v$, we jump through some extra hoops to ensure that our rewriter cannot cheat by rewriting away the ${}+0$ before reducing the recursion on $n$.

We can reduce this expression in three ways.

\subsubsection{Our Rewriter}
One lemma is required for rewriting with our rewriter:
\begin{minted}[fontsize=\small]{coq}
Lemma Z.add_0_r : forall z, z + 0 = z.
\end{minted}

Creating the rewriter takes about 12 seconds on the machine we used for running the performance experiments:
\begin{minted}[fontsize=\small]{coq}
Make myrew := Rewriter For (Z.add_0_r, eval_rect nat, eval_rect prod).
\end{minted}
Recall from \autoref{sec:explain-eval-rect} that \mintinline{coq}{eval_rect} is a definition provided by our framework for eagerly evaluating recursion associated with certain types.
It functions by triggering typeclass resolution for the lemmas reducing the recursion principle associated to the given type.
We provide instances for \texttt{nat}, \texttt{prod}, \texttt{list}, \texttt{option}, and \texttt{bool}.
Users may add more instances if they desire.

\subsubsection{\texorpdfstring{\texttt{setoid\_rewrite}}{setoid\_rewrite} and \texorpdfstring{\texttt{rewrite\_strat}}{rewrite\_strat}}
To give as many advantages as we can to the preexisting work on rewriting, we pre-reduce the recursion on \mintinline{coq}{nat}s using \texttt{cbv} before performing \texttt{setoid\_rewrite}.
(Note that \texttt{setoid\_rewrite} cannot itself perform reduction without generating large proof terms, and \texttt{rewrite\_strat} is not currently capable of sequencing reduction with rewriting internally due to bugs such as \coqbug{10923}.)
Rewriting itself is easy; we may use any of \texttt{repeat setoid\_rewrite Z.add\_0\_r}, \texttt{rewrite\_strat topdown Z.add\_0\_r}, or \texttt{rewrite\_strat bottomup Z.add\_0\_r}.

\subsection{Plus0Tree} \label{sec:Plus0Tree-more}

This is a version of \autoref{sec:UnderLetsPlus0-more} without any let binders, discussed in \autoref{sec:micro:Plus0Tree} but not displayed in \autoref{fig:multi-timing}.

We use two definitions for this microbenchmark:
\begin{minted}[fontsize=\small]{coq}
Definition iter (m : nat) (acc v : Z) :=
  @nat_rect (fun _ => Z -> Z)
    (fun acc => acc)
    (fun _ rec acc => rec (acc + v))
    m
    acc.
Definition make_tree (n m : nat) (v acc : Z) :=
 Eval cbv [iter] in
  @nat_rect (fun _ => Z * Z -> Z)
    (fun '(v, acc) => iter m (acc + acc) v)
    (fun _ rec '(v, acc) =>
      iter m (rec (v, acc) + rec (v, acc)) v)
    n
    (v, acc).
\end{minted}

\subsection{LiftLetsMap} \label{sec:LiftLetsMap-more}

We now discuss in more detail the ``binders and recursive functions'' example from \autoref{sec:micro:LiftLetsMap}.

The expression we want to get out at the end looks like:
\begin{align*}
    & \letin[{v_{1,1} := v + v}{}] \\
    & \vdots \\
    & \letin[{v_{1,n} := v + v}{}] \\
    & \letin[{v_{2,1} := v_{1,1} + v_{1,1}}{}] \\
    & \vdots \\
    & \letin[{v_{2,n} := v_{1,n} + v_{1,n}}{}] \\
    & \vdots \\
    & [v_{m,1}, \ldots, v_{m,n}]
\end{align*}

Recall that we make this example with the code
\begin{minted}[fontsize=\small]{coq}
Definition map_double (ls : list Z) :=
  list_rect _ [] (λ x xs rec, let y := x + x in y :: rec) ls.
Definition make (n : nat) (m : nat) (v : Z) :=
  nat_rect _ (List.repeat v n) (λ _ rec, map_double rec) m.
\end{minted}

We can perform this rewriting in four ways; see \autoref{fig:timing-LiftLetsMap}.

\subsubsection{Our Rewriter}
One lemma is required for rewriting with our rewriter:
\begin{minted}[fontsize=\small]{coq}
Lemma eval_repeat A x n
: @List.repeat A x ('n) = ident.eagerly nat_rect _ [] (λ k repeat_k, x :: repeat_k) ('n).
\end{minted}
Recall that the apostrophe marker (\verb|'|) is explained in \autoref{sec:explain-'}.
Recall again from \autoref{sec:explain-ident.eagerly} that we use \mintinline{coq}{ident.eagerly} to ask the reducer to simplify a case of primitive recursion by complete traversal of the designated argument's constructor tree.
Our current version only allows a limited, hard-coded set of eliminators with \mintinline{coq}{ident.eagerly} (\texttt{nat\_rect} on return types with either zero or one arrows, \texttt{list\_rect} on return types with either zero or one arrows, and \texttt{List.nth\_default}), but nothing in principle prevents automatic generation of the necessary code.
% Note that we use eliminators in the \mintinline{coq}{Thunked} namespace, which we provide in our library, to encode a version of the eliminator whose non-function cases are thunked rather than eagerly evaluated.
% That is, the type of, e.g., \mintinline{coq}{Thunked.nat_rect} is \mintinline{coq}{forall P, (unit -> P) -> (nat -> P -> P) -> nat -> P}.
% This is required in general for good performance with a call-by-value strategy, where you don't want to compute recursive cases that are then thrown away.
% Note that we don't really support any dependently typed constants; the use of \mintinline{coq}{nat_rect} in the first lemma is turned into a \mintinline{coq}{Thunked.nat_rect} automatically.
% Finally, we will note that our rewriter does not support matching on $\lambda$s on the left-hand-side, which is why the thunking cannot be done entirely internally on the left-hand-side.
% (It is performed during a preprocessing stage internally on the right-hand-side.)
% That is, the rewrite rule \emph{must} bind the thunked cases as part of the pattern.

We construct our rewriter with
\begin{minted}[fontsize=\small]{coq}
Make myrew := Rewriter For (eval_repeat, eval_rect list, eval_rect nat)
  (with extra idents (Z.add)).
\end{minted}
On the machine we used for running all our performance experiments, this command takes about 13 seconds to run.
Note that all identifiers which appear in any goal to be rewritten must either appear in the type of one of the rewrite rules or in the tuple passed to \texttt{with extra idents}.

Rewriting is relatively simple, now.
Simply invoke the tactic \mintinline{coq}{Rewrite_for myrew}.
We support rewriting on only the left-hand-side and on only the right-hand-side using either the tactic \mintinline{coq}{Rewrite_lhs_for myrew} or else the tactic \mintinline{coq}{Rewrite_rhs_for myrew}, respectively.

\subsubsection{\texorpdfstring{\texttt{rewrite\_strat}}{rewrite\_strat}}

To reduce adequately using \texttt{rewrite\_strat}, we need the following two lemmas:
\begin{minted}[fontsize=\small]{coq}
Lemma lift_let_list_rect T A P N C (v : A) fls
: @list_rect T P N C (Let_In v fls) = Let_In v (fun v => @list_rect T P N C (fls v)).
Lemma lift_let_cons T A x (v : A) f
: @cons T x (Let_In v f) = Let_In v (fun v => @cons T x (f v)).
\end{minted}

Note that \mintinline{coq}{Let_In} is the constant we use for writing \letin{} expressions that do not reduce under $\zeta$.
Throughout most of this paper, anywhere that \letin{} appears, we have actually used \mintinline{coq}{Let_In} in the code.
It would alternatively be possible to extend the reification preprocessor to automatically convert \letin{} to \mintinline{coq}{Let_In}, but this may cause problems when converting the interpretation of the reified term with the pre-reified term, as Coq's conversion does not allow fine-tuning of when to inline or unfold \mintinline{coq}{let}s.

To rewrite, we start with \mintinline{coq}{cbv [example make map_dbl]} to expose the underlying term to rewriting.
One would hope that one could just add these two hints to a database \mintinline{coq}{db} and then write \texttt{rewrite\_strat (repeat (eval cbn [list\_rect]; try bottomup hints db))}, but unfortunately this does not work due to a number of bugs in Coq: \coqbug{10934}, \coqbug{10923}, \coqbug{4175}, \coqbug{10955}, and the potential to hit \coqbug{10972}.
Instead, we must put the two lemmas in separate databases, and then write \texttt{repeat (cbn [list\_rect]; (rewrite\_strat (try repeat bottomup hints db1)); (rewrite\_strat (try repeat bottomup hints db2)))}.
Note that the rewriting with \mintinline{coq}{lift_let_cons} can be done either top-down or bottom-up, but \texttt{rewrite\_strat} breaks if the rewriting with \mintinline{coq}{lift_let_list_rect} is done top-down.

\subsubsection{CPS and the VM}
If we want to use Coq's built-in VM reduction without our rewriter, to achieve the prior state-of-the-art performance, we can do so on this example, because it only involves partial reduction and not equational rewriting.
However, we must (a) module-opacify the constants which are not to be unfolded, and (b) rewrite all of our code in CPS.

Then we are looking at
\begin{align*}
    \text{map\_dbl\_cps}(\ell,k) & \defeq \begin{cases} k([]) & \text{if }\ell = [] \\
        \letin[{y := h +_\text{ax} h}{}] & \text{if }\ell = h::t \\
        \text{map\_dbl\_cps}(t, \\
        \qquad(\lambda ys, k(y :: ys)))
    \end{cases} \\
    \text{make\_cps}(n, m, v, k) & \defeq \begin{cases} k([\underbrace{v, \ldots, v}_n]) & \text{if }m = 0 \\
        \text{make\_cps}(n, m-1, v, & \text{if }m > 0 \\
        \quad(\lambda \ell,\text{map\_dbl\_cps}(\ell, k))
    \end{cases} \\
    \text{example\_cps}_{n, m} & \defeq \forall v,\ \text{make\_cps}(n, m, v, \lambda x.\,x) = []
\end{align*}

Then we can just run \texttt{vm\_compute}.
Note that this strategy, while quite fast, results in a stack overflow when $n \cdot m$ is larger than approximately $2.5\cdot 10^4$.
This is unsurprising, as we are generating quite large terms.
Our framework can handle terms of this size but stack-overflows on only slightly larger terms.

\subsubsection{Takeaway}

From this example, we conclude that \texttt{rewrite\_strat} is unsuitable for computations involving large terms with many binders, especially in cases where reduction and rewriting need to be interwoven, and that the many bugs in \texttt{rewrite\_strat} result in confusing gymnastics required for success.
The prior state of the art---writing code in CPS---suitably tweaked by using module pacity to allow \texttt{vm\_compute}, remains the best performer here, though the cost of rewriting everything is CPS may be prohibitive.
Our method soundly beats \texttt{rewrite\_strat}.
We are additionally bottlenecked on \texttt{cbv}, which is used to unfold the goal post-rewriting and costs about a minute on the largest of terms; see Coq bug \coqbug{11151} for a discussion on what is wrong with Coq's reduction here.

\subsection{SieveOfEratosthenes} \label{sec:Eratosthenes}

We define the sieve using \mintinline{coq}{PositiveMap.t} and \mintinline{coq}{list Z}:
\begin{minted}[fontsize=\small]{coq}
Definition sieve' (fuel : nat) (max : Z) :=
 List.rev
  (fst
   (@nat_rect
    (λ _, list Z (* primes *) *
     PositiveSet.t (* composites *) *
     positive (* np (next_prime) *) ->
     list Z (* primes *) *
     PositiveSet.t (* composites *))
    (λ '(primes, composites, next_prime),
     (primes, composites))
    (λ _ rec '(primes, composites, np),
      rec
       (if (PositiveSet.mem np composites ||
            (Z.pos np >? max))%bool%Z
        then
         (primes, composites, Pos.succ np)
        else
         (Z.pos np :: primes,
          List.fold_right
           PositiveSet.add
           composites
           (List.map
            (λ n, Pos.mul (Pos.of_nat (S n)) np)
            (List.seq 0 (Z.to_nat(max/Z.pos np)))),
          Pos.succ np)))
    fuel
    (nil, PositiveSet.empty, 2%positive))).

Definition sieve (n : Z)
  := Eval cbv [sieve'] in sieve' (Z.to_nat n) n.
\end{minted}

We need four lemmas and an additional instance to create the rewriter:
\begin{minted}[fontsize=\small]{coq}
Lemma eval_fold_right A B f x ls :
@List.fold_right A B f x ls
= ident.eagerly list_rect _ _
    x
    (λ l ls fold_right_ls, f l fold_right_ls)
    ls.

Lemma eval_app A xs ys :
xs ++ ys
= ident.eagerly list_rect A _
    ys
    (λ x xs app_xs_ys, x :: app_xs_ys)
    xs.

Lemma eval_map A B f ls :
@List.map A B f ls
= ident.eagerly list_rect _ _
    []
    (λ l ls map_ls, f l :: map_ls)
    ls.

Lemma eval_rev A xs :
@List.rev A xs
= (@list_rect _ (fun _ => _))
    []
    (λ x xs rev_xs, rev_xs ++ [x])%list
    xs.

Scheme Equality for PositiveSet.tree.

Definition PositiveSet_t_beq
   : PositiveSet.t -> PositiveSet.t -> bool
  := tree_beq.

Global Instance PositiveSet_reflect_eqb
 : reflect_rel (@eq PositiveSet.t) PositiveSet_t_beq
 := reflect_of_brel
      internal_tree_dec_bl internal_tree_dec_lb.
\end{minted}

We then create the rewriter with
\begin{minted}[fontsize=\small]{coq}
Make myrew := Rewriter For
  (eval_rect nat, eval_rect prod, eval_fold_right,
   eval_map, do_again eval_rev, eval_rect bool,
   @fst_pair, eval_rect list, eval_app)
   (with extra idents (Z.eqb, orb, Z.gtb,
    PositiveSet.elements, @fst, @snd,
    PositiveSet.mem, Pos.succ, PositiveSet.add,
    List.fold_right, List.map, List.seq, Pos.mul,
    S, Pos.of_nat, Z.to_nat, Z.div, Z.pos, O,
    PositiveSet.empty))
  (with delta).
\end{minted}

To get \texttt{cbn} and \texttt{simpl} to unfold our term fully, we emit
\begin{minted}[fontsize=\small]{coq}
Global Arguments Pos.to_nat !_ / .
\end{minted}

\section{Experience vs.\ Lean and \texorpdfstring{\texttt{setoid\_rewrite}}{setoid\_rewrite}\label{sec:lean}}

Although all of our toy examples work with \texttt{setoid\_rewrite} or \texttt{rewrite\_strat} (until the terms get too big), even the smallest of examples in Fiat Cryptography fell over using these tactics.
When attempting to use \texttt{rewrite\_strat} for partial evaluation and rewriting on unsaturated Solinas with 1 limb on small primes (such as 29), we were able to get \texttt{rewrite\_strat} to finish after about 90 seconds.
The bugs in \texttt{rewrite\_strat} made finding the right magic invocation quite painful, nonetheless; the invocation we settled on involved \emph{sixteen} consecutive calls to \texttt{rewrite\_strat} with varying arguments and strategies.
Trying to synthesize code for two limbs on slightly larger primes (such as 113, which needs two limbs on a 64-bit machine) took about three hours.
The widely used primes tend to have around five to ten limbs; we leave extrapolating this slowdown to the reader.

We have attached this experiment using \verb|rewrite_strat| as \verb|fiat_crypto_via_rewrite_strat.v|, which is meant to be run in emacs/PG from inside the \verb|fiat-crypto| directory, or in \verb|coqc| by setting \verb|COQPATH| to the value emitted by \texttt{make printenv} in \verb|fiat-crypto| and then invoking the command \texttt{coqc -q -R /path/to/fiat-crypto/src Crypto /path/to/fiat\_crypto\_via\_rewrite\_strat.v}.
To test with the two-limb prime 113, change \verb|of_string "2^5-3" 8| in the definition of \verb|p| to \verb|of_string "2^7-15" 64|.

We also tried Lean, in the hopes that rewriting in Lean, specifically optimized for performance, would be up to the challenge.
Although Lean performed about 30\% better than Coq on the 1-limb example, taking a bit under a minute, it did not complete on the two-limb example even after four hours (after which we stopped trying), and a five-limb example was still going after 40 hours.

We have attached our experiments with running \texttt{rewrite} in Lean on the Fiat Cryptography code as a supplement as well.
We used Lean version 3.4.2, commit cbd2b6686ddb, Release.
Run \texttt{make} in \texttt{fiat-crypto-lean} to run the one-limb example;
change \texttt{open ex} to \texttt{open ex2} to try the two-limb example, or to \texttt{open ex5} to try the five-limb example.

\section{Reading the Code Supplement} \label{sec:CodeSupplement-more}

% allow breaks after slashes, for long paths
\catcode`\/=\active
\def/{\slash}%

We have attached both the code for implementing the rewriter, as well as a copy of Fiat Cryptography adapted to use the rewriting framework.
Both code supplements build with Coq 8.9 and Coq 8.10, and they require that whichever OCaml was used to build Coq be installed on the system to permit building plugins.
(If Coq was installed via opam, then the correct version of OCaml will automatically be available.)
Both code bases can be built by running \texttt{make} in the top-level directory.

The performance data for both repositories are included at the top level as \texttt{.txt} and \texttt{.csv} files.

The performance data for the microbenchmarks can be rebuilt using \texttt{make perf-SuperFast perf-Fast perf-Medium} followed by \texttt{make perf-csv} to get the \texttt{.txt} and \texttt{.csv} files.
The microbenchmarks should run in about 24 hours when run with \texttt{-j5} on a 3.5 GHz machine.
There also exist targets \texttt{perf-Slow} and \texttt{perf-VerySlow}, but these take significantly longer.

The performance data for the macrobenchmark can be rebuilt from the Fiat Cryptography copy included by running \texttt{make perf -k}.
We ran this with \texttt{PERF\_MAX\_TIME=3600} to allow each benchmark to run for up to an hour; the default is 10 minutes per benchmark.
Expect the benchmarks to take over a week of time with an hour timeout and five cores.
Some tests are expected to fail, making \texttt{-k} a necessary flag.
Again, the \texttt{perf-csv} target will aggregate the logs and turn them into \texttt{.txt} and \texttt{.csv} files.

The entry point for the rewriter is the Coq source file \texttt{rewriter/src/Rewriter/Util/plugins/RewriterBuild.v}.

The rewrite rules used in Fiat Cryptography are defined in \texttt{fiat-crypto/src/Rewriter/Rules.v} and proven in \texttt{fiat-crypto/src/Rewriter/RulesProofs.v}.
Note that the Fiat Cryptography copy uses \verb|COQPATH| for dependency management, and \verb|.dir-locals.el| to set \verb|COQPATH| in emacs/PG; you must accept the setting when opening a file in the directory for interactive compilation to work.
Thus interactive editing either requires ProofGeneral or manual setting of \verb|COQPATH|.
The correct value of \verb|COQPATH| can be found by running \verb|make printenv|.

We will now go through this paper and describe where to find each reference in the code base.

\newcommand{\autocommanameref}[1]{\autoref{#1}, \nameref{#1}}

\subsection{Code from \autocommanameref{sec:intro}}

\subsubsection{Code from \autocommanameref{sec:motivating-example}}

The \texttt{prefixSums} example appears in the Coq source file \texttt{rewriter/src/Rewriter/Rewriter/Examples/PrefixSums.v}.
Note that we use \texttt{dlet} rather than \mintinline{coq}{let} in binding \texttt{acc'} so that we can preserve the \mintinline{coq}{let} binder even under $\iota$ reduction, which much of Coq's infrastructure performs eagerly.
Because we attempt to isolate the dependency on the axiom of functional extensionality as much as possible, we also in practice require \texttt{Proper} instances for each higher-order identifier saying that each constant respects function extensionality.
We hope to remove the dependency on function extensionality altogether in the future.
Although we glossed over this detail in the body of this paper, we also prove
\begin{minted}[fontsize=\small]{coq}
Global Instance: forall A B,
 Proper ((eq ==> eq ==> eq) ==> eq ==> eq ==> eq)
        (@fold_left A B).
\end{minted}

The \mintinline{coq}{Make} command is exposed in \texttt{rewriter/src/Rewriter/Util/plugins/RewriterBuild.v} and defined in \texttt{rewriter/src/Rewriter/Util/plugins/rewriter\_build\_plugin.mlg}.
Note that one must run \texttt{make} to create this latter file; it is copied over from a version-specific file at the beginning of the build.

\label{sec:code:eval-rect}%
\label{sec:code:ident.eagerly}%
The \verb|do_again|, \verb|eval_rect|, and \verb|ident.eagerly| constants are defined at the bottom of module \verb|RewriteRuleNotations| in \texttt{rewriter/src/Rewriter/Language/Pre.v}.

\subsubsection{Code from \autocommanameref{sec:trusted-code-base-size}}

There is no code mentioned in this section.

\subsubsection{Code from \autocommanameref{sec:our-solution}}

We claimed that our solution meets five criteria.
We briefly justify each criterion with a sentence or a pointer to code:
\begin{itemize}
  \item
    We claimed that we \textbf{did not grow the trusted base} (excepting the axiom of functional extensionality).
    In any example file (of which a couple can be found in \texttt{rewriter/src/Rewriter/Rewriter/Examples/}), the \mintinline{coq}{Make} command creates a rewriter package.
    Running \texttt{Print Assumptions} on this new constant (often named \texttt{rewriter} or \texttt{myrew}) should demonstrate a lack of axioms other than functional extensionality.
    \texttt{Print Assumptions} may also be run on the proof that results from using the rewriter.
  \item
    We claimed \textbf{fast} partial evaluation with reasonable memory use; we assume that the performance graphs stand on their own to support this claim.
    Note that memory usage can be observed by making the benchmarks while passing \texttt{TIMED=1} to \texttt{make}.
  \item
    We claimed to allow reduction that \textbf{mixes} \emph{rules of the definitional equality} with \emph{equalities proven explicitly as theorems}; the ``rules of the definitional equality'' are, for example, $\beta$ reduction, and we assert that it should be self-evident that our rewriter supports this.
  \item
    We claimed common-subterm \textbf{sharing preservation}.
    This is implemented by supporting the use of the \texttt{dlet} notation which is defined in \texttt{rewriter/src/Rewriter/Util/LetIn.v} via the \texttt{Let\_In} constant.
    We will come back to the infrastructure that supports this.
  \item
    We claimed \textbf{extraction of standalone partial evaluators}.
    The extraction is performed in the files \texttt{perf\_unsaturated\_solinas.v} and \texttt{perf\_word\_by\_word\_montgomery.v}, and the files \texttt{saturated\_solinas.v}, \texttt{unsaturated\_solinas.v}, and \texttt{word\_by\_word\_montgomery.v}, all in the directory \texttt{fiat-crypto/src/ExtractionOCaml/}.
    The OCaml code can be extracted and built using the target \texttt{make standalone-ocaml} (or \texttt{make perf-standalone} for the \texttt{perf\_} binaries).
    There may be some issues with building these binaries on Windows as some versions of \texttt{ocamlopt} on Windows seem not to support outputting binaries without the \texttt{.exe} extension.
\end{itemize}

The P-384 curve is mentioned.
This is the curve with modulus $2^{384} - 2^{128} - 2^{96} + 2^{32} - 1$; its benchmarks can be found in files matching the glob \texttt{fiat-crypto/src/Rewriter/PerfTesting/Specific/generated/p2384m2128m296p232m1\_\_*\_\_word\_by\_word\_montgomery\_*}.
The output \texttt{.log} files are included in the tarball; the \texttt{.v} and \texttt{.sh} files are automatically generated in the course of running \texttt{make perf -k}.

We mention integration with abstract interpretation; the abstract-interpretation pass is implemented in \texttt{fiat-crypto/src/AbstractInterpretation/}.

\subsection{Code from \autocommanameref{sec:trust}}

The individual rewritings mentioned are implemented via the \texttt{Rewrite\_*} tactics exported at the top of \texttt{rewriter/src/Rewriter/Util/plugins/RewriterBuild.v}.
These tactics bottom out in tactics defined at the bottom of \texttt{rewriter/src/Rewriter/Rewriter/AllTactics.v}.

\subsubsection{Code from \autocommanameref{sec:nine-steps}} \label{sec:code:nine-steps}

We match the nine steps with functions from the source code:
\begin{enumerate}
  \item
    The given lemma statements are scraped for which named functions and types the rewriter package will support.
    This is performed by \texttt{rewriter\_scrape\_data} in the file \texttt{rewriter/src/Rewriter/Util/plugins/rewriter\_build.ml} which invokes the \Ltac{} tactic named \texttt{make\_scrape\_data} in a submodule in the source file \texttt{rewriter/src/Rewriter/Language/IdentifiersBasicGenerate.v} on a goal headed by the constant we provide under the name \texttt{Pre.ScrapedData.t\_with\_args} in \texttt{rewriter/src/Rewriter/Language/PreCommon.v}.
  \item
    Inductive types enumerating all available primitive types and functions are emitted.
    This step is performed by \texttt{rewriter\_emit\_inductives} in file \texttt{rewriter/src/Rewriter/Util/plugins/rewriter\_build.ml} invoking tactics, like \texttt{make\_base\_elim} in \texttt{rewriter/src/Rewriter/Language/IdentifiersBasicGenerate.v}, on goals headed by constants from \texttt{rewriter/src/Rewriter/Language/IdentifiersBasicLibrary.v}, including the constant \texttt{base\_elim\_with\_args} for example, to turn scraped data into eliminators for the inductives.
    The actual emitting of inductives is performed by code in the file \texttt{rewriter/src/Rewriter/Util/plugins/inductive\_from\_elim.ml}.
  \item
    Tactics generate all of the necessary definitions and prove all of the necessary lemmas for dealing with this particular set of inductive codes.
    This step is performed by the tactic \texttt{make\_rewriter\_of\_scraped\_and\_ind} in the source file \texttt{rewriter/src/Rewriter/Util/plugins/rewriter\_build.ml} which invokes the tactic \texttt{make\_rewriter\_all} defined in the file \texttt{rewriter/src/Rewriter/Rewriter/AllTactics.v} on a goal headed by the provided constant \texttt{VerifiedRewriter\_with\_ind\_args} defined in \texttt{rewriter/src/Rewriter/Rewriter/ProofsCommon.v}.
    The definitions emitted can be found by looking at the tactic \texttt{Build\_Rewriter} in \texttt{rewriter/src/Rewriter/Rewriter/AllTactics.v}, the \Ltac{} tactics \texttt{build\_package} in \texttt{rewriter/src/Rewriter/Language/IdentifiersBasicGenerate.v} and also in \texttt{rewriter/src/Rewriter/Language/IdentifiersGenerate.v} (there is a different tactic named \texttt{build\_package} in each of these files), and \texttt{prove\_package\_proofs\_via} which can be found in \texttt{rewriter/src/Rewriter/Language/IdentifiersGenerateProofs.v}.
  \item
    The statements of rewrite rules are reified and soundness and syntactic-well-formedness lemmas are proven about each of them.
    This is done as part of the previous step, when the tactic \texttt{make\_rewriter\_all} transitively calls \texttt{Build\_Rewriter} from \texttt{rewriter/src/Rewriter/Rewriter/AllTactics.v}.
    Reification is handled by the tactic \texttt{Build\_RewriterT} in \texttt{rewriter/src/Rewriter/Rewriter/Reify.v}, while soundness and the syntactic-well-formedness proofs are handled by the tactics \texttt{prove\_interp\_good} and \texttt{prove\_good} respectively, both in the source file \texttt{rewriter/src/Rewriter/Rewriter/ProofsCommonTactics.v}.
  \item
    The definitions needed to perform reification and rewriting and the lemmas needed to prove correctness are assembled into a single package that can be passed by name to the rewriting tactic.
    This step is also performed by \texttt{make\_rewriter\_of\_scraped\_and\_ind} in the source file \texttt{rewriter/src/Rewriter/Util/plugins/rewriter\_build.ml}.
\end{enumerate}

When we want to rewrite with a rewriter package in a goal, the following steps are performed, with code in the following places:
\begin{enumerate}
  \item
    We rearrange the goal into a closed logical formula: all free-variable quantification in the proof context is replaced by changing the equality goal into an equality between two functions (taking the free variables as inputs).
    Note that it is not actually an equality between two functions but rather an \texttt{equiv} between two functions, where \texttt{equiv} is a custom relation we define indexed over type codes that is equality up to function extensionality.
    This step is performed by the tactic \texttt{generalize\_hyps\_for\_rewriting} in \texttt{rewriter/src/Rewriter/Rewriter/AllTactics.v}.
  \item
    We reify the side of the goal we want to simplify, using the inductive codes in the specified package.  That side of the goal is then replaced with a call to a denotation function on the reified version.
    This step is performed by the tactic \texttt{do\_reify\_rhs\_with} in \texttt{rewriter/src/Rewriter/Rewriter/AllTactics.v}.
  \item
    We use a theorem stating that rewriting preserves denotations of well-formed terms to replace the denotation subterm with the denotation of the rewriter applied to the same reified term.
    We use Coq's built-in full reduction (\texttt{vm\_compute}) to reduce the application of the rewriter to the reified term.
    This step is performed by the tactic \texttt{do\_rewrite\_with} in \texttt{rewriter/src/Rewriter/Rewriter/AllTactics.v}.
  \item
    Finally, we run \texttt{cbv} (a standard call-by-value reducer) to simplify away the invocation of the denotation function on the concrete syntax tree from rewriting.
    This step is performed by the tactic \texttt{do\_final\_cbv} in \texttt{rewriter/src/Rewriter/Rewriter/AllTactics.v}.
\end{enumerate}
These steps are put together in the tactic \texttt{Rewrite\_for\_gen} in \texttt{rewriter/src/Rewriter/Rewriter/AllTactics.v}.

\subsubsection{Our Approach in More Than Nine Steps}

As the nine steps of \autoref{sec:nine-steps} do not exactly match the code, we describe here a more accurate version of what is going on.
For ease of readability, we do not clutter this description with references to the code supplement, instead allowing the reader to match up the steps here with the more coarse-grained ones in \autoref{sec:nine-steps} or \autoref{sec:code:nine-steps}.

In order to allow easy invocation of our rewriter, a great deal of code (about 6500 lines) needed to be written.
Some of this code is about reifying rewrite rules into a form that the rewriter can deal with them in.
Other code is about proving that the reified rewrite rules preserve interpretation and are well-formed.
We wrote some plugin code to automatically generate the inductive type of base-type codes and identifier codes, as well as the two variants of the identifier-code inductive used internally in the rewriter.
One interesting bit of code that resulted was a plugin that can emit an inductive declaration given the Church encoding (or eliminator) of the inductive type to be defined.
We wrote a great deal of tactic code to prove basic properties about these inductive types, from the fact that one can unify two identifier codes and extract constraints on their type variables from this unification, to the fact that type codes have decidable equality.
Additional plugin code was written to invoke the tactics that construct these definitions and prove these properties, so that we could generate an entire rewriter from a single command, rather than having the user separately invoke multiple commands in sequence.

In order to build the precomputed rewriter, the following actions are performed:
\begin{enumerate}
    \item
    The terms and types to be supported by the rewriter package are scraped from the given lemmas.
    \item
    An inductive type of codes for the types is emitted, and then three different versions of inductive codes for the identifiers are emitted (one with type arguments, one with type arguments supporting pattern type variables, and one without any type arguments, to be used internally in pattern-matching compilation).
    \item
    Tactics generate all of the necessary definitions and prove all of the necessary lemmas for dealing with this particular set of inductive codes.
    Definitions cover categories like ``Boolean equality on type codes'' and ``how to extract the pattern type variables from a given identifier code,'' and lemma categories include ``type codes have decidable equality'' and ``the types being coded for have decidable equality'' and ``the identifiers all respect function extensionality.''
    \item
    The rewrite rules are reified, and we prove interpretation-correctness and well-formedness lemmas about each of them.
    \item
    The definitions needed to perform reification and rewriting and the lemmas needed to prove correctness are assembled into a single package that can be passed by name to the rewriting tactic.
    \item
    The denotation functions for type and identifier codes are marked for early expansion in the kernel via the \mintinline{coq}{Strategy} command;
    this is necessary for conversion at \mintinline{coq}{Qed}-time to perform reasonably on enormous goals.
\end{enumerate}

When we want to rewrite with a rewriter package in a goal, the following steps are performed:
\begin{enumerate}
    \item
    We use \mintinline{coq}{etransitivity} to allow rewriting separately on the left- and right-hand-sides of an equality.
    Note that we do not currently support rewriting in non-equality goals, but this is easily worked around using \texttt{let v := open\_constr:(\_) in replace <some term> with v} and then rewriting in the second goal.
    \item
    We revert all hypotheses mentioned in the goal, and change the form of the goal from a universally quantified statement about equality into a statement that two functions are extensionally equal.
    Note that this step will fail if any hypotheses are functions not known to respect function extensionality via typeclass search.
    \item
    We reify the side of the goal that is not an existential variable using the inductive codes in the specified package; the resulting goal equates the denotation of the newly reified term with the original evar.
    \item
    We use a lemma stating that rewriting preserves denotations of well-formed terms to replace the goal with the rewriter applied to our reified term.
    We use \texttt{vm\_compute} to prove the well-formedness side condition reflectively.
    We use \texttt{vm\_compute} again to reduce the application of the rewriter to the reified term.
    \item
    Finally, we run \texttt{cbv} to unfold the denotation function, and we instantiate the evar with the resulting rewritten term.
\end{enumerate}

There are a couple of steps that contribute to the trusted base.
We must trust that the rewriter package we generate from the rewrite rules in fact matches the rewrite rules we want to rewrite with.
This involves partially trusting the scraper, the reifier, and the glue code.
We must also trust the VM we use for reduction at various points in rewriting.
Otherwise, everything is checked by Coq.
We do, however, depend on the axiom of function extensionality in one place in the rewriter proof; after spending a couple of hours trying to remove this axiom, we temporarily gave up.

\subsection{Code from \autocommanameref{sec:structure}}

The expression language $e$ corresponds to the inductive \texttt{expr} type defined in module \texttt{Compilers.expr} in \texttt{rewriter/src/Rewriter/Language/Language.v}.

\subsubsection{Code from \autocommanameref{sec:pattern-matching-compilation-and-evaluation}}

The pattern-matching compilation step is done by the tactic \texttt{CompileRewrites} in \texttt{rewriter/src/Rewriter/Rewriter/Rewriter.v}, which just invokes the Gallina definition named \texttt{compile\_rewrites} with ever-increasing amounts of fuel until it succeeds.
(It should never fail for reasons other than insufficient fuel, unless there is a bug in the code.)
The workhorse function here is \texttt{compile\_rewrites\_step}.

The decision-tree evaluation step is done by the definition \texttt{eval\_rewrite\_rules}, also in the file \texttt{rewriter/src/Rewriter/Rewriter/Rewriter.v}.
The correctness lemmas are the theorem \texttt{eval\_rewrite\_rules\_correct} in the file \texttt{rewriter/src/Rewriter/Rewriter/InterpProofs.v} and the theorem \texttt{wf\_eval\_rewrite\_rules} in \texttt{rewriter/src/Rewriter/Rewriter/Wf.v}.
Note that the second of these lemmas, not mentioned in the paper, is effectively saying that for two related syntax trees, \texttt{eval\_rewrite\_rules} picks the same rewrite rule for both.
(We actually prove a slightly weaker lemma, which is a bit harder to state in English.)

The third step of rewriting with a given rule is performed by the definition \texttt{rewrite\_with\_rule} in \texttt{rewriter/src/Rewriter/Rewriter/Rewriter.v}.
The correctness proof goes by the name \texttt{interp\_rewrite\_with\_rule} in \texttt{rewriter/src/Rewriter/Rewriter/InterpProofs.v}.
Note that the well-formedness-preservation proof for this definition in inlined into the proof of the lemma \verb|wf_eval_rewrite_rules| mentioned above.

The inductive description of decision trees is \verb|decision_tree| in \texttt{rewriter/src/Rewriter/Rewriter/Rewriter.v}.

The pattern language is defined as the inductive \verb|pattern| in \texttt{rewriter/src/Rewriter/Rewriter/Rewriter.v}.
Note that we have a \verb|Raw| version and a typed version; the pattern-matching compilation and decision-tree evaluation of \citet{Aehlig} is an algorithm on untyped patterns and untyped terms.
We found that trying to maintain typing constraints led to headaches with dependent types.
Therefore when doing the actual decision-tree evaluation, we wrap all of our expressions in the dynamically typed \verb|rawexpr| type and all of our patterns in the dynamically typed \verb|Raw.pattern| type.
We also emit separate inductives of identifier codes for each of the \verb|expr|, \verb|pattern|, and \verb|Raw.pattern| type families.

We partially evaluate the partial evaluator defined by \verb|eval_rewrite_rules| in the \Ltac{} tactic \verb|make_rewrite_head| in \texttt{rewriter/src/Rewriter/Rewriter/Reify.v}.

\subsubsection{Code from \autocommanameref{sec:thunk-eval-subst-term}}

The type $\text{NbE}_t$ mentioned in this paper is not actually used in the code; the version we have is described in \autoref{sec:under-lets} as the definition \verb|value'| in \texttt{rewriter/src/Rewriter/Rewriter/Rewriter.v}.

The functions \verb|reify| and \verb|reflect| are defined in \texttt{rewriter/src/Rewriter/Rewriter/Rewriter.v} and share names with the functions in the paper.
The function \texttt{reduce} is named \verb|rewrite_bottomup| in the code, and the closest match to NbE is \verb|rewrite|.

\subsection{Code from \autocommanameref{sec:scaling}}

\subsubsection{Code from \autocommanameref{sec:PHOAS}}

The inductives \verb|type|, \verb|base_type| (actually the inductive type \verb|base.type.type| in the supplemental code), and \verb|expr|, as well as the definition \verb|Expr|, are all defined in \texttt{rewriter/src/Rewriter/Language/Language.v}.
The definition \verb|denoteT| is the fixpoint \verb|type.interp| (the fixpoint \verb|interp| in the module \verb|type|) in \texttt{rewriter/src/Rewriter/Language/Language.v}.
The definition \verb|denoteE| is \verb|expr.interp|, and \verb|DenoteE| is the fixpoint \verb|expr.Interp|.

As mentioned above, \verb|nbeT| does not actually exist as stated but is close to \verb|value'| in \texttt{rewriter/src/Rewriter/Rewriter/Rewriter.v}.
The functions \verb|reify| and \verb|reflect| are defined in \texttt{rewriter/src/Rewriter/Rewriter/Rewriter.v} and share names with the functions in the paper.
The actual code is somewhat more complicated than the version presented in the paper, due to needing to deal with converting well-typed-by-construction expressions to dynamically typed expressions for use in decision-tree evaluation and also due to the need to support early partial evaluation against a concrete decision tree.
Thus the version of \verb|reflect| that actually invokes rewriting at base types is a separate definition \verb|assemble_identifier_rewriters|, while \verb|reify| invokes a version of \verb|reflect| (named \verb|reflect|) that does not call rewriting.
The function named \texttt{reduce} is what we call \verb|rewrite_bottomup| in the code; the name \verb|Rewrite| is shared between this paper and the code.
Note that we eventually instantiate the argument \verb|rewrite_head| of \verb|rewrite_bottomup| with a partially evaluated version of the definition named \verb|assemble_identifier_rewriters|.
Note also that we use fuel to support \verb|do_again|, and this is used in the definition \verb|repeat_rewrite| that calls \verb|rewrite_bottomup|.

The correctness proofs are \verb|InterpRewrite| in the Coq source file \texttt{rewriter/src/Rewriter/Rewriter/InterpProofs.v} and \verb|Wf_Rewrite| in \texttt{rewriter/src/Rewriter/Rewriter/Wf.v}.

Packages containing rewriters and their correctness theorems are in the record \verb|VerifiedRewriter| in \texttt{rewriter/src/Rewriter/Rewriter/ProofsCommon.v};
a package of this type is then passed to the tactic \verb|Rewrite_for_gen| from \texttt{rewriter/src/Rewriter/Rewriter/AllTactics.v} to perform the actual rewriting.
The correspondence of the code to the various steps in rewriting is described in the second list of \autoref{sec:code:nine-steps}.

\subsubsection{Code from \autocommanameref{sec:under-lets}}

To run the P-256 example in the copy of Fiat Cryptography attached as a code supplement, after building the library, run the code
\begin{minted}[fontsize=\small]{coq}
Require Import Crypto.Rewriter.PerfTesting.Core.
Require Import Crypto.Util.Option.

Import WordByWordMontgomery.
Import Core.RuntimeDefinitions.

Definition p : params
  := Eval compute in invert_Some (of_string "2^256-2^224+2^192+2^96-1" 64).

Goal True.
  (* Successful run: *)
  Time let v := (eval cbv
    -[Let_In
      runtime_nth_default
      runtime_add runtime_sub runtime_mul runtime_opp runtime_div runtime_modulo
      RT_Z.add_get_carry_full RT_Z.add_with_get_carry_full RT_Z.mul_split]
    in (GallinaDefOf p)) in
    idtac.
  (* Unsuccessful OOM run: *)
  Time let v := (eval cbv
    -[(*Let_In*)
      runtime_nth_default
      runtime_add runtime_sub runtime_mul runtime_opp runtime_div runtime_modulo
      RT_Z.add_get_carry_full RT_Z.add_with_get_carry_full RT_Z.mul_split]
    in (GallinaDefOf p)) in
    idtac.
Abort.
\end{minted}

The \verb|UnderLets| monad is defined in the file \texttt{rewriter/src/Rewriter/Language/UnderLets.v}.

The definitions \verb|nbeT'|, \verb|nbeT|, and \verb|nbeT_with_lets| are in \texttt{rewriter/src/Rewriter/Rewriter/Rewriter.v} and are named \verb|value'|, \verb|value|, and \verb|value_with_lets|, respectively.

\subsubsection{Code from \autocommanameref{sec:side-conditions}}

The ``variant of pattern variable that only matches constants'' is actually special support for the reification of \verb|ident.literal| (defined in the module \verb|RewriteRuleNotations| in \texttt{rewriter/src/Rewriter/Language/Pre.v}) threaded throughout the rewriter.
The apostrophe notation \verb|'| is also introduced in the module \verb|RewriteRuleNotations| in \texttt{rewriter/src/Rewriter/Language/Pre.v}.
The support for side conditions is handled by permitting rewrite-rule-replacement expressions to return \verb|option expr| instead of \verb|expr|, allowing the function \verb|expr_to_pattern_and_replacement| in the file \texttt{rewriter/src/Rewriter/Rewriter/Reify.v} to fold the side conditions into a choice of whether to return \verb|Some| or \verb|None|.

\subsubsection{Code from \autocommanameref{sec:abs-int}}

The abstract-interpretation pass is defined in \texttt{fiat-crypto/src/AbstractInterpretation/}, and the rewrite rules handling abstract-interpretation results are the Gallina definitions \verb|arith_with_casts_rewrite_rulesT|, as well as \verb|strip_literal_casts_rewrite_rulesT|, as well as \verb|fancy_with_casts_rewrite_rulesT|, and finally as well as \verb|mul_split_rewrite_rulesT|, all defined in \texttt{fiat-crypto/src/Rewriter/Rules.v}.

The \verb|clip| function is the definition \verb|ident.cast| in \texttt{fiat-crypto/src/Language/PreExtra.v}.

\subsubsection{Code from \autocommanameref{sec:implementation-and-usage}} \label{sec:code-from-implementation-and-usage}

The \Ltac{} hooks for extending the preprocessing of eliminators are \mintinline{coq}{reify_preprocess_extra} and \mintinline{coq}{reify_ident_preprocess_extra} in a submodule of \texttt{rewriter/src/Rewriter/Language/PreCommon.v}.
These hooks are called by \mintinline{coq}{reify_preprocess} and \mintinline{coq}{reify_ident_preprocess} in a submodule of \texttt{rewriter/src/Rewriter/Language/Language.v}.
Some recursion lemmas for use with these tactics are defined in the \verb|Thunked| module in \texttt{fiat-crypto/src/Language/PreExtra.v}.
These tactics are overridden in the file \texttt{fiat-crypto/src/Language/IdentifierParameters.v}.

The typeclass associated to \mintinline{coq}{eval_rect} ({c.f.} \autoref{sec:code:eval-rect}) is \mintinline{coq}{rules_proofs_for_eager_type} defined in \texttt{rewriter/src/Rewriter/Language/Pre.v}.
The instances we provide by default are defined in a submodule of \texttt{src/Rewriter/Language/PreLemmas.v}.

The hard-coding of the eliminators for use with \mintinline{coq}{ident.eagerly} ({c.f.} \autoref{sec:code:ident.eagerly}) is done in the tactics \mintinline{coq}{reify_ident_preprocess} and \mintinline{coq}{rewrite_interp_eager} in \texttt{rewriter/src/Rewriter/Language/Language.v}, in the inductive type \mintinline{coq}{restricted_ident} and the typeclass \mintinline{coq}{BuildEagerIdentT} in \texttt{rewriter/src/Rewriter/Language/Language.v}, and in the \Ltac{} tactic \mintinline{coq}{handle_reified_rewrite_rules_interp} defined in the file \texttt{rewriter/src/Rewriter/Rewriter/ProofsCommonTactics.v}.

The \mintinline{coq}{Let_In} constant is defined in \texttt{rewriter/src/Rewriter/Util/LetIn.v}.

\subsection{Code from \autocommanameref{sec:evaluation}}

\subsubsection{Code from \autocommanameref{sec:micro}}

This code is found in the files in \texttt{rewriter/src/Rewriter/Rewriter/Examples/}.
We ran the microbenchmarks using the code in \texttt{rewriter/src/Rewriter/Rewriter/Examples/PerfTesting/Harness.v} together with some \texttt{Makefile} cleverness.

The code from \autocommanameref{sec:micro:Plus0Tree} can be found in \texttt{Plus0Tree.v}.

The code from \autocommanameref{sec:micro:UnderLetsPlus0} can be found in \texttt{UnderLetsPlus0.v}.

The code used for the performance investigation mentioned in \autocommanameref{sec:micro:setoid-rewrite-bottlenecks-lite} and detailed in \autoref{sec:setoid-rewrite-bottlenecks} is not part of the framework we are presenting, and thus not in the supplement.

The code from \autocommanameref{sec:micro:LiftLetsMap} can be found in \texttt{LiftLetsMap.v}.

The code from \autocommanameref{sec:micro:Eratosthenes} can be found in \texttt{SieveOfEratosthenes.v}.

\subsubsection{Code from \autocommanameref{sec:macro}}

The rewrite rules are defined in \texttt{fiat-crypto/src/Rewriter/Rules.v} and proven in the file \texttt{fiat-crypto/src/Rewriter/RulesProofs.v}.
They are turned into rewriters in the various files in \texttt{fiat-crypto/src/Rewriter/Passes/}.
The shared inductives and definitions are defined in the Coq source file \texttt{fiat-crypto/src/Language/IdentifiersBasicGENERATED.v}, the Coq source file \texttt{fiat-crypto/src/Language/IdentifiersGENERATED.v}, and finally also the Coq source file \texttt{fiat-crypto/src/Language/IdentifiersGENERATEDProofs.v}.
Note that we invoke the subtactics of the \mintinline{coq}{Make} command manually to increase parallelism in the build and to allow a shared language across multiple rewriter packages.

\end{document}

### 2020-07-30 papers/source/category-coq-experience/keep_only_good_urls.py

@@ -1,0 +1,35 @@
+#!/usr/bin/env python
+from __future__ import with_statement
+import sys, re
+
+VALID_URLS = ('https://github.com',
+              'http://coq.inria.fr/pylons',
+              'http://coq.inria.fr',
+              'http://www.cs.berkeley.edu/~megacz',
+              'http://mattam.org/repos',
+              'http://ncatlab.org',
+              'http://homotopytypetheory.org/2011/',
+              'http://mathoverflow.net/')
+
+def is_doi_line(line):
+    return re.match(r'^[^a-zA-Z]+doi = .+$', line) is not None
+
+def is_url_line(line):
+    return re.match(r'^[^a-zA-Z]+url = .+$', line) is not None \
+        or re.match(r'^[^a-zA-Z]+eprint = .+$', line) is not None
+
+def url_line_is_valid(line):
+    for url in VALID_URLS:
+        if re.match(r'^[^a-zA-Z]+url = {' + url + '.+$', line) is not None:
+            return True
+    return False
+
+def fixline(line):
+    return line.replace(r'@MASTERSTHESIS{spiwackverified,', r'@THESIS{spiwackverified,')
+
+if __name__ == '__main__':
+    with open(sys.argv[1], 'r') as f:
+        contents = f.readlines()
+    for i in contents:
+        if (not (is_url_line(i) or is_doi_line(i))) or url_line_is_valid(i):
+            print(fixline(i.strip('\n')))

### 2020-09-14 presentations/coqpl-2021/reification-by-type-inference.tex
Update submodules and add CoqPL 2021

\documentclass{article}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{verbatim}
\newcommand{\Ltac}{\texorpdfstring{\ensuremath{\mathcal{L}_{\text{\textit{tac}}}}}{Ltac}}
\title{A Limited Case for Reification by Type Inference}
\author{Jason Gross}
\date{CoqPL 2021}
% Key Words: reification, reflection, type inference
\begin{document}
\maketitle
\begin{abstract}
%\subsection*{Typical Reification}
Proof by reflection is a common and well-studied automation tool.
Reification---generally written using \Ltac, OCaml, typeclasses, or canonical structures---is the means by which a structured representation is derived from an unstructured representation.
The reflective automation then operates on the structured representation, relying on an interpretation or denotation function to justify a correspondence between the structured and unstructured representations.

A couple of years ago, I presented a trick for blazing fast reification in two lines of \Ltac---using the \texttt{pattern} tactic---which I termed reification by parametricity.
While I still advocate for parametricity as the preferred method of domain-specific reification, I would like to present here yet another method.

While reification typically requires meta-programming features, I was surprised and delighted to discover that, in some restricted cases, reification can be performed entirely by a combination of the notation system and type inference.
In some sense, this is trivial: by redefining the basic syntactic notations, a term can be ``reified'' merely by writing the same symbols in another scope.
In another sense, though, this trick is quite surprising: we use the notation system merely to insert ``reify here'' functions at every atom, and the reification itself is in fact performed by type inference.
My hope is that the audience will walk away with this new trick in their toolbox, and that some day some problem will come along demanding a slight generalization of this trick, and that generalization will be new and interesting in its own right.
This, after all, is how reification by parametricity was discovered.

I propose to present the one example I have for this trick: reifying the type structure of a function in a way that allows manipulations of the arguments, such as uncurrying, reassociation of the uncurried structure, and reordering.
I will present the simple code for this example in detail.
My goal will be that the audience understand completely how it works, why it works, and how it might be used elsewhere.

We reproduce the code here without explanation on the following page.

\clearpage

\verbatiminput{reification_by_type_inference_tex.v}
\end{abstract}
\end{document}

### 2020-09-14 update-submodules.sh

@@ -1,0 +1,3 @@
+#!/usr/bin/env bash
+
+git submodule foreach --recursive bash -c 'if [[ "$(pwd)" != *"/coq" ]]; then echo "$(pwd)" && git pull origin master; fi'

### 2021-08-15 presentations/coq-workshop-2018/coq-workshop-proposal-notations.tex
Copy over built files

\documentclass{article}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{newunicodechar}
\newunicodechar{λ}{\ensuremath{\lambda}}
\title{Teaching Your Rooster to Crow in C}
\author{Jason Gross}
\date{Coq Workshop 2018}
% Keywords: Coq, notation, pretty printing, DSL
\begin{document}
\maketitle
\begin{abstract}
  Coq's notation system is both extremely powerful and confusingly ad-hoc.
  While powerful enough to pretty-print abstract syntax trees in most domain-specific languages, how to do so does not seem to be common knowledge.
  Typical questions arising from such an endeavor might include ``How do I pick notation levels?'', ``Why are these notations clashing?'', ``Which things should be marked as symbols?'', ``How do I use boxes in \texttt{format}?'', and ``How do I get parentheses to show up (only) where I want them to?''
  This interactive presentation aims to serve as a guide to these questions and more, by demonstrating and explaining how to pretty-print subsets of C using only Coq's \texttt{Notation} mechanism.
\end{abstract}

My goal in this presentation is to give the audience enough familiarity with \texttt{Notation}s to use them in their own projects for pretty-printing DSLs, or at least to know what sorts of experiments to execute to discover what they want.
As such, I intend this to be a very interactive presentation.  I plan to explain the basics of using \texttt{Notation}s, a selection of features, and present a number of tricks for getting things to work.
The audience will be encouraged to ask questions and propose things they'd like to see me do with notations, and the presentation will take place in a prepared ProofGeneral buffer.

I expect to cover the following things, in roughly the following order, as time permits:
\begin{itemize}
  \item Syntax of notations (e.g., \texttt{Infix "+" := add.}; \texttt{Notation "a + b" := (add a b).})
    \begin{itemize}
        \item Notation bodies must be parenthesized
    \end{itemize}
  \item How to use levels and associativity, how to pick levels, how to discover levels of existing notations (\texttt{Print Grammar constr})
  \item Controlling display: \texttt{format}, extra spaces, quoting symbols, boxes in \texttt{format} for display of indentation, newlines, and whitespace
  \item Recursive notations: \texttt{..} for pair- and $\lambda$-like things
  \item \texttt{Reserved Notation} and \texttt{only printing}, useful for ensuring that notations don't conflict
  \item Using single-variable only-printing notations for hiding identifiers
  \item Abbreviations vs notations (what it means when your \texttt{Notation} doesn't have quotes, and why you might want this)
  \item Notation scopes, useful for overloading notations and selecting the right one automatically (\texttt{Bind Scope}, \texttt{Delimit Scope}, \texttt{Open Scope})
  \item Using binders for destructuring pairs
  \item Parenthesizing notations: when to use parentheses, and when to use levels
  \item Deliberate ordering of notations: how to express wildcard and negation in notation matching
\end{itemize}

\paragraph{Examples}
By the end of the presentation, I intend for the audience to understand, or at least feel comfortable experimenting with, notations that look like any of the following:
\begin{verbatim}
Infix "+" := Add.
Notation "T x = A ; b" := (LetIn (tx:=T) A (fun x => b))
  (at level 200, b at level 200, format "T  x  =  A ; '//' b").
Notation "T x = A ; 'return' ( b0 , b1 , .. , b2 )"
   := (LetIn (tx:=T) A (fun x => Pair .. (Pair b0%expr b1%expr) .. b2%expr))
  (at level 200, format "T  x  =  A ; '//' 'return'  ( b0 ,  b1 ,  .. ,  b2 )").
Notation "x" := (Var x) (only printing).
Notation "'slet' x .. y := A 'in' b"
  := (LetIn A%expr (fun x => .. (fun y => b%expr) .. )) : expr_scope.
Notation "( x * y )" := (Op (Mul _ _ _) (Pair x y))
  (format "( x  *  y )") : expr_scope.
Notation "( x * '(uint8_t)' y )"
 := (Op (Mul (TWord _) (TWord (S _)) (TWord 1)) (Pair x (Var y)))
  (at level 40, y at level 9, format "( x  *  '(uint8_t)' y )") : expr_scope.
\end{verbatim}

\paragraph{Motivating Example} A sort of ``motivating example'' driving this presentation, an answer to the question of ``what can we do with this?'', is turning a syntax tree that looks like
\begin{verbatim}
Abs (fun '(x8, x9, x7, x5, x3, (x16, x17, x15, x13, x11)) =>
LetIn (Op (Mul (TWord 6) (TWord 6) (TWord 7)) (Pair (Var x3) (Var x11)))
(fun x18 : var (TWord 7) => ...
\end{verbatim}
into
\begin{verbatim}
 λ '(x8, x9, x7, x5, x3, (x16, x17, x15, x13, x11))%core,
 uint128_t x18 = ((uint128_t)x3 * x11);
 ...
\end{verbatim}
\end{document}

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/coq-makefile/timing/run.sh

@@ -1,0 +1,104 @@
+#!/usr/bin/env bash
+
+#set -x
+set -e
+
+. ../template/path-init.sh
+
+cd precomputed-time-tests
+./run.sh || exit $?
+
+cd ../error
+coq_makefile -f _CoqProject -o Makefile
+make cleanall
+if make pretty-timed TGTS="all" -j1; then
+    echo "Error: make pretty-timed should have failed"
+    exit 1
+fi
+
+cd ../aggregate
+coq_makefile -f _CoqProject -o Makefile
+make cleanall
+make pretty-timed TGTS="all" -j1 || exit $?
+
+cd ../before
+coq_makefile -f _CoqProject -o Makefile
+make cleanall
+make make-pretty-timed-before TGTS="all" -j1 || exit $?
+
+cd ../after
+coq_makefile -f _CoqProject -o Makefile
+make cleanall
+make make-pretty-timed-after TGTS="all" -j1 || exit $?
+rm -f time-of-build-before.log
+make print-pretty-timed-diff TIMING_SORT_BY=diff TIME_OF_BUILD_BEFORE_FILE=../before/time-of-build-before.log
+cp ../before/time-of-build-before.log ./
+make print-pretty-timed-diff TIMING_SORT_BY=diff || exit $?
+
+INFINITY="∞"
+INFINITY_REPLACEMENT="+.%" # assume that if the before time is zero, we expected the time to increase
+
+TO_SED_IN_BOTH=(
+    -e s"/${INFINITY}/${INFINITY_REPLACEMENT}/g" # Whether or not something shows up as ∞ depends on whether a time registers as 0.s or as 0.001s, so we can't rely on this being consistent
+    -e s':|\s*N/A\s*$:| '"${INFINITY_REPLACEMENT}"':g' # Whether or not something shows up as N/A depends on whether a time registers as 0.s or as 0.001s, so we can't rely on this being consistent
+    -e s'/ *$//g' # the number of trailing spaces depends on how many digits percentages end up being; since this varies across runs, we remove trailing spaces
+    -e s'/[0-9]*\.[0-9]*//g' # the precise timing numbers vary, so we strip them out
+    -e s'/^-*$/------/g' # When none of the numbers get over 100 (or 1000, in per-file), the width of the table is different, so we normalize the number of dashes for table separators
+)
+
+TO_SED_IN_PER_FILE=(
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/misc/deps-order.sh

@@ -1,0 +1,20 @@
+# Check that both coqdep and coqtop/coqc supports -R
+# Check that both coqdep and coqtop/coqc takes the later -R
+# See bugs 2242, 2337, 2339
+rm -f misc/deps/lib/*.vo misc/deps/client/*.vo
+tmpoutput=`mktemp /tmp/coqcheck.XXXXXX`
+$coqdep -R misc/deps/lib lib -R misc/deps/client client misc/deps/client/bar.v 2>&1 | head -n 1 > $tmpoutput
+diff -u --strip-trailing-cr misc/deps/deps.out $tmpoutput 2>&1
+R=$?
+times
+$coqc -R misc/deps/lib lib misc/deps/lib/foo.v 2>&1
+$coqc -R misc/deps/lib lib -R misc/deps/client client misc/deps/client/foo.v 2>&1
+$coqtop -R misc/deps/lib lib -R misc/deps/client client -load-vernac-source misc/deps/client/bar.v 2>&1
+S=$?
+if [ $R = 0 -a $S = 0 ]; then
+    printf "coqdep and coqtop agree\n"
+    exit 0
+else
+    printf "coqdep and coqtop disagree\n"
+    exit 1
+fi

### 2021-08-15 update-submodules.sh

@@ -1,0 +1,3 @@
+#!/usr/bin/env bash
+
+git submodule foreach --recursive bash -c 'if [[ "$(pwd)" != */*"coq" ]] && [[ "$(pwd)" != *"/lob-paper" ]]; then echo "$(pwd)" && git pull origin master; fi'

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/misc/printers.sh
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@20b0f0bcf3b2dec413042a845be83385ca63dd06 🚀

@@ -1,0 +1,3 @@
+printf "Drop. #use\"include\";; #quit;;\n" | $coqtopbyte 2>&1 | egrep "Error|Unbound"
+if [ $? = 0 ]; then exit 1; else exit 0; fi
+

### 2021-08-15 papers/source/lob-paper/review-responses.md
Add more deploy

Thank you all for the feedback; we really appreciate it.

It's now clear that one of the most subtle parts of Lӧb's theorem is distinguishing between semantics, syntax, quoted syntax, and doubly-quoted syntax, and that our paper does not make it easy enough to make these distinctions; we will edit the paper to do far more hand-holding here, and will drop the cute notations that blur this distinction.  The points about cute notations and magic tricks, and about consistently using the same words for the same concepts, are well-taken, and we will fix these things.

The suggestion to begin with the Y combinator is a good one; thank you reviewers B and C!

We now realize that the traditional presentation of Curry-Howard is classical and identifies provability with truth; to extend it neatly to talking about provability logic in a computational manner, we've presented it in from within the lens of Brouwer’s intuitionism; we will clarify our use of the word "truth".

Reviewer-specific responses, followed by an appendix of reviewer-specific responses that didn't fit in the first 500 words.

========================

Reviewer A:
------------------------

Indeed, it seems impossible at first glance to prove Lӧb's theorem without some sort of `QUINE` primitive.  But it turns out it is possible!  Section 10 sketches how to do this, and we flesh it out for those interested in the supplemental material.

Your point about cute names is well-taken, and the use of capitalization to distinguish the syntax for types from the syntax for terms, and using names rather than symbols to distinguish syntax from semantics, are both great ideas.

========================

Reviewer B:
------------------------

- The Y combinator is unsound in a proof assistant; non-termination leads to proofs of false.  Indeed, Lӧb's theorem is nothing but a variant of the Y combinator with (un)quoting, *which implicitly forbids infinite loops*.  We will make clear that this paper is implementing this variant of the Y combinator in varying degrees of detail, in a way that is terminating-by-construction.

- The type `□ (‘□’ X ‘→’ X) → □ X` corresponds to the deduction rule "from `⊢ □ X → X`, we deduce `⊢ X`".  An alternative would be "`⊢ □ (□ X → X) → □ X`".

========================

Reviewer C:
------------------------

If we understand correctly, you want to see us separate the proof of Lӧb's theorem from its semantics, using a variant of S4.

This is a good idea, and we may replace the section that assumes a quine with this formalization (see the appendix to this note for a 40 line version).

- We assumed that for those unfamiliar with encodings of type theory in type theory, working with strings and string replacement and quotation in Python would be more familiar than working with quotation and substitution on abstract syntax trees.  We acknowledge that it might be clearer to present the quotation operation from scratch, rather than relying on intuition from Python.  (Note that internally implementing quotation on well-typed syntax trees is, as far as we are aware, still an open problem, implementation-wise.)

========================
Appendix
========================

========================

Reviewer A:
------------------------

I'm a bit confused by your second bit of code; the type of `LÖB` should be `EXP → EXP`, and `löb` should have type

```
löb : {T : TYP} {interp : EXP} → interp ∶ (ARR (BOX T) T) → (LÖB interp ∶ T)
```

Your hole can be filled with `Σ EXP (λ e → e ∶ T)`, and the denotation function for `EXP` is the same as in your first block of code.

What you are doing here, I think, is separating raw terms from their typing judgments.  What we do is remove the `LÖB` primitive and reimplement it using a `QUINE` primitive, and then sketch how that can be removed entirely.

We'd be curious to see these references in the context of quantum computation; I've not seen any physical interpretation of the modal □ operator.

========================

Reviewer B:
------------------------

- The proof of Lӧb's theorem uses self-reference in an essential way; it results from analyzing the self-referential sentence "If this sentence is provable, then A".  We explain that this is used to prove Lӧb's theorem in the second-to-last paragraph in the introduction, containing "This, in a nutshell, is Löb’s theorem."  We'd very much appreciate suggestions on what made this unclear.

- Thank you for the suggestions about unifying the synonyms and clarifying types where they are not obvious; we'll do this.

- The relation between the program `Lob(X)` on page 2 and `□(□A → A) → □A` is that `Lob(X)` is the key ingredient of the proof of Löb’s theorem; it is the type corresponding to the sentence "If this sentence is provable, then A" which is mentioned above.

- Interpreting `□A` as `1 → A` does not fix the termination problem.  It is possible to have a total language with a Löbian combinator `(□A → A) → A` if we interpret `□A` as syntax, but if we interpret it as `1 → A`, we get `((1 → A) → A) → A`, which is just the Y combinator.  Said another way, `1 → A` suspends computation at *runtime*, but to make Lӧb's theorem work, we must suspend computation at *compile-time*.

- Regarding `repr`, we can *add* a level of quotation to anything that's already quoted.  We plan to make it easier to see at a glance the distinction between semantics, (singly-quoted) syntax, and (doubly) quoted syntax.

- Regarding "There is no syntactic proof of absurdity", we again need to make clearer the distinction between syntax and semantics.  Agda has no proof of absurdity.  If we build our syntax in a stupid way, we might permit a syntactic proof of absurdity in the object language (which is not the full extent of Agda, and which we have no guarantee is identical to a subset of Agda); we must prove that it is impossible to derive contradictions in our object language.

- Regarding the footnote, `□(□X → X) → □X` is a valid expression of the type of Lӧb's theorem, but it's not a valid type for a constructor of an inductive datatype `□`, again for termination (positivity) reasons.

- Thank you for the feedback on the Prisoner's dilemma section; we will add a bit more exposition to make it more self-contained.

- Agda does not have a notion of truth distinct from provability; this is what we meant by "banish [the word] 'truth'".  We will add more hand-holding here; the reason to talk about "provability" is that it is extremely confusing to say "P is false, and 'P is false' is false" while it is only mildly confusing to say "we can prove that P is absurd, but we can also prove that proving "WE CAN PROVE THAT P IS ABSURD" is absurd".

========================

Reviewer C:
------------------------

The formalization and semantics of S4, specializing the quine/fixed-point theorem to the Lӧbian sentence.  We may use this to replace the formalization based on the Quine primitive (obviously with added explanation of the Lӧb′s-Theorem proof term).

```
data TYP : Set where
  ARR : TYP → TYP → TYP -- the type of implications, or function types
  BOX : TYP → TYP -- the modal □ operator, denoted to TERM
  LӦB-SENTENCE : TYP → TYP -- the Lӧbian sentence "If this sentence is provable, then A"
                           -- this is the modal fixpoint of λ Ψ. □ Ψ → A

data TERM : TYP → Set where
  nec : {A : TYP} → TERM A → TERM (BOX A)                                             -- from A, we deduce □ A
  distr : {A B : TYP} → TERM (ARR (BOX (ARR A B)) (ARR (BOX A) (BOX B)))              -- we deduce □ (A → B) → □ A → □ B
  s4 : {A : TYP} → TERM (ARR (BOX A) (BOX (BOX A)))                                   -- we deduce □ A →  □ □ A
  app : {A B : TYP} → TERM (ARR A B) → TERM A → TERM B                                -- from A → B, and A, we deduce B
  lӧb→ : {A : TYP} → TERM (ARR (LӦB-SENTENCE A) (ARR (BOX (LӦB-SENTENCE A)) A))       -- LӦB-SENTENCE A is Ψ such that Ψ → (□ Ψ → A)
  lӧb← : {A : TYP} → TERM (ARR (ARR (BOX (LӦB-SENTENCE A)) A) (LӦB-SENTENCE A))       -- LӦB-SENTENCE A is Ψ such that (□ Ψ → A) → Ψ
  compose : {A B C : TYP} → TERM (ARR A B) → TERM (ARR B C) → TERM (ARR A C)          -- from A → B and B → C, we deduce A → C
  compose2 : {A B C : TYP} → TERM (ARR A B) → TERM (ARR A (ARR B C)) → TERM (ARR A C) -- from A → B and A → B → C, we deduce A → C

Lӧb′s-Theorem : {A : TYP} → TERM (ARR (BOX A) A) → TERM A -- from □ A → A, we deduce A
Lӧb′s-Theorem {A} refl-rule = app prog (nec (app lӧb← prog))
  where prog : TERM (ARR (BOX (LӦB-SENTENCE A)) A)
        prog = compose (compose2 s4 (compose (app distr (nec lӧb→)) distr)) refl-rule

⟦_⟧ᵀ : TYP → Set
⟦ ARR A B ⟧ᵀ = ⟦ A ⟧ᵀ → ⟦ B ⟧ᵀ
⟦ BOX T ⟧ᵀ = TERM T
⟦ LӦB-SENTENCE A ⟧ᵀ = TERM (LӦB-SENTENCE A) → ⟦ A ⟧ᵀ

⟦_⟧ᵗ : {T : TYP} → TERM T → ⟦ T ⟧ᵀ
⟦ nec e ⟧ᵗ = e
⟦ distr ⟧ᵗ box-a-b box-a = app box-a-b box-a
⟦ s4 ⟧ᵗ = nec
⟦ app f x ⟧ᵗ = ⟦ f ⟧ᵗ ⟦ x ⟧ᵗ
⟦ lӧb→ ⟧ᵗ = λ x → x -- this implication is true because on denotation, the two are judgmentally equal
⟦ lӧb← ⟧ᵗ = λ x → x -- this implication is true because on denotation, the two are judgmentally equal
⟦ compose f g ⟧ᵗ = λ x → ⟦ g ⟧ᵗ (⟦ f ⟧ᵗ x)
⟦ compose2 f g ⟧ᵗ = λ x → ⟦ g ⟧ᵗ x (⟦ f ⟧ᵗ x)
```

- Section 3 is intended to be a crash course in formalizing the syntax and semantics of type theory for those who have not seen it before.  Perhaps your point is that it does not provide enough explanation for those unfamiliar with it, that it is (intentionally) irrelevant for those familiar with formalizing type theory in type theory, and that what little content in the section falls into neither category should be merged into the other sections, to improve flow.  This is a good point.

### 2021-08-15 presentations/coq-workshop-2014/coq-workshop-proposal-tactics-in-terms.tex
Add more deploy

@@
 \usepackage{fancyhdr}
-\title{Presentation of Three Neat Tricks in Coq 8.5}
+\title{Presentation Proposal for Three Neat Tricks in Coq 8.5}
 \author{Jason Gross}
@@
             pdfkeywords={Coq 8.5, Ltac, Gallina, binders, notations, tactics, recursion, recursive tactics, typeclasses, universe polymorphism, universes}]{hyperref}
-\chead{Presentation of Three Neat Tricks in Coq 8.5 \\ Jason Gross \\ ITP 2014: 6th Coq Workshop}
+\chead{Presentation Proposal for Three Neat Tricks in Coq 8.5 \\ Jason Gross \\ ITP 2014: 6th Coq Workshop}
 \thispagestyle{fancy}

### 2021-08-15 presentations/coq-workshop-2014-coq/dev/make-installer-win32.sh

@@ -1,0 +1,22 @@
+#!/bin/sh
+
+set -e
+
+NSIS="$BASE/NSIS/makensis"
+ZIP=_make.zip
+URL1=http://sourceforge.net/projects/gnuwin32/files/make/3.81/make-3.81-bin.zip/download
+URL2=http://sourceforge.net/projects/gnuwin32/files/make/3.81/make-3.81-dep.zip/download
+
+[ -e config/Makefile ] || ./configure -debug -prefix ./ -with-doc no
+make -j2 
+if [ ! -e bin/make.exe ]; then
+  wget -O $ZIP $URL1 && 7z x $ZIP "bin/*"
+  wget -O $ZIP $URL2 && 7z x $ZIP "bin/*"
+  rm -rf $ZIP
+fi
+VERSION=`grep ^VERSION= config/Makefile | cut -d = -f 2`
+cd dev/nsis
+"$NSIS" -DVERSION=$VERSION -DGTK_RUNTIME="`cygpath -w $BASE`" coq.nsi
+echo Installer:
+ls -h $PWD/*exe
+cd ../..

### 2021-08-15 presentations/coq-workshop-2014-coq/dev/make-macos-dmg.sh

@@ -1,0 +1,31 @@
+#!/bin/bash
+
+# Fail on first error
+set -e
+
+# Configuration setup
+eval `opam config env`
+make distclean
+OUTDIR=$PWD/_install
+DMGDIR=$PWD/_dmg
+./configure -debug -prefix $OUTDIR
+VERSION=$(sed -n -e '/^let coq_version/ s/^[^"]*"\([^"]*\)"$/\1/p' configure.ml)
+APP=bin/CoqIDE_${VERSION}.app
+
+# Create a .app file with CoqIDE
+~/.local/bin/jhbuild run make -j -l2 $APP
+
+# Build Coq and run test-suite
+make && make check
+
+# Add Coq to the .app file
+make OLDROOT=$OUTDIR COQINSTALLPREFIX=$APP/Contents/Resources/ install-coq install-ide-toploop
+
+# Sign the .app file
+codesign -f -s - $APP
+
+# Create the dmg bundle
+mkdir -p $DMGDIR
+ln -sf /Applications $DMGDIR/Applications
+cp -r $APP $DMGDIR
+hdiutil create -imagekey zlib-level=9 -volname CoqIDE_$VERSION -srcfolder $DMGDIR -ov -format UDZO CoqIDE_$VERSION.dmg

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/build/windows/difftar-folder.sh

@@ -1,0 +1,89 @@
+#!/bin/bash
+
+###################### COPYRIGHT/COPYLEFT ######################
+
+# (C) 2016 Intel Deutschland GmbH
+# Author: Michael Soegtrop
+#
+# Released to the public by Intel under the
+# GNU Lesser General Public License Version 2.1 or later
+# See https://www.gnu.org/licenses/old-licenses/lgpl-2.1.html
+
+###################### DIFF A TAR FILE AND A FOLDER ######################
+
+set -o nounset
+
+# Print usage
+
+if [ "$#" -lt 2 ] ; then
+  echo 'Diff a tar (or compressed tar) file with a folder'
+  echo 'difftar-folder.sh <tarfile> <folder> [strip]'
+  echo '<tarfile> is the name of the tar file do diff with (required)'
+  echo '<folder> is the name of the folder to diff with (required)'
+  echo '<strip> is the number of path components to strip from tar file (default is 0)'
+  echo 'All files in the tar file must have at least <strip> path components.'
+  echo 'This also adds new files from folder.new, if folder.new exists'
+  exit 1
+fi
+
+# Parse parameters
+
+tarfile=$1
+folder=$2
+
+if [ "$#" -ge 3 ] ; then
+  strip=$3
+else
+  strip=0
+fi
+
+# Get path prefix if --strip is used
+
+if [ "$strip" -gt 0 ] ; then
+  # Get the path/name of the first file from teh tar and extract the first $strip path components
+  # This assumes that the first file in the tar file has at least $strip many path components
+  prefix=$(tar -t -f $tarfile | head -1 | cut -d / -f -$strip)/
+else
+  prefix=
+fi
+
+# Original folder
+
+orig=$folder.orig
+mkdir -p "$orig"
+
+# New amd empty filefolder
+
+new=$folder.new
+empty=$folder.empty
+mkdir -p "$empty"
+
+# Print information (this is ignored by patch)
+
+echo diff/patch file created on $(date) with:
+echo difftar-folder.sh $@
+echo TARFILE=    $tarfile
+echo FOLDER=     $folder
+echo TARSTRIP=   $strip
+echo TARPREFIX=  $prefix
+echo ORIGFOLDER= $orig
+
+# Make sure tar uses english output (for Mod time differs)
+export LC_ALL=C
+
+# Search all files with a deviating modification time using tar --diff
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-basic-overlay.sh

@@ -1,0 +1,152 @@
+#!/usr/bin/env bash
+
+# This is the basic overlay set for repositories in the CI.
+
+# Maybe we should just use Ruby to have real objects...
+
+# : "${foo:=bar}" sets foo to "bar" if it is unset or null
+
+########################################################################
+# MathComp
+########################################################################
+: "${mathcomp_CI_BRANCH:=master}"
+: "${mathcomp_CI_GITURL:=https://github.com/math-comp/math-comp.git}"
+
+########################################################################
+# UniMath
+########################################################################
+: "${UniMath_CI_BRANCH:=master}"
+: "${UniMath_CI_GITURL:=https://github.com/UniMath/UniMath.git}"
+
+########################################################################
+# Unicoq + Metacoq
+########################################################################
+: "${unicoq_CI_BRANCH:=master}"
+: "${unicoq_CI_GITURL:=https://github.com/unicoq/unicoq.git}"
+
+: "${metacoq_CI_BRANCH:=master}"
+: "${metacoq_CI_GITURL:=https://github.com/MetaCoq/MetaCoq.git}"
+
+########################################################################
+# Mathclasses + Corn
+########################################################################
+: "${math_classes_CI_BRANCH:=master}"
+: "${math_classes_CI_GITURL:=https://github.com/math-classes/math-classes.git}"
+
+: "${Corn_CI_BRANCH:=master}"
+: "${Corn_CI_GITURL:=https://github.com/c-corn/corn.git}"
+
+########################################################################
+# Iris
+########################################################################
+: "${stdpp_CI_BRANCH:=master}"
+: "${stdpp_CI_GITURL:=https://gitlab.mpi-sws.org/robbertkrebbers/coq-stdpp.git}"
+
+: "${Iris_CI_BRANCH:=master}"
+: "${Iris_CI_GITURL:=https://gitlab.mpi-sws.org/FP/iris-coq.git}"
+
+: "${lambdaRust_CI_BRANCH:=master}"
+: "${lambdaRust_CI_GITURL:=https://gitlab.mpi-sws.org/FP/LambdaRust-coq.git}"
+
+########################################################################
+# HoTT
+########################################################################
+: "${HoTT_CI_BRANCH:=master}"
+: "${HoTT_CI_GITURL:=https://github.com/HoTT/HoTT.git}"
+
+########################################################################
+# Ltac2
+########################################################################
+: "${ltac2_CI_BRANCH:=v8.8}"
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-fiat-crypto.sh

@@ -1,0 +1,11 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+fiat_crypto_CI_DIR=${CI_BUILD_DIR}/fiat-crypto
+
+git_checkout ${fiat_crypto_CI_BRANCH} ${fiat_crypto_CI_GITURL} ${fiat_crypto_CI_DIR}
+( cd ${fiat_crypto_CI_DIR} && git submodule update --init --recursive )
+
+( cd ${fiat_crypto_CI_DIR} && make lite )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-flocq.sh

@@ -1,0 +1,10 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+Flocq_CI_DIR=${CI_BUILD_DIR}/flocq
+
+git_checkout ${Flocq_CI_BRANCH} ${Flocq_CI_GITURL} ${Flocq_CI_DIR}
+
+( cd ${Flocq_CI_DIR} && ./autogen.sh && ./configure && ./remake -j${NJOBS} )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-geocoq.sh

@@ -1,0 +1,12 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+GeoCoq_CI_DIR=${CI_BUILD_DIR}/GeoCoq
+
+git_checkout ${GeoCoq_CI_BRANCH} ${GeoCoq_CI_GITURL} ${GeoCoq_CI_DIR}
+
+( cd ${GeoCoq_CI_DIR}                                     && \
+  ./configure-ci.sh                                       && \
+  make )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-tlc.sh

@@ -1,0 +1,10 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+tlc_CI_DIR=${CI_BUILD_DIR}/tlc
+
+git_checkout ${tlc_CI_BRANCH} ${tlc_CI_GITURL} ${tlc_CI_DIR}
+
+( cd ${tlc_CI_DIR} && make )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-unimath.sh

@@ -1,0 +1,14 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+UniMath_CI_DIR=${CI_BUILD_DIR}/UniMath
+
+git_checkout ${UniMath_CI_BRANCH} ${UniMath_CI_GITURL} ${UniMath_CI_DIR}
+
+( cd ${UniMath_CI_DIR}                        && \
+  sed -i.bak '/Folds/d'              Makefile && \
+  sed -i.bak '/HomologicalAlgebra/d' Makefile && \
+  make BUILD_COQ=no )
+

### 2021-08-15 presentations/coq-workshop-2018-coq/doc/sphinx/conf.py

@@ -1,0 +1,414 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+##########################################################################
+##         #   The Coq Proof Assistant / The Coq Development Team       ##
+##  v      #   INRIA, CNRS and contributors - Copyright 1999-2018       ##
+## <O___,, #       (see CREDITS file for the list of authors)           ##
+##   \VV/  ###############################################################
+##    //   #    This file is distributed under the terms of the         ##
+##         #     GNU Lesser General Public License Version 2.1          ##
+##         #     (see LICENSE file for the text of the license)         ##
+##########################################################################
+#
+# Coq documentation build configuration file, created by
+# sphinx-quickstart on Wed May 11 11:23:13 2016.
+#
+# This file is execfile()d with the current directory set to its
+# containing dir.
+#
+# Note that not all possible configuration values are present in this
+# autogenerated file.
+#
+# All configuration values have a default; values that are commented out
+# serve to show the default.
+
+import sys
+import os
+
+# Increase recursion limit for sphinx
+sys.setrecursionlimit(1500)
+
+# If extensions (or modules to document with autodoc) are in another directory,
+# add these directories to sys.path here. If the directory is relative to the
+# documentation root, use os.path.abspath to make it absolute, like shown here.
+sys.path.append(os.path.abspath('../tools/'))
+sys.path.append(os.path.abspath('../../config/'))
+
+import coq_config
+
+# -- General configuration ------------------------------------------------
+
+# If your documentation needs a minimal Sphinx version, state it here.
+#needs_sphinx = '1.0'
+
+# Add any Sphinx extension module names here, as strings. They can be
+# extensions coming with Sphinx (named 'sphinx.ext.*') or your custom
+# ones.
+extensions = [
+    'sphinx.ext.mathjax',
+    'sphinx.ext.todo',
+    'sphinxcontrib.bibtex',
+    'coqrst.coqdomain'
+]
+
+# Add any paths that contain templates here, relative to this directory.
+templates_path = ['_templates']
+
+# The suffix(es) of source filenames.
+# You can specify multiple suffix as a list of string:
+# source_suffix = ['.rst', '.md']
+source_suffix = '.rst'
+
+# The encoding of source files.
+#source_encoding = 'utf-8-sig'
+
+# The master toctree document.
+master_doc = 'index'
+
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/kernel/byterun/coq_memory.c

@@ -1,0 +1,212 @@
+/***********************************************************************/
+/*                                                                     */
+/*                           Coq Compiler                              */
+/*                                                                     */
+/*        Benjamin Gregoire, projets Logical and Cristal               */
+/*                        INRIA Rocquencourt                           */
+/*                                                                     */
+/*                                                                     */
+/***********************************************************************/
+
+#include <stdio.h>
+#include <string.h> 
+#include "coq_gc.h"
+#include "coq_instruct.h"
+#include "coq_fix_code.h"
+#include "coq_memory.h"
+#include "coq_interp.h"
+
+/* stack */
+
+value * coq_stack_low;
+value * coq_stack_high;
+value * coq_stack_threshold;
+asize_t coq_max_stack_size = Coq_max_stack_size;
+/* global_data */
+
+
+value coq_global_data;
+value coq_atom_tbl;
+
+int drawinstr;
+/* interp state */
+
+long coq_saved_sp_offset;
+value * coq_sp;
+/* Some predefined pointer code */ 
+code_t accumulate;
+
+/* functions over global environment */
+
+void coq_stat_free (void * blk)
+{
+  free (blk);
+}
+
+value coq_static_alloc(value size) /* ML */
+{
+  return (value) coq_stat_alloc((asize_t) Long_val(size));
+}
+
+value accumulate_code(value unit) /* ML */
+{
+  return (value) accumulate;
+}
+
+static void (*coq_prev_scan_roots_hook) (scanning_action);
+
+static void coq_scan_roots(scanning_action action)
+{
+  register value * i;
+  /* Scan the global variables */
+  (*action)(coq_global_data, &coq_global_data);
+  (*action)(coq_atom_tbl, &coq_atom_tbl);
+  /* Scan the stack */
+  for (i = coq_sp; i < coq_stack_high; i++) {
+    (*action) (*i, i);
+  };
+  /* Hook */
+  if (coq_prev_scan_roots_hook != NULL) (*coq_prev_scan_roots_hook)(action);
+
+
+}
+
+void init_coq_stack()
+{
[…]

### 2021-08-15 media/drawing_videos/make_character_videos.m
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@3be4690818a2d6abaa3b5b24bd1709251fe32229 🚀

@@ -1,57 +1,0 @@
-
-%   <tr>
-%     <td><img src="drawing_videos/char9_Tagalog_images_small.jpg" width="390" height="417"></td>
-%     <td>
-% 	<video width="390" height="417" controls>
-% 	 <source src="drawing_videos/char9_Tagalog_high.mp4"  type='video/mp4; codecs="avc1.42E01E, mp4a.40.2"'>
-% 	 <source src="drawing_videos/char9_Tagalog_high.theora.ogg"  type='video/ogg; codecs="theora, vorbis"'>
-% 	</video>
-% 	</td>
-%   </tr>
-
-function make_character_videos
-    
-    dr = 'drawing_videos';
-    files = inspect_dir(dr);
-    nfiles = length(files);
-    
-    % Get all the unique base names
-    unique_base = [];
-    stop_str = {'_images','_high','_strokes'};
-    for i=1:nfiles
-        fn = files{i};    
-        fn_base = [];
-        for j=1:length(stop_str)
-            indxj = strfind(fn,stop_str{j});
-            if ~isempty(indxj)
-               fn_base = fn(1:indxj-1);
-            end
-        end
-        assert(~isempty(fn_base) || isequal(fn,'key.jpg'));
-        if ~isempty(fn_base)
-            unique_base = unique([unique_base; {fn_base}]); 
-        end
-    end
-    
-       
-    nunique = length(unique_base);
-    fid = fopen('table.txt','w');
-    for i=1:nunique
-        bn = unique_base{i};
-        fprintf(fid,'  <tr>\n');
-        
-        img = [bn,'_images_small.jpg'];
-        vid_mp4 = [bn,'_high.mp4'];
-        vid_ogg = [bn,'_high.theora.ogg'];
-        
-        fprintf(fid,'    <td><img src="drawing_videos/%s" width="390" height="417"></td>\n',img);
-        fprintf(fid,'    <td>\n');
-        fprintf(fid,'    <video width="390" height="417" controls>\n');
-        fprintf(fid,'      <source src="drawing_videos/%s"  type=''video/mp4; codecs="avc1.42E01E, mp4a.40.2">''',vid_mp4);
-        fprintf(fid,'      <source src="drawing_videos/%s"  type=''video/ogg; codecs="theora, vorbis"''>',vid_ogg);
-        fprintf(fid,'    </video>\n');
-        fprintf(fid,'    </td>\n');
-        fprintf(fid,'  </tr>\n');        
-    end
-    fclose(fid);    
-end

### 2021-08-15 papers/lob-paper/etc/ci/add_upstream.sh

@@ -1,11 +1,0 @@
-#!/bin/bash
-
-# in case we're run from out of git repo
-DIR="$( cd "$( dirname "${BASH_SOURCE[0]}" )" && pwd )"
-pushd "$DIR" 1>/dev/null
-
-echo "Adding remote..."
-git remote add upstream git://github.com/JasonGross/lob-paper
-git remote update
-
-popd 1>/dev/null

### 2021-08-15 papers/source/category-coq-experience/keep_only_good_urls.py

@@ -1,35 +1,0 @@
-#!/usr/bin/env python
-from __future__ import with_statement
-import sys, re
-
-VALID_URLS = ('https://github.com',
-              'http://coq.inria.fr/pylons',
-              'http://coq.inria.fr',
-              'http://www.cs.berkeley.edu/~megacz',
-              'http://mattam.org/repos',
-              'http://ncatlab.org',
-              'http://homotopytypetheory.org/2011/',
-              'http://mathoverflow.net/')
-
-def is_doi_line(line):
-    return re.match(r'^[^a-zA-Z]+doi = .+$', line) is not None
-
-def is_url_line(line):
-    return re.match(r'^[^a-zA-Z]+url = .+$', line) is not None \
-        or re.match(r'^[^a-zA-Z]+eprint = .+$', line) is not None
-
-def url_line_is_valid(line):
-    for url in VALID_URLS:
-        if re.match(r'^[^a-zA-Z]+url = {' + url + '.+$', line) is not None:
-            return True
-    return False
-
-def fixline(line):
-    return line.replace(r'@MASTERSTHESIS{spiwackverified,', r'@THESIS{spiwackverified,')
-
-if __name__ == '__main__':
-    with open(sys.argv[1], 'r') as f:
-        contents = f.readlines()
-    for i in contents:
-        if (not (is_url_line(i) or is_doi_line(i))) or url_line_is_valid(i):
-            print(fixline(i.strip('\n')))

### 2021-08-15 papers/source/lob-paper/etc/ci/update_nightlies.sh

@@ -1,110 +1,0 @@
-#!/bin/bash
-
-PS4='$ '
-set -x
-
-# in case we're run from out of git repo
-DIR="$( cd "$( dirname "${BASH_SOURCE[0]}" )" && pwd )"
-pushd "$DIR" 1>/dev/null
-
-# now change to the git root
-ROOT_DIR="$(git rev-parse --show-toplevel)"
-cd "$ROOT_DIR"
-
-# only make if we should ($UPDATE_NIGHTLIES is not empty) and we're the same as origin/master
-if [ -z "$UPDATE_NIGHTLIES" ]; then
-    echo 'Not updating nightlies because $UPDATE_NIGHTLIES is not set'
-    exit 0
-fi
-
-git reset --hard || exit 1
-
-"$DIR"/add_upstream.sh || exit 1
-
-git remote update || exit 1
-
-SUPPORTS_UNSHALLOW="$(git fetch --help | grep -c -- '--unshallow')"
-if [ "$SUPPORTS_UNSHALLOW" -eq 0 ]; then
-    echo "Your git ($(git --version)) does not support the --unshallow option to fetch"
-    git fetch --depth 1000000000 || exit 1
-else
-    git fetch --unshallow
-fi
-
-git fetch --tags || exit 1
-
-PDFS="lob.pdf lob-preprint.pdf"
-
-"$DIR"/configure_commit.sh || exit 1
-git remote -v
-git branch -a
-git --no-pager diff HEAD
-git --no-pager diff HEAD..origin/master
-git --no-pager diff HEAD..upstream/master
-ls
-
-BAD_REMOTES="$(git remote -v | grep origin | grep -v 'github.com/JasonGross/lob-paper')"
-UPSTREAM_LOG="$(git log HEAD..upstream/master)"
-#MASTER_LOG="$(git log HEAD..master)"
-#ORIGIN_LOG="$(git log HEAD..origin/master)"
-
-MASTER_COMMIT="$(git rev-parse HEAD)"
-
-git checkout -b gh-pages upstream/gh-pages || exit 1
-
-git rm -rf nightly || true
-rm -rf nightly || true
-
-mkdir nightly
-
-git add -f $PDFS || exit 1
-
-git mv $PDFS nightly/ || exit 1
-
-git rm -rf nightly/html
-
-git add -f html || exit 1
-
-git mv -f html nightly/ || exit 1
-
[…]

### 2021-08-15 papers/source/lob-paper/etc/ci/wget_retry.sh

@@ -1,7 +1,0 @@
-#!/bin/sh
-
-for i in 1 2 3 4 5 6 7 8 9 10 11 12 13 14 15 16 17 18 19 20
-do
-    echo "try $i"
-    wget -t 1 "$@" && exit 0
-done

### 2021-08-15 papers/source/reification-by-parametricity/benchmark/Benchmarks/make_test_parsing.sh

@@ -1,28 +1,0 @@
-#!/usr/bin/env bash
-
-function construct() {
-    echo 'Require Export Reify.Benchmarks.Parsing.'
-    echo ''
-    for i in "$@"; do
-	echo 'Goal True.'
-	echo "idtac \"Doing parsing (n= $i ) for \"\"Parsing\"\" with big :\";"
-	echo 'restart_timer "parsing".'
-	echo 'let x := constr:(fun var : Type =>'
-	echo "elet a'0 := NatS (@NatO var) * NatS (@NatO var) in"
-	for v in $(seq 0 $(expr $i - 1)); do
-	    echo "elet a'$(expr $v + 1) := Var a'$v * Var a'$v in"
-	done
-	echo "Var a'$i"
-	echo ') in'
-	echo 'finish_timing ( "Tactic call" ) "parsing".'
-	echo 'Abort.'
-    done
-}
-
-if [ ! -z "$1" ]; then
-    construct "$@"
-else
-    construct $(seq 1 500) 600 700 800 900 1000 > Parsing_quick.v
-    construct $(seq 1500 500 5000) > Parsing_medium.v
-    construct $(seq 6000 1000 10000) 20000 > Parsing_slow.v
-fi

### 2021-08-15 papers/source/reification-by-parametricity/benchmark/Benchmarks/make_test_parsing_elaborated.sh

@@ -1,31 +1,0 @@
-#!/usr/bin/env bash
-
-function construct() {
-    echo 'Require Export Reify.Benchmarks.ParsingElaborated.'
-    echo ''
-    for i in "$@"; do
-	echo 'Goal True.'
-	echo "idtac \"Doing parsing (n= $i ) for \"\"ParsingElaborated\"\" with big :\";"
-	echo 'restart_timer "parsing".'
-	echo 'let x := constr:(fun var : Type =>'
-	echo "@LetIn var (@NatMul var (@NatS var (@NatO var)) (@NatS var (@NatO var))) (fun a'0 : var =>"
-	for v in $(seq 0 $(expr $i - 1)); do
-	    echo "@LetIn var (@NatMul var (@Var var a'$v) (@Var var a'$v)) (fun a'$(expr $v + 1) : var =>"
-	done
-	echo "@Var var a'$i"
-	for v in $(seq 0 $(expr $i - 1)); do
-	    echo -n ')'
-	done
-	echo ')) in'
-	echo 'finish_timing ( "Tactic call" ) "parsing".'
-	echo 'Abort.'
-    done
-}
-
-if [ ! -z "$1" ]; then
-    construct "$@"
-else
-    construct $(seq 1 500) 600 700 800 900 1000 > ParsingElaborated_quick.v
-    construct $(seq 1500 500 5000) > ParsingElaborated_medium.v
-    construct $(seq 6000 1000 10000) 20000 > ParsingElaborated_slow.v
-fi

### 2021-08-15 papers/source/reification-by-parametricity/benchmark/Benchmarks/make_test_parsing_flat.py

@@ -1,47 +1,0 @@
-#!/usr/bin/env python2
-
-import sys
-
-def expr_of_count(i, elaborated):
-    if i <= 1:
-        if elaborated:
-            return '(@NatS var (@NatO var))'
-        else:
-            return '(NatS (@NatO var))'
-    elif i % 2 == 0:
-        subterm = expr_of_count(i / 2, elaborated)
-        if elaborated:
-            return '(@NatMul var %s %s)' % (subterm, subterm)
-        else:
-            return '(%s * %s)' % (subterm, subterm)
-    else:
-        one = expr_of_count(1, elaborated)
-        subterm = expr_of_count(i - 1, elaborated)
-        if elaborated:
-            return '(@NatMul var %s %s)' % (one, subterm)
-        else:
-            return '(%s * %s)' % (one, subterm)
-
-def construct(nums):
-    print('Require Export Reify.Benchmarks.ParsingFlat.')
-    print('')
-    for i in nums:
-        print('Goal True.')
-        print('idtac "Doing parsing (n= %d ) for ""ParsingElaborated"" with big_flat :";' % i)
-        print('restart_timer "parsing".')
-	print('let x := constr:(fun var : Type =>')
-        print('(%s)%%expr' % expr_of_count(i, True))
-        print(') in')
-	print('finish_timing ( "Tactic call" ) "parsing".')
-	print('Abort.')
-	print('Goal True.')
-	print('idtac "Doing parsing (n= %d ) for ""Parsing"" with big_flat :";' % i)
-	print('restart_timer "parsing".')
-	print('let x := constr:(fun var : Type =>')
-        print('(%s)%%expr' % expr_of_count(i, False))
-        print(') in')
-	print('finish_timing ( "Tactic call" ) "parsing".')
-	print('Abort.')
-
-if __name__ == '__main__':
-    construct(map(int, sys.argv[1:]))

### 2021-08-15 papers/source/reification-by-parametricity/comparison.py

@@ -1,102 +1,0 @@
-#!/usr/bin/env python2
-from __future__ import with_statement
-import re, sys
-
-STATS_RES = [re.compile(r'Doing (%s)\s+\(n= ([0-9]+) \) on ([^ ]+) with ([^ ]+) :\nTactic call ran for ([0-9\.]+) secs \(([0-9\.]+)u,([0-9\.]+)s\)' % ty,
-                        flags=re.DOTALL|re.MULTILINE)
-             for ty in ('identity cbv', 'identity lazy', 'identity vm_compute', 'identity native_compute', 'identity simpl', 'identity cbn', 'refine let', 'cbv Denote', 'lazy Denote', 'cbn Denote', 'simpl Denote')]
-
-NON_REIF_RES = [re.compile(res, flags=re.DOTALL|re.MULTILINE) for res in
-                [r'(Aggregate) time \(n= ([0-9]+) \) for "([^"]+)" with ([^ ]+) :\nTactic call aggregate ran for ([0-9\.]+) secs \(([0-9\.]+)u,([0-9\.]+)s\)'] +
-                [(r'Doing (%s)\s+\(n= ([0-9]+) \) for "([^"]+)" with ([^ ]+) :\nTactic call %s ran for ([0-9\.]+) secs \(([0-9\.]+)u,([0-9\.]+)s\)' % (ty, ty))
-                 for ty in ('cbv', 'pre', 'post')]]
-
-REIF_PRE_RE = re.compile(r'Doing reif\s+\(n= ([0-9]+) \) for "([^"]+)" with ([^ ]+) :\n' +
-                         r'(.*?\n' +
-                         r'Tactic call reif ran for [^\n]+)',
-                         flags=re.DOTALL|re.MULTILINE)
-REIF_RE = re.compile('Tactic call ([^\n]*?) ran for ([0-9\.]+) secs \(([0-9\.]+)u,([0-9\.]+)s\)',
-                     flags=re.DOTALL|re.MULTILINE)
-
-def fix_num(val):
-    if val == '0.': return '0'
-    return val
-
-def fix_times(real, user, system):
-    return (fix_num(real), fix_num(user), fix_num(system))
-
-def dict_set(d, value, *keys):
-    for k in keys[:-1]:
-        d[k] = d.get(k, {})
-        d = d[k]
-    d[keys[-1]] = value
-
-def process(fname):
-    with open(fname, 'r') as f:
-        contents = f.read()
-
-
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-compcert.sh

@@ -1,11 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-CompCert_CI_DIR=${CI_BUILD_DIR}/CompCert
-
-opam install -j "$NJOBS" -y menhir
-git_checkout ${CompCert_CI_BRANCH} ${CompCert_CI_GITURL} ${CompCert_CI_DIR}
-
-( cd ${CompCert_CI_DIR} && ./configure -ignore-coq-version x86_32-linux && make && make check-proof )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-coquelicot.sh

@@ -1,12 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-Coquelicot_CI_DIR=${CI_BUILD_DIR}/coquelicot
-
-install_ssreflect
-
-git_checkout ${Coquelicot_CI_BRANCH} ${Coquelicot_CI_GITURL} ${Coquelicot_CI_DIR}
-
-( cd ${Coquelicot_CI_DIR} && ./autogen.sh && ./configure && ./remake -j${NJOBS} )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-cpdt.sh

@@ -1,10 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-wget http://adam.chlipala.net/cpdt/cpdt.tgz
-tar xvfz cpdt.tgz
-
-( cd cpdt && make clean && make )
-

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-flocq.sh

@@ -1,10 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-Flocq_CI_DIR=${CI_BUILD_DIR}/flocq
-
-git_checkout ${Flocq_CI_BRANCH} ${Flocq_CI_GITURL} ${Flocq_CI_DIR}
-
-( cd ${Flocq_CI_DIR} && ./autogen.sh && ./configure && ./remake -j${NJOBS} )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-formal-topology.sh

@@ -1,10 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-formal_topology_CI_DIR=${CI_BUILD_DIR}/formal-topology
-
-git_checkout ${formal_topology_CI_BRANCH} ${formal_topology_CI_GITURL} ${formal_topology_CI_DIR}
-
-( cd ${formal_topology_CI_DIR} && make )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-hott.sh

@@ -1,10 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-HoTT_CI_DIR=${CI_BUILD_DIR}/HoTT
-
-git_checkout ${HoTT_CI_BRANCH} ${HoTT_CI_GITURL} ${HoTT_CI_DIR}
-
-( cd ${HoTT_CI_DIR} && ./autogen.sh && ./configure && make )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-template.sh

@@ -1,12 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-Template_CI_BRANCH=master
-Template_CI_GITURL=https://github.com/Template/Template
-Template_CI_DIR=${CI_BUILD_DIR}/Template
-
-git_checkout ${Template_CI_BRANCH} ${Template_CI_GITURL} ${Template_CI_DIR}
-
-( cd ${Template_CI_DIR} && make )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/lint-repository.sh

@@ -1,34 +1,0 @@
-#!/usr/bin/env bash
-
-# A script to check prettyness over the repository.
-
-# lint-commits.sh seeks to prevent the worsening of already present
-# problems, such as tab indentation in ml files. lint-repository.sh
-# seeks to prevent the (re-)introduction of solved problems, such as
-# newlines at the end of .v files.
-
-CODE=0
-
-if [ -n "${TRAVIS_PULL_REQUEST}" ] && [ "${TRAVIS_PULL_REQUEST}" != false ];
-then
-    # skip PRs from before the linter existed
-    if [ -z "$(git ls-tree --name-only "${TRAVIS_PULL_REQUEST_SHA}" dev/lint-commits.sh)" ];
-    then
-        1>&2 echo "Linting skipped: pull request older than the linter."
-        exit 0
-    fi
-
-    # Some problems are too widespread to fix in one commit, but we
-    # can still check that they don't worsen.
-    CUR_HEAD=${TRAVIS_COMMIT_RANGE%%...*}
-    PR_HEAD=${TRAVIS_COMMIT_RANGE##*...}
-    MERGE_BASE=$(git merge-base "$CUR_HEAD" "$PR_HEAD")
-    dev/lint-commits.sh "$MERGE_BASE" "$PR_HEAD" || CODE=1
-fi
-
-# Check that the files with 'whitespace' gitattribute end in a newline.
-# xargs exit status is 123 if any file failed the test
-find . "(" -path ./.git -prune ")" -o -type f -print0 |
-    xargs -0 dev/tools/check-eof-newline.sh || CODE=1
-
-exit $CODE

### 2021-08-15 presentations/coq-workshop-2018-coq/kernel/byterun/coq_interp.c

@@ -1,1523 +1,0 @@
-/***********************************************************************/
-/*                                                                     */
-/*                           Coq Compiler                              */
-/*                                                                     */
-/*        Benjamin Gregoire, projets Logical and Cristal               */
-/*                        INRIA Rocquencourt                           */
-/*                                                                     */
-/*                                                                     */
-/***********************************************************************/
-
-/* The bytecode interpreter */
-
-/* Spiwack: expanded the virtual machine with operators used
-   for fast computation of bounded (31bits) integers */
-
-#include <stdio.h>
-#include <signal.h>
-#include <stdint.h>
-#include "coq_gc.h"
-#include "coq_instruct.h"
-#include "coq_fix_code.h"
-#include "coq_memory.h" 
-#include "coq_values.h" 
-
-/* spiwack: I append here a few macros for value/number manipulation */
-#define uint32_of_value(val) (((uint32_t)(val)) >> 1)
-#define value_of_uint32(i)   ((value)((((uint32_t)(i)) << 1) | 1))
-#define UI64_of_uint32(lo) ((uint64_t)((uint32_t)(lo)))
-#define UI64_of_value(val) (UI64_of_uint32(uint32_of_value(val)))
-/* /spiwack */
-
-
-
-/* Registers for the abstract machine:
-        pc         the code pointer
-        sp         the stack pointer (grows downward)
-        accu       the accumulator
-        env        heap-allocated environment
-        trapsp     pointer to the current trap frame
-        extra_args number of extra arguments provided by the caller
-
-sp is a local copy of the global variable extern_sp. */
-
-
-
-/* Instruction decoding */
-
-
-#ifdef THREADED_CODE
-#  define Instruct(name) coq_lbl_##name: 
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/coq-makefile/multiroot/run.sh

@@ -1,0 +1,56 @@
+#!/usr/bin/env bash
+
+. ../template/init.sh
+
+cp -r theories theories2
+mv src/test_plugin.mlpack src/test_plugin.mllib
+coq_makefile -f _CoqProject -o Makefile
+cat Makefile.conf
+make
+make html mlihtml
+make install DSTROOT="$PWD/tmp"
+make install-doc DSTROOT="$PWD/tmp"
+#make debug
+(for d in `find tmp -name user-contrib` ; do pushd $d >/dev/null && find . && popd >/dev/null; done) | sort -u > actual
+sort > desired <<EOT
+.
+./test
+./test/test.glob
+./test/test.cmi
+./test/test.cmx
+./test/test_aux.cmi
+./test/test_aux.cmx
+./test/test_plugin.cmxa
+./test/test_plugin.cmxs
+./test/test.v
+./test/test.vo
+./test2
+./test2/test.glob
+./test2/test.v
+./test2/test.vo
+./orphan_test_test2_test
+./orphan_test_test2_test/html
+./orphan_test_test2_test/html/coqdoc.css
+./orphan_test_test2_test/html/index.html
+./orphan_test_test2_test/html/test2.test.html
+./orphan_test_test2_test/html/test.test.html
+./orphan_test_test2_test/html/toc.html
+./orphan_test_test2_test/mlihtml
+./orphan_test_test2_test/mlihtml/index_attributes.html
+./orphan_test_test2_test/mlihtml/index_classes.html
+./orphan_test_test2_test/mlihtml/index_class_types.html
+./orphan_test_test2_test/mlihtml/index_exceptions.html
+./orphan_test_test2_test/mlihtml/index_extensions.html
+./orphan_test_test2_test/mlihtml/index.html
+./orphan_test_test2_test/mlihtml/index_methods.html
+./orphan_test_test2_test/mlihtml/index_modules.html
+./orphan_test_test2_test/mlihtml/index_module_types.html
+./orphan_test_test2_test/mlihtml/index_types.html
+./orphan_test_test2_test/mlihtml/index_values.html
+./orphan_test_test2_test/mlihtml/style.css
+./orphan_test_test2_test/mlihtml/Test_aux.html
+./orphan_test_test2_test/mlihtml/Test.html
+./orphan_test_test2_test/mlihtml/type_Test_aux.html
+./orphan_test_test2_test/mlihtml/type_Test.html
+EOT
+exec diff -u desired actual

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/coq-makefile/plugin3/run.sh

@@ -1,0 +1,26 @@
+#!/usr/bin/env bash
+
+. ../template/init.sh
+
+mv src/test_plugin.mlpack src/test_plugin.mllib
+coq_makefile -f _CoqProject -o Makefile
+cat Makefile.conf
+make
+make html mlihtml
+make install DSTROOT="$PWD/tmp"
+#make debug
+(cd `find tmp -name user-contrib` && find .) | sort > actual
+sort > desired <<EOT
+.
+./test
+./test/test.glob
+./test/test.cmi
+./test/test.cmx
+./test/test_aux.cmi
+./test/test_aux.cmx
+./test/test_plugin.cmxa
+./test/test_plugin.cmxs
+./test/test.v
+./test/test.vo
+EOT
+exec diff -u desired actual

### 2021-08-15 papers/source/rewriting/pldi-2020-reviews-response-rejection.txt
Re-add deployed files

PLDI 2020 Paper #226 Reviews and Comments
===========================================================================
Paper #226 A Framework for Building Verified Partial Evaluators

Review #226A
===========================================================================

Overall merit
-------------
B. I support accepting this paper but will not champion it.

Reviewer expertise
------------------
Z. Some familiarity

Paper summary
-------------
This paper presents a new approach for partial evaluation in Coq. The
goal is to generate more efficient extracted code using partial
evaluation, but most existing partial evaluators for verified code
either increase the size of the trusted code base or cannot scale to
large programs. The new approach combines a different term encoding with
verified (potentially conditional) optimizations. The new framework is
capable of extracted efficient implementations of larger elliptic curve
crypto primitives than previous work.

Comments for author
-------------------
This is a well-written paper that was relatively easy to follow for me,
a total non-expert in interactive theorem proving. It walks the reader
through the motivation for a better partial evaluator (Fiat Crypto
couldn't extract code for larger curves), the desiderata, and the
solutions for each problem that arises. It's hard to argue with the
results -- the new extraction process is much faster, allowing it to
generate code for curves that the old implementation couldn't.

As a non-expert, one challenge I had was that this paper reads more like
a Coq experience report than a new conceptual contribution. There were a
number of places in the paper where I (and, it sounds like, the authors)
wondered if all this effort was just to work around some limitations of
the Coq implementation. Concretely, there are several points (like the
end of Section 3) where it seems like things would have been much easier
if not for termination checking; then, later on, it's revealed that we
got lucky(?) and the chosen PHOAS representation satisfies the
termination checker without any effort. I wonder how much those in the
PLDI audience who don't hack a lot of Coq will be able to learn from
this paper.

Section 3.1 discusses the need for efficient pattern matching for a
partial evaluator, and how pattern matching in Coq's logic isn't very
sophisticated, so the authors rolled their own based on [20]. That makes
sense. However, Section 5.2 concedes that actually, the performance of
the partial evaluator running inside Coq is pretty poor. To get that
performance back up, we need to extract the evaluator to OCaml and run
it natively. It seems like if we're going to make that concession, we
don't need to worry much about the efficiency of the pattern matching
engine -- we can just delegate to OCaml's, just like [1]?

Something I was unclear about throughout the paper was what the user
experience is for a tool like this. One of the cool things about LMS was
that it was trivial to adopt -- just changing the types of the code you
wanted to stage was enough to get the partial evaluation effect. I'm
less clear on the story here -- if I've written and proven an
implementation of some code, and now want to specialize it to a
particular input, what exactly do I have to write and/or do? Section 4.4
left me wondering how often I would actually need to do a lot more than
just write down some rewrite rules -- it seems like depending on the
problem, I would need to (1) write a new abstract interpretation, (2)
prove that correct, and (3) write new rules with the results of that
interpretation in mind. How laborious is that effort?

Post author-response comments
-----------------------------
Thanks for a well-written and helpful response. Ultimately, as a non-expert, I needed more help understanding how this paper would appeal to the PLDI audience broadly. In the paper's current form, I had a hard time distinguishing conceptual insights from Coq engineering tricks.

Review #226B
===========================================================================

Overall merit
-------------
A. I will champion accepting this paper.

Reviewer expertise
------------------
Y. Knowledgeable

Paper summary
-------------
The paper addresses the problem of performing rewrite-intensive program transformation in the context of a provably semantics-preserving system, with proofs generated in checked with a proof assistant. Since type theory-based proof assistants, by design, have limited kernels, and the primitive operations therein are small (e.g. beta-reduction and unfolding of definitions), the proofs quickly become too large too manage and generating them infeasible. The approach combines several techniques to avoid expensive rewriting under binders (such as Coq’s setoid_rewrite) but still use the native kernel for lambda calculus normalization.

Comments for author
-------------------
This work takes a very practical standpoint: if verified toolchains and proof-carrying code are to become mainstream, then someone has to do something about those proof term sizes. The solution is very creative, too. In some sense, it’s even counterintuitive: implement another layer of embedding when what you’re striving for is performance? — but the paper eventually delivers and the way that it does that is clever, by capitalizing on existing normalization mechanics that are already built into Coq.

I find the treatment of the problem systematic and one that presents several different facets of proof-assistant engineering. Also, the writing is superb.

= Specific comments =

p.1 [75] “fancier reduction strategies grow the trusted code base” — this is a bit over-generalizing. They don’t necessarily do so, and I understand after reading this paper that it’s either that or really blowing up the proof size, but IMHO it should be stated as a tradeoff.

p.2 [213] “reduction included in .. need not be requested explicitly.” What does it mean for a reduction to be “requested”?

p.5 [457] “rely on no such embeddings” — it’s perhaps besides the point, but Coq’s extraction mechanism must rely on the semantics of OCaml, and probably any attempt to build a certified extraction procedure will need to have such embedding, right?

p.5 [498] “set of inductive codes” — what are these codes? Is it just an implementation detail?

p.5 [545] “does not need to look into definitions” — so this rewriting scheme never unfolds the definitions of symbols? Why is that an acceptable limitation?

p.6 [630] “Wildcard”. The holes are not named nor indexed. Is it irrelevant or are they implicitly indexed left-to-right?
p.6 [636] “App” — I am guessing the edge in the decision tree should read “Switch” (as in [609])? Or is “Switch” supposed to be written inside the circle?

p.6 [654] “match e with etc.” — Where did the vectors go in this expression? Were they all eliminated by Coq’s Compute or similar?

p.11 [1169] “including reification, cbv, etc.” — It is not clear how important these are, and whether the reader should consider them as part of the running time or not. Surely reification is a vital part of the process, so what is the significance of “only rewriting”?

p.11 [1205] “Macrobenchmark”: maybe it’s just me but I read this as “Microbenchmark” (following 5.1) and was confused.

Post author-response comments
-----------------------------
*Outcome of PC discussion*

While all the reviewers appreciated the effort involved in leveraging the agreeably weak beta-reduction in Coq’s kernel without resorting to OCaml implementation, the contribution is found to be too narrow and the vast majority of the PLDI audience will not be able to relate to it.

Review #226C
===========================================================================

Overall merit
-------------
B. I support accepting this paper but will not champion it.

Reviewer expertise
------------------
Z. Some familiarity

Paper summary
-------------
Partial evaluation is an optimization technique that aims at generating faster, specialized variants of a function by reducing its application domain. An example in cryptography is a specialization of a function to a certain key size. The resulting function is faster than the more generic function, but a proof is required to demonstrate that the behaviour of the new function is the same as the general function. This paper presents a way to efficiently prove the correctness of partial evaluators. Although previous approaches exist, they introduced a trusted component in the Trusted Computing Base (TCB), and had performance issue due to the lack of common sub-expression handling. This paper proposes an approach that does not add to the TCB and that takes common sub-expression into account to reduce the redundancy in generated proofs.

In order to prove correctness of partial evaluators, the authors have developed a Coq library, which works in two steps: first a rewriter is generated from equality lemmas, and then it can be applied to prove the equality of two functions or synthesize a specialized version of a function with its correctness proof. The paper presents a quick overview of the features of the library, followed by a more in-depth explanation of the techniques used to build it, comparing with existing implementations, as their work is heavily inspired by work by Aehlig et al. The paper evaluates the library speed compared to coq standard tactics on two microbenchmarks, and on a macrobenchmark involving code generation for the Fiat cryptography library.

Comments for author
-------------------
Positives
----------- 
- The paper is well-structured and easy to follow. 
- The authors are positioning well their work in the context of existing approaches and techniques that they use. The paper explains when they adapted them and justifies all of their choices. 
They also compared how their own use-case impacts existing approaches, and how their implementation overcomes the difficulty. Overall, the formalization flows almost naturally.

Negatives
---------------

- Through the paper there is a constant comparison between the introduced  approach and the work of Aehlig et al [1]. One expects that those two approaches will also be empirically compared and evaluated. 
- Although important to the presented work, side conditions (in Sections 4.3 and 4.4) are not as well explained and justified as the rest of the paper. 
- The abstract interpreter is not evaluated.

Suggestions
----------------------------------------------
- Figure 3 (benchmark results) is really hard to read and evaluate since the text is small and the graphs are barely readable. Slightly reducing the size of Figure 1 might help you get enough space for Figure 3.

Review #226D
===========================================================================

Overall merit
-------------
B. I support accepting this paper but will not champion it.

Reviewer expertise
------------------
Y. Knowledgeable

Paper summary
-------------
This paper proposes a new approach to do partial evaluation in Coq.  A large component of it is a term rewriter, which is used to simplify terms as knowledge about some parameters is given.  Unlike previous work by Achlig etal [1], where an operational semantics of an ML-like language for terms was defined via a deep embedding and had to be trusted as part of the overall proof checking, here the authors propose an approach where the language is defined at the core level via translation back and forth to Coq itself.  While this translation-based definition still needs to be given, the advantage is that Coq can now take over a lot of the checking work (like termination) and thus the overall trust base stays the same, without any need to modify or extend the proof checker.

Coq being Coq, a lot of challenges had to be overcome in order to achieve usable performance.  First, a less-than-naive rewrite engine had to be implemented, where some limited indexing across rewrite rules was used, to avoid doing work from scratch each time a rule is attempted to be matched.    Second, translations to Coq and normalization of evaluation of terms had to be implemented in a certain way, to convince Coq that everything terminates.  Third, a shallow encoding of terms' binders into Coq's existing binders had to be implemented, to avoid the cost of de Bruijn encodings.  Then subterm sharing and side conditions to rules were necessary, which are also considered basic features of and implemented in other rewrite engines outhere.

Evaluation was done both on microbenchmarks and on a larger example, the Fiat Crytocurrency.  The results show good results.

Comments for author
-------------------
The paper is overall well-written.  The choice of what to include in the paper and what in the appendix is fair, although there is a lot of material in the appendix that is required to appreciate the contribution.  Not sure what you can do about it, though, but at least make sure that it will be all available in a companion report if the paper is accepted.  I do not have any major questions, only a few minor ones that made me have some reservations about the paper and not give it an A.

Why not implement, once and for all in Coq, the efficient rewriting engine algorithms, based on indexing and pattern match automata, that the term rewriting community has developed over the years and implemented in truly fast rewrite engines, performing millions, or even tens of millions, or rewrite steps per second?  That would benefit the entire Coq community.

Why should one trust the rewrite rules as correct?  Don't they become part of the trust base anyway?  Also, isn't your term translation back and forth to Coq the same as like giving a semantics to an ML-like language like in [1]?  One still has to trust these translations, including the fact that the ML-like language and Coq share the semantics of their similar constructs., which to me counts as the same level of trust that [1] requires from their users.  You should clarify that in the paper.

Lines 701-704: this becomes part of the trust base, no?

How is the approach in [11] different from the approach here?  Is it only a performance improvement?  How about the trust-base?  I was a bit disappointed to read that you generate essentially the same code as [11] and you only care about the performance of the partial evaluator; it would have been a really nice application where correctness is guaranteed through Coq, but if the application already had all these guarantees, then the work in this paper feels less critical and more incremental.

Response by Jason Gross <jgross@mit.edu>
---------------------------------------------------------------------------
Thank you all for your feedback and comments.  We'll start by
clarifying a few most crucial elements of our approach and its
advantages, after which we'll return to answer further questions on a
review-by-review basis (some of which are forward-referenced by our
initial highlights).

First, a quick response to Review A's questions about what the user
experience is really like.  Section 1.1 does provide a complete
example; we are not eliding any additional user overhead of applying
our rewrite engine.  (Abstract interpretation is a separate activity
that we say a bit more about later, but this paper is not about
abstract interpretation, but rather a flexible framework that can
interface with abstract intepretation.)

Next: Review D revealed a number of places where we left it unclear
how we avoid growing the trusted base, beyond what one accepts anyway
in using Coq, and we're glad to have the opportunity to clarify.
Briefly:
- *Must one trust that the registered rewrite rules are correct*?
   No, the verified rewriting engine requires a proof that the
   interpretation of the AST of each rewrite rule is a correct
   proposition in the logic.
- *Must one trust the translations back and forth between normal Coq and our ASTs?*
  No, the translation *into* ASTs is inherently untrusted, as we ask
  the Coq kernel to "run it backwards" through interpretation of ASTs,
  for instance in formulating the semantic correctness condition for a
  rewrite rule. The kernel naturally checks that interpretation takes
  us back to the terms (e.g., rewrite-rule statements) we started
  with.
- *Are the central NbE function definitions (lines 701-704) trusted?*
  No, they receive full proofs of functional correctness (which are
  used as lemmas in the overall engine proof) and thus need not be
  trusted.
Summing up, in any particular use of our rewrite engine, the final
theorem statement proved truly does not mention or depend on any part
of our framework; only the *proof* depends on our framework.  Since we
do not add any axioms in this project, the trusted base is not
expanded.  And see our more detailed Review D response below for more.

Finally, a few brief remarks on performance, which indeed is at the
heart of our project's motivation. Review D expresses disappointment
that, in comparison to [11], our project is only about improving the
performance of a compilation pipeline; we should have emphasized that
the tool from [11] timed out (running longer than 100 hours) on many
important inputs, while our new version handles all known real-world
elliptic curves usually within a few minutes or at most two hours. Also,
Review A summarizes Section 5.2 as concluding that our engine's
performance is poor. We want to emphasize again that there we compare
against reduction on programs (from [11]) that have been
tailor-written in CPS just to improve reduction performance, while
our engine works on much more readable, direct-style code, in addition
to integrating flexible rewriting with reduction.  Still, our
within-Coq evaluation doesn't pass 90 seconds of running time until
primes grow to about 10 times as many bits as native machine words.

Responses to Review A
=====================

Regarding termination checking, indeed, normalization by evaluation in
PHOAS is, in some sense, specifically crafted to appease the
termination checker in an elegant way.

Broadly, none of the individual techniques are new.  What's new is
that we combine all of them to get a flexible verified tool that
operates efficiently at scale, without extending the trusted code
base.

> However, Section 5.2 concedes that actually, the performance of the
> partial evaluator running inside Coq is pretty poor.

While the performance is poor compared to the extracted version and is
not stellar compared to the "rewrite all the code in CPS" version
(which is inflexible), the performance is still quite good relative to
the prior state-of-the-art equational reasoning in Coq of
`setoid_rewrite` / `rewrite_strat`, which take around 90 seconds for
primes equal to the bitwidth, and about 3 hours for primes with twice
as many bits as the machine architecture.  By contrast, even within
Coq, our approach only approaches 90 seconds for primes with about 10
times as many bits as the machine architecture.  We'll aim to
emphasize this more in future versions of the paper.

> It seems like if we're going to make that concession, we don't need
> to worry much about the efficiency of the pattern matching engine --
> we can just delegate to OCaml's, just like [1]?

If we delegate to OCaml's pattern matching engine, then we can't do
conditional rewrite rules, because there's not currently a way to
extract Coq pattern matches to OCaml's conditional matches.
Additionally, we tried eliminating pattern-matching compilation from
our engine.  This results in out-of-memory errors during compile-time
(because we end up being quadratic in the number of rewrite rules for
reasons that are slightly hard to explain), and if we disable the
partial evaluation of rewrite rule selection (trusting in OCaml's
pattern matching engine), we lose approximately a factor of 8 in
performance of extracted code.

The intended user experience for a tool like this is detailed at the
end of section 1.1.  The couple lines of code described there are all
that is necessary for equational reasoning.  The work of abstract
interpretation was required for Fiat Cryptography to handle automatic
specialization from unbounded integers to fixed-width integers.  This
is more than just equational reasoning, and, indeed, was a somewhat
laborious process.  We have yet to find any better alternative to
this, however, when abstract interpretation is, in fact, required.

Responses to Review B
=====================

Thank you for your detailed comments about places where we were
unclear or misstating things.

Responding to the questions:

> p.2 [213] “reduction included in .. need not be requested
> explicitly.” What does it mean for a reduction to be “requested”?

Imagine a proof assistant without any computational reduction rules,
only axioms.  In such a system, every reduction step needs to
correspond to a use of an axiom in the proof term.  This is what we
mean by "requested".

> p.5 [457] “rely on no such embeddings” — it’s perhaps besides the
> point, but Coq’s extraction mechanism must rely on the semantics of
> OCaml, and probably any attempt to build a certified extraction
> procedure will need to have such embedding, right?

Yes.

> p.5 [498] “set of inductive codes” — what are these codes? Is it
> just an implementation detail?

The inductive codes are the inductive types from step (2), "Inductive
types enumerating all available primitive types and functions are
emitted."  For example, we might have the inductive codes for types
`Inductive base := nat_code | bool_code | Z_code` and the inductive
codes for functions `Inductive ident := nat_add | Z_add | bool_if`.
Details beyond this are basically just implementation details.

> p.5 [545] “does not need to look into definitions” — so this
> rewriting scheme never unfolds the definitions of symbols? Why is
> that an acceptable limitation?

The rewriter itself has no native support for unfolding definitions,
but you can include a rewrite rule that corresponds to an "unfolding",
such as `?n + S ?m = S (n + m)` or `fst (?x, ?y) = x`.

> p.6 [630] “Wildcard”. The holes are not named nor indexed. Is it
> irrelevant or are they implicitly indexed left-to-right?

Holes are indexed left-to-right, but this is an implementation detail;
the machinery for reifying Gallina rewrite rules into patterns and
replacements takes care of indexing.

> p.6 [636] “App” — I am guessing the edge in the decision tree should
> read “Switch” (as in [609])? Or is “Switch” supposed to be written
> inside the circle?

We should have been more clear that every un-labeled node is a
"Switch" node.  The "App" edges are for the `app_case` on [609], and
the identifier-labeled edges are part of `icases`.

> p.6 [654] “match e with etc.” — Where did the vectors go in this
> expression? Were they all eliminated by Coq’s Compute or similar?

Yes, they are all eliminated under `cbv` / `cbn` during the partial
evaluation described on lines [644]-[646].

> p.11 [1169] “including reification, cbv, etc.” — It is not clear how
> important these are, and whether the reader should consider them as
> part of the running time or not. Surely reification is a vital part
> of the process, so what is the significance of “only rewriting”?

The response to this is hinted at on [1120]-[1123]: "As the
reification method was not especially optimized, and there exist fast
reification methods [13]".  However, refication is not actually the
bottleneck.  The bottleneck is the final cbv at the end, and this is
only due to a bug in Coq's implementation of NbE reduction that makes
substitution quadratic in the number of
binders (we wish we could link to the bug tracker, but it might reveal
something about our identities).
While we felt it would be misleading to give only the "only rewriting"
numbers, we do believe this is really the benchmark that matters, Coq
bugs aside.  (And we hope this bug in Coq will be fixed soon.)

Responses to Review C
=====================

Thank you for your points; we will aim to incorporate your feedback in
the next revision of the paper.

One clarification we'd like to make is that the abstract interpreter
itself is beyond the scope of this project; we merely want to
demonstrate that it is possible to integrate separate abstract
interpretation passes with the partial evaluator and rewriting
framework.

Responses to Review D
=====================

> Why not implement, once and for all in Coq, the efficient rewriting
> engine algorithms, based on indexing and pattern match automata,
> that the term rewriting community has developed over the years and
> implemented in truly fast rewrite engines, performing millions, or
> even tens of millions, or rewrite steps per second?  That would
> benefit the entire Coq community.

Indeed, this would be a great next step.  We started with a somewhat
more modest goal, and already found the proof effort significant.

> Why should one trust the rewrite rules as correct? Don't they become
> part of the trust base anyway?

To be accepted by the engine, the rewrite rules must be proven in Coq.
So no unsound rewrites are permitted (unless you use unsound axioms to
prove them). The checking happens as part of the step from line 502 of
the paper, where we bring all the rewrite rules together in a list of
dependently typed records, each one pairing the syntax of a rewrite
rule with a required proof of its soundness, stated in terms of
interpretation of the reified syntax tree.

> Also, isn't your term translation back and forth to Coq the same as
> like giving a semantics to an ML-like language like in [1]?  One
> still has to trust these translations, including the fact that the
> ML-like language and Coq share the semantics of their similar
> constructs., which to me counts as the same level of trust that [1]
> requires from their users.  You should clarify that in the paper.

The translations in our tool are *not* trusted. This is because every
time our code reifies a term, the Coq kernel checks that the embedded
AST our code produced has the same semantics as the original term by
unfolding the interpreter applied to the AST and observing that it
results in the original term. Our rewriting transformation is proven to
be semantics-preserving, so the theorem we get at the end is that the
new term, post-rewriting, has the same semantics as the original term.
Since the kernel checks the correctness of the semantics of the initial
term, there's nothing additional here to trust. We should perhaps
clarify this in the paper.

> Lines 701-704: this becomes part of the trust base, no?

Thankfully, it does not.  In this case, reify and reflect are internal
operations which are not exposed in the end-to-end theorem (any moreso
than code for CompCert's intermediate phases appears in a generated
theorem witnessing correct compilation of a concrete C program), and
we prove the properties that we need of these operations.

> How is the approach in [11] different from the approach here?

> Is it only a performance improvement?

[11] relies on a delicate ad-hoc combination of hand-writing all of
the code in CPS to take advantage of built-in reduction for partial
evaluation, and then using a hand-rolled domain-specific rewriter to do
the necessary algebraic simplifications. The authors of [11] describe
the need to write all algorithms in CPS as "arbitrary and unfortunate."
Further, this two-stage approach requires reifying large blocks of code,
and reifying repeatedly for every prime, which was prohibitively slow in
many cases ([11] Appendix B).

The approach described here lets us reify the generic code once, and
combine all of the rewriting and partial evaluation into the same
generic framework. We believe this is much easier to use. Probably the
clearest example of this is that our rewriting and partial evaluation
can be applied to code that was written without considering our tool as
a target, whereas the code repository accompanying [11] contains two
versions of many functions: one human-readable, and one for partial
evaluation (which they prove equivalent each time).

The performance improvement is also essential, in the sense that some of
the primes that did not finish even after waiting nearly a thousand
hours in [11] finished for us in seconds, minutes, or, at most, an hour
or two.

> How about the trust-base?

Running our rewriter in Coq has the same trusted base as [11].  The
faster version of our code also relies on extraction to OCaml, but this
is not essential.

> I was a bit disappointed to read that you generate essentially the
> same code as [11] and you only care about the performance of the
> partial evaluator; it would have been a really nice application
> where correctness is guaranteed through Coq, but if the application
> already had all these guarantees, then the work in this paper feels
> less critical and more incremental.

Again, we would like to emphasize that many primes could not be
handled at all by [11] even with hundreds of hours and 64 GB of RAM.
The "performance improvement" here allows us to not just handle the
same primes faster, but also allows us to handle primes that could not
be handled at all.

### 2021-08-15 papers/source/coqpl-15/coq-bug-minimizer.tex
Re-add deployed files

\documentclass{article}
\usepackage{amsmath}
\usepackage{hyperref}
\title{Coq Bug Minimizer}
\author{Jason Gross}
\date{CoqPL 2015}
% Key Words: bugs, test-suite, automation, tools
\begin{document}
\maketitle
\begin{abstract}
\subsection*{Sales Pitch}
Are bugs the bane of your existence?  Do you dread Coq upgrades, because they mean you'll have to spend days tracking down subtle failures deep in your developments?  Have you ever hit an anomaly that just wouldn't go away, and wished you understood what triggered it?  Have you ever been tormented by two blocks of code that looked identical, but behaved differently?  Do you wish you could submit more helpful error reports, but don't want to put in the time to construct minimal examples?  If you answered ``yes'' to any of these questions, then the Coq Bug Minimizer is for you!  Clone your own copy at \url{https://github.com/JasonGross/coq-bug-finder}.

\subsection*{What does it do?  How do I use it?}
The Coq Bug Minimizer is an external script that will take any error-generating Coq file in the middle of a development, and after confirming that it has picked up the right error message, will automatically produce a small, stand-alone file which is guaranteed to generate the same error message.

The transformations automatically performed by the bug minimizer include:
\begin{itemize}
    \item inlining all of the dependencies of a file
    \item removing unnecessary definitions, lemmas, hints, tactics, section variables, notations, and other lines
    \item (optionally) replacing unnecessary proofs and definition bodies with \texttt{admit}
    \item flattening module and section structure
    \item removing unnecessary imports
\end{itemize}

The bug minimizer explicitly does not perform the following transformations:
\begin{itemize}
    \item removing unnecessary definition-specific variables
    \item removing constructors of inductive types nor fields of records
    \item simplifying the error-generating proof script itself
\end{itemize}
The first two are, unfortunately, changes that are often required to happen in multiple places at once; the bug minimizer, currently, only considers changes that can be made in isolation.  The third of these is actually a feature; it allows you to check for a particular state and then \texttt{fail} with an appropriate message.\footnote{If you have case where \texttt{rewrite} used to progress but no longer does, you would want to be able to use \texttt{progress rewrite foo} to trigger the error.  But if the minimizer was allowed to change anything, it could inform you that \texttt{progress idtac} produces the same error message.  That is certainly not the test case you were looking for!}

Hence, typical interaction will consist of a few rounds of automatic minimization alternated with manual simplification of the final proof script and removal of record fields.  In practice, each round of automatic minimization decreases the number of lines by a factor of 5--10, bottoming out at around 50--500 lines.

\subsection*{This sounds amazing!  Is it fast, too?}
Not yet, unfortunately.  Because it re-runs Coq for every change it attempts, it tends to take a few minutes to minimize around a thousand lines of code, a few hours to handle a 5\,000 line file, and a few weeks to handle a 40\,000 line file.

However, by the time the workshop comes, the Coq Bug Minimizer will hopefully be 100\% faster!\footnote{How is this possible, you may ask?  Currently, the bug minimizer takes time roughly $\mathcal O((\text{lines of code})^3)$.  By taking advantage of \texttt{BackTo}, it should be possible to chop off at least a factor of the number of lines of code.}

\subsection*{What's left to do?}
Possible future avenues of development include:
\begin{itemize}
    \item automatic anomaly reporting integrated into CoqIDE
    \item automatic bug minimization service, where an end user can upload a tarball of their development, and the Coq developers can get a standalone test-case
    \item smarter heuristics for minimization, e.g., remove slower lines first
    \item integration of multiple versions of Coq simultaneously
    \item using the new \texttt{-ideslave} XML protocol rather than the old and error-prone \texttt{-emacs} option
    \item translation from Python into a saner language (like OCaml or Haskell)
\end{itemize}

\subsection*{What's left to say?}
The proposed presentation will consist of a demo of using the Coq Bug Minimizer and a brief explanation of how it works and what scenarios it can be used in.  This will be followed by an audience driven presentation or discussion of other use-cases and details of its functioning, possibly including the warts of Coq that workarounds were required for.  The presentation will close with a brief period for the audience to make feature requests.
\end{abstract}
\end{document}

### 2021-08-15 papers/source/category-coq-experience/make_comparison_table.py

@@ -1,0 +1,211 @@
+#!/usr/bin/env python
+from __future__ import with_statement
+
+#  grep -R '^\s*\(Global\s\+\|Program\s\+\)*\(Definition\|Lemma\|Theorem\|Corollary\|Remark\|Fact\|Proposition\|Example\|Fixpoint\|CoFixpoint\|Inductive\|CoInductive\|Instance\)' . | sort | wc -l
+
+
+def median(l):
+    l = sorted(l)
+    if len(l) % 2 == 1:
+        # length 3 -> index 1
+        return l[(len(l) - 1) / 2]
+    else:
+        # length 2 -> average of index 1 and index 2
+        return (l[(len(l) - 2) / 2] + l[(len(l) / 2)]) / 2
+
+def format_sections(secs):
+    return '%dm%ds' % (secs / 60, secs % 60)
+
+def compute_attrs(table):
+    rtn = []
+    for row in table:
+        new_row = dict(row)
+        if 'file times' in row:
+            times = tuple(j['user'] for j in row['file times'].values())
+            new_row['files'] = len(times)
+            new_row['total time'] = str(int(sum(times))) + 's'
+            new_row['total time no m'] = str(int(sum(v['user'] for k, v in row['file times'].items()
+                                                     if 'monoidal' not in k.lower() and 'enrich' not in k.lower()))) + 's'
+            new_row['median time'] = '%0.01fs' % median(times)
+            if 'big times' in row:
+                for feature, time in row['big times'].items():
+                    new_row[feature + '-frac'] = time / 20.0 / 10 # sum(times)
+        if 'file stats' in row:
+            stats = tuple(row['file stats'].values())
+            new_row['lpf-mean'] = int(sum(i['lines'] for i in stats) / len(stats))
+            new_row['lpf-median'] = int(median(i['lines'] for i in stats))
+            if all('definitions' in i for i in stats):
+                new_row['constr'] = sum(i['definitions'] for i in stats)
+        rtn.append(new_row)
+    return tuple(rtn)
+
+PRETABLE = (
[…]

### 2021-08-15 papers/source/category-coq-experience/presentation/media/many_slow_f_equal.sh

@@ -1,0 +1,108 @@
+#!/bin/bash
+
+if [ -z "$4" ]; then
+    echo "Usage: $0 HEADER FILENAME TOTAL-ITERATIONS MIN-DUP MAX-DUP"
+    exit 1
+fi
+
+if [ -z "$COQC" ]; then
+    COQC=coqc
+fi
+
+HEADER="$1"
+HEADER_FILENAME="$1.v"
+OUT="$2"
+FILENAME="$2.v"
+TOTAL="$3"
+MIN_DUP="$4"
+MAX_DUP="$5"
+
+#echo "$FILENAME"
+
+rm -f "$HEADER_FILENAME" "$FILENAME"
+
+for i in $(seq $MIN_DUP $MAX_DUP)
+do
+    echo -n "Axiom iter$i : " >> "$HEADER_FILENAME"
+    for j in $(seq 1 $i)
+    do
+        echo -n "bool -> " >> "$HEADER_FILENAME"
+    done
+    echo "bool." >> "$HEADER_FILENAME"
+    cat >> "$HEADER_FILENAME" <<EOF
+Fixpoint do$i (n : nat) (A : bool) :=
+  match n with
+    | 0 => A
+EOF
+    echo -n "    | S n' => do$i n' (iter$i" >> "$HEADER_FILENAME"
+    for j in $(seq 1 $i)
+    do
+        echo -n " A" >> "$HEADER_FILENAME"
+    done
+    echo ")" >> "$HEADER_FILENAME"
+    echo "  end." >> "$HEADER_FILENAME"
+
+    echo "Ltac display$i := match goal with |- let n := ?k in _ => "'idtac "['"$i"'][" k "] :="'" end; hnf; unfold do$i; cbv beta." >> "$HEADER_FILENAME"
+done
+
+
+
+for i in $(seq $MIN_DUP $MAX_DUP)
+do
+    TAC_I=0
+    for tac in "apply f_equal" \
+	"lazymatch goal with |- ?f ?a = ?g ?b => idtac end" \
+	"lazymatch goal with |- ?f ?a = ?g ?b => let H := constr:(@f_equal bool bool f a b) in idtac end" \
+	"lazymatch goal with |- ?f ?a = ?g ?b => let H := constr:(@f_equal bool bool f a b (@eq_refl bool a)) in idtac end" \
[…]

### 2021-08-15 papers/source/lob-paper/etc/ci/update_nightlies.sh

@@ -1,0 +1,110 @@
+#!/bin/bash
+
+PS4='$ '
+set -x
+
+# in case we're run from out of git repo
+DIR="$( cd "$( dirname "${BASH_SOURCE[0]}" )" && pwd )"
+pushd "$DIR" 1>/dev/null
+
+# now change to the git root
+ROOT_DIR="$(git rev-parse --show-toplevel)"
+cd "$ROOT_DIR"
+
+# only make if we should ($UPDATE_NIGHTLIES is not empty) and we're the same as origin/master
+if [ -z "$UPDATE_NIGHTLIES" ]; then
+    echo 'Not updating nightlies because $UPDATE_NIGHTLIES is not set'
+    exit 0
+fi
+
+git reset --hard || exit 1
+
+"$DIR"/add_upstream.sh || exit 1
+
+git remote update || exit 1
+
+SUPPORTS_UNSHALLOW="$(git fetch --help | grep -c -- '--unshallow')"
+if [ "$SUPPORTS_UNSHALLOW" -eq 0 ]; then
+    echo "Your git ($(git --version)) does not support the --unshallow option to fetch"
+    git fetch --depth 1000000000 || exit 1
+else
+    git fetch --unshallow
+fi
+
+git fetch --tags || exit 1
+
+PDFS="lob.pdf lob-preprint.pdf"
+
+"$DIR"/configure_commit.sh || exit 1
+git remote -v
+git branch -a
+git --no-pager diff HEAD
+git --no-pager diff HEAD..origin/master
+git --no-pager diff HEAD..upstream/master
+ls
+
+BAD_REMOTES="$(git remote -v | grep origin | grep -v 'github.com/JasonGross/lob-paper')"
+UPSTREAM_LOG="$(git log HEAD..upstream/master)"
+#MASTER_LOG="$(git log HEAD..master)"
+#ORIGIN_LOG="$(git log HEAD..origin/master)"
+
+MASTER_COMMIT="$(git rev-parse HEAD)"
+
+git checkout -b gh-pages upstream/gh-pages || exit 1
+
+git rm -rf nightly || true
+rm -rf nightly || true
+
+mkdir nightly
+
+git add -f $PDFS || exit 1
+
+git mv $PDFS nightly/ || exit 1
+
+git rm -rf nightly/html
+
+git add -f html || exit 1
+
+git mv -f html nightly/ || exit 1
+
[…]

### 2021-08-15 papers/source/reification-by-parametricity/comparison.py

@@ -1,0 +1,102 @@
+#!/usr/bin/env python2
+from __future__ import with_statement
+import re, sys
+
+STATS_RES = [re.compile(r'Doing (%s)\s+\(n= ([0-9]+) \) on ([^ ]+) with ([^ ]+) :\nTactic call ran for ([0-9\.]+) secs \(([0-9\.]+)u,([0-9\.]+)s\)' % ty,
+                        flags=re.DOTALL|re.MULTILINE)
+             for ty in ('identity cbv', 'identity lazy', 'identity vm_compute', 'identity native_compute', 'identity simpl', 'identity cbn', 'refine let', 'cbv Denote', 'lazy Denote', 'cbn Denote', 'simpl Denote')]
+
+NON_REIF_RES = [re.compile(res, flags=re.DOTALL|re.MULTILINE) for res in
+                [r'(Aggregate) time \(n= ([0-9]+) \) for "([^"]+)" with ([^ ]+) :\nTactic call aggregate ran for ([0-9\.]+) secs \(([0-9\.]+)u,([0-9\.]+)s\)'] +
+                [(r'Doing (%s)\s+\(n= ([0-9]+) \) for "([^"]+)" with ([^ ]+) :\nTactic call %s ran for ([0-9\.]+) secs \(([0-9\.]+)u,([0-9\.]+)s\)' % (ty, ty))
+                 for ty in ('cbv', 'pre', 'post')]]
+
+REIF_PRE_RE = re.compile(r'Doing reif\s+\(n= ([0-9]+) \) for "([^"]+)" with ([^ ]+) :\n' +
+                         r'(.*?\n' +
+                         r'Tactic call reif ran for [^\n]+)',
+                         flags=re.DOTALL|re.MULTILINE)
+REIF_RE = re.compile('Tactic call ([^\n]*?) ran for ([0-9\.]+) secs \(([0-9\.]+)u,([0-9\.]+)s\)',
+                     flags=re.DOTALL|re.MULTILINE)
+
+def fix_num(val):
+    if val == '0.': return '0'
+    return val
+
+def fix_times(real, user, system):
+    return (fix_num(real), fix_num(user), fix_num(system))
+
+def dict_set(d, value, *keys):
+    for k in keys[:-1]:
+        d[k] = d.get(k, {})
+        d = d[k]
+    d[keys[-1]] = value
+
+def process(fname):
+    with open(fname, 'r') as f:
+        contents = f.read()
+
+
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-coquelicot.sh

@@ -1,0 +1,12 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+Coquelicot_CI_DIR=${CI_BUILD_DIR}/coquelicot
+
+install_ssreflect
+
+git_checkout ${Coquelicot_CI_BRANCH} ${Coquelicot_CI_GITURL} ${Coquelicot_CI_DIR}
+
+( cd ${Coquelicot_CI_DIR} && ./autogen.sh && ./configure && ./remake -j${NJOBS} )

### 2021-08-15 presentations/coq-workshop-2018-coq/kernel/byterun/coq_interp.c

@@ -1,0 +1,1523 @@
+/***********************************************************************/
+/*                                                                     */
+/*                           Coq Compiler                              */
+/*                                                                     */
+/*        Benjamin Gregoire, projets Logical and Cristal               */
+/*                        INRIA Rocquencourt                           */
+/*                                                                     */
+/*                                                                     */
+/***********************************************************************/
+
+/* The bytecode interpreter */
+
+/* Spiwack: expanded the virtual machine with operators used
+   for fast computation of bounded (31bits) integers */
+
+#include <stdio.h>
+#include <signal.h>
+#include <stdint.h>
+#include "coq_gc.h"
+#include "coq_instruct.h"
+#include "coq_fix_code.h"
+#include "coq_memory.h" 
+#include "coq_values.h" 
+
+/* spiwack: I append here a few macros for value/number manipulation */
+#define uint32_of_value(val) (((uint32_t)(val)) >> 1)
+#define value_of_uint32(i)   ((value)((((uint32_t)(i)) << 1) | 1))
+#define UI64_of_uint32(lo) ((uint64_t)((uint32_t)(lo)))
+#define UI64_of_value(val) (UI64_of_uint32(uint32_of_value(val)))
+/* /spiwack */
+
+
+
+/* Registers for the abstract machine:
+        pc         the code pointer
+        sp         the stack pointer (grows downward)
+        accu       the accumulator
+        env        heap-allocated environment
+        trapsp     pointer to the current trap frame
+        extra_args number of extra arguments provided by the caller
+
+sp is a local copy of the global variable extern_sp. */
+
+
+
+/* Instruction decoding */
+
+
+#ifdef THREADED_CODE
+#  define Instruct(name) coq_lbl_##name: 
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/build/windows/difftar-folder.sh
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@cbda5049e77b1054b88d69e586a21a32885f0aa8 🚀

@@ -1,89 +1,0 @@
-#!/bin/bash
-
-###################### COPYRIGHT/COPYLEFT ######################
-
-# (C) 2016 Intel Deutschland GmbH
-# Author: Michael Soegtrop
-#
-# Released to the public by Intel under the
-# GNU Lesser General Public License Version 2.1 or later
-# See https://www.gnu.org/licenses/old-licenses/lgpl-2.1.html
-
-###################### DIFF A TAR FILE AND A FOLDER ######################
-
-set -o nounset
-
-# Print usage
-
-if [ "$#" -lt 2 ] ; then
-  echo 'Diff a tar (or compressed tar) file with a folder'
-  echo 'difftar-folder.sh <tarfile> <folder> [strip]'
-  echo '<tarfile> is the name of the tar file do diff with (required)'
-  echo '<folder> is the name of the folder to diff with (required)'
-  echo '<strip> is the number of path components to strip from tar file (default is 0)'
-  echo 'All files in the tar file must have at least <strip> path components.'
-  echo 'This also adds new files from folder.new, if folder.new exists'
-  exit 1
-fi
-
-# Parse parameters
-
-tarfile=$1
-folder=$2
-
-if [ "$#" -ge 3 ] ; then
-  strip=$3
-else
-  strip=0
-fi
-
-# Get path prefix if --strip is used
-
-if [ "$strip" -gt 0 ] ; then
-  # Get the path/name of the first file from teh tar and extract the first $strip path components
-  # This assumes that the first file in the tar file has at least $strip many path components
-  prefix=$(tar -t -f $tarfile | head -1 | cut -d / -f -$strip)/
-else
-  prefix=
-fi
-
-# Original folder
-
-orig=$folder.orig
-mkdir -p "$orig"
-
-# New amd empty filefolder
-
-new=$folder.new
-empty=$folder.empty
-mkdir -p "$empty"
-
-# Print information (this is ignored by patch)
-
-echo diff/patch file created on $(date) with:
-echo difftar-folder.sh $@
-echo TARFILE=    $tarfile
-echo FOLDER=     $folder
-echo TARSTRIP=   $strip
-echo TARPREFIX=  $prefix
-echo ORIGFOLDER= $orig
-
-# Make sure tar uses english output (for Mod time differs)
-export LC_ALL=C
-
-# Search all files with a deviating modification time using tar --diff
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/appveyor.sh

@@ -1,9 +1,0 @@
-#!/bin/bash
-set -e -x
-wget https://github.com/fdopen/opam-repository-mingw/releases/download/0.0.0.1/opam64.tar.xz
-tar -xf opam64.tar.xz
-bash opam64/install.sh
-opam init -a mingw https://github.com/fdopen/opam-repository-mingw.git --comp 4.02.3+mingw64c --switch 4.02.3+mingw64c
-eval $(opam config env)
-opam install -y ocamlfind camlp5
-cd $APPVEYOR_BUILD_FOLDER && ./configure -local && make && make byte && make -C test-suite all INTERACTIVE= && make validate

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-basic-overlay.sh

@@ -1,152 +1,0 @@
-#!/usr/bin/env bash
-
-# This is the basic overlay set for repositories in the CI.
-
-# Maybe we should just use Ruby to have real objects...
-
-# : "${foo:=bar}" sets foo to "bar" if it is unset or null
-
-########################################################################
-# MathComp
-########################################################################
-: "${mathcomp_CI_BRANCH:=master}"
-: "${mathcomp_CI_GITURL:=https://github.com/math-comp/math-comp.git}"
-
-########################################################################
-# UniMath
-########################################################################
-: "${UniMath_CI_BRANCH:=master}"
-: "${UniMath_CI_GITURL:=https://github.com/UniMath/UniMath.git}"
-
-########################################################################
-# Unicoq + Metacoq
-########################################################################
-: "${unicoq_CI_BRANCH:=master}"
-: "${unicoq_CI_GITURL:=https://github.com/unicoq/unicoq.git}"
-
-: "${metacoq_CI_BRANCH:=master}"
-: "${metacoq_CI_GITURL:=https://github.com/MetaCoq/MetaCoq.git}"
-
-########################################################################
-# Mathclasses + Corn
-########################################################################
-: "${math_classes_CI_BRANCH:=master}"
-: "${math_classes_CI_GITURL:=https://github.com/math-classes/math-classes.git}"
-
-: "${Corn_CI_BRANCH:=master}"
-: "${Corn_CI_GITURL:=https://github.com/c-corn/corn.git}"
-
-########################################################################
-# Iris
-########################################################################
-: "${stdpp_CI_BRANCH:=master}"
-: "${stdpp_CI_GITURL:=https://gitlab.mpi-sws.org/robbertkrebbers/coq-stdpp.git}"
-
-: "${Iris_CI_BRANCH:=master}"
-: "${Iris_CI_GITURL:=https://gitlab.mpi-sws.org/FP/iris-coq.git}"
-
-: "${lambdaRust_CI_BRANCH:=master}"
-: "${lambdaRust_CI_GITURL:=https://gitlab.mpi-sws.org/FP/LambdaRust-coq.git}"
-
-########################################################################
-# HoTT
-########################################################################
-: "${HoTT_CI_BRANCH:=master}"
-: "${HoTT_CI_GITURL:=https://github.com/HoTT/HoTT.git}"
-
-########################################################################
-# Ltac2
-########################################################################
-: "${ltac2_CI_BRANCH:=v8.8}"
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-bignums.sh

@@ -1,16 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-
-# This script could be included inside other ones
-# Let's avoid to source ci-common twice in this case
-if [ -z "${CI_BUILD_DIR}" ];
-then
-    source ${ci_dir}/ci-common.sh
-fi
-
-bignums_CI_DIR=${CI_BUILD_DIR}/Bignums
-
-git_checkout ${bignums_CI_BRANCH} ${bignums_CI_GITURL} ${bignums_CI_DIR}
-
-( cd ${bignums_CI_DIR} && make && make install)

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-coq-dpdgraph.sh

@@ -1,10 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-coq_dpdgraph_CI_DIR=${CI_BUILD_DIR}/coq-dpdgraph
-
-git_checkout ${coq_dpdgraph_CI_BRANCH} ${coq_dpdgraph_CI_GITURL} ${coq_dpdgraph_CI_DIR}
-
-( cd ${coq_dpdgraph_CI_DIR} && autoconf && ./configure && make && make test-suite )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-geocoq.sh

@@ -1,12 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-GeoCoq_CI_DIR=${CI_BUILD_DIR}/GeoCoq
-
-git_checkout ${GeoCoq_CI_BRANCH} ${GeoCoq_CI_GITURL} ${GeoCoq_CI_DIR}
-
-( cd ${GeoCoq_CI_DIR}                                     && \
-  ./configure-ci.sh                                       && \
-  make )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-math-classes.sh

@@ -1,10 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-math_classes_CI_DIR=${CI_BUILD_DIR}/math-classes
-
-git_checkout ${math_classes_CI_BRANCH} ${math_classes_CI_GITURL} ${math_classes_CI_DIR}
-
-( cd ${math_classes_CI_DIR} && make && make install )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-math-comp.sh

@@ -1,15 +1,0 @@
-#!/usr/bin/env bash
-
-# $0 is not the safest way, but...
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-mathcomp_CI_DIR=${CI_BUILD_DIR}/math-comp
-
-checkout_mathcomp ${mathcomp_CI_DIR}
-
-# odd_order takes too much time for travis.
-( cd ${mathcomp_CI_DIR}/mathcomp                  && \
-  sed -i.bak '/PFsection/d'                  Make && \
-  sed -i.bak '/stripped_odd_order_theorem/d' Make && \
-  make Makefile.coq && make -f Makefile.coq all )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-vst.sh

@@ -1,13 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-VST_CI_DIR=${CI_BUILD_DIR}/VST
-
-# opam install -j ${NJOBS} -y menhir
-git_checkout ${VST_CI_BRANCH} ${VST_CI_GITURL} ${VST_CI_DIR}
-
-# Targets are: msl veric floyd progs , we remove progs to save time
-# Patch to avoid the upper version limit
-( cd ${VST_CI_DIR} && make IGNORECOQVERSION=true .loadpath version.vo msl veric floyd )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/user-overlays/00669-maximedenes-ssr-merge.sh

@@ -1,4 +1,0 @@
-if [ "$CI_PULL_REQUEST" = "669" ] || [ "$CI_BRANCH" = "ssr-merge" ]; then
-    mathcomp_CI_BRANCH=ssr-merge
-    mathcomp_CI_GITURL=https://github.com/maximedenes/math-comp.git
-fi

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/tools/backport-pr.sh

@@ -1,74 +1,0 @@
-#!/usr/bin/env bash
-
-# Usage: dev/tools/backport-pr.sh <PR number> [--stop-before-merging]
-
-set -e
-
-PRNUM=$1
-OPTION=$2
-
-if ! git log master --grep "Merge PR #${PRNUM}" | grep "." > /dev/null; then
-    echo "PR #${PRNUM} does not exist."
-    exit 1
-fi
-
-SIGNATURE_STATUS=$(git log master --grep "Merge PR #${PRNUM}" --format="%G?")
-git log master --grep "Merge PR #${PRNUM}" --format="%GG"
-if [[ "${SIGNATURE_STATUS}" != "G" ]]; then
-    echo
-    read -p "Merge commit does not have a good (valid) signature. Bypass? [y/N] " -n 1 -r
-    echo
-    if [[ ! $REPLY =~ ^[Yy]$ ]]; then
-        exit 1
-    fi
-fi
-
-BRANCH=backport-pr-${PRNUM}
-RANGE=$(git log master --grep "Merge PR #${PRNUM}" --format="%P" | sed 's/ /../')
-MESSAGE=$(git log master --grep "Merge PR #${PRNUM}" --format="%s" | sed 's/Merge/Backport/')
-
-if git checkout -b ${BRANCH}; then
-
-    if ! git cherry-pick -x ${RANGE}; then
-        echo "Please fix the conflicts, then exit."
-        bash
-        while ! git cherry-pick --continue; do
-            echo "Please fix the conflicts, then exit."
-            bash
-        done
-    fi
-    git checkout -
-
-else
-
-    echo
-    read -p "Skip directly to merging phase? [y/N] " -n 1 -r
-    echo
-    if [[ ! $REPLY =~ ^[Yy]$ ]]; then
-        exit 1
-    fi
-
-fi
-
-if ! git diff --exit-code HEAD ${BRANCH} -- "*.mli"; then
-    echo
-    read -p "Some mli files are modified. Bypass? [y/N] " -n 1 -r
-    echo
-    if [[ ! $REPLY =~ ^[Yy]$ ]]; then
-        exit 1
-    fi
-fi
-
-if [[ "${OPTION}" == "--stop-before-merging" ]]; then
-    exit 0
-fi
-
-git merge -S --no-ff ${BRANCH} -m "${MESSAGE}"
-git branch -d ${BRANCH}
-
-# To-Do:
-# - Support for backporting a PR before it is merged
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/doc/tools/coqrst/coqdoc/main.py

@@ -1,86 +1,0 @@
-##########################################################################
-##         #   The Coq Proof Assistant / The Coq Development Team       ##
-##  v      #   INRIA, CNRS and contributors - Copyright 1999-2018       ##
-## <O___,, #       (see CREDITS file for the list of authors)           ##
-##   \VV/  ###############################################################
-##    //   #    This file is distributed under the terms of the         ##
-##         #     GNU Lesser General Public License Version 2.1          ##
-##         #     (see LICENSE file for the text of the license)         ##
-##########################################################################
-"""
-Use CoqDoc to highlight Coq snippets
-====================================
-
-Pygment's Coq parser isn't the best; instead, use coqdoc to highlight snippets.
-Only works for full, syntactically valid sentences; on shorter snippets, coqdoc
-swallows parts of the input.
-
-Works by reparsing coqdoc's output into the output that Sphinx expects from a
-lexer.
-"""
-
-import os
-from tempfile import mkstemp
-from subprocess import check_output
-
-from bs4 import BeautifulSoup
-from bs4.element import NavigableString
-
-COQDOC_OPTIONS = ['--body-only', '--no-glob', '--no-index', '--no-externals',
-                  '-s', '--html', '--stdout', '--utf8']
-
-COQDOC_SYMBOLS = ["->", "<-", "<->", "=>", "<=", ">=", "<>", "~", "/\\", "\\/", "|-", "*", "forall", "exists"]
-COQDOC_HEADER = "".join("(** remove printing {} *)".format(s) for s in COQDOC_SYMBOLS)
-
-def coqdoc(coq_code, coqdoc_bin = os.path.join(os.getenv("COQBIN"),"coqdoc")):
-    """Get the output of coqdoc on coq_code."""
-    fd, filename = mkstemp(prefix="coqdoc-", suffix=".v")
-    try:
-        os.write(fd, COQDOC_HEADER.encode("utf-8"))
-        os.write(fd, coq_code.encode("utf-8"))
-        os.close(fd)
-        return check_output([coqdoc_bin] + COQDOC_OPTIONS + [filename], timeout = 2).decode("utf-8")
-    finally:
-        os.remove(filename)
-
-def is_whitespace_string(elem):
-    return isinstance(elem, NavigableString) and elem.strip() == ""
-
-def strip_soup(soup, pred):
-    """Strip elements maching pred from front and tail of soup."""
-    while soup.contents and pred(soup.contents[-1]):
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/doc/tools/coqrst/repl/coqtop.py

@@ -1,101 +1,0 @@
-##########################################################################
-##         #   The Coq Proof Assistant / The Coq Development Team       ##
-##  v      #   INRIA, CNRS and contributors - Copyright 1999-2018       ##
-## <O___,, #       (see CREDITS file for the list of authors)           ##
-##   \VV/  ###############################################################
-##    //   #    This file is distributed under the terms of the         ##
-##         #     GNU Lesser General Public License Version 2.1          ##
-##         #     (see LICENSE file for the text of the license)         ##
-##########################################################################
-"""
-Drive coqtop with Python!
-=========================
-
-This module is a simple pexpect-based driver for coqtop, based on the old
-REPL interface.
-"""
-
-import os
-import re
-
-import pexpect
-
-class CoqTop:
-    """Create an instance of coqtop.
-
-    Use this as a context manager: no instance of coqtop is created until
-    you call `__enter__`.  coqtop is terminated when you `__exit__` the
-    context manager.
-
-    Sentence parsing is very basic for now (a "." in a quoted string will
-    confuse it).
-    """
-
-    COQTOP_PROMPT = re.compile("\r\n[^< ]+ < ")
-
-    def __init__(self, coqtop_bin=None, color=False, args=None) -> str:
-        """Configure a coqtop instance (but don't start it yet).
-
-        :param coqtop_bin: The path to coqtop; uses $COQBIN by default, falling back to "coqtop"
-        :param color:      When True, tell coqtop to produce ANSI color codes (see
-                           the ansicolors module)
-        :param args:       Additional arugments to coqtop.
-        """
-        self.coqtop_bin = coqtop_bin or os.path.join(os.getenv('COQBIN'),"coqtop")
-        self.args = (args or []) + ["-boot", "-color", "on"] * color
-        self.coqtop = None
-
-    def __enter__(self):
-        if self.coqtop:
-            raise ValueError("This module isn't re-entrant")
-        self.coqtop = pexpect.spawn(self.coqtop_bin, args=self.args, echo=False, encoding="utf-8")
-        # Disable delays (http://pexpect.readthedocs.io/en/stable/commonissues.html?highlight=delaybeforesend)
-        self.coqtop.delaybeforesend = 0
-        self.next_prompt()
-        return self
-
-    def __exit__(self, type, value, traceback):
-        self.coqtop.kill(9)
-
-    def next_prompt(self):
-        "Wait for the next coqtop prompt, and return the output preceeding it."
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/kernel/byterun/coq_fix_code.c

@@ -1,170 +1,0 @@
-/***********************************************************************/
-/*                                                                     */
-/*                           Coq Compiler                              */
-/*                                                                     */
-/*        Benjamin Gregoire, projets Logical and Cristal               */
-/*                        INRIA Rocquencourt                           */
-/*                                                                     */
-/*                                                                     */
-/***********************************************************************/
-
-/* Arnaud Spiwack: expanded the virtual machine with operators used
-   for fast computation of bounded (31bits) integers */
-
-#include <stdio.h>
-#include <stdlib.h> 
-#include <stdint.h>
-#include <caml/config.h>
-#include <caml/misc.h>
-#include <caml/mlvalues.h>
-#include <caml/fail.h>
-#include <caml/memory.h>
-#include "coq_instruct.h"
-#include "coq_fix_code.h"
-
-#ifdef THREADED_CODE
-char ** coq_instr_table;
-char * coq_instr_base;
-int arity[STOP+1];
-
-void init_arity () {
-  /* instruction with zero operand */
-  arity[ACC0]=arity[ACC1]=arity[ACC2]=arity[ACC3]=arity[ACC4]=arity[ACC5]=
-    arity[ACC6]=arity[ACC7]=arity[PUSH]=arity[PUSHACC0]=arity[PUSHACC1]=
-    arity[PUSHACC2]=arity[PUSHACC3]=arity[PUSHACC4]=arity[PUSHACC5]=
-    arity[PUSHACC6]=arity[PUSHACC7]=arity[ENVACC1]=arity[ENVACC2]=
-    arity[ENVACC3]=arity[ENVACC4]=arity[PUSHENVACC1]=arity[PUSHENVACC2]=
-    arity[PUSHENVACC3]=arity[PUSHENVACC4]=arity[APPLY1]=arity[APPLY2]=
-    arity[APPLY3]=arity[RESTART]=arity[OFFSETCLOSUREM2]=
-    arity[OFFSETCLOSURE0]=arity[OFFSETCLOSURE2]=arity[PUSHOFFSETCLOSUREM2]=
-    arity[PUSHOFFSETCLOSURE0]=arity[PUSHOFFSETCLOSURE2]=
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/kernel/byterun/coq_memory.c

@@ -1,212 +1,0 @@
-/***********************************************************************/
-/*                                                                     */
-/*                           Coq Compiler                              */
-/*                                                                     */
-/*        Benjamin Gregoire, projets Logical and Cristal               */
-/*                        INRIA Rocquencourt                           */
-/*                                                                     */
-/*                                                                     */
-/***********************************************************************/
-
-#include <stdio.h>
-#include <string.h> 
-#include "coq_gc.h"
-#include "coq_instruct.h"
-#include "coq_fix_code.h"
-#include "coq_memory.h"
-#include "coq_interp.h"
-
-/* stack */
-
-value * coq_stack_low;
-value * coq_stack_high;
-value * coq_stack_threshold;
-asize_t coq_max_stack_size = Coq_max_stack_size;
-/* global_data */
-
-
-value coq_global_data;
-value coq_atom_tbl;
-
-int drawinstr;
-/* interp state */
-
-long coq_saved_sp_offset;
-value * coq_sp;
-/* Some predefined pointer code */ 
-code_t accumulate;
-
-/* functions over global environment */
-
-void coq_stat_free (void * blk)
-{
-  free (blk);
-}
-
-value coq_static_alloc(value size) /* ML */
-{
-  return (value) coq_stat_alloc((asize_t) Long_val(size));
-}
-
-value accumulate_code(value unit) /* ML */
-{
-  return (value) accumulate;
-}
-
-static void (*coq_prev_scan_roots_hook) (scanning_action);
-
-static void coq_scan_roots(scanning_action action)
-{
-  register value * i;
-  /* Scan the global variables */
-  (*action)(coq_global_data, &coq_global_data);
-  (*action)(coq_atom_tbl, &coq_atom_tbl);
-  /* Scan the stack */
-  for (i = coq_sp; i < coq_stack_high; i++) {
-    (*action) (*i, i);
-  };
-  /* Hook */
-  if (coq_prev_scan_roots_hook != NULL) (*coq_prev_scan_roots_hook)(action);
-
-
-}
-
-void init_coq_stack()
-{
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/coq-makefile/template/path-init.sh

@@ -1,5 +1,0 @@
-set -e
-set -o pipefail
-
-export PATH="$COQBIN:$PATH"
-export LC_ALL=C

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/coq-makefile/timing/precomputed-time-tests/run.sh

@@ -1,10 +1,0 @@
-#!/usr/bin/env bash
-
-set -x
-set -e
-
-cd "$(dirname "${BASH_SOURCE[0]}")"
-export COQLIB="$(cd ../../../.. && pwd)"
-
-./001-correct-diff-sorting-order/run.sh || exit $?
-./002-single-file-sorting/run.sh || exit $?

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/misc/deps-utf8.sh

@@ -1,17 +1,0 @@
-# Check reading directories matching non pure ascii idents
-# See bug #5715 (utf-8 working on macos X and linux)
-# Windows is still not compliant
-a=`uname`
-if [ "$a" = "Darwin" -o "$a" = "Linux" ]; then
-rm -f misc/deps/théorèmes/*.v
-tmpoutput=`mktemp /tmp/coqcheck.XXXXXX`
-$coqc -R misc/deps AlphaBêta misc/deps/αβ/γδ.v
-R=$?
-$coqtop -R misc/deps AlphaBêta -load-vernac-source misc/deps/αβ/εζ.v
-S=$?
-if [ $R = 0 -a $S = 0 ]; then
-    exit 0
-else
-    exit 1
-fi
-fi

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/misc/printers.sh

@@ -1,3 +1,0 @@
-printf "Drop. #use\"include\";; #quit;;\n" | $coqtopbyte 2>&1 | egrep "Error|Unbound"
-if [ $? = 0 ]; then exit 1; else exit 0; fi
-

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/misc/universes.sh

@@ -1,8 +1,0 @@
-# Sort universes for the whole standard library
-EXPECTED_UNIVERSES=4 # Prop is not counted
-$coqc -R misc/universes Universes misc/universes/all_stdlib 2>&1
-$coqc -R misc/universes Universes misc/universes/universes 2>&1
-mv universes.txt misc/universes
-N=`awk '{print $3}' misc/universes/universes.txt | sort -u | wc -l`
-printf "Found %s/%s universes\n" $N $EXPECTED_UNIVERSES
-if [ "$N" -eq $EXPECTED_UNIVERSES ]; then exit 0; else exit 1; fi

### 2021-08-15 resume/Resume.tex
Revert "Deploying to gh-pages from @ JasonGross/JasonGross.github.io@cbda5049e77b1054b88d69e586a21a32885f0aa8 🚀"
This reverts commit 8d3ef534c8997c7ef4e469dcf58ccdd69af0346b.

% LaTeX file for resume
% This file uses the resume document class (res.cls)

\documentclass[11pt]{res}
\usepackage[T1]{fontenc}
\usepackage{xifthen,verbatim}
\usepackage{hyperref}
\usepackage{etoolbox}[2011/01/03]

\usepackage[backend=biber,bibencoding=utf-8,maxbibnames=99,sorting=ymdnt,noerroretextools=true,style=alphabetic]{biblatex}
% http://tex.stackexchange.com/a/140641
\DeclareSortingTemplate{ymdnt}{
    \sort{
        \field{presort}
    }
    \sort[final]{
        \field{sortkey}
    }
    \sort[direction=descending]{
        \field{sortyear}
        \field{year}
        \literal{9999}
    }
    \sort[direction=descending]{
        \field[padside=left,padwidth=2,padchar=0]{month}
        \literal{99}
    }
    \sort[direction=descending]{
        \field[padside=left,padwidth=2,padchar=0]{day}
        \literal{99}
    }
    \sort{
        \field{sortname}
        \field{author}
        \field{editor}
        \field{translator}
        \field{sorttitle}
        \field{title}
    }
    \sort{
        \field{sorttitle}
        \field{title}
    }
}

\makeatletter

\let\origifempty\ifempty
\undef\ifempty
\let\origforlistloop\forlistloop
\usepackage{etextools}
\let\eifempty\ifempty
\let\ifempty\origifempty
\let\forlistloop\origforlistloop
\def\preetextools{\let\ifempty\eifempty}

\newif\ifmultitexcompile
\multitexcompiletrue

\newcommand\multitex@separator{-}
\newwrite\multitex@temp
% \immediate\openout\multitex@temp{\jobname.mtx}
\newtoks\multitex@names
\multitex@names={}
\newcommand{\multitex@postprocess}[1]{%
  \immediate\expandafter\closeout\csname @multitex@#1@file\endcsname
  \ifmultitexcompile
    \immediate\write18{pdflatex \csname @multitex@#1@name\endcsname}%
    \immediate\write18{biber \csname @multitex@#1@barename\endcsname}%
    \immediate\write18{pdflatex \csname @multitex@#1@name\endcsname}%
  \fi
}
\newcommand{\onlymultitex}[1]{\FE@ifstar{#1}{\only@other@multitex}{\only@multitex}}
\newcommand{\only@multitex}[2]{#2}
\newcommand{\only@other@multitex}[2]{}
{\catcode`\#=12
  \global\def\@pound{#}
}
\newcommand{\DeclareFile}[2][]{
  \global\multitex@names=\expandafter{\the\multitex@names,#2}
  \expandafter\xdef\csname @multitex@#2\endcsname{\xifempty{#1}{#2}{#1}}
  \expandafter\xdef\csname @multitex@#2@name\endcsname{"\jobname\multitex@separator\csname @multitex@#2\endcsname.tex"}
  \expandafter\xdef\csname @multitex@#2@auxname\endcsname{"\jobname\multitex@separator\csname @multitex@#2\endcsname.aux"}
  \expandafter\xdef\csname @multitex@#2@barename\endcsname{"\jobname\multitex@separator\csname @multitex@#2\endcsname"}
  \expandafter\newwrite\csname @multitex@#2@file\endcsname
  \immediate\expandafter\openout\csname @multitex@#2@file\expandafter\endcsname\csname @multitex@#2@name\endcsname
  \AtBeginDocument{%
    \multitex@write{#2}{%
      \detokenize{\newcommand{\onlymultitex}[1]}%
        {\detokenize{\csname FE@ifstar\endcsname}%
          {\@pound1}%
          \detokenize{%
            {\csname only@multitex\endcsname}{\csname only@multitex\endcsname}%
          }%
        }%
      \detokenize{%
        \expandafter\newcommand\csname only@multitex\endcsname[2]}%
        {%
          \noexpand\renewcommand*{\noexpand\do}[1]{%
            \noexpand\ifstrequal{\@pound\@pound1}{#2}%
              {\@pound2\noexpand\listbreak}%
              {}%
            }%
          {\noexpand\edef\noexpand\foo{\@pound1}\noexpand\expandafter}%
          \noexpand\expandafter\noexpand\docsvlist\noexpand\expandafter{\noexpand\foo}%
        }
      \detokenize{%
        \expandafter\let\csname only@other@multitex\expandafter\endcsname\csname only@multitex\endcsname
      }%
    }%
  }
  \AtEndDocument{\multitex@postprocess{#2}}
}
\newcommand{\multitex@write}[2]{\xifblank{#1}{}{\immediate\expandafter\write\csname @multitex@#1@file\endcsname{#2}}}
\edef\multitex@tempfilename{\jobname.mtx}
\newcommand{\multitex@pre@writetothis}{\immediate\openout\multitex@temp\multitex@tempfilename}%\long\gdef\multitex@thiscode{}}%
\newcommand{\multitex@writetothis}[1]{\immediate\write\multitex@temp{#1}}%\ExpandNext{\long\gappto\multitex@thiscode}{\expandnext{\relax\scantokens}{#1}}}%
\newcommand{\multitex@post@writetothis}{\immediate\closeout\multitex@temp\relax\input{\multitex@tempfilename}}%\show\multitex@thiscode\multitex@thiscode\relax}%
\newcommand{\multitex@writetoothers}[2][\the\multitex@names]{{\preetextools\ExpandNext{\forcsvloop}{#1}\do{\multitex@write{##1}{#2}}}}
\newcommand{\DeclareDocumentClass}[1]{\multitex@writetoothers{\detokenize{#1}}}
\AtBeginDocument{\multitex@writetoothers{\detokenize{\begin{document}}}}
\AtEndDocument{\multitex@writetoothers{\detokenize{\end{document}}}}
\newcommand{\includein}[2][]{\multitex@pre@writetothis\multitex@writetothis{\detokenize{#2}}\multitex@post@writetothis\multitex@writetoothers[#1]{\detokenize{#2}}}
\AfterEndEnvironment{multitex}{\multitex@post@writetothis}
\newenvironment{multitex}[1][\the\multitex@names]{%
  \@bsphack
  \begingroup% Lets Keep the Changes Local
    \multitex@pre@writetothis
    \let\do\@makeother\dospecials\catcode`\^^M\active
    \def\verbatim@processline{\multitex@writetothis{\the\verbatim@line}\multitex@writetoothers[#1]{\the\verbatim@line}}%
    \verbatim@start\space % I'm not sure why I need the space, but I seem to need it.
}{\endgroup
  \@esphack
}
% only others
\newenvironment{multitex*}[1][\the\multitex@names]{%
  \@bsphack
  \begingroup% Lets Keep the Changes Local
    \let\do\@makeother\dospecials\catcode`\^^M\active
    \def\verbatim@processline{\multitex@writetoothers[#1]{\the\verbatim@line}}%
    \verbatim@start\space % I'm not sure why I need the space, but I seem to need it.
}{\endgroup
  \@esphack
}
\makeatother

\DeclareFile{curriculum-vitae}
\DeclareFile{mathematics}
\DeclareFile{computer-science}
\DeclareFile{curriculum-vitae-long}
\DeclareFile{curriculum-vitae-mathematics}
\DeclareFile{curriculum-vitae-computer-science}
\DeclareDocumentClass{\documentclass[11pt]{res}}

\begin{multitex}
  \edef\resumefiles{mathematics,computer-science}
  \edef\cvshortspecialfiles{curriculum-vitae-mathematics,curriculum-vitae-computer-science}
  \edef\cvshortfiles{curriculum-vitae,\cvshortspecialfiles}
  \edef\cvlongfiles{curriculum-vitae-long}
  \edef\cvfiles{\cvshortfiles,\cvlongfiles}
  \edef\highschoolfiles{curriculum-vitae-long}
  \edef\mathfiles{mathematics,curriculum-vitae-mathematics}
  \edef\cvlongandmathfiles{\mathfiles,curriculum-vitae-long}
  \edef\cvandmathfiles{\cvlongandmathfiles,curriculum-vitae,curriculum-vitae-computer-science}
  \edef\computersciencefiles{computer-science,curriculum-vitae-computer-science}
  \edef\cvlongandcomputersciencefiles{\computersciencefiles,curriculum-vitae-long}
  \edef\cvandcomputersciencefiles{curriculum-vitae,curriculum-vitae-mathematics,\cvlongandcomputersciencefiles}
  \edef\allexceptcv{\resumefiles,\cvshortspecialfiles}
  \edef\allexceptcvlongfiles{\resumefiles,\cvshortfiles}
  \edef\allexceptcvandcvlongfiles{\resumefiles,\cvshortspecialfiles}
  %\usepackage[utf8x]{inputenc}
  %\usepackage[T1]{fontenc}
  %\usepackage{helvetica} % uses helvetica postscript font (download helvetica.sty)
  %\usepackage{newcent}   % uses new century schoolbook postscript font
  \usepackage[T1]{fontenc}
  \usepackage[margin=1in,top=0.5in,bottom=0.5in]{geometry}
  \usepackage{hyperref}
  \usepackage{calc}
  \usepackage{amsmath}
  \usepackage{etoolbox}[2011/01/03]
  %\bibliographystyle{plainyr-rev}
  \usepackage[backend=biber,maxbibnames=99,sorting=ymdnt,noerroretextools=true,style=alphabetic]{biblatex}
% http://tex.stackexchange.com/a/140641
\DeclareSortingTemplate{ymdnt}{
    \sort{
        \field{presort}
    }
    \sort[final]{
        \field{sortkey}
    }
    \sort[direction=descending]{
        \field{sortyear}
        \field{year}
        \literal{9999}
    }
    \sort[direction=descending]{
        \field[padside=left,padwidth=2,padchar=0]{month}
        \literal{99}
    }
    \sort[direction=descending]{
        \field[padside=left,padwidth=2,padchar=0]{day}
        \literal{99}
    }
    \sort{
        \field{sortname}
        \field{author}
        \field{editor}
        \field{translator}
        \field{sorttitle}
        \field{title}
    }
    \sort{
        \field{sorttitle}
        \field{title}
    }
}
  \addbibresource{jason-gross.bib}
  \DeclareBibliographyCategory{exclude}
  \makeatletter
  \@ifpackageloaded{etextools}{%
  }{%
    \let\origifempty\ifempty
    \let\origforlistloop\forlistloop
    \undef\ifempty
    \usepackage{etextools}
    \let\eifempty\ifempty
    \let\ifempty\origifempty
    \let\forlistloop\origforlistloop
    \def\preetextools{\let\ifempty\eifempty}
  }%
  \makeatother
  \begin{format}
    \employer{l}\dates{r}\\
    \title{l}\location{r} \\
    \body
  \end{format}
  \let\oldemployer=\employer
  \let\oldtitle=\title
  \renewcommand{\employer}[1]{\oldemployer{\textbf{#1}}}
  \renewcommand{\title}[1]{\oldtitle{\textit{#1}}}

  % \let\oldposition=\position
  % \let\oldendposition=\endposition
  % \renewenvironment{position}{\oldposition\vspace*{-2\baselineskip}}{\endoldposition}

  \makeatletter
  \let\old@itemize=\itemize
  \def\itemize{\old@itemize\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}\setlength{\leftskip}{-1em}}
  \renewenvironment{position}
    {%
    \begingroup
      \par
        \the\tabular@head
  %     \addpenalty{-\@secpenalty}% bad place for a page break
      \penalty -\@secpenalty % bad place for a page break
      \penalty 10000
      \ignorespaces
      \vspace*{-\baselineskip}
    }{%
        \the\tabular@tail
  %     \addpenalty{\@secpenalty}% good place for a page break
      \penalty \@secpenalty % good place for a page break
      \endgroup
  }
  \makeatother

  \newsectionwidth{1em}
  \resumewidth=7.5in

  \newcommand{\aswidthof}[2]{\rlap{#1}\hphantom{#2}}
\end{multitex}
\begin{multitex*}[\cvlongfiles]
  \name{\LARGE \bf \hspace{\stretch{1}} Jason Gross \hspace{\stretch{1}} \\
    \hspace{\stretch{1}}\large \href{mailto:jgross@mit.edu}{jgross@mit.edu} --- (631) 790-8962 --- \url{https://people.csail.mit.edu/jgross/}\hspace{\stretch{1}} \\}

  \address{\textsc{Present Address}\\258 Prospect Street, Apt \# 1L\\Cambridge, MA 02139}
  \address{\textsc{Permanent Address}\\126 Hayrick Lane \\  Commack, NY 11725}
\end{multitex*}
\begin{multitex}[\allexceptcvlongfiles]
  \name{\LARGE \bf \parbox{\widthof{github.com/JasonGross}}{%
      \centering Jason Gross \\
      \vspace{0.2cm}
      \large \href{https://github.com/JasonGross/}{github.com/JasonGross} \\
      \href{https://people.csail.mit.edu/jgross/}{people.csail.mit.edu/jgross} \\
      $\left.\right.$ \\
      $\left.\right.$ \\%
  }}

  \vspace{5cm}

  %\address{}%\textsc{Address}\\258 Prospect Street, Apt \# 1L\\Cambridge, MA 02139}
  %\address{\textsc{Contact}\\\href{mailto:jgross@mit.edu}{jgross@mit.edu}\\(631) 790-8962}
  \newcommand{\addressaswidth}[1]{\makebox[\widthof{\textsc{jgross@mit.edu}}][c]{#1}}
  \address{\addressaswidth{\textsc{Contact}}\\\addressaswidth{\href{mailto:jgross@mit.edu}{jgross@mit.edu}}\\\addressaswidth{(631) 790-8962}}
\end{multitex}

\begin{document}

\begin{multitex}
  \resume
\end{multitex}

\begin{multitex*}[\allexceptcvandcvlongfiles]
  \section{\textsc{Research Interests}}
  Programming Languages, Formal Verification, Cryptography, Performance of Automation in Interactive Proof Assistants, Homotopy Type Theory, Category Theory, Program Synthesis
\end{multitex*}
\begin{multitex}[curriculum-vitae]
  \section{\textsc{Research Interests}}
  Programming Languages, Formal Verification, Cryptography, Performance Engineering
\end{multitex}

\begin{multitex}
  %\section{JOB OBJECTIVE}
    %A summer position that will use my accounting and computer
    %skills.
    %
  \section{\textsc{Education}}
\end{multitex}

\begin{multitex}[\allexceptcvlongfiles]
  \employer{Massachusetts Institute of Technology}
  \dates{2013--2021}
  \title{{\rm PhD in Computer Science}}
  \location{Cambridge, MA}
  \begin{position}\\
      Advisor: Adam Chlipala\\
    \textit{Thesis: Performance Engineering of Proof-Based Software Systems at Scale}\\
    \textit{SM Thesis: An Extensible Framework for Synthesizing Efficient, Verified Parsers}
  \end{position}

  \employer{Massachusetts Institute of Technology}
  \dates{2009--2013}
  \title{{\rm BS in Mathematics and Physics}}
  \location{Cambridge, MA}
  \begin{position}\\
      GPA: 4.6/5
  \end{position}
\end{multitex}

\begin{multitex*}[\cvlongfiles]
    \textbf{Massachusetts Institute of Technology}, Cambridge, MA  \\
    Began Doctorate of Philosophy in Computer Science in September 2013 \\
    \onlymultitex*{\cvfiles}{Research Interests: Programming Languages, Performance of Automation in Interactive Proof Assistants, Formal Verification, Homotopy Type Theory, Category Theory, Program Synthesis, Type Theory \\}
    G.P.A. 4.9/5.0 \\
    \\
    Bachelor of Science in Mathematics and Physics, June 2013   \\
    G.P.A. 4.6/5.0 \\
\end{multitex*}
\begin{multitex*}[\resumefiles]%
    \\
    Relevant Coursework: %Computer Systems
        Security, %Foundations of
        Program Analysis, Performance Engineering of Software% Systems
        , Statistical Physics, Quantum% Physics
        , Topology, Analysis, Waves \& Vibrations, Special Relativity% and the Physics of Spacetime%
        , \mbox{Algebra}% force no hyphenation
\end{multitex*}
\begin{multitex*}[\cvlongfiles]%
    \\
		Coursework:
		\begin{itemize}
			\item Computer Science: Inference and Information, Computer Systems Security, Geometric Computing, Foundations of Program Analysis, Performance Engineering of Software Systems, Structure and Interpretation of Computer Programs
			\item Mathematics: Paradox and Infinity, Category Theory for Scientists, Algebraic Topology I, Seminar in Topology, Introduction to Topology, Real Analysis, Abstract Algebra I \& II, Differential Equations, Calculus I \& II, Linear Algebra
			\item Physics: Modern Astrophysics, General Relativity (Graduate), Statistical Physics I \& II, Quantum Physics I, II, \& III, Classical Mechanics II, Waves and Vibrations, Special Relativity and the Physics of Spacetime
			\item Other Sciences: Introductory Biology, Introductory Chemistry, Design a Concentrated Solar Power Water Heater (seminar)
			\item Other: The Art and Science of Happiness, Metaphysics, Introduction to Musical Composition, Capitalism and Its Critics, Moral Psychology, Philosophy of Quantum Mechanics, Moral Problems and the Good Life, Philosophy of Love, Kaufman Teaching Certificate Program
		\end{itemize}
\end{multitex*}

% \item Presented ``Computational Higher Inductive Types: Computing with Custom Equalities'', October 6, 2013, The CSAIL Student Workshop

\begin{multitex*}[\cvlongfiles]
\section{\textsc{Experience}}
\end{multitex*}
\begin{multitex*}[\cvlongfiles]
  \employer{MIT CSAIL}
  \dates{September 2013--February 2021}
  \title{Researcher}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \item Researching program synthesis and verification with Adam Chlipala
      \item Collaboratively working on \onlymultitex*{\cvfiles}{implementing }one of the world's first algorithm-level-optimizing compilers
      \item Collaboratively implemented proven-correct crypto code now used by Google Chrome
    \end{itemize}
  \end{position}
\end{multitex*}

\begin{multitex}
\section{\textsc{Experience}}
\end{multitex}

\begin{multitex*}[\cvfiles]
  \employer{Coq Development Team, INRIA}
  \dates{June 2021--Present}
  \title{Core Team Member}
  \location{Nantes, France (remote)}
  \begin{position}
    \begin{itemize}
      \item Stress Testing, Bug Reporting, Bug Minimizer, Compatibility Assurance
      \item Engineering and maintaining a bug minimizer for automatically producing minimized stand-alone test-cases from buggy code
      \item Researching performance issues that impact scalability of automated verification
      \item Have reported the plurality of all-time bugs in Coq (since 2012)
    \end{itemize}
  \end{position}

  \employer{Machine Intelligence Research Institute}
  \dates{February 2021--Present}
  \title{Research Staff}
  \location{Berkeley, CA (remote)}
  \begin{position}
    \begin{itemize}
      \item Performing self-directed research into topics in fundamental programming language theory and mathematics%, bringing insights and understanding back to my supervisor
    \end{itemize}
  \end{position}

  \employer{MIT CSAIL}
  \dates{September 2013--February 2021}
  \title{PhD Researcher}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \item Main Research Project: Fiat Cryptography (\href{https://github.com/mit-plv/fiat-crypto}{github.com/mit-plv/fiat-crypto})
      \item Collaboratively implemented proven-correct cryptographic code now used by Google Chrome, and in the majority of secure connects from web browsers
      \item Led development of one of the world's first algorithm-level-optimizing compilers
      \item Wrote backends to C, Go, Java, and JSON; managed development of backends to Rust and Zig
    \end{itemize}
  \end{position}
\end{multitex*}

\begin{multitex}
\section{\textsc{Internships}}
\end{multitex}

\begin{multitex*}[\cvfiles]
  \employer{Machine Intelligence Research Institute}
  \dates{June 2019--August 2019}
  \title{Type Theory Intern}
  \location{Berkeley, CA}
  \begin{position}
    \begin{itemize}
      \item Worked on formalizing type theories and on proving things within proof assistants
    \end{itemize}
  \end{position}
\end{multitex*}
\begin{multitex}
  \bgroup
  \interlinepenalty 10000
  \employer{Google}
  \dates{June 2018--August 2018}
  \title{Software Engineering Intern}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \item Worked with BoringSSL on integration of proven-correct low-level ECC primitives into Chrome
    \end{itemize}
  \end{position}
  \egroup
\end{multitex}
\begin{multitex*}[\cvfiles]
  \bgroup
  \interlinepenalty 10000
  \employer{Google}
  \dates{June 2016--September 2016}
  \title{Software Engineering Intern}
  \location{Mountain View, CA}
  \begin{position}
    \begin{itemize}
      \item Formalized low-level ECC primitives with proofs of correctness
    \end{itemize}
  \end{position}
  \egroup
\end{multitex*}
\begin{multitex}[\cvlongandmathfiles]
  \employer{MIT}
  \dates{Fall 2009--Present}
  \title{Teacher}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \item Taught classes on \LaTeX, philosophy, linear algebra, and quantum mechanics for MIT Educational Studies Program's Splash, Spark, and Summer HSSP (High School Studies Program)
      \onlymultitex{\cvandmathfiles}{\item Teaching Assistant for 8.012 (Physics I) and 8.022 (Physics II) in Experimental Study Group}
    \end{itemize}
  \end{position}
\end{multitex}
\begin{multitex*}[\cvlongandcomputersciencefiles]
  \employer{MIT}
  \dates{September 2015--December 2015}
  \title{TA for 6.172 (Performance Engineering% of Software% Systems
  )}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \item Created and led recitations, taught students, and helped run class
      \item Analyzed and explained assembly output of \texttt{gcc -O3} to teach vectorization
    \end{itemize}
  \end{position}
\end{multitex*}
\begin{multitex*}[\cvlongfiles]
    \employer{MIRI}
    \dates{June 12--14, 2015}
    \title{Decision Theory Workshop Attendee}
    \location{Berkeley, CA}
    \begin{position}
        \begin{itemize}
            \item Formalized various versions of L\"ob's theorem in Agda and Coq
            \item Learned about and worked on problems on the frontiers of decision theory
        \end{itemize}
    \end{position}
\end{multitex*}
\begin{multitex}
  \employer{Microsoft Research}
  \dates{June 2014--August 2014}
  \title{Intern}
  \location{Cambridge, United Kingdom}
  \begin{position}
    \begin{itemize}
      \item \onlymultitex*{\allexceptcvlongfiles}{Collaboratively c}\onlymultitex*{\cvlongfiles}{C}reated a language for specifying input/output behavior of x86 assembly programs\onlymultitex*{\cvlongfiles}{with Andrew Kennedy and Nick Benton}; Verified the I/O behavior of a number of simple programs
      \onlymultitex*{\cvfiles}{\item Improved automation of the x86proved library}
    \end{itemize}
  \end{position}
\end{multitex}
\begin{multitex*}[\cvfiles]
    \employer{MIT CSAIL}
    \dates{April 2012--June 2014}
    \title{Researcher}
    \location{Cambridge, MA}
    \begin{position}
        \begin{itemize}
            \item Entered a significant amount of category theory into the automated proof assistant Coq \onlymultitex*{\cvfiles}{(\url{https://github.com/HoTT/HoTT/tree/master/theories/Categories})}
            \onlymultitex{\cvfiles}{\item Made progress towards an interface for databases and database migration on top of category theory in Coq\onlymultitex*{\cvlongfiles}{ with David Spivak and Adam Chlipala}}
            \onlymultitex*{\cvlongfiles}{\item Presented ``Building Database Management on top of Category Theory in Coq'', January 25, 2013, POPL 2013: 40th ACM SIGPLAN-SIGACT Symposium on Principles of Programming Languages}
            \onlymultitex*{\cvlongfiles}{\item Presented	 ``Experience Implementing a Performant Category-Theory Library in Coq'', July 14, 2014, Interactive Theorem Proving 2014 Conference}
        \end{itemize}
    \end{position}
\end{multitex*}
\begin{multitex*}[\cvfiles]
  \employer{MIT CSAIL}
  \dates{November 2009--September 2011}
  \title{Researcher}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \onlymultitex*{\cvfiles}{\item Designed from scratch a data collection webpage, collected data for, and helped with research \onlymultitex*{\cvlongfiles}{ of Brenden Lake, Ruslan Salakhutdinov, and Josh Tenenbaum, }on categorical and transfer learning\onlymultitex*{\cvlongfiles}{ (\url{https://jgross.scripts.mit.edu/alphabets/})}}
      \onlymultitex*{\cvlongfiles,\resumefiles}{%
      \item Co-author of ``One shot learning of simple visual concepts''\onlymultitex*{\cvfiles}{ published in \emph{Proceedings of the 33rd Annual Meeting of the Cognitive Science Society}}.
        }
    \end{itemize}
  \end{position}
\end{multitex*}
\begin{multitex*}[\cvlongfiles]
  \employer{MIT OCW}
    \dates{May--August 2010}
    \title{}
    \location{Cambridge, MA}
    \begin{position}
        \begin{itemize}
            \item Evaluated and categorized videos of Walter Lewin solving physics problems for updating the 8.01 (Physics I) OpenCourseWare (OCW) website.
        \end{itemize}
    \end{position}
\end{multitex*}
\begin{multitex*}[\cvfiles]
  \employer{Commack High School}
  \dates{Fall 2006--Summer 2009}
  \title{Independent Researcher}
  \location{Commack, NY}
  \begin{position}
    \begin{itemize}
      \item Independently researched circuits over sets of natural numbers for three years.
      \item Won fourth place award in mathematics in ISEF (Intel International Science and Engineering Fair) in 2009, third place award in ISEF 2008.
    \end{itemize}
  \end{position}
\end{multitex*}

\begin{multitex*}[\cvlongfiles]
  \employer{Turnpike Total Appliance}
  \dates{Fall 2006--Summer 2009}
  \title{Web Page Designer}
  \location{Commack, NY}
  \begin{position}
    \begin{itemize}
      \item Designed and improved main company website, took initiative to make more improvements.
    \end{itemize}
  \end{position}

\end{multitex*}

\begin{multitex}[curriculum-vitae]
\section{\textsc{Programming Languages}}
\end{multitex}
\begin{multitex*}[\allexceptcv]
\section{\textsc{Computer Skills}}
\end{multitex*}
\begin{multitex}
    \begin{itemize}
      \item
        Proficient: Coq, \onlymultitex*{\cvlongfiles}{\TeX{} macro language, }Mathematica, git, Python, JavaScript, BASIC
      \item
        Working knowledge: \onlymultitex*{\cvlongfiles}{\LaTeX{}, }C, C++, Agda, OCaml, Haskell, Scheme, HTML, CSS, Perl, Java
      \item
        Basic knowledge: Matlab, Lean, Idris, Ruby, Go, Ur/Web, x86 Assembly
    \end{itemize}
\end{multitex}

\begin{multitex}
  \section{\textsc{Teaching}}
  \begin{itemize}
  \item Instructor at Monsoon Math Camp: category theory, linear logic, L\"ob's theorem
  \item TA for 6.172 (Performance Engineering% of Software% Systems
    ): Led recitations, analyzed and explained assembly output of \texttt{gcc -O3} to teach vectorization
  \item TA for 8.012 (Physics I) and 8.022 (Physics II) in Experimental Study Group
  \item Teacher at MIT ESP Programs: \LaTeX, philosophy, linear algebra, quantum mechanics
  \end{itemize}
\end{multitex}

\begin{multitex*}[\cvlongfiles]
\section{\textsc{Honors and Awards}}
\end{multitex*}
\begin{multitex*}[\cvlongfiles]
    \begin{itemize}
      \onlymultitex*{\highschoolfiles,\cvfiles}{%
      \item
        Mathematics Honor Society (Commack High School)
      }
      \item
        Collection of 12 original K'NEX synagogue models exhibited in various museums (2004--2008)
      \onlymultitex*{\highschoolfiles}{%
        \begin{itemize}
          \item
            Collection of 12 original K'NEX synagogue models
          \item
            Volunteered to give speeches and conduct K'NEX workshops for children at opening of exhibits, as well as in many locations in NY area
          \item
            April 2008 --- Gave interview for April edition of BabagaNewz Magazine and audio interview for BabagaNewz.com
          \item
            October 2006 --- Displayed in Gurwin Nursing Home on Long Island
          \item
            February 2006 --- Exhibited at Lisa Ann Watson Children's Discovery Museum in Miami, Florida
          \item
            October 2005 --- Displayed at Shomrim (Society for New York City Jewish Police Officers)
          \item
            February--April 2005 --- Exhibited in Yeshiva University Museum in Manhattan
          \item
            December 2004 --- Displayed at Suffolk Y-JCC in Commack
        \end{itemize}
        }
      \onlymultitex*{\cvfiles}{
      \item
        Graduated Cum Laude from Commack High School (June 2009)
      }
      \onlymultitex*{\highschoolfiles}{%
      \item
        Commack High School Outstanding Academic Performance (June 2009)
      \item
        Florence Burns Scholarship for Achievement of Excellence in Computer Science (June 2009)
      \item
        Theodore R. Spedalle Award for Achievement of Excellence in Mathematics (June 2009)
      \item
        Commack High School Mathematics Research Award (June 2009)
      \item
        President's Award for Educational Excellence (June 2009)
      \item
        AP Scholar with Distinction (2009)
      \item
        Robert C. Byrd Honors Scholarship (June 2009)
      \item
        ISEF (Intel International Science \& Engineering Fair)
        \begin{itemize}
          \item
            Math research paper, ``Circuits over Sets of Natural Numbers'' won $3^\text{rd}$ Place in Math in Atlanta, May 2008
          \item
            Math research paper, ``An Investigation of the Closure of the Set of Singleton Sets of Natural Numbers under $\cup$, $\cap$, $\overline{\phantom{\emptyset}}$, $+$, $\times$'' won $4^\text{th}$ place in Math in Reno, May 2009
       \end{itemize}
      \item
        LISEF (Long Island Science \& Engineering Fair) --- Won First Place Grand Award in Math in March 2009; selected to participate in ISEF 2009
      \item
        2009 National Merit Scholarship Finalist (February 2009)
      \item
        JSHS (Junior Science \& Humanities Symposium) --- Invited to present math research paper, February 2009 and February 2008
      \item
        Rensselaer Medal --- Awarded the Rensselaer Polytechnic Institute's Rensselaer Medal, June 2008
      \item
        Outstanding Junior Award --- sole recipient, awarded by Suffolk County Math Teacher's Association (June 2008)
      \item
        County All-Star Interscholastic Math League --- competed at NYSML (New York State Mathematics League) and ARML (American Regions Mathematics League) during junior, sophomore and freshmen years, 2006--2009; Achieved individual high score for team in 2009
      \item
        Suffolk County Senior High Interscholastic Math League --- $2^\text{nd}$ in county in 2009, $1^\text{st}$ in county in 2008, $3^\text{rd}$ in county in 2007
      \item
        NYSSEF (New York State Science and Engineering Fair)
        \begin{itemize}
          \item
            2008 ($11^\text{th}$ grade) --- $1^\text{st}$ in math in Intel division, one of four first place individual winners in fair, Mu Alpha Theta award, Army award, Ohio Wesleyan University Scholarship
          \item
            2007 ($10^\text{th}$ grade) --- $1^\text{st}$ in math in Andromeda Division, earned Mu Alpha Theta Award from National Mathematics Honor Society
        \end{itemize}
      \item
        Long Island Math Fair
        \begin{itemize}
          \item
            2008 ($11^\text{th}$ grade) --- Earned gold medal for ``Group Theory'' project
          \item
            2007 ($10^\text{th}$ grade) --- Earned gold medal for ``Circuits over Sets of Natural Numbers'' project
          \item
            2006 ($9^\text{th}$ grade) --- Earned gold medal for ``Prime Numbers'' project
          \item
            2005 ($8^\text{th}$ grade) --- Earned gold medal for ``The Fourth Spatial Dimension'' project
          \item
            2004 ($7^\text{th}$ grade) --- Earned gold medal for ``The Seven Bridges of K\"onigsberg'' project
        \end{itemize}
      \item
        AIME (American Invitational Mathematics Examination)
        \begin{itemize}
          \item
            2009 ($12^\text{th}$ grade) - ???
          \item
            2008 ($11^\text{th}$ grade) - score of 8
          \item
            2007 ($10^\text{th}$ grade) - score of 4
          \item
            2006 ($9^\text{th}$ grade) - score of 1
        \end{itemize}
      \item
        American Mathematics Contest 12 (AMC-12)
        \begin{itemize}
          \item
            2009 ($12^\text{th}$ grade) - score of ???
          \item
            2008 ($11^\text{th}$ grade) - score of 120, top 1\% nationwide, earned title of ``School Winner''
          \item
            2007 ($10^\text{th}$ grade) - score of 103.5, top 1\% nationwide, earned title of ``School Winner''
        \end{itemize}
      \item
        American Mathematics Contest 10 (AMC-10)
        \begin{itemize}
          \item
            2006 ($9^\text{th}$ grade) - score of 127.5, top 1\% nationwide, earned title of ``School Winner''
          \item
            2005 ($8^\text{th}$ grade) - score of 118.5, earned title of ``School Winner'' in high school as an $8^\text{th}$ grader in the $9^\text{th}$ grade honors math class
        \end{itemize}
      \item
        Suffolk County Mathematics Contest
        \begin{itemize}
          \item
            2009 ($12^\text{th}$ grade) - ???
          \item
            2008 ($11^\text{th}$ grade) - $1^\text{st}$ place in county competing with $12^\text{th}$ graders
          \item
            2007 ($10^\text{th}$ grade) - $4^\text{th}$ place in county competing with $11^\text{th}$ graders
          \item
            2006 ($9^\text{th}$ grade) - $4^\text{th}$ place in county competing with $10^\text{th}$ graders
          \item
            2005 ($8^\text{th}$ grade) - $5^\text{th}$ place in county competing with $9^\text{th}$ graders
          \item
            2004 ($7^\text{th}$ grade) - $1^\text{st}$ place in county
        \end{itemize}
      \item
        Suffolk County Mathematics Teachers Association (SCMTA) Tournament
        \begin{itemize}
          \item
            2009 ($12^\text{th}$ grade) - ??????th in county
          \item
            2008 ($11^\text{th}$ grade) - $4^\text{th}$ in county
          \item
            2007 ($10^\text{th}$ grade) - $2^\text{nd}$ in county
          \item
            2006 ($9^\text{th}$ grade) - $1^\text{st}$ in county
        \end{itemize}
      \item
        Commack High School Mathematics Department Awards --- June 2008, June 2007, and June 2006
      \item
        Commack Middle School Mathematics Departmental Award --- was only student of 605 students to receive Commack Middle School mathematics departmental award, 2005
      \item
        Suffolk County Mathcounts --- placed first in county as individual, 2005
      \item
        USA Mathematics Talent Search (USAMTS), 2007--2008; Placed on National List of Students with 6 Commended Solutions in 2008, silver medal in 2008
      \item
        Harvard-MIT Mathematics Tournament (HMMT), February 2009, 2008, and 2007
      \item
        New York State Science Olympiads
        \begin{itemize}
          \item
            February 2008 --- $2^\text{nd}$ place in Environmental Chemistry, $4^\text{th}$ place in Circuit Lab, $4^\text{th}$ place in Five Star Science
          \item
            February 2007 --- $4^\text{th}$ place in Circuit Lab
        \end{itemize}
      \item
        Mandelbrot Competition, team that competes in national competitions, 2006--2009; Placed on National Leaderboard (Top Tier) in 2008; co-captain in 2007--2008, captain in 2008--2009
      \item
        American Computer Science League (ACSL) Club, 2005--present; President in 2008--2009
      \item
        United States of America Computer Olympiad (USACO), 2005--present
      \item
        Suffolk County Senior High Interscholastic Math League, 2005--present
      \item
        Article was published in Teen Ink Magazine (November 2007)
      }
    \end{itemize}

\end{multitex*}

\begin{multitex*}[\highschoolfiles]
\section{\textsc{Community Service and Involvement}}
  \begin{itemize}
    \item
      Member of National Math Honor Society; Tutor students in math both in and out of school (2007--2009)
    \item
      Member of National Technology Honor Society (2007--2009)
  \end{itemize}
\end{multitex*}

\begin{multitex}
\section{\textsc{Other Activities}}
\end{multitex}
\begin{multitex}
    \begin{itemize}
      \onlymultitex*{\cvfiles,computer-science}{%
      \item Co-maintainer of the Fiat Cryptography project (\href{https://github.com/mit-plv/fiat-crypto}{\texttt{mit-plv/fiat-crypto}} on GitHub)
      \item Co-maintainer of the homotopy type theory Coq repository (\href{https://github.com/HoTT/HoTT}{\texttt{HoTT/HoTT} on GitHub})
      \item Program Committee Member of CoqPL 2022
      \onlymultitex*{\cvlongfiles}{%
        \item MIRI Decision Theory Workshop Attendee: Formalized various versions of L\"ob's theorem in Agda and Coq
        }
      \item Commiter to the SIPB BarnOwl project (\url{https://barnowl.mit.edu})
      }
      \onlymultitex*{\cvfiles}{%
      \item SIPB (Student Information and Processing Board) Member
      }
      \onlymultitex*{\cvlongfiles}{
      \item Was project leader for MITeX, an online interface for composing \LaTeX
      }
      \onlymultitex*{\cvfiles}{%
      \item President of Tech Squares, MIT's Square Dancing Club (May 2013--October 2014)
      }
      \onlymultitex*{\cvlongfiles}{%
      \item HMMT Solutions Editor (2010)
      }
      \item Canada/USA Mathcamp (Summers 2006--2009)
    \end{itemize}

\end{multitex}

\begin{multitex*}[\cvlongfiles]
% this is a hack
%\clearpage
\section{\textsc{Interests}}
    \begin{itemize}
      \item Philosophy
      \onlymultitex*{\cvlongfiles}{%
      \item Programming
      \item Dancing, especially tango, contra, and square dancing
        }
      \item Neuroscience
      \onlymultitex*{\cvlongfiles}{%
      \item Hiking
      %\onlymultitex*{\highschoolfiles}{
      \item MIT Mystery Hunt, Participated on Manic Sages Team, January 2008--2012
        %  }
        }
    \end{itemize}
\end{multitex*}

\begin{multitex}
  \nocite{*}
\end{multitex}

\begin{multitex*}[\cvshortfiles]
  \clearpage
  \section{\textsc{Selected Presentations and Publications}}
  \addtocategory{exclude}{Evaluation2021Huhmann}
  \printbibliography[title={$\left.\right.$},notcategory=exclude]
\end{multitex*}

\begin{multitex*}[\cvlongfiles]
  \section{\textsc{Presentations and Publications}}
  \printbibliography[title={$\left.\right.$}]
\end{multitex*}

\begin{multitex}
\endresume
\end{multitex}
\end{document}

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-bignums.sh

@@ -1,0 +1,16 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+
+# This script could be included inside other ones
+# Let's avoid to source ci-common twice in this case
+if [ -z "${CI_BUILD_DIR}" ];
+then
+    source ${ci_dir}/ci-common.sh
+fi
+
+bignums_CI_DIR=${CI_BUILD_DIR}/Bignums
+
+git_checkout ${bignums_CI_BRANCH} ${bignums_CI_GITURL} ${bignums_CI_DIR}
+
+( cd ${bignums_CI_DIR} && make && make install)

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-fiat-parsers.sh

@@ -1,0 +1,10 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+fiat_parsers_CI_DIR=${CI_BUILD_DIR}/fiat
+
+git_checkout ${fiat_parsers_CI_BRANCH} ${fiat_parsers_CI_GITURL} ${fiat_parsers_CI_DIR}
+
+( cd ${fiat_parsers_CI_DIR} && make parsers parsers-examples && make fiat-core )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-iris-lambda-rust.sh

@@ -1,0 +1,41 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+stdpp_CI_DIR=${CI_BUILD_DIR}/coq-stdpp
+Iris_CI_DIR=${CI_BUILD_DIR}/iris-coq
+lambdaRust_CI_DIR=${CI_BUILD_DIR}/lambdaRust
+
+install_ssreflect
+
+# Add or update the opam repo we need for dependency resolution
+opam repo add iris-dev https://gitlab.mpi-sws.org/FP/opam-dev.git -p 0 || opam update iris-dev
+
+# Setup lambdaRust first
+git_checkout ${lambdaRust_CI_BRANCH} ${lambdaRust_CI_GITURL} ${lambdaRust_CI_DIR}
+
+# Extract required version of Iris
+Iris_VERSION=$(cat ${lambdaRust_CI_DIR}/opam | fgrep coq-iris | egrep 'dev\.([0-9.-]+)' -o)
+Iris_URL=$(opam show coq-iris.$Iris_VERSION -f upstream-url)
+read -a Iris_URL_PARTS <<< $(echo $Iris_URL | tr '#' ' ')
+
+# Setup Iris
+git_checkout ${Iris_CI_BRANCH} ${Iris_URL_PARTS[0]} ${Iris_CI_DIR} ${Iris_URL_PARTS[1]}
+
+# Extract required version of std++
+stdpp_VERSION=$(cat ${Iris_CI_DIR}/opam | fgrep coq-stdpp | egrep 'dev\.([0-9.-]+)' -o)
+stdpp_URL=$(opam show coq-stdpp.$stdpp_VERSION -f upstream-url)
+read -a stdpp_URL_PARTS <<< $(echo $stdpp_URL | tr '#' ' ')
+
+# Setup std++
+git_checkout ${stdpp_CI_BRANCH} ${stdpp_URL_PARTS[0]} ${stdpp_CI_DIR} ${stdpp_URL_PARTS[1]}
+
+# Build std++
+( cd ${stdpp_CI_DIR} && make && make install )
+
+# Build and validate (except on Travis, i.e., skip if TRAVIS is non-empty) Iris
+( cd ${Iris_CI_DIR} && make && (test -n "${TRAVIS}" || make validate) && make install )
+
+# Build lambdaRust
+( cd ${lambdaRust_CI_DIR} && make && make install )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-ltac2.sh

@@ -1,0 +1,10 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+ltac2_CI_DIR=${CI_BUILD_DIR}/ltac2
+
+git_checkout ${ltac2_CI_BRANCH} ${ltac2_CI_GITURL} ${ltac2_CI_DIR}
+
+( cd ${ltac2_CI_DIR} && make && make tests && make install )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-math-comp.sh

@@ -1,0 +1,15 @@
+#!/usr/bin/env bash
+
+# $0 is not the safest way, but...
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+mathcomp_CI_DIR=${CI_BUILD_DIR}/math-comp
+
+checkout_mathcomp ${mathcomp_CI_DIR}
+
+# odd_order takes too much time for travis.
+( cd ${mathcomp_CI_DIR}/mathcomp                  && \
+  sed -i.bak '/PFsection/d'                  Make && \
+  sed -i.bak '/stripped_odd_order_theorem/d' Make && \
+  make Makefile.coq && make -f Makefile.coq all )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-sf.sh

@@ -1,0 +1,18 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+mkdir -p ${CI_BUILD_DIR} && cd ${CI_BUILD_DIR}
+wget -qO- ${sf_lf_CI_TARURL}  | tar xvz
+wget -qO- ${sf_plf_CI_TARURL} | tar xvz
+wget -qO- ${sf_vfa_CI_TARURL} | tar xvz
+
+sed -i.bak '1i From Coq Require Extraction.' lf/Extraction.v
+sed -i.bak '1i From Coq Require Extraction.' vfa/Extract.v
+
+( cd lf && make clean && make )
+
+( cd plf && sed -i.bak 's/(K,N)/((K,N))/' LibTactics.v && make clean && make )
+
+( cd vfa && make clean && make )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-template.sh

@@ -1,0 +1,12 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+Template_CI_BRANCH=master
+Template_CI_GITURL=https://github.com/Template/Template
+Template_CI_DIR=${CI_BUILD_DIR}/Template
+
+git_checkout ${Template_CI_BRANCH} ${Template_CI_GITURL} ${Template_CI_DIR}
+
+( cd ${Template_CI_DIR} && make )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-vst.sh

@@ -1,0 +1,13 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+VST_CI_DIR=${CI_BUILD_DIR}/VST
+
+# opam install -j ${NJOBS} -y menhir
+git_checkout ${VST_CI_BRANCH} ${VST_CI_GITURL} ${VST_CI_DIR}
+
+# Targets are: msl veric floyd progs , we remove progs to save time
+# Patch to avoid the upper version limit
+( cd ${VST_CI_DIR} && make IGNORECOQVERSION=true .loadpath version.vo msl veric floyd )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/lint-repository.sh

@@ -1,0 +1,34 @@
+#!/usr/bin/env bash
+
+# A script to check prettyness over the repository.
+
+# lint-commits.sh seeks to prevent the worsening of already present
+# problems, such as tab indentation in ml files. lint-repository.sh
+# seeks to prevent the (re-)introduction of solved problems, such as
+# newlines at the end of .v files.
+
+CODE=0
+
+if [ -n "${TRAVIS_PULL_REQUEST}" ] && [ "${TRAVIS_PULL_REQUEST}" != false ];
+then
+    # skip PRs from before the linter existed
+    if [ -z "$(git ls-tree --name-only "${TRAVIS_PULL_REQUEST_SHA}" dev/lint-commits.sh)" ];
+    then
+        1>&2 echo "Linting skipped: pull request older than the linter."
+        exit 0
+    fi
+
+    # Some problems are too widespread to fix in one commit, but we
+    # can still check that they don't worsen.
+    CUR_HEAD=${TRAVIS_COMMIT_RANGE%%...*}
+    PR_HEAD=${TRAVIS_COMMIT_RANGE##*...}
+    MERGE_BASE=$(git merge-base "$CUR_HEAD" "$PR_HEAD")
+    dev/lint-commits.sh "$MERGE_BASE" "$PR_HEAD" || CODE=1
+fi
+
+# Check that the files with 'whitespace' gitattribute end in a newline.
+# xargs exit status is 123 if any file failed the test
+find . "(" -path ./.git -prune ")" -o -type f -print0 |
+    xargs -0 dev/tools/check-eof-newline.sh || CODE=1
+
+exit $CODE

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/save-logs.sh

@@ -1,0 +1,19 @@
+#!/usr/bin/env bash
+
+SAVEDIR="logs"
+
+# reset for local builds
+rm -rf "$SAVEDIR"
+mkdir "$SAVEDIR"
+
+# keep this synced with test-suite/Makefile
+FAILMARK="==========> FAILURE <=========="
+
+FAILED=$(mktemp /tmp/coq-check-XXXXX)
+find . '(' -path ./bugs/opened -prune ')' -o '(' -name '*.log' -exec grep "$FAILMARK" -q '{}' ';' -print0 ')' > "$FAILED"
+
+rsync -a --from0 --files-from="$FAILED" . "$SAVEDIR"
+cp summary.log "$SAVEDIR"/
+
+# cleanup
+rm "$FAILED"

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-color.sh
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@849f2e7ac141efa62e594e461c33cc60a245b123 🚀

@@ -1,10 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-CoLoR_CI_DIR=${CI_BUILD_DIR}/color
-
-# Compile CoLoR
-git_checkout ${CoLoR_CI_BRANCH} ${CoLoR_CI_GITURL} ${CoLoR_CI_DIR}
-( cd ${CoLoR_CI_DIR} && make )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-fiat-crypto.sh

@@ -1,11 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-fiat_crypto_CI_DIR=${CI_BUILD_DIR}/fiat-crypto
-
-git_checkout ${fiat_crypto_CI_BRANCH} ${fiat_crypto_CI_GITURL} ${fiat_crypto_CI_DIR}
-( cd ${fiat_crypto_CI_DIR} && git submodule update --init --recursive )
-
-( cd ${fiat_crypto_CI_DIR} && make lite )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-iris-lambda-rust.sh

@@ -1,41 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-stdpp_CI_DIR=${CI_BUILD_DIR}/coq-stdpp
-Iris_CI_DIR=${CI_BUILD_DIR}/iris-coq
-lambdaRust_CI_DIR=${CI_BUILD_DIR}/lambdaRust
-
-install_ssreflect
-
-# Add or update the opam repo we need for dependency resolution
-opam repo add iris-dev https://gitlab.mpi-sws.org/FP/opam-dev.git -p 0 || opam update iris-dev
-
-# Setup lambdaRust first
-git_checkout ${lambdaRust_CI_BRANCH} ${lambdaRust_CI_GITURL} ${lambdaRust_CI_DIR}
-
-# Extract required version of Iris
-Iris_VERSION=$(cat ${lambdaRust_CI_DIR}/opam | fgrep coq-iris | egrep 'dev\.([0-9.-]+)' -o)
-Iris_URL=$(opam show coq-iris.$Iris_VERSION -f upstream-url)
-read -a Iris_URL_PARTS <<< $(echo $Iris_URL | tr '#' ' ')
-
-# Setup Iris
-git_checkout ${Iris_CI_BRANCH} ${Iris_URL_PARTS[0]} ${Iris_CI_DIR} ${Iris_URL_PARTS[1]}
-
-# Extract required version of std++
-stdpp_VERSION=$(cat ${Iris_CI_DIR}/opam | fgrep coq-stdpp | egrep 'dev\.([0-9.-]+)' -o)
-stdpp_URL=$(opam show coq-stdpp.$stdpp_VERSION -f upstream-url)
-read -a stdpp_URL_PARTS <<< $(echo $stdpp_URL | tr '#' ' ')
-
-# Setup std++
-git_checkout ${stdpp_CI_BRANCH} ${stdpp_URL_PARTS[0]} ${stdpp_CI_DIR} ${stdpp_URL_PARTS[1]}
-
-# Build std++
-( cd ${stdpp_CI_DIR} && make && make install )
-
-# Build and validate (except on Travis, i.e., skip if TRAVIS is non-empty) Iris
-( cd ${Iris_CI_DIR} && make && (test -n "${TRAVIS}" || make validate) && make install )
-
-# Build lambdaRust
-( cd ${lambdaRust_CI_DIR} && make && make install )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-ltac2.sh

@@ -1,10 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-ltac2_CI_DIR=${CI_BUILD_DIR}/ltac2
-
-git_checkout ${ltac2_CI_BRANCH} ${ltac2_CI_GITURL} ${ltac2_CI_DIR}
-
-( cd ${ltac2_CI_DIR} && make && make tests && make install )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-unimath.sh

@@ -1,14 +1,0 @@
-#!/usr/bin/env bash
-
-ci_dir="$(dirname "$0")"
-source ${ci_dir}/ci-common.sh
-
-UniMath_CI_DIR=${CI_BUILD_DIR}/UniMath
-
-git_checkout ${UniMath_CI_BRANCH} ${UniMath_CI_GITURL} ${UniMath_CI_DIR}
-
-( cd ${UniMath_CI_DIR}                        && \
-  sed -i.bak '/Folds/d'              Makefile && \
-  sed -i.bak '/HomologicalAlgebra/d' Makefile && \
-  make BUILD_COQ=no )
-

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/tools/merge-pr.sh

@@ -1,50 +1,0 @@
-#!/usr/bin/env bash
-
-set -e
-
-# This script depends (at least) on git and jq.
-# It should be used like this: dev/tools/merge-pr.sh /PR number/
-
-#TODO: check arguments and show usage if relevant
-
-PR=$1
-
-CURRENT_LOCAL_BRANCH=$(git rev-parse --abbrev-ref HEAD)
-REMOTE=$(git config --get "branch.$CURRENT_LOCAL_BRANCH.remote")
-git fetch "$REMOTE" "refs/pull/$PR/head"
-
-API=https://api.github.com/repos/coq/coq
-
-BASE_BRANCH=$(curl -s "$API/pulls/$PR" | jq -r '.base.label')
-
-COMMIT=$(git rev-parse FETCH_HEAD)
-STATUS=$(curl -s "$API/commits/$COMMIT/status" | jq -r '.state')
-
-if [ "$BASE_BRANCH" != "coq:$CURRENT_LOCAL_BRANCH" ]; then
-  echo "Wrong base branch"
-  read -p "Bypass? [y/N] " -n 1 -r
-  echo
-  if [[ ! $REPLY =~ ^[Yy]$ ]]
-  then
-    exit 1
-  fi
-fi;
-
-if [ "$STATUS" != "success" ]; then
-  echo "CI status is \"$STATUS\""
-  read -p "Bypass? [y/N] " -n 1 -r
-  echo
-  if [[ ! $REPLY =~ ^[Yy]$ ]]
-  then
-    exit 1
-  fi
-fi;
-
-git merge -S --no-ff FETCH_HEAD -m "Merge PR #$PR: $(curl -s "$API/pulls/$PR" | jq -r '.title')" -e
-
-# TODO: improve this check
-if ! git diff --quiet "$REMOTE/$CURRENT_LOCAL_BRANCH" -- dev/ci; then
-  echo "******************************************"
-  echo "** WARNING: does this PR have overlays? **"
-  echo "******************************************"
-fi

### 2021-08-15 presentations/coq-workshop-2018-coq/doc/sphinx/conf.py

@@ -1,414 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-##########################################################################
-##         #   The Coq Proof Assistant / The Coq Development Team       ##
-##  v      #   INRIA, CNRS and contributors - Copyright 1999-2018       ##
-## <O___,, #       (see CREDITS file for the list of authors)           ##
-##   \VV/  ###############################################################
-##    //   #    This file is distributed under the terms of the         ##
-##         #     GNU Lesser General Public License Version 2.1          ##
-##         #     (see LICENSE file for the text of the license)         ##
-##########################################################################
-#
-# Coq documentation build configuration file, created by
-# sphinx-quickstart on Wed May 11 11:23:13 2016.
-#
-# This file is execfile()d with the current directory set to its
-# containing dir.
-#
-# Note that not all possible configuration values are present in this
-# autogenerated file.
-#
-# All configuration values have a default; values that are commented out
-# serve to show the default.
-
-import sys
-import os
-
-# Increase recursion limit for sphinx
-sys.setrecursionlimit(1500)
-
-# If extensions (or modules to document with autodoc) are in another directory,
-# add these directories to sys.path here. If the directory is relative to the
-# documentation root, use os.path.abspath to make it absolute, like shown here.
-sys.path.append(os.path.abspath('../tools/'))
-sys.path.append(os.path.abspath('../../config/'))
-
-import coq_config
-
-# -- General configuration ------------------------------------------------
-
-# If your documentation needs a minimal Sphinx version, state it here.
-#needs_sphinx = '1.0'
-
-# Add any Sphinx extension module names here, as strings. They can be
-# extensions coming with Sphinx (named 'sphinx.ext.*') or your custom
-# ones.
-extensions = [
-    'sphinx.ext.mathjax',
-    'sphinx.ext.todo',
-    'sphinxcontrib.bibtex',
-    'coqrst.coqdomain'
-]
-
-# Add any paths that contain templates here, relative to this directory.
-templates_path = ['_templates']
-
-# The suffix(es) of source filenames.
-# You can specify multiple suffix as a list of string:
-# source_suffix = ['.rst', '.md']
-source_suffix = '.rst'
-
-# The encoding of source files.
-#source_encoding = 'utf-8-sig'
-
-# The master toctree document.
-master_doc = 'index'
-
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/coq-makefile/plugin1/run.sh

@@ -1,26 +1,0 @@
-#!/usr/bin/env bash
-
-. ../template/init.sh
-
-mv src/test_plugin.mlpack src/test_plugin.mllib
-coq_makefile -f _CoqProject -o Makefile
-cat Makefile.conf
-make
-make html mlihtml
-make install DSTROOT="$PWD/tmp"
-#make debug
-(cd `find tmp -name user-contrib` && find .) | sort > actual
-sort > desired <<EOT
-.
-./test
-./test/test.glob
-./test/test.cmi
-./test/test.cmx
-./test/test_aux.cmi
-./test/test_aux.cmx
-./test/test_plugin.cmxa
-./test/test_plugin.cmxs
-./test/test.v
-./test/test.vo
-EOT
-exec diff -u desired actual

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/coq-makefile/timing/run.sh

@@ -1,104 +1,0 @@
-#!/usr/bin/env bash
-
-#set -x
-set -e
-
-. ../template/path-init.sh
-
-cd precomputed-time-tests
-./run.sh || exit $?
-
-cd ../error
-coq_makefile -f _CoqProject -o Makefile
-make cleanall
-if make pretty-timed TGTS="all" -j1; then
-    echo "Error: make pretty-timed should have failed"
-    exit 1
-fi
-
-cd ../aggregate
-coq_makefile -f _CoqProject -o Makefile
-make cleanall
-make pretty-timed TGTS="all" -j1 || exit $?
-
-cd ../before
-coq_makefile -f _CoqProject -o Makefile
-make cleanall
-make make-pretty-timed-before TGTS="all" -j1 || exit $?
-
-cd ../after
-coq_makefile -f _CoqProject -o Makefile
-make cleanall
-make make-pretty-timed-after TGTS="all" -j1 || exit $?
-rm -f time-of-build-before.log
-make print-pretty-timed-diff TIMING_SORT_BY=diff TIME_OF_BUILD_BEFORE_FILE=../before/time-of-build-before.log
-cp ../before/time-of-build-before.log ./
-make print-pretty-timed-diff TIMING_SORT_BY=diff || exit $?
-
-INFINITY="∞"
-INFINITY_REPLACEMENT="+.%" # assume that if the before time is zero, we expected the time to increase
-
-TO_SED_IN_BOTH=(
-    -e s"/${INFINITY}/${INFINITY_REPLACEMENT}/g" # Whether or not something shows up as ∞ depends on whether a time registers as 0.s or as 0.001s, so we can't rely on this being consistent
-    -e s':|\s*N/A\s*$:| '"${INFINITY_REPLACEMENT}"':g' # Whether or not something shows up as N/A depends on whether a time registers as 0.s or as 0.001s, so we can't rely on this being consistent
-    -e s'/ *$//g' # the number of trailing spaces depends on how many digits percentages end up being; since this varies across runs, we remove trailing spaces
-    -e s'/[0-9]*\.[0-9]*//g' # the precise timing numbers vary, so we strip them out
-    -e s'/^-*$/------/g' # When none of the numbers get over 100 (or 1000, in per-file), the width of the table is different, so we normalize the number of dashes for table separators
-)
-
-TO_SED_IN_PER_FILE=(
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/coq-makefile/vio2vo/run.sh

@@ -1,13 +1,0 @@
-#!/usr/bin/env bash
-a=`uname`
-
-. ../template/init.sh
-
-coq_makefile -f _CoqProject -o Makefile
-make quick
-# vio2vo is broken on Windows (#6720)
-if [ "$a" = "Darwin" -o "$a" = "Linux" ]; then
-    make vio2vo J=2
-    test -f theories/test.vo
-    make validate
-fi

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/misc/deps-order.sh

@@ -1,20 +1,0 @@
-# Check that both coqdep and coqtop/coqc supports -R
-# Check that both coqdep and coqtop/coqc takes the later -R
-# See bugs 2242, 2337, 2339
-rm -f misc/deps/lib/*.vo misc/deps/client/*.vo
-tmpoutput=`mktemp /tmp/coqcheck.XXXXXX`
-$coqdep -R misc/deps/lib lib -R misc/deps/client client misc/deps/client/bar.v 2>&1 | head -n 1 > $tmpoutput
-diff -u --strip-trailing-cr misc/deps/deps.out $tmpoutput 2>&1
-R=$?
-times
-$coqc -R misc/deps/lib lib misc/deps/lib/foo.v 2>&1
-$coqc -R misc/deps/lib lib -R misc/deps/client client misc/deps/client/foo.v 2>&1
-$coqtop -R misc/deps/lib lib -R misc/deps/client client -load-vernac-source misc/deps/client/bar.v 2>&1
-S=$?
-if [ $R = 0 -a $S = 0 ]; then
-    printf "coqdep and coqtop agree\n"
-    exit 0
-else
-    printf "coqdep and coqtop disagree\n"
-    exit 1
-fi

### 2021-08-15 _site/papers/2015-parsing-parse-trees-reviews/detailed-review-response-notes.txt
Add back deploy

Reviewer A

> Splitting oracles: the essential algorithmic component of parsing is
  the association of the input into meaningful trees. Deferring the
  split choice to an "oracle" means that all non-trivial computation
  is in it. So I would not say that you're performing parsing.

This is a good point.  We are intentionally not doing anything interesting with regard to algorithmic parsing theory; that is reserved for future papers on currently in-progress work on synthesizing efficient parsers.  We picked the simplest, most obvious parsing algorithm we could think of, as a starting point.  That you claim our algorithm contains only trivial computation speaks to our success in picking a simple starting point.  Allow me to add a trivial splitter, which allows us to really perform parsing, trivially: the brute-force splitter, which returns the sequence of numbers [0, 1, ..., length of string] is sufficient for correctness, because we filter the list of splits to accept only valid ones.  We believe that the idea of parsing parse trees is complicated enough as it stands, and does not need extra complexity in exposition introduced by chosing a non-trivial parsing algorithm as a starting point.

> The work that remains appear to be mostly about detection of
  recursion in the grammar, which is easier than the above step.

Certainly not easier, when good performance is not demanded (as it is not, for code that is never run and only compiled).  It is straightforward, bordering on trivial, to prove that returning every possible split of the string does not miss any splits (and is thus complete), and it does not add any complexity to the completeness proof to do it this way.

> The textbook approach to detect loops is by reduction to Chomsky
  Normal Form. (This can be done statically on the grammar.) In
  comparison your approach feels awkward. At the minimum, discuss
  your choice!

This seems more complicated, but I am not sure.  Computing the Chompsky Normal Form of a grammar in a manifestly terminating way, proving that every grammar has a Chompsky Normal Form, that a grammar has a parse iff its CNF has a parse, this seems more complicated than proving that our way of keeping track of which nonterminals have been seen does not lose us any valid parses.  Perhaps it is not more complicated, but it seems less ovbious, and not obviously simpler nor more straightforward.

> I don't think that your interface to the parser is sensible. Why is
  it so long (fig 1)? In my mind I should just supply the grammar
  rules. (And perhaps your oracle.) I can also not reconcile your
  explanation (reading it off the minimal parse tree) with fig 1.

The interface is not a user-facing one; it is the internal one, to be partially instantiated to generate minimal parse trees, or parse trees, or booleans.  Once you fix the type of the return, indeed, all that is left is to specify the rules of the grammar, and, perhaps, the splitter.  The length of the interface results from the fact that the algorithm, a priori, knows nothing about minimal parse trees; the interface needs to have enough generality to specify exactly how to build a minimal parse tree, and how to prove that no such parse tree exists.  Hence it has enough cases for all the constructors, and enough cases to prove absurdity from impossible parse trees.

>   1. Define a "ParseTreeOf nt s" (the spec)
  2. Write an algorithm to decide the relation:

     parse : (nt : NT) -> (s : String) -> Decide (MinParseTreeOf ns t)

  And to get to the ParseTreeOf version only the de-loop lemma is
  needed (in one direction).

The de-loop lemma would need to be strengthened to do this; the current version only de-loops at one level, which is not enough to get ParseTreeOf -> MinParseTreeOf.  When proving correctness of the parser returning booleans directly, the full delooping construction is longer than the entire rest of the soundness and completeness proofs, combined!  Note that the full delooping construction includes the generation of ParseTreeOf from MinParseTreeOf as a way of getting well-foundedness evidence.

Reviewer B
> I had problems in matching the paper to the accompanying Coq codes.

I agree, this is a problem.  The accompanying code is a half-assed proof-of-concept, not an elegant match for the description.

I agree, this is a problem.  We did not find enough time to clean up the code.  It was an explicit goal in writing the paper to hide as many of the warts of our code-base as we could: long names, long lines, costly mistakes that we didn't have time to fix, unenlightening details forced on us by Coq.

> E.g., grep could not find has_parse or min_parse -- I was interested in a proof of min_parse.)

We apologize.  [min_parse] is named [minimal_parse_nonterminal__of__parse] in our code base.  We should have done more to make the code match the paper, and intend to do so, should the paper be accepted.
[FIXME: Adam: Is this okay to promise?]

> I miss technical comparisons
with the work by Ridge in HOL4. The authors say their approach is different from Ridge's in that they use dependent types which have a different aesthetic appeal 
-- the statement is too general to be informative.

FIXME: How to respond?

Reviewer C:
> what I was wanting to see was a set of versions of increasing complexity (with diffs indicated) that gave more of an indication of the way in which the development had proceeded.

To a first approximation, there is the code in 1.1, the slightly more complicated code in 1.2, and the fuller code of Figure 2.  Perhaps we could have indicated diffs more clearly, or rewritten 1.1 and 1.2 when it came time to introduce Figure 2, perhaps indicating which bits were already there in 1.1, which bits where there to prevent infinite regress from 1.2, and which bits are just there to handle the more complicated type signature.  In the actual code, there is BooleanRecognizer.v, which corresponds to the grammar-independent version of the code in 1.2 (equivalently, to the code in Figure 2 specialized to booleans), while DependentlyTyped.v gives the full code of the interface and implementation of Figures 1 and 2.

> there is a substantial amount to do on defining splitters and their effect on performance (more discussion on splitters below); this is reported as planned work;

Yes, this paper is about parsers that can prove their own correctness.  We should have mentioned that we believe this can be extended to other parsing algorithms.

> when splitting strings, this is not the case and it was not clear (at least to me) that for every grammar a splitter of the right kind exists

The brute-force splitter, which returns the sequence of numbers [0, 1, ..., length of string] is sufficient for correctness, because we filter the list of splits to accept only valid ones.

> did editors and any other tools support the development, or not?

We used ProofGeneral, with it's interactive display of coqtop feedback, syntax highlighting, and saving capabilities.  But most of these features are built into coqtop, and are assumed when developing Coq code; an interactive compiler seems essential when developing proof scripts and dependently typed code.  Is it worth mentioning explicitly?

Reviewer D

> The biggest criticism is about the code. The presented code in the
paper is not intact with the real development which makes it
impossible to really see if the proposed approach is as simple and
elegant as authors claim. I was unable to find functions like
"has_parse", "has_no_parse", "min_parse", "parse_full", 
"parse_sound", "parse_complete", etc.

Yes, this needs to be fixed, thank you for the suggestions.

> You are using provocative statement that parser is proving its
correctness and this is not what is happening.

I am tempted to try to do this precisely, rather than approximately, as I suggest in the last section.

> At page 2 you say that your strategy works only with finite sets of
nonterminals.  However, set of nonterminals could be infinite, but
then you should add the step of extracting the "used" nonterminals
from the grammar.

Consider A := B, B := C, C := D, and so on, ad infinitum.  There is no finite set of "used" nonterminals to be extracted, not even a decision strategy to see if it terminates.  The strategy really only works for manifestly finite sets of explicitly given "used nonterminals", which is what is meant by "finite sets of nonterminals".

> In subsection 2.3 it would be nice if you would make precise which
parse tree is returned by the parser for ambiguous grammars. In that
sense your notion of "completeness" is very special.

It depends on the splitter.  It returns the first loopless parse tree that the splitter provides it.  It would be relatively easy to make it return all parse trees, roughly by replacing a "fold" with a "flat_map".

### 2021-08-15 _site/papers/source/fiat-crypto/sigops-osr-2020/fiat-crypto.tex
Add back deploy

--- 2020-07-30 papers/source/fiat-crypto/fiat-crypto.tex
@@
-\documentclass[conference,letterpaper]{IEEEtran}
-%\overfullrule=1mm
-\bibliographystyle{IEEEtran}
+%Page: single-spaced 8.5″ x 11″ pages
+%Font and Columns: 10-point type on 12-point (single-spaced) leading, two-column format, Times Roman or a similar font, within a text block 6.5″ wide x 9″ deep.
+%Number of pages: OSR does not enforce strict minimum or maximum page limits. We suggest 2-6 pages as a rough guideline, but authors should feel free to devote more space as needed..
+%Other considerations: Pages should NOT be numbered, and figures and tables should be legible in black and white, without requiring magnification.
+%Please feel free to use any Systems conference template (e.g., USENIX ATC, OSDI, SOSP, NSDI, etc.) to prepare drafts.
+
+\documentclass[letterpaper,twocolumn,10pt]{article}
+\usepackage{usenix2019_v3}
+
+\bibliographystyle{plain}
+\pagestyle{empty}
 
@@
 \usepackage{mathtools}
-\include{digitboxed}
 
@@
 
+%-------------------------------------------------------------------------------
 \begin{document}
-
-\title{Simple High-Level Code For Cryptographic Arithmetic -- With Proofs, Without Compromises}
-
-\author{%
- \IEEEauthorblockN{Andres Erbsen \quad Jade Philipoom \quad Jason Gross \quad Robert Sloan \quad Adam Chlipala}
- \IEEEauthorblockA{%
-   MIT CSAIL, \\
-   Cambridge, MA, USA \\
-   \{\texttt{andreser}, \texttt{jadep}, \texttt{jgross}\}\texttt{@mit.edu}, \texttt{rob.sloan@alum.mit.edu}, \texttt{adamc@csail.mit.edu} %TODO(@csail.mit.edu for all?)
- }
-}
+%-------------------------------------------------------------------------------
+
+%don't want date printed
+\date{}
+
+% make title bold and 14 pt font (Latex default is non-bold, 16 pt)
+\title{\Large \bf Simple High-Level Code For Cryptographic Arithmetic: \\ With Proofs, Without Compromises}
+
+% The template didn't include email addresses; do we want to?
+\author{
+{\rm Andres Erbsen} \\
+MIT CSAIL \\
+{\rm \texttt{andreser@mit.edu}}
+\and
+{\rm Jade Philipoom} \\
+MIT CSAIL \\
+{\rm \texttt{jadep@mit.edu}}
+\and
+{\rm Jason Gross} \\
+MIT CSAIL \\
+{\rm \texttt{jgross@mit.edu}}
+\and
+{\rm Robert Sloan} \\
+MIT CSAIL \\
+{\rm \texttt{rob.sloan@alum.mit.edu}}
+\and
+{\rm Adam Chlipala} \\
+MIT CSAIL \\
+{\rm \texttt{adamc@csail.mit.edu}}
+} % end author
 
@@
 \begin{abstract}
-  We introduce a new approach for implementing cryptographic arithmetic in short high-level code with machine-checked proofs of functional correctness.
-  We further demonstrate that simple partial evaluation is sufficient to transform such initial code into the fastest-known C code,
-  breaking the decades-old pattern that the only fast implementations are those whose instruction-level steps were written out by hand.
-
-  These techniques were used to build an elliptic-curve library that achieves competitive performance for 80 prime fields and multiple CPU architectures, showing that implementation and proof effort scales with the number and complexity of conceptually different algorithms, not their use cases.
+  We introduce an unusual approach for implementing cryptographic arithmetic in short high-level code with machine-checked proofs of functional correctness.
+  We further demonstrate that simple partial evaluation is sufficient to transform such initial code into highly competitive C code, breaking the decades-old pattern that the only fast implementations are those whose instruction-level steps were written out by hand.
+
+These techniques were used to build an elliptic-curve library that achieves competitive performance for a wide range of prime fields and multiple CPU architectures, showing that implementation and proof effort scales with the number and complexity of conceptually different algorithms, not their use cases.
 As one outcome, we present the first verified high-performance implementation of P-256, the most widely used elliptic curve.
 Implementations from our library were included in BoringSSL to replace existing specialized code, for inclusion in several large deployments for Chrome, Android, and CloudFlare.
+
+This is an abridged version of the full paper originally presented in IEEE S\&P 2019~\cite{FiatCryptoSP19}.
+We have omitted most proof-engineering details in favor of a focus on the system's functional capabilities.
 \end{abstract}
 
-\IEEEpeerreviewmaketitle
-
-\section{Motivation}
-
-The predominant practice today is to implement cryptographic primitives at the lowest possible level of abstraction: usually flat C code, ideally assembly language if engineering time is plentiful.
-Low-level implementation has been the only way to achieve good performance, ensure that execution time does not leak secret information, and enable the code to be called from any programming language.
-Yet placing the burden of every detail on the programmer also leaves a lot to be desired.
-The demand for implementation experts' time is high, resulting in an ever-growing backlog of algorithms to be reimplemented in specialized code ``as soon as possible.''
-Worse, bugs are also common enough that it is often not economical to go beyond fixing the behavior for problematic inputs and understand the root cause and impact of the defect.
-
-There is hope that these issues can eventually be overcome by redoubling community efforts to produce high-quality crypto code.
-However, even the most renowned implementors are not infallible, and if a typo among thousands of lines of monotonous code can only be recognized as incorrect by top experts,\footnote{We refer to the \texttt{ed25519-amd64-64-24k} bug that is discussed along with other instructive real-world implementation defects in Appendix \ref{bugsapp}.} it is reasonable to expect diminishing returns for achieving correctness through manual review.
-A number of tools for computer-aided verification have been proposed~\cite{verif25519,gfverif,ECCstar,hacl,vale,jasmin}, enabling developers to eliminate incorrect-output bugs conclusively.
-Three solid implementations of the X25519 Diffie-Hellman function were verified.
-In the beautiful world where X25519 was the only arithmetic-based crypto primitive we need, now would be the time to declare victory and go home.
-Yet most of the Internet still uses P-256, and the current proposals for post-quantum cryptosystems are far from Curve25519's combination of performance and simplicity.
-
-A closer look into the development and verification process (with an eye towards replicating it for another function) reveals a discouraging amount of duplicated effort.
-First, expert time is spent on mechanically writing out the instruction-level steps required to perform the desired arithmetic, drawing from experience and folklore.
-Then, expert time is spent to optimize the code by applying a carefully chosen combination of simple transformations.
-Finally, expert time is spent on annotating the now-munged code with assertions relating values of run-time variables to specification variables, until the SMT solver (used as a black box) is convinced that each assertion is implied by the previous ones.
-As the number of steps required to multiply two numbers is quadratic in the size of those numbers, effort grows accordingly: for example ``Verifying Curve25519 Software''~\cite{verif25519} reports that multiplication modulo $2^{255}-19$ required 27 times the number of assertions used for $2^{127}-1$.
-It is important to note that both examples use \emph{the same} simple modular multiplication algorithm, and the same correctness argument should apply, just with different parameters.
-This is just one example of how manual repetition of algorithmic steps inside primitive cryptographic arithmetic implementations makes them notoriously difficult to write, review, or prove correct.
-
-Hence we suggest a rather different way of implementing cryptographic primitives:
-to implement general algorithms in the style that best illustrates their correctness arguments
-and then derive parameter-specific low-level code through partial evaluation.
-For example, the algorithm used for multiplication modulo $2^{127}-1$ and $2^{255}-19$ is first encoded as a purely functional program operating on lists of digit-weight pairs: it generates a list of partial products, shifts down the terms with weight greater than the modulus, and scales them accordingly.
-It is then instantiated with lists of concrete weights and length, for example ten limbs with weights $2^{\lceil 25.5 i\rceil}$, and partially evaluated so that no run-time manipulation of lists and weights remains, resulting in straight-line code very similar to what an expert would have written.
-As the optimization pipeline is proven correct once and for all, correctness of all code generated from a high-level algorithm can be proven directly on its high-level implementation.
-A nice side benefit of this approach is that the same high-level algorithm and proof can be used for multiple target configurations with very little effort, enabling the use of fast implementations in cases that were previously not worth the effort.
-Finally, the generated code is compiled with a conventional C compiler.
-As current compiler support for different CPU features varies wildly in optimization quality, this step can limit the performance of our code (especially when the best compilation would make careful use of carry flags or vector instructions).
-We aim for the fastest code in terms of ALU functions, but we leave instruction selection, scheduling, and register allocation out of scope.
-
-In practice, we use the fast arithmetic implementations reported on here in conjunction with a library of verified cryptographic algorithms, the rest of which deviates much less from common implementation practice.
-All formal reasoning is done in the Coq proof assistant, and the overall trusted computing base also includes a simple pretty-printer and the C language toolchain.
-The development, called \textbf{Fiat Cryptography}, is available under the MIT license at:
+\section{Introduction}
+
+As verification makes the transition from research topic to component of real-world software ecosystems, the field must face challenges of scale, in terms of both amount of code and conceptual complexity.
+Each system, of course, has its own constraints and objectives, and these needs will best be met with a broad toolbox of approaches.
+In this paper, we sketch a case study following a relatively unusual approach, in the hope that it will help illuminate the broad array of possibilities for verification projects.
+The full paper was originally presented in IEEE S\&P 2019~\cite{FiatCryptoSP19}; what follows is a version abridged, selectively updated, and adapted for a more systems-oriented audience.
+
+That audience is probably most familiar with projects fitting the classic mold of proving theorems about programs written by hand, as exemplified by sEL4~\cite{seL4SOSP09}, IronClad~\cite{hawblitzel2014ironclad,hawblitzel2015ironfleet}, FSCQ~\cite{FscqSOSP15,FscqSOSP17,FscqOSDI18}, CertiKOS~\cite{CertiKOS}, and various case studies in push-button verification~\cite{Pbfs,Hyperkernel,Nickel,Serval}.
+Often the implementers of these projects do rewrite systems from scratch in tandem with proofs, but the final code artifacts implement very similar pseudocode to off-the-shelf systems.
+A moderately widely known alternative is \emph{program synthesis}, which writes a program automatically from its logical specification.
+Most colloquially, this idea is associated with relatively intensive combinatorial search through spaces of possibilities, often taking advantage of SMT solvers like Z3~\cite{Z3}.
+However, some systems-code domains are focused enough that we can write very predictable and \emph{verified} generic implementations.
+With the help of a verified domain-specific compiler, we can translate these general problem descriptions into practical, performance-competitive code.
+This paper considers a successful example of the template-based code-generation strategy.
+
+Our project verifies efficient modular-arithmetic routines for elliptic-curve cryptography, a domain to which much optimization effort has been applied.
+This means that competitive implementations use highly specialized techniques, previously accessible only to a small number of experts.
+We aimed to create a system that generalized these approaches and proved, with minimal trusted code, that the detailed low-level routines implemented their high-level arithmetic specifications.
+To that end, we first studied handwritten implementations and extracted fully general descriptions of the complex design choices they required.
+Then, we encoded these algorithms in the Coq proof assistant~\cite{coq} as templates, generic over compile-time parameters like the modulus and integer width of the underlying architecture.
+Once these parameters are plugged in, a domain-specific compiler transforms the templates into efficient low-level code.
+
+The output code is encoded in a specialized low-level abstract syntax tree (AST) that includes only very basic fixed-width integer operations, at the approximate abstraction level of assembly instructions, and is easily translatable to a number of output languages.
+At current count we can produce C, Go, and Rust, along with two other unusual specialized languages for specific applications.
+The C code, our primary target, is competitive with existing handwritten C, although it still falls short of the performance of handwritten assembly.
+However, our toolchain, with no input other than the compile-time parameters, produces efficient and completely verified code, and we believe it presents a very interesting case study in a verification approach that is unusual in \emph{generating} code, rather than working backwards from existing code.
+(It is worthwhile to note here that one can still verify specific existing code by proving it equivalent to the generated code, and the equivalence is much more straightforward than equivalence with a high-level specification, and we have used this technique to verify existing production code.)
+Our verification stretches all the way from very low-level C to the natural high-level specifications of modular arithmetic (e.g. $\forall x, y.\ \text{add}(x, y) = (x + y) \bmod p$).
+The output code uses a restricted set of instructions by construction, preventing timing side channels.
+The overall trusted computing base includes the Coq proof checker, a simple pretty-printer, and the C language toolchain.
+
+In addition to being interesting research, our approach has proven to be practically applicable, as it is already used in widely deployed code.
+Most applications relying on the implementations described in this paper are using them through BoringSSL, Google's OpenSSL-derived crypto library that includes our implementations of Curve25519 and P-256, the two most widely used elliptic curves via Internet standards.
+A high-profile user is the Chrome Web browser, so today about half of HTTPS connections opened by Web browsers worldwide use our fast verified code (Chrome has about 60\% market share~\cite{browserMarketShare2020}, and 90\% of connections use X25519 or P-256~\cite{davidenTLSECC}).
+% todo
+% now Edge is just rebranded chrome, so +9% if we can find a citation
+% Safari also uses BoringSSL, so +15%, but again nothing to cite, and Apple also may have put their own crypto implementations into BoringSSL and just kept the TLS parts
+BoringSSL is also used in Android and on a majority of Google's servers, so a Google search from Chrome would likely be relying on our code on both sides of the connection.
+Maintainers appreciate the reduced need to worry about bugs.
+
+We believe our approach is particularly suited to design domains that a) involve methodical low-level transformations that are nonetheless finicky and easy for humans to get wrong, and b) apply these transformations to a wide range of setups with different parameters.
+Elliptic-curve field arithmetic fits this description, because most efficient implementations use the same family of optimization techniques, while there are hundreds of combinations of moduli and native-integer sizes.
+A change to either one usually dramatically changes the resulting code, even though the underlying techniques are similar.
+Other opportunities to apply these ideas may include families of very closely related data structures being used in OS kernels and other systems software.
+
+This abridged paper will be a whirlwind tour through the most interesting and informative aspects of our approach.
+First, Section \ref{templates} will go over a common optimization technique for modular reduction and integer representation, as an example of the kinds of algorithms our templates encode.
+Then, in Section \ref{compiler}, we will explain the stages of our domain-specific compiler and its capabilities.
+Finally, Section \ref{experiments} will show a sample of our benchmarks.
+For more details, particularly for proof-engineering techniques, explorations of related work, and full benchmarks, please look to the full-length paper~\cite{FiatCryptoSP19}.
+For even \emph{more} details, the entire development is called \textbf{Fiat Cryptography}, and is available under the MIT license at:
 \begin{center}
@@
 
-The approach we advocate for is already used in widely deployed code.
-Most applications relying on the implementations described in this paper are using them through BoringSSL, Google's OpenSSL-derived crypto library that includes our Curve25519 and P-256 implementations.
-A high-profile user is the Chrome Web browser, so today about half of HTTPS connections opened by Web browsers worldwide use our fast verified code (Chrome versions since 65 have about 60\% market share~\cite{browserMarketShare}, and 90\% of connections use X25519 or P-256~\cite{davidenTLSECC}).
-BoringSSL is also used in Android and on a majority of Google's servers, so a Google search from Chrome would likely be relying on our code on both sides of the connection.
-Maintainers appreciate the reduced need to worry about bugs.
-
-\begin{figure}
-\footnotesize
-\enableverbatimcolors
-\textbf{Input:}
-\begin{verbatim}
-Definition wwmm_step A B k S ret :=
-  divmod_r_cps A (λ '(A, T1),
-  !A[[@scmul_cps r _ T1 B _ (λ aB,]]
-  !B[[@add_cps r _ S aB _ (λ S,]]
-  divmod_r_cps S (λ '(_, s),
-  mul_split_cps' r s k (λ '(q, _),
-  !C[[@scmul_cps r _ q M _ (λ qM,]]
-  !D[[@add_longer_cps r _ S qM _ (λ S,]]
-  divmod_r_cps S (λ '(S, _),
-  @drop_high_cps (S R_numlimbs) S _ (λ S,
-  ret (A, S) ))))))))).
-\end{verbatim}
-\begin{verbatim}
-Fixpoint wwmm_loop A B k len_A ret :=
-  match len_A with
-  | O => ret
-  | S len_A' => λ '(A, S),
-    wwmm_step A B k S (wwmm_loop A B k len_A' ret)
-  end.
-\end{verbatim}
-\textbf{Output:}
-\begin{verbatim}
-void wwmm_p256(u64 out[4], u64 x[4], u64 y[4]) {
-  !A[[u64 x17, x18 = mulx_u64(x[0], y[0]);]]
-  !A[[u64 x20, x21 = mulx_u64(x[0], y[1]);]]
-  !A[[u64 x23, x24 = mulx_u64(x[0], y[2]);]]
-  !A[[u64 x26, x27 = mulx_u64(x[0], y[3]);]]
-  !B[[u64 x29, u8 x30 = addcarryx_u64(0x0, x18, x20);]]
-  !B[[u64 x32, u8 x33 = addcarryx_u64(x30, x21, x23);]]
-  !B[[u64 x35, u8 x36 = addcarryx_u64(x33, x24, x26);]]
-  !B[[u64 x38, u8 _ = addcarryx_u64(0x0, x36, x27);]]
-  !C[[u64 x41, x42 = mulx_u64(x17, 0xffffffffffffffff);]]
-  !C[[u64 x44, x45 = mulx_u64(x17, 0xffffffff);]]
-  !C[[u64 x47, x48 = mulx_u64(x17, 0xffffffff00000001);]]
-  !D[[u64 x50, u8 x51 = addcarryx_u64(0x0, x42, x44);]]
-\end{verbatim}
-  \vspace{-1.8mm}
-\verb|  // |\textbf{\texttt{100 more lines...}}
-  \caption{\label{top-level-example} Word-by-word Montgomery Multiplication, output specialized to P-256 ($2^{256} - 2^{224} + 2^{192} + 2^{96} - 1$)}
-\end{figure}
-
-\section{\label{overview}Overview}
-
-We will explain how we write algorithm implementations in a high-level way that keeps the structural principles as clear and readable as possible, which facilitates proving functional correctness.
-Our explanation builds up to showing how other security constraints can be satisfied without sacrificing the appeal of our new strategy.
-Our output code follows industry-standard ``constant-time'' coding practices and is only bested in throughput by platform-specific assembly-language implementations (and only by a modest margin).
-This section will give a tour of the essential ingredients of our methodology, and the next section will explain in detail how to implement field arithmetic for implementation-friendly primes like $2^{255}-19$.
-Building on this intuition (and code!), Section \ref{saturated} culminates in a specializable implementation of the word-by-word Montgomery reduction algorithm for multiplication modulo any prime.
-Section \ref{bounds} will explain the details of the compiler that turns the described implementations into the kind of code that implementation experts write by hand for each arithmetic primitive.
-We report performance measurements in Section \ref{experiments} and discuss related work and future directions in Sections \ref{related} and \ref{conclusion}.
-
-Fig. \ref{top-level-example} gives a concrete example of what our framework provides.
-The algorithm in the top half of the figure is a transcription of the word-by-word Montgomery reduction algorithm, following Gueron's Algorithm 4~\cite{montexp2011}.
-As is required by the current implementation of our optimization pipeline, this parameter-agnostic code is written in \emph{continuation-passing style} where functions do not return normally but instead take as arguments callbacks to invoke when answers become available.
-The notation \mbox{\small\texttt{λ '(x, y), ...}} encodes an anonymous function with arguments \texttt{\small x} and \texttt{\small y}.
-
-The bottom half of the figure shows the result of specializing the general algorithm to P-256, the elliptic curve most commonly used in TLS and unfortunately not supported by any of the verified libraries mentioned earlier.
-Note that some operations influence the generated code by creating explicit instruction sequences (highlighted in matching colors), while other operations instead control which input expressions should be fed into others.
-The number of low-level lines per high-level line varies between curves and target hardware architectures.
-The generated code is accompanied by a Coq proof of correctness stated in terms of the straightline-code language semantics and the Coq standard-library definitions of integer arithmetic.
-
-\subsection{High-Level Arithmetic Library Designed For Specialization}\label{ex-basesys}
-
-At the most basic level, all cryptographic arithmetic algorithms come down to implementing large mathematical objects (integers or polynomials of several hundred bits) in terms of arithmetic provided by the hardware platform (for example integers modulo $2^{64}$).
-To start out with, let us consider the example of addition, with the simplifying precondition that no digits exceed half of the hardware-supported maximum.
-
-\noindent {\small \begin{align*}
-  &\kw{type} \; \id{num} = \id{list} \; \nat \\
-  &\id{add} : \id{num} \to \id{num} \to \id{num} \\
-  &\id{add} \; (a :: as) \; (b :: bs) = \kw{let} \; x = a + b \; \kw{in} \; x :: \id{add} \; as \; bs \\
-  &\id{add} \; as \; [] = as \\
-  &\id{add} \; [] \; bs = bs
-\end{align*}}%
-
-\noindent To give the specification of this function acting on a base-$2^{64}$ little-endian representation, we define an abstraction function evaluating each digit list $\ell$ into a single number.
-\noindent {\small \begin{align*}
-  &\floor{\_} : \id{num} \to \nat \\
-  &\floor{a::as} = a + 2^{64}\floor{as}\\
-  &\floor{[]} = 0
-\end{align*}}%
-To prove correctness, we show (by naive induction on $a$) that evaluating the result of $\id{add}$ behaves as expected:
-$$\forall a, b. \; \floor{\id{add} \; a \; b} = \floor{a} + \floor{b}$$
-
-\noindent
-Using the standard list type in Coq, the theorem becomes
-{
-\begin{verbatim}
-Lemma eval_add : ∀a b, ⌊add a b⌋ = ⌊a⌋+⌊b⌋.
- induction a,b;cbn; rewrite ?IHa;nia. Qed.
-\end{verbatim}
-}
-
-The above proof is only possible because we are intentionally avoiding details of the underlying machine arithmetic --
-if the list of digits were defined to contain $64$-bit words instead of $\nat$, repeated application of $\id{add}$ would eventually result in integer overflow.
-This simplification is an instance of the pattern of anticipating low-level optimizations in writing high-level code: we do expect to avoid overflow, and our choice of a digit representation is motivated precisely by that aim.
-It is just that the proofs of overflow-freedom will be injected in a later stage of our pipeline, as long as earlier stages like our current one are implemented correctly.
-There is good reason for postponing this reasoning: generally we care about the \emph{context} of higher-level code calling our arithmetic primitives.
-In an arbitrary context, an $\id{add}$ implemented using $64$-bit words would need to propagate carries from each word to the next, causing an unnecessary slowdown when called with inputs known to be small.
-
-\subsection{Partial Evaluation For Specific Parameters}\label{partial}
-
-It is impossible to achieve competitive performance with arithmetic code that manipulates dynamically allocated lists at runtime.
-The fastest code will implement, for example, a single numeric addition with straightline code that keeps as much state as possible in registers.
-Expert implementers today write that straightline code manually, applying various rules of thumb.
-Our alternative is to use \emph{partial evaluation} in Coq to generate all such specialized routines, beginning with a single library of high-level functional implementations that generalize the patterns lurking behind hand-written implementations favored today.
-
-Consider the case where we know statically that each number we add will have 3 digits.
-A particular addition in our top-level algorithm may have the form $\id{add} \; [a_1, a_2, a_3] \; [b_1, b_2, b_3]$, where the $a_i$s and $b_i$s are unknown program inputs.
-While we cannot make compile-time simplifications based on the values of the digits, we \emph{can} reduce away all the overhead of dynamic allocation of lists.
-We use Coq's term-reduction machinery, which allows us to choose $\lambda$-calculus-style reduction rules to apply until reaching a normal form.
-Here is what happens with our example, when we ask Coq to leave $\kw{let}$ and $+$ unreduced but apply other rules.
-
-\noindent {\small \begin{eqnarray*}
-  \id{add} \; [a_1, a_2, a_3] \; [b_1, b_2, b_3] &\Downarrow& \kw{let} \; n_1 = a_1 + b_1 \; \kw{in} \; n_1 :: \\
-  && \kw{let} \; n_2 = a_2 + b_2 \; \kw{in} \; n_2 :: \\
-  && \kw{let} \; n_3 = a_3 + b_3 \; \kw{in} \; n_3 :: []
-\end{eqnarray*}}%
-
-\noindent We have made progress: no run-time case analysis on lists remains.
-Unfortunately, $\kw{let}$ expressions are intermixed with list constructions, leading to code that looks rather different than assembly.
-To persuade Coq's built-in term reduction to do what we want, we first translate arithmetic operations to \emph{continuation-passing style}.
-Concretely, we can rewrite $\id{add}$.
-
-\noindent {\small \begin{align*}
-  &\id{add'} : \forall \alpha. \; \id{num} \to \id{num} \to (\id{num} \to \alpha) \to \alpha \\
-  &\id{add'} \; (a :: as) \; (b :: bs) \; k =\\
-  &\hspace{.1in} \kw{let} \; n = a + b \; \kw{in} \\
-  &\hspace{.1in} \id{add'} \; as \; bs \; (\lambda \ell. \; k \; (n :: \ell)) \\
-  &\id{add'} \; as \; [] \; k = k \; as \\
-  &\id{add'} \; [] \; bs \; k = k \; bs
-\end{align*}}%
-
-\noindent Reduction turns this function into straightline code.
-
-\noindent {\small \begin{eqnarray*}
-  \id{add'} \; [a_1, a_2, a_3] \; [b_1, b_2, b_3] \; (\lambda \ell. \; \ell) &\Downarrow& \kw{let} \; n_1 = a_1 + b_1 \; \kw{in} \\
-  && \kw{let} \; n_2 = a_2 + b_2 \; \kw{in} \\
-  && \kw{let} \; n_3 = a_3 + b_3 \; \kw{in} \\
-  && [n_1, n_2, n_3]
-\end{eqnarray*}}%
-
-\noindent When $\id{add'}$ is applied to a particular continuation, we can also reduce away the result list.
-Chaining together sequences of function calls leads to idiomatic and efficient straightline code, based just on Coq's normal term reduction, preserving sharing of $\kw{let}$-bound variables.
-This level of inlining is common for the inner loops of crypto primitives, and it will also simplify the static analysis described in the next subsection.
-
 \begin{figure*}
@@
 \end{figure*}
-
-\subsection{Word-Size Inference}\label{ex-bounds}
-
-Up to this point, we have derived code that looks almost exactly like the C code we want to produce.
-The code is structured to avoid overflows when run with fixed-precision integers, but so far it is only proven correct for natural numbers.
-The final major step is to infer a range of possible values for each variable, allowing us to assign each one a register or stack-allocated variable of the appropriate bit width.
-
-The bounds-inference pass works by standard abstract interpretation with intervals.
-As inputs, we require lower and upper bounds for the integer values of all arguments of a function.
-These bounds are then pushed through all operations to infer bounds for temporary variables.
-Each temporary is assigned the smallest bit width that can accommodate its full interval.
-
-As an artificial example, assume the input bounds $a_1, a_2, a_3, b_1 \in [0, 2^{31}]$; $b_2, b_3 \in [0, 2^{30}]$.
-The analysis concludes $n_1 \in [0, 2^{32}]$; $n_2, n_3 \in [0, 2^{30} + 2^{31}]$.
-The first temporary is just barely too big to fit in a 32-bit register, while the second two will fit just fine.
-Therefore, assuming the available temporary sizes are 32-bit and 64-bit, we can transform the code with precise size annotations.
-
-\noindent {\small \begin{align*}
-  &\kw{let} \; n_1 : \nat_{2^{64}} = a_1 + b_1 \; \kw{in} \\
-  &\kw{let} \; n_2 : \nat_{2^{32}} = a_2 + b_2 \; \kw{in} \\
-  &\kw{let} \; n_3 : \nat_{2^{32}} = a_3 + b_3 \; \kw{in} \\
-  &[n_1, n_2, n_3]
-\end{align*}}%
-
-\noindent Note how we may infer different temporary widths based on different bounds for arguments.
-As a result, the same primitive inlined within different larger procedures may get different bounds inferred.
-World-champion code for real algorithms takes advantage of this opportunity.
-
-This phase of our pipeline is systematic enough that we chose to implement it as a certified compiler.
-That is, we define a type of abstract syntax trees (ASTs) for the sorts of programs that earlier phases produce, we reify those programs into our AST type, and we run compiler passes written in Coq's Gallina functional programming language.
-Each pass is proved correct once and for all, as Section \ref{bounds} explains in more detail.
-
-\subsection{Compilation To Constant-Time Machine Code}\label{ex-asm}
-
-What results is straightline code very similar to that written by hand by experts, represented as ASTs in a simple language with arithmetic and bitwise operators.
-Our correctness proofs connect this AST to specifications in terms of integer arithmetic, such as the one for $\id{add}$ above.
-All operations provided in our lowest-level AST are implemented with input-independent execution time in many commodity compilers and processors, and if so, our generated code is trivially free of timing leaks.
-Each function is pretty-printed as C code and compiled with a normal C compiler, ready to be benchmarked or included in a library.
-We are well aware that top implementation experts can translate C to assembly better than the compilers, and we do not try to compete with them: while better instruction scheduling and register allocation for arithmetic-heavy code would definitely be valuable, it is outside the scope of this project.
-Nevertheless, we are excited to report that our library generates the fastest-known C code for all operations we have benchmarked (Section \ref{experiments}).
 
@@
 
-\section{Arithmetic Template Library}\label{basesystems}
-
-Recall Section \ref{ex-basesys}'s toy $\id{add}$ function for little-endian big integers.
-We will now describe our full-scale library, starting with the core \emph{unsaturated} arithmetic subset, the foundation for all big-integer arithmetic in our development.
-For those who prefer to read code, we suggest \texttt{src/Demo.v} in the framework's source code, which contains a succinct standalone development of the unsaturated-arithmetic library up to and including prime-shape-aware modular reduction.
+\section{Arithmetic Template Library}\label{templates}
+
+In this section, we will describe the arithmetic optimization strategies we encoded in our template library.
+For those who prefer to read code, we suggest \texttt{src/Demo.v} in the framework's source code, which contains a succinct standalone development of the unsaturated-arithmetic library up to and including an implementation of modular reduction specialized to take advantage of the specific shape of its fixed modulus.
 The concrete examples derived in this section are within established implementation practice, and an expert would be able to reproduce them given an informal description of the strategy.
-Our contribution is to encode this general wisdom in concrete algorithms and provide these with correctness certificates without sacrificing the programmer's sanity.
+Our contribution is to encode this general wisdom in concrete algorithms and provide correctness certificates.
 
@@
 
-Cryptographic modular arithmetic implementations distribute very large numbers across smaller ``limbs'' of 32- or 64-bit integers.
+Cryptographic modular arithmetic implementations distribute very large numbers across smaller \emph{limbs} of 32- or 64-bit integers.
 Fig. \ref{fig:representationexamples} shows a small sample of fast representations for different primes.
@@
 Conversely, the ones using all available bits are called \emph{saturated}.
+This abridged paper will only describe the unsaturated arithmetic in our development; for saturated arithmetic, please refer to the full paper.
 
@@
 
-These unorthodox big integers are fast primarily because of a specific modular-reduction trick, which is most efficient when the number of bits in the prime corresponds to a limb boundary.
-For instance, reduction modulo $2^{255}-19$ is fastest when the bitwidths of a prefix of limbs sum to 255.
-Every unsaturated representation in our examples is designed to fulfill this criterion.
-
-\subsection{Partial Modular Reduction}
-\label{solinasreduce}
-
-Suppose we are working modulo a $k$-bit prime $m$.
-Multiplying two $k$-bit numbers can produce up to $2k$ bits, potentially much larger than $m$.
-However, if we only care about what the result is mod $m$, we can perform a \emph{partial modular reduction} to reduce the upper bound while preserving modular equivalence.
-(The reduction is ``partial'' because the result is not guaranteed to be the minimal residue.)
-
-The most popular choices of primes in elliptic-curve cryptography are of the form $m = 2^k - c_l 2^{t_l} - \ldots - c_0 2^{t_0}$, encompassing what have been called ``generalized Mersenne primes,'' ``Solinas primes,'' ``Crandall primes,'' ``pseudo-Mersenne primes,'' and ``Mersenne primes.''
-Although any number could be expressed this way, and the algorithms we describe would still apply, choices of $m$ with relatively few terms ($l \ll k$) and small $c_i$ facilitate fast arithmetic.
-
-Set $s=2^k$ and $c = c_l 2^{t_l} + \ldots + c_0 2^{t_0}$, so $m = s - c$.
-To reduce $x$ mod $m$, first \emph{split} $x$ by finding $a$ and $b$ such that $x = a + sb$.
-Then, a simple derivation yields a division-free procedure for partial modular reduction.
-% TODO(jgross): fix mod alignment?
-\begin{align*}
-x \bmod m &= (a + sb) \bmod (s - c) \\
-                                             % &= (a(s-c+c) + b) \bmod (s - c) \\
-         &= (a + (s-c)b + cb) \bmod (s - c) \\
-         &= (a + cb) \bmod m
-\end{align*}
-
-The choice of $a$ and $b$ does not further affect the correctness of this formula, but it does influence how much the input is reduced: picking $a=x$ and $b=0$ would make this formula a no-op.
-One might pick $a = x \bmod s$, although the formula does not require it.
-Here is where careful choices of big-integer representation help: having $s$ be at a limb boundary allows for a good split without any computation!
-
-Our Coq proof of this trick is reproduced here:
-{%\footnotesize
-\begin{verbatim}
-Lemma reduction_rule a b s c (_: s-c <> 0)
- : (a + s * b) mod (s - c) =
-   (a + c * b) mod (s - c).
-Proof.
- replace (a+s*b) with ((a+c*b)+b*(s-c)).
- rewrite add_mod,mod_mult,add_0_r,mod_mod.
- all: auto; nsatz.
-Qed.
-\end{verbatim}}
-
-\subsection{Example: Multiplication Modulo $2^{127}-1$}
-
-Before describing the general modular-reduction algorithm implemented in our library, we will walk through multiplication specialized to just one modulus and representation.
-To simplify matters a bit, we use the (relatively) small modulus $2^{127} - 1$.
-Say we want to multiply 2 numbers $s$ and $t$ in its field, with those inputs broken up as $s = s_0 + 2^{43} s_1 + 2^{85} s_2$ and $t = t_0 + 2^{43} t_1 + 2^{85} t_2$.
-Distributing multiplication repeatedly over addition gives us the answer form shown in Fig. \ref{distribute}.
-
-We format the first intermediate term suggestively: down each column, the powers of two are very close together, differing by at most one.
-Therefore, it is easy to add down the columns to form our final answer, split conveniently into digits with integral bit widths.
-%The number of primitive addition and multiplication operations is reasonable (and the primitive multiplications can be implemented as self-additions, which are often faster in hardware).
-
-At this point we have a double-wide answer for multiplication, and we need to do modular reduction to shrink it down to single-wide.
-For our example, the last two digits can be rearranged so that the modular-reduction rule applies:
-% N.B.
-We use \phantom{\vphantom{.}=\vphantom{.}} to make sure that ``='' is aligned with space on either side; the space only shows up when LaTeX registers a character before/after the character, so we fool it into thinking that there is a character there.
-\begin{align*}
-    &\phantom{\vphantom{.}=\vphantom{.}} 2^{127}(2s_1t_2 + 2s_2t_1) + 2^{170}s_2t_2 & \pmod{2^{127}-1} \\
-    &= 2^{127}((2s_1t_2 + 2s_2t_1) + 2^{43}s_2t_2) & \pmod{2^{127}-1} \\
-    &= 1((2s_1t_2 + 2s_2t_1) + 2^{43}s_2t_2) & \pmod{2^{127}-1}
-\end{align*}
-
-As a result, we can merge the second-last digit into the first and merge the last digit into the second, leading to this final formula specifying the limbs of a single-width answer:
-\begin{align*}
-  (s_0t_0 + 2s_1t_2 + 2s_2t_1) &+ 2^{43}(s_0t_1 + s_1t_0 + s_2t_2) \\
-  &+ 2^{85}(s_0t_2 + 2s_1t_1 + s_2t_0)
-\end{align*}
-
-\subsection{Associational Representation}
-% The Central Insight to Simplify Coding and Verifying Generic Arithmetic
-
-As is evident by now, the most efficient code makes use of sophisticated and specific big-number representations, but many concrete implementations operate on the same set of underlying principles.
-Our strategy of writing high-level templates allows us to capture the underlying principles in a fully general way, without committing to a run-time representation.
-Abstracting away the implementation-specific details, like the exact number of limbs or whether the base system is mixed- or uniform-radix, enables simple high-level proofs of all necessary properties.
-
-A key insight that allows us to simplify the arithmetic proofs is to use multiple different representations for big integers, as long as changes between these representations can be simplified away once the prime shape is known.
-We have two main representations of big integers in the core library: \emph{associational} and \emph{positional}.
-A big integer in associational form is represented as a list of pairs; the first element in each pair is a \emph{weight}, known at compile time, and the second is a runtime value.
-The decimal number 95 might be encoded as \texttt{[(16, 5); (1, 15)]}, or equivalently as \texttt{[(1, 5); (10, 9)]} or \texttt{[(1, 1); (1, 6); (44, 2)]}.
-As long as the sum of products is correct, we do not care about the order, whether the weights are powers of each other, or whether there are multiple pairs of equal weight.
-
-In associational format, arithmetic operations are extremely easy to reason about.
-Addition is just concatenation of the lists.
-Schoolbook multiplication (see Fig.
-\ref{assocmul}) is also simple: $(a_1\cdot x_1 + \ldots)(b_1\cdot y_1 + \ldots) = (a_1b_1\cdot x_1y_1 + \ldots)$, where $a_1b_1$ is the new compile-time-known weight.
-Because of the flexibility allowed by associational representation, we do not have to worry about whether the new compile-time weight produced is actually in the base.
-If we are working with the base $2^{25.5}$ in associational representation, and we multiply two pairs that both have weight $2^{26}$, we can just keep the $2^{52}$ rather than the base's $2^{51}$ weight.
-
-\begin{figure}
-  %\footnotesize
-  \begin{verbatim}
-Definition mul (p q:list(Z*Z)) : list(Z*Z)
-:= flat_map (λt,
-    map (λt',(fst t*fst t', snd t*snd t'))
-   q) p.
-	\end{verbatim}
-  \begin{verbatim}
-Lemma eval_map_mul (a x:Z) (p:list (Z*Z))
-  : eval (map (λt, (a*fst t, x*snd t)) p)
-  = a*x * eval p.
-Proof. induction p; push; nsatz.  Qed.
-	\end{verbatim}
-  \begin{verbatim}
-Lemma eval_mul p q
-  : eval (mul p q) = eval p * eval q.
-Proof.
-induction p; cbv [mul]; push; nsatz. Qed.
-\end{verbatim}
-  \caption{\label{assocmul}Definition and correctness proof of multiplication}
-\end{figure}
-
-Positional representation, on the other hand, uses lists of runtime values with weight functions.
-There is only one runtime value for each weight, and order matters.
-This format is closer to how a number would actually be represented at runtime.
-Separating associational from positional representations lets us separate reasoning about arithmetic from reasoning about making weights line up.
-Especially in a system that must account for mixed-radix bases, the latter step is nontrivial and is better not mixed with other reasoning.
-
-\begin{figure}
-  \begin{verbatim}
-Fixpoint place (t:Z*Z) (i:nat) : nat*Z :=
- if dec (fst t mod weight i = 0) then
-  let c:=fst t / weight i in (i, c*snd t)
- else match i with
-      | S i' => place t i'
-      | O => (O, fst t*snd t) end.
-	\end{verbatim}
-  \begin{verbatim}
-Definition from_associational n p :=
-  List.fold_right (λt,
-      let wx := place t (pred n) in
-      add_to_nth (fst wx) (snd wx) )
-    (zeros n) p.
-	\end{verbatim}
-  \begin{verbatim}
-Lemma place_in_range (t:Z*Z) (n:nat)
-: (fst (place t n) < S n)%nat.
-Proof.
- cbv [place]; induction n; break_match;
- autorewrite with cancel_pair; omega. Qed.
-	\end{verbatim}
-  \begin{verbatim}
-Lemma weight_place t i
-: weight (fst(place t i)) * snd(place t i)
-  = fst t * snd t.
-Proof. (* ... 4 lines elided ... *) Qed.
-	\end{verbatim}
-  \begin{verbatim}
-Lemma eval_from_associational {n} p
-: eval (from_associational (S n) p)
-  = Associational.eval p.
-Proof.
- cbv [from_associational]; induction p;
- push;
- try pose proof place_in_range a n;
- try omega; try nsatz.                Qed.
- \end{verbatim}
-  \caption{\label{fromassoc}Coq definition and selected correctness proofs of conversion from associational to positional representation}
-\end{figure}
-
-Fig. \ref{fromassoc} shows the code and proofs for conversion from associational to positional format, including the fixing of misaligned weights.
-The helper function \texttt{place} takes a pair $t = (w, x)$ (compile-time weight and runtime value) and recursively searches for the highest-index output of the weight function that $w$ is a multiple of.
-After finding it, \texttt{place} returns an index $i$ and the runtime value to be added at that index: ($w / weight_i \cdot x$).
-Then, \texttt{from\_associational} simply adds each limb at the correct index.
-
-\begin{figure}
-%\footnotesize
-\begin{verbatim}
-Definition split (s:Z) (p:list (Z*Z))
-  : list(Z*Z)*list(Z*Z) :=
- let hl :=
-   partition (λt, fst t mod s =? 0) p in
- (snd hl,
-  map (λt, (fst t / s, snd t)) (fst hl)).
-\end{verbatim}
-\begin{verbatim}
-Definition reduce(s:Z)(c:list _)(p:list _)
-  : list(Z*Z) :=
- let lh := split s p in
- fst lh ++ mul c (snd lh).
-\end{verbatim}
-\begin{verbatim}
-Lemma eval_split s p (s_nz:s<>0) :
-    eval (fst (split s p))
-     + s * eval (snd (split s p))
-  = eval p.
-Proof. (* ... (7 lines elided) *) Qed.
-\end{verbatim}
-\begin{verbatim}
-Lemma eval_reduce s c p
-       (s_nz:s<>0)(m_nz:s-eval c<>0) :
-    eval (reduce s c p) mod (s - eval c)
-  = eval p mod (s - eval c).
-Proof. cbv [reduce]; push.
- rewrite<-reduction_rule,eval_split; auto.
-Qed.
-\end{verbatim}
-\caption{\label{splitreduce}Coq code for prime-shape-sensitive modular reduction}
-%TODO: if we need to cut a figure, I propose this one
-\end{figure}
-
-The modular-reduction trick we explained in Section \ref{solinasreduce} can also be succinctly represented in associational format (Fig. \ref{splitreduce}).
-First we define \texttt{split}, which partitions the list, separating out pairs with compile-time weights that are multiples of the number at which we are splitting (\texttt{s}, in the code).
-Then it divides all the weights that are multiples of \texttt{s} by \texttt{s} and returns both lists.
-The \texttt{reduce} function simply multiplies the ``high'' list by \texttt{c} and adds the result to the ``low'' list.
-Using the \texttt{mul} we defined earlier recreates the folklore modular-reduction pattern for prime shapes in which \texttt{c} is itself multi-limb.
-
-% The original motivation behind saturated arithmetic is that the sum (or difference, or product) of two $s$-bit values no longer fits in a $s$-bit register: instead, the low and high half of the output are returned separately.
-% However, the correspnding implementation techniques are not locked into operating on radix $w_i = 2^{si}$, they are also useful (for non-performance-critical operations) on the mixed-radix basesystems designed for efficient unsaturated arithmetic.
-% Furthermore, the high level flows for saturated addition and subtraction do not require the inputs to be non-redundant, but nevertheless produce non-redundant output.
-% This allows for a generic way of putting unsaturated representations of field elements into a canonical form, which can then be converted into a saturated representation with simple arithmetic and bitwise operations.
-% In particular: it is sufficient to proceed in three steps.
-% First, our code ensures that the limb $x_n$ with the highest weight $w_n$ is represented canonically by replacing it with two limbs $x_n \bmod w_n$ and $\lfloor x_n / w_n \rfloor$ and applying the modular reduction procedure described above to eliminate the limb at $w_{n+1}$.
-% Assuming $x_i < w_{i+2}/W_i$, this establishes that the overall value is $\le 2w_n$ (we do not actually prove this in general, we run bounds analysis and then prove the specific inequality by \texttt{omega}).
-% Second, our code subtracts the modulus $m$, producing a carry bit.
-% Third, our code adds the modulus, producing a carry bit if and only if the previous subtraction did not produce one.
-% The output limbs of the saturated operation that did not produce a carry bit (and thus represents the original value modulo $m$).
-% We learned this technique from Mike Hamburg (todo:cite Goldilocks code); we would not surprised if it has been presented before.
-%
-% We wrote generic implementations of these algorithms, proved them correct, and generated C code for basesystems used in cryptography.
-% However, even leaving aside verification, we would like to call the reader's attention to the fact that the above descriptions of the algorithms, unlike any implementation of them we have seen, do not include arithmetic on indices of limbs.
-
-\subsection{Carrying}
+Why accept all of this added complication, and waste available bits, instead of just using saturated representations for everything?
+These unorthodox big integers are fast primarily because of a specific modular-reduction algorithm, pseudo-Mersenne reduction, which becomes extremely fast when the number of bits in the prime \emph{corresponds to a limb boundary}.
+For instance, reduction modulo $2^{255}-19$ is fastest when the 255th bit of a bignum is the first bit of a limb.
+In a saturated implementation, limb boundaries would fall on multiples of the integer size (for instance, a 64-bit saturated representation would have limb boundaries at bits 0, 64, 128, etc.).
+But in an unsaturated implementation, with a funky base like $2^{25.5}$, we get limb boundaries at bits 0, 26, 51, 77, ... and 255.
+The vast speed improvement in modular reduction is worth keeping a few extra registers around to represent each large number.
+
+\subsection{A Note on Carrying}
 
@@
 
-\subsection{Saturated Arithmetic}\label{saturated}
-
-The code described so far was designed with unsaturated representations in mind.
-However, unsaturated-arithmetic performance degrades rapidly when used with moduli that it is not a good match for, so older curves such as P-256 need to be implemented using saturated arithmetic.
-Basic arithmetic on saturated digits is complicated by the fact that addition and multiplication outputs may not fit in single words, necessitating the use of two-output addition and multiplication instructions.
-While the partial products in Fig. \ref{assocmul} also need to be stored in pairs of registers, the halves are never separated, avoiding the need to reason about their representation.
-On the other hand, saturated-arithmetic algorithms can be seen as treating carry bits (and the low and high halves of a multiplication) as separate limbs.
-This conceptual framework allows us to reuse most of our unsaturated-associational-arithmetic procedures on saturated representations while still generating the expected low-level code that differs significantly from that of unsaturated arithmetic.
-
-\begin{figure}
-  %\footnotesize
-\begin{verbatim}
-Definition from_associational
-    n (p:list (list Z)) :=
- List.fold_right (λt,
-  let p := place t (pred n) in
-  cons_to_nth (fst p) (snd p)) (nils n) p.
-\end{verbatim}
-  \caption{\label{columnsfromassoc}Conversion from associational to columns}
-\end{figure}
-
-First, we write another version of \texttt{mul} that looks very similar to the unsaturated one but uses a two-output multiplication instruction to compute the partial product of $(a,x)$ and $(b,y)$ as \texttt{let xy := mul x y in [(a*b, fst xy); (a*b*bound, snd xy)]} instead of \texttt{[(a*b, x*y)]}.
-%TODO: is the meaning of bound clear?
-
-Second, rather than using the unsaturated \texttt{from\_associational}, we first convert to an intermediate representation we call \emph{columns}, named after the conventional layout of partial products during pencil-and-paper multiplication.
-While associational representations are lists of pairs of integers (\texttt{list (Z * Z)}) and positional ones are lists of integers (\texttt{list Z}), columns representations are lists of lists of integers (\texttt{list (list Z)}).
-Columns representations are very similar to positional ones, except that the values at each position have not yet been added together.
-Starting from the columns representation, we can add terms together in a specific order so that carry bits are used right after they are produced, enabling them to be stored in flag registers in assembly code.
-The code to convert from associational to columns is very similar to associational-to-positional conversion (Fig. \ref{columnsfromassoc}).
-
-With these changes in place, much of the code shown before for associational works on positional representations as well.
-For instance, the partial modular-reduction implementation still applies unchanged; using the new \texttt{from\_associational} is sufficient.
-The takeaway here is that even completely changing the underlying hardware instructions we used for basic arithmetic did not require redoing all the work from unsaturated representations.
-
-Our most substantial use of saturated arithmetic was for \emph{Montgomery modular reduction}, shown in full earlier in Fig. \ref{top-level-example}.
-For moduli $m$ without convenient special structure, there is not much room to optimize the computation of $ab \bmod m$.
-Instead, it pays off to keep all intermediate values $a$ in a scaled form $aR \bmod m$ for some parameter $R$ and instead optimize the computation of $(aR)(bR)R^{-1} \bmod m = (ab)R \bmod m$.
-We will use underlines to denote the (unique) Montgomery forms of numbers: $\bunderline{a} = aR \bmod m$.
-
-\begin{figure}
-  %\footnotesize
-\begin{verbatim}
-Lemma S2_mod
-: eval (sat_add S1 (scmul q m)) mod m
-  = eval S1 mod m.
-Proof. push; zsimplify; auto. Qed.
-\end{verbatim}
-  \caption{\label{wwmod}Proof that adjusting $S$ to be $0 \bmod m$ preserves the correct value $\bmod\, m$}
-\end{figure}
-
-First, consider the simple case where $S = \bunderline{a}\bunderline{b}$ is divisible by $R$: we can simply divide it by $R$, ignoring the modulus $m$ completely.
-To make division by $R$ fast, we use a minimal saturated representation for $S$ and pick $R$ to match the weight of some limb.
-This way, dividing by $R$ needs no code at all: simply forget about the less significant limbs!
-
-Of course, $S$ is extremely unlikely to be divisible by $R$.
-However, since we only care about the answer $\bmod\, m$, we can add a multiple of $m$ to adjust $S$ to be $0 \bmod R$.
-In particular, we want to add $qm$ such that $qm \bmod R = -(S \bmod R)$.
-Multiplying both sides by $m^{-1} \bmod R$ gives $q = (-m^{-1} \bmod R)(S \bmod R) \bmod R$, which is very cheap to compute: the first factor is a compile-time constant, and $S \bmod R$ is simply the limbs of $S$ below $R$.
-The Coq proof of this step of our word-by-word implementation is in Fig. $\ref{wwmod}$.
-
-\begin{figure}
-  % TODO: get rid of this example?
-  \setlength{\abovedisplayskip}{0pt}%
-  \setlength{\belowdisplayskip}{0pt}%
-  Given \mbox{\small $\texttt{eval}\ S < \texttt{eval}\ m + \texttt{eval}\ B$};\\
-  To show: \par% must be \par not \\ for line spacing, see https://tex.stackexchange.com/a/142950/2066
-  \small
-  $\texttt{eval}\ (\texttt{div\_r}\ (\texttt{add}\ (\texttt{add}\ S\ (\texttt{scmul}\ a\ B))\ (\texttt{scmul}\ q\ m)))$
+\subsection{Example: Multiplication Modulo $2^{127}-1$}
+
+To give a more concrete sense of how the representation described above works in practice, we will walk through a modular multiplication procedure specialized to the (relatively) small modulus $2^{127} - 1$ and a 64-bit architecture.
+Say we want to multiply two numbers $s$ and $t$ in its field, with those inputs broken up into limbs as $s = s_0 + 2^{43} s_1 + 2^{85} s_2$ and $t = t_0 + 2^{43} t_1 + 2^{85} t_2$.
+Distributing multiplication repeatedly over addition gives us the answer form shown in Fig. \ref{distribute}.
+
+We have formatted the calculation suggestively: down each column, the powers of two are very close together, differing by at most one.
+Therefore, it is easy to add down the columns to form our final answer, split conveniently into digits with integral bit widths.
+%The number of primitive addition and multiplication operations is reasonable (and the primitive multiplications can be implemented as self-additions, which are often faster in hardware).
+
+At this point the multiplication step is complete, but we still need to do modular reduction; we don't want our answer to have five limbs when it should need only three.
+Here we use the pseudo-Mersenne reduction trick: $2^k \bmod {(2^k -c)} = c$, so $a + 2^kb \equiv a + cb \pmod{2^k - c}$.
+We ``divide'' the last two limbs of the product, the terms with weights $2^{127}$ and $2^{170}$, by $2^{127}$.
+Because we chose our base system wisely, this division is actually a no-op; we just now consider these limbs to have different weights.
+This is the big trick that makes pseudo-Mersenne reduction fast: eliminating division by clever choice of base system.
+Then we can multiply the high limbs by $c$, which in this case is 1, and add them to the first three digits, providing the following final formula:
 \begin{align*}
-    & < \texttt{eval}\ m + \texttt{eval}\ B \\
-    (\texttt{eval}\ S + a*\texttt{eval}\ B + q*\texttt{eval}\ m)/r
-    & < \texttt{eval}\ m + \texttt{eval}\ B \\
-(\texttt{eval}\ m * r + r * \texttt{eval}\ B - 1)/r
-    & < \texttt{eval}\ m + \texttt{eval}\ B \\
-\texttt{eval}\ m + \texttt{eval}\ B - 1
-    & < \texttt{eval}\ m + \texttt{eval}\ B
+  (s_0t_0 + 2s_1t_2 + 2s_2t_1) &+ 2^{43}(s_0t_1 + s_1t_0 + s_2t_2) \\
+  &+ 2^{85}(s_0t_2 + 2s_1t_1 + s_2t_0)
 \end{align*}
-  \caption{\label{wwbound} Intermediate goals of the proof that the word-by-word Montgomery reduction state value remains bounded by a constant, generated by single-stepping our proof script}
-\end{figure}
-
-This procedure is further optimized based on the observation that it only acts on the higher limbs of $S$ additively.
-For $n$-limb $\bunderline{a}$, instead of first computing $\bunderline{a}\bunderline{b}$ and then dividing it by $R=r^n$, it is fine to divide eagerly by $r$ after multiplying $\bunderline{b}$ with a single limb of $\bunderline{a}$ -- adding the partial products with the more significant limbs of $\bunderline{a}$ commutes with adding multiples of $m$.
-Even better, adding a multiple of $m$ and a multiple of $\bunderline{b}$ can only increase the length of $S$ by one limb, balancing out the loss of one limb when dividing $S$ by $r$ and allowing for an implementation that loops through limbs of $\bunderline{a}$ while operating in-place on $S$.
-The proof of the last fact is delicate -- for example, it is not true that one of the additions always produces a newly nonzero value in the most significant limb and the other never does.
-In our library, this subtlety is handled by a 15-line/100-word proof script; the high-level steps, shown in Fig.
-\ref{wwbound}, admittedly fail to capture the amount of thought that was required to find the invariant that makes the bounds line up exactly right.
-
-%% \subsection{The Story So Far}
-
-%% Let us take stock of how the final packaging of this section's results fits into the overall story of the paper.
-%% We have \emph{generic} implementations of key modular-arithmetic algorithms coded as functional programs in Coq.
-%% Each operation is proved correct once and for all.
-%% Moreover, the operations' input and output formats have been chosen to enforce a clear separation between compile-time (e.g., weight of a digit) and runtime (e.g., a digit itself) values, though nothing formal has been proved about that separation.
-%% Instead, in the context of the top-level code for a primitive, where the compile-time values are apparent, we apply \emph{partial evaluation} as explained in Section \ref{partial}.
-%% The result is, for each instance of an arithmetic operation, relatively simple first-order code.
-%% That code is the input to our final compiler phases, described next.
-
-\section{Certified Bounds Inference}\label{bounds}
-
-Recall from Section \ref{partial} how we use partial evaluation to specialize the functions from the last section to particular parameters.
-The resulting code is elementary enough that it becomes more practical to apply relatively well-understood ideas from \emph{certified compilers}.
-That is, as sketched in Section \ref{ex-bounds}, we can define an explicit type of program abstract syntax trees (ASTs), write compiler passes over it as Coq functional programs, and prove those passes correct.
-%once and for all. The contrast is with earlier transformation phases, which work using Coq's proof-generating tactics.
-
-\subsection{Abstract Syntax Trees}
-
-The results of partial evaluation fit, with minor massaging, into this intermediate language:
-$$\begin{array}{rrcl}
-  \textrm{Base types} & b \\
-  \textrm{Types} & \tau &::=& b \mid \kw{unit} \mid \tau \times \tau \\
-  \textrm{Variables} & x \\
-  \textrm{Operators} & o \\
-  \textrm{Expressions} & e &::=& x \mid o(e) \mid () \mid (e, e) \\
-  &&& \mid \kw{let} \; (x_1, \ldots, x_n) = e \; \kw{in} \; e
-\end{array}$$
-
-Types are trees of pair-type operators $\times$ where the leaves are one-element $\kw{unit}$ types and base types $b$, the latter of which come from a domain that is a parameter to our compiler.
-It will be instantiated differently for different target hardware architectures, which may have different primitive integer types.
-When we reach the certified compiler's part of the pipeline, we have converted earlier uses of lists into tuples, so we can optimize away any overhead of such value packaging.
-
-Another language parameter is the set of available primitive operators $o$, each of which takes a single argument, which is often a tuple of base-type values.
-Our $\kw{let}$ construct bakes in destructuring of tuples, in fact using typing to ensure that all tuple structure is deconstructed fully, with variables bound only to the base values at a tuple's leaves.
-Our deep embedding of this language in Coq uses dependent types to enforce that constraint, along with usual properties like lack of dangling variable references and type agreement between operators and their arguments.
-
-Several of the key compiler passes are polymorphic in the choices of base types and operators, but bounds inference is specialized to a set of operators.
-We assume that each of the following is available for each type of machine integers (e.g., 32-bit vs. 64-bit).
-$$\begin{array}{@{}l@{}}
-  \textrm{Integer literals: } n \\
-  \textrm{Unary arithmetic operators: } -e \\
-  \textrm{Binary arithmetic operators: } e_1 + e_2, e_1 - e_2, e_1 \times e_2 \\
-  \textrm{Bitwise operators: } e_1 \ll e_2, e_1 \gg e_2, e_1 \; \& \; e_2, e_1 \mid e_2 \\
-  \textrm{Conditionals: } \mathsf{if} \; e_1 \neq 0 \; \mathsf{then} \; e_2 \; \mathsf{else} \; e_3 \\
-  \textrm{Carrying: } \mathsf{addWithCarry}(e_1, e_2, c), \mathsf{carryOfAdd}(e_1, e_2, c) \\
-  \textrm{Borrowing: } \mathsf{subWithBorrow}(c, e_1, e_2), \mathsf{borrowOfSub}(c, e_1, e_2) \\
-  \textrm{Two-output multiplication: } \mathsf{mul2}(e_1, e_2)
-%| IdWithAlt T1 T2 Tout : op (Tbase T1 * Tbase T2) (Tbase Tout)
-\end{array}$$
-
-We explain only the last three categories, since the earlier ones are familiar from C programming.
-%TODO: agl drew a \rfloor symbol here, idk why
-To chain together multiword additions, as discussed in the prior section, we need to save overflow bits (i.e., carry flags) from earlier additions, to use as inputs into later additions.
-The $\mathsf{addWithCarry}$ operation implements this three-input form, while $\mathsf{carryOfAdd}$ extracts the new carry flag resulting from such an addition.
-Analogous operators support subtraction with \emph{borrowing}, again in the grade-school-arithmetic sense.
-Finally, we have $\mathsf{mul2}$ to multiply two numbers to produce a two-number result, since multiplication at the largest available word size may produce outputs too large to fit in that word size.
-
-All operators correspond directly to common assembly instructions.
-Thus the final outputs of compilation look very much like assembly programs, just with unlimited supplies of temporary variables, rather than registers.
-$$\begin{array}{rrcl@{}}
-  \textrm{Operands} & O &::=& x \mid n \\
-  \textrm{Expressions} & e &::=& \kw{let} \; (x_1, \ldots, x_n) = o(O, \ldots, O) \; \kw{in} \; e \\
-  &&& \mid (O, \ldots, O)
-\end{array}$$
-
-We no longer work with first-class tuples.
-Instead, programs are sequences of primitive operations, applied to constants and variables, binding their (perhaps multiple) results to new variables.
-A function body, represented in this type, ends in the function's (perhaps multiple) return values.
-
-Such functions are easily pretty-printed as C code, which is how we compile them for our experiments.
-Note also that the language enforces the \emph{constant time} security property by construction: the running time of an expression leaks no information about the values of the free variables.
-We do hope, in follow-on work, to prove that a particular compiler preserves the constant-time property in generated assembly.
-For now, we presume that popular compilers like GCC and Clang do preserve constant time, modulo the usual cycle of occasional bugs and bug fixes.
-(One additional source-level restriction is important, forcing conditional expressions to be those supported by native processor instructions like conditional move.)
-
-\subsection{Phases of Certified Compilation}
-
-%% \begin{figure*}
-%%   \centering
-%%   \tikzstyle{rf} = [rectangle, draw=black, inner sep=3pt]
-%%   \tikzstyle{rrf} = [rounded rectangle, draw=black, inner sep=3pt]
-%%   \tikzstyle{line}=[-]
-%%   \tikzstyle{arrow}=[<-]
-%%   \begin{tikzpicture}
-%%     \node[anchor=south] (root) at (0, 0) [rf] {
-%%       Gallina
-%%     };
-%%     \node[anchor=south] (n1) at (1.9, 0) [rf] {
-%%       AST
-%%     };
-%%     \node[anchor=south] (n2) at (4.15, 0) [rf] {
-%%       flat
-%%     };
-%%     \node[anchor=south] (n3) at (7.9, 0) [rf] {
-%%       flat
-%%     };
-%%     \node[anchor=south] (n4) at (10.00, 0) [rf] {
-%%       flat
-%%     };
-%%     \node[anchor=south] (n5) at (11.8, 0) [rf] {
-%%       flat
-%%     };
-
-%%     \node[anchor=east] (n1exp) at (1.5, 0.5) {reify};
-%%     \node[anchor=east] (n2exp) at (3.75, 0.5) {linearize};
-%%     \node[anchor=east] (n3exp) at (7.5, 0.5) {const.-fold \& inline};
-%%     \node[anchor=east] (n4exp) at (9.5, 0.5) {bounds};
-%%     \node[anchor=east] (n5exp) at (11.4, 0.5) {$\eta$-exp.};
-
-%%     \draw [arrow] (n1) to (root);
-%%     \draw [arrow] (n2) to (n1);
-%%     \draw [arrow] (n3) to (n2);
-%%     \draw [arrow] (n4) to (n3);
-%%     \draw [arrow] (n5) to (n4);
-%%   \end{tikzpicture}
-%%   \caption{Phases of certified compilation}
-%%   \label{fig-certified}
-%% \end{figure*}
-
-To begin the certified-compilation phase of our pipeline, we need to \emph{reify} native Coq programs as terms of this AST type.
-To illustrate the transformations we perform on ASTs, we walk through what the compiler does to an example program:
+
+Note that this is a \emph{partial} modular-reduction algorithm; it is not guaranteed that the result is less than $2^{127}-1$, just that the result fits in three limbs.
+
+In the code, \texttt{src/Demo.v} includes another example that walks through modular multiplication with pseudo-Mersenne reduction, using the modulus from Curve25519~\cite{curve25519}.
+The full generic template library is found under \texttt{src/Arithmetic}.
+
+\section{Domain-Specific Compiler}\label{compiler}
+
+\subsection{Partial Evaluation For Specific Parameters}\label{partial}
+
+It is impossible to achieve competitive performance with arithmetic code that manipulates dynamically allocated lists at runtime.
+The fastest code will implement, for example, a single numeric addition with straightline code that keeps as much state as possible in registers.
+Expert implementers today write that straightline code manually, applying various rules of thumb.
+Our alternative is to use \emph{partial evaluation} in Coq to generate all such specialized routines, beginning with our library of high-level functional implementations that generalize the patterns lurking behind hand-written implementations today.
+
+Consider the case where we know statically that each number we add will have 3 digits.
+A particular addition in our top-level algorithm may have the form $\id{add} \; [a_1, a_2, a_3] \; [b_1, b_2, b_3]$, where the $a_i$s and $b_i$s are unknown program inputs.
+While we cannot make compile-time simplifications based on the values of the digits, we \emph{can} reduce away all the overhead of dynamic allocation of lists.
+We could use Coq's term-reduction machinery, which allows us to choose $\lambda$-calculus-style reduction rules to apply until reaching a normal form.
+Here is what happens with our example, when we ask Coq to leave $\kw{let}$ and $+$ unreduced but apply other rules.
+(Note that :: is a notation for appending to the front of a list.)
+
+\noindent {\small \begin{eqnarray*}
+  \id{add} \; [a_1, a_2, a_3] \; [b_1, b_2, b_3] &\Downarrow& \kw{let} \; n_1 = a_1 + b_1 \; \kw{in} \; n_1 :: \\
+  && \kw{let} \; n_2 = a_2 + b_2 \; \kw{in} \; n_2 :: \\
+  && \kw{let} \; n_3 = a_3 + b_3 \; \kw{in} \; n_3 :: []
+\end{eqnarray*}}%
+
+\noindent We have made progress: no run-time case analysis on lists remains.
+Unfortunately, $\kw{let}$ expressions are intermixed with list constructions, leading to code that looks rather different than assembly.
+To work around this issue, we chose to implement this phase of our pipeline as a certified compiler\footnote{The compiler was ongoing work at the time of our original publication and was only briefly discussed as a future improvement; in the original paper, we described using Coq's built-in term-reduction machinery.
+However, although the proof-engineering techniques have changed dramatically, the function is the same.
+}.
+That is, we define a type of abstract syntax trees (ASTs) for the sorts of programs that earlier phases produce, we reify those programs into our AST type, and we run compiler passes written in Coq's Gallina functional programming language.
+Each pass is proved correct once and for all.
+
+Our certified compiler handles partial evaluation and $\kw{let}$-lifting, turning this function into the straightline code
+\noindent {\small \begin{eqnarray*}
+  && \kw{let} \; n_1 = a_1 + b_1 \; \kw{in} \\
+  && \kw{let} \; n_2 = a_2 + b_2 \; \kw{in} \\
+  && \kw{let} \; n_3 = a_3 + b_3 \; \kw{in} \\
+  && [n_1, n_2, n_3]
+\end{eqnarray*}}%
+
+Chaining together sequences of function calls leads to idiomatic and efficient straightline code, preserving sharing of $\kw{let}$-bound variables.
+This level of inlining is common for the inner loops of crypto primitives, and it will also simplify the static analysis described in the next subsection.
+
+\subsection{Word-Size Inference}\label{ex-bounds}
+
+Up to this point, we have derived code that looks almost exactly like the C code we want to produce.
+The code is structured to avoid overflows when run with fixed-precision integers, but so far it is only proven correct for natural numbers.
+The final major step is to infer a range of possible values for each variable, allowing us to assign each one a register or stack-allocated variable of the appropriate bit width.
+
+The bounds-inference pass works by standard abstract interpretation with intervals.
+As inputs, we require lower and upper bounds for the integer values of all arguments of a function.
+These bounds are then pushed through all operations to infer bounds for temporary variables.
+Each temporary is assigned the smallest bit width that can accommodate its full interval.
+
+As an artificial example, assume the input bounds $a_1, a_2, a_3, b_1 \in [0, 2^{31}]$; $b_2, b_3 \in [0, 2^{30}]$.
+The analysis concludes $n_1 \in [0, 2^{32}]$; $n_2, n_3 \in [0, 2^{30} + 2^{31}]$.
+The first temporary is just barely too big to fit in a 32-bit register, while the second two will fit just fine.
+Therefore, assuming the available temporary sizes are 32-bit and 64-bit, we can transform the code with precise size annotations.
 
 \noindent {\small \begin{align*}
-  &\kw{let} \; (x_1, x_2, x_3) = x \; \kw{in} \\
-  &\kw{let} \; (y_1, y_2) = ((\kw{let} \; z = x_2 \times 1 \times x_3 \; \kw{in} \; z + 0), x_2) \; \kw{in} \\
-  &y_1 \times y_2 \times x_1
+  &\kw{let} \; n_1 : \nat_{2^{64}} = a_1 + b_1 \; \kw{in} \\
+  &\kw{let} \; n_2 : \nat_{2^{32}} = a_2 + b_2 \; \kw{in} \\
+  &\kw{let} \; n_3 : \nat_{2^{32}} = a_3 + b_3 \; \kw{in} \\
+  &[n_1, n_2, n_3]
 \end{align*}}%
 
-\noindent The first phase is \emph{linearize}, which cancels out all intermediate uses of tuples and immediate $\kw{let}$-bound variables and moves all $\kw{let}$s to the top level.
-%Optionally, this phase can also create temporaries for all subexpressions of multiple-operation terms, putting the output in A-normal form (ANF)~\cite{ANF}.
-
-\noindent {\small \begin{align*}
-  &\kw{let} \; (x_1, x_2, x_3) = x \; \kw{in} \\
-  &\kw{let} \; z = x_2 \times 1 \times x_3 \; \kw{in} \\
-  &\kw{let} \; y_1 = z + 0 \; \kw{in} \\
-  &y_1 \times x_2 \times x_1
-\end{align*}}%
-
-\noindent Next is \emph{constant folding}, which applies simple arithmetic identities and inlines constants and variable aliases.
-
-\noindent {\small \begin{align*}
-  &\kw{let} \; (x_1, x_2, x_3) = x \; \kw{in} \\
-  &\kw{let} \; z = x_2 \times x_3 \; \kw{in} \\
-  &z \times x_2 \times x_1
-\end{align*}}%
-
-\noindent At this point we run the core phase, \emph{bounds inference}, the one least like the phases of standard C compilers.
-The phase is parameterized over a list of available fixed-precision base types with their ranges; for our example, assume the hardware supports bit sizes 8, 16, 32, and 64.
-Intervals for program inputs, like $x$ in our running example, are given as additional inputs to the algorithm.
-Let us take them to be as follows: $x_1 \in [0, 2^8], x_2 \in [0, 2^{13}], x_3 \in [0, 2^{18}]$.
-The output of the algorithm has annotated each variable definition and arithmetic operator with a finite type.
-
-\noindent {\small \begin{align*}
-  &\kw{let} \; (x_1 : \nat_{2^{16}}, \; x_2 : \nat_{2^{16}}, \; x_3 : \nat_{2^{32}}) = x \; \kw{in} \\
-  &\kw{let} \; z : \nat_{2^{32}} = x_2 \times_{\nat_{2^{32}}} x_3 \; \kw{in} \\
-  &z \times_{\nat_{2^{64}}} x_2 \times_{\nat_{2^{64}}} x_1
-\end{align*}}%
-
-\noindent Our biggest proof challenge here was in the interval rules for bitwise operators applied to negative numbers, a subject mostly missing from Coq's standard library.
-%% Bitwise operators are essential for full-scale crypto primitives, and these operators pose an additional challenge in our full implementation.
-%% In particular, Coq computes bitwise operations on negative integers using an infinite-width two's-complement representation.
-%% Much of our time in the bounds-analysis proof was spent on appropriate bounds for negative as well as nonnegative inputs, since Coq's standard library contains few theorems about applying bitwise operators to negative numbers (but they can show up in crypto algorithms).
-
-%The final compiler phase is \emph{$\eta$-expansion}.
-%Because our $\kw{let}$ construct bakes in the decomposition of tuples, execution of the previous passes will block on any $\kw{let}$ binding for an opaque operator that returns a tuple (e.g., two-output multiplication), because Coq cannot ``peek underneath'' the operator to determine the rest of the syntax tree.
-%The $\eta$-expansion phase destructures all binders, passing each reconstructed tuple to the rest of the syntax tree, allowing reduction of the previous phases to proceed.
-
-\subsection{Important Design Choices}
-
-%\paragraph{Encoding term structure in Coq.}
-
-Most phases of the compiler use a term encoding called parametric higher-order abstract syntax (PHOAS)~\cite{PHOAS}.
-Briefly, this encoding uses variables of the metalanguage (Coq's Gallina) to encode variables of the object language, to avoid most kinds of bookkeeping about variable environments; and for the most part we found that it lived up to that promise.
-However, we needed to convert to a first-order representation (de Bruijn indices) and back for the bounds-inference phase, essentially because it calls for a forward analysis followed by a backward transformation: calculate intervals for variables, then rewrite the program bottom-up with precise base types for all variables.
-We could not find a way with PHOAS to write a recursive function that returns both bounds information and a new term, taking better than quadratic time, while it was trivial to do with first-order terms.
-We also found that the established style of term well-formedness judgment for PHOAS was not well-designed for large, automatically generated terms like ours: proving well-formedness would frequently take unreasonably long, as the proof terms are quadratic in the size of the syntax tree.
-The fix was to switch well-formedness from an inductive definition into an executable recursive function that returns a linear number of simple constraints in propositional logic.
-% where atomic formulas are equalities between variables, which admit more compact proofs.
-
-\subsection{Extensibility with Nonobvious Algebraic Identities}
-
-Classic abstract interpretation with intervals works surprisingly well for most of the cryptographic code we have generated.
-However, a few spans of code call for some algebra to establish the tightest bounds.
-We considered extending our abstract interpreter with general algebraic simplification, perhaps based on E-graphs as in SMT solvers~\cite{simplify}, but in the end we settled on an approach that is both simpler and more accommodating of unusual reasoning patterns.
-The best-known among our motivating examples is Karatsuba's multiplication, which expresses a multi-digit multiplication in a way with fewer single-word multiplications (relatively slow) but more additions (relatively fast), where bounds checking should recognize algebraic equivalence to a simple multiplication.
-Naive analysis would treat $(a + b)(c+d) - ac$ equivalently to $(a+b)(c+d) - xy$, where $x$ and $y$ are known to be within the same ranges as $a$ and $c$ but are not necessarily equal to them.
-The latter expression can underflow.
-We add a new primitive operator $\mathsf{equivalently}$ such that, within the high-level functional program, we can write the above expression as $\mathsf{equivalently}\left((a+b)(c+d)-ac, ad+bc+bd\right)$.
-The latter argument is used for bounds analysis.
-The soundness theorem of bounds analysis requires, as a side condition, equality for all pairs of arguments to $\mathsf{equivalently}$.
-These side conditions are solved automatically during compilation using Coq tactics, in this case \texttt{ring} algebra.
-Compilation then replaces any $\mathsf{equivalently}\left(e_1, e_2\right)$ with just $e_1$.
+\noindent Note how we may infer different temporary widths based on different bounds for arguments.
+As a result, the same primitive inlined within different larger procedures may get different bounds inferred.
+World-champion code for real algorithms takes advantage of this opportunity.
+
+This phase of our pipeline is systematic enough that we chose to implement it too as another phase in our certified compiler.
+
+\subsection{Compilation To Constant-Time Machine Code}\label{ex-asm}
+
+What results is straightline code very similar to that written by hand by experts, represented as ASTs in a simple language with arithmetic and bitwise operators.
+Our correctness proofs connect this AST to specifications in terms of integer arithmetic, such as the one for $\id{add}$ above.
+All operations provided in our lowest-level AST are implemented with input-independent execution time in many commodity compilers and processors, and if so, our generated code is trivially free of timing leaks.
+Each function is pretty-printed as C code and compiled with a normal C compiler, ready to be benchmarked or included in a library.
+We are well aware that top implementation experts can translate C to assembly better than the compilers, and we do not try to compete with them: while better instruction scheduling and register allocation for arithmetic-heavy code would definitely be valuable, it is outside the scope of this project.
+But this is the entire extent of the compromise: as described in the next section, we have been able to match or exceed the performance of all C code we sought to replicate.
 
@@
 
-The first purpose of this section is to confirm that implementing optimized algorithms in high-level code and then separately specializing to concrete parameters actually achieves the expected performance.
+The purpose of this section is to confirm that implementing optimized algorithms in high-level code and then separately specializing to concrete parameters actually achieves the expected performance.
 Given the previous sections, this conclusion should not be surprising: as code generation is extremely predictable, it is fair to think of the high-level implementations as simple templates for low-level code.
-Thus, it would be possible write new high-level code to mimic every low-level implementation.
-Such a strategy would, of course, be incredibly unsatisfying, although perhaps still worthwhile as a means to reduce effort required for correctness proofs.
-Thus, we wish additionally to demonstrate that the two simple generic-arithmetic implementations discussed in Sections \ref{solinasreduce} and \ref{saturated} are sufficient to achieve good performance in all cases.
-Furthermore, we do our best to call out differences that enable more specialized implementations to be faster, speculating on how we might modify our framework in the future based on those differences.
+We will detail how this works out for the two most prominent applications of our framework, the TLS ECC fields modulo $2^{255}-19$ and $2^{256}-2^{224}+2^{192}+2^{96}-1$.
+In our full paper, we additionally demonstrate that the two simple templates generalized from these case studies are sufficient to achieve good performance across a broad sample of finite fields proposed for elliptic-curve-cryptography use.
 
@@
 
-We measure the number of CPU cycles different implementations take to multiply a secret scalar and a public Curve25519 point (represented by the $x$ coordinate in Montgomery coordinates).
-Despite the end-to-end task posed for this benchmark, we believe the differences between implementations we compare against lie in the field-arithmetic implementation --
-all use the same scalar-multiplication algorithm and two almost-identical variants of the Montgomery-curve $x$-coordinate differential-addition formulas.
-
-The benchmarks were run on an Intel Broadwell i7-5600U processor in a kernel module
+We measured the number of CPU cycles different implementations take to multiply a secret scalar and a public Curve25519 point (represented by the $x$ coordinate in Montgomery coordinates).
+Despite the end-to-end task posed for this benchmark, we believe the differences between implementations we compare against lie in the field-arithmetic implementation.
+% They all use the same scalar-multiplication algorithm and two almost-identical variants of the Montgomery-curve $x$-coordinate differential-addition formulas.
+
+The benchmarks were run using \texttt{gcc} 7.3 on an Intel Broadwell i7-5600U processor in a kernel module
 %\footnote{Thanks to Jason Donenfeld for teaching us how and setting up benchmarks for popular Curve25519 implementations! %TODO:rephrase? Moved to acknowledgments
@@
 with interrupts, power management, Hyper Threading, and Turbo Boost features disabled.
-Each measurement represents the average of a batch of 15,000 consecutive trials, with time measured using the \texttt{RDTSC} instruction and converted to CPU cycles by multiplying by the ratio of CPU and timestamp-counter clock speeds.
-C code was compiled using \texttt{gcc} 7.3 with \texttt{-O3 -march=native -mtune=native -fwrapv} (we also tried \texttt{clang} 5.0, but it produced $\sim$10\% slower code for the implementations here).
+We are presenting them as a rather arbitrary illustration of the performance that can be achieved using our approach; it is expected that the relative speeds could differ by as much as a couple of times across different microarchitectures, compilers, and compiler flags.
+% Each measurement represents the average of a batch of 15,000 consecutive trials, with time measured using the \texttt{RDTSC} instruction and converted to CPU cycles by multiplying by the ratio of CPU and timestamp-counter clock speeds.
+% C code was compiled using \texttt{gcc} 7.3 with \texttt{-O3 -march=native -mtune=native -fwrapv} (we also tried \texttt{clang} 5.0, but it produced $\sim$10\% slower code for the implementations here).
 
@@
 \texttt{amd64-64}, asm       & \texttt{151586} & \makeothercpu{151586} \\ % 158029
-\textit{this work B}, 64-bit & \texttt{152195} & \makethiscpu{152195} \\ % 158664
+\textit{this work}, 64-bit & \texttt{152195} & \makethiscpu{152195} \\ % 158664
+%\textit{this work B}, 64-bit & \texttt{152195} & \makethiscpu{152195} \\ % 158664
 \texttt{sandy2x}, asm        & \texttt{154313} & \makeothercpu{154313} \\ % 160872
@@
 \texttt{donna64}, 64-bit C   & \texttt{168502} & \makeothercpu{168502} \\ % 175664
-\textit{this work A}, 64-bit & \texttt{174637} & \makethiscpu{174637} \\ % 182060
-\textit{this work, 32-bit  } & \texttt{310585} & \makethiscpu{310585} \\ % 323786
-\texttt{donna32}, 32-bit C   & \texttt{529812} & \makeothercpu{529812} \\ % 552331
+%\textit{this work A}, 64-bit & \texttt{174637} & \makethiscpu{174637} \\ % 182060
+%\textit{this work, 32-bit  } & \texttt{310585} & \makethiscpu{310585} \\ % 323786
+%\texttt{donna32}, 32-bit C   & \texttt{529812} & \makeothercpu{529812} \\ % 552331
 \hline
@@
 
+We report on our code generated using the standard representations for both 32-bit and 64-bit, though we are primarily interested in the latter, since we benchmark on a 64-bit processor.
 In order, we compare against \texttt{OLHFR}, the non-precomputation-based implementation from \cite{oliveira_sac2017}; \texttt{amd64-64} and \texttt{sandy2x}, the fastest assembly implementations from SUPERCOP~\cite{SUPERCOP} that use scalar and vector instructions respectively; the verified X25519 implementation from the HACL$^*$ project~\cite{hacl}; and the best-known high-performance C implementation \texttt{curve25519-donna}, in both 64-bit and 32-bit variants.
-We measured cycle counts for all comparison implementations from a checkout of kbench9000~\cite{kbench} at commit \texttt{59063a52}.
 The field arithmetic in both \texttt{amd64-64} and \texttt{hacl-star} has been verified using SMT solvers~\cite{verif25519,ECCstar}.
-Of course, care should be taken when extrapolating these results to other compilers and CPUs: each assembly implementation is likely to be fastest on the microarchitecture it was optimized for, and performance of C code additionally varies with the quality of compiler optimizations for the microarchitecture in question.
-
-We report on our code generated using the standard representations for both 32-bit and 64-bit, though we are primarily interested in the latter, since we benchmark on a 64-bit processor.
-We actually report on two of our 64-bit variants, both of which have correctness proofs of the same strength.
-Variant \emph{A} is the output of the toolchain described in Section \ref{basesystems}.
-Variant \emph{B} contains an additional implementation grouping multiplications by the modular-reduction coefficient $c=19$ with one of the inputs of the multiplication instead of the output ($19\times (a\times b) \longrightarrow (19\times a)\times b$).
-This algebra pays off because $19\times a$ can be computed using a 64-bit multiplication, whereas $19\times (ab)$ would require a significantly more expensive 128-bit multiplication.
-We currently implement this trick for just $2^{255}-19$ as an ad-hoc code replacement after partial evaluation but before word-size inference, using the Coq \texttt{ring} algebra tactic to prove that the code change does not change the output.
-We are planning on incorporating a more general version of this optimization into the main pipeline, after a planned refactoring to add automatic translation to continuation-passing style (see Section \ref{conclusion}).
-
-The results of a similar benchmark on an earlier version of our pipeline were good enough to convince the maintainers of the BoringSSL library to adopt our methodology, resulting in this Curve25519 code being shipped in Chrome 64 and used by default for TLS connection establishment in other Google products and services.
+Our code is generated from a general template for unsaturated pseudo-mersenne arithmetic which is proven correct for all parameters.
+We previously applied a few optimizations to $2^{255}-19$ specifically, but have since integrated the generalized forms of these into our template.
+
+The results of a similar benchmark on an early prototype of our methodology were good enough to convince the maintainers of the BoringSSL library to adopt it, resulting in this Curve25519 code being shipped since Chrome 64 and used by default for TLS connection establishment in other Google products and services.
 Previously, BoringSSL included the \texttt{amd64-64} assembly code and a 32-bit C implementation as a fallback, which was the first to be replaced with our generated code.
@@
 
-Next, we benchmark our Montgomery modular arithmetic as used for in-place point addition on the P-256 elliptic curve with one precomputed input (Jacobian += affine).
+Next, we benchmarked our Montgomery modular arithmetic as used for in-place point addition on the P-256 elliptic curve with one precomputed input (Jacobian += affine).
 A scalar-multiplication algorithm using precomputed tables would use some number of these additions depending on the table size.
-The assembly-language implementation \texttt{nistz256} was reported on by Gueron and Krasnov~\cite{nistz256} and included in OpenSSL; we also measure its newer counterpart that makes use of the ADX instruction-set extension.
+The assembly-language implementation \texttt{nistz256} was reported on by Gueron and Krasnov~\cite{nistz256} and included in OpenSSL; we also measured its newer counterpart that makes use of the ADX instruction-set extension.
 % Our choice to use Montgomery reduction was inspired by the impressive performance of that implementation.
@@
 \end{tabular}\end{center}
-
- \begin{figure*}
- \begin{tikzpicture}
- 	\begin{axis}[
- 		height=3.4cm,
- 		title style={font=\small},
- 		title=64-Bit Field Arithmetic Benchmarks,
- 		width=\textwidth,
- 		legend pos= north west,
-% 		xtick distance=64,
- 		legend style={font=\tiny},
- 		label style={font=\footnotesize},
- 		xlabel style={at={(0.5,0.1)}, anchor=north},
- 		legend columns=2,
- 		tick label style={font=\footnotesize},
- 		grid=major,
- 		ymin=0,
-         xmin=90,
-         xmax=560,
- 		xlabel=log2(prime),
- 		ylabel=Time (seconds)]
-		\addplot[color=red,mark=*,mark size=2pt] coordinates {
-			(127.0, 0.26)
-			(129.0, 0.38)
-			(130.0, 0.39)
-			(137.0, 0.37)
-			(140.0, 0.38)
-			(141.0, 0.39)
-			(150.0, 0.38)
-			(150.0, 0.39)
-			(152.0, 0.38)
-			(158.0, 0.37)
-			(165.0, 0.38)
-			(166.0, 0.39)
-			(171.0, 0.38)
-			(174.0, 0.38)
-			(174.0, 0.38)
-			(189.0, 0.39)
-			(190.0, 0.38)
-			(191.0, 0.36)
-			(192.0, 0.31)
-			(194.0, 0.5)
-			(196.0, 0.5)
-			(198.0, 0.51)
-			(204.37503943134692, 0.51)
-			(206.0, 0.5)
-			(212.0, 0.49)
-			(213.0, 0.49)
-			(216.0, 0.51)
-			(221.0, 0.51)
-			(222.0, 0.53)
-			(224.0, 0.5)
-			(226.0, 0.5)
-			(230.0, 0.54)
-			(235.0, 0.5)
-			(243.0, 0.5)
-			(251.0, 0.5)
-			(253.98877343250717, 0.5)
-			(255.0, 0.48)
-			(255.0, 0.52)
-			(256.0, 0.38)
-			(255.99999999966408, 0.38)
-			(256.0, 0.38)
-			(256.0, 0.37)
-			(255.9980614856364, 0.39)
-			(266.0, 0.66)
-			(285.0, 0.73)
-			(291.0, 0.68)
-			(321.0, 0.8)
-			(322.0, 0.83)
-			(336.0, 0.8)
-			(336.0, 0.86)
-			(338.0, 0.8)
-			(369.0, 0.79)
-			(379.0, 0.79)
-			(382.0, 0.92)
-			(383.0, 0.75)
-			(383.0, 0.75)
-			(383.0, 0.76)
-			(384.0, 0.64)
-			(384.0, 0.67)
-			(383.9998899269044, 0.63)
-			(383.467605550083, 0.62)
-			(389.0, 0.97)
-			(401.0, 0.97)
-			(413.0, 0.99)
-			(414.0, 0.98)
-			(416.0, 0.98)
-			(444.0, 0.96)
-			(448.0, 0.79)
-			(450.0, 1.22)
-			(452.0, 1.24)
-			(468.0, 1.23)
-			(480.0, 1.18)
-			(488.0, 1.2)
-			(489.0, 1.17)
-			(495.0, 1.17)
-			(509.97423531735535, 1.2)
-			(511.0, 1.13)
-			(511.0, 1.12)
-			(511.9891505409899, 0.99)
-			(512.0, 0.95)
-			(521.0, 1.4)
-		};
-		\addlegendentry{GMP mpn\_sec API}
-
-		\addplot[color=red,mark=o,mark size=2pt] coordinates {
-			(127.0, 0.15)
-			(129.0, 0.27)
-			(130.0, 0.28)
-			(137.0, 0.27)
-			(140.0, 0.27)
-			(141.0, 0.27)
-			(150.0, 0.3)
-			(150.0, 0.29)
-			(152.0, 0.27)
-			(158.0, 0.27)
-			(165.0, 0.27)
-			(166.0, 0.27)
-			(171.0, 0.27)
-			(174.0, 0.28)
-			(174.0, 0.27)
-			(189.0, 0.27)
-			(190.0, 0.27)
-			(191.0, 0.26)
-			(192.0, 0.24)
-			(194.0, 0.34)
-			(196.0, 0.34)
-			(198.0, 0.34)
-			(204.37503943134692, 0.34)
-			(206.0, 0.34)
-			(212.0, 0.35)
-			(213.0, 0.37)
-			(216.0, 0.35)
-			(221.0, 0.36)
-			(222.0, 0.35)
-			(224.0, 0.35)
-			(226.0, 0.35)
-			(230.0, 0.35)
-			(235.0, 0.34)
-			(243.0, 0.34)
-			(251.0, 0.35)
-			(253.98877343250717, 0.35)
-			(255.0, 0.35)
-			(255.0, 0.34)
-			(256.0, 0.34)
-			(255.99999999966408, 0.33)
-			(256.0, 0.34)
-			(256.0, 0.34)
-			(255.9980614856364, 0.34)
-			(266.0, 0.45)
-			(285.0, 0.43)
-			(291.0, 0.42)
-			(321.0, 0.54)
-			(322.0, 0.53)
-			(336.0, 0.53)
-			(336.0, 0.53)
-			(338.0, 0.54)
-			(369.0, 0.52)
-			(379.0, 0.55)
-			(382.0, 0.57)
-			(383.0, 0.5)
-			(383.0, 0.51)
-			(383.0, 0.51)
-			(384.0, 0.47)
-			(384.0, 0.48)
-			(383.9998899269044, 0.46)
-			(383.467605550083, 0.46)
-			(389.0, 0.6)
-			(401.0, 0.61)
-			(413.0, 0.62)
-			(414.0, 0.6)
-			(416.0, 0.6)
-			(444.0, 0.6)
-			(448.0, 0.52)
-			(450.0, 0.74)
-			(452.0, 0.71)
-			(468.0, 0.71)
-			(480.0, 0.71)
-			(488.0, 0.7)
-			(489.0, 0.69)
-			(495.0, 0.69)
-			(509.97423531735535, 0.7)
-			(511.0, 0.66)
-			(511.0, 0.66)
-			(511.9891505409899, 0.62)
-			(512.0, 0.62)
-			(521.0, 0.81)
-		};
-		\addlegendentry{GMP mpn API}
-
-		\addplot[color=blue,mark=triangle*,mark size=2pt] coordinates {
-			(127.0, 0.03)
-			(129.0, 0.03)
-			(130.0, 0.03)
-			(137.0, 0.03)
-			(140.0, 0.03)
-			(141.0, 0.03)
-			(150.0, 0.03)
-			(150.0, 0.03)
-			(152.0, 0.03)
-			(158.0, 0.03)
-			(165.0, 0.03)
-			(166.0, 0.03)
-			(171.0, 0.03)
-			(174.0, 0.03)
-			(174.0, 0.03)
-			(189.0, 0.04)
-			(190.0, 0.04)
-			(191.0, 0.04)
-			(192.0, 0.05)
-			(194.0, 0.04)
-			(196.0, 0.04)
-			(198.0, 0.04)
-			(204.37503943134692, 0.14)
-			(206.0, 0.04)
-			(212.0, 0.05)
-			(213.0, 0.04)
-			(216.0, 0.04)
-			(221.0, 0.05)
-			(222.0, 0.05)
-			(224.0, 0.13)
-			(226.0, 0.04)
-			(230.0, 0.05)
-			(235.0, 0.06)
-			(243.0, 0.06)
-			(251.0, 0.06)
-			(253.98877343250717, 0.12)
-			(255.0, 0.06)
-			(255.0, 0.06)
-			(256.0, 0.06)
-			(255.99999999966408, 0.11)
-			(256.0, 0.1)
-			(256.0, 0.13)
-			(255.9980614856364, 0.11)
-			(266.0, 0.06)
-			(285.0, 0.06)
-			(291.0, 0.07)
-			(321.0, 0.1)
-			(322.0, 0.07)
-			(336.0, 0.1)
-			(336.0, 0.09)
-			(338.0, 0.1)
-			(369.0, 0.13)
-			(379.0, 0.12)
-			(382.0, 0.13)
-			(383.0, 0.13)
-			(383.0, 0.13)
-			(383.0, 0.13)
-			(384.0, 0.25)
-			(384.0, 0.13)
-			(383.9998899269044, 0.23)
-			(383.467605550083, 0.23)
-			(389.0, 0.13)
-			(401.0, 0.14)
-			(413.0, 0.16)
-			(414.0, 0.15)
-			(416.0, 0.14)
-			(444.0, 0.17)
-			(448.0, 0.12)
-			(450.0, 0.13)
-			(452.0, 0.16)
-			(468.0, 0.16)
-			(480.0, 0.13)
-			(488.0, 0.19)
-			(489.0, 0.2)
-			(495.0, 0.19)
-			(511.0, 0.25)
-			(511.0, 0.25)
-			(512.0, 0.24)
-			(521.0, 0.18)
-		};
-		\addlegendentry{this work}
-
-	\end{axis}
-\end{tikzpicture}
- \begin{tikzpicture}
- \begin{axis}[
- ymin=0,
- ymax=15.5,
- xmin=90,
- xmax=560,
- ticks=none,
- height=3.4cm,
- width=\textwidth,
- ytick={0,1.67,3.33,5,...,15},
- %label style={font=\footnotesize},
- %xlabel style={at={(0.5,0.1)}, anchor=north},
- %tick label style={font=\footnotesize},
- grid=major,
- ]
- \end{axis}
- 	\begin{axis}[
-        height=3.4cm,
- 		title style={font=\small},
- 		title=32-Bit Field Arithmetic Benchmarks,
- 		width=\textwidth,
- 		legend pos= north west,
-% 		xtick distance=64,
- 		legend style={font=\tiny},
- 		label style={font=\footnotesize},
- 		xlabel style={at={(0.5,0.1)}, anchor=north},
- 		legend columns=2,
- 		tick label style={font=\footnotesize},
- 		grid=major,
- 		ymin=0,
-        ymax=15.5,
-        xmin=90,
-        xmax=560,
- 		xlabel=log2(prime),
- 		ylabel=Time (seconds)]
-		\addplot[color=red,mark=*,mark size=2pt] coordinates {
-			(127.0, 2.86)
-			(129.0, 3.38)
-			(130.0, 3.56)
-			(137.0, 3.41)
-			(140.0, 3.43)
-			(141.0, 3.43)
-			(150.0, 3.56)
-			(150.0, 3.38)
-			(152.0, 3.4)
-			(158.0, 3.4)
-			(165.0, 4.02)
-			(166.0, 4.02)
-			(171.0, 4.04)
-			(174.0, 4.03)
-			(174.0, 4.02)
-			(189.0, 4.05)
-			(190.0, 4.1)
-			(191.0, 4.03)
-			(192.0, 3.56)
-			(194.0, 4.66)
-			(196.0, 4.64)
-			(198.0, 4.72)
-			(204.37503943134692, 4.66)
-			(206.0, 4.62)
-			(212.0, 4.68)
-			(213.0, 4.68)
-			(216.0, 4.67)
-			(221.0, 4.68)
-			(222.0, 4.72)
-			(224.0, 4.13)
-			(226.0, 5.25)
-			(230.0, 5.29)
-			(235.0, 5.31)
-			(243.0, 5.29)
-			(251.0, 5.3)
-			(253.98877343250717, 5.26)
-			(255.0, 5.25)
-			(255.0, 5.27)
-			(256.0, 4.71)
-			(255.99999999966408, 4.7)
-			(256.0, 4.72)
-			(256.0, 4.77)
-			(255.9980614856364, 4.78)
-			(266.0, 6.1)
-			(285.0, 6.13)
-			(291.0, 6.94)
-			(321.0, 7.6)
-			(322.0, 7.66)
-			(336.0, 7.64)
-			(336.0, 7.58)
-			(338.0, 7.66)
-			(369.0, 8.41)
-			(379.0, 8.44)
-			(382.0, 8.41)
-			(383.0, 8.44)
-			(383.0, 8.47)
-			(383.0, 8.45)
-			(384.0, 7.62)
-			(384.0, 7.62)
-			(383.9998899269044, 7.64)
-			(383.467605550083, 7.66)
-			(389.0, 9.41)
-			(401.0, 9.35)
-			(413.0, 9.48)
-			(414.0, 9.4)
-			(416.0, 8.54)
-			(444.0, 10.31)
-			(448.0, 9.57)
-			(450.0, 11.37)
-			(452.0, 11.33)
-			(468.0, 11.37)
-			(480.0, 10.47)
-			(488.0, 12.23)
-			(489.0, 12.26)
-			(495.0, 12.2)
-			(509.97423531735535, 12.17)
-			(511.0, 12.21)
-			(511.0, 12.23)
-			(511.9891505409899, 11.26)
-			(512.0, 11.23)
-			(521.0, 13.3)
-		};
-		\addlegendentry{GMP mpn\_sec API}
-
-		\addplot[color=red,mark=o,mark size=2pt] coordinates {
-			(127.0, 3.23)
-			(129.0, 3.77)
-			(130.0, 3.79)
-			(137.0, 3.78)
-			(140.0, 3.77)
-			(141.0, 3.81)
-			(150.0, 3.79)
-			(150.0, 3.8)
-			(152.0, 3.82)
-			(158.0, 3.77)
-			(165.0, 4.45)
-			(166.0, 4.43)
-			(171.0, 4.51)
-			(174.0, 4.39)
-			(174.0, 4.4)
-			(189.0, 4.4)
-			(190.0, 4.42)
-			(191.0, 4.4)
-			(192.0, 4.23)
-			(194.0, 4.94)
-			(196.0, 4.94)
-			(198.0, 4.97)
-			(204.37503943134692, 5.03)
-			(206.0, 4.91)
-			(212.0, 4.91)
-			(213.0, 4.94)
-			(216.0, 4.92)
-			(221.0, 4.92)
-			(222.0, 4.87)
-			(224.0, 4.85)
-			(226.0, 5.65)
-			(230.0, 5.71)
-			(235.0, 5.69)
-			(243.0, 5.62)
-			(251.0, 5.65)
-			(253.98877343250717, 5.7)
-			(255.0, 5.7)
-			(255.0, 5.71)
-			(256.0, 5.49)
-			(255.99999999966408, 5.46)
-			(256.0, 5.45)
-			(256.0, 5.48)
-			(255.9980614856364, 5.46)
-			(266.0, 6.32)
-			(285.0, 6.34)
-			(291.0, 6.98)
-			(321.0, 7.66)
-			(322.0, 7.74)
-			(336.0, 7.74)
-			(336.0, 7.69)
-			(338.0, 7.67)
-			(369.0, 9.03)
-			(379.0, 9.25)
-			(382.0, 9.04)
-			(383.0, 9.11)
-			(383.0, 9.13)
-			(383.0, 9.11)
-			(384.0, 8.8)
-			(384.0, 8.82)
-			(383.9998899269044, 8.94)
-			(383.467605550083, 8.84)
-			(389.0, 9.93)
-			(401.0, 9.92)
-			(413.0, 9.93)
-			(414.0, 9.86)
-			(416.0, 9.67)
-			(444.0, 10.89)
-			(448.0, 10.51)
-			(450.0, 11.63)
-			(452.0, 11.63)
-			(468.0, 11.63)
-			(480.0, 11.33)
-			(488.0, 12.92)
-			(489.0, 12.81)
-			(495.0, 13.1)
-			(509.97423531735535, 12.9)
-			(511.0, 13.07)
-			(511.0, 12.9)
-			(511.9891505409899, 12.58)
-			(512.0, 12.55)
-			(521.0, 13.91)
-		};
-		\addlegendentry{GMP mpn API}
-
-		\addplot[color=blue,mark=triangle*,mark size=2pt] coordinates {
-			(127.0, 0.3)
-			(129.0, 0.35)
-			(130.0, 0.44)
-			(137.0, 0.48)
-			(140.0, 0.51)
-			(141.0, 0.51)
-			(150.0, 0.42)
-			(150.0, 0.49)
-			(152.0, 0.5)
-			(158.0, 0.52)
-			(165.0, 0.59)
-			(166.0, 0.61)
-			(171.0, 0.57)
-			(174.0, 0.58)
-			(174.0, 0.61)
-			(189.0, 0.7)
-			(190.0, 0.71)
-			(191.0, 0.66)
-			(192.0, 2.41)
-			(194.0, 0.75)
-			(196.0, 0.77)
-			(198.0, 0.76)
-			(206.0, 0.76)
-			(212.0, 0.86)
-			(213.0, 0.7)
-			(216.0, 0.7)
-			(221.0, 0.8)
-			(222.0, 0.87)
-			(226.0, 0.87)
-			(230.0, 0.83)
-			(235.0, 0.9)
-			(243.0, 0.86)
-			(251.0, 1.12)
-			(253.98877343250717, 3.97)
-			(255.0, 1.01)
-			(255.0, 1.43)
-			(256.0, 1.2)
-			(256.0, 1.65)
-			(266.0, 1.01)
-			(285.0, 1.13)
-			(291.0, 1.33)
-			(321.0, 1.72)
-			(322.0, 1.37)
-			(336.0, 1.67)
-			(336.0, 1.59)
-			(338.0, 1.7)
-			(369.0, 2.44)
-			(379.0, 2.47)
-			(382.0, 2.66)
-			(383.0, 2.63)
-			(383.0, 2.6)
-			(383.0, 3.58)
-			(384.0, 3.95)
-			(389.0, 2.89)
-			(401.0, 2.85)
-			(413.0, 3.53)
-			(414.0, 3.72)
-			(416.0, 2.48)
-			(444.0, 3.7)
-			(448.0, 3.18)
-			(452.0, 3.23)
-			(468.0, 3.2)
-			(480.0, 3.58)
-			(488.0, 7.99)
-			(489.0, 7.7)
-			(495.0, 6.07)
-			(511.0, 9.73)
-			(521.0, 3.9)
-		};
-		\addlegendentry{this work}
-
-	\end{axis}
-\end{tikzpicture}
-\caption{\label{graphs}Performance comparison of our generated C code vs. handwritten using \texttt{libgmp}}
-\end{figure*}
 
@@
 \texttt{nistz256} remains as an option for use in server contexts where performance is critical and where patches can be applied quickly when new bugs are found.
-The latter is not a purely theoretical concern -- Appendix \ref{bugsapp} contains an incomplete list of issues discovered in previous \texttt{nistz256} versions.
-
-The two curves thus generated with our framework for BoringSSL together account for over 99\% of ECDH connections.
-Chrome version 65 is the first release to use our P-256 code.
+The latter is not a purely theoretical concern -- the appendices of our full paper contain a sample of issues discovered in previous \texttt{nistz256} versions.
+The two curves thus generated with our framework for BoringSSL together account for over 99\% of ECDH connections initiated by Google Chrome.
+%% In fact, given that very recent versions of Microsoft Edge are based on Chromium~\cite{EdgeChromiumAnnouncement}, it's possible that connections initiated through that browser also use our generated code, although we cannot find a detailed enough source to confirm.
+%% \todo{(Adam) Is the previous sentence too handwavy? Should we just not mention Edge? Yeah, I'd skip mentioning it, if we can't track down confirmation before submitting.}
 
 % https://boringssl-review.googlesource.com/c/boringssl/+/23244
-
-% maybe write a section answering openssl mailing lists skepticism about
-% low-level compilers, "what to do about the compiler?", basically saying it is
-% critical infrastructure no matter what?
-
-\subsection{Benchmarking All the Primes}\label{instantiate}
-
-Even though the primes supported by the modular-reduction algorithms we implemented vastly outnumber those actually used for cryptography, we could not resist the temptation to do a breadth-oriented benchmark for plausible cryptographic parameters.
-The only other implementation of constant-time modular arithmetic for arbitrary modulus sizes we found is the \verb|mpn_sec| section of the GNU Multiple Precision Arithmetic Library\footnote{\url{https://gmplib.org/}}.
-We also include the non-constant-time \texttt{mpn} interface for comparison.
-The list of primes to test was generated by scraping the archives of the ECC discussion list \texttt{curves@moderncrypto.org}, finding all supported textual representations of prime numbers.
-
-One measurement in Fig.
-\ref{graphs} corresponds to 1000 sequential computations of a 256-bit Montgomery ladder, aiming to simulate cryptographic use accurately while still constraining the source of performance variation to field arithmetic.
-A single source file was used for all \verb|mpn_sec|-based implementations; the modulus was specified as a preprocessor macro to allow compiler optimizations to try their best to specialize GMP code.
-Our arithmetic functions were instantiated using standard heuristics for picking the compile-time parameters based on the shorthand textual representation of the prime, picking a champion configuration out of all cases where correctness proof succeeds, with two unsaturated and one Montgomery-based candidates.
-The 64-bit trials were run on an x86 Intel Haswell processor and 32-bit trials were run on an ARMv7-A Qualcomm Krait (LG Nexus 4).
-
-We see a significant performance advantage for our code, even compared to the GMP version that ``cheats'' by leaking secrets through timing.
-Speedups range between 1.25X and 10X.
-Appendix \ref{data} includes the full details, with Tables \ref{table64} and \ref{table32} recording all experimental data points, including for an additional comparison implementation in C++.
-In our current experiments, compilation of saturated arithmetic in Coq times out for a handful of larger primes.
-There is no inherent reason for this slowness, but our work-in-progress reimplementation of this pipeline (Section \ref{conclusion}) is designed to be executable using the Haskell extraction feature of Coq,
-aiming to sidestep such performance unpredictability conclusively.
-
-\section{Related Work}\label{related}
-
-% I think it might make sense to organize/split this section based on whether some project or plan goes up or down in terms of abstraction level.
-% First, verification projects:
-
-% I think verif25519 is the predominant prior work, and we should give adaquate exposure unless political reasons demand otherwise.
-% From the perspective of our high-level implementations, there is very little difference between the abstraction levels of verif25519 (qhasm) and HACl-star (common subset of F* with mutable state and C).
-% HACL-star specifications are much nicer, but only about halfway to ours
-% HACL-star proofs are slightly nicer, but only questionably so.
-% HACL-star covers more code, but the additional higher-level code it covers acts as its own specification during verification.
-
-% The second most interesting piece of prior work is gfverif, because it is fully automatic, but it is not published and is alpha-alpha quality according to the authors.
-% of course, it still requires writing all the code by hand
-% gröbner basis solvers are just really good for addition and multiplication, which make up the vast majority of unsaturated implementations.
-% it is unclear how it could be extended to do anything beyond unsatruated modular arithmetic
-
-% Scala LMS and other templating systems should be mentioned here
-
-% Our main comparison to them is that our code is actually a natural expression of the algorithm
-% because of this, verification is much less painful
-% because of this, genericity (to the extent supported by the algorithm) is freely available
-
-% Separately, we might think of extending our library to deal with assembly as opposde to C.
-% there are some efforts that already do pretty well on this
-
-% Vale and Jasmin deserve the current mentions
-
-% verifiedBignum, compcert, cakeml
-
-% a simple success criteria for a compiler: if you take handwritten assembly that is line-by-line expressible as compiler input and translate it to the input language, the compiler should not generate code slower than that assembly.
-This would help our P-256 code greatly.
-
-% supporting loops is not a todo-item, we already have proofs about loops and we don't want the current C extraction to deal with them anyway
-
-\subsection{Verified Elliptic-Curve Cryptography}
-
-Several past projects have produced verified ECC implementations.
-We first summarize them and then give a unified comparison against our new work.
-
-Chen et al.~\cite{verif25519} verified an assembly implementation of Curve25519, using a mix of automatic SAT solving and manual Coq proof for remaining goals.
-A later project by Bernstein and Schwabe~\cite{gfverif}, described as an ``alpha test,'' explores an alternative workflow using the Sage computer-algebra system.
-This line of work benefits from the chance for direct analysis of common C and assembly programs for unsaturated arithmetic, so that obviously established standards of performance can be maintained.
-
-Vale~\cite{vale} supports compile-time metaprogramming of assembly code, with a cleaner syntax to accomplish the same tasks done via Perl scripts in OpenSSL.
-There is a superficial similarity to the flexible code generation used in our own work.
-However, Vale and OpenSSL use comparatively shallow metaprogramming, essentially just doing macro substitution, simple compile-time offset arithmetic, and loop unrolling.
-Vale has not been used to write code parameterized on a prime modulus (and OpenSSL includes no such code).
-%Vale connects to Dafny in a proof-assistant-inspired style that keeps Vale itself out of the trusted code base.
-A verified static analysis checks that assembly code does not leak secrets, including through timing channels.
-
-HACL$^*$~\cite{hacl} is a cryptographic library implemented and verified in the F$^*$ programming language, providing all the functionality needed to run TLS 1.3 with the latest primitives.
-Primitives are implemented in the Low$^*$ imperative subset of F$^*$~\cite{kremlin}, which supports automatic semantics-preserving translation to C.
-As a result, while taking advantage of F$^*$'s high-level features for specification, HACL$^*$ performs comparably to leading C libraries.
-Additionally, abstract types are used for secret data to avoid side-channel leaks.
-
-Jasmin~\cite{jasmin} is a low-level language that wraps assembly-style straightline code with C-style control flow.
-It has a Coq-verified compiler to 64-bit x86 assembly (with other targets planned), along with support for verification of memory safety and absence of information leaks, via reductions to Dafny.
-A Dafny reduction for functional-correctness proof exists but has not yet been used in a significant case study.
-
-We compare against these past projects in a few main claims.
-
-\emph{Advantage of our work:} \textbf{first verified high-performance implementation of P-256}.
-Though this curve is the most widely used in TLS today, no prior projects had demonstrated performance-competitive versions with correctness proofs.
-A predecessor system to HACL$^*$~\cite{ECCstar} verified more curves, including P-256, but in high-level F$^*$ code, incurring performance overhead above 100X.
-
-\emph{Advantage of our work:} \textbf{code and proof reuse across algorithms and parameters}.
-A more fundamental appeal of our approach is that it finally enables the high-performance-cryptography domain to benefit from pleasant software-engineering practices of abstraction and code reuse.
-The code becomes easier to understand and maintain (and therefore to prove), and we also gain the ability to compile fast versions of new algorithm variants automatically, with correctness proofs.
-Considering the cited past work, it is perhaps worth emphasizing how big of a shift it is to do implementation and proving at a parameterized level: as far as we are aware, no past project on high-performance verified ECC has ever \emph{verified implementations of multiple elliptic curves} or \emph{verified multiple implementations of one curve, targeted at different hardware characteristics}.
-With our framework, all of those variants can be generated automatically, at low developer cost; and indeed the BoringSSL team took advantage of that flexibility.
-
-\emph{Advantage of our work:} \textbf{small trusted code base}.
-Every past project we mentioned includes either an SMT solver or a computer-algebra system in the trusted code base.
-Furthermore, each past project also trusts some program-verification tool that processes program syntax and outputs symbolic proof obligations.
-We trust only the standard Coq theorem prover, which has a proof-checking kernel significantly smaller than the dependencies just mentioned, and the (rather short, whiteboard-level) statements of the formal claims that we make (plus, for now, the C compiler; see below).
-
-\emph{Disadvantage of our work:} \textbf{stopping at C rather than assembly}.
-Several past projects apply to assembly code rather than C code.
-As a result, they can achieve higher performance and remove the C compiler from the trusted base.
-Our generated code is already rather close to assembly, differing only in the need for some register allocation and instruction scheduling.
-In fact, it seems worth pointing out that every phase of our code-generation pipeline is necessary to get code \emph{low-level enough} to be accepted as input by Jasmin or Vale.
-We have to admit to nontrivial frustration with popular C compilers in the crypto domain, keeping us from using them to bridge this gap with low performance cost.
-We wish compiler authors would follow the maxim that, if a programmer takes handwritten assembly and translates it line-by-line to the compiler's input language, the compiler should not generate code slower than the original assembly.
-Unfortunately, no popular compiler seems to obey that maxim today, so we are looking into writing our own (with correctness proof) or integrating with one of the projects mentioned above.
-
-\emph{Disadvantage of our work:} \textbf{constant-time guarantee mechanized only for straightline code}.
-Our development also includes higher-level cryptographic code that calls the primitives described here, with little innovation beyond comprehensive functional-correctness proofs.
-That code is written following ``constant-time'' coding rules and thus should be free of timing side channels barring human error.
-However, we do not yet do certified compilation or static analysis for timing leaks on the higher-level code.
-To this end, it could be advantageous to take inspiration from, or adopt, one of the tools mentioned above.
-
-\subsection{Other Related Work}
-
-Performance-oriented synthesis of domain-specific code (without proofs of correctness) has previously been done using explicit templates (e.g.
-Template Haskell~\cite{TemplateHaskell}) and more sophisticated multistage programming (e.g.
-Lightweight Modular Staging (LMS)~\cite{LMS}).
-More specialized frameworks along these lines include FFTW~\cite{FFTW} and Spiral~\cite{SPIRAL}.
-Out of these, our strategy is most similar to the one from LMS, differing mainly in the choice of using existing (proof-generating) Coq facilities for partial evaluation and rewriting rather than implementing them ourselves.
-While verified compilers (e.g., CakeML~\cite{CakeML}, CompCert~\cite{Compcert}) and translation validators~\cite{NeculaTransVal} are useful for creating soundly optimized versions of a reference program, we are not aware of any that could cope with abstraction-collapsing synthesis as done in this work or LMS.
-
-Myreen and Curello verified a general-purpose big-integer library~\cite{verifiedBignum}.
-The code uses a single uniform base system, does not include specialized modular-reduction optimizations, and does not run in constant time.
-However, their verification extends down to AMD64 assembly using verified decompilation.
-%The proof effort is roughly similar to ours (6227 lines of HOL).
-
-Verification of cryptographic protocols (e.g., FCF~\cite{FCF}, CertiCrypt~\cite{CertiCrypt}) is complementary to this work: given a good formal specification of a protocol, it can be shown separately that an implementation corresponds to the protocol (as we do for elliptic-curve primitives used in TLS) and that the protocol is secure (out of scope for this paper).
-The work by Beringer et al.~\cite{HMAC} is a good example of this pattern, composing a protocol-security proof, a correctness proof for its C-level implementation, and a correctness proof for the C compiler.
 
@@
 
-On the other hand, we do not believe that the phase distinction between arithmetic and bounds analysis is fundamental, although it is convenient.
-When viewing uniform positional representations as polynomials, this distinction corresponds to finding a bound on the degree of the polynomial using partial evaluation and finding bounds on its coefficients using certified compilation, but cryptographic implementation literature (e.g.~\cite{curve25519}) has treated degree and coefficient ranges rather similarly.
-We are aware of one example that would be simpler if the high-level algorithms operated on tuples $(l, x, h)$ s.t. $l \le x \le h$ instead of arbitrary-precision integers $x$.
-The BoringSSL P-256 implementation that was replaced with our code used prime-shape-specific modular reduction with negative $c_i$ and avoided limb underflow by adding the modulus to a field element before any addition where the input limb ranges did not already rule out underflow.
-A certified compiler that needs to preserve the exact value of each limb cannot use this strategy, but it could be encoded in the framework so far by figuring out when to add the modulus for balance in some other way in the high-level algorithm.
-
-The bounds-analysis pipeline described in this paper only works on straight-line code, greatly simplifying the implementation.
-It is geared towards being used on the bodies of inner loops where naively picking the largest available integer type for each variable would be costly, but temporaries are numerous enough that picking sizes by hand would be tedious and error-prone.
-The analysis could be generalized to loops and recursion by iterating until convergence, but it is unclear whether that is called for in any application.
-
-While we do appreciate being able to support a number of plausibly useful elliptic curves with the same (parameterized) code and proof,
-we believe that engineering benefits, especially simplicity of code and proof, make this approach worth considering even when targeting a single configuration.
-
-We find it arbitrary and unfortunate that arithmetic algorithms need to be written in continuation-passing style.
-As CPS was only necessary for using Coq's term reduction for partial evaluation, writing a certified partial evaluator would be sufficient to avoid it.
-Preliminary experiments have given promising results, but we have not finalized the interface yet.
-We also expect that replacing the controlled term reduction with a certified compilation pass would significantly improve the compilation time.
-%\todo{say something about having to remove dependent types because we don't know how to reify them?}
-
-We would like to shrink our trusted base by connecting to a certified compiler targeting assembly.
-However, existing compilers are not smart enough at mapping the data flow between instructions onto time and space available on the hardware.
-Thus another fruitful future-work area is studying those optimizations, principally combined register allocation and instruction scheduling, even independently of proof.
+Looking towards the future, we would like to extend our pipeline to bypass the C compiler and target assembly, shrinking our trusted code base.
+This would require studying combined register allocation and instruction scheduling; since our output AST is very low-level, the C compiler is doing little else for us.
 Alternatively, hand-optimized assembly code could be checked to implement the same machine-word-level computation as our generated code using an equality-checking engine like those found in modern SMT solvers (e.g., \cite{simplify}), allowing verification of optimizations that are out-of-reach for compilers, for example effective use of vector instructions.
 
+% TODO: any updates here?
 \section*{Acknowledgments}
@@
 
-\newpage
-
-\appendices
-
-\section{\label{bugsapp}Study of Bugs in Similar Crypto Code}
-\begin{figure*}
-  \begin{tabular}{llll}
-    Reference/\emph{Category}                                                           & Specification               & Implementation              & Defect       \\
- \hline \emph{Carrying}\\ \hline
- \href{https://github.com/golang/go/issues/13515}{go\#13515}               & Modular exponentiation      & \texttt{uintptr}-sized Montgomery form, Go & carry handling\\
- \href{https://tweetnacl.cr.yp.to/tweetnacl-20131229.pdf}{\cite[p.
-2]{TweetNaCl}\textsuperscript{\ref{tweetnacl-correction}}} & F25519 mul, square          & \texttt{ed25519-amd64-64-24k} & carry handling\\
- \href{https://github.com/openssl/openssl/commit/29851264f11ccc70c6c0140d7e3d8d93ef5c9b11}{openssl\#ef5c9b11} & Modular exponentiation & 64-bit Montgomery form, AMD64 & carry handling\\
- \href{https://git.openssl.org/gitweb/?p=openssl.git;a=commitdiff;h=4b8736a22e758c371bc2f8b3534dc0c274acf42c;ds=sidebyside}{openssl\#74acf42c} & Poly1305 & multiple implementations & carry handling \\
- \href{https://git.lysator.liu.se/nettle/nettle/commit/c71d2c9d20eeebb985e3872e4550137209e3ce4d}{nettle\#09e3ce4d} & secp-256r1 modular reduction & & carry handling\\
- \href{https://www.openssl.org/news/secadv/20170126.txt}{CVE-2017-3732} & $x^2 \bmod m$ & Montgomery form, AMD64 assembly & \href{https://boringssl.googlesource.com/boringssl/+/d103616db14ca9587f074efaf9f09a48b8ca80cb%5E%21/}{carry, exploitable}\\
- \href{https://rt.openssl.org/Ticket/Display.html?id=1593&user=guest&pass=guest}{openssl\#1593} & P384 modular reduction & & \href{https://eprint.iacr.org/2011/633.pdf}{carry, exploitable}\\
-    \href{https://web.archive.org/web/20160305001036/http://blog.skylable.com/2014/05/tweetnacl-carrybit-bug/}{tweetnacl-U32} & irrelevant & bit-twiddly C & assumed 32-bit \texttt{long}\\
-    \hline \emph{Canonicalization} \\ \hline
- \href{https://github.com/agl/curve25519-donna/commit/2647eeba59fb628914c79ce691df794a8edc799f}{donna\#8edc799f} & GF($2^{255}-19$) internal to wire &  32-bit pseudo-Mersenne, C & non-canonical\\
- \href{https://git.openssl.org/gitweb/?p=openssl.git;a=commitdiff;h=b62b2454fadfccaf5e055a1810d72174c2633b8f;ds=sidebyside}{openssl\#c2633b8f} & a + b mod p256 &  Montgomery form, AMD64 assembly & \href{https://mta.openssl.org/pipermail/openssl-dev/2016-August/008179.html}{non-canonical}\\
- \href{http://seb.dbzteam.org/blog/2014/04/28/tweetnacl_arithmetic_bug.html}{tweetnacl-m15} & GF($2^{255}-19$) freeze & bit-twiddly C & bounds? typo?\\
-    \hline \emph{Misc. number system} \\ \hline
- \href{https://rt.openssl.org/Ticket/Display.html?id=3607}{openssl\#3607}  & P-256 field element squaring & 64-bit Montgomery form, AMD64 & limb overflow\\
- \href{https://git.openssl.org/gitweb/?p=openssl.git;a=commitdiff;h=dc3c5067cd90f3f2159e5d53c57b92730c687d7e;ds=sidebyside}{openssl\#0c687d7e} & Poly1305 & 32-bit pseudo-Mersenne, x86 and ARM & bad truncation\\
- \href{https://github.com/openssl/openssl/commit/a7a44ba55cb4f884c6bc9ceac90072dea38e66d0}{CVE-2014-3570} & Bignum squaring & asm &  limb overflow\\
- \href{https://github.com/golang/crypto/commit/84e98f45760e87786b7f24603b8166a6fa09811d}{go\#fa09811d} & poly1305 reduction & AMD64 asm, missing subtraction of 3 & found quickly\\
- \href{https://git.openssl.org/gitweb/?p=openssl.git;a=commitdiff;h=bbe9769ba66ab2512678a87b0d9b266ba970db05;ds=sidebyside}{openssl\#a970db05} & Poly1305 &  Lazy reduction in x86 asm & lost bit 59\\
- \href{https://git.openssl.org/gitweb/?p=openssl.git;a=commitdiff;h=1ea8ae5090f557fea2e5b4d5758b10566825d74b;ds=sidebyside}{openssl\#6825d74b} & Poly1305 & AVX2 addition and reduction & bounds?\\
- \href{https://github.com/bitcoin-core/secp256k1/commit/5de4c5dffd22aa4510a5c97d0ad4a9c2eed71d85}{bitcoin\#eed71d85} & ECDSA-secp256k1 x*B & mixed addition Jacobian+Affine & missing case\\
-    \hline \emph{Elliptic Curves} \\ \hline
- \href{https://ed25519.cr.yp.to/python/ed25519.py}{ed25519.py} & Ed25519 & accepts signatures other impls reject & missing $h \bmod l$\\
- \href{http://hg.openjdk.java.net/jdk8u/jdk8u/jdk/rev/d99101781d7e}{openjdk\#01781d7e} & EC scalarmult & mixed addition Jacobian+Affine & \href{https://twitter.com/asanso/status/887691580018176000}{missing case}\\
- \href{https://blogs.adobe.com/security/2017/03/critical-vulnerability-uncovered-in-json-encryption.html}{jose-adobe} & ECDH-ES & 5 different libraries affected & not on curve\\
- \href{http://euklid.org/pdf/ECC_Invalid_Curve.pdf}{invalid-curve} & NIST ECDH & irrelevant & not on curve\\
- \href{https://github.com/google/end-to-end/issues/340}{end-to-end\#340} & Curve25519 library & twisted Edwards coordinates & $(0, 1) = \infty$\\
- \href{https://git.openssl.org/gitweb/?p=openssl.git;a=commitdiff;h=e3057a57caf4274ea1fb074518e4714059dfcabf;ds=sidebyside}{openssl\#59dfcabf} & Weier.
-affine \texttt{<->} Jacobian &  Montgomery form, AMD64 and C & $\infty$ confusion\\
-    \hline \emph{Crypto Primitives} \\ \hline
- \href{http://www.dest-unreach.org/socat/contrib/socat-secadv7.html}{socat\#7} & DH in Z*p & irrelevant & non-prime $p$\\
- \href{https://web.archive.org/web/20071010042708/http://www.imc.org/ietf-openpgp/mail-archive/msg14307.html}{CVE-2006-4339} & RSA-PKCS-1 sig. verification & irrelevant & padding check\\
- \href{https://getmonero.org/2017/05/17/disclosure-of-a-major-bug-in-cryptonote-based-currencies.html}{CryptoNote}
-    & Anti-double-spending tag &  additive curve25519 curve point & missed order$(P) \neq l$\\
-  \end{tabular}
-  \caption{Survey of bugs in algebra-based cryptography implementations}\label{bugs}
-\end{figure*}
-
-Highly optimized handwritten cryptographic arithmetic implementations have an uninspiring history of serious bugs, even when written and audited by respected experts.
-To get a sense of the details, we surveyed project bug trackers and other Internet sources, stopping after finding 26 bugs (each hyperlinked to its bug report) in implementation of nontrivial cryptography-specific optimizations.
-Fig. \ref{bugs} summarizes our findings, in terms of 5 emergent categories.
-The first three categories have to do with representing large integers using multiple machine-word-sized integers, with custom positional number systems.
-Carrying is fairly analogous to the same concept in grade-school arithmetic, and canonicalization is a part of converting back from a custom representation into a standard one.
-Elliptic curve formulas are part of high-level algebraic optimizations, above the level of operations on single large integers.
-Crypto primitives bring it all together to provide functionality like digital signatures.
-
-Here is a sampling of root causes behind these bugs.
-\begin{itemize}
-\item Mechanical errors:
-One of the two bugs uncovered in OpenSSL issue 3607 was summarized by its author as ``Got math wrong :-('', which we think refers to a pencil-and-paper execution of numerical range analysis.
-    The discussion was concluded when the patched version was found to be ``good for \textasciitilde{}6B random tests'' and the reviewer saw that ``there aren't any low-hanging bugs left.''
-In \texttt{ed25519-amd64-64-24k}, one of 16,184 repetitive (and handwritten) lines should have been \texttt{r2 += 0 + carry} instead of \texttt{r1 += 0 + carry} \cite[p. 2]{TweetNaCl}.
-\item Confusion over intermediate specifications:
-OpenSSL bug 1953 was traced back to confusion between the postconditions of exact division with remainder and an operation that produces a $q$ and $r$ s.t.
-$x = qm + r$ but does not guarantee that $r$ is the smallest possible.
-    The probability of a random test triggering this bug was bounded to $10\cdot{}2^{-29}$~\cite{OpenSSLBignumExploit}.
-\item Mathematical misconceptions:
-  The
-    \href{https://getmonero.org/2017/05/17/disclosure-of-a-major-bug-in-cryptonote-based-currencies.html}{CryptoNote
-    doublespending bug} arose from use of an algorithm on a composite-order
-    elliptic curve when it is only applicable in a prime-order group.
-\end{itemize}
-
-\section{\label{data}Full Results of Many-Primes Experiments}
-
-Tables \ref{table64} and \ref{table32} contain the full results of our performance experiments on many primes.
-Recall the basic experimental setup:
-\begin{itemize}
-\item Scrape all prime numbers mentioned in the archives of the ECC mailing list \texttt{curves@moderncrypto.org}.
-Crucially, we record not just the numeric values of the primes but also the \emph{ways in which they are expressed} in terms of additions and subtractions of powers of 2 and small multiples thereof.
-\item We run a small Python script (shorter than 300 lines) to inspect the shapes of these prime formulas, using simple heuristics to choose the parameters to our Coq library: not just a prime modulus of arithmetic but also how to divide a big integer into digits and which sequence of carry operations to perform in modular reduction.
-Actually, the script generates multiple variants, by considering 64-bit vs.
-32-bit hardware architectures and by considering the Montgomery and Solinas arithmetic strategies.
-The main operation defined in each case is a Montgomery ladder step.
-\item We run our Coq pipeline on every variant, culminating in C code pretty-printed for each one.
-\item 64-bit configurations are compiled and run on an x86 Linux desktop machine, while 32-bit configurations are compiled and run on an ARM Android mobile device.
-We save the running time of each variation.
-\item We also compile and run fixed C and C++ implementations using \texttt{libgmp}.
-\end{itemize}
-
-The three comparison implementations are:
-\begin{itemize}
-\item GMP C constant-time, the best comparison with the goals of the code we generate, since running time is required to be independent of integer inputs
-\item GMP C variable time, taking advantage of additional optimizations that leak input values through timing
-\item GMP C++, the only one of the comparison implementations that does not include manual memory management
-\end{itemize}
-
-All three comparison programs are conventional in that they are fixed C or C++ programs, where the prime modulus is set as a preprocessor macro.
-It is up to the C compiler and \texttt{libgmp} to take advantage of properties of each modulus.
-The final column in each table shows how much better our specialized generation does.
-We take the ratio of variable-time C GMP (the fastest GMP code) to whichever of our generated variants is faster.
-
-Some columns in the tables contain dashes in place of numbers of seconds needed for one trial.
-Those spots indicate configurations where our Coq compilation currently times out or exhausts available memory.
-Considering that Coq is not designed as a platform for executing an optimizing compiler, we are pleased that we get as many successful compilations as we do!
-However, we continue working on optimizations to our implementation, to push up the size of prime whose code we can compile quickly.
-The timing bottleneck is generally in reification, where repeated inefficient manipulation of terms and contexts by Ltac incurs significant overhead.
-The memory bottleneck generally shows up at \texttt{Qed}-time.
-Note also that some configurations are expected to fail to build, for instance when applying the Solinas strategy to so-called ``Montgomery-friendly'' primes like $2^{256} - 88 \cdot 2^{240} - 1$, where implementation experts would never choose Solinas.
-
-Among successful compilations, time ranges between tens of seconds and levels best run overnight.
-
-% \section{\label{weighing}Weighing Our Code Base}
-%
-% %Lacking an established method for quantifying how difficult some piece of code is to maintain, we will resort to a mix of presenting solid (but rather uninformative) numbers and providing anecdotes about our experience developing this library.
-%
-% % While we started working on the first version of this library two years ago, real progress was made rather sproadically.
-% % We implemented new features as we figured out how to do them nicely, making no attempt to work continuously.
-% % Even so, the vast majority of the work we did has been replaced with alternatives that in hindsight seem obviously superior.
-% % For these reasons, we find it rather meaningless to provide any numeric estimate of developer time -- numbers from 1 person-year to 8 person years could constructed using various thresholds for counting activity.
-%
-% % We also do not see our library as a feat in large-scale implementation or a collection of programming pearls.
-% % Out of the 38000 lines of our code relevant to this paper, around 15000 would probably be best migrated to the Coq standard library.
-% % For example, this includes numerous simple lemmas about the basic operations on natural numbers and integers along with tactic code for manipulating standard library objects.
-% % Maybe a half of these were additions were created for this project, with the rest being salvaged from other projects.
-% % Maybe a quarter of our code (in each file) is braindead boilerplate, working around deficencies in Coq's metaprogramming facilities.
-% % To maintain compatability with tactic-based code generation, we cannot use the Coq module system for parametrization and need to import definitions from other files one-by-one, specifiying their parameters again each time.
-% % Similarly, to preserve sanity while writing tactics that operate on algebraic code, we wrote a 2000-line algebraic hierarchy library that mirror's Coq's standard library in most aspects but doesn't require all uses of (for example) an operation that happens to form a ring to be marked as such ahead of time.
-% % Every definition in our basesytem library is duplicated in CPS and non-CPS form, and the two are proven equivalent -- while avoiding this isn't trivial, a certified CPS-compiler for the relecant fragment of Gallina would do the trick.
-% % Some of this is already improving: every time Coq developers fix a bug in the \texttt{nsatz} tactic, one of our elliptic curve proofs drops from several hundred lines to one line.
-% % This list could go on; measuring this library by weight is meaningless again.
-%
-% We can also give a short quantitative summary of our code base, as a proxy for work required to develop and maintain it.
-% Just the code for unsaturated arithmetic, described in Section \ref{basesystems}, requires 160 lines (each 80 characters or less) of code and proof, which can then trivially be used to generate all unsaturated reduction examples in this paper.
-% However, trying to synthesize a chained carry operation using this code would result in exponential blow-up due to loss of sharing.
-% Rewriting functions in continuation-passing style to force sharing of subexpressions (as demonstrated in Section \ref{partial}), the same library grows to around 1000 lines.
-% The extensions for saturated arithmetic add 800 lines, and all proofs and code we needed to add for Montgomery reduction total 1500 lines.
-% For comparison, the \texttt{amd64-51} implementation of X25519 contains 1900
-% lines of assembly code (with P-256 implementations being substantially longer),
-% and the translator from assembly to SMT-solver formulas used by \cite{verif25519} is 8800 lines of OCaml.
-% The elliptic-curves library is rather thin: a total of 1300 lines of code for 3 curve shapes and 8 point formats, including all equivalence proofs.
-% Certified-compiler-phase implementations are rather verbose and sometimes include more lines for parameters than actual code: the largest one is bounds inference (about 2500 lines code+proof), adding up to a total around 15,000 lines.
-% With another 15,000 lines of utility lemmas and tactics that seem reasonable candidates to move into Coq's standard library, the total development adds up to 38,000 lines.
-
-\pagebreak
-\footnotetext{\label{tweetnacl-correction}An earlier version of this paper (prepublication) incorrectly referred to this implementation as NaCl ed25519.
-In fact, the bug was discovered as the code underwent NaCl's auditing process and never made it into a release.}
-\ %
-\clearpage
-\nobalance
-
-\bottomcaption{\label{table64}Full 64-bit benchmark data.
-Our code tried both Solinas and Montgomery implementations for each prime, and we test against three GMP-based implementations: one that is constant-time (gmpsec), one that is variable time (gmpvar), and GMP's C++ API.
-Our code is constant-time, so gmpsec is the best comparison; however, even with that constraint removed from GMP and not us, we compare favorably to gmpvar.}
-
-\tablehead{%
-  \hline
-  & \multicolumn{2}{c|}{\textbf{Our Code}} & \multicolumn{3}{c|}{\textbf{GMP Code}} & \\
-  \cline{2-6}
-  \textbf{Prime} & \textbf{Sol.} & \textbf{Mont.} & \textbf{const time} & \textbf{var time} & \textbf{C++} & \textbf{Speed\allowbreak-up} \\ \hline}
-\footnotesize
-\centering
-\begin{xtabular}{|l|p{0.475cm}|p{0.75cm}@{}|p{0.7cm}@{}|p{0.6cm}@{}|p{0.6cm}@{}|p{0.75cm}@{}|@{}}
-	$2^{127}-1$ & 0.03 & 0.04 & 0.26 & 0.15 & 0.67 & 5.0 \\
-	$2^{129}-25$ & 0.03 & 0.07 & 0.38 & 0.27 & 0.8 & 9.0 \\
-	$2^{130}-5$ & 0.03 & 0.09 & 0.39 & 0.28 & 0.79 & 9.33 \\
-	$2^{137}-13$ & 0.03 & 0.08 & 0.37 & 0.27 & 0.8 & 9.0 \\
-	$2^{140}-27$ & 0.03 & 0.08 & 0.38 & 0.27 & 0.8 & 9.0 \\
-	$2^{141}-9$ & 0.03 & 0.08 & 0.39 & 0.27 & 0.83 & 9.0 \\
-	$2^{150}-3$ & 0.03 & 0.08 & 0.38 & 0.3 & 0.8 & 10.0 \\
-	$2^{150}-5$ & 0.03 & 0.08 & 0.39 & 0.29 & 0.84 & 9.67 \\
-	$2^{152}-17$ & 0.03 & 0.08 & 0.38 & 0.27 & 0.82 & 9.0 \\
-	$2^{158}-15$ & 0.03 & 0.08 & 0.37 & 0.27 & 0.76 & 9.0 \\
-	$2^{165}-25$ & 0.03 & 0.08 & 0.38 & 0.27 & 0.78 & 9.0 \\
-	$2^{166}-5$ & 0.03 & 0.08 & 0.39 & 0.27 & 0.79 & 9.0 \\
-	$2^{171}-19$ & 0.03 & 0.08 & 0.38 & 0.27 & 0.79 & 9.0 \\
-	$2^{174}-17$ & 0.03 & 0.08 & 0.38 & 0.28 & 0.78 & 9.33 \\
-	$2^{174}-3$ & 0.03 & 0.08 & 0.38 & 0.27 & 0.78 & 9.0 \\
-	$2^{189}-25$ & 0.04 & 0.08 & 0.39 & 0.27 & 0.8 & 6.75 \\
-	$2^{190}-11$ & 0.04 & 0.08 & 0.38 & 0.27 & 0.78 & 6.75 \\
-	$2^{191}-19$ & 0.04 & 0.09 & 0.36 & 0.26 & 0.78 & 6.5 \\
-	$2^{192}-2^{64}-1$ & 0.05 & 0.07 & 0.31 & 0.24 & 0.79 & 4.8 \\
-	$2^{194}-33$ & 0.04 & 0.12 & 0.5 & 0.34 & 0.93 & 8.5 \\
-	$2^{196}-15$ & 0.04 & 0.12 & 0.5 & 0.34 & 0.89 & 8.5 \\
-	$2^{198}-17$ & 0.04 & 0.12 & 0.51 & 0.34 & 0.87 & 8.5 \\
-	$2^{205}-45\cdot 2^{198}-1$ & - & 0.14 & 0.51 & 0.34 & 0.87 & 2.43 \\
-	$2^{206}-5$ & 0.04 & 0.14 & 0.5 & 0.34 & 0.84 & 8.5 \\
-	$2^{212}-29$ & 0.05 & 0.12 & 0.49 & 0.35 & 0.87 & 7.0 \\
-	$2^{213}-3$ & 0.04 & 0.13 & 0.49 & 0.37 & 0.88 & 9.25 \\
-	$2^{216}-2^{108}-1$ & 0.04 & 0.12 & 0.51 & 0.35 & 0.88 & 8.75 \\
-	$2^{221}-3$ & 0.05 & 0.15 & 0.51 & 0.36 & 0.89 & 7.2 \\
-	$2^{222}-117$ & 0.05 & 0.12 & 0.53 & 0.35 & 0.91 & 7.0 \\
-	$2^{224}-2^{96}+1$ & - & 0.13 & 0.5 & 0.35 & 0.88 & 2.69 \\
-	$2^{226}-5$ & 0.04 & 0.13 & 0.5 & 0.35 & 0.92 & 8.75 \\
-	$2^{230}-27$ & 0.05 & 0.13 & 0.54 & 0.35 & 0.91 & 7.0 \\
-	$2^{235}-15$ & 0.06 & 0.13 & 0.5 & 0.34 & 0.89 & 5.67 \\
-	$2^{243}-9$ & 0.06 & 0.13 & 0.5 & 0.34 & 0.89 & 5.67 \\
-	$2^{251}-9$ & 0.06 & 0.13 & 0.5 & 0.35 & 0.94 & 5.83 \\
-	$2^{254}-127\cdot 2^{240}-1$ & - & 0.12 & 0.5 & 0.35 & 0.92 & 2.92 \\
-	$2^{255}-19$ & 0.06 & 0.13 & 0.48 & 0.35 & 0.9 & 5.83 \\
-	$2^{255}-765$ & 0.06 & 0.13 & 0.52 & 0.34 & 0.9 & 5.67 \\
-	$2^{256}-189$ & 0.06 & 0.14 & 0.38 & 0.34 & 0.87 & 5.67 \\
-	\makecell{$2^{256}-2^{224}$ \\ \quad $+2^{192}+2^{96}-1$} & - & 0.11 & 0.38 & 0.33 & 0.84 & 3.0 \\
-	$2^{256}-2^{32}-977$ & 0.1 & 0.12 & 0.38 & 0.34 & 0.87 & 3.4 \\
-	$2^{256}-4294968273$ & 0.14 & 0.13 & 0.37 & 0.34 & 0.86 & 2.62 \\
-	$2^{256}-88\cdot 2^{240}-1$ & - & 0.11 & 0.39 & 0.34 & 0.88 & 3.09 \\
-	$2^{266}-3$ & 0.06 & 0.18 & 0.66 & 0.45 & 1.13 & 7.5 \\
-	$2^{285}-9$ & 0.06 & 0.18 & 0.73 & 0.43 & 0.97 & 7.17 \\
-	$2^{291}-19$ & 0.07 & 0.18 & 0.68 & 0.42 & 1.0 & 6.0 \\
-	$2^{321}-9$ & 0.1 & 0.26 & 0.8 & 0.54 & 1.18 & 5.4 \\
-	$2^{322}-2^{161}-1$ & 0.07 & 0.27 & 0.83 & 0.53 & 1.15 & 7.57 \\
-	$2^{336}-17$ & 0.1 & 0.27 & 0.8 & 0.53 & 1.11 & 5.3 \\
-	$2^{336}-3$ & 0.09 & 0.27 & 0.86 & 0.53 & 1.08 & 5.89 \\
-	$2^{338}-15$ & 0.1 & 0.25 & 0.8 & 0.54 & 1.06 & 5.4 \\
-	$2^{369}-25$ & 0.13 & 0.26 & 0.79 & 0.52 & 1.1 & 4.0 \\
-	$2^{379}-19$ & 0.12 & 0.26 & 0.79 & 0.55 & 1.07 & 4.58 \\
-	$2^{382}-105$ & 0.13 & 0.25 & 0.92 & 0.57 & 1.11 & 4.38 \\
-	$2^{383}-187$ & 0.13 & 0.28 & 0.75 & 0.5 & 1.05 & 3.85 \\
-	$2^{383}-31$ & 0.13 & 0.26 & 0.75 & 0.51 & 1.05 & 3.92 \\
-	$2^{383}-421$ & 0.13 & 0.25 & 0.76 & 0.51 & 1.06 & 3.92 \\
-	\makecell{$2^{384}-2^{128}$ \\ \quad $-2^{96}+2^{32}-1$} & - & 0.25 & 0.64 & 0.47 & 0.98 & 1.88 \\
-	$2^{384}-317$ & 0.13 & 0.26 & 0.67 & 0.48 & 1.0 & 3.69 \\
-	$2^{384}-5\cdot 2^{368}-1$ & - & 0.23 & 0.63 & 0.46 & 0.99 & 2.0 \\
-	$2^{384}-79\cdot 2^{376}-1$ & - & 0.23 & 0.62 & 0.46 & 0.99 & 2.0 \\
-	$2^{389}-21$ & 0.13 & - & 0.97 & 0.6 & 1.22 & 4.62 \\
-	$2^{401}-31$ & 0.14 & - & 0.97 & 0.61 & 1.17 & 4.36 \\
-	$2^{413}-21$ & 0.16 & - & 0.99 & 0.62 & 1.22 & 3.88 \\
-	$2^{414}-17$ & 0.15 & - & 0.98 & 0.6 & 1.21 & 4.0 \\
-	$2^{416}-2^{208}-1$ & 0.14 & - & 0.98 & 0.6 & 1.16 & 4.29 \\
-	$2^{444}-17$ & 0.17 & - & 0.96 & 0.6 & 1.2 & 3.53 \\
-	$2^{448}-2^{224}-1$ & 0.12 & - & 0.79 & 0.52 & 1.06 & 4.33 \\
-	$2^{450}-2^{225}-1$ & 0.13 & - & 1.22 & 0.74 & 1.34 & 5.69 \\
-	$2^{452}-3$ & 0.16 & - & 1.24 & 0.71 & 1.32 & 4.44 \\
-	$2^{468}-17$ & 0.16 & - & 1.23 & 0.71 & 1.29 & 4.44 \\
-	$2^{480}-2^{240}-1$ & 0.13 & - & 1.18 & 0.71 & 1.28 & 5.46 \\
-	$2^{488}-17$ & 0.19 & - & 1.2 & 0.7 & 1.28 & 3.68 \\
-	$2^{489}-21$ & 0.2 & - & 1.17 & 0.69 & 1.27 & 3.45 \\
-	$2^{495}-31$ & 0.19 & - & 1.17 & 0.69 & 1.3 & 3.63 \\
-	$2^{510}-290\cdot 2^{496}-1$ & - & - & 1.2 & 0.7 & 1.28 & - \\
-	$2^{511}-187$ & 0.25 & - & 1.13 & 0.66 & 1.21 & 2.64 \\
-	$2^{511}-481$ & 0.25 & - & 1.12 & 0.66 & 1.24 & 2.64 \\
-	$2^{512}-491\cdot 2^{496}-1$ & - & - & 0.99 & 0.62 & 1.15 & - \\
-	$2^{512}-569$ & 0.24 & - & 0.95 & 0.62 & 1.14 & 2.58 \\
-	$2^{521}-1$ & 0.18 & - & 1.4 & 0.81 & 1.44 & 4.5 \\
-\hline
-\end{xtabular}
-
-\clearpage
-
-\bottomcaption{\label{table32}Full 32-bit benchmark data.
-Many of the 32-bit Montgomery implementations exceeded the one-hour timeout for proofs, because 32-bit code involves approximately four times as many operations.
-The C++ GMP program was not benchmarked on 32-bit.}
-
-\tablehead{%
-  \hline
-  & \multicolumn{2}{c|}{\textbf{Our Code}} & \multicolumn{2}{c|}{\textbf{GMP Code}} & \\
-  \cline{2-5}
-  \textbf{Prime} & \textbf{Sol.} & \textbf{Mont.} & \textbf{const time} & \textbf{var time} & \textbf{Speed\allowbreak-up} \\ \hline}
-\footnotesize
-\centering
-\begin{xtabular}{|l|p{0.6cm}@{}|p{0.75cm}@{}|p{0.75cm}@{}|p{0.7cm}@{}|p{0.75cm}@{}|}
-	$2^{127}-1$ & 0.3 & 1.19 & 2.86 & 3.23 & 9.53 \\
-	$2^{129}-25$ & 0.35 & 1.7 & 3.38 & 3.77 & 9.66 \\
-	$2^{130}-5$ & 0.44 & 1.87 & 3.56 & 3.79 & 8.09 \\
-	$2^{137}-13$ & 0.48 & 2.06 & 3.41 & 3.78 & 7.1 \\
-	$2^{140}-27$ & 0.51 & 1.98 & 3.43 & 3.77 & 6.73 \\
-	$2^{141}-9$ & 0.51 & 2.0 & 3.43 & 3.81 & 6.73 \\
-	$2^{150}-3$ & 0.42 & 2.0 & 3.56 & 3.79 & 8.48 \\
-	$2^{150}-5$ & 0.49 & 1.99 & 3.38 & 3.8 & 6.9 \\
-	$2^{152}-17$ & 0.5 & 1.96 & 3.4 & 3.82 & 6.8 \\
-	$2^{158}-15$ & 0.52 & 2.04 & 3.4 & 3.77 & 6.54 \\
-	$2^{165}-25$ & 0.59 & 2.46 & 4.02 & 4.45 & 6.81 \\
-	$2^{166}-5$ & 0.61 & 2.43 & 4.02 & 4.43 & 6.59 \\
-	$2^{171}-19$ & 0.57 & 2.68 & 4.04 & 4.51 & 7.09 \\
-	$2^{174}-17$ & 0.58 & 2.63 & 4.03 & 4.39 & 6.95 \\
-	$2^{174}-3$ & 0.61 & 2.62 & 4.02 & 4.4 & 6.59 \\
-	$2^{189}-25$ & 0.7 & 2.65 & 4.05 & 4.4 & 5.79 \\
-	$2^{190}-11$ & 0.71 & 2.64 & 4.1 & 4.42 & 5.77 \\
-	$2^{191}-19$ & 0.66 & 2.69 & 4.03 & 4.4 & 6.11 \\
-	$2^{192}-2^{64}-1$ & - & 2.41 & 3.56 & 4.23 & 1.48 \\
-	$2^{194}-33$ & 0.75 & - & 4.66 & 4.94 & 6.21 \\
-	$2^{196}-15$ & 0.77 & - & 4.64 & 4.94 & 6.03 \\
-	$2^{198}-17$ & 0.76 & - & 4.72 & 4.97 & 6.21 \\
-	$2^{205}-45\cdot 2^{198}-1$ & - & - & 4.66 & 5.03 & - \\
-	$2^{206}-5$ & 0.76 & - & 4.62 & 4.91 & 6.08 \\
-	$2^{212}-29$ & 0.86 & - & 4.68 & 4.91 & 5.44 \\
-	$2^{213}-3$ & 0.7 & - & 4.68 & 4.94 & 6.69 \\
-	$2^{216}-2^{108}-1$ & 0.7 & - & 4.67 & 4.92 & 6.67 \\
-	$2^{221}-3$ & 0.8 & - & 4.68 & 4.92 & 5.85 \\
-	$2^{222}-117$ & 0.87 & - & 4.72 & 4.87 & 5.43 \\
-	$2^{224}-2^{96}+1$ & - & - & 4.13 & 4.85 & - \\
-	$2^{226}-5$ & 0.87 & - & 5.25 & 5.65 & 6.03 \\
-	$2^{230}-27$ & 0.83 & - & 5.29 & 5.71 & 6.37 \\
-	$2^{235}-15$ & 0.9 & - & 5.31 & 5.69 & 5.9 \\
-	$2^{243}-9$ & 0.86 & - & 5.29 & 5.62 & 6.15 \\
-	$2^{251}-9$ & 1.12 & - & 5.3 & 5.65 & 4.73 \\
-	$2^{254}-127\cdot 2^{240}-1$ & - & 3.97 & 5.26 & 5.7 & 1.32 \\
-	$2^{255}-19$ & 1.01 & - & 5.25 & 5.7 & 5.2 \\
-	$2^{255}-765$ & 1.43 & - & 5.27 & 5.71 & 3.69 \\
-	$2^{256}-189$ & 1.2 & - & 4.71 & 5.49 & 3.93 \\
-	$2^{256}-2^{224}+2^{192}+2^{96}-1$ & - & - & 4.7 & 5.46 & - \\
-	$2^{256}-2^{32}-977$ & 1.65 & - & 4.72 & 5.45 & 2.86 \\
-	$2^{256}-4294968273$ & - & - & 4.77 & 5.48 & - \\
-	$2^{256}-88\cdot 2^{240}-1$ & - & - & 4.78 & 5.46 & - \\
-	$2^{266}-3$ & 1.01 & - & 6.1 & 6.32 & 6.04 \\
-	$2^{285}-9$ & 1.13 & - & 6.13 & 6.34 & 5.42 \\
-	$2^{291}-19$ & 1.33 & - & 6.94 & 6.98 & 5.22 \\
-	$2^{321}-9$ & 1.72 & - & 7.6 & 7.66 & 4.42 \\
-	$2^{322}-2^{161}-1$ & 1.37 & - & 7.66 & 7.74 & 5.59 \\
-	$2^{336}-17$ & 1.67 & - & 7.64 & 7.74 & 4.57 \\
-	$2^{336}-3$ & 1.59 & - & 7.58 & 7.69 & 4.77 \\
-	$2^{338}-15$ & 1.7 & - & 7.66 & 7.67 & 4.51 \\
-	$2^{369}-25$ & 2.44 & - & 8.41 & 9.03 & 3.45 \\
-	$2^{379}-19$ & 2.47 & - & 8.44 & 9.25 & 3.42 \\
-	$2^{382}-105$ & 2.66 & - & 8.41 & 9.04 & 3.16 \\
-	$2^{383}-187$ & 2.63 & - & 8.44 & 9.11 & 3.21 \\
-	$2^{383}-31$ & 2.6 & - & 8.47 & 9.13 & 3.26 \\
-	$2^{383}-421$ & 3.58 & - & 8.45 & 9.11 & 2.36 \\
-	$2^{384}-2^{128}-2^{96}+2^{32}-1$ & - & - & 7.62 & 8.8 & - \\
-	$2^{384}-317$ & 3.95 & - & 7.62 & 8.82 & 1.93 \\
-	$2^{384}-5\cdot 2^{368}-1$ & - & - & 7.64 & 8.94 & - \\
-	$2^{384}-79\cdot 2^{376}-1$ & - & - & 7.66 & 8.84 & - \\
-	$2^{389}-21$ & 2.89 & - & 9.41 & 9.93 & 3.26 \\
-	$2^{401}-31$ & 2.85 & - & 9.35 & 9.92 & 3.28 \\
-	$2^{413}-21$ & 3.53 & - & 9.48 & 9.93 & 2.69 \\
-	$2^{414}-17$ & 3.72 & - & 9.4 & 9.86 & 2.53 \\
-	$2^{416}-2^{208}-1$ & 2.48 & - & 8.54 & 9.67 & 3.44 \\
-	$2^{444}-17$ & 3.7 & - & 10.31 & 10.89 & 2.79 \\
-	$2^{448}-2^{224}-1$ & 3.18 & - & 9.57 & 10.51 & 3.01 \\
-	$2^{450}-2^{225}-1$ & - & - & 11.37 & 11.63 & - \\
-	$2^{452}-3$ & 3.23 & - & 11.33 & 11.63 & 3.51 \\
-	$2^{468}-17$ & 3.2 & - & 11.37 & 11.63 & 3.55 \\
-	$2^{480}-2^{240}-1$ & 3.58 & - & 10.47 & 11.33 & 2.92 \\
-	$2^{488}-17$ & 7.99 & - & 12.23 & 12.92 & 1.53 \\
-	$2^{489}-21$ & 7.7 & - & 12.26 & 12.81 & 1.59 \\
-	$2^{495}-31$ & 6.07 & - & 12.2 & 13.1 & 2.01 \\
-	$2^{510}-290\cdot 2^{496}-1$ & - & - & 12.17 & 12.9 & - \\
-	$2^{511}-187$ & 9.73 & - & 12.21 & 13.07 & 1.25 \\
-	$2^{511}-481$ & - & - & 12.23 & 12.9 & - \\
-	$2^{512}-491\cdot 2^{496}-1$ & - & - & 11.26 & 12.58 & - \\
-	$2^{512}-569$ & - & - & 11.23 & 12.55 & - \\
-	$2^{521}-1$ & 3.9 & - & 13.3 & 13.91 & 3.41 \\
-\hline
-\end{xtabular}
 \end{document}

### 2021-08-15 _site/presentations/csw-2013/jgross-presentation.tex
Add back deploy

\documentclass{beamer}
\usepackage{hyperref}
\usepackage{multicol}
\title[HITs]{Computational Higher Inductive Types}
\subtitle{Computing with Custom Equalities}
\author[Jason Gross]{Jason Gross\texorpdfstring{ \\ \texttt{\href{mailto:jgross@mit.edu}{jgross@mit.edu}}}{}}
\institute{MIT CSAIL Student Workshop}

\begin{document}
\begin{frame}
  \maketitle
\end{frame}

\begin{frame}
  \tableofcontents
\end{frame}

\section{Properties of Equality}
\begin{frame}{Properties of Equality}
  \pause
  \begin{itemize}
    \item Reflexivity: $x = x$ \pause
    \item Symmetry: if $x = y$ then $y = x$ \pause
    \item Transitivity: if $x = y$ and $y = z$, then $x = z$ \pause
    \item Leibniz rule: if $x = y$, then $f(x) = f(y)$
  \end{itemize}
\end{frame}

\section{Warm Up: Linked Lists}
\begin{frame}{Warm Up: Linked Lists}
  \begin{itemize}
    \item Two constructors: \pause \texttt{nil}, or \texttt{[]}, and \texttt{cons} \pause
    \item Two accessors on non-nil lists: \pause \texttt{head} and \texttt{tail} \pause
    \item Equality is defined on an element-by-element basis
       \begin{itemize}
         \item $[] = []$
         \item $[] \ne [a,\ldots]$
         \item $[a,\ldots] \ne []$
         \item $[x_0,x_1,\ldots, x_n] = [y_0,y_1,\ldots,y_m]$ iff $[x_1,\ldots,x_n] = [y_1,\ldots,y_m]$ and $x_0 = y_0$
       \end{itemize} \pause
     \item Fairly easy to prove the properties of equality
       \begin{itemize}
         \item In Coq, Agda, and Idris, you get all of these properties for free
       \end{itemize}
   \end{itemize}
\end{frame}

\section{Example: Unordered Sets}
\begin{frame}{Example: Unordered Sets}
  \pause
  \begin{itemize}
    \item \texttt{nil}, or $\emptyset$ \pause
    \item \texttt{add} \pause
    \item \texttt{remove} \pause
    \item \texttt{contains} \pause
    \item Often implemented internally as a list or a tree \pause
    \item Equality is then implemented as ``is one a permutation of the other?'' \pause
    \item Fairly easy to prove that it's an equivalence relation \pause
    \item Leibniz rule (if $x = y$, then $f(x) = f(y)$) is harder
    \item In Haskell, Agda, Coq, and Idris, the Leibniz rule is false! \pause (or at least not internally provable) \pause
      \begin{itemize}
        \item The problem is that either you don't have private
          fields, or you can't make use of the fact that everything is
          defined in terms of your public methods.
      \end{itemize}
  \end{itemize}
\end{frame}

\subsection{Canonical Inhabitants}
\begin{frame}{Example: Unordered Sets}{Solution 1: Canonical Inhabitants}
  \begin{itemize}
    \item Give up private fields, but use element-wise equality \pause
    \item Define a type of ``sorted lists without duplication'', and call them sets \pause
    \item Now we can use element-wise equality, and get Leibniz (and other properties) for free \pause
    \item What if we don't have an ordering on the elements, only equality? \pause
    \item Is this really what we wanted?  We asked for unordered sets, and instead made sorted lists.
  \end{itemize}
\end{frame}

\subsection{Higher Inductive Types}
\begin{frame}{Example: Unordered Sets}{Solution 2: Higher Inductive Types}
  \begin{itemize}
    \item Higher Inductive Types \pause
    \item Keep the built-in equality (so we get the properties for free), but turn it into equality up to permutation \pause
    \item How do we get that it's an equivalence relation for free? \pause
      \begin{itemize}
        \item Take the reflexive symmetric transitive closure of the given relation
      \end{itemize} \pause
    \item How do we get Leibniz for free? \pause
      \begin{itemize}
        \item Require proving it each time you define a particular function
        \item To define a function that deals with unordered sets,
          you have to simultaneously prove that your function is
          invariant under permutations
      \end{itemize}
  \end{itemize}
\end{frame}

\section{Computing with Higher Inductive Types}
\begin{frame}{Computing with Higher Inductive Types}
  \begin{itemize}
    \item It seems simple enough, so what's the problem? \pause
    \item Having higher inductive types gives you functional extensionality (if $f(x) = g(x)$ for all $x$, then $f = g$), which doesn't yet have a good computational interpretation in Coq nor Agda nor Idris \pause
    \item Equality in Coq and Agda (\texttt{--without-K}) actually has a rich structure \pause
    \item If you look at proofs of equality, and equality of these proofs, and you iterate this process, you get enough math to do topology! \pause
    \item This is Homotopy Type Theory
  \end{itemize}
\end{frame}

\section{Thank you}
\begin{frame}{Thank you}
  \noindent \Huge
  \noindent \begin{center}
    \noindent Thanks! \\ $\left.\right.$ \pause \\
    \noindent Questions?
  \end{center}
\end{frame}

\section*{Parametricity}
\begin{frame}{Example: Unordered Sets}{Solution 3: Parametricity}
  \begin{itemize}
    \item Make use of the fact that private fields are private \pause
    \item Very hard to do! \pause
    \item Can probably be done by way of parametricity (aka ``theorems for free''), or a generalization of it \pause
    \item Parametricity can be given a computational interpretation,
      but it's very non-trivial to do so
  \end{itemize}
\end{frame}

\end{document}

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-color.sh

@@ -1,0 +1,10 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+CoLoR_CI_DIR=${CI_BUILD_DIR}/color
+
+# Compile CoLoR
+git_checkout ${CoLoR_CI_BRANCH} ${CoLoR_CI_GITURL} ${CoLoR_CI_DIR}
+( cd ${CoLoR_CI_DIR} && make )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-common.sh

@@ -1,0 +1,106 @@
+#!/usr/bin/env bash
+
+set -xe
+
+# default value for NJOBS
+: "${NJOBS:=1}"
+export NJOBS
+
+if [ -n "${GITLAB_CI}" ];
+then
+    export COQBIN="$PWD/_install_ci/bin"
+    export CI_BRANCH="$CI_COMMIT_REF_NAME"
+else
+    if [ -n "${TRAVIS}" ];
+    then
+        export CI_PULL_REQUEST="$TRAVIS_PULL_REQUEST"
+        export CI_BRANCH="$TRAVIS_BRANCH"
+    elif [ -n "${CIRCLECI}" ];
+    then
+        export CI_PULL_REQUEST="$CIRCLE_PR_NUMBER"
+        export CI_BRANCH="$CIRCLE_BRANCH"
+    else # assume local
+        export CI_BRANCH="$(git rev-parse --abbrev-ref HEAD)"
+    fi
+    export COQBIN="$PWD/bin"
+fi
+export PATH="$COQBIN:$PATH"
+
+# Coq's tools need an ending slash :S, we should fix them.
+export COQBIN="$COQBIN/"
+
+ls "$COQBIN"
+
+# Where we clone and build external developments
+CI_BUILD_DIR="$PWD/_build_ci"
+
+# shellcheck source=ci-basic-overlay.sh
+source "${ci_dir}/ci-basic-overlay.sh"
+for overlay in "${ci_dir}"/user-overlays/*.sh; do
+    # shellcheck source=/dev/null
+    source "${overlay}"
+done
+
+mathcomp_CI_DIR="${CI_BUILD_DIR}/math-comp"
+
+# git_checkout branch url dest will create a git repository
+# in <dest> (if it does not exist already) and checkout the
+# remote branch <branch> from <url>
+git_checkout()
+{
+  local _BRANCH=${1}
+  local _URL=${2}
+  local _DEST=${3}
+
+  # Allow an optional 4th argument for the commit
+  local _COMMIT=${4:-FETCH_HEAD}
+  local _DEPTH=()
+  if [ -z "${4}" ]; then _DEPTH=(--depth 1); fi
+
+  mkdir -p "${_DEST}"
+  ( cd "${_DEST}"                                                     && \
+    if [ ! -d .git ] ; then git clone "${_DEPTH[@]}" "${_URL}" . ; fi && \
+    echo "Checking out ${_DEST}"                                      && \
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-compcert.sh

@@ -1,0 +1,11 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+CompCert_CI_DIR=${CI_BUILD_DIR}/CompCert
+
+opam install -j "$NJOBS" -y menhir
+git_checkout ${CompCert_CI_BRANCH} ${CompCert_CI_GITURL} ${CompCert_CI_DIR}
+
+( cd ${CompCert_CI_DIR} && ./configure -ignore-coq-version x86_32-linux && make && make check-proof )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-coq-dpdgraph.sh

@@ -1,0 +1,10 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+coq_dpdgraph_CI_DIR=${CI_BUILD_DIR}/coq-dpdgraph
+
+git_checkout ${coq_dpdgraph_CI_BRANCH} ${coq_dpdgraph_CI_GITURL} ${coq_dpdgraph_CI_DIR}
+
+( cd ${coq_dpdgraph_CI_DIR} && autoconf && ./configure && make && make test-suite )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-cpdt.sh

@@ -1,0 +1,10 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+wget http://adam.chlipala.net/cpdt/cpdt.tgz
+tar xvfz cpdt.tgz
+
+( cd cpdt && make clean && make )
+

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-hott.sh

@@ -1,0 +1,10 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+HoTT_CI_DIR=${CI_BUILD_DIR}/HoTT
+
+git_checkout ${HoTT_CI_BRANCH} ${HoTT_CI_GITURL} ${HoTT_CI_DIR}
+
+( cd ${HoTT_CI_DIR} && ./autogen.sh && ./configure && make )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/ci-metacoq.sh

@@ -1,0 +1,19 @@
+#!/usr/bin/env bash
+
+ci_dir="$(dirname "$0")"
+source ${ci_dir}/ci-common.sh
+
+unicoq_CI_DIR=${CI_BUILD_DIR}/unicoq
+metacoq_CI_DIR=${CI_BUILD_DIR}/MetaCoq
+
+# Setup UniCoq
+
+git_checkout ${unicoq_CI_BRANCH} ${unicoq_CI_GITURL} ${unicoq_CI_DIR}
+
+( cd ${unicoq_CI_DIR} && coq_makefile -f Make -o Makefile && make && make install )
+
+# Setup MetaCoq
+
+git_checkout ${metacoq_CI_BRANCH} ${metacoq_CI_GITURL} ${metacoq_CI_DIR}
+
+( cd ${metacoq_CI_DIR} && coq_makefile -f _CoqProject -o Makefile && make )

### 2021-08-15 presentations/coq-workshop-2018-coq/dev/ci/user-overlays/00669-maximedenes-ssr-merge.sh

@@ -1,0 +1,4 @@
+if [ "$CI_PULL_REQUEST" = "669" ] || [ "$CI_BRANCH" = "ssr-merge" ]; then
+    mathcomp_CI_BRANCH=ssr-merge
+    mathcomp_CI_GITURL=https://github.com/maximedenes/math-comp.git
+fi

### 2021-08-15 presentations/coq-workshop-2018-coq/doc/tools/coqrst/coqdomain.py

@@ -1,0 +1,865 @@
+##########################################################################
+##         #   The Coq Proof Assistant / The Coq Development Team       ##
+##  v      #   INRIA, CNRS and contributors - Copyright 1999-2018       ##
+## <O___,, #       (see CREDITS file for the list of authors)           ##
+##   \VV/  ###############################################################
+##    //   #    This file is distributed under the terms of the         ##
+##         #     GNU Lesser General Public License Version 2.1          ##
+##         #     (see LICENSE file for the text of the license)         ##
+##########################################################################
+"""A Coq domain for Sphinx.
+
+Currently geared towards Coq's manual, rather than Coq source files, but one
+could imagine extending it.
+"""
+
+# pylint: disable=too-few-public-methods
+
+import re
+from itertools import chain
+from collections import defaultdict
+
+from docutils import nodes, utils
+from docutils.transforms import Transform
+from docutils.parsers.rst import Directive, directives
+from docutils.parsers.rst.roles import code_role #, set_classes
+from docutils.parsers.rst.directives.admonitions import BaseAdmonition
+
+from sphinx import addnodes
+from sphinx.roles import XRefRole
+from sphinx.util.nodes import set_source_info, set_role_source_info, make_refnode
+from sphinx.util.logging import getLogger
+from sphinx.directives import ObjectDescription
+from sphinx.domains import Domain, ObjType, Index
+from sphinx.domains.std import token_xrefs
+from sphinx.ext.mathbase import MathDirective, displaymath
+
+from . import coqdoc
+from .repl import ansicolors
+from .repl.coqtop import CoqTop
+from .notations.sphinx import sphinxify
+from .notations.plain import stringify_with_ellipses
+
+def parse_notation(notation, source, line, rawtext=None):
+    """Parse notation and wrap it in an inline node"""
+    node = nodes.inline(rawtext or notation, '', *sphinxify(notation), classes=['notation'])
+    node.source, node.line = source, line
+    return node
+
+def highlight_using_coqdoc(sentence):
+    """Lex sentence using coqdoc, and yield inline nodes for each token"""
+    tokens = coqdoc.lex(utils.unescape(sentence, 1))
+    for classes, value in tokens:
+        yield nodes.inline(value, value, classes=classes)
+
+def make_target(objtype, targetid):
+    """Create a target to an object of type objtype and id targetid"""
+    return "coq:{}.{}".format(objtype, targetid)
+
+class CoqObject(ObjectDescription):
[…]

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/coq-makefile/mlpack1/run.sh

@@ -1,0 +1,23 @@
+#!/usr/bin/env bash
+
+. ../template/init.sh
+
+coq_makefile -f _CoqProject -o Makefile
+cat Makefile.conf
+make
+make html mlihtml
+make install DSTROOT="$PWD/tmp"
+#make debug
+(cd `find tmp -name user-contrib` && find .) | sort > actual
+sort > desired <<EOT
+.
+./test
+./test/test.glob
+./test/test_plugin.cmi
+./test/test_plugin.cmx
+./test/test_plugin.cmxa
+./test/test_plugin.cmxs
+./test/test.v
+./test/test.vo
+EOT
+exec diff -u desired actual

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/coq-makefile/native1/run.sh

@@ -1,0 +1,33 @@
+#!/usr/bin/env bash
+
+NATIVECOMP=`grep "let no_native_compiler = false" ../../../config/coq_config.ml`||true
+if [[ `which ocamlopt` && $NATIVECOMP ]]; then
+
+. ../template/init.sh
+	
+coq_makefile -f _CoqProject -o Makefile
+cat Makefile.conf
+make
+make html mlihtml
+make install DSTROOT="$PWD/tmp"
+#make debug
+(cd `find tmp -name user-contrib` && find .) | sort > actual
+sort > desired <<EOT
+.
+./test
+./test/test.glob
+./test/test_plugin.cmi
+./test/test_plugin.cmx
+./test/test_plugin.cmxa
+./test/test_plugin.cmxs
+./test/test.v
+./test/test.vo
+./test/.coq-native
+./test/.coq-native/Ntest_test.cmi
+./test/.coq-native/Ntest_test.cmx
+./test/.coq-native/Ntest_test.cmxs
+EOT
+exec diff -u desired actual
+
+fi
+exit 0 # test skipped

### 2021-08-15 presentations/coq-workshop-2018-coq/test-suite/coq-makefile/uninstall1/run.sh

@@ -1,0 +1,18 @@
+#!/usr/bin/env bash
+
+. ../template/init.sh
+
+coq_makefile -f _CoqProject -o Makefile
+cat Makefile.conf
+make
+make html mlihtml
+make install DSTROOT="$PWD/tmp"
+make install-doc DSTROOT="$PWD/tmp"
+make uninstall DSTROOT="$PWD/tmp"
+make uninstall-doc DSTROOT="$PWD/tmp"
+#make debug
+(for d in `find tmp -name user-contrib` ; do pushd $d >/dev/null && find . && popd >/dev/null; done) | sort -u > actual
+sort -u > desired <<EOT
+.
+EOT
+exec diff -u desired actual

### 2022-01-31 papers/2021-rewriting-popl/response-point-by-point.txt
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@7118518b955b1dd8e7147ba0dfcf7e9d8b00d430 🚀

Responses to Review #81A
===========================================================================

Responses to Weaknesses
-----------------------

> - The exact rewrite strategy produced for a given rewrite system
>   is not entirely clear to me.

The rewrite strategy produced is a slight variant on running a single
bottom-up pass with eager βι-reduction:

- Any rule tagged with `do_again`, on success, triggers an extra
  bottom-up pass on the output of that rewrite rule.  This repetition
  is fueled, with the fuel currently hard-coded to the number of
  rewrite rules (which works well when the rules form a dag), but this
  could easily be a parameter of the rewriter.

> - The supported language of goals and rewrite rules is rather
>   restricted compared to the one supported by Malecha and Bengston's
>   reflexive rewriter, which can deal with quantified formulas for
>   example and more general side conditions.

In fact our rewriter could support both of these if the user is
willing to put in the work to adjust the encoding by hand.  We do not
advertise support for these because the convenience machinery, which
allows users to make use of our rewriter with nearly the same ease
that they could use `autorewrite`, does not support them.

There are in fact a couple of features of the rewriter of Malecha and
Bengston which would require fundamental design changes for us to support:

- setoid relations (we only use pointwise equality)

- supporting dependent types and computation at the type level (I'm
  actually not sure if they support this, but it seems plausible)

- matching application patterns that are not in eta-long form (such as
  rewriting `flat_map f` without having it applied to a list)

- matching application patterns where the head symbol is a unification
  variable

There are a couple more features that they support which we could
support with only relatively straightforward engineering work:

- repeatedly rewriting until there are no more matches, using fuel

- supporting non-linear patterns out-of-the-box (this would require
  some tweaking to the PHOAS types to support deciding equality of
  terms)

Finally, there are the things that we don't claim support for just
because our convenience code doesn't support them, but which are
supported in principle by the core reflective framework:

- discharging side-conditions through other reflective tactics

- supporting arbitrary type families

- support for quantified formulas

Roughly, the first one merely requires reifying the rewriting lemma by
hand, rather than using the convenience machinery; the success or
failure of rewrite replacement can then be conditioned on the result
of running another reflective procedure which decides the
side-condition.

Supporting arbitrary type families just requires extending the
datatype of non-base-types; the parts which we did not know how to
automate were (a) automatically turning the eliminator on a container
type into an eliminator in the UnderLets monad (more precisely, in the
NbE-value type augmeted with the UnderLets monad), and (b) generating
the section-retraction commutation of the container type and the expr
type (i.e., automatically generating the functions `list (expr T) →
expr (list T) → option (list (expr T))`, `expr A * expr B → expr (A *
B) → option (expr A * expr B)`, `option (expr T) → expr (option T) →
option (option (expr T))`, etc).  We'd be excited to know about a
general theory that handles this sort of thing.

Supporting quantified formulas would merely require tweaking the
current convenience code to treat `Prop`-valued things as term-like
rather than type-like, and adding a bit of pre-processing for `forall`
to turn it into a constant analogous to `ex`.  This would not be very
useful, though, unless users were assuming propositional
extensionality, because we can only rewrite with pointwise equality.
Nevertheless, we're not aware of any barriers in theory.

To sum up, as I understand it, Malecha and Bengston's rewriter is
parameterized over a user-supplied type of symbols for types and
terms, and the user needs to reify the lemmas themselves.  If rather
than relying on our plugin, our user writes the type codes by hand,
then we can quite easily support a much richer language of types.
Similarly, if the user codes the reified rewrite rule by hand rather
than relying on our preprocessor, then it's quite trivial to solve
side conditions with arbitrary other reflective tactics (though we
don't provide infrastructure for sequencing them).  So the differences
here are that (a) we provide much more convenience to the user, and
the cost of this is that we lose some of the expressivity that is
available by writing the lower-level versions by hand, and (b) their
rewriter fits into a much larger framework of reflective tactics that
they can use for discharging side-conditions, while our framework is
just the rewriter.

Finally, there are a couple of features that we support that we
believe Malecha and Bengtson do not:

- sharing preservation

- a rewriting strategy that allows fine-grained control of which
  subterms are re-walked looking for more rewriting

- automatically specializing the reflective language to the rewriting
  problem in a way that allows drastic speedups (I haven't seen this
  anywhere in the literature)

- constructing the reflective tactic and applying it with syntax
  that's approximately as simple as the built-in autorewrite and
  doesn't require that the user know about reflection

- interleaving βι reduction with rewriting (I'm actually not sure if
  they support this; the paper doesn't seem to say)

The first one, in particular, is quite essential in the Fiat-Crypto
example, as not having native support for lifting let-binders results
in asymptotically worse performance.  This is already apparent in the
LiftLetsMap / "Binders and Recursive Functions" example in the paper.
We achieve linear performance in `n×m` when you ignore the final cbv;
`setoid_rewrite` and `rewrite_strat` are both cubic.  The rewriter of
Malecha and Bengtson cannot possibly achieve better than O(n×m²)
unless it can be sublinear in the number of rewrites.  This is because
our rewriter gets away with a constant number of rewrites (four), plus
evaluating recursion principles for a total amount of work O(n×m).
But without primitive support for let-lifting, you instead need to
lift the lets by rewrite rules, and you end up needing to perform
O(n×m^2) rewrites just to lift the lets.  The analysis is thus:
running `make` simply gives us `m` nested applications of `map_dbl` to
a length `n` list.  To reduce a given call to `map_dbl`, you must
first lift all existing let-binders (there are n×k of them on the
k-innermost-call) across map_dbl, one at a time.  Then the map_dbl
adds another n let binders.  So we end up doing $\sum_{k=0}^{m} n×k$
lifts, i.e., n×m(m+1)/2 rewrites just to lift the lets.

Responses to Comments for author
--------------------------------
> L93 "apostropher mark" does that refer to l' on L97?

We mistakenly left this sentence in from a previous version of the
example.  The apostrophe mark in fact does not show up until the
example on line 737--738, where `?l` and `?u` are prefixed with `'` to
denote that they are literals.  It shows up again on line 1371 where
we prefix `n` with `'` to indicate that `List.repeat` should only be
simplified when the length of the output list is a ground term.

> L219: "before reaching normal forms".
>    By strong normalization, you can always reach the normal form
>    of an open term w.r.t. the reduction rules making up
>    definitional equality.
>    You seem to be referring to normal forms w.r.t. a larger rewrite
>    system here, please be more specific.

We had a mistaken understanding of the technical meaning of "normal
forms"; the setence should read "before reaching ground terms" or
"before reaching ground terms".

> L623: Maybe you can just say that reify and reflect are structurally recursive
>   on types and reduce on the expression?

Sure.

> L557 Let binders are not included here, unlike the grammar L359, why?

This is an oversight, they should be included, thanks.

> L762 Explains that you use a constant to represent them, it would be
>   good to give a hint earlier about this encoding trick.

We actually use this constant on the Gallina side, not on the expr
side.  They are, in fact, hard-coded into the expr type.

> L709 Again we did not see any example of this syntax. Maybe you
>   can provide an example here?

This is the apostrophe mark described first on L93.  There is an
example on L737--738 and L1371, but indeed we should provide the
example here of `∀ n1 m n2, 'n1 = 'n2 → 'n1 + m - 'n2 = m`.

> L710. If I understand correctly this means that `Expr t` contains
>     a boolean type among the base types, and certainly if-then-else and
>     constants as well. Can this be added simply by updating the denotation
>     of base types and constants (`denote_base_type` and `denoteI`)?

This is actually not required.  The boolean expression and the
if-then-else lives at the Gallina level.  Rewrite rules return an
optional expression, so, for example, te aforementioned `∀ n1 m n2,
'n1 = 'n2 → 'n1 + m - 'n2 = m` rule would be encoded in pseudo-Gallina
as something like:

```
{| pattern := '?1 + ?2 - '?3
 ; replacement := λ (n1 : nat) (m : expr nat) (n2 : nat),
       if Nat.eqb n1 n2
       then Some m
       else None |}
```

The transformation into a boolean statement happens entirely during
reification, and the reflective procedure cannot inspect the condition
in any way.  (This, too, is how general side-conditions could be
handled, by replacing the condition of the `if` statement with any
reflective procedure.)

> L762 If Coq's `let ... in ..` are not reified as `Let_In`, are they then
>     reduced during reification?

Yes, but this behavior is tweakable using a preprocessing tactic
(which can, for example, replace `let ... in ...` with `Let_In
... ...`).

>     I was a bit confused here because it is not clear in which world `Let_In`
>     is leaving, as a Coq identifier or as something used only by the reification/
>     interpretation. IIUC, you define a Coq constant
>     `Let_In: forall {A B} (a : A) (b : A -> B) := b a` that is used to
>     represent all lets in your goals and definitions to be reified?

Yes.

>     This `Let_In` application is then reified simply as an `App (Ident "Let_In") ...`
>     in the deeply embedded language?

No, there is a special constructor of `expr` that encodes this
constant.  But the fact that this is not just an `Ident` is largely a
convenience choice, and is not integral to the design.

>     Interpretation of `UnderLets` in the semantic domain will hence target that
>     constant too?

Yes.

>     Please provide a more detailed presentation of the treatments of lets, it
>     sounds unnecessarily vague as it is.

Undoubtedly some of this confusion is due to our attempt to present a
sufficiently simplified version of the rewriter so as to be
comprehensible.  It seems we simplified too far, or else
inconsistently.

> L883: I'd like to know here what `cbv` does when it encounters
>     `Let_In`, do you make it so that it does not reduce it,
>     and only recurses in the value and body of the let independently?

We pass `cbv` an explicit whitelist which does not include `Let_In`,
hence indeed recursion is performed independently in the value and
body.  The source of quadratic behavior is explained by Pierre-Marie
Pédrot in
https://github.com/coq/coq/issues/11151#issuecomment-557435656.

> L929: What exactly happens in the extracted OCaml version? You are
>   running the compiled rewriter extracted to OCaml applied to the
>   extraction of the deeply-embedded language?

That's correct.  We define a constant which is the application of the
rewriter to a particular reified constant, and then apply a
transformation which produces a string or other flat term (as the
OCaml runtime does not normalize under binders).  We then extract this
constant, compile it, and run the program.

>   Why didn't you also compare the native compilation mode for
>   the prime example? It would also avoid the problem of doubly
>   interpreting of the VM.

The reification of the prime examples is quite large, and the native
compiler is quite bad at large input terms, and spends most of its
time in compilation.  The comparison is not worth making without
having access to the fine-grained details of native timing (available
with `Set NativeCompute Timing` in 8.12 but not before).  For a point
of comparison, synthesizing the approximately 5 lines of code for
unsaturated addition on the prime 2^255-19 with 5 limbs on a 64-bit
machine takes 0.225s in the VM, 4.6s in cbv, 6.8s in lazy, and fails
with a stack overflow after running for 168s in the native compiler.
The smaller example of addition for 2^127-1 with 2 limbs on 64 bits
takes 0.056s in the VM, 0.519s in cbv, 0.57s in lazy, and the native
compiler again fails with a stack overflow after 173s.

> Typos:
>
> L883: refied

Thanks!  Now fixed.

Responses to Questions for the response period
----------------------------------------------
> - There seems to be no handling of "alien" terms (of "alien" type)
>   in the reification/interpretation setup? Is it a difficulty that you
>   side-step or can it be supported easily? Is it the purpose of the
>   extra idents parameter of the Make command? In that case, the tactic
>   would not be able to reify goals mentioning variables of alien types
>   then which cannot be mentioned as extras?

That's correct.  One of the selling points and, we believe,
innovations of this framework is that (a) reflective syntax can be
automatically specialized to a particular domain on-the-fly with minor
but non-trivial overhead, and (b) performing this specialization can
be looked at as part of a form of pre-compilation of rewrite databases
and allows significant speed-ups (about 2×, in our experience) by
reducing identifier lookup to Gallina pattern matching.

We could parameterize the entire framework on extra "alien" terms and
types.  Indeed this is the point of the `extra idents` parameter of
`Make`.  We don't currently have any examples where there are types in
the goal which cannot be added to the `extra ident` list, and hence
don't currently see a pressing need to add direct support for this.
Note also that supporting alien terms and types would would result in
significant overhead unless very carefully engineered (to take a
trivial example of this problem, we require decidable equality between
constants/identifiers, and `decide equality` is cubic, if I recall
correctly).

> - The handling of LetIn is a bit strange: we switch between
>   a grammar with a LetIn constructor to a presentation of `expr`
>   which doesn't, but looking at the code (L588 of Rewriter/Language.v),
>   `LetIn` does appear in the `expr` type which is interpreted to `dlet`,
>   a notation for a `Let_In` constant. In the same file, L774, the
>   preprocessing step of reification apparently inlines dependent lets
>   or turn them into beta-redexes (depending on the dependency),
>   so let-ins are not preserved by reification? Please clarify the situation.

Appologies for the lack of clarity.  The presentation of `expr` in the
paper should contain a `LetIn` constructor.  Indeed currently `let
... in ...` are inlined and `Let_In`s / `dlet ... in ...` are kept.
This is quite easily configurable, though, by changing the relevant
line of `reify_preprocess` in `src/Rewriter/Language/Language.v`, or
indeed just overwriting the preprocesser with a different one using
Ltac's reassignment (`::=`) feature.

> - The framework requires no properties of the rewrite system
>   except for semantics preservation, as far as I understand.
>   In particular you don't guarantee that the final terms
>   cannot be rewritten anymore by the rewrite rules, as
>   `rewrite_head` only performs one rewrite at the head.
>   Similarly, you use the `do_again` option to force the
>   rewriter/reducer to perform an additional bottom-up pass.
>   So, do I understand correctly that the NbE algorithm
>   produces normal forms w.r.t. β, but it does not necessarily
>   compute a normal form w.r.t. the rewrite rules?
>   Does it also mean that the rewriter is resistant to looping
>   rewrite rules like commutativity, e.g. `?x + ?y -> y + x`?
>   Finally, if my assumption is correct, to compare with
>   other rewriting strategies like `rewrite !` which loop
>   until no rewrite matches appear, don't you need to
>   repeatedly call your rewriter to obtain the same simplified
>   terms in general?

Yes, that is all correct.  (And commutativity rules will only loop if
you mark them with `do_again`.)  On the particular examples where we
compare, a single bottom-up pass (modulo the `do_again` rules) is
indeed sufficient to fully normalize the term.  For goals like
quantifier-lifting or reassociation, where a single bottom-up pass is
not sufficient, you would indeed need to repeatedly call the rewriter
(guarded by some form of progress check).  This is not too hard to
implement, but we have not implemented it.

Responses to Review #81B
===========================================================================

Responses to Comments for author
-------------------
> The paper presents a Coq library for building verified partial evaluators and shows that it performs better than primitive Coq reduction on examples such as FIAT crypto. The library is carefully designed to preserve subterm sharing.

> Though I understand the scheme at a high-level, I find it hard to understand the research contributions and how all the pieces in the paper fit together. NBE has been implemented in Coq (Boespflug et al.) and in other proof assistants, so is the contribution here to make it usable from a tactic in a way it can be proof-checked?

> Most of the techniques used in the paper are already well-known (pattern matching using decision trees, Nbe, PHOAS), the paper has put them together to make them perform better in a verified way, however, it is not evaluated well.There is only a single macro-benchmark, leaving open its applicability to other domains. The tweaks for discharging side-conditions in 4.3 and 4.4 also seems specific to that single domain.

> There are other details that the paper glosses over, e.g. around line 618, what is the correctness theorem for pattern matching and how does it show up in the proof of this lemma? There are several key places in the paper where the details are left to the Appendix, which is surprising given that the paper does not use its full budget.

> Overall I think the paper could do a better job at explaining the main contributions and the technique. It is not quite convincing in its current form.

TODO: I'm not sure what to say, I basically agree with this reviewer.
The new contributions, as I see them, are (a) actually assembling all
the pieces, and (b) the approach of creating on-the-fly a datatype of
constants as a sort of early compilation phase.  The aforementioned
paper of Boespflug et al. ("Efficient normalization by evaluation") in
fact discusses that representing constructors of the term language as
constructors of the metalanguage is one of the best optimizations they
found, but seems to go on to misinterpret their own performance
results.  Section 3.2 says "Replacing the constructor names with
integers rather than strings to avoid string comparison cost does
spare some computation, but it is better to avoid the \emph{Const}
constructor altogether. Rather than representing a datatype as an
in-memory tree, with \emph{App} constructors at branch nodes and
\emph{Const} constructors at the leaves, each in its own memory cell,
it is much more memory efficient to add all data constructors found in
the syntax as additional constructors to the metalanguage
interpretation, effectively flattening the representation in memory."
However, as I understand it, this memory optimization is already
effectively present in the evalapply benchmark, which in most cases
shows not nearly as significant a gain as the constructors benchmark.
Instead, as suggested by the performance numbers on page 14 of the
Aehlig et al. paper, I expect that most of this gain actually comes
from not using strings as constant names.  Alas, Boespflug et
al. neither include benchmarking results for replacing strings with
integers without going all the way to constructors, and apparently nor
did they include this optimization when benchmarking the strategy of
Aehlig et al.  While they correctly describe how making constructors
of the object language into constructors of the metalanguage breaks
modularity, their proposal to recover modularity is about the memory
layout, not about the encoding of constructor identity.  Furthermore,
the proposal of using integers works fine for untyped NbE, but incurs
significant overhead (proportional to the size of the term times the
number of constants) when defining typed NbE.  (Whether rewriting is
better performed with typed or untyped NbE is a different question,
not addressed in our paper.)  Our contribution (b) from the beginning
of this paragraph is to provide a way to have constructors in the
object language be represented as constructors in the metalanguage, in
a typed NbE setting, without breaking modularity nor incurring extra
work on the user, for a minor but non-trivial up-front time cost.

> Some other comments:
>
> - Please explain the notation [| E |] used on line 618 (or better show the Coq statement of the lemma?).

The notation `[| E |]` stands for `DenoteE E`, with `DenoteE` defined
on lines 582--583.  Sorry for not making this clear.

> - The paper could also elaborate/simplify the steps in 2.1 with the help of a small running example.

We agree that a small running example would help elaborate the steps
in 2.1, but we disagree that this would be a simplification.  In
particular, the framework is sufficiently intricate that even an
example with no lemmas to rewrite involves emitting nearly a hundred
definitions.  Nevertheless, we will consider including it.  Here is
such an example:

We use the example of subsection 1.1, where we write:
\begin{minted}[fontsize=\small]{coq}
Make rewriter := Rewriter For (zero_plus, plus_zero, times_zero, times_one, eval_map,
  eval_fold_left, do_again eval_length, do_again eval_combine,
  eval_rect nat, eval_rect list, eval_rect prod) (with delta) (with extra idents (seq)).
\end{minted}

\begin{enumerate}
\item
  The given lemma statements are scraped for which named functions and types the rewriter package will support.
  In this case, the lemma statements are \mintinline{coq}{zero_plus}, \mintinline{coq}{plus_zero}, etc.; the types are \mintinline{coq}{nat}, \mintinline{coq}{list}, and \mintinline{coq}{prod}; and the named functions are \mintinline{coq}{Nat.add}, \mintinline{coq}{Nat.mul}, \mintinline{coq}{List.map}, \mintinline{coq}{List.fold_left}, \mintinline{coq}{List.length}, \mintinline{coq}{List.combine}, \mintinline{coq}{nat_rect}, \mintinline{coq}{list_rect}, \mintinline{coq}{prod_rect}, and the extra identifier given \mintinline{coq}{List.seq}.
\item
  Inductive types enumerating all available primitive types and functions are emitted.
  In this example, the inductive types are approximately
\begin{minted}{coq}
Inductive base := nat | list (_ : base).
Inductive ident := Nat_add | Nat_mul | List_map | List_fold_left | List_length | List_combine | nat_rect | list_rect | prod_rect | List_seq.
\end{minted}
  In fact this list is not accurate, because some types are forced to be included (\mintinline{coq}{nat} and \mintinline{coq}{bool}) while others have hard-coded support separate from the list of base types (\mintinline{coq}{list}, \mintinline{coq}{option}, \mintinline{coq}{unit}), and the list of identifiers is pre-filled with some hard-coded support necessary for making the rewriter work at all.
  The true inductive types generated are:
\begin{minted}{coq}
Inductive base : Set :=  base_nat : base | base_bool : base.
Inductive ident : type (base.type base) → Type :=
| ident_seq : ident (type.base (base.type.type_base base_nat) -> type.base (base.type.type_base base_nat) -> type.base (base.type.list (base.type.type_base base_nat)))
| ident_combine : ∀ t t0 : base.type base, ident (type.base (base.type.list t) -> type.base (base.type.list t0) -> type.base (base.type.list (t * t0)))
| ident_length : ∀ t : base.type base, ident (type.base (base.type.list t) -> type.base (base.type.type_base base_nat))
| ident_fold_left : ∀ t t0 : base.type base, ident ((type.base t -> type.base t0 -> type.base t) -> type.base (base.type.list t0) -> type.base t -> type.base t)
| ident_map : ∀ t t0 : base.type base, ident ((type.base t -> type.base t0) -> type.base (base.type.list t) -> type.base (base.type.list t0))
| ident_succ : ident (type.base (base.type.type_base base_nat) -> type.base (base.type.type_base base_nat))
| ident_mul : ident (type.base (base.type.type_base base_nat) -> type.base (base.type.type_base base_nat) -> type.base (base.type.type_base base_nat))
| ident_O : ident (type.base (base.type.type_base base_nat))
| ident_add : ident (type.base (base.type.type_base base_nat) -> type.base (base.type.type_base base_nat) -> type.base (base.type.type_base base_nat))
| ident_literal : ∀ t : base, match t with | base_nat => nat | base_bool => bool end → ident (type.base (base.type.type_base t))
| ident_nil : ∀ t : base.type base, ident (type.base (base.type.list t))
| ident_cons : ∀ t : base.type base, ident (type.base t -> type.base (base.type.list t) -> type.base (base.type.list t))
| ident_Some : ∀ t : base.type base, ident (type.base t -> type.base (base.type.option t))
| ident_None : ∀ t : base.type base, ident (type.base (base.type.option t))
| ident_pair : ∀ t t0 : base.type base, ident (type.base t -> type.base t0 -> type.base (t * t0))
| ident_tt : ident (type.base ())
| ident_prod_rect_nodep : ∀ t t0 t1 : base.type base, ident ((type.base t -> type.base t0 -> type.base t1) -> type.base (t * t0) -> type.base t1)
| ident_bool_rect : ∀ t : base.type base, ident ((type.base () -> type.base t) -> (type.base () -> type.base t) -> type.base (base.type.type_base base_bool) -> type.base t)
| ident_list_case : ∀ t t0 : base.type base, ident ((type.base () -> type.base t0) -> (type.base t -> type.base (base.type.list t) -> type.base t0) -> type.base (base.type.list t) -> type.base t0)
| ident_option_rect : ∀ t t0 : base.type base, ident ((type.base t -> type.base t0) -> (type.base () -> type.base t0) -> type.base (base.type.option t) -> type.base t0)
| ident_nat_rect : ∀ t : base.type base, ident ((type.base () -> type.base t) -> (type.base (base.type.type_base base_nat) -> type.base t -> type.base t) -> type.base (base.type.type_base base_nat) -> type.base t)
| ident_eager_nat_rect : ∀ t : base.type base, ident ((type.base () -> type.base t) -> (type.base (base.type.type_base base_nat) -> type.base t -> type.base t) -> type.base (base.type.type_base base_nat) -> type.base t)
| ident_nat_rect_arrow : ∀ t t0 : base.type base, ident ((type.base t -> type.base t0) -> (type.base (base.type.type_base base_nat) -> (type.base t -> type.base t0) -> type.base t -> type.base t0) -> type.base (base.type.type_base base_nat) -> type.base t -> type.base t0)
| ident_eager_nat_rect_arrow : ∀ t t0 : base.type base, ident ((type.base t -> type.base t0) -> (type.base (base.type.type_base base_nat) -> (type.base t -> type.base t0) -> type.base t -> type.base t0) -> type.base (base.type.type_base base_nat) -> type.base t -> type.base t0)
| ident_list_rect : ∀ t t0 : base.type base, ident ((type.base () -> type.base t0) -> (type.base t -> type.base (base.type.list t) -> type.base t0 -> type.base t0) -> type.base (base.type.list t) -> type.base t0)
| ident_eager_list_rect : ∀ t t0 : base.type base, ident ((type.base () -> type.base t0) -> (type.base t -> type.base (base.type.list t) -> type.base t0 -> type.base t0) -> type.base (base.type.list t) -> type.base t0)
| ident_list_rect_arrow : ∀ t t0 t1 : base.type base, ident ((type.base t0 -> type.base t1) -> (type.base t -> type.base (base.type.list t) -> (type.base t0 -> type.base t1) -> type.base t0 -> type.base t1) -> type.base (base.type.list t) -> type.base t0 -> type.base t1)
| ident_eager_list_rect_arrow : ∀ t t0 t1 : base.type base, ident ((type.base t0 -> type.base t1) -> (type.base t -> type.base (base.type.list t) -> (type.base t0 -> type.base t1) -> type.base t0 -> type.base t1) -> type.base (base.type.list t) -> type.base t0 -> type.base t1)
| ident_List_nth_default : ∀ t : base.type base, ident (type.base t -> type.base (base.type.list t) -> type.base (base.type.type_base base_nat) -> type.base t)
| ident_eager_List_nth_default : ∀ t : base.type base, ident (type.base t -> type.base (base.type.list t) -> type.base (base.type.type_base base_nat) -> type.base t)
.
\end{minted}
  We believe that this is long enough as to be confusing rather than illuminating.
\item
  Tactics generate all of the necessary definitions and prove all of the necessary lemmas for dealing with this particular set of inductive codes.
  Definitions include operations like Boolean equality on type codes and lemmas like ``all representable primitive types have decidable equality.''
  There are actually 69 definitions and lemmas generated during this step, and it's needlessly confusing to list all of them.
  They can be found by inspecting the records \mintinline{coq}{Rewriter.Language.IdentifiersBasicLibrary.Compilers.Basic.GoalType.package} and \mintinline{coq}{Rewriter.Language.IdentifiersLibrary.Compilers.pattern.ident.GoalType.package}.
  Note that the first of these has a couple of fields which are themselves records, and the definition of these records must also be inspected for the full picture.
\item
  The statements of rewrite rules are reified and soundness and syntactic-well-formedness lemmas are proven about each of them.
  Each instance of the former involves wrapping the user-provided proof with the right adapter to apply to the reified version.
  It is not clear that examples would help here.
\item
  The definitions needed to perform reification and rewriting and the lemmas needed to prove correctness are assembled into a single package that can be passed by name to the rewriting tactic.
  This step is semantically trivial but extremely verbose.
\end{enumerate}

Overall, it's not clear what this reviewer was hoping to get from a
worked out example.

> - The micro-benchmark 4(a) does not include the reification time, which seems a little strange given that reification is very much part of the technique.

The micro-benchmark 4(a) includes lines both with and without
reification.  The former is labeled "Our approach including
reification, cbv, etc." while the latter is labeled "Our approach
(only rewriting)".  The lines are quite close, and both are
significantly better than other approaches.  Moreover, as explained in
the benchmark of 4(c) (the only benchmark where the time of just
rewriting diverges significantly from the full time), more than 99% of
the overhead is due to *interpretation* in Coq's cbv tactic.

In all microbenchmarks, we have constructed the input so that it is
trivial to reify, and the overhead is negligable when there is any
real rewriting work to be done.  Reification time in the
macrobenchmark is significant due to the fact that we reify modulo δ
without sharing of subterms, but this is an artifact of the way that
we've integrated the rewriter into Fiat-Crypto.

Finally, as we explain at the end of the first paragraph of 5.1.1 on
lines 785--788, the particular method of reification is not integral
to the framework, and we have prior work demonstrating how to achieve
extremely fast reification.

> - Does the technique allow for compilation of P-384 in Fiat Crypto? The introduction mentions it, but the experiments never come back to it.

Yes.  P-384 is a prime close to 2^384, and figure 4(d) shows that our
benchmarks extend well past 2^400.  In fact, though, as described in
section 5.2, "in the course of running this experiment, we found a way
to improve the old approach for a fairer comparison."  This
improvement allows the old approach to compile P-384 (32-bit) in about
55 seconds.  We achieve about 16 seconds if you extract and compile
our code, or about 117 seconds if you run our code in Coq's VM.

Responses to Review #81C
===========================================================================

> That said, I doubt that even an expert could reproduce what they have
> done, based on this paper.  There's an overview in 2.1, aspects of
> which are fleshed out later, but a great deal is unspecified.  In a
> paper of this length that's probably inevitable.  But it's still sad.

This is true.  We have another 28 pages of content merely describing
the trickiest implementation and design choices, and the document
describing the full design of the core of the framework (not including
the code that makes the rewriter convenient to use, i.e., only
including the rewriter itself and the correctness theorem used in step
(3) of the second half of our nine-step approach in section 2.1) is a
bit over 40 pages.  Note that this document does not describe any of
the proof techniques, only the definitions and proof statements.
Documenting the full rewriter is undoubtedly a much larger effort, and
would not fit in a paper such as this.

> It's a description of a substantial and sophisticated engineering
> achivement, not a crisp, pearly idea.

Indeed, and this is unfortunate.  If there is a crisp, pearly idea to
be found, beyond assembling preexisting techniques, it is that in the
tradeoff between slow generic procedures and fast domain-specific
code, we can have our pie and eat it too by specializing the generic
reflective method to our specific domain and partially evaluating it
in a pre-compilation phase.  However, this paper is not about this
pearl, but about the larger system in which it is applied.

Responses to Comments for author
-------------------
Some small notes

> * Fiat Crypto.  Line 897 says you replicate Erbsen; and yet on line 55
>   you say you can do something they can't.  Which is right?  If they
>   can't do P-384, how is it done in deployed systems?  Do the deployed
>   systems use verified code?  If not, how does the verified code
>   compared with the (presumably hand-written) deployed code?  Please
>   be clearer about all this.

We say on lines 910--911, "[a]ctually, in the course of running this
experiment, we found a way to improve the old approach for a fairer
comparison."  This modification of their approach allows synthesis of
P-384.

Perhaps the following modifications would make the text clearer?  To
the first sentence ("[...] replicating the generation of
performance-competitive finite-field-arithmetic code for all popular
elliptic curves by Erbsen et al. [2019]"), there is an implicit "as
well as generating code for a number of curves they were unable to
generate code for".  The second sentence, "In all cases, we generate
essentially the same code as they did," should instead read "In all
cases where they succeeded in generating code, [...]".

The point of this comparison is that we are generating code from the
same starting point and with the same ending point, but we can handle
more and larger primes than could be handled in Erbsen et al. [2019].

> * Line 93.  Special marker.  Is this part of Coq already, or are you
>   inventing it?
>
> * Line 96 "Further marker".  Same question.

Both of these are just the identity function under a different name.  That is, we write
```
Definition literal {T} (v : T) := v.
Notation "' x" := (literal x).
```
for the apostrophe marker and we write
```
Definition eagerly {T} (v : T) := v.
```
for the further marker.
```

> * Line 106. I'm already.  I'm not a Coq user so I don't know
>   what plugins do.  Is do_again defined by you or built-in?  "request
>   that some rules trigger a bottom-up pass.."  Who implements that
>   request.  It's all very baffling.

Appologies for the confusion.

A Coq plugin allows us to extend not just the tactic language but the
vernacular language with new commands; we use it to create a single
command that the user can type which will define the needed
definitions and inductive types.

When we speak of "requests" and "triggers", we are speaking of
annotations which the user gives and which our rewrite procedure
inspects to determine what to do.  As a very simple example, you could
write a tactic which takes in a pair of a boolean and a lemma, and
rewrites with the lemma once when the boolean is false, but rewrites
repeatedly when the boolean is true.  This procedure could be written
in Ltac as
```
Ltac my_procedure argument :=
  lazymatch argument with
  | (true, ?lemma) => repeat rewrite lemma
  | (false, ?lemma) => rewrite lemma
  end.
```
You could say that passing the true boolean is a request to the
procedure which triggers repeated rewriting, and this is in fact
almost exactly how our rewriter functions.

To answer the last part of this question in full detail, we define
`do_again` as another identity function:
```
Notation do_again := (@id (@snd bool Prop (pair true _))) (only parsing).
```
(Note the use of `pair true`, which is us annotating the lemma with a
boolean indicating that rewriting should be repeated.  We perform the
annotation as an argument to the identity function merely for
convenience of recognition by our tactic.)

> * Line 226 little examples like this, illustrating what the perf
>   challenges are, are really helpful.  More please.

Glad to hear this!  We like small examples like this too, but they
tend to be notoriously hard to come by, especially in Coq and other
proof assistants where there is a great deal going on under the hood,
and it's not always clear where the bottlenecks are.  We'd be
interested in knowing if there are specific other parts of the paper
that could benefit from examples like these.

### 2022-01-31 papers/2022-rewriting-popl/response.txt
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@7118518b955b1dd8e7147ba0dfcf7e9d8b00d430 🚀

Thanks for the reviews, and we'll address the seemingly largest concerns before returning to answer additional questions.

The biggest concern seems to be about generality of the approach.
There is enough going on in the Fiat Cryptography example (almost 200 distinct rewrite rules covering quite a range of functional-programming patterns) that quantifying it as just "one compiler" may be misleading.
Still, we acknowledge that, beyond the microbenchmarks, we've only assessed this approach as an improvement to Fiat Cryptography.
However, our tool is a fully free-standing Coq plugin, and making use of it requires only a couple lines of code in the simplest cases (see Appendix C in the supplement for details).
We're pretty confident that we can get any Coq expert started using it on their own intuitive examples in 10 minutes; a sanity check of this kind could even be reasonable as a required revision.

We expect that our approach would apply smoothly to any compiler which can be phrased in terms of rewrite rules.
Our particular tool is subject to some minor engineering limitations, such as requiring that the source, target, and all intermediate languages be simply typed, and some less-minor limitations, such as, at present, requiring that the source and target language be total and that the semantics be expressed in terms of a denotation function.
The reviewers are probably right that our approach applies more naturally to domain-specific languages/compilers (e.g., we don't save much effort when it comes to program-analysis phases), but we are interested to discuss with authors of the cited compiler-verification projects about opportunities to streamline with our tooling.

Additionally, two reviewers asked about the rewriting strategy we implement and how this strategy is controlled by the user.

Our rewriting strategy, described in detail in section 3.3, is normalization-by-evaluation-order rewriting.
This is closest to a bottom-up strategy, except that it has much better behavior on lambda-abstractions, achieving asymptotically better performance than either top-down or bottom-up rewriting in some cases.
We now plan to add a paragraph calling out this strategy more explicitly, emphasizing that we believe it's the correct one for rewriting higher-order terms.
Rewrite rules are then performed in-order; more granular phasing of rewrite rules can be performed by sequencing invocations of the rewriter.
As described in section 2, users can specify that certain rewrite rules should trigger additional rewriting in their output (with `do_again`), allowing for a limited form of top-down rewriting.

As an example, consider rewriting with the rule `map f xs = list_rect _ [] (fun x _ map_tl => f x :: map_tl) xs` together with the rule that says to evaluate list recursion, `list_rect`, on concrete lists.
If we want to rewrite in an expression such as `map succ (map succ [0; 1; 2; 3])`, we must mark the `map` rule as `do_again` to achieve complete rewriting in a single pass.

# More responses past 500 words (other important high-level points)

Reviewer B asked:
> What about conditional rewriting rules in the style of OCaml's "when" or Agda's "with abstractions"?

The side conditions described in section 4.3 support the full scope of OCaml's "when" and Agda's "with abstractions".
The only restriction is that conditions can only be applied to fully evaluated terms (we cannot, for example, say that `x + y` should be rewritten into `2 * x` when `x == y` if `x` and `y` are not known to be fully evaluated).
Users specify side conditions for our engine by implication, so OCaml's `| pat when cond -> body` becomes `cond = true -> pat = body`, and Agda's `f args with term` followed by `f pat1 | pat2 = body` becomes, roughly `forall args, term = pat2 -> pat1 = body`.

Reviewer C asked:
> A claimed advantage over Aehlig et al. is that the trusted computing base is not grown.
> However, Aehlig et al. claim (page 2 of their paper) that their approach does not grow the trusted base either, "if the compiler used at runtime is the same as the one compiling the theorem prover".
> Could you respond to this claim?

As we understand it, Aehlig et al. formalize a semantics of ML, code up their rewriting engine relative to that semantics of ML, and then use a (trusted, unverified) compiler for ML to compile their rewrite engine and rewrite rules.
In section 3 (Model and Verification), they say "we over-approximate the operational semantics of ML."
This model that is an over-approximation of ML is an extension of the TCB, unless the ML compiler is proven with respect to this model.
(Granted, using a different compiler for rewriting and for the proof assistant would grow the TCB much more.)

Reviewer D asked:
> Could you explain how the compiler works (and is used) when it is extracted to OCaml code?
> In this scenario, what source language and target language does it use?

Fundamentally, the extracted compiler consumes and produces ASTs of the kind introduced in this paper, for a subset of Gallina.
Building a full real-world compiler requires statically composing the extracted heart of the compiler with two additional pieces: a way to input or generate deeply-embedded syntax, and a way to print or consume the resulting deeply-embedded output.
In Fiat Cryptography, we reify once and for all a template library of arithmetic routines, extracting the deeply embedded syntax trees in addition to the compiler.
We additionally have (mostly unverified) backends that pretty-print the compiled AST to strings, which we also extract to OCaml.
This results in a binary that can parse elliptic curve parameters from the command line, apply pre-reified ASTs to these parameters, run the rewriter to produce compiled code, and pretty-print the resulting ASTs to stdout.
Hence the source language of the compiler is Gallina (though in some sense the source language of our extracted binary can be said to be a tiny DSL of elliptic curve parameters), and the target language of the compiler is ASTs, which are then processed to emit C, Java, Rust, or Go code.
In theory, we could, for example, write a translator from our ASTs to CompCert C and then compose our pipeline with the CompCert compiler, which would allow us to have extracted binaries which produce object files from elliptic curve parameters.

# Answering remaining questions per review

## Review A

> line 232: This sentence seems to be missing a word.

Perhaps the subject change in "That compiler execution ran inside of Coq, but even more pragmatic is to *extract* the compiler as a standalone program in OCaml or Haskell." is confusing, and we should have instead written "That compiler execution ran inside of Coq, but an even more pragmatic approach is to *extract* the compiler as a standalone program in OCaml or Haskell."

> line 749: You refer to the developers of Fiat Crypto as 'we'; this is inconsistent with other places in the paper.

Oops, thanks.
I think for the camera-ready submission we will probably be changing all instances of "The Fiat Cryptography authors" to "we".

> line 798: Please define IR (I suppose it is 'intermediate representation')

Yes, that's correct.  We'll fix this.

## Review B

> Section 3.3: the motivating example could be treated by first beta-reducing to $z + 0$, then rewriting as usual.
> I thought you wanted to demonstrate rewriting under lambdas?
> In which case, $\lambda x. x + 0$ would be a better example.

In fact we achieve something much better than just rewriting under lambdas: we are able to interleave reduction and rewriting in a way that only a single pass over the term is required for full reduction, in most cases, even when rewrite rules expose new beta redexes and even when beta reduction exposes new rewriting opportunities.
A more complete example might have been $\texttt{map }(\lambda x. x + 1)\ [0, x, y]$, where rewriting exposes the $\beta$-redexes $[(\lambda x. x + 1)\ 0,\ (\lambda x. x + 1)\ 1,\ (\lambda x. x + 1)\ 2]$, $\beta$-reduction exposes rewriting opportunities in $[0 + 1, x + 1, y + 1]$, and finally rewriting reduces this to $[1, x + 1, y + 1]$.
We were concerned that this might be too complicated or nuanced of an example just for orienting the reader to NbE.

> Section 4.2: yes, subterm sharing is crucial, but sometimes terms must be unshared so that rewriting / reduction can proceed.
> (That's the optimal reduction dilemma.)
> Strategies such as "let-reduce when the bound term is a constant" are simplistic.
> There are many other heuristics in the literature on compiling functional languages, e.g. Appel's book "Compiling with continuations", but maybe what is needed is ways for users to control "let" reductions?

That is a fair point, which we have indeed avoided confronting because we are exploring a rather different compilation style than is most standard for functional languages, thanks to the opportunity to reduce away all higher-order features at compile time.
It is still true that any pattern that can be expressed as a rewrite rule with Boolean side conditions can be used for inlining "let"-binders, and we also support some additional heuristics.

> Section 4.4 and section 5.1.4: it is really unclear how bit-width of parameters and intermediate results is being tracked.
> I would have expected types "bits(N)" to be used for this purpose, with specific passes to infer appropriate bit widths, but this is not what happens, apparently.

Bitwidths are tracked using the arguments to the $\text{clip}_{l,u}$ functions in section 4.4.
There is a specific pass, outside the rewriting engine, to infer these bitwidths via abstract interpretation (though, as mentioned in 5.1.6, having abstract interpretation outside the rewriting engine results in quartic performance where we ought to be able to achieve quadratic performance).
We could have used "bits(N)" types, as long as we hard-coded the permissible values of N ahead of time and defined separate versions of each function for each value of N, because our engine doesn't yet support dependent types.
However, we saw no benefit to using bits(N) types over the more general, less-dependently typed solution of using `clip` functions.

## Review C

> I was quite confused by the third paragraph of the intro.
> The paragraph is explaining the scalability challenges of _proof-producing_ compilers.
> But the next paragraph goes on to say that this paper is about a _proven-correct_ compiler, which doesn't have these challenges.
> So I didn't understand why that third paragraph is necessary.
> The reader has already been primed to expect proven-correct compilers from the very first line of the introduction, so can't proof-producing compilers be deferred to Related Work?

We believe this paragraph is important for motivating our work.
Proof-producing is almost always simpler, easier to manage, easier to debug, and easier to maintain than proven-correct.
Hickey and Nogin's [2006] work, when cast in the frame of modern proof assistants such as Coq, Agda, or Isabelle/HOL, is about proof-producing compilers.
We originally tried to have Fiat Cryptography be a proof-producing compiler (see Appendix D).
Scalability challenges of proof-producing compilers are what motivated us to develop a toolkit for building proven-correct compilers in the first place, and performance concerns are essentially the only reason we see to maintain a proven-correct compiler for Fiat Cryptography rather than a proof-producing one.
Since performance concerns are so front-and-center in motivating our particular tool, we believe this paragraph deserves its place in the introduction.

> I was a bit surprised that Hickey and Nogin's [2006] work on compilers built using rewrite rules is criticised for its rewriting engine being _unproven_.
> As I understand Hickey and Nogin's work, that's kinda the point -- once the rewrites are proven, it doesn't matter which order they're applied in, so the rewriting engine can be unproven (and hence easier to change, etc).
> But your point about them not supporting side-conditions seems believable.

As we understand it, Hickey and Nogin's claim is not that the rewriting engine can be unproven, but that it can be divided into unproven heuristic parts (determining the order of rule application, performing pattern matching) and a trusted Logical Framework for scope preservation, binder tracking, etc.
Insofar as the Logical Framework maintains any kind of proof object (whether abstract as in Isabelle/HOL or concrete as in Coq), such a rewriting engine is inherently proof-producing rather than proven-correct, and hence subject to hard-to-avoid asymptotic issues in the presence of deeply nested binders.
Indeed, in our framework, order selection and term decomposition for pattern matching are unproven; we only need to prove the part of the framework that composes the rewrite rules; tracks scope preservation, binders, and variable instantiation; and performs reduction.

> Line 312: I'm not sure what "nearly simply typed" means.

"Nearly simply typed" means that we support prenex type-polymorphism in constants so long as the types are always concretely instantiated whenever the constant actually appears in an expression.

## Review D

> My description is probably inaccurate, reflecting my limited understanding of the details.

Your description is accurate.

> The presentation is [...] not self-contained enough, not gentle enough.
> The technical presentation is incomplete: reification is not described at all, I think; the use of decision trees is only briefly described; the treatment of `LetIn` constructs in the NbE algorithm is barely described; etc.

We tried to strike a balance between being overly technical and inaccessible to all but the most expert readers on the one hand, and sweeping so many details under the rug that readers would come away with an inaccurate picture on the other.
To give a sense of scale, we have another thirty pages of written content on the various design decisions, tradeoffs, and lessons of the core of the framework, and approximately another sixty pages sketching out the design of the rewriting engine in full detail.
We aimed to give an overview of the novel features of the rewriting engine, describe the lessons we learned, and present the engine in enough detail that a motivated reader could reconstruct most of the key components.
By and large, we chose to omit descriptions of techniques where there is already a rich existing literature (our engine is compatible with any method of reification, and the particular method we chose is unimportant), where we make no innovations over the existing literature (our implemention of decision trees is a transcription to Coq of the algorithm described in Maranget's "Compiling Pattern Matching to Good Decision Trees"), or where the treatment is straightforward and without pitfalls (the `LetIn` constructs follow the obvious combination of application and abstraction in the NbE algorithm of `reduce(let v := e_1 in e_2) = let x := reduce(e_1) in reduce([x/v]e_2)`).
In Appendix E, we give a detailed map between the description in the paper and the code in the code supplement, which we hope will fill in the gaps which we had no space to cover in the paper.

> lines 29-31, "most desired iteration on the compiler can be achieved through iteration on the rewrite rules".
> The authors seem to suggest that all changes to the compiler can take the form of new rewrite rules, or modifications to existing rewrite rules.
> However, much of the "foresight" that a compiler writer must display resides in the designing of appropriate intermediate languages, separated by conceptual gaps that must be just large enough to be interesting and small enough to be tractable; and in the ordering of these intermediate languages.
> One might argue that, once the design of these ILs has been frozen, some optimisations may become difficult to express, as they might require extending the definition of one or more ILs, or (worse) they might require revisiting the order in which certain transformations take place.

As we discuss in 5.1.2, 5.1.3, and especially 5.1.4, design of the appropriate intermediate languages falls within the scope of adding or tweaking rewrite rules in our framework.
The plugin code will automatically generate identifiers for any definitions mentioned in any rewrite rule.
Users are free to change which definitions they use for a shallowly-embedded IR; no freezing of design is necessary.

> line 33, "Proof-producing compilers usually operate on the functional languages of the proof assistants that they are written in".
> Perhaps I am missing something, but I don't see why that is or must be the case.
> In fact, there is probably a spectrum of "proof-producing compilers", where the "proof" takes the form of a witness that can be checked by a certified verifier; the more complex the verifier, the closer we are to a "certified" compiler, and the simpler the verifier, the closer we are to a "certifying" compiler.
> When the authors write "Proof-producing compilers usually operate [...]", which end of the spectrum do they have in mind?
> (Perhaps the extreme where there is no user-defined verifier and the witness must be a term of type Prop?)

Yes, we are considering the end where the type of a witness has type Prop.
Propositions are about terms in the functional language of the proof assistant.
While it's possible to add extra indirection (as is done, for example, in Iris' proof mode), the simplest solution is to just operate directly on such terms.

> line 60, "Their execution timed out": the execution of what?

Execution of the old Fiat Cryptography framework / code generator.
(We let it run for over 300 hours; it didn't finish.)

> line 90, "preserving subterm sharing".
> It is not clear what kind of sharing the authors have in mind.
> Sharing of program terms, or sharing of proof terms?
> Why are there important opportunities for sharing?
> Why cannot R_tac preserve sharing?

Sharing of program terms.
R_tac does not have native support for subterm sharing, and without the let-lifting optimization described in 4.2, rewriting requires asymptotically more term traversals.

> line 135, "we expect it to be performance-prohibitive to perform bookkeeping operations on first-order-encoded terms".
> Why is that?
> Is it because the authors assume that de Bruijn indices are encoded as unary numbers?
> Or would the cost be prohibitive even with an efficient representation of de Bruijn indices (either as natural integers in binary notation or as machine integers)?

Merely having an efficient representations of de Bruijn indices is insufficient; subsitution is a very significant cost.
In Figure 3(c), the gap between the blue line (our method with the final conversion from a deeply-embedded AST to a shallowly-embedded term) and the green line (just the rewriting and reduction portion of our method) is entirely due to an inefficiency in Coq's handling of de Bruijn indices, which are nevertheless encoded as machine integers.
More details are available in [Coq Issue #11151](https://github.com/coq/coq/issues/11151#issuecomment-557435656), where the foremost expert on Coq performance says:

> My understanding so far is that the cbv / lazy reduction machine are completely broken performance-wise on open terms.
> The explicit substitutions for the CBV machine for instance seem to be enforcing a call-by-name behaviour for lifts, as every time we access a variable we have to crawl the substitution and recompute the lifted result.
>
> Compare this to the VM which uses an accumulator, where closures are represented by mere arrays.
> I think that the NbE approach is inherently more efficient because object-language values are indeed represented by meta-language values.
>
> If anything, this is probably a theoretical issue that requires more pondering.

The fix for this [required using "a variant of skewed lists enriched over a monoid" to track substitutions](https://github.com/coq/coq/pull/13537).
We are happy to defer issues involving efficient handling of binders to the Coq kernel.

One point that this reviewer might be missing is that we are dealing with enormous terms, and that many of the scaling challenges that arise in Fiat Cryptography are not present in any other project we've yet encountered.

> line 142, "Fiat Cryptography's arithmetic routines rely on significant sharing".
> It is still not clear what is meant.
> If sharing matters, why not explicitly represent it via an object-level `let` construct?
> Does the difficulty stem from the fact that a shallow embedding is used, therefore it is difficult to control whether and when Coq decides to unfold its meta-level `let` construct?

We do explicitly represent sharing via an explicit object-level `let` construct.
The issue is that this construct blocks reduction unless either the code is rewritten in CPS or the reduction engine has native support for lifting `let`-binders, and that attempting to support `let`-lifting via rewrite rules results in poor asymptotic performance.
Again, the primary point here is about performance and scaling challenges.

> line 178, I suppose I could have understood this earlier, but it is becoming clear only now that the paper is about partial evaluation, and about a form of supercompilation.

Yes, perhaps we should have brought in the role of partial evaluation earlier.

> line 203, the role of `ident.eagerly` is unclear.
> Is it an identity function?
> I assume that it serves as a marker.
> It would be good to point precisely to the place in the paper where the recognition of this marker is explained.
> Why is `ident.eagerly` *not* used in the lemmas `eval_combine` and `eval_length`?

Yes, it is an identity function that serves as a marker.
Recognition of this marker is hard-coded into the reification of replacement patterns of rewrite rules into NbE-typed values, where these identifiers are replaced by custom definitions which await concrete terms to recurse on.
Appendices E.4.5 and E.2 describe some of the code associated to this feature.
We decided that this detail was too technical and required too much contextual explanation to include in this paper.
The `ident.eagerly` marker could be used in `eval_length`; there is no reason not to do so, and we would then not need to mark `eval_length` with `do_again`.
However, while we could mark the `list_rect` in `eval_combine` with `ident.eagerly`, we would still need `do_again` to handle `list_case`, because we did not build in special support for "eager" evaluation of `list_case`.
(Such support is unnecessary, because the two rewrite rules for the `cons` and `nil` cases for `list_case` are complete with only a single pass of rewriting, while no rewrite rule reified from a Gallina expression can express complete reduction of `list_rect` under a single pass of rewriting, or even under a fixed number of rewriting passes.)

> lines 220-225, most of the details mentioned in this paragraph remain mysterious.
> Why would "extra bottom-up passes" be needed?
> (Doesn't rewriting produce a normal form?)
> What is a "permitted" identifier?
> What is `seq`? etc.

We have apparently not made it clear that our rewriter performs only a *single* NbE-ordered pass over the term.
Much like NbE is capable of fully reducing a term to normal form in a single pass, we've found that nearly all of our rewriting and compilation can be done in just a single pass over the AST.
For the sake of convenience, we allow limited localized exceptions to this single-pass rule, primarily useful for cases where the RHS of one rewrite rule is expressed in terms of identifiers that allow immediate additional rewriting.
Multiple passes are supported by wrapping the rewriter in a fueled loop at the Gallina level, but neither Fiat Cryptography nor our microbenchmarks required such looping.

Our plugin emits on-the-fly an inductive type of identifiers which can be used in the AST.
Any identifiers not in this inductive type cannot appear in goals to be reified.
By default, this list is composed of the identifiers found in any rewrite rule mentioned.
The `with extra idents` construct allows extending this list so that more goals can be reified.
The `seq` identifier is `Coq.Lists.List.seq`, which shows up in the definition of `prefixSums` but not in any rewrite rule.
It might seem like such identifiers are useless, because they will not be mentioned in any rewrite rule, but the `with delta` generates rules even for identifiers which show up only in the `with extra idents` list, and in fact in this example the rule that says to reduce `seq` to normal form only when its applied to fully concrete (ground / closed) arguments is adequate.

> line 290, what is "the former"?

Soundness.
"The latter" would be "syntactic-well-formedness", both from the preceding sentence "The statements of rewrite rules are reified and soundness and syntactic-well-formedness lemmas are proven about each of them."

> line 292, "Automating this step allows rewrite rules to be proven in terms of their shallow embedding".
> I don't know what this means.

In engines like R_tac, users must provide proofs about the reified, syntactic form of rewrite rules.
This is a significant burden for users unfamiliar with the ins and outs of R_tac.
In our engine, users never need to see the underlying reflective engine, nor perform any reasoning about it, thanks to automation.

More concretely, a user merely needs to prove a lemma like `forall n, n + 0 = n`, not `forall n, Expr.Interp (expr.App (expr.App (expr.Ident expr.Z_add) n) (expr.Ident (expr.Literal 0))) = Expr.Interp n` nor `forall n, Expr.Wf n -> Expr.Wf (expr.App (expr.App (expr.Ident expr.Z_add) n) (expr.Ident (expr.Literal 0)))`.

> line 312, "limited support".
> Limited in what ways?

We support only prenex type-polymorphism in constants, and furthermore the types must always be concretely instantiated whenever the constant actually appears in an expression.
For example, we support `List.append`, which is polymorphic over the type of the list, but we only support it on concretely typed lists.
That is, we can use `List.append` to concatenate two lists of `nat`s, but we can't talk about lists whose types are unknown or variable.

> line 325, "we instead build our rewriter in Coq’s logic".
> Not sure what this means.
> Is your rewrite a program expressed in Gallina?
> Is it something else?

Yes, our rewriter is a Gallina program.
We are contrasting this with Aehlig et al.'s rewriter, which is an OCaml program.

> line 330, you adopt Maranget's approach, because you fear that a simpler approach would involve "duplicate work".
> Have you compared the two approaches in practice?

We're about 8x faster than the simpler approach.
(We didn't include any numbers or plots because we've only done a very rough comparison, and didn't consider it worth it to do extensive performance-testing here.)
This 8x can be broken down into a 4x factor and a 2x factor:
Using decision trees gives us a 4x speed-up over the naive approach, and furthermore allows us to partially evaluate the rewriting engine itself to specialize it to the particular rewriting rules.
(This partial evaluation is not easily possible without the decision tree; we needed to carefully craft our engine to be written in CPS subject to a few other constraints to allow such partial evaluation to be useful, and enabling partial evaluation on the naive approach would require an entirely different set of constraints on how we wrote the engine.
Furthermore, partial evaluation done naively on the simpler approach would result in a roughly quadratic code-size blowup with no gain: it would duplicate the list of rewrite rules for every identifier in the language without actually specializing the logic to the particular rewrite rules.)
Actually performing the early partial evaluation, so that the decision tree does not need to be evaluated when the rewriting engine is run, buys us another 2x speedup.

> line 335, "There are three steps to turn a set of rewrite rules into a functional program that takes in an expression and reduces according to the rules."
> Is the goal even well-defined?
> What if the set of rewrite rules is not confluent or not terminating?

Here is another place where we would have benefited from being clearer that the goal is a *single* NbE-ordered pass over the term, which is nearly sufficient for the vast majority of our use cases.
Handling of non-confluent rewrite rules is determined first by NbE-order term traversal, and subsequently by applying all rules with the same head identifier in order and taking the first one that matches.
Any looping must be fueled, and non-terminating rules would run repeatedly until fuel runs out.

> line 379, what is "Coq's logic"?
> What is "Coq's normal partial evaluation"?

"Coq's logic" is Gallina.
"Coq's normal partial evaluation" is `cbv` / `lazy` / `cbn` with a whitelist or blacklist.

> line 440, is this an amendment to the code in Figure 1?
> It would be clearer to have just one version of the code.

Yes.
Thanks for the suggestion.

> line 482, I don't understand the implied opposition between "closed programs" and "thousands of lines of low-level code".
> One aspect is whether the program has free variables; another aspect is the size of the program.

In our experience, most routines scale super-linearly in the number of binders.
Hence a rewriting engine or reduction procedure that scales linearly in the size of the output term, in the absence of binders, might scale quadratically or even cubically in the number of binders in the output term.
This behavior will be invisible when the engine is only tested on closed programs and on open programs that produce only a couple lines of code.

> line 496, "it is difficult to integrate arrays soundly in a logic".
> This is a naïve question, but could Coq's primitive arrays be used here?

Coq's primitive arrays have no recursion scheme and, as far as we're aware, have not been used in any major development.
(Additionally, Coq had no primitive arrays when we completed work on this project.)
We would be interested in seeing a performance analysis of various ways of doing binder bookkeeping that included primitive arrays.

> line 504, "Here is the actual inductive definition".
> Could this definition come earlier?
> The definition of reify & reflect would be clearer if it did.

It could, though we were wary of overwhelming the reader with large amounts of code.

> line 517, "A good example of encoding adequacy".
> I believe I know what "adequacy" means in this context, but I don't think every reader will.

Noted, thanks.

> line 577, "we needed to convince Coq that the function terminates".
> Coq is convinced already, if it accepts your definition.
> The phrasing is somewhat odd.
>
> line 581, "the Coq kernel is ready to run our Rewrite procedure during checking."
> During checking of what?
> If you mean that the Coq type-checker performs reduction during type-checking, that is a well-known fact which perhaps does not need to be recalled here.

Indeed, the phrasing here is odd, and we'll reword it.

> line 584, "These strategies have".
> Not clear which strategies are meant here.

The reduction strategies built in to Coq's kernel which are referred to at the end of the previous sentence, "[...] it is important to note how the Coq kernel builds in reduction strategies."
In particular, we are referring to `lazy`, `vm_compute`, and `native_compute`.

> line 590, "“known-good” goals".
> I have no idea what this means.

We mean goals that line up with the sorts of goals on which Coq's built-in reduction strategies have been tuned.
Granted, this knowledge is probably only known to expert Coq users, and perhaps we should rephrase.

> line 616, "Then we modified reduction to inline let binders instead of preserving them, at which point the reduction job terminated with an out-of-memory error [...]".
> I don't really get the point.
> Even if the compiler had been able to terminate, the resulting "optimized" code would be practically unusable, as it would perform many redundant computations.
> Right?

That's right.
This is yet another argument that subterm sharing is crucial, which is what we're trying to argue in this section.

> the "NbE strategy that outputs UnderLets telescopes" in Section 4.2 is never fully defined, as far as I can tell.

The UnderLets type is defined on lines 639--641.
This type is a monad, and it is straightforward to sprinkle the monadic `bind` and `return` operators in the NbE code given from lines 546--567 in such a way that the code typechecks.
There is very little of note in the interplay of this monad with NbE, for all that it seems quite essential to the performance of our rewriting engine, and we felt that including this code in our initial presentation of NbE would overwhelm the reader with too many unfamiliar moving parts.

> line 650, why is `with_lets` needed?
> Is it used to implement the heuristics discussed on line 658?

No, it's unrelated to the heuristic on line 658.
Strictly speaking, `with_lets` is not needed.
However, including `with_lets` allows us to avoid monadic binds in a number of places, which simplifies some of the code.
Perhaps we should have dropped `with_lets` entirely from our presentation, since even our additional thirty pages of written content on the various design decisions, tradeoffs, and lessons does not delve into the tradeoffs of having `with_lets` vs avoiding it.
However, we felt that it was better to present exactly the code that we actually use, in this instance.

> I believe the authors could perhaps better explain why the rewriting rule at line 706 is preferable (somehow more tractable) than the rewriting rule at line 696.
> At first sight, they seem similar.
> I guess the point is that the variable `n` is unknown (i.e., *not* a compile-time constant) whereas the variable `u` is known (a compile-time constant), so the second rule falls within the scope of the technology described in Section 4.3, whereas the first rule does not.

Yes, this is correct.

> The `clip` trick seems to be a nice way of attaching statically computable information to variables.
> It may be worth discussing how general this trick is (can it be used to carry arbitrary information, not just an interval?) and whether there might be other ways of achieving the same result (could the information be stored in some sort of environment?).

Thanks!
Yes, it can carry arbitrary information.
However, we've neither needed nor evaluated it on any other sort of information, and expect that the reader can think up uses just as well as we can.
Rewrite rules don't have access to any sort of environment, though I suppose that section 5.1.6 can be seen as suggesting that we provide rewrite rules with a sort of environment of abstract information.

> lines 721 and 725, "Recall from section 2 [...]".
> This was never really explained.
> All of Section 4.5 is relatively obscure.

Explaining the mechanism by which the rewriter makes use of these markers is far beyond the scope of this paper, and including all such details of the rewriting engine would probably result in upwards of a hundred pages of content.
We were not sure how to make this section less obscure; these features implement relatively straightforward capabilities with highly technical code, and it's hard to say much more than "our framework supports this feature" without ending up exposing the reader to a great deal of detailed implementation minutiae.

> line 736, "when converting the interpretation of the reified term with the prereified term".
> I don't know what this means.

As per your description of our engine, the user gives a program in Gallina and asks for the optimized version.
This optmized version is equal to the original term.
Since our rewriting engine operates on deeply-embedded syntax, we must equate the user-given program in Gallina with the interpretation of its own reification.
This equality is proven by reflexivity, as the two terms are convertible, and we are referring to the user-given Gallina program as "the prereified term".

> line 755, I don't understand the meaning of "iterating on the Fiat Cryptography compiler".

Fiat Cryptography development is ongoing.
Porting the core of the compiler to this framework did not end development.
Some of the development since then has taken the form of tweaking the set of rewrite rules and rewriter passes, and we are referring to this development as "iterating on the Fiat Cryptography compiler".

> Section 5.1.2 seems relatively uninteresting.

Perhaps, though there's a sense in which section 5.1.2 might be more properly labeled "New Constructs in the Frontend and Intermediate Representations", and this easy ability to change the IRs seems to be something you argued is important above, when discussing lines 29-31.

> line 808, "designating Z.combine_at_bitwidth as an identifier that should be inlined rather than let-bound".
> I am not sure at which point it should be inlined away.
> Certainly not too soon, otherwise the rewrite rules mentioned at lines 805-806 would never fire.

Inlining occurs after all applicable rewrite rules have triggered at a location, and rewrite rules are allowed to let-bind subterms.
However, you may be misinterpreting what we mean by "inlined", since inlining would not prevent the rewrite rules mentioned at lines 805-806 from firing.
We mean that whenever the rewriter would emit a term such as `let x := Z.combine_at_bitwidth y in z`, it should instead emit `[Z.combine_at_bitwidth y/x]z`.

> line 834-847, I think I understand this paragraph (more or less), but I do not see how it explains the quartic behavior discussed in the previous paragraph.

The particular algorithm that we use contains quartically many variable assignments, but only quadratically many of them are distinct.
However, this constant folding combined with limited common subexpression elimination (realzing that you don't need to have a variable holding `x + 0` if you already have a variable `x`) can only be realized after bounds analysis.
The paragraph on lines 834-847 is a simplified example aimed at making concrete how the fusion can impact the number of operations used, rather than being precisely accurate about the source of quartic behavior.

> line 852, the forward pointer to section 5.2.4 is somewhat unfortunate; the text cannot be fully understood without looking ahead, it seems.
> We do not know what the parameters `n` and `m` represents.

Indeed, we should rearrange the text so that we're not referring to variables before we display the code.

> On line 853, it is not clear what "cubic" means (cubic in what parameter?).

Cubic in the same variable that the first half of the sentence expressed linear performance in terms of, `n · m`:
"We achieve linear performance in n · m when ignoring the final `cbv`, while `setoid_rewrite` and `rewrite_strat` are both cubic"

> On line 858, what is `make`?
> What is `map_dbl`? The whole paragraph (848-861) is extremely technical.
> It does not clearly explain why it is the fundamentally difficult to perform let-lifting in an efficient way.

We will move this paragraph after Figure 5, where `make` is defined.
Hopefully this will make the explanation clear, since all we are doing is arguing that a particular reduction strategy (performing all rewriting and reduction with the exception of let-lifing, followed by lifting the let binders, and repeating this sequence of two steps) has a particular asymptotic performance.

> It also does not explain its opening sentence, "Another essential pass to fuse with rewriting and partial evaluation is let-lifting".
> Why is it essential?
> Is it more efficient if fused than if performed separately?
> Why is that?

Setting aside the missing clarity of the proof of asymptotic behavior, which is hampered by the fact that this paragraph comes before Figure 5, we clearly state that performing let-lifting via rewrite rules results in $\mathcal O(n\cdot m^2)$ let-liftings, while we are able to achieve performance of $\mathcal O(n\cdot m)$ by fusing the passes.
The analysis is identical if let-lifting is performed in any other way, so long as it alternates with reduction rather than being fused with it.

> In Figure 2, is `m` in fact a known constant?
> Otherwise I am not sure what the ellipsis means.

For any particular data-point, both `m` and `n` are known constants.
However, we do evaluate the code for multiple values of `m`.

> line 958 (and elsewhere), "continuation-passing style [...] allows standard VM-based reduction to achieve good performance".
> Is it folklore knowledge why writing in CPS style helps?
> It might be worth briefly recalling why this is so.

This is alluded to on lines 619--636.
In the expression `map (λ x. y + x) (let z := e in [0; 1; 2; z; z + 1])`, the `map` will not reduce unless let-binders are inlined.
However, in the equivalent CPS'd expression `(λ K. map_cps (λ x K. K (y + x)) (λ K. let z := e in K [0; 1; 2; z; z + 1]) K) (λ x. x)` reduces to `let z := e in [y; y + 1; y + 2; y + z; y + (z + 1)]` as desired.
I'm not sure if there's any folklore citation to be had, and I'm also not sure how to express the general principle here.

### 2022-05-03 papers/source/rewriting/arxiv/rewriting.tex
Add more final versions of 2022 ITP papers

--- 2020-07-30 papers/source/rewriting/rewriting.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
+\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
 \RequirePackage{luatex85}
 \PassOptionsToPackage{dvipsnames}{xcolor}
-\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2021}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-%\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-\usepackage{wrapfig}
+\RequirePackage{amssymb}
+% order matters, see https://tex.stackexchange.com/a/83051/2066
+\RequirePackage{varioref}
+\PassOptionsToPackage{nameinlink}{cleverref}%[2012/02/15]% v0.18.4;
+\documentclass[a4paper,USenglish,cleveref,autoref,thm-restate]{lipics-v2021}
+\usepackage{hyperref}
+%This is a template for producing LIPIcs articles.
+%See lipics-v2021-authors-guidelines.pdf for further information.
+%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
+%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
+%for section-numbered lemmas etc., use "numberwithinsect"
+%for enabling cleveref support, use "cleveref"
+%for enabling autoref support, use "autoref"
+%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
+%for enabling thm-restate support, use "thm-restate"
+%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
+%for producing a PDF according the PDF/A standard, add "pdfa"
+
+%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
+%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
+
+%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
+
+\bibliographystyle{plainurl}% the mandatory bibstyle
+
+%\usepackage{wrapfig}% forbidden in ITP
 \setlength{\intextsep}{0pt} % don't add space above nor below
-\usepackage{subfig}
-\newcommand{\subfigureautorefname}{\figureautorefname}
-%% http://ctan.org/pkg/subcaption
-
-\usepackage{hyperref}
-\usepackage{minted}
+%\usepackage{subfig}
+%\newcommand{\subfigureautorefname}{\figureautorefname}
+
+\nolinenumbers% needef for arXiv
+\usepackage[frozencache]{minted}
 \usemintedstyle{bw}
-\usepackage{amsthm}
-%\usepackage{amssymb}
+\usepackage{amsmath}
+%\usepackage{amsthm}
+\usepackage{amssymb}
+\usepackage{comment}
 \usepackage[all]{xy}
@@
 \usepackage{apptools}
+\usepackage{placeins}
 \usepackage{adjustbox}
+\usepackage{xparse}
 \usepgfplotslibrary{units}
@@
 
-% https://tex.stackexchange.com/a/183682/2066
+\provideboolean{tobesplit}
+
+\ifthenelse{\boolean{tobesplit}}{%
+  \AtBeginDocument{\renewcommand{\appendixautorefname}{Online Appendix}}%
+}{}
+%https://tex.stackexchange.com/a/149897/2066
 \makeatletter
-
-% define a macro \Autoref to allow multiple references to be passed to \autoref
-\newcommand\Autoref[1]{\@first@ref#1,@}
-\def\@throw@dot#1.#2@{#1}% discard everything after the dot
-\def\@set@refname#1{%    % set \@refname to autoefname+s using \getrefbykeydefault
-    \edef\@tmp{\getrefbykeydefault{#1}{anchor}{}}%
-    \xdef\@tmp{\expandafter\@throw@dot\@tmp.@}%
-    \ltx@IfUndefined{\@tmp autorefnameplural}%
-         {\def\@refname{\@nameuse{\@tmp autorefname}s}}%
-         {\def\@refname{\@nameuse{\@tmp autorefnameplural}}}%
+\patchcmd{\hyper@makecurrent}{%
+    \ifx\Hy@param\Hy@chapterstring
+        \let\Hy@param\Hy@chapapp
+    \fi
+}{%
+    \iftoggle{inappendix}{%true-branch
+        % list the names of all sectioning counters here
+        \@checkappendixparam{chapter}%
+        \@checkappendixparam{section}%
+        \@checkappendixparam{subsection}%
+        \@checkappendixparam{subsubsection}%
+        \@checkappendixparam{paragraph}%
+        \@checkappendixparam{subparagraph}%
+    }{}%
+}{}{\show\hyper@makecurrent\errmessage{failed to patch}}
+
+\newcommand*{\@checkappendixparam}[1]{%
+    \def\@checkappendixparamtmp{#1}%
+    \ifx\Hy@param\@checkappendixparamtmp
+        \let\Hy@param\Hy@appendixstring
+    \fi
 }
-\def\@first@ref#1,#2{%
-  \ifx#2@\autoref{#1}\let\@nextref\@gobble% only one ref, revert to normal \autoref
-  \else%
-    \@set@refname{#1}%  set \@refname to autoref name
-    \@refname~\ref{#1}% add autoefname and first reference
-    \let\@nextref\@next@ref% push processing to \@next@ref
-  \fi%
-  \@nextref#2%
+\makeatletter
+
+\newtoggle{inappendix}
+\togglefalse{inappendix}
+
+\apptocmd{\appendix}{\toggletrue{inappendix}\renewcommand{\appendixautorefname}{Appendix}}{}{\show\appendix\errmessage{failed to patch}}
+%\apptocmd{\subappendices}{\toggletrue{inappendix}}{}{\show\subappendices\errmessage{failed to patch}}
+% end appendix autoref patch
+
+% https://tex.stackexchange.com/a/98016/2066
+\pgfplotsset{
+    discard if not/.style 2 args={
+        x filter/.code={
+            \edef\tempa{\thisrow{#1}}
+            \edef\tempb{#2}
+            \ifx\tempa\tempb
+            \else
+                \def\pgfmathresult{inf}
+            \fi
         }
-\def\@next@ref#1,#2{%
-   \ifx#2@ and~\ref{#1}\let\@nextref\@gobble% at end: print and+\ref and stop
-   \else, \ref{#1}% print  ,+\ref and continue
-   \fi%
-   \@nextref#2%
     }
-
-\makeatother
+}
 
@@
 \newbibfield{title}
+\newbibfield{citet}
 \bibinput{rewriting-lower}
 \newcommand{\citetitle}[1]{\emph{\usebibentry{#1}{title}}\nocite{#1}}
-\newcommand{\todo}[1]{\textbf{TODO: #1}}
+\newcommand{\citet}[1]{\usebibentry{#1}{citet}~\cite{#1}}
+\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO: #1}}}
 \newcommand{\coqbug}[1]{\href{https://github.com/coq/coq/issues/#1}{\##1}}
+\newcommand{\coqpr}[1]{\href{https://github.com/coq/coq/pull/#1}{\##1}}
+\newcommand{\coqissue}[1]{\href{https://github.com/coq/coq/issues/#1}{\##1}}
+\NewDocumentCommand{\githublink}{m m o}{%
+  \IfNoValueTF{#3}%
+    {\href{https://github.com/#1/#2}{\texttt{#1/#2}}}%
+    {\href{https://github.com/#1/#2/tree/#3}{\texttt{#1/#2@#3}}}%
+}
+\NewDocumentCommand{\githublinkurl}{m m o}{%
+  \IfNoValueTF{#3}%
+    {\url{https://github.com/#1/#2}}%
+    {\url{https://github.com/#1/#2/tree/#3}}%
+}
 \newcommand{\Rtac}{\ensuremath{\mathcal{R}_\text{\emph{tac}}}}
 \newcommand{\Ltac}{\ensuremath{\mathcal{L}_\text{\emph{tac}}}}
+\newcommand{\tacvmcompute}{\mintinline{coq}{vm_compute}}
+\newcommand{\tacnativecompute}{\mintinline{coq}{native_compute}}
+\newcommand{\taccbv}{\mintinline{coq}{cbv}}
+\newcommand{\taclazy}{\mintinline{coq}{lazy}}
+\newcommand{\tacsimpl}{\mintinline{coq}{simpl}}
+\newcommand{\taccbn}{\textbf{\texttt{cbn}}}
 \newcommand{\defeq}{=}
@@
   \newwrite\@appendixpagefile
-  \immediate\openout\@appendixpagefile=\jobname.appendix.startpage
-  \immediate\write\@appendixpagefile{\detokenize{\def\appendixstartpage}{\thepage}}
+  \immediate\openout\@appendixpagefile=\jobname.appendix.startpage\relax
+  \immediate\write\@appendixpagefile{\detokenize{\def\appendixstartpage}{\thepage}}%
 }
-\newcommand{\supplement}{%
+\newcommand{\startsupplement}{%
   \clearpage
   \newwrite\@supplementpagefile
-  \immediate\openout\@supplementpagefile=\jobname.supplement.startpage
-  \immediate\write\@supplementpagefile{\detokenize{\def\supplementstartpage}{\thepage}}
+  \immediate\openout\@supplementpagefile=\jobname.supplement.startpage\relax
+  \immediate\write\@supplementpagefile{\detokenize{\def\supplementstartpage}{\thepage}}%
+  \newwrite\@supplementonlypagefile
+  \immediate\openout\@supplementonlypagefile=\jobname.supplement.onlystartpage\relax
+  \immediate\write\@supplementonlypagefile{\thepage}%
 }%
+\AtEndDocument{%
+  \newwrite\@endsupplementonlypagefile
+  \immediate\openout\@endsupplementonlypagefile=\jobname.supplement.onlyendpage\relax
+  \immediate\write\@endsupplementonlypagefile{\thepage}%
+}
 \makeatother
 
+\title{Accelerating Verified-Compiler Development with a Verified Rewriting Engine}
+
+%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line
+
+\author{Jason Gross}{CSAIL, Massachusetts Institute of Technology, 77 Massachusetts Ave., Cambridge, MA 02139, USA \and MIRI, USA \and \url{https://jasongross.github.io/} }{jgross@mit.edu}{https://orcid.org/0000-0002-9427-4891}{}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty.
+Please provide at least the name of the affiliation and the country.
+The full address is optional.
+Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.
+
+\author{Andres Erbsen}{CSAIL, Massachusetts Institute of Technology, 77 Massachusetts Ave., Cambridge, MA 02139, USA \and \url{https://andres.systems/} }{andreser@mit.edu}{}{}
+
+\author{Jade Philipoom%\footnote{now with Google}
+}{CSAIL, Massachusetts Institute of Technology, 77 Massachusetts Ave., Cambridge, MA 02139, USA \and Google}{jade.philipoom@gmail.com}{}{}
+% should we \footnote{now with Google}?
+
+\author{Miraya Poddar-Agrawal}{Reed College, 3203 SE Woodstock Blvd, Portland, OR 97202, USA% \and \url{https://github.com/miraya-pa}
+}{ragrawal@reed.edu}{https://orcid.org/0000-0001-7617-9180}{}
+
+\author{Adam Chlipala}{CSAIL, Massachusetts Institute of Technology, 77 Massachusetts Ave., Cambridge, MA 02139, USA \and \url{http://adam.chlipala.net/} }{adamc@csail.mit.edu}{https://orcid.org/0000-0001-7085-9417}{%[funding]
+}
+
+\authorrunning{J.
+Gross and A.
+Erbsen and J.
+Philipoom and M.
+Poddar-Agrawal and A.
+Chlipala} %TODO mandatory.
+First: Use abbreviated first/middle names.
+Second (only in severe cases): Use first author plus 'et al.'
+
+\Copyright{Jason Gross \& Andres Erbsen \& Jade Philipoom \& Miraya Poddar-Agrawal \& Adam Chlipala} %TODO mandatory, please use full first names.
+LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
+
+\begin{CCSXML}
+<ccs2012>
+   <concept>
+       <concept_id>10003752.10003790.10002990</concept_id>
+       <concept_desc>Theory of computation~Logic and verification</concept_desc>
+       <concept_significance>300</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10003752.10003790.10003798</concept_id>
+       <concept_desc>Theory of computation~Equational logic and rewriting</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10011007.10011006.10011041</concept_id>
+       <concept_desc>Software and its engineering~Compilers</concept_desc>
+       <concept_significance>300</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10011007.10011006.10011041.10011046</concept_id>
+       <concept_desc>Software and its engineering~Translator writing systems and compiler generators</concept_desc>
+       <concept_significance>300</concept_significance>
+       </concept>
+ </ccs2012>
+\end{CCSXML}
+
+\ccsdesc[300]{Theory of computation~Logic and verification}
+\ccsdesc[500]{Theory of computation~Equational logic and rewriting}
+\ccsdesc[300]{Software and its engineering~Compilers}
+\ccsdesc[300]{Software and its engineering~Translator writing systems and compiler generators}
+
+\keywords{compiler verification, rewriting engines, cryptography} %TODO mandatory; please add comma-separated list of keywords
+
+%\category{} %optional, e.g. invited paper
+
+\ifthenelse{\boolean{tobesplit}}{%
+%  \relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
+%  \relatedversiondetails%[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]
+%      {Full Version}{https://arxiv.org/abs/2205.00862} %linktext and cite are optional
+}{}
+
+\supplement{%
+%  \url{https://jasongross.github.io/papers/2022-rewriting-itp-supplement.pdf},
+  \githublinkurl{mit-plv}{rewriter}[ITP-2022-perf-data],
+  \githublinkurl{mit-plv}{fiat-crypto}[perf-testing-data-ITP-2022-rewriting]%
+}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
+%\supplementdetails[linktext={opt.
+text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g.
+Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional
+
+\funding{%
+  This work was supported in part by a Google Research Award, National Science Foundation grants CCF-1253229, CCF-1512611, and CCF-1521584, and the National Science Foundation Graduate Research Fellowship under Grant Nos.\ 1122374 and 1745302.
+  Any opinion, findings, and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the National Science Foundation.%
+}
+%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors.
+Please enter author specific funding statements as fifth argument of the \author macro.
+
+%\acknowledgements{I want to thank \dots}%optional
+
+%\nolinenumbers %uncomment to disable line numbering
+
+%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\EventEditors{June Andronick and Leonardo de Moura}
+\EventNoEds{2}
+\EventLongTitle{13th International Conference on Interactive Theorem Proving (ITP 2022)}
+\EventShortTitle{ITP 2022}
+\EventAcronym{ITP}
+\EventYear{2022}
+\EventDate{August 7--10, 2022}
+\EventLocation{Haifa, Israel}
+\EventLogo{}
+\SeriesVolume{237}
+\ArticleNo{5}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{document}
 
-%% Title information
-\title{A Framework for Building Verified Partial Evaluators}         %% [Short Title] is optional;
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-%\titlenote{with title note}             %% \titlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-%\subtitle{Subtitle}                     %% \subtitle is optional
-%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{Jason Gross}
-%\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{0000-0002-9427-4891}             %% \orcid is optional
-\affiliation{
-  %\position{Position1}
-  \department{CSAIL}              %% \department is recommended
-  \institution{MIT}            %% \institution is required
-  %\streetaddress{Street1 Address1}
-  \city{Cambridge}
-  \state{MA}
-  \postcode{02139}
-  \country{USA}                    %% \country is recommended
-}
-\email{jgross@mit.edu}          %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{Andres Erbsen}
-%\authornote{with author2 note}          %% \authornote is optional;
-%% can be repeated if necessary
-%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-    %\position{Position2a}
-    \department{CSAIL}             %% \department is recommended
-    \institution{MIT}           %% \institution is required
-    %\streetaddress{Street2a Address2a}
-    \city{Cambridge}
-    \state{MA}
-    \postcode{02139}
-    \country{USA}                    %% \country is recommended
-}
-\email{andreser@mit.edu}         %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{Adam Chlipala}
-%\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  %\position{Position2a}
-  \department{CSAIL}             %% \department is recommended
-  \institution{MIT}           %% \institution is required
-  %\streetaddress{Street2a Address2a}
-  \city{Cambridge}
-  \state{MA}
-  \postcode{02139}
-  \country{USA}                    %% \country is recommended
-}
-\email{adamc@csail.mit.edu}         %% \email is recommended
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
+\maketitle
+
+%TODO mandatory: add short abstract of the document
 \begin{abstract}
-  \emph{Partial evaluation} is a classic technique for generating lean, customized code from libraries that start with more bells and whistles.
-  It is also an attractive approach to creation of \emph{formally verified} systems, where theorems can be proved about libraries, yielding correctness of all specializations ``for free.''
-  However, it can be challenging to make library specialization both performant (at compile time and runtime) and trustworthy.
-  We present a new approach, prototyped in the Coq proof assistant, which supports specialization at the speed of native-code execution, without adding to the trusted code base.
-  Our extensible engine, which combines the traditional concepts of tailored term reduction and automatic rewriting from hint databases, is also of interest to replace these ingredients in proof assistants' proof checkers and tactic engines, at the same time as it supports extraction to standalone compilers from library parameters to specialized code.
+  Compilers are a prime target for formal verification, since compiler bugs invalidate higher-level correctness guarantees, but compiler changes may become more labor-intensive to implement, if they must come with proof patches.
+  One appealing approach is to present compilers as sets of algebraic rewrite rules, which a generic engine can apply efficiently.
+  Now each rewrite rule can be proved separately, with no need to revisit past proofs for other parts of the compiler.
+  We present the first realization of this idea, in the form of a framework for the Coq proof assistant.
+  Our new Coq command takes normal proved theorems and combines them automatically into fast compilers with proofs.
+  We applied our framework to improve the Fiat Cryptography toolchain for generating cryptographic arithmetic, producing an extracted command-line compiler that is about 1000$\times$ faster while actually featuring simpler compiler-specific proofs.
 \end{abstract}
 
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-%% \begin{CCSXML}
-%% <ccs2012>
-%% <concept>
-%% <concept_id>10011007.10011006.10011008</concept_id>
-%% <concept_desc>Software and its engineering~General programming languages</concept_desc>
-%% <concept_significance>500</concept_significance>
-%% </concept>
-%% <concept>
-%% <concept_id>10003456.10003457.10003521.10003525</concept_id>
-%% <concept_desc>Social and professional topics~History of programming languages</concept_desc>
-%% <concept_significance>300</concept_significance>
-%% </concept>
-%% </ccs2012>
-%% \end{CCSXML}
-
-%% \ccsdesc[500]{Software and its engineering~General programming languages}
-%% \ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-%% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction} \label{sec:intro}
-
-Mechanized proof is gaining in importance for development of critical software infrastructure.
-Oft-cited examples include the CompCert verified C compiler~\cite{Compcert} and the seL4 verified operating-system microkernel~\cite{seL4SOSP09}.
-Here we have very flexible systems that are ready to adapt to varieties of workloads, be they C source programs for CompCert or application binaries for seL4.
-For a verified operating system, such adaptation takes place at \emph{runtime}, when we launch the application.
-However, some important bits of software infrastructure commonly do adaptation at \emph{compile time}, such that the fully general infrastructure software is not even installed in a deployed system.
-
-Of course, compilers are a natural example of that pattern, as we would not expect CompCert itself to be installed on an embedded system whose application code was compiled with it.
-The problem is that writing a compiler is rather labor-intensive, with its crafting of syntax-tree types for source, target, and intermediate languages, its fine-tuning of code for transformation passes that manipulate syntax trees explicitly, and so on.
-An appealing alternative is \emph{partial evaluation}~\cite{PartialEvaluation}, which relies on reusable compiler facilities to specialize library code to parameters, with no need to write that library code in terms of syntax-tree manipulations.
-Cutting-edge tools in this tradition even make it possible to use high-level functional languages to generate performance-competitive low-level code, as in Scala's Lightweight Modular Staging~\cite{LMS}.
-
-It is natural to try to port this approach to construction of systems with mechanized proofs.
-On one hand, the typed functional languages in popular proof assistants' logics make excellent hosts for flexible libraries, which can often be specialized through means as simple as partial application of curried functions.
-Term-reduction systems built into the proof assistants can then generate the lean residual programs.
-On the other hand, it is surprisingly difficult to realize the last sentence with good performance.
-The challenge is that we are not just implementing algorithms; we also want a proof to be checked by a small proof checker, and there is tension in designing such a checker, as fancier reduction strategies grow the trusted code base.
-It would seem like an abandonment of the spirit of proof assistants to bake in a reduction strategy per library, yet effective partial evaluation tends to be rather fine-tuned in this way.
-Performance tuning matters when generated code is thousands of lines long.
-
-In this paper, we present an approach to verified partial evaluation in proof assistants, which requires no changes to proof checkers.
-To make the relevance concrete, we use the example of Fiat Cryptography~\cite{FiatCryptoSP19}, a Coq library that generates code for big-integer modular arithmetic at the heart of elliptic-curve-cryptography algorithms.
-This domain-specific compiler has been adopted, for instance, in the Chrome Web browser, such that about half of all HTTPS connections from browsers are now initiated using code generated (with proof) by Fiat Cryptography.
-However, Fiat Cryptography was only used successfully to build C code for the two most widely used curves (P-256 and Curve25519).
-Their method of partial evaluation timed out trying to compile code for the third most widely used curve (P-384).
-Additionally, to achieve acceptable reduction performance, the library code had to be written manually in continuation-passing style.
-We will demonstrate a new Coq library that corrects both weaknesses, while maintaining the generality afforded by allowing rewrite rules to be mixed with partial evaluation.
-
-\subsection{A Motivating Example} \label{sec:motivating-example} \label{sec:explain-ident.eagerly} \label{sec:explain-eval-rect} \label{sec:explain-'}
-
-We are interested in partial-evaluation examples that mix higher-order functions, inductive datatypes, and arithmetic simplification.
-For instance, consider the following Coq code.
-
-\begin{minted}[fontsize=\small]{coq}
-Definition prefixSums (ls:list nat) : list nat :=
+\section{Introduction}\label{sec:intro}
+% TODO(Jason): Add section/info somewhere about NbE-order being the right order
+
+Formally verified compilers like CompCert~\cite{Compcert} and CakeML~\cite{CakeML} are success stories for proof assistants, helping close a trust gap for one of the most important categories of software infrastructure.
+A popular compiler cannot afford to stay still; developers will add new backends, new language features, and better optimizations.
+Proofs must be adjusted as these improvements arrive.
+It makes sense that the author of a new piece of compiler code must prove its correctness, but ideally there would be no need to revisit old proofs.
+There has been limited work, though, on avoiding that kind of coupling.
+\citet{xcert} demonstrated a streamlined way to extend CompCert with new verified optimizations driven by dataflow analysis, but we are not aware of past work that supports easy extension for compilers from functional languages to C code.
+We present our work targeting that style.
+
+One strategy for writing compilers modularly is to exercise foresight in designing a core that will change very rarely, such that feature iteration happens outside the core.
+Specifically, phrasing the compiler in terms of rewrite rules allows clean abstractions and conceptual boundaries~\cite{Hickey2006}.
+Then, most desired iteration on the compiler can be achieved through iteration on the rewrite rules.
+
+It is surprisingly difficult to realize this modular approach with good performance.
+Verified compilers can either be proof-producing (certifying) or proven-correct (certified).
+Proof-producing compilers usually operate on the functional languages of the proof assistants that they are written in, and variable assignments are encoded as let binders.
+All existing proof-producing rewriting strategies scale at least quadratically in the number of binders.
+This performance scaling is inadequate for applications like Fiat Cryptography~\cite{FiatCryptoSP19} where the generated code has 1000s of variables in a single function.
+Proven-correct compilers do not suffer from this asymptotic blowup in the number of binders.
+
+In this paper, we present \textbf{the first proven-correct compiler-builder toolkit parameterized on rewrite rules}.
+Arbitrary sets of Coq theorems (quantified equalities) can be assembled by a single new Coq command into an extraction-ready verified compiler.
+We did not need to extend the trusted code base, so our compiler compiler need not be trusted.
+We achieve both good performance of compiler runs and good performance of generated code, via addressing a number of scale-up challenges vs.\ past work.
+%including efficient handling of intermediate code with thousands of nested binders and critical subterm sharing.
+
+We evaluate our toolkit by replacing a key component of Fiat Cryptography~\cite{FiatCryptoSP19}, a Coq library that generates code for big-integer modular arithmetic at the heart of elliptic-curve-cryptography algorithms.
+Routines generated (with proof) with Fiat Cryptography now ship with all major Web browsers and all major mobile operating systems.
+With our improved compiler architecture, it became easy to add two new backends and a variety of new supported source-code features, and we were easily able to try out new optimizations.
+%and later replace unsound optimizations with equivalent sound optimizations.
+
+Replacing Fiat Cryptography's original compiler with the compiler generated by our toolkit has two additional benefits.
+Fiat Cryptography was previously only used successfully to build C code for two of the three most widely used curves (P-256 and Curve25519).
+Our prior version's execution timed out trying to compile code for the third most widely used curve (P-384).
+Using our new toolkit has made it possible to generate compiler-synthesized code for P-384 while generating completely identical code for the primes handled by the previous version, about 1000$\times$ more quickly.
+Additionally, Fiat Cryptography previously required source code to be written in continuation-passing style, and our compiler has enabled a direct-style approach, which pays off in simplifying theorem statements and proofs.
+
+%There are a number of subtleties that arise in building a verified compiler around rewrite rules.
+%There are two important performance criteria: performance of the compiler in generating code, and the performance of the generated code itself.
+
+%The typed functional languages in popular proof assistants' logics make excellent hosts for flexible libraries, which can often be specialized through means as simple as partial application of curried functions.
+%Term-reduction systems built into the proof assistants can then generate the lean residual programs.
+%It is surprisingly difficult to realize the last sentence with good performance.
+%The challenge is that we are not just implementing algorithms; we also want a proof to be checked by a small proof checker, and there is tension in designing such a checker, as fancier reduction strategies grow the trusted code base.
+%It would seem like an abandonment of the spirit of proof assistants to bake in a reduction strategy per library, yet effective partial evaluation tends to be rather fine-tuned in this way.
+%Performance tuning matters when generated code is thousands of lines long.
+
+% While there are nascent efforts in proof repair to mitigate the cost of iteration in a verified setting~\cite{Ringer2021}, we
+
+%Mechanized proof is gaining in importance for development of critical software infrastructure.
+%Oft-cited examples include the CompCert verified C compiler~\cite{Compcert} and the seL4 verified operating-system microkernel~\cite{seL4SOSP09}.
+%Here we have very flexible systems that are ready to adapt to varieties of workloads, be they C source programs for CompCert or application binaries for seL4.
+%For a verified operating system, such adaptation takes place at \emph{runtime}, when we launch the application.
+%However, some important bits of software infrastructure commonly do adaptation at \emph{compile time}, such that the fully general infrastructure software is not even installed in a deployed system.
+
+%Of course, compilers are a natural example of that pattern, as we would not expect CompCert itself to be installed on an embedded system whose application code was compiled with it.
+%The problem is that writing a compiler is rather labor-intensive, with its crafting of syntax-tree types for source, target, and intermediate languages, its fine-tuning of code for transformation passes that manipulate syntax trees explicitly, and so on.
+%An appealing alternative is \emph{partial evaluation}~\cite{PartialEvaluation}, which relies on reusable compiler facilities to specialize library code to parameters, with no need to write that library code in terms of syntax-tree manipulations.
+%Cutting-edge tools in this tradition even make it possible to use high-level functional languages to generate performance-competitive low-level code, as in Scala's Lightweight Modular Staging~\cite{LMS}.
+
+%It is natural to try to port this approach to construction of systems with mechanized proofs.
+%On one hand, the typed functional languages in popular proof assistants' logics make excellent hosts for flexible libraries, which can often be specialized through means as simple as partial application of curried functions.
+%Term-reduction systems built into the proof assistants can then generate the lean residual programs.
+%On the other hand, it is surprisingly difficult to realize the last sentence with good performance.
+%The challenge is that we are not just implementing algorithms; we also want a proof to be checked by a small proof checker, and there is tension in designing such a checker, as fancier reduction strategies grow the trusted code base.
+%It would seem like an abandonment of the spirit of proof assistants to bake in a reduction strategy per library, yet effective partial evaluation tends to be rather fine-tuned in this way.
+%Performance tuning matters when generated code is thousands of lines long.
+
+\subsection{Related Work}\label{sec:related-one}
+% POPL 2022 review TODO:
+% Section 1.1, it may be a little early at this point to discuss the related work, as I don't think it is quite clear yet exactly what problem this paper solves.
+% From the first 1.5 pages, I understand that the Fiat Cryptography compiler has been improved, but it is not clear yet how this is achieved and what are the merits or demerits of the approach presented in this paper.
+
+Assume our mission is to take libraries of purely functional combinators, apply them to compile-time parameters, and compile the results down to lean C code.
+Furthermore, we ask for machine-checked proofs that the C programs preserve the behavior of the higher-order functional programs we started with.
+What good ideas from the literature can we build on?
+
+\citet{Hickey2006} discuss at length how to build compilers around rewrite rules.
+``All program transformations, from parsing to code generation, are cleanly isolated and specified as term rewrites.''
+While they note that the correctness of the compiler is thus reduced to the correctness of the rewrite rules, they did not prove correctness mechanically.
+Furthermore, it is not clear that they manage to avoid the asymptotic blow-up associated with proof-producing rewriting of deeply nested let-binders.
+They give no performance numbers, so it is hard to say whether or not their compiler performs at the scale necessary for Fiat Cryptography.
+Their rewrite-engine driver is unproven OCaml code, while we will produce custom drivers with Coq proofs.
+%This suggests also that they cannot use rewrite rules which are only sound in the presence of side conditions, as such side-condition checking could not happen in the engine driving the rewrites.
+%Overall, their work complements ours; we focus on how to build a scalable engine for building a proven-correct compiler around shallowly embedded rewrite rules in a proof assistant like Coq; they focus on what rewrite rules are necessary for implementing a general-purpose compiler.
+
+\Rtac{}~\cite{rtac} is a more general framework for verified proof tactics in Coq, including an experimental reflective version of \texttt{rewrite\_strat} supporting arbitrary setoid relations, unification variables, and arbitrary semidecidable side conditions solvable by other verified tactics, using de Bruijn indexing to manage binders.
+We found that \Rtac{} misses a critical feature for compiling large programs: preserving subterm sharing.
+As a result, our experiments with compiler construction yielded clear asymptotic slowdown vs.\ what we eventually accomplished.
+\Rtac{} is also more heavyweight to use, for instance requiring that theorems be restated manually in a deep embedding to bring them into automation procedures.
+Furthermore, we are not aware of any past experiments driving verified compilers with \Rtac{}.
+%, and we expect there would be other bottlenecks working with large, deeply nested terms.
+
+\citet{Aehlig} came closest to a fitting approach, using \emph{normalization by evaluation (NbE)}~\cite{NbE} to bootstrap reduction of open terms on top of full reduction, as built into a proof assistant.
+% POPL 2022 review TODO: Line 99: I'd drop the "it was simultaneously true that".
+However, it was simultaneously true that they expanded the proof-assistant trusted code base in ways specific to their technique, and that they did not report any experiments actually using the tool for partial evaluation (just traditional full reduction), potentially hiding performance-scaling challenges or other practical issues.
+For instance, they also do not preserve subterm sharing explicitly, and they represent variable references as unary natural numbers (de Bruijn-style).
+They also require that rewrite rules be embodied in ML code, rather than stated as natural ``native'' lemmas of the proof assistant.
+We will follow their basic outline with important modifications.
+
+Our implementation builds on fast full reduction in Coq's kernel, via a virtual machine~\cite{vmcompute} or compilation to native code~\cite{nativecompute} (neither verified).
+Especially the latter is similar in adopting NbE for full reduction, simplifying even under $\lambda$s, on top of a more traditional implementation of OCaml that never executes preemptively under $\lambda$s.
+Neither approach unifies support for rewriting with proved rules, and partial evaluation only applies in very limited cases.
+
+A variety of forms of pragmatic partial evaluation have been demonstrated, with Lightweight Modular Staging~\cite{LMS} in Scala as one of the best-known current examples.
+The LMS-Verify system~\cite{LMSVerify} can be used for formal verification of generated code after-the-fact.
+Typically LMS-Verify has been used with relatively shallow properties (though potentially applied to larger and more sophisticated code bases than we tackle), not scaling to the kinds of functional-correctness properties that concern us here.
+%justifying investment in verified partial evaluators.
+
+So, overall, to our knowledge, no past compiler as a set of rewrite rules has come with a full proof of correctness as a standalone functional program.
+Related prior work with mechanized proofs suffered from both performance bottlenecks and usability problems, the latter in requiring that eligible rewrite rules be stated in special deep embeddings.
+%We will demonstrate all of these pieces working together well enough to be adopted by several high-profile open-source projects.
+
+\subsection{Our Solution}\label{sec:our-solution}
+
+% POPL 2022 review TODO:
+% line 114, "Our variant on the technique of Aehlig et al. [...]".
+% The authors have not yet recalled what is Aehlig et al.'s technique, nor explained how this work differs from it, so it is difficult to appreciate the list of claimed advantages at this point.
+Our variant on the technique of \citet{Aehlig} has these advantages:
+
+\begin{itemize}
+\item It integrates with a general-purpose, foundational proof assistant, \textbf{without growing the trusted code base}.
+\item For a wide variety of initial functional programs, it provides \textbf{fast} partial evaluation with reasonable memory use.
+\item It allows reduction that \textbf{mixes} \emph{rules of the definitional equality} with \emph{equalities proven explicitly as theorems}.
+\item It allows \textbf{rapid iteration} on rewrite rules with \emph{minimal verification overhead}.
+\item It \textbf{preserves sharing} of common subterms.
+\item It also allows \textbf{extraction of standalone compilers}.
+\end{itemize}
+
+Our contributions include answers to a number of challenges that arise in scaling NbE-based partial evaluation in a proof assistant.
+First, we rework the approach of \citet{Aehlig} to function \emph{without extending a proof assistant's trusted code base}, which, among other challenges, requires us to prove termination of reduction and encode pattern matching explicitly (leading us to adopt the performance-tuned approach of \citet{maranget2008compiling}).
+We also improve on Coq-specific related work (e.g., of \citet{rtac}) by allowing rewrites to be written in natural Coq form (not special embedded syntax-tree types), while supporting optimizations associated with past unverified engines (e.g., \citet{Boespflug2009}).
+
+Second, using partial evaluation to generate residual terms thousands of lines long raises \emph{new scaling challenges}:
+\begin{itemize}
+\item
+  Output terms may contain so \emph{many nested variable binders} that we expect it to be performance-prohibitive to perform bookkeeping operations on first-order-encoded terms (e.g., with de Bruijn indices, as is done in \Rtac{} by \citet{rtac}).
+  For instance, while the reported performance experiments of \citet{Aehlig} generate only closed terms with no binders, Fiat Cryptography may generate a single routine (e.g., multiplication for curve P-384) with nearly a thousand nested binders.
+%\todo{Consider being more specific here about expected cost of first-order representation, maybe using some experiment.}
+%  \todo{Note from Jason: We haven't actually checked that the bookkeeping is performance-prohibitive.
+I suspect there will be a cost, but that if when we made fiat-crypto there was an existing system that used de Bruijn indices and did this, (along with built-in let-lifting, etc), I suspect we might have been a bit more sad, but not prohibitively more sad.
+I think we should be much more tentative with this bullet point.}
+\item
+  Naive representation of terms without proper \emph{sharing of common subterms} can lead to fatal term-size blow-up.
+  %Fiat Cryptography's arithmetic routines rely on significant sharing of this kind.
+\item
+  Unconditional rewrite rules are in general insufficient, and we need \emph{rules with side conditions}.
+  E.g., Fiat Cryptography depends on checking lack-of-overflow conditions.
+\item
+  However, it is also not reasonable to expect a general engine to discharge all side conditions on the spot.
+  We need integration with \emph{abstract interpretation}.
+  %that can analyze whole programs to support reduction.
+\end{itemize}
+
+Briefly, our respective solutions to these problems are the \emph{parametric higher-order abstract syntax (PHOAS)}~\cite{PhoasICFP08} term encoding, a \emph{let-lifting} transformation threaded throughout reduction, extension of rewrite rules with executable Boolean side conditions, and a design pattern that uses decorator function calls to include analysis results in a program.
+
+Finally, we carry out the \emph{first large-scale performance-scaling evaluation} of a verified rewrite-rule-based compiler, covering all elliptic curves from the published Fiat Cryptography experiments, along with microbenchmarks.
+
+We pause to give a motivating example before presenting the core structure of our engine (\autoref{sec:structure}), the additional scaling challenges we faced (\autoref{sec:scaling}), experiments (\autoref{sec:evaluation}), and conclusions.
+Our implementation is attached.
+
+\section{A Motivating Example}\label{sec:motivating-example}\label{sec:explain-ident.eagerly}\label{sec:explain-eval-rect}
+
+Our compilation style involves source programs that mix higher-order functions and inductive types.
+We want to compile to C code, reducing away uses of fancier features while seizing opportunities for arithmetic simplification.
+Here is a small but illustrative example.
+
+\begin{minted}[fontsize=\small]{coq}
+Definition prefixSums (ls : list nat) : list nat :=
   let ls' := combine ls (seq 0 (length ls)) in
@@
   let '(_, ls''') := fold_left (λ '(acc, ls''') n,
-    let acc' := acc + n in (acc', acc' :: ls''')) ls'' (0, []) in
-  ls'''.
+    let acc' := acc + n in (acc', acc' :: ls''')) ls'' (0, []) in ls'''.
 \end{minted}
@@
 Then we map over the list of pairs, multiplying the components at each position.
-Finally, we traverse that list, building up a list of all prefix sums.
-
+Finally, we compute all prefix sums.
+
+% POPL 2022 review TODO:
+% line 178, I suppose I could have understood this earlier, but it is becoming clear only now that the paper is about partial evaluation, and about a form of supercompilation.
 We would like to specialize this function to particular list lengths.
@@
 \begin{minted}[fontsize=\small]{coq}
-let acc := b + c * 2 in
-let acc' := acc + d * 3 in
-  [acc'; acc; b; 0]
-\end{minted}
+let acc := b + c * 2 in let acc' := acc + d * 3 in [acc'; acc; b; 0]
+\end{minted}
+
+We do not quite have C code yet, but, composing this code with another routine to consume the output list, we easily arrive at a form that looks almost like three-address code and is quite easy to translate to C and many other languages.
+%That actual translation we leave to other small compiler phases outside the scope of this paper.
 
@@
 
-To set up our partial evaluator, we prove the algebraic laws that it should use for simplification, starting with basic arithmetic identities.
-\begin{minted}[fontsize=\small]{coq}
-Lemma zero_plus : forall n, 0 + n = n.       Lemma times_zero : forall n, n * 0 = 0.
-Lemma plus_zero : forall n, n + 0 = n.       Lemma times_one  : forall n, n * 1 = n.
+To set up our compiler, we prove the algebraic laws that it should use for simplification, starting with basic arithmetic identities.
+\begin{minted}[fontsize=\small]{coq}
+Lemma zero_plus : ∀ n, 0 + n = n.      Lemma times_zero : ∀ n, n * 0 = 0.
+Lemma plus_zero : ∀ n, n + 0 = n.      Lemma times_one  : ∀ n, n * 1 = n.
 \end{minted}
@@
 Next, we prove a law for each list-related function, connecting it to the primitive-recursion combinator for some inductive type (natural numbers or lists, as appropriate).
-We use a special apostrophe marker to indicate a quantified variable that may only match with \emph{compile-time constants}.
-We also use a further marker \mintinline{coq}{ident.eagerly} to ask the reducer to simplify a case of primitive recursion by complete traversal of the designated argument's constructor tree.
+We also use a further marker \mintinline{coq}{ident.eagerly} to ask the compiler to simplify a case of primitive recursion by complete traversal of the designated argument's constructor tree.
 %\todo{Hey, it sure would be nice to generate all of these to-eliminator lemmas automatically, too!  Next step for the plugin?}
@@
 
-With all the lemmas available, we can package them up into a rewriter, which triggers generation of a specialized rewrite procedure and its soundness proof.
+With all the lemmas available, we can package them up into a rewriter, which triggers generation of a specialized compiler and its soundness proof.
 Our Coq plugin introduces a new command \mintinline{coq}{Make} for building rewriters
@@
 Most inputs to \mintinline{coq}{Rewriter For} list quantified equalities to use for left-to-right rewriting.
-However, we also use options \mintinline{coq}{do_again}, to request that some rules trigger an extra bottom-up pass after being used for rewriting; \mintinline{coq}{eval_rect}, to queue up eager evaluation of a call to a primitive-recursion combinator on a known recursive argument; \mintinline{coq}{with delta}, to request evaluation of all monomorphic operations on concrete inputs; and \mintinline{coq}{with extra idents}, to inform the engine of further permitted identifiers that do not appear directly in any of the rewrite rules.
-
-Our plugin also provides new tactics like \mintinline{coq}{Rewrite_rhs_for}, which applies a rewriter to the righthand side of an equality goal.
-That last tactic is just what we need to synthesize a specialized \mintinline{coq}{prefixSums} for list length four, along with a proof of its equivalence to the original function.
+However, we also use options \mintinline{coq}{do_again}, to request that some rules trigger extra bottom-up passes after being used for rewriting; \mintinline{coq}{eval_rect}, to queue up eager evaluation of a call to a primitive-recursion combinator on a known recursive argument; \mintinline{coq}{with delta}, to request evaluation of all monomorphic operations on concrete inputs; and \mintinline{coq}{with extra idents}, to inform the engine of further permitted identifiers that do not appear directly in any of the rewrite rules.
+
+Our plugin also provides new tactics like \mintinline{coq}{Rewrite_rhs_for}, which applies a rewriter to the right-hand side of an equality goal.
+That last tactic is just what we need to synthesize a specialized \mintinline{coq}{prefixSums} for list length four, along with its correctness proof.
 \begin{minted}[fontsize=\small]{coq}
 Definition prefixSums4 :
- {f : nat → nat → nat → nat → list nat | ∀ a b c d, f a b c d = prefixSums [a;b;c;d]}
+{f:nat→nat→nat→nat→list nat | ∀ a b c d, f a b c d = prefixSums [a;b;c;d]}
   := ltac:(eexists; Rewrite_rhs_for rewriter; reflexivity).
@@
 
-\subsection{Concerns of Trusted-Code-Base Size} \label{sec:trusted-code-base-size}
-
-Crafting a reduction strategy is challenging enough in a standalone tool.
-A large part of the difficulty in a proof assistant is reducing in a way that leaves a proof trail that can be checked efficiently by a small kernel.
-Most proof assistants present user-friendly surface tactic languages that generate proof traces in terms of more-elementary tactic steps.
-The trusted proof checker only needs to know about the elementary steps, and there is pressure to be sure that these steps are indeed elementary, not requiring excessive amounts of kernel code.
-However, hardcoding a new reduction strategy in the kernel can bring dramatic performance improvements.
-Generating thousands of lines of code with partial evaluation would be intractable if we were outputting sequences of primitive rewrite steps justifying every little term manipulation, so we must take advantage of the time-honored feature of type-theoretic proof assistants that reductions included in the definitional equality need not be requested explicitly.
-We discuss the performance issues in more detail in \autoref{sec:setoid-rewrite-bottlenecks}.
-
-Which kernel-level reductions \emph{does} Coq support today?
-Currently, the trusted code base knows about four different kinds of reduction: left-to-right conversion, right-to-left conversion, a virtual machine (VM) written in C based on the OCaml compiler, and a compiler to native code.
-Furthermore, the first two are parameterized on an arbitrary user-specified ordering of which constants to unfold when, in addition to internal heuristics about what to do when the user has not specified an unfolding order for given constants.
-Recently, native support for 63-bit integers~\cite{coq-pr-ints} and IEEE 754-2008 binary64 floats~\cite{coq-pr-floats} have been added to the VM and native machines.
-A recent pull request proposes adding support for native arrays~\cite{coq-pr-arrays}.
-
-% Not sure if we want to include the following. -Adam
-%% Some users of Coq are concerned by this increase in the TCB.
-%% Out of the 34\,000 lines of code in the Coq kernel (about 5\,000 of them comments), about 4\,250 lines are the native compiler\footnote{In files matching \texttt{kernel/*native*}.}, about 6\,650 additional lines are dedicated to the VM\footnote{In files matching \texttt{kernel/*vm*}, \texttt{kernel/*byte*}, \texttt{kernel/cemitcodes*}, \texttt{kernel/clambda*}, \texttt{kernel/csymtable*}, \texttt{kernel/vconv*}, and \texttt{kernel/genOpcodeFiles*}.}, and an additional 3\,100 lines are devoted to conversion and reduction\footnote{In \texttt{kernel/reduction*}, \texttt{kernel/cClosure*}, \texttt{kernel/primred*}, and \texttt{kernel/conv\_oracle*}.}.
-
-To summarize, there has been quite a lot of ``complexity creep'' in the Coq trusted base, to support efficient reduction, and yet realistic partial evaluation has \emph{still} been rather challenging.
-Even the additional three reduction mechanisms outside Coq's kernel (\texttt{cbn}, \texttt{simpl}, \texttt{cbv}) are not at first glance sufficient for verified partial evaluation.
+That compiler execution ran inside of Coq, but an even more pragmatic approach is to \emph{extract} the compiler as a standalone program in OCaml or Haskell.
+Such a translation is possible because the \mintinline{coq}{Make} command produces a proved program in Gallina, Coq's logic.
+As a result, our reworking of Fiat Cryptography compilation culminated in extraction of a command-line OCaml program that developers in industry have been able to run without our help, where Fiat Cryptography previously required installing and running Coq, with an elaborate build process to capture its output.
+It is also true that the standalone program is about 10$\times$ as fast as execution within Coq, though the trusted code base is larger.
+% for the extracted version, since extraction itself is not proved.
 
@@
 
-\subsection{Our Solution} \label{sec:our-solution}
-
-\citet{Aehlig} presented a very relevant solution to a related problem, using \emph{normalization by evaluation (NbE)}~\cite{NbE} to bootstrap reduction of open terms on top of full reduction, as built into a proof assistant.
-However, it was simultaneously true that they expanded the proof-assistant trusted code base in ways specific to their technique, and that they did not report any experiments actually using the tool for partial evaluation (just traditional full reduction), potentially hiding performance-scaling challenges or other practical issues.
-We have adapted their approach in a new Coq library embodying \textbf{the first partial-evaluation approach to satisfy the following criteria}.
-
+\section{The Structure of a Rewriter}\label{sec:structure}
+% POPL 2022 review TODO:
+% line 243, "We are [...] but made [...]".
+% Stylistically, I suggest sticking with present tense everywhere.
+% See also "a type is defined [...] they also wrote [...]" on the next page.
+We are mostly guided by \citet{Aehlig} but made a number of crucial changes.
+Let us review the basic idea of the approach of Aehlig et al.
+% POPL 2022 review TODO:
+% The description of Aehlig et al.'s approach on page 6 is difficult to follow (for a non-expert reader, at least).
+% Why not describe your approach first and contrast it with Aehlig et al.'s approach in the Related Work section?
+First, their supporting library contains:
+\begin{enumerate}
+\item
+  Within the logic of the proof assistant (Isabelle/HOL, in their case), a type of syntax trees for ML programs is defined, with an associated (trusted) operational semantics.
+\item
+  They also wrote a reduction function in (deeply embedded) ML, parameterized on a function to choose the next rewrite, and proved it sound once-and-for-all.
+\end{enumerate}
+
+Given a set of rewrite rules and a term to simplify, their main tactic must:
+\begin{enumerate}
+\item
+  \emph{Generate a (deeply embedded) ML program that decides which rewrite rule, if any, to apply at the top node of a syntax tree}, along with a proof of its soundness.
+\item
+  \emph{Generate a (deeply embedded) ML term standing for the term we set out to simplify}, with a proof that it means the same as the original.
+\item
+  Combining the general proof of the rewrite engine with proofs generated by reification (the prior two steps), conclude that an application of the reduction function to the reified rules and term is indeed an ML term that generates correct answers.
+\item
+  ``Throw the ML term over the wall,'' using a general code-generation framework for Isabelle/HOL~\cite{CodeGen}.
+  Trusted code compiles the ML code into the concrete syntax of Standard ML, and compiles it, and runs it, asserting an axiom about the outcome.
+\end{enumerate}
+
+Here is where our approach differs at that level of detail:
 \begin{itemize}
-\item It integrates with a general-purpose, foundational proof assistant, \textbf{without growing the trusted base}.
-\item For a wide variety of initial functional programs, it provides \textbf{fast} partial evaluation with reasonable memory use.
-\item It allows reduction that \textbf{mixes} \emph{rules of the definitional equality} with \emph{equalities proven explicitly as theorems}.
-\item It \textbf{preserves sharing} of common subterms.
-\item It also allows \textbf{extraction of standalone partial evaluators}.
+\item
+  Our reduction engine is written \emph{as a normal Gallina functional program}, rather than within a deeply embedded language.
+  As a result, we are able to prove its type-correctness and termination, and we are able to run it within Coq's kernel.
+\item
+  We do \emph{compile-time specialization of the reduction engine} to sets of rewrite rules, removing overheads of generality.
 \end{itemize}
 
-Our contributions include answers to a number of challenges that arise in scaling NbE-based partial evaluation in a proof assistant.
-First, we rework the approach of \citet{Aehlig} to function \emph{without extending a proof assistant's trusted code base}, which, among other challenges, requires us to prove termination of reduction and encode pattern matching explicitly (leading us to adopt the performance-tuned approach of \citet{maranget2008compiling}).
-
-Second, using partial evaluation to generate residual terms thousands of lines long raises \emph{new scaling challenges}:
-\begin{itemize}
-\item Output terms may contain so \emph{many nested variable binders} that we expect it to be performance-prohibitive to perform bookkeeping operations on first-order-encoded terms (e.g., with de Bruijn indices, as is done in \Rtac{} by \citet{rtac}).
-For instance, while the reported performance experiments of \citet{Aehlig} generate only closed terms with no binders, Fiat Cryptography may generate a single routine (e.g., multiplication for curve P-384) with nearly a thousand nested binders.
-%\todo{Consider being more specific here about expected cost of first-order representation, maybe using some experiment.}
-%  \todo{Note from Jason: We haven't actually checked that the bookkeeping is performance-prohibitive.
-I suspect there will be a cost, but that if when we made fiat-crypto there was an existing system that used de Bruijn indices and did this, (along with built-in let-lifting, etc), I suspect we might have been a bit more sad, but not prohibitively more sad.
-I think we should be much more tentative with this bullet point.}
-\item Naive representation of terms without proper \emph{sharing of common subterms} can lead to fatal term-size blow-up.
-Fiat Cryptography's arithmetic routines rely on significant sharing of this kind.
-\item Unconditional rewrite rules are in general insufficient, and we need \emph{rules with side conditions}.
-For instance, in Fiat Cryptography, some rules for simplifying modular arithmetic depend on proofs that operations in subterms do not overflow.
-\item However, it is also not reasonable to expect a general engine to discharge all side conditions on the spot.
-We need integration with \emph{abstract interpretation} that can analyze whole programs to support reduction.
-\end{itemize}
-
-Briefly, our respective solutions to these problems are the \emph{parametric higher-order abstract syntax (PHOAS)}~\cite{PhoasICFP08} term encoding, a \emph{let-lifting} transformation threaded throughout reduction, extension of rewrite rules with executable Boolean side conditions, and a design pattern that uses decorator function calls to include analysis results in a program.
-
-Finally, we carry out the \emph{first large-scale performance-scaling evaluation} of partial evaluation in a proof assistant, covering all elliptic curves from the published Fiat Cryptography experiments, along with microbenchmarks.
-
-This paper proceeds through explanations of the trust stories behind our approach and earlier ones (\autoref{sec:trust}), the core structure of our engine (\autoref{sec:structure}), the additional scaling challenges we faced (\autoref{sec:scaling}), performance experiments (\autoref{sec:evaluation}), and related work (\autoref{sec:related}) and conclusions.
-Our implementation is included as an anonymous supplement.
-
-\section{Trust, Reduction, and Rewriting} \label{sec:trust}
-
-Since much of the narrative behind our design process depends on tradeoffs between performance and trustworthiness, we start by reviewing the general situation in proof assistants.
-
-Across a variety of proof assistants, simplification of functional programs is a workhorse operation.
-Proof assistants like Coq that are based on type theory typically build in \emph{definitional equality} relations, identifying terms up to reductions like $\beta$-reduction and unfolding of named identifiers.
-What looks like a single ``obvious'' step in an on-paper equational proof may require many of these reductions, so it is handy to have built-in support for checking a claimed reduction.
-\autoref{subfig:change} diagrams how such steps work in a system like Coq, where the system implementation is divided between a trusted \emph{kernel}, for checking \emph{proof terms} in a minimal language, and additional untrusted support, like a \emph{tactic} engine evaluating a language of higher-level proof steps, in the process generating proof terms out of simpler building blocks.
-It is standard to include a primitive proof step that validates any reduction compatible with the definitional equality, as the latter is decidable.
-The figure shows a tactic that simplifies a goal using that facility.
-
-In proof goals containing free variables, executing subterms can get stuck before reaching normal forms.
-However, we can often achieve further simplification by using equational rules that we prove explicitly, rather than just relying on the rules built into the definitional equality and its decidable equivalence checker.
-Coq's \texttt{autorewrite} tactic, as diagrammed in \autoref{subfig:autorewrite}, is a good example: it takes in a database of quantified equalities and applies them repeatedly to rewrite in a goal.
-It is important that Coq's kernel does not trust the \texttt{autorewrite} tactic.
-Instead, the tactic must output a proof term that, in some sense, is the moral equivalent of a line-by-line equational proof.
-It can be challenging to keep these proof terms small enough, as naive rewrite-by-rewrite versions repeatedly copy large parts of proof goals, justifying a rewrite like $C[e_1] = C[e_2]$ for some context $C$ given a proof of $e_1 = e_2$, with the full value of $C$ replicated in the proof term for that single rewrite.
-Overcoming these challenges while retaining decidability of proof checking is tricky, since we may use \texttt{autorewrite} with rule sets that do not always lead to terminating reduction.
-Coq includes more experimental alternatives like \texttt{rewrite\_strat}, which use bottom-up construction of multi-rewrite proofs, with sharing of common contexts.
-Still, as \autoref{sec:evaluation} will show, these methods that generate substantial proof terms are at significant performance disadvantages.
-We also experimented with the corresponding tactics in the Lean proof assistant, with similarly dissapointing results (\autoref{sec:lean}).
-
-\begin{figure} \centering
-  \subfloat[Reduction via the definitional equality]{%
-    \includegraphics[width=0.48\textwidth]{trust1}
-    \label{subfig:change}} \hfill %\\
-  \subfloat[Rewriting via proved rules]{%
-    \includegraphics[width=0.48\textwidth]{trust2}
-    \label{subfig:autorewrite}} \\
-  \subfloat[Approach of \citet{Aehlig}]{%
-    \includegraphics[width=0.48\textwidth]{trust3}
-    \label{subfig:aehlig}} \hfill %\\
-  \subfloat[Our approach]{%
-    \includegraphics[width=0.48\textwidth]{trust4}
-    \label{subfig:perfect}}
-
-  \caption{Different approaches to reduction and rewriting}
-  \label{fig:trust}
-\end{figure}
-
-Now we summarize how \citet{Aehlig} provide flexible and fast interleaving of standard $\lambda$-calculus reduction and use of proved equalities (the next section will go into more detail).
-\autoref{subfig:aehlig} demonstrates a workflow based on \emph{a deep embedding of a core ML-like language}.
-That is, within the logic of the proof assistant (Isabelle/HOL, in their case), a type of syntax trees for ML programs is defined, with an associated operational semantics.
-The basic strategy is, for a particular set of rewrite rules and a particular term to simplify, to \emph{generate a (deeply embedded) ML program that, if it terminates, produces a syntax tree for the simplified term}.
-Their tactic uses \emph{reification} to create ML versions of rule sets and terms.
-They also wrote a reduction function in ML and proved it sound once and for all, against the ML operational semantics.
-Combining that proof with proofs generated by reification, we conclude that an application of the reduction function to the reified rules and term is indeed an ML term that generates correct answers.
-The tactic then ``throws the ML term over the wall,'' using a general code-generation framework for Isabelle/HOL~\cite{CodeGen}.
-Trusted code compiles the ML code into the concrete syntax of a mainstream ML language, Standard ML in their case, and compiles it with an off-the-shelf compiler.
-The output of that compiled program is then passed back over to the tactic, in terms of an axiomatic assertion that the ML semantics really yields that answer.
-
-As \citet{Aehlig} argue, their use of external compilation and evaluation of ML code adds no real complexity on top of that required by the proof assistant -- after all, the proof assistant itself must be compiled and executed somehow.
-However, the perceived increase of trusted code base is not spurious:
-it is one thing to trust that the toolchain and execution environment used by the proof assistant and the partial evaluator are well-behaved,
-and another to rely on two descriptions of ML (one deeply embedded in the proof assistant and another implied by the compiler) to agree on every detail of the semantics.
-Furthermore, there still is new trusted code to translate from the deeply embedded ML subset into the concrete syntax of the full-scale ML language.
-% Aehlig et al.
-defined an ML language quite specialized to their use case, building in exactly the one algebraic datatype they need to represent NbE results, and it is unsatisfying to add such a specialized translator to the trusted base.
-% However, we could imagine using a very general deep embedding of the proof assistant in itself, as with Template-Coq~\cite{TemplateCoq}.
-% In that case, the expansion of the trusted code base would be even more worrying, to include an interpretation function from a complete syntactic embedding of the language within itself.
-% Any inconsistency between the mechanized operational semantics (which would not be small, for a full-scale practical logic) and the actual implementation could break logical soundness.
-The vast majority of proof-assistant developments today rely on no such embeddings with associated mechanized semantics, so need we really add one to a proof-checking kernel to support efficient partial evaluation?
-
-Our answer, diagrammed in \autoref{subfig:perfect}, shows a different way.
-We still reify terms and rules into a deeply embedded language.
-However, \emph{the reduction engine is implemented directly in the logic}, rather than as a deeply embedded syntax tree of an ML program.
-As a result, the kernel's own reduction engine is prepared to execute our reduction engine for us -- using an operation that would be included in a type-theoretic proof assistant in any case, with no special support for a language deep embedding.
-We also stage the process for performance reasons.
-First, the \mintinline{coq}{Make} command creates a rewriter out of a list of rewrite rules, by specializing a generic partial-evaluation engine, which has a generic proof that applies to any set of proved rewrite rules.
-We perform partial evaluation on the specialized partial evaluator, using Coq's normal reduction mechanisms, under the theory that we can afford to pay performance costs at this stage because we only need to create new rewriters relatively infrequently.
-Then individual rewritings involve reifying terms, asking the kernel to execute the specialized evaluator on them, and simplifying an application of an interpretation function to the result (this last step must be done using Coq's normal reduction, and it is the bottleneck for outputs with enormous numbers of nested binders as discussed in section \ref{sec:micro}).
-
-We would like to emphasize that, while we prototyped our implementation in Coq in particular, the trade-off space that we navigate seems fundamental, so that it should be the case both that our approach can be adapted to other proof assistants and that this case study may inform proof-assistant design.
-The general game here is to stock the trusted proof-checking kernel with as few primitive rules as we can get away with, while still providing enough flexibility and performance.
-Every proof assistant we are aware of has a small functional language at its core, and we argue that is quite natural to include a primitive for efficient full reduction of programs.
-Our empirical result is that such a primitive can form the basis for bootstrapping other kinds of efficient reduction, perhaps suggesting that a future Coq version could fruitfully shrink its kernel by eliminating other built-in reduction strategies.
-
-\subsection{Our Approach in Nine Steps} \label{sec:nine-steps}
+\subsection{Our Approach in Ten Steps}\label{sec:nine-steps}
 
 Here is a bit more detail on the steps that go into applying our Coq plugin, many of which we expand on in the following sections.
-In order to build a precomputed rewriter with the \mintinline{coq}{Make} command, the following actions are performed:
+For \mintinline{coq}{Make} to precompute a rewriter:
 \begin{enumerate}
 \item
-  The given lemma statements are scraped for which named functions and types the rewriter package will support.
+  The given lemma statements are scraped for which named identifiers to encode.
 \item
   Inductive types enumerating all available primitive types and functions are emitted.
+  This allows us to achieve the performance gains attributed in \citet{Boespflug2009} to having native metalanguage constructors for all constants, without manual coding.
 \item
   Tactics generate all of the necessary definitions and prove all of the necessary lemmas for dealing with this particular set of inductive codes.
-  Definitions include operations like Boolean equality on type codes and lemmas like ``all representable primitive types have decidable equality.''
+  Definitions include operations like Boolean equality on type codes and lemmas like ``all types have decidable equality.''
 \item
   The statements of rewrite rules are reified and soundness and syntactic-well-formedness lemmas are proven about each of them.
-  Each instance of the former involves wrapping the user-provided proof with the right adapter to apply to the reified version.
-\item
-  The definitions needed to perform reification and rewriting and the lemmas needed to prove correctness are assembled into a single package that can be passed by name to the rewriting tactic.
+  %Each instance of the former involves wrapping the user-provided proof with the right adapter to apply to the reified version.
+  %Automating this step allows rewrite rules to be proven in terms of their shallow embedding, which drastically accelerates iteration on the set of rewrite rules.
+\item
+  Definitions and lemmas needed to prove correctness are assembled into a single package.
 \end{enumerate}
 
-When we want to rewrite with a rewriter package in a goal, the following steps are performed:
+Then, to rewrite in a goal, the following steps are performed:
 \begin{enumerate}
 \item
-  We rearrange the goal into a single logical formula: all free-variable quantification in the proof context is replaced by changing the equality goal into an equality between two functions (taking the free variables as inputs).
+  Rearrange the goal into a single quantifier-free logical formula.%: free-variable quantification is replaced by equality between functions (taking free variables as inputs).
   %\todo{Pedantic Note From Jason: it's not actually an equality between two functions, it's an \mintinline{coq}{equiv} between two functions where \mintinline{coq}{equiv} is a custom relation we define indexed over type codes that is equality up to function extensionality}
 \item
-  We reify the side of the goal we want to simplify, using the inductive codes in the specified package.
-That side of the goal is then replaced with a call to a denotation function on the reified version.
-\item
-  We use a theorem stating that rewriting preserves denotations of well-formed terms to replace the denotation subterm with the denotation of the rewriter applied to the same reified term.
-  We use Coq's built-in full reduction (\texttt{vm\_compute}) to reduce the application of the rewriter to the reified term.
-\item
-  Finally, we run \texttt{cbv} (a standard call-by-value reducer) to simplify away the invocation of the denotation function on the concrete syntax tree from rewriting.
-  %\todo{Is it worth mentioning that we'd instead use \texttt{vm\_compute} if it supported whitelists (and also said whitelists were recorded in the casts)?}
+  Reify a selected subterm and replace it with a call to our denotation function.
+  %We reify the goal side we want to simplify, using inductive codes from the specified package.
+  %That goal side is replaced with a denotation-function call on the reified version.
+\item
+  Rewrite with a theorem, into a form calling our rewriter.
+\item Call Coq's built-in full reduction (\tacvmcompute{}) to reduce this application.
+\item
+  Run standard call-by-value reduction to simplify away use of the denotation function. %on the concrete syntax tree from rewriting.
+  %\todo{Is it worth mentioning that we'd instead use \tacvmcompute{} if it supported whitelists (and also said whitelists were recorded in the casts)?}
 \end{enumerate}
 
-\section{The Structure of a Rewriter} \label{sec:structure}
-
-We now simultaneously review the approach of \citet{Aehlig} and introduce some notable differences in our own approach, noting similarities to the reflective rewriter of \citet{rtac} where applicable.
-
-First, let us describe the language of terms we support rewriting in.
-Note that, while we support rewriting in full-scale Coq proofs, where the metalanguage is dependently typed, the object language of our rewriter is nearly simply typed, with limited support for calling polymorphic functions.
-However, we still support identifiers whose definitions use dependent types, since our reducer does not need to look into definitions.
+%We now simultaneously review the approach of \citet{Aehlig} and introduce some notable differences in our own approach, noting similarities to the reflective rewriter of \citet{rtac} where applicable.
+
+%Let us now describe the language of terms we support rewriting in.
+%Note that, while we support rewriting in full-scale Coq proofs, where the metalanguage is dependently typed,
+% POPL 2022 review TODO: Line 312: I'm not sure what "nearly simply typed" means.
+The object language of our rewriter is nearly simply typed. %with limited support for calling polymorphic functions.
+%However, we still support identifiers whose definitions use dependent types, since our reducer does not need to look into definitions.
 \begin{align*}
@@
 For example, the identifiers might be codes for $+$, $\cdot$, and literal constants.
-We write $\llbracket e \rrbracket$ for a standard denotational semantics.
-
-\subsection{Pattern-Matching Compilation and Evaluation} \label{sec:pattern-matching-compilation-and-evaluation}
+We write $\llbracket e \rrbracket$ for a standard denotational semantics.\label{sec:denote-brackets-def}
+
+\subsection{Pattern-Matching Compilation and Evaluation}\label{sec:pattern-matching-compilation-and-evaluation}
 
@@
 As we instead build our rewriter in Coq's logic, we have no such option to defer to ML.
-Indeed, Coq's logic only includes primitive pattern-matching constructs to match one constructor at a time.
+%Indeed, Coq's logic only includes primitive pattern-matching constructs to match one constructor at a time.
 %The frontend desugars nested pattern matching into trees of matches, but we do not have access to that desugaring within the language.
-%It is arguably a good design decision to keep Coq working this way, since putting efficient nested pattern matching into the kernel would grow the trusted base substantially.
+%It is arguably a good design decision to keep Coq working this way, since putting efficient nested pattern matching into the kernel would grow the trusted code base substantially.
 
@@
 Instead, we adopted the approach of \citet{maranget2008compiling}, who describes compilation of pattern matches in OCaml to decision trees that eliminate needless repeated work (for example, decomposing an expression into $x + y + z$ only once even if two different rules match on that pattern).
-We have not yet implemented any of the optimizations described therein for finding \emph{minimal} decision trees.
+%We have not yet implemented his optimizations for finding \emph{minimal} decision trees.
 
 There are three steps to turn a set of rewrite rules into a functional program that takes in an expression and reduces according to the rules.
-The first step is pattern-matching compilation: we must compile the lefthand sides of the rewrite rules to a decision tree that describes how and in what order to decompose the expression, as well as describing which rewrite rules to try at which steps of decomposition.
+The first step is pattern-matching compilation: we must compile the left-hand sides of the rewrite rules to a decision tree that describes how and in what order to decompose the expression, as well as describing which rewrite rules to try at which steps of decomposition.
 Because the decision tree is merely a decomposition hint, we require no proofs about it to ensure soundness of our rewriter.
@@
 Here the correctness condition is that we must not change the semantics of the expression.
-Said another way, any rewrite-rule replacement expression must match the semantics of the rewrite-rule pattern.
+%Said another way, any rewrite-rule replacement expression must match the semantics of the rewrite-rule pattern.
 
@@
   \item \texttt{Switch icases app\_case default}:
-    With the first element of the vector, match on its kind; if it is an identifier matching something in \texttt{icases}, which is a list of pairs of identifiers and decision trees, remove the first element of the vector and run that decision tree; if it is an application and \texttt{app\_case} is not \texttt{None}, try the \texttt{app\_case} decision tree, replacing the first element of each vector with the two elements of the function and the argument it is applied to; otherwise, do not modify the vectors and use the \texttt{default} decision tree.
+    With the first element of the vector, match on its kind; if it is an identifier matching something in \texttt{icases}, which is a list of pairs of identifiers and decision trees, remove the first element of the vector and run that decision tree; if it is an application and \texttt{app\_case} is not \texttt{None}, try the \texttt{app\_case} decision tree, replacing the first element of each vector with the two elements of the function and the argument it is applied to; otherwise, do not modify the vectors and use the \texttt{default}.
   \item \texttt{Swap i cont}: Swap the first element of the vector with the $i^\texttt{th}$ element (0-indexed) and keep going with \texttt{cont}.
@@
 \end{verbatim}
-The decision tree produced is \label{sec:compiled-pattern}
+The decision tree produced is\label{sec:compiled-pattern}
 \[\resizebox{230px}{!}{\xymatrix@R-1pc{
@@
 }}\]
-\noindent where every non-swap node implicitly has a ``default'' case arrow to \texttt{Failure} and circles represent \texttt{Switch} nodes.
+\noindent where every nonswap node implicitly has a ``default'' case arrow to \texttt{Failure} and circles represent \texttt{Switch} nodes.
 
@@
 Note that we use Coq's normal partial evaluation to turn our general decision-tree evaluator into a specialized matcher to get reasonable efficiency.
-Although this partial evaluation of our partial evaluator is subject to the same performance challenges we highlighted in the introduction, it only has to be done once for each set of rewrite rules, and we are targeting cases where the time of per-goal reduction dominates this time of meta-compilation.
+Although this partial evaluation of our partial evaluator is subject to the same performance challenges we highlighted in the introduction, it only has to be done once for each set of rewrite rules, and we are targeting cases where the time of per-goal reduction dominates this time of metacompilation.
 
@@
 
-\subsection{Adding Higher-Order Features} \label{sec:thunk-eval-subst-term}
+\subsection{Adding Higher-Order Features}\label{sec:thunk-eval-subst-term}
 
@@
 
-\citet{Aehlig} also use NbE, and we begin by reviewing its most classic variant, for performing full $\beta$-reduction in a simply typed term in a guaranteed-terminating way.
-The simply typed $\lambda$-calculus syntax we use is:
+We begin by reviewing NbE's most classic variant, for performing full $\beta$-reduction in a simply typed term in a guaranteed-terminating way.
+Our simply typed $\lambda$-calculus syntax is:
 \begin{align*}
@@
 We can now define normalization by evaluation.
-First, we choose a ``semantic'' representation for each syntactic type, which serves as the result type of an intermediate interpreter.
+First, we choose a ``semantic'' representation for each syntactic type, which serves as an interpreter's result type.
 \begin{align*}
-  \text{NbE}_t(t_1 \to t_2) & \defeq \text{NbE}_t(t_1) \to \text{NbE}_t(t_2) \\
-  \text{NbE}_t(b) & \defeq \texttt{expr}(b)
+  \text{NbE}_t(t_1 \to t_2) & \defeq \text{NbE}_t(t_1) \to \text{NbE}_t(t_2)
+  & \text{NbE}_t(b) & \defeq \texttt{expr}(b)
 \end{align*}
@@
 \begin{figure}%[b]
+  \begin{subfigure}[b]{0.5\textwidth}
     \begin{align*}
@@
       \text{reflect}_{b}(e) & \defeq e \\ \noalign{\vskip7pt}
+    \end{align*}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.5\textwidth}
+    \begin{align*}
       \text{reduce} & : \text{expr}(t) \to \text{NbE}_t(t) \\
@@
     \end{align*}
+  \end{subfigure}
 \caption{\label{fig:nbe}Implementation of normalization by evaluation}
@@
 We split out analysis of term syntax in a separate function reduce, defined by primitive recursion on term syntax, when usually this functionality would be mixed in with reflect.
-The reason for this choice will become clear when we extend NbE to handle our full problem domain.
+The reason for this choice will become clear when we extend NbE.
 
@@
 To fuse NbE with rewriting, we only modify the constant case of \texttt{reduce}.
-First, we bind our specialized decision-tree engine under the name rewrite-head.
-Recall that this function only tries to apply rewrite rules at the top level of its input.
+% POPL 2022 review TODO: line 433, "First, [...]" but there is no "Second, [...]"?
+First, we bind our specialized decision-tree engine (which rewrites \emph{at the root of an AST only}) under the name rewrite-head.
 
@@
 In other words, we change this one function-definition clause:
+% POPL 2022 review TODO:
+% line 440, is this an amendment to the code in Figure 1?
+% It would be clearer to have just one version of the code.
 \begin{align*}
@@
 It is important to note that a constant of function type will be $\eta$-expanded only once for each syntactic occurrence in the starting term, though the expanded function is effectively a thunk, waiting to perform rewriting again each time it is called.
-From first principles, it is not clear why such a strategy terminates on all possible input terms, though we work up to convincing Coq of that fact.
+From first principles, it is not clear why such a strategy terminates on all possible input terms.
+%, though we work up to convincing Coq of that fact.
 
@@
 Recall that their rewriter was implemented in a deeply embedded ML, while ours is implemented in Coq's logic, which enforces termination of all functions.
-Aehlig et al.
-did not prove termination, which indeed does not hold for their rewriter in general, which works with untyped terms, not to mention the possibility of rule-specific ML functions that diverge themselves.
+Aehlig et al.\ did not prove termination, which indeed does not hold for their rewriter in general, which works with untyped terms, not to mention the possibility of divergent rule-specific ML functions.
 In contrast, we need to convince Coq up-front that our interleaved $\lambda$-term normalization and algebraic simplification always terminate.
-Additionally, we need to prove that our rewriter preserves denotations of terms, which can easily devolve into tedious binder bookkeeping, depending on encoding.
+Additionally, we must prove that rewriting preserves term denotations, which can easily devolve into tedious binder bookkeeping.
+%, depending on encoding.
 
@@
 
-\section{Scaling Challenges} \label{sec:scaling}
+\section{Scaling Challenges}\label{sec:scaling}
 
@@
 
-\subsection{Variable Environments Will Be Large} \label{sec:PHOAS}
+\subsection{Variable Environments Will Be Large}\label{sec:PHOAS}
 We should think carefully about representation of ASTs, since many primitive operations on variables will run in the course of a single partial evaluation.
@@
 
-Our strategy is to use a variable encoding that pushes all first-order bookkeeping off on Coq's kernel, which is itself performance-tuned with some crucial pieces of imperative code.
+Our strategy is to use a variable encoding that pushes all first-order bookkeeping off on Coq's kernel or the implementation of the language we extract to, which are themselves performance-tuned with some crucial pieces of imperative code.
 Parametric higher-order abstract syntax (PHOAS)~\cite{PhoasICFP08} is a dependently typed encoding of syntax where binders are managed by the enclosing type system.
@@
 
+% POPL 2022 review TODO:
+% line 504, "Here is the actual inductive definition".
+% Could this definition come earlier?
+% The definition of reify & reflect would be clearer if it did.
 Here is the actual inductive definition of term syntax for our object language, PHOAS-style.
@@
 | App {s d} (f : expr var (s → d)) (x : expr var s) : expr var d
-  | Const {t} (c : const t) : expr var t
+| LetIn {a b} (x : expr var a) (f : var a -> expr var b) : expr var b
+| Const {t} (c : const t) : expr var t.
 Definition Expr (t : type) : Type := forall var, expr var t.
@@
 
+% POPL 2022 review TODO:
+% line 517, "A good example of encoding adequacy".
+% I believe I know what "adequacy" means in this context, but I don't think every reader will.
 A good example of encoding adequacy is assigning a simple denotational semantics.
@@
 \begin{minted}[fontsize=\small]{coq}
-Fixpoint denoteT (t : type) : Type
-  := match t with
+Fixpoint denoteT (t : type) : Type := match t with
   | arrow s d => denoteT s -> denoteT d
-     | base b    => denote_base_type b
-     end.
+  | base b    => denote_base_type b  end.
 \end{minted}
@@
 Next we see the convenience of being able to \emph{use} an expression by choosing how it should represent variables.
-Specifically, it is natural to choose \emph{the type-denotation function itself} as the variable representation.
-Especially note how this choice makes rigorous the convention we followed in the prior section (e.g., in the suspicious function-abstraction clause of function reduce), where a recursive function enforces that values have always been substituted for variables early enough.
-\begin{minted}[fontsize=\small]{coq}
-Fixpoint denoteE {t} (e : expr denoteT t) : denoteT t
-  := match e with
+Specifically, it is natural to choose \emph{the type-denotation function itself} as variable representation.
+% POPL 2022 review TODO:
+% line 528, "the suspicious function-abstraction clause".
+% I believe I would prefer to see correct code up front.
+Especially note how this choice makes rigorous last section's convention (e.g., in the suspicious function-abstraction clause of reduce), where a recursive function enforces that values have always been substituted for variables early enough.
+\begin{minted}[fontsize=\small]{coq}
+Fixpoint denoteE {t} (e : expr denoteT t) : denoteT t := match e with
   | Var v     => v
@@
   | App f x   => (denoteE f) (denoteE x)
-     | Ident c   => denoteI c
-     end.
-Definition DenoteE {t} (E : Expr t) : denoteT t
-  := denoteE (E denoteT).
-\end{minted}
-
-It is now easy to follow the same script in making our rewriting-enabled NbE fully formal.
+  | LetIn x f => let xv := denoteE x in denoteE f xv
+  | Ident c   => denoteI c  end.
+Definition DenoteE {t} (E : Expr t) : denoteT t := denoteE (E denoteT).
+\end{minted}
+
+It is now easy to follow the same script in making our rewriting-enabled NbE fully formal, in \autoref{fig:nbe2}.
 Note especially the first clause of \texttt{reduce}, where we avoid variable substitution precisely because we have chosen to represent variables with normalized semantic values.
@@
 The final definition $\lambda$-quantifies over that choice.
-\begin{minted}[fontsize=\small]{coq}
-Fixpoint nbeT var (t : type) : Type
-  := match t with
+
+\begin{figure}
+  \begin{subfigure}[b]{0.5\textwidth}
+    \begin{minted}[fontsize=\small]{coq}
+Fixpoint nbeT var (t : type) : Type :=
+match t with
 | arrow s d => nbeT var s -> nbeT var d
@@
 end.
-Fixpoint reify {var t} : nbeT var t -> expr var t
+
+Fixpoint reify {var t}
+  : nbeT var t -> expr var t :=
+match t with
+| arrow s d => λ f, Abs (λ x,
+    reify (f (reflect (Var x))))
+| base b    => λ e, e            end
+    \end{minted}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.5\textwidth}
+    \begin{minted}[fontsize=\small]{coq}
+with reflect{var t}:expr var t->nbeT var t
   := match t with
-     | arrow s d => λ f, Abs (λ x, reify (f (reflect (Var x))))
-     | base b    => λ e, e
-     end
-with reflect {var t} : expr var t -> nbeT var t
-  := match t with
-     | arrow s d => λ e, λ x, reflect (App e (reify x))
-     | base b    => rewrite_head
-     end.
-Fixpoint reduce {var t}
-  (e : expr (nbeT var) t) : nbeT var t
-  := match e with
+| arrow s d => λ e, λ x,
+  reflect (App e (reify x))
+| base b    => rewrite_head      end.
+Fixpoint reduce{var t}(e:expr (nbeT var) t)
+  : nbeT var t := match e with
 | Abs e     => λ x, reduce (e (Var x))
@@
 | Var x     => x
-     | Ident c   => reflect (Ident c)
-     end.
-Definition Rewrite {t} (E : Expr t) : Expr t
+| Ident c   => reflect (Ident c) end.
+Definition Rewrite {t} (E:Expr t) : Expr t
   := λ var, reify (reduce (E (nbeT var t))).
     \end{minted}
-
-One subtlety hidden above in implicit arguments is in the final clause of \texttt{reduce}, where the two applications of the \texttt{Ident} constructor use different variable representations.
+  \end{subfigure}
+  \caption{\label{fig:nbe2}PHOAS implementation of normalization by evaluation}
+\end{figure}
+
+One subtlety hidden in \autoref{fig:nbe2} in implicit arguments is in the final clause of \texttt{reduce}, where the two applications of the \texttt{Ident} constructor use different variable representations.
 With all those details hashed out, we can prove a pleasingly simple correctness theorem, with a lemma for each main definition, with inductive structure mirroring recursive structure of the definition, also appealing to correctness of last section's pattern-compilation operations.
+(We now use syntax $\llbracket \cdot \rrbracket$ for calls to \texttt{DenoteE}.)
 $$\forall t, E : \mintinline{coq}{Expr t}. \; \llbracket \mintinline{coq}{Rewrite}(E) \rrbracket = \llbracket E \rrbracket$$
 
-Even before getting to the correctness theorem, we needed to convince Coq that the function terminates.
-While for \citet{Aehlig}, a termination proof would have been a whole separate enterprise, it turns out that PHOAS and NbE line up so well that Coq accepts the above code with no additional termination proof.
-As a result, the Coq kernel is ready to run our \mintinline{coq}{Rewrite} procedure during checking.
+% POPL 2022 review TODO:
+% line 577, "we needed to convince Coq that the function terminates".
+% Coq is convinced already, if it accepts your definition.
+% The phrasing is somewhat odd.
+%Even before getting to the correctness theorem, we needed to convince Coq that the function terminates.
+%While for \citet{Aehlig}, a termination proof would have been a whole separate enterprise, it turns out that PHOAS and NbE line up so well that Coq accepts the above code with no additional termination proof; each key function is obviously structurally recursive on either a type or an expression.
+% POPL 2022 review TODO:
+% line 581, "the Coq kernel is ready to run our Rewrite procedure during checking."
+% During checking of what?
+% If you mean that the Coq type-checker performs reduction during type-checking, that is a well-known fact which perhaps does not need to be recalled here.
+%As a result, the Coq kernel is ready to run our \mintinline{coq}{Rewrite} procedure during checking.
 
@@
 Now we use the \mintinline{coq}{Rewrite} correctness theorem to replace $\llbracket E \rrbracket$ with $\llbracket \mintinline{coq}{Rewrite}(E) \rrbracket$.
-Next we may ask the Coq kernel to simplify $\mintinline{coq}{Rewrite}(E)$ by \emph{full reduction via compilation to native code}, since we carefully designed $\mintinline{coq}{Rewrite}(E)$ and its dependencies to produce closed syntax trees, so that reduction will not get stuck pattern-matching on free variables.
-Finally, where $E'$ is the result of that reduction, we simplify $\llbracket E' \rrbracket$ with standard reduction, producing a normal-looking Coq term.
+Next we ask the Coq kernel to simplify $\mintinline{coq}{Rewrite}(E)$ by \emph{full reduction via native compilation}.
+%since we carefully designed $\mintinline{coq}{Rewrite}(E)$ and its dependencies to produce closed syntax trees, so that reduction will not get stuck pattern-matching on free variables.
+Finally, where $E'$ is the result of that reduction, we simplify $\llbracket E' \rrbracket$ with standard reduction.
+
+We have been discussing representation of bound variables.
+Also important is representation of constants (e.g., library functions mentioned in rewrite rules).
+They could also be given some explicit first-order encoding, but dispatching on, say, strings or numbers for constants would be rather inefficient in our generated code.
+Instead, we chose to have our Coq plugin generate a custom inductive type of constant codes, for each rewriter that we ask it to build with \texttt{Make}.
+As a result, dispatching on a constant can happen in constant time, based on whatever pattern-matching is built into the execution language (either the Coq kernel or the target language of extraction).
+To our knowledge, no past verified reduction tool in a proof assistant has employed that optimization.
 
@@
 
-\subsection{Subterm Sharing is Crucial} \label{sec:under-lets}
+\subsection{Subterm Sharing Is Crucial}\label{sec:under-lets}
 
@@
 The library has been converted manually to continuation-passing style, allowing proper generation of \mintinline{coq}{let} binders, whose variables are often mentioned multiple times.
-We ran their code generator (actually just a subset of its functionality, but optimized by us a bit further, as explained in \autoref{sec:macro}) on the P-256 example and found it took about 15 seconds to finish.
-Then we modified reduction to inline \mintinline{coq}{let} binders instead of preserving them, at which point the reduction job terminated with an out-of-memory error, on a machine with 64 GB of RAM.
-(The successful run uses under 2 GB.)
+% POPL 2022 review(ish) TODO: actually it's no longer explained further in \autoref{sec:macro}
+We ran that old code generator (actually just a subset of its functionality, but optimized by us a bit further, as explained in \autoref{sec:macro}) on the P-256 example and found it took about 15 seconds to finish.
+Then we modified reduction to inline \mintinline{coq}{let} binders instead of preserving them, at which point the job terminated with an out-of-memory error, on a machine with 64 GB of RAM.
+%(The successful run uses under 2 GB.)
+% POPL 2022 review TODO (maybe just include this point):
+% Even if the compiler had been able to terminate, the resulting "optimized" code would be practically unusable, as it would perform many redundant computations.
 
 We see a tension here between performance and niceness of library implementation.
-The Fiat Cryptography authors found it necessary to CPS-convert their code to coax Coq into adequate reduction performance.
-Then all of their correctness theorems were complicated by reasoning about continuations.
+When we built the original Fiat Cryptography, we found it necessary to CPS-convert the code to coax Coq into adequate reduction performance.
+Then all of our correctness theorems were complicated by reasoning about continuations.
+In fact, the CPS reasoning was so painful that at one point most algorithms in the template library were defined twice, once in continuation-passing style and once in direct-style code, because it was easier to prove the two equivalent and work with the non-CPS version than to reason about the CPS version directly.
 It feels like a slippery slope on the path to implementing a domain-specific compiler, rather than taking advantage of the pleasing simplicity of partial evaluation on natural functional programs.
@@
 \begin{align*}
-  \texttt{map}\ (\lambda x.\, y+x)\ (\letin[{z:=e}{[0;1;2;z;z+1]}]) \\
-  \leadsto \; \letin[{z:=e}{[y;y+1;y+2;y+z;y+(z+1)]}]
+  \texttt{map}\ (\lambda x.\, y+x)\ (\letin[{z:=e}{[0;1;z+1]}]) & %\\
+  \; \leadsto \;
+  \letin[{z:=e}{[y;y+1;y+(z+1)]}]
 \end{align*}
@@
 
-Our approach is to define a telescope-style type family called \mintinline{coq}{UnderLets}:
-\begin{minted}[fontsize=\small]{coq}
-Inductive UnderLets {var} (T : Type) :=
-| Base (v : T)
-| UnderLet {A}(e : @expr var A)(f : var A -> UnderLets T).
+We define a telescope-style type family called \mintinline{coq}{UnderLets}:
+\begin{minted}[fontsize=\small]{coq}
+Inductive UnderLets {var} (T : Type) := Base (v : T)
+| UnderLet {A} (e : @expr var A) (f : var A -> UnderLets T).
 \end{minted}
 A value of type \mintinline{coq}{UnderLets T} is a series of \texttt{let} binders (where each expression \mintinline{coq}{e} may mention earlier-bound variables) ending in a value of type \mintinline{coq}{T}.
-It is easy to build various ``smart constructors'' working with this type, for instance to construct a function application by lifting the \texttt{let}s of both function and argument to a common top level.
-
-Such constructors are used to implement an NbE strategy that outputs \mintinline{coq}{UnderLets} telescopes.
+%It is easy to build various ``smart constructors'' working with this type, for instance to construct a function application by lifting the \texttt{let}s of both function and argument to a common top level.
+
+%Such constructors are used to implement an NbE strategy that outputs \mintinline{coq}{UnderLets} telescopes.
 Recall that the NbE type interpretation mapped base types to expression syntax trees.
-We now parameterize that type interpretation by a Boolean declaring whether we want to introduce telescopes.
-
-\begin{minted}[fontsize=\small]{coq}
-Fixpoint nbeT' {var} (with_lets : bool) (t : type)
-  := match t with
+We add flexibility, parameterizing by a Boolean declaring whether to introduce telescopes.
+
+\begin{minted}[fontsize=\small]{coq}
+Fixpoint nbeT' {var} (with_lets : bool) (t : type) := match t with
   | base t => if with_lets then @UnderLets var (@expr var t) else @expr var t
-     | arrow s d => nbeT' false s -> nbeT' true d
-     end.
-Definition nbeT := nbeT' false.
-Definition nbeT_with_lets := nbeT' true.
+  | arrow s d => nbeT' false s -> nbeT' true d  end.
+Definition nbeT := nbeT' false.      Definition nbeT_with_lets := nbeT' true.
 \end{minted}
@@
 
-\subsection{Rules Need Side Conditions} \label{sec:side-conditions}
+\subsection{Rules Need Side Conditions}\label{sec:side-conditions}
 
 Many useful algebraic simplifications require side conditions.
-One simple case is supporting \emph{nonlinear} patterns, where a pattern variable appears multiple times.
-We can encode nonlinearity on top of linear patterns via side conditions.
+For example, bit-shifting operations are faster than divisions, so we might want a rule such as
+%One simple case is supporting \emph{nonlinear} patterns, where a pattern variable appears multiple times.
+%We can encode nonlinearity on top of linear patterns via side conditions.
 \begin{align*}
-  {?n_1} + {?m} - {?n_2} & \to m\text{\ \ if\ \ }n_1 = n_2
+  {?n} / {?m} & \to n \gg \log_2 m\text{\quad if\quad }2^{\lfloor \log_2 m \rfloor} = m
 \end{align*}
@@
 We decided to sidestep this problem by allowing side conditions only as executable Boolean functions, to be applied only to variables that are confirmed as \emph{compile-time constants}, unlike \citet{rtac} who support general unification variables.
+% POPL 2022 review TODO:
+% line 672, "We added a variant of pattern variable [...]".
+% This passage is rather obscure.
 We added a variant of pattern variable that only matches constants.
-Semantically, this variable style has no additional meaning, and in fact we implement it as a special identity function that should be called in the right places within Coq lemma statements.
+Semantically, this variable style has no additional meaning, and in fact we implement it as a special identity function (notated as an apostrophe) that should be called in the right places within Coq lemma statements.
 Rather, use of this identity function triggers the right behavior in our tactic code that reifies lemma statements.
-We introduce a notation where a prefixed apostrophe signals a call to the ``constants only'' function.
-
-Our reification inspects the hypotheses of lemma statements, using type classes to find decidable realizations of the predicates that are used, synthesizing one Boolean expression of our deeply embedded term language, standing for a decision procedure for the hypotheses.
+%We introduce a notation where a prefixed apostrophe signals a call to the ``constants only'' function.
+\label{sec:explain-'}
+%We use a special apostrophe marker to indicate a quantified variable that may only match with \emph{compile-time constants}.
+
+Our reification inspects the hypotheses of lemma statements, using type classes to find decidable realizations of the predicates that are used, thereby synthesizing one Boolean expression of our deeply embedded term language, which stands for a decision procedure for the hypotheses.
 The \mintinline{coq}{Make} command fails if any such expression contains pattern variables not marked as constants.
-Therefore, matching of rules can safely run side conditions, knowing that Coq's full-reduction engine can determine their truth efficiently.
-
-\subsection{Side Conditions Need Abstract Interpretation} \label{sec:abs-int}
-
-With our limitation that side conditions are decided by executable Boolean procedures, we cannot yet handle directly some of the rewrites needed for realistic partial evaluation.
+%Therefore, matching of rules can safely run side conditions, knowing that Coq's full-reduction engine can determine their truth efficiently.
+
+Hence, we encode the above rule as $\forall n, m. \; 2^{\lfloor \log_2(\texttt{'}m)\rfloor} = \texttt{'}m \to n / \texttt{'}m = n \gg \texttt{'}(\log_2 m)$.
+
+\subsection{Side Conditions Need Abstract Interpretation}\label{sec:abs-int}
+
+With our limitation that side conditions are decided by executable Boolean procedures, we cannot yet handle directly some of the rewrites needed for realistic compilation.
 For instance, Fiat Cryptography reduces high-level functional to low-level code that only uses integer types available on the target hardware.
@@
 
+% POPL 2022 review TODO:
+% I believe the authors could perhaps better explain why the rewriting rule at line 706 is preferable (somehow more tractable) than the rewriting rule at line 696.
+% At first sight, they seem similar.
+% I guess the point is that the variable `n` is unknown (i.e., *not* a compile-time constant) whereas the variable `u` is known (a compile-time constant), so the second rule falls within the scope of the technology described in Section 4.3, whereas the first rule does not.
 Second, we proceed with our example rule refactored:
 \begin{align*}
-  \texttt{add\_with\_carry}_{64}(\texttt{clip}_{'{?l},'{?u}}({?n}), 0) & \to (0, \texttt{clip}_{l,u}(n)) %\\
+  \texttt{add\_with\_carry}_{64}(\texttt{clip}_{\texttt{'}{?l},\texttt{'}{?u}}({?n}), 0) & \to (0, \texttt{clip}_{l,u}(n)) %\\
   %&
@@
 
-\subsection{Limitations and Preprocessing} \label{sec:implementation-and-usage}
-
-%\todo{It's not clear to Jason what in this section belongs here, and what isn't interesting (yet?) to readers}
-%\todo{I (Jason) have no idea what the story for this section should be}
-%\todo{Insofar as text is actually kept in this section, make sure any duplication between here and the out-of-paper appendix is deliberate}
-%\todo{If this section is moved or removed, be sure to restructure \autoref{sec:code-from-implementation-and-usage} accordingly}
-
-We now note some details of the rewriting framework that were previously glossed over, which are useful for using the code or implementing something similar, but which do not add fundamental capabilities to the approach.
-Although the rewriting framework does not support dependently typed constants, we can automatically preprocess uses of eliminators like \mintinline{coq}{nat_rect} and \mintinline{coq}{list_rect} into non-dependent versions.
-The tactic that does this preprocessing is extensible via \Ltac{}'s reassignment feature.
-Since pattern-matching compilation mixed with NbE requires knowing how many arguments a constant can be applied to, internally we must use a version of the recursion principle whose type arguments do not contain arrows; current preprocessing can handle recursion principles with either no arrows or one arrow in the motive.
-
-Recall from \autoref{sec:explain-eval-rect} that \mintinline{coq}{eval_rect} is a definition provided by our framework for eagerly evaluating recursion associated with certain types.
-It functions by triggering typeclass resolution for the lemmas reducing the recursion principle associated to the given type.
-We provide instances for \texttt{nat}, \texttt{prod}, \texttt{list}, \texttt{option}, and \texttt{bool}.
-Users may add more instances if they desire.
-
-Recall again from \autoref{sec:explain-ident.eagerly} that we use \mintinline{coq}{ident.eagerly} to ask the reducer to simplify a case of primitive recursion by complete traversal of the designated argument's constructor tree.
-Our current version only allows a limited, hard-coded set of eliminators with \mintinline{coq}{ident.eagerly} (\texttt{nat\_rect} on return types with either zero or one arrows, \texttt{list\_rect} on return types with either zero or one arrows, and \texttt{List.nth\_default}), but nothing in principle prevents automatic generation of the necessary code.
-
-Note that \mintinline{coq}{Let_In} is the constant we use for writing \letin{} expressions that do not reduce under $\zeta$ (Coq's reduction rule for \mintinline{coq}{let}-inlining).
-Throughout most of this paper, anywhere that \letin{} appears, we have actually used \mintinline{coq}{Let_In} in the code.
-It would alternatively be possible to extend the reification preprocessor to automatically convert \letin{} to \mintinline{coq}{Let_In}, but this strategy may cause problems when converting the interpretation of the reified term with the pre-reified term, as Coq's conversion does not allow fine-tuning of when to inline or unfold \mintinline{coq}{let}s.
+See \autoref{sec:implementation-and-usage}\ifthenelse{\boolean{tobesplit}}{%
+  \footnote{%
+  The appendix of this paper is available online at \url{https://jasongross.github.io/papers/2022-rewriting-itp-supplement.pdf}.%
+  }%
+}{}
+for discussion of some further twists in the implementation.
 
@@
 % TODO: Presumably ``attached to this submission as an anonymized supplement'' will change for the final submission
-Our implementation, attached to this submission as an anonymized supplement with a roadmap in \autoref{sec:CodeSupplement-more}, includes a mix of Coq code for the proved core of rewriting, tactic code for setting up proper use of that core, and OCaml plugin code for the manipulations beyond the current capabilities of the tactic language.
-We report here on experiments to isolate performance benefits for rewriting under binders and reducing higher-order structure.
-
-\subsection{Microbenchmarks} \label{sec:micro}
+Our implementation, available on GitHub at \githublink{mit-plv}{rewriter}[ITP-2022-perf-data] and with a roadmap in \autoref{sec:CodeSupplement-more}, includes a mix of Coq code for the proved core of rewriting, tactic code for setting up proper use of that core, and OCaml plugin code for the manipulations beyond the tactic language's current capabilities.
+We report here on evidence that the tool is effective, first in terms of productivity by users and then in terms of compile-time performance.
+
+\subsection{Iteration on the Fiat Cryptography Compiler}\label{sec:iteration}
+
+We ported Fiat Cryptography's core compiler functionality to use our framework.
+The result is now used in production by a number of open-source projects.
+We were glad to retire the CPS versions of verified arithmetic functions, which had been present only to support predictable reduction with subterm sharing.
+More importantly, it became easy to experiment with new transformations via proving new rewrite theorems, directly in normal Coq syntax, including the following, all justified by demand from real users:
+\begin{itemize}
+\item Reassociating arithmetic to minimize the bitwidths of intermediate results
+\item Multiplication primitives that separately return high halves and low halves
+\item Strings and a ``comment'' function of type $\forall A. \; \texttt{string} \to A \to A$
+\item Support for bitwise exclusive-or
+\item A special marker to block C compilers from introducing conditional jumps in code that should be constant-time
+\item Eliding bitmask-with-constant operations that can be proved as no-ops
+\item Rules to introduce conditional moves (on supported platforms)
+\item New hardware backend, via rules that invoke special instructions of a cryptographic accelerator
+\item New hardware backend, with a requirement that all intermediate integers have the same bitwidth, via rules to break wider operations down into several narrower operations
+\end{itemize}
+
+% This is an odd one to include in my summarized list.
+% It sounds like more of an observation on the software-engineering experience,
+% rather than extending fiat-crypto.  Maybe BRING BACK if we get more space.
+%% \subsubsection{Moving Rules Involving Carries}\label{sec:iteration:carries}
+%% We originally had rules like ``adding 0 to $x$ produces just $x$, with no carry.''
+%% This rule is not true in general, because we must encode carrying as happening at a particular bitwidth, while we have no guarantee that $x$ fits within that bitwidth.
+%% The correct rule has a precondition: that $x$ is between 0 and $2^{64}$.
+
+%% We discovered this issue when trying to prove our rewrite rules correct.
+%% As a result, we had to move this sort of rule from happening before abstract interpretation to happening after abstract interpretation.
+%% Since the passes are just defined as lists of rewrite rules, moving the rules was quite simple.
+%(Rephrasing them to talk about the outputs of abstract interpretation was somewhat painful, though, because the proof assistant did not enforce the conventions we were using for where to store abstract-interpretation information.
+%We hypothesize that a more uniform approach to integrating abstract interpretation with rewriting and partial evaluation would solve this problem.)
+
+%% BRING BACK \label{sec:fusing-compiler-passes} subsubsection from appendix when we have plenty of space.
+
+\subsection{Microbenchmarks}\label{sec:micro}
 
 \def\NoBindersSubfloatNval{3}%
-\def\NoBindersSubfloatXRow{\thisrowno{0}*(2^(\nval+1)-1)}%
-
-We start with microbenchmarks focusing attention on particular aspects of reduction and rewriting, with \autoref{sec:additionalMicro} going into more detail.
-
-\subsubsection{Rewriting Without Binders} \label{sec:micro:Plus0Tree}
-
-Consider the code defined by the expression $\text{tree}_{n,m}(v)$ in \autoref{fig:micro:Plus0Tree:code}.
-%\todo{make sure the figure spacing is okay}
-%RSolve[{t[0]==m, t[n]==2*t[n-1] + m},t[n],n]
-%t[n]=m(2^{n+1}-1)
-We want to remove all of the${}+0$s.
-There are $\Theta(m \cdot 2^n)$ such rewriting locations.
-We can start from this expression directly, in which case reification alone takes as much time as Coq's \texttt{rewrite}.
-\begin{wrapfigure}[7]{r}{0pt}
-{\small $\begin{aligned}
-\text{iter}_m(v) & = v + \underbrace{0 + 0 + \cdots + 0}_m \\
-\text{tree}_{0,m}(v) &= \text{iter}_m(v + v) \\
-\text{tree}_{n+1,m}(v) &= \text{iter}_m(\text{tree}_{n,m}(v) + \text{tree}_{n,m}(v))
-\end{aligned}$}%
-\caption{\label{fig:micro:Plus0Tree:code}Expressions computing initial code}
-\end{wrapfigure}
-As the reification method was not especially optimized, and there exist fast reification methods~\cite{ReificationITP18}, we instead start from a call to a recursive function that generates such an expression.
-
-\begin{figure*}[b]
+\def\NoBindersSubfloatXRow{\thisrow{param-2-m}*(2^(\thisrow{param-1-n}+1)-1)}%
+
+\begin{figure*}
   \newsavebox{\NestedBindersSubfloat}%
   \sbox{\NestedBindersSubfloat}{%
-    \adjustbox{valign=t}{\resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
-      \einput{perf-UnderLetsPlus0-cbv;rewrite-strat(bottomup).txt}
-      \einput{perf-UnderLetsPlus0-cbv;rewrite-strat(topdown).txt}
-      \einput{perf-UnderLetsPlus0-cbv;setoid-rewrite.txt}
-      \einput{perf-UnderLetsPlus0-Rewrite-for.txt}
-      \einput{perf-UnderLetsPlus0-rewriting.txt}
+    \adjustbox{valign=t}{\resizebox{0.32\textwidth}{!}{\beginTikzpictureStamped[only marks]{
+      \einput{perf-UnderLetsPlus0.csv.md5}
     }
@@
           anchor=north}}
-      \begin{axis}[xlabel=\# of let binders,ylabel=time (s),ymax=65]
-        \addplot[mark=o,color=red] table{perf-UnderLetsPlus0-cbv;rewrite-strat(bottomup).txt};
-        \addplot[mark=triangle,color=red] table{perf-UnderLetsPlus0-cbv;rewrite-strat(topdown).txt};
-        \addplot[mark=square,color=red] table{perf-UnderLetsPlus0-cbv;setoid-rewrite.txt};
-        \addplot[mark=+,color=blue] table{perf-UnderLetsPlus0-Rewrite-for.txt};
-        \addplot[mark=x,color=ForestGreen] table{perf-UnderLetsPlus0-rewriting.txt};
+      \begin{axis}[xlabel=\# of let binders,
+          ylabel=time (s),
+          scaled x ticks=false,
+          ymax=65,
+          xmax=5000,
+          table/col sep=comma,
+          table/x=param-n
+        ]
+        \addplot[mark=o,color=red]         table[y=rewrite-strat(bottomup)-regression-exponential-user]{perf-UnderLetsPlus0.csv};
+        \addplot[mark=triangle,color=red]  table[y=rewrite-strat(topdown)-regression-exponential-user] {perf-UnderLetsPlus0.csv};
+        \addplot[mark=square,color=red]    table[y=setoid-rewrite-regression-cubic-user]               {perf-UnderLetsPlus0.csv};
+        \addplot[mark=+,color=blue]        table[y=Rewrite-for-gen-user]                               {perf-UnderLetsPlus0.csv};
+        \addplot[mark=x,color=ForestGreen] table[y=rewriting-user]                                     {perf-UnderLetsPlus0.csv};
         \legend{rewrite\_strat bottomup,rewrite\_strat topdown,setoid\_rewrite,{Our approach including reification, cbv, etc.},Our approach (only rewriting)}
@@
   \sbox{\BindersAndRecursiveFunctionsSubfloat}{%
-    \adjustbox{valign=t}{\resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
-        \einput{perf-LiftLetsMap-rewrite-strat(bottomup,bottomup).txt}
-        \einput{perf-LiftLetsMap-rewrite-strat(topdown,bottomup).txt}
-        \einput{perf-LiftLetsMap-setoid-rewrite.txt}
-        \einput{perf-LiftLetsMap-Rewrite-for.txt}
-        \einput{perf-LiftLetsMap-rewriting.txt}
-        \einput{perf-LiftLetsMap-cps+vm-compute.txt}
+    \adjustbox{valign=t}{\resizebox{0.32\textwidth}{!}{\beginTikzpictureStamped[only marks]{
+        \einput{perf-UnderLetsPlus0.csv.md5}
     }
@@
           anchor=north}}
-      \begin{axis}[xlabel=$n\cdot m$,ylabel=time (s),scaled x ticks=false,ymax=27]
-        \addplot[mark=o,color=red] table{perf-LiftLetsMap-rewrite-strat(bottomup,bottomup).txt};
-        \addplot[mark=triangle,color=red] table{perf-LiftLetsMap-rewrite-strat(topdown,bottomup).txt};
-        \addplot[mark=square,color=red] table{perf-LiftLetsMap-setoid-rewrite.txt};
-        \addplot[mark=+,color=blue] table{perf-LiftLetsMap-Rewrite-for.txt};
-        \addplot[mark=x,color=ForestGreen] table{perf-LiftLetsMap-rewriting.txt};
-        \addplot[mark=*,color=purple] table{perf-LiftLetsMap-cps+vm-compute.txt};
+      \begin{axis}[xlabel=$n\cdot m$,
+          ylabel=time (s),
+          scaled x ticks=false,
+          ymax=27,
+          xmax=12000,
+          table/col sep=comma,
+          table/x=param-0-nm]
+        \addplot[mark=o,color=red]         table[y=rewrite-strat(bottomup-bottomup)-regression-exponential-user]{perf-LiftLetsMap.csv};
+        \addplot[mark=triangle,color=red]  table[y=rewrite-strat(topdown-bottomup)-regression-exponential-user]{perf-LiftLetsMap.csv};
+        \addplot[mark=square,color=red]    table[y=setoid-rewrite-regression-cubic-user]{perf-LiftLetsMap.csv};
+        \addplot[mark=+,color=blue]        table[y=Rewrite-for-gen-user]{perf-LiftLetsMap.csv};
+        \addplot[mark=x,color=ForestGreen] table[y=rewriting-user]{perf-LiftLetsMap.csv};
+        \addplot[mark=*,color=purple]      table[y=cps+vm-compute-regression-quadratic-user]{perf-LiftLetsMap.csv};
         \legend{rewrite\_strat bottomup,rewrite\_strat topdown,repeat setoid\_rewrite,{Our approach including reification, cbv, etc.},Our approach (only rewriting),cps+vm\_compute}
@@
     \edef\nval{\NoBindersSubfloatNval}%
-    \adjustbox{valign=t}{\resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
-        \einput{perf-Plus0Tree-only-n-\nval-cbv;rewrite-strat(bottomup).txt}
-        \einput{perf-Plus0Tree-only-n-\nval-cbv;setoid-rewrite.txt}
-        \einput{perf-Plus0Tree-only-n-\nval-cbv;rewrite-strat(topdown).txt}
-        \einput{perf-Plus0Tree-only-n-\nval-cbv;rewrite!.txt}
-        \einput{perf-Plus0Tree-only-n-\nval-Rewrite-for.txt}
-        \einput{perf-Plus0Tree-only-n-\nval-rewriting.txt}
+    \adjustbox{valign=t}{\resizebox{0.32\textwidth}{!}{\beginTikzpictureStamped[only marks]{
+        \einput{perf-Plus0Tree.csv.md5}
         \nval
@@
       %,xticklabel={\pgfkeys{/pgf/fpu}\pgfmathparse{2^\nval*\tick}$\mathsf{\pgfmathprintnumber{\pgfmathresult}}$}
-      \begin{axis}[xlabel=\# of rewrite locations,scaled x ticks=false,ylabel=time (s),ymax=7]% ymax=10]
-        \addplot[mark=square,color=red] table[x expr={\NoBindersSubfloatXRow}]{perf-Plus0Tree-only-n-\nval-cbv;rewrite-strat(bottomup).txt};
-        \addplot[mark=*,color=red] table[x expr={\NoBindersSubfloatXRow}]{perf-Plus0Tree-only-n-\nval-cbv;setoid-rewrite.txt};
-        \addplot[mark=triangle,color=red] table[x expr={\NoBindersSubfloatXRow}]{perf-Plus0Tree-only-n-\nval-cbv;rewrite-strat(topdown).txt};
-        \addplot[mark=o,color=red] table[x expr={\NoBindersSubfloatXRow}]{perf-Plus0Tree-only-n-\nval-cbv;rewrite!.txt};
-        \addplot[mark=+,color=blue] table[x expr={\NoBindersSubfloatXRow}]{perf-Plus0Tree-only-n-\nval-Rewrite-for.txt};
-        \addplot[mark=x,color=ForestGreen] table[x expr={\NoBindersSubfloatXRow}]{perf-Plus0Tree-only-n-\nval-rewriting.txt};
+      \begin{axis}[xlabel=\# of rewrite locations,
+          scaled x ticks=false,
+          ylabel=time (s),
+          ymax=7,
+          xmax=15000,
+          xtick distance=3000,
+          table/col sep=comma,
+          table/x expr={\NoBindersSubfloatXRow}]% ymax=10]
+        \addplot[discard if not={param-1-n}{\nval},mark=square,color=red] table[y=rewrite-strat(bottomup)-regression-cubic-user]{perf-Plus0Tree.csv};
+        \addplot[discard if not={param-1-n}{\nval},mark=*,color=red] table[y=setoid-rewrite-regression-cubic-user]{perf-Plus0Tree.csv};
+        \addplot[discard if not={param-1-n}{\nval},mark=triangle,color=red] table[y=rewrite-strat(topdown)-regression-cubic-user]{perf-Plus0Tree.csv};
+        \addplot[discard if not={param-1-n}{\nval},mark=o,color=red] table[y=rewrite!-regression-cubic-user]{perf-Plus0Tree.csv};
+        \addplot[discard if not={param-1-n}{\nval},mark=+,color=blue] table[y=Rewrite-for-gen-user]{perf-Plus0Tree.csv};
+        \addplot[discard if not={param-1-n}{\nval},mark=x,color=ForestGreen] table[y=rewriting-user]{perf-Plus0Tree.csv};
         \legend{rewrite\_strat bottomup,setoid\_rewrite,rewrite\_strat topdown,rewrite!,{Our approach including reification, cbv, etc.},Our approach (only rewriting)}
@@
   \sbox{\FiatCryptoSubfloat}{%
-    \adjustbox{valign=t}{\resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
-      \einput{perf-new-vm-over-old-vm.txt.md5}
-      \einput{perf-old-vm-over-old-vm.txt.md5}
-      \einput{perf-new-extraction-over-old-vm.txt.md5}
+    \adjustbox{valign=t}{\resizebox{0.32\textwidth}{!}{\beginTikzpictureStamped[only marks]{
+      \einput{perf-fiat-crypto.csv.md5}
     }
@@
           anchor=north}}
-      \begin{axis}[xlabel=prime,ylabel=slowdown factor over old approach,xmode=log, ymode=log,log basis x={2}]
-        \addplot[mark=+] table{perf-new-vm-over-old-vm.txt};
-        \addplot[mark=o] table{perf-old-vm-over-old-vm.txt};
-        \addplot[mark=*] table{perf-new-extraction-over-old-vm.txt};
-        \legend{Our approach w/ Coq's VM,Old approach (handwritten-CPS+VM),Our approach w/ extracted OCaml}
+%      \def\formatextratick#1{\raisebox{-1.5\baselineskip}{\rotatebox{-90}{#1}}}%
+%      \def\formatextratickLower#1{\raisebox{-2.5\baselineskip}{\rotatebox{-90}{#1}}}%
+      \begin{axis}[xlabel=prime,ylabel=time relative to original Fiat Crypto,xmode=log, ymode=log,log basis x={2},
+%          extra x ticks={2^255,2^256,2^384},
+%          extra x tick labels={\formatextratick{Curve25519},\formatextratickLower{P-256},\formatextratick{P-384}},
+%          extra x tick label style={yshift={-1em}},
+        ]
+        \addplot[mark=*,color=red] table[x=prime,y=OldSynthesisAndPackage-over-OldSynthesisAndPackage-real,col sep=comma]{perf-fiat-crypto.csv};
+        \addlegendentry{Original Fiat Crypto (includes reification+rewriting)}
+        \addplot[mark=+,color=blue] table[x=prime,y=NewVMFull-over-OldSynthesisAndPackage-real,col sep=comma]{perf-fiat-crypto.csv};
+        \addlegendentry{Our approach w/ Coq's VM}
+
+        %\addplot[mark=o,color=purple] table[x=prime,y=OldVM-over-OldSynthesisAndPackage-real,col sep=comma]{perf-fiat-crypto.csv};
+        %\addlegendentry{Old approach (handwritten-CPS+VM, only reduction)}
+
+        \addplot[mark=x,color=ForestGreen] table[x=prime,y=NewExtractionFull-over-OldSynthesisAndPackage-real,col sep=comma]{perf-fiat-crypto.csv};
+        \addlegendentry{Our approach w/ extracted OCaml}
       \end{axis}
@@
   \subfloat[Nested binders]{\usebox{\NestedBindersSubfloat}\vphantomSubfloatOne\label{fig:timing-UnderLetsPlus0}}%
-  \quad
-  \subfloat[Binders and recursive functions]{\usebox{\BindersAndRecursiveFunctionsSubfloat}\vphantomSubfloatTwo\label{fig:timing-LiftLetsMap}}%
+  %% \quad
+  %% \subfloat[Binders and recursive functions]{\usebox{\BindersAndRecursiveFunctionsSubfloat}\vphantomSubfloatTwo\label{fig:timing-LiftLetsMap}}%
   \quad
@@
 
-  \caption{Timing of different partial-evaluation implementations} \label{fig:multi-timing}
+  \caption{Timing of different partial-evaluation implementations}\label{fig:multi-timing}
 \end{figure*}
 
-%RSolve[{t[0]==m, t[n]==2*t[n-1] + m},t[n],n]
-%t[n]=m(2^{n+1}-1)
+% POPL 2022 review TODO:
+% The graphs in Figure 3 are pretty ugly because they have so many data points packed together.
+% It is nonetheless possible to distinguish the various series involved though, with a bit of geometrical deduction.
+% And I suppose you'd lose some information if you thinned out the data points or plotted lines only.
+% So I don't really have a concrete suggestion for improving the graphs; I just don't find them very pleasing to look at.
+
+Now we turn to evaluating performance of generated compilers.
+We start with microbenchmarks focusing attention on particular aspects of reduction and rewriting, with \autoref{sec:additionalMicro} going into more detail, including on a few more benchmarks.
+
+Our first example family, \emph{nested binders}, has two integer parameters $n$ and $m$.
+An expression tree is built with $2^n$ copies of an expression, which is itself a free variable with $m$ ``useless'' additions of zero.
+We want to see all copies of this expression reduced to just the variable.
 \autoref{fig:timing-Plus0Tree} shows the results for $n = \NoBindersSubfloatNval$ as we scale $m$.
@@
 For our own approach, we list both the total time and the time taken for core execution of a verified rewrite engine, without counting reification (converting goals to ASTs) or its inverse (interpreting results back to normal-looking goals).
-
 The comparison here is very favorable for our approach so long as $m > 2$.
-The competing tactics spike upward toward timeouts at just around a thousand rewrite locations, while our engine is still under two seconds for examples with tens of thousands of rewrite locations.
-When $m < 2$, Coq's \texttt{rewrite!} tactic does a little bit better than our engine, corresponding roughly to the overhead incurred by our term representation (which, for example, stores the types at every application node) when most of the term is in fact unchanged by rewriting.
-See \autoref{sec:additionalPlots:Plus0Tree}\footnote{Like several forward references in this section, this one goes to an appendix included within the main submission page limit, to avoid interrupting the flow in presenting the most important results.} for more detailed plots.
-
-\subsubsection{Rewriting Under Binders} \label{sec:micro:UnderLetsPlus0}
-
-\begin{wrapfigure}[7]{r}{0pt}
-{\small $\begin{aligned}
-  & \letin[{v_1 := v_0 + v_0 + 0}{}] \\ \noalign{\vskip-7pt}
-  & \vdots \\ \noalign{\vskip-3pt}
-  & \letin[{v_n := v_{n-1} + v_{n-1} + 0}{}] \\
-  & v_n + v_n + 0
-\end{aligned}$}%
-%  & \letin[{v_2 := v_1 + v_1 + 0}{}] \\
-\caption{\label{fig:micro:UnderLetsPlus0:code}Initial code}
-\end{wrapfigure}
-
-Consider now the code in \autoref{fig:micro:UnderLetsPlus0:code}, which is a version of the code above where redundant expressions are shared via \mintinline{coq}{let} bindings.
-
-\autoref{fig:timing-UnderLetsPlus0} shows the results.
+%% The competing tactics spike upward toward timeouts at just around a thousand rewrite locations, while our engine is still under two seconds for examples with tens of thousands of rewrite locations.
+%% When $m < 2$, Coq's \texttt{rewrite!} tactic does a little bit better than our engine, corresponding roughly to the overhead incurred by our term representation (which, for example, stores the types at every application node) when most of the term is in fact unchanged by rewriting.
+(See \autoref{sec:additionalPlots:Plus0Tree} %\footnote{Like several forward references in this section, this one goes to an appendix included within the main submission page limit, to avoid interrupting the flow in presenting the most important results.}
+for more detailed plots.)
+
+Now consider what happens when we use \mintinline{coq}{let} binders to share subterms within repeated addition of zero, incorporating exponentially many additions with linearly sized terms.
+\Vref{fig:timing-UnderLetsPlus0} shows the results.
 The comparison here is again very favorable for our approach.
@@
 
-\subsubsection{Performance Bottlenecks of Proof-Producing Rewriting} \label{sec:micro:setoid-rewrite-bottlenecks-lite}
-
 Although we have made our comparison against the built-in tactics \mintinline{coq}{setoid_rewrite} and \mintinline{coq}{rewrite_strat}, by analyzing the performance in detail, we can argue that these performance bottlenecks are likely to hold for any proof assistant designed like Coq.
-Detailed debugging reveals five performance bottlenecks in the existing rewriting tactics, which we discuss in \autoref{sec:setoid-rewrite-bottlenecks}.\footnote{Also included within the submission page limit, though of interest mostly to proof-assistant-implementation experts.}
-
-\subsubsection{Binders and Recursive Functions} \label{sec:micro:LiftLetsMap}
-
-\begin{wrapfigure}[11]{r}{0pt}
-{\small %\allowdisplaybreaks
-$\begin{aligned}
-  \text{map\_dbl}(\ell) & \defeq \begin{cases} [] & \text{if }\ell = [] \\
-      \letin[{y := h + h}{}] & \text{if }\ell = h::t \\
-      y :: \text{map\_dbl}(t) &
-      \end{cases} \\
-  \text{make}(n, m, v) & \defeq \begin{cases} [\underbrace{v, \ldots, v}_n] & \text{if }m = 0 \\
-      \text{map\_dbl}(\text{make}(n, m-1, v)) & \text{if }m > 0
-      \end{cases} \\
-  \text{example}_{n, m} & \defeq \forall v,\ \text{make}(n, m, v) = []
-\end{aligned}$}%
-\caption{\label{fig:micro:LiftLetsMap:code}Initial code for binders and recursive functions}
-\end{wrapfigure}
-
-The next experiment uses the code in \autoref{fig:micro:LiftLetsMap:code}.
-%Note that the ${}=[]$ at the end is just a dummy placeholder to allow us to check reduction and rewriting of this term.
-Note that the \letin{} binding blocks further reduction of map\_dbl when we iterate it $m$ times in make, and so we need to take care to preserve sharing when reducing here.
-
-\autoref{fig:timing-LiftLetsMap} compares performance between our approach, \texttt{repeat setoid\_rewrite}, and two variants of \texttt{rewrite\_strat}.
-Additionally, we consider another option, which was adopted by Fiat Cryptography at a larger scale: rewrite our functions to improve reduction behavior.
-Specifically, both functions are rewritten in continuation-passing style, which makes them harder to read and reason about but allows standard VM-based reduction to achieve good performance.
-The figure shows that \texttt{rewrite\_strat} variants are essentially unusable for this example, with \texttt{setoid\_rewrite} performing only marginally better, while our approach applied to the original, more readable definitions loses ground steadily to VM-based reduction on CPSed code.
-On the largest terms ($n \cdot m > 20,000$), the gap is 6s vs.\ 0.1s of compilation time, which should often be acceptable in return for simplified coding and proofs, plus the ability to mix proved rewrite rules with built-in reductions.
-Note that about 99\% of the difference between the full time of our method and just the rewriting is spent in the final \texttt{cbv} at the end, used to denote our output term from refied syntax.
-We blame this performance on the unfortunate fact that reduction in Coq is quadratic in the number of nested binders present; see Coq bug \coqbug{11151}.
-%traded for simplified coding and proofs, plus the ability to mix proved rewrite rules with built-in reductions.
-See \autoref{sec:LiftLetsMap-more} for more on this microbenchmark.
-
-\subsubsection{Full Reduction} \label{sec:micro:Eratosthenes}
-
-\begin{wrapfigure}[18]{r}{0pt}
-  \resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
-      \einput{perf-SieveOfEratosthenes-simpl.txt}
-      \einput{perf-SieveOfEratosthenes-cbn.txt}
-      \einput{perf-SieveOfEratosthenes-lazy.txt}
-      \einput{perf-SieveOfEratosthenes-Rewrite-for.txt}
-      \einput{perf-SieveOfEratosthenes-rewriting.txt}
-      \einput{perf-SieveOfEratosthenes-cbv.txt}
-      \einput{perf-SieveOfEratosthenes-vm-compute.txt}
-      \einput{perf-SieveOfEratosthenes-native(2)(real).txt}
-  }
-    \pgfplotsset{every axis legend/.append style={
-        at={(0.5,-0.2)},
-        anchor=north}}
-    \begin{axis}[xlabel=upper bound,ylabel=time (s),ymax=65]
-      \addplot[mark=square,color=red] table{perf-SieveOfEratosthenes-simpl.txt};
-      \addplot[mark=triangle,color=red] table{perf-SieveOfEratosthenes-cbn.txt};
-      \addplot[mark=square*,color=red] table{perf-SieveOfEratosthenes-lazy.txt};
-      \addplot[mark=+,color=blue] table{perf-SieveOfEratosthenes-Rewrite-for.txt};
-      \addplot[mark=x,color=ForestGreen] table{perf-SieveOfEratosthenes-rewriting.txt};
-      \addplot[mark=triangle*,color=orange] table{perf-SieveOfEratosthenes-cbv.txt};
-      \addplot[mark=*,color=purple] table{perf-SieveOfEratosthenes-vm-compute.txt};
-      \addplot[mark=-,color=red] table{perf-SieveOfEratosthenes-native(2)(real).txt};
-      \legend{simpl,cbn,lazy,{Our approach including reification, cbv, etc.},Our approach (only rewriting),cbv,vm\_compute,native\_compute}
-    \end{axis}
-  \end{tikzpicture}}
-  \caption{\label{fig:timing-SieveOfEratosthenes}Full evaluation, Sieve of Eratosthenes}
-\end{wrapfigure}
-
-The final experiment involves full reduction in computing the Sieve of Eratosthenes, taking inspiration on benchmark choice from \citet{Aehlig}.
-We find in \autoref{fig:timing-SieveOfEratosthenes} that we are slower than \texttt{vm\_compute}, \mintinline{coq}{native_compute}, and \texttt{cbv}, but faster than \mintinline{coq}{lazy}, and of course much faster than \texttt{simpl} and \texttt{cbn}, which are quite slow.
-
-\subsection{Macrobenchmark: Fiat Cryptography} \label{sec:macro}
+Detailed debugging reveals five performance bottlenecks in the existing tactics, discussed in \autoref{sec:setoid-rewrite-bottlenecks}.
+
+\subsection{Macrobenchmark: Fiat Cryptography}\label{sec:macro}
 
@@
 
-\autoref{fig:scaling} graphs running time of three different partial-evaluation methods for Fiat Cryptography, as the prime modulus of arithmetic scales up.
-Times are normalized to the performance of the original method, which relied entirely on standard Coq reduction.
-Actually, in the course of running this experiment, we found a way to improve the old approach for a fairer comparison.
-It had relied on Coq's configurable \texttt{cbv} tactic to perform reduction with selected rules of the definitional equality, which the Fiat Cryptography developers had applied to blacklist identifiers that should be left for compile-time execution.
-By instead hiding those identifiers behind opaque module-signature ascription, we were able to run Coq's more-optimized virtual-machine-based reducer.
-
-As the figure shows, our approach running partial evaluation inside Coq's kernel begins with about a 10$\times$ performance disadvantage vs.\ the original method.
-With log scale on both axes, we see that this disadvantage narrows to become nearly negligible for the largest primes, of around 500 bits.
-(We used the same set of prime moduli as in the experiments run by \citet{FiatCryptoSP19}, which were chosen based on searching the archives of an elliptic-curves mailing list for all prime numbers.)
-It makes sense that execution inside Coq leaves our new approach at a disadvantage, as we are essentially running an interpreter (our normalizer) within an interpreter (Coq's kernel), while the old approach ran just the latter directly.
+\Vref{fig:scaling} graphs running time of three different partial-evaluation and rewriting methods for Fiat Cryptography, as the prime modulus of arithmetic scales up.
+Times are normalized to the performance of the original method of \citet{FiatCryptoSP19}, which relied on standard Coq reduction to evaluate code that had been manually written in CPS, followed by reification and a custom ad-hoc simplification and rewriting engine.
+%In the course of running this experiment, we found a way to improve the reduction part of the old approach for a somewhat fairer comparison.
+%It had relied on Coq's configurable \taccbv{} tactic to perform reduction with selected rules of the definitional equality, which the Fiat Cryptography developers had applied to blacklist identifiers that should be left for compile-time execution.
+%By instead hiding those identifiers behind opaque module-signature ascription, we were able to run Coq's more-optimized virtual-machine-based reducer.
+
+%As the figure shows, our approach running partial evaluation inside Coq's kernel begins with about a 10$\times$ performance disadvantage vs.\ the original method.
+%With log scale on both axes, we see that this disadvantage narrows to become nearly negligible for the largest primes, of around 500 bits.
+As the figure shows, our approach gives about a 10$\times$--1000$\times$ speed-up over the original Fiat Cryptography pipeline.
+%(We used the same set of prime moduli as in the experiments run by \citet{FiatCryptoSP19}, which were chosen based on searching the archives of an elliptic-curves mailing list for all prime numbers.)
+Inspection of the timing profiles of the original pipeline reveals that reification dominates the timing profile; since partial evaluation is performed by Coq's kernel, reification must happen \emph{after} partial evaluation, and hence the size of the term being reified grows with the size of the output code.
 Also recall that the old approach required rewriting Fiat Cryptography's library of arithmetic functions in continuation-passing style, enduring this complexity in library correctness proofs, while our new approach applies to a direct-style library.
-Finally, the old approach included a custom reflection-based arithmetic simplifier for term syntax, run after traditional reduction, whereas now we are able to apply a generic engine that combines both, without requiring more than proving traditional rewrite rules.
-
-The figure also confirms clear performance advantage of running reduction in code extracted to OCaml, which is possible because our plugin produces verified code in Coq's functional language.
-By the time we reach middle-of-the-pack prime size around 300 bits, the extracted version is running about 10$\times$ as quickly as the baseline.
-
-\section{Related Work}\label{sec:related}
-
-We have already discussed the work of \citet{Aehlig}, which introduced the basic structure that our engine shares, but which required a substantially larger trusted code base, did not tackle certain challenges in scaling to large partial-evaluation problems, and did not report any performance experiments in partial evaluation.
-
-We have also mentioned \Rtac{}~\cite{rtac}, which implements an experimental reflective version of \texttt{rewrite\_strat} supporting arbitrary setoid relations, unification variables, and arbitrary semi-decidable side conditions solvable by other reflective tactics, using de Bruijn indexing to manage binders.
-We were unfortunately unable to get the rewriter to work with Coq 8.10 and were also not able to determine from the paper how to repurpose the rewriter to handle our benchmarks.
-
-Our implementation builds on fast full reduction in Coq's kernel, via a virtual machine~\cite{vmcompute} or compilation to native code~\cite{nativecompute}.
-Especially the latter is similar in adopting an NbE style for full reduction, simplifying even under $\lambda$s, on top of a more traditional implementation of OCaml that never executes preemptively under $\lambda$s.
-Neither approach unifies support for rewriting with proved rules, and partial evaluation only applies in very limited cases, where functions that should not be evaluated at compile time must have properly opaque definitions that the evaluator will not consult.
-Neither implementation involved a machine-checked proof suitable to bootstrap on top of reduction support in a kernel providing simpler reduction.
-
-% A more limited form of code generation for cryptography was already widespread through libraries like OpenSSL, which specifically uses Perl scripts to generate assembly code.
-% The Vale tool suite~\cite{Vale} formalizes these practices with a more principled language and associated verification tools.
-% However, code generation done in this style is significantly simpler than what we treat here, amounting mostly to loop unrolling, macro substitution, and computation of compile-time constants.
-% Also, Vale involves a significantly larger trusted code base than with our approach, with no reduction to some kernel proof checker, instead placing trust in language-specific tooling and an SMT solver.
-
-A variety of forms of pragmatic partial evaluation have been demonstrated, with Lightweight Modular Staging~\cite{LMS} in Scala as one of the best-known current examples.
-A kind of type-based overloading for staging annotations is used to smooth the rough edges in writing code that manipulates syntax trees.
-The LMS-Verify system~\cite{LMSVerify} can be used for formal verification of generated code after-the-fact.
-Typically LMS-Verify has been used with relatively shallow properties (though potentially applied to larger and more sophisticated code bases than we tackle), not scaling to the kinds of functional-correctness properties that concern us here, justifying investment in verified partial evaluators.
+Finally, the old approach included a custom reflection-based arithmetic simplifier for term syntax, run after traditional reduction, whereas now we are able to apply a generic engine that combines both, without requiring more than proving traditional rewrites.
+
+The figure also confirms a clear performance advantage of running reduction in code extracted to OCaml, which is possible because our plugin produces verified code in Coq's functional language.
+The extracted version is about 10$\times$ faster than running in Coq's kernel.
+%By the time we reach middle-of-the-pack prime size around 300 bits, the extracted version, which includes arithmetic simplification, is faster even than the best performance we could achieve for the old
 
@@
 
-There are a number of natural extensions to our engine.
-For instance, we do not yet allow pattern variables marked as ``constants only'' to apply to container datatypes; we limit the mixing of higher-order and polymorphic types, as well as limiting use of first-class polymorphism; we do not support rewriting with equalities of non-fully-applied functions; we only support decidable predicates as rule side conditions, and the predicates may only mention pattern variables restricted to matching constants; we have hardcoded support for a small set of container types and their eliminators; we support rewriting with equality and no other relations (e.g., subset inclusion); and we require decidable equality for all types mentioned in rules.
-It may be helpful to design an engine that lifts some or all of these limitations, building on the basic structure that we present here.
-
-%% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-                                        %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
+By far the biggest next step for our engine is to integrate abstract interpretation with rewriting and partial evaluation.
+We expect this would net us asymptotic performance gains
+as described in \autoref{sec:fusing-compiler-passes}.
+Additionally, it would allow us to simplify the phrasing of many of our post-abstract-interpretation rewrite rules, by relegating bounds information to side conditions rather than requiring that they appear in the syntactic form of the rule.
+
+There are also a number of natural extensions to our engine.
+For instance, we do not yet allow pattern variables marked as ``constants only'' to apply to container datatypes; we limit the mixing of higher-order and polymorphic types, as well as limiting use of first-class polymorphism; we do not support rewriting with equalities of nonfully-applied functions; we only support decidable predicates as rule side conditions, and the predicates may only mention pattern variables restricted to matching constants; we have hardcoded support for a small set of container types and their eliminators; we support rewriting with equality and no other relations% (e.g., subset inclusion)
+; and we require decidable equality for all types mentioned in rules.
+%It may be helpful to design an engine that lifts some or all of these limitations, building on the basic structure that we present here.
 
@@
 \clearpage
-\nocite{*}
+%\nocite{*}
 \bibliography{rewriting}
@@
 
+\startsupplement
+%% Appendix
 \appendix
 
-\section{Performance Bottlenecks of Proof-Producing Rewriting} \label{sec:setoid-rewrite-bottlenecks}
+\FloatBarrier\section{Performance Bottlenecks of Proof-Producing Rewriting}\label{sec:setoid-rewrite-bottlenecks}
 
@@
 We posit that this overhead comes from \mintinline{coq}{setoid_rewrite} looking for head-symbol matches and then creating evars (existential variables) to instantiate the arguments of the lemmas for each head-symbol-match location; hence even if there are no matches of the rule as a whole, there may still be head-symbol matches.
-Since Coq uses a locally nameless representation~\cite{LocallyNameless} for its terms, evar contexts are necessarily representated as \emph{named} contexts.
+Since Coq uses a locally nameless representation~\cite{LocallyNameless} for its terms, evar contexts are necessarily represented as \emph{named} contexts.
 Representing a substitution between named contexts takes linear space, even when the substitution is trivial, and hence each evar incurs overhead linear in the number of binders above it.
@@
 One way to ask this question in Coq is to see how long it takes to run \mintinline{coq}{Qed}.
-While \mintinline{coq}{Qed}-time is asymptotically better, it is still quadratic in the number of binders.
+While \mintinline{coq}{Qed} time is asymptotically better, it is still quadratic in the number of binders.
 This outcome is unsurprising, because the proof-term size is quadratic in the number of binders.
-On this microbenchmark, we found that \mintinline{coq}{Qed}-time hits one second at about 250 binders, and using the best-fit quadratic line suggests that it would hit 10 seconds at about 800 binders and 100 seconds at about 2\,500 binders.
+On this microbenchmark, we found that \mintinline{coq}{Qed} time hits one second at about 250 binders, and using the best-fit quadratic line suggests that it would hit 10 seconds at about 800 binders and 100 seconds at about 2\,500 binders.
 While this may be reasonable for the microbenchmarks, which only contain as many rewrite occurrences as there are binders, it would become unwieldy to try to build and typecheck such a proof with a rule for every primitive reduction step, which would be required if we want to avoid manually CPS-converting the code in Fiat Cryptography.
@@
 Note that \mintinline{coq}{g x} occurs three times (and will continue to occur in every subsequent step).
+
+\subsection{Poor subterm sharing}
+
+How easy is it to share subterms and create a linearly sized proof?
+While it is relatively straightforward to share subterms using \mintinline{coq}{let} binders when the rewrite locations are not under any binders, it is not at all obvious how to share subterms when the terms occur under different binders.
+Hence any rewriting algorithm that does not find a way to share subterms across different contexts will incur a quadratic factor in proof-building and proof-checking time, and we expect this factor will be significant enough to make applications to projects as large as Fiat Crypto infeasible.
+
+\subsection{Overhead from the \mintinline{coq}{let} typing rule}
+
+Suppose we had a proof-producing rewriting algorithm that shared subterms even under binders.
+Would it be enough?
+It turns out that even when the proof size is linear in the number of binders, the cost to typecheck it in Coq is still quadratic!
+The reason is that when checking that \texttt{f : T} in a context \mintinline{coq}{x := v}, to check that \mintinline{coq}{let x := v in f} has type \texttt{T} (assuming that \mintinline{coq}{x} does not occur in \texttt{T}), Coq will substitute \mintinline{coq}{v} for \mintinline{coq}{x} in \texttt{T}.
+So if a proof term has $n$ \mintinline{coq}{let} binders (e.g., used for sharing subterms), Coq will perform $n$ substitutions on the type of the proof term, even if none of the \mintinline{coq}{let} binders are used.
+If the number of \mintinline{coq}{let} binders is linear in the size of the type, there is quadratic overhead in proof-checking time, even when the proof-term size is linear.
+
+We performed a microbenchmark on a rewriting goal with no binders (because there is an obvious algorithm for sharing subterms in that case) and found that the proof-checking time reached about one second at about 2\,000 binders and reached 10 seconds at about 7\,000 binders.
+While these results might seem good enough for Fiat Cryptography, we expect that there are hundreds of thousands of primitive reduction/rewriting steps even when there are only a few hundred binders in the output term, and we would need \mintinline{coq}{let} binders for each of them.
+Furthermore, we expect that getting such an algorithm correct would be quite tricky.
+
+Fixing this quadratic bottleneck would, as far as we can tell, require deep changes in how Coq is implemented; it would either require reworking all of Coq to operate on some efficient representation of delayed substitutions paired with unsubstituted terms, or else it would require changing the typing rules of the type theory itself to remove this substitution from the typing rule for \mintinline{coq}{let}.
+Note that there is a similar issue that crops up for function application and abstraction.
+
+\subsection{Inherent advantages of reflection}
+
+Finally, even if this quadratic bottleneck were fixed, \citet{Aehlig} reported a $10\times$--$100\times$ speed-up over the \emph{simp} tactic in Isabelle, which performs all of the intermediate rewriting steps via the kernel API.
+Their results suggest that even if all of the superlinear bottlenecks were fixed---no small undertaking---rewriting and partial evaluation via reflection might still be orders of magnitude faster than any proof-term-generating tactic.
+
+\clearpage
+
+\section{Additional Benchmarking Plots}\label{sec:additionalPlots}
+
+\subsection{Rewriting Without Binders}\label{sec:additionalPlots:Plus0Tree}
 
@@
   \resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
-    \einput{perf-Plus0Tree-only-m-\mval-cbv;rewrite-strat(bottomup).txt}
-    \einput{perf-Plus0Tree-only-m-\mval-cbv;setoid-rewrite.txt}
-    \einput{perf-Plus0Tree-only-m-\mval-cbv;rewrite-strat(topdown).txt}
-    \einput{perf-Plus0Tree-only-m-\mval-cbv;rewrite!.txt}
-    \einput{perf-Plus0Tree-only-m-\mval-Rewrite-for.txt}
-    \einput{perf-Plus0Tree-only-m-\mval-rewriting.txt}
+    \einput{perf-Plus0Tree.csv.md5}
     \mval
@@
         anchor=north}}
-    \begin{axis}[xlabel=$n$,ylabel=time (s),ymode=log,ymin=0.002,ymax=110]
-      \addplot[mark=square,color=red] table{perf-Plus0Tree-only-m-\mval-cbv;rewrite-strat(bottomup).txt};
-      \addplot[mark=*,color=red] table{perf-Plus0Tree-only-m-\mval-cbv;setoid-rewrite.txt};
-      \addplot[mark=triangle,color=red] table{perf-Plus0Tree-only-m-\mval-cbv;rewrite-strat(topdown).txt};
-      \addplot[mark=o,color=red] table{perf-Plus0Tree-only-m-\mval-cbv;rewrite!.txt};
-      \addplot[mark=+,color=blue] table{perf-Plus0Tree-only-m-\mval-Rewrite-for.txt};
-      \addplot[mark=x,color=ForestGreen] table{perf-Plus0Tree-only-m-\mval-rewriting.txt};
+    \begin{axis}[xlabel=$n$,
+        ylabel=time (s),
+        ymode=log,
+        ymin=0.002,
+        ymax=110,
+        xmax=17,
+        table/col sep=comma,
+        table/x=param-1-n]
+      \addplot[discard if not={param-2-m}{\mval},mark=square,color=red] table[y=rewrite-strat(bottomup)-regression-cubic-user]{perf-Plus0Tree.csv};
+      \addplot[discard if not={param-2-m}{\mval},mark=*,color=red] table[y=setoid-rewrite-regression-cubic-user]{perf-Plus0Tree.csv};
+      \addplot[discard if not={param-2-m}{\mval},mark=triangle,color=red] table[y=rewrite-strat(topdown)-regression-cubic-user]{perf-Plus0Tree.csv};
+      \addplot[discard if not={param-2-m}{\mval},mark=o,color=red] table[y=rewrite!-regression-cubic-user]{perf-Plus0Tree.csv};
+      \addplot[discard if not={param-2-m}{\mval},mark=+,color=blue] table[y=Rewrite-for-gen-user]{perf-Plus0Tree.csv};
+      \addplot[discard if not={param-2-m}{\mval},mark=x,color=ForestGreen] table[y=rewriting-user]{perf-Plus0Tree.csv};
       \legend{rewrite\_strat bottomup,setoid\_rewrite,rewrite\_strat topdown,rewrite!,{Our approach including reification, cbv, etc.},Our approach (only rewriting)}
@@
   \caption{Timing of different partial-evaluation implementations for code with no binders for fixed $m$.
-Note that we have a logarithmic time scale, because term size is proportional to $2^n$.} \label{fig:timing-Plus0Tree-fixed-m}
+Note that we have a logarithmic time scale, because term size is proportional to $2^n$.}\label{fig:timing-Plus0Tree-fixed-m}
 \end{figure*}
-
-\subsection{Poor subterm sharing}
-
-How easy is it to share subterms and create a linearly sized proof?
-While it is relatively straightforward to share subterms using \mintinline{coq}{let} binders when the rewrite locations are not under any binders, it is not at all obvious how to share subterms when the terms occur under different binders.
-Hence any rewriting algorithm that does not find a way to share subterms across different contexts will incur a quadratic factor in proof-building and proof-checking time, and we expect this factor will be significant enough to make applications to projects as large as Fiat Crypto infeasible.
-
-\subsection{Overhead from the \mintinline{coq}{let} typing rule}
-
-Suppose we had a proof-producing rewriting algorithm that shared subterms even under binders.
-Would it be enough?
-It turns out that even when the proof size is linear in the number of binders, the cost to typecheck it in Coq is still quadratic!
-The reason is that when checking that \texttt{f : T} in a context \mintinline{coq}{x := v}, to check that \mintinline{coq}{let x := v in f} has type \texttt{T} (assuming that \mintinline{coq}{x} does not occur in \texttt{T}), Coq will substitute \mintinline{coq}{v} for \mintinline{coq}{x} in \texttt{T}.
-So if a proof term has $n$ \mintinline{coq}{let} binders (e.g., used for sharing subterms), Coq will perform $n$ substitutions on the type of the proof term, even if none of the \mintinline{coq}{let}-binders are used.
-If the number of \mintinline{coq}{let}-binders is linear in the size of the type, there is quadratic overhead in proof-checking time, even when the proof-term size is linear.
-
-We performed a microbenchmark on a rewriting goal with no binders (because there is an obvious algorithm for sharing subterms in that case) and found that the proof-checking time reached about one second at about 2\,000 binders and reached 10 seconds at about 7\,000 binders.
-While these results might seem good enough for Fiat Cryptography, we expect that there are hundreds of thousands of primitive reduction/rewriting steps even when there are only a few hundred binders in the output term, and we would need \mintinline{coq}{let}-binders for each of them.
-Furthermore, we expect that getting such an algorithm correct would be quite tricky.
 
@@
   \resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
-      \einput{perf-Plus0Tree-only-n-\nval-cbv;rewrite-strat(bottomup).txt}
-      \einput{perf-Plus0Tree-only-n-\nval-cbv;setoid-rewrite.txt}
-      \einput{perf-Plus0Tree-only-n-\nval-cbv;rewrite-strat(topdown).txt}
-      \einput{perf-Plus0Tree-only-n-\nval-cbv;rewrite!.txt}
-      \einput{perf-Plus0Tree-only-n-\nval-Rewrite-for.txt}
-      \einput{perf-Plus0Tree-only-n-\nval-rewriting.txt}
+      \einput{perf-Plus0Tree.csv.md5}
       \nval
@@
         anchor=north}}
-    \begin{axis}[xlabel=$m$,ylabel=time (s),ymax=10]
-      \addplot[mark=square,color=red] table{perf-Plus0Tree-only-n-\nval-cbv;rewrite-strat(bottomup).txt};
-      \addplot[mark=*,color=red] table{perf-Plus0Tree-only-n-\nval-cbv;setoid-rewrite.txt};
-      \addplot[mark=triangle,color=red] table{perf-Plus0Tree-only-n-\nval-cbv;rewrite-strat(topdown).txt};
-      \addplot[mark=o,color=red] table{perf-Plus0Tree-only-n-\nval-cbv;rewrite!.txt};
-      \addplot[mark=+,color=blue] table{perf-Plus0Tree-only-n-\nval-Rewrite-for.txt};
-      \addplot[mark=x,color=ForestGreen] table{perf-Plus0Tree-only-n-\nval-rewriting.txt};
+    \begin{axis}[xlabel=$m$,
+        ylabel=time (s),
+        ymax=10,
+        xmax=1100,
+        table/col sep=comma,
+        table/x=param-2-m]
+      \addplot[discard if not={param-1-n}{\nval},mark=square,color=red] table[y=rewrite-strat(bottomup)-regression-cubic-user]{perf-Plus0Tree.csv};
+      \addplot[discard if not={param-1-n}{\nval},mark=*,color=red] table[y=setoid-rewrite-regression-cubic-user]{perf-Plus0Tree.csv};
+      \addplot[discard if not={param-1-n}{\nval},mark=triangle,color=red] table[y=rewrite-strat(topdown)-regression-cubic-user]{perf-Plus0Tree.csv};
+      \addplot[discard if not={param-1-n}{\nval},mark=o,color=red] table[y=rewrite!-regression-cubic-user]{perf-Plus0Tree.csv};
+      \addplot[discard if not={param-1-n}{\nval},mark=+,color=blue] table[y=Rewrite-for-gen-user]{perf-Plus0Tree.csv};
+      \addplot[discard if not={param-1-n}{\nval},mark=x,color=ForestGreen] table[y=rewriting-user]{perf-Plus0Tree.csv};
       \legend{rewrite\_strat bottomup,setoid\_rewrite,rewrite\_strat topdown,rewrite!,{Our approach including reification, cbv, etc.},Our approach (only rewriting)}
@@
 \PlusZeroTreeNValSubfloat{9}
-  \caption{Timing of different partial-evaluation implementations for code with no binders for fixed $n$ (1, 2, 3, and then we jump to 9)} \label{fig:timing-Plus0Tree-fixed-n}
+  \caption{Timing of different partial-evaluation implementations for code with no binders for fixed $n$ (1, 2, 3, and then we jump to 9)}\label{fig:timing-Plus0Tree-fixed-n}
 \end{figure*}
@@
 
-Fixing this quadratic bottleneck would, as far as we can tell, require deep changes in how Coq is implemented; it would either require reworking all of Coq to operate on some efficient representation of delayed substitutions paired with unsubstituted terms, or else it would require changing the typing rules of the type theory itself to remove this substitution from the typing rule for \mintinline{coq}{let}.
-Note that there is a similar issue that crops up for function application and abstraction.
-
-\subsection{Inherent advantages of reflection}
-
-Finally, even if this quadratic bottleneck were fixed, \citet{Aehlig} reported a $10\times$--$100\times$ speed-up over the \emph{simp} tactic in Isabelle, which performs all of the intermediate rewriting steps via the kernel API.
-Their results suggest that even if all of the super-linear bottlenecks were fixed---no small undertaking---rewriting and partial evaluation via reflection might still be orders of magnitude faster than any proof-term-generating tactic.
-
-\section{Additional Benchmarking Plots} \label{sec:additionalPlots}
-
-\subsection{Rewriting Without Binders} \label{sec:additionalPlots:Plus0Tree}
-
 The code in \autoref{fig:micro:Plus0Tree:code} in \autoref{sec:micro:Plus0Tree} is parameterized on both $n$, the height of the tree, and $m$, the number of rewriting occurrences per node.
@@
 
-\subsection{Additional Information on the Fiat Cryptography Benchmark} \label{sec:additionalPlots:FiatCrypto} \label{sec:additionalMacro}
+\subsection{Additional Information on the Fiat Cryptography Benchmark}\label{sec:additionalPlots:FiatCrypto}\label{sec:additionalMacro}
+
+The data for this benchmark can be found on GitHub at \githublink{mit-plv}{fiat-crypto}[perf-testing-data-ITP-2022-rewriting].
 
@@
   \resizebox{0.45\textwidth}{!}{\beginTikzpictureStamped[only marks]{
-      \einput{perf-new-vm-times--only-#1-#3.txt}
-      \einput{perf-old-vm-times--only-#1-#3.txt}
-      \einput{perf-new-extraction-times--only-#1-#3.txt}
+      \einput{perf-fiat-crypto.csv.md5}
   }
@@
     \begin{axis}[xlabel=prime,ylabel=time (s),xmode=log, ymode=log,log basis x={2}]
-      \addplot[mark=+] table{perf-new-vm-times--only-#1-#3.txt};
-      \addplot[mark=o] table{perf-old-vm-times--only-#1-#3.txt};
-      \addplot[mark=*] table{perf-new-extraction-times--only-#1-#3.txt};
-      \legend{Our approach w/ Coq's VM,Old approach (handwritten-CPS+VM),Our approach w/ extracted OCaml}
+      \addplot[mark=*,color=red] table[x=prime,y=#1-#3-OldSynthesisAndPackage-real,col sep=comma]{perf-fiat-crypto.csv};
+      \addlegendentry{Old approach (including reification+rewriting)}
+      \addplot[mark=+,color=blue] table[x=prime,y=#1-#3-NewVMFull-real,col sep=comma]{perf-fiat-crypto.csv};
+      \addlegendentry{Our approach w/ Coq's VM}
+
+      %\addplot[mark=o,color=purple] table[x=prime,y=#1-#3-OldVM-real,col sep=comma]{perf-fiat-crypto.csv};
+      %\addlegendentry{Old approach (handwritten-CPS+VM, only reduction)}
+
+      \addplot[mark=x,color=ForestGreen] table[x=prime,y=#1-#3-NewExtractionFull-real,col sep=comma]{perf-fiat-crypto.csv};
+      \addlegendentry{Our approach w/ extracted OCaml}
     \end{axis}
@@
   \MacroSubfloat{WordByWordMontgomery}{word-by-word Montgomery}{x64}%
-  \caption{\label{fig:timing-macro-various}Timing of different partial-evaluation implementations for Fiat Cryptography {vs.} prime modulus}
+  \caption{\label{fig:timing-macro-various}Timing of different partial-evaluation implementations for Fiat Cryptography vs.\ prime modulus}
 \end{figure*}
 
-It may also be useful to see performance results with absolute times, rather than normalized execution ratios {vs.} the original Fiat Cryptography implementation.
+It may also be useful to see performance results with absolute times, rather than normalized execution ratios vs.\ the original Fiat Cryptography implementation.
 Furthermore, the benchmarks fit into four quite different groupings: elements of the cross product of two algorithms (unsaturated Solinas and word-by-word Montgomery) and bitwidths of target architectures (32-bit or 64-bit).
@@
 \clearpage
-\supplement
-
-\section{Additional Information on Microbenchmarks} \label{sec:additionalMicro}
-
-We performed all benchmarks on a 3.5 GHz Intel Haswell running Linux and Coq 8.10.0.
-We name the subsections here with the names that show up in the code supplement.
-
-\subsection{UnderLetsPlus0} \label{sec:UnderLetsPlus0-more}
-
-We provide more detail on the ``nested binders'' microbenchmark of \autoref{sec:micro:UnderLetsPlus0} displayed in \autoref{fig:timing-UnderLetsPlus0}.
-
-Recall that we are removing all of the${}+0$s from
-{\small \begin{align*}
-  & \letin[{v_1 := v_0 + v_0 + 0}{}] \\
-%  & \letin[{v_2 := v_1 + v_1 + 0}{}] \\
-  & \vdots \\
+\begin{figure}[t]
+  \subfloat[Expressions computing initial code for Rewriting Without Binders]{%
+  {\small $\begin{aligned}
+  \text{iter}_m(v) & = v + \underbrace{0 + 0 + \cdots + 0}_m \\
+  \text{tree}_{0,m}(v) &= \text{iter}_m(v + v) \\
+  \text{tree}_{n+1,m}(v) &= \text{iter}_m(\text{tree}_{n,m}(v) + \text{tree}_{n,m}(v))
+    \end{aligned}$}%
+  \label{fig:micro:Plus0Tree:code}}
+  \qquad
+  \subfloat[Initial code for Rewriting Under Binders]{%
+  {\small $\begin{aligned}
+  & \letin[{v_1 := v_0 + v_0 + 0}{}] \\ \noalign{\vskip-7pt}
+  & \vdots \\ \noalign{\vskip-3pt}
   & \letin[{v_n := v_{n-1} + v_{n-1} + 0}{}] \\
   & v_n + v_n + 0
-\end{align*}}%
-
-The code used to define this microbenchmark is
-\begin{minted}[fontsize=\small]{coq}
-Definition make_lets_def (n:nat) (v acc : Z) :=
- @nat_rect (fun _ => Z * Z -> Z)
-   (fun '(v, acc) => acc + acc + v)
-   (fun _ rec '(v, acc) =>
-     dlet acc := acc + acc + v in rec (v, acc))
-   n
-   (v, acc).
-\end{minted}
-We note some details of the rewriting framework that were glossed over in the main body of the paper, which are useful for using the code:
-Although the rewriting framework does not support dependently typed constants, we can automatically preprocess uses of eliminators like \mintinline{coq}{nat_rect} and \mintinline{coq}{list_rect} into non-dependent versions.
-The tactic that does this preprocessing is extensible via \Ltac{}'s reassignment feature.
-Since pattern-matching compilation mixed with NbE requires knowing how many arguments a constant can be applied to, we must internally use a version of the recursion principle whose type arguments do not contain arrows; current preprocessing can handle recursion principles with either no arrows or one arrow in the motive.
-Even though we will eventually plug in 0 for $v$, we jump through some extra hoops to ensure that our rewriter cannot cheat by rewriting away the ${}+0$ before reducing the recursion on $n$.
-
-We can reduce this expression in three ways.
-
-\subsubsection{Our Rewriter}
-One lemma is required for rewriting with our rewriter:
-\begin{minted}[fontsize=\small]{coq}
-Lemma Z.add_0_r : forall z, z + 0 = z.
-\end{minted}
-
-Creating the rewriter takes about 12 seconds on the machine we used for running the performance experiments:
-\begin{minted}[fontsize=\small]{coq}
-Make myrew := Rewriter For (Z.add_0_r, eval_rect nat, eval_rect prod).
-\end{minted}
-Recall from \autoref{sec:explain-eval-rect} that \mintinline{coq}{eval_rect} is a definition provided by our framework for eagerly evaluating recursion associated with certain types.
-It functions by triggering typeclass resolution for the lemmas reducing the recursion principle associated to the given type.
-We provide instances for \texttt{nat}, \texttt{prod}, \texttt{list}, \texttt{option}, and \texttt{bool}.
-Users may add more instances if they desire.
-
-\subsubsection{\texorpdfstring{\texttt{setoid\_rewrite}}{setoid\_rewrite} and \texorpdfstring{\texttt{rewrite\_strat}}{rewrite\_strat}}
-To give as many advantages as we can to the preexisting work on rewriting, we pre-reduce the recursion on \mintinline{coq}{nat}s using \texttt{cbv} before performing \texttt{setoid\_rewrite}.
-(Note that \texttt{setoid\_rewrite} cannot itself perform reduction without generating large proof terms, and \texttt{rewrite\_strat} is not currently capable of sequencing reduction with rewriting internally due to bugs such as \coqbug{10923}.)
-Rewriting itself is easy; we may use any of \texttt{repeat setoid\_rewrite Z.add\_0\_r}, \texttt{rewrite\_strat topdown Z.add\_0\_r}, or \texttt{rewrite\_strat bottomup Z.add\_0\_r}.
-
-\subsection{Plus0Tree} \label{sec:Plus0Tree-more}
-
-This is a version of \autoref{sec:UnderLetsPlus0-more} without any let binders, discussed in \autoref{sec:micro:Plus0Tree} but not displayed in \autoref{fig:multi-timing}.
+\end{aligned}$}%
+%  & \letin[{v_2 := v_1 + v_1 + 0}{}] \\
+  \label{fig:micro:UnderLetsPlus0:code}%
+  }
+  \caption{\label{fig:micro:Plus0Tree+UnderLetsPlus0:code}Code for rewriting without and under binders}
+\end{figure}
+
+\FloatBarrier\section{Additional Information on Microbenchmarks}\label{sec:additionalMicro}
+
+We performed all benchmarks on a 3.5 GHz Intel Haswell running Linux and Coq 8.11.1.
+We name the subsections here with the names that show up in the code supplement.
+
+\subsection{Rewriting Without Binders: Plus0Tree}\label{sec:Plus0Tree-more}\label{sec:micro:Plus0Tree}
+
+Consider the code defined by the expression $\text{tree}_{n,m}(v)$ in \autoref{fig:micro:Plus0Tree:code}.
+%\todo{make sure the figure spacing is okay}
+%RSolve[{t[0]==m, t[n]==2*t[n-1] + m},t[n],n]
+%t[n]=m(2^{n+1}-1)
+We want to remove all of the${}+0$s.
+There are $\Theta(m \cdot 2^n)$ such rewriting locations.
+We can start from this expression directly, in which case reification alone takes as much time as Coq's \texttt{rewrite}.
+As the reification method was not especially optimized, and there exist fast reification methods~\cite{ReificationITP18}, we instead start from a call to a recursive function that generates such an expression.
+
+%This is a version of \autoref{sec:UnderLetsPlus0-more} without any let binders, discussed in \autoref{sec:micro:Plus0Tree} but not displayed in \autoref{fig:multi-timing}.
 
@@
 
-\subsection{LiftLetsMap} \label{sec:LiftLetsMap-more}
-
-We now discuss in more detail the ``binders and recursive functions'' example from \autoref{sec:micro:LiftLetsMap}.
-
-The expression we want to get out at the end looks like:
-\begin{align*}
-    & \letin[{v_{1,1} := v + v}{}] \\
-    & \vdots \\
-    & \letin[{v_{1,n} := v + v}{}] \\
-    & \letin[{v_{2,1} := v_{1,1} + v_{1,1}}{}] \\
-    & \vdots \\
-    & \letin[{v_{2,n} := v_{1,n} + v_{1,n}}{}] \\
-    & \vdots \\
-    & [v_{m,1}, \ldots, v_{m,n}]
-\end{align*}
-
-Recall that we make this example with the code
-\begin{minted}[fontsize=\small]{coq}
-Definition map_double (ls : list Z) :=
-  list_rect _ [] (λ x xs rec, let y := x + x in y :: rec) ls.
-Definition make (n : nat) (m : nat) (v : Z) :=
-  nat_rect _ (List.repeat v n) (λ _ rec, map_double rec) m.
-\end{minted}
-
-We can perform this rewriting in four ways; see \autoref{fig:timing-LiftLetsMap}.
+\subsection{Rewriting Under Binders: UnderLetsPlus0}\label{sec:UnderLetsPlus0-more}\label{sec:micro:UnderLetsPlus0}
+
+Consider now the code in \autoref{fig:micro:UnderLetsPlus0:code}, which is a version of the code above where redundant expressions are shared via \mintinline{coq}{let} bindings.
+
+The code used to define this microbenchmark is
+\begin{minted}[fontsize=\small]{coq}
+Definition make_lets_def (n:nat) (v acc : Z) :=
+ @nat_rect (fun _ => Z * Z -> Z)
+   (fun '(v, acc) => acc + acc + v)
+   (fun _ rec '(v, acc) =>
+     dlet acc := acc + acc + v in rec (v, acc))
+   n
+   (v, acc).
+\end{minted}
+We note some details of the rewriting framework that were glossed over in the main body of the paper, which are useful for using the code:
+Although the rewriting framework does not support dependently typed constants, we can automatically preprocess uses of eliminators like \mintinline{coq}{nat_rect} and \mintinline{coq}{list_rect} into nondependent versions.
+The tactic that does this preprocessing is extensible via \Ltac{}'s reassignment feature.
+Since pattern-matching compilation mixed with NbE requires knowing how many arguments a constant can be applied to, we must internally use a version of the recursion principle whose type arguments do not contain arrows; current preprocessing can handle recursion principles with either no arrows or one arrow in the motive.
+Even though we will eventually plug in 0 for $v$, we jump through some extra hoops to ensure that our rewriter cannot cheat by rewriting away the ${}+0$ before reducing the recursion on $n$.
+
+We can reduce this expression in three ways.
+
+\subsubsection{Our Rewriter}
+One lemma is required for rewriting with our rewriter:
+\begin{minted}[fontsize=\small]{coq}
+Lemma Z.add_0_r : forall z, z + 0 = z.
+\end{minted}
+
+Creating the rewriter takes about 12 seconds on the machine we used for running the performance experiments:
+\begin{minted}[fontsize=\small]{coq}
+Make myrew := Rewriter For (Z.add_0_r, eval_rect nat, eval_rect prod).
+\end{minted}
+Recall from \autoref{sec:explain-eval-rect} that \mintinline{coq}{eval_rect} is a definition provided by our framework for eagerly evaluating recursion associated with certain types.
+It functions by triggering typeclass resolution for the lemmas reducing the recursion principle associated to the given type.
+We provide instances for \texttt{nat}, \texttt{prod}, \texttt{list}, \texttt{option}, and \texttt{bool}.
+Users may add more instances if they desire.
+
+\subsubsection{\texorpdfstring{\texttt{setoid\_rewrite}}{setoid\_rewrite} and \texorpdfstring{\texttt{rewrite\_strat}}{rewrite\_strat}}
+To give as many advantages as we can to the preexisting work on rewriting, we pre-reduce the recursion on \mintinline{coq}{nat}s using \taccbv{} before performing \texttt{setoid\_rewrite}.
+(Note that \texttt{setoid\_rewrite} cannot itself perform reduction without generating large proof terms, and \texttt{rewrite\_strat} is not currently capable of sequencing reduction with rewriting internally due to bugs such as \coqbug{10923}.)
+Rewriting itself is easy; we may use any of \texttt{repeat setoid\_rewrite Z.add\_0\_r}, \texttt{rewrite\_strat topdown Z.add\_0\_r}, or \texttt{rewrite\_strat bottomup Z.add\_0\_r}.
+
+\subsection{Binders and Recursive Functions: LiftLetsMap}\label{sec:LiftLetsMap-more}\label{sec:micro:LiftLetsMap}
+
+\begin{figure}
+  {\small %\allowdisplaybreaks
+  $\begin{aligned}
+    \text{map\_dbl}(\ell) & \defeq \begin{cases} [] & \text{if }\ell = [] \\
+        \letin[{y := h + h}{}] & \text{if }\ell = h::t \\
+        y :: \text{map\_dbl}(t) &
+        \end{cases} \\
+    \text{make}(n, m, v) & \defeq \begin{cases} [\underbrace{v, \ldots, v}_n] & \text{if }m = 0 \\
+        \text{map\_dbl}(\text{make}(n, m-1, v)) & \text{if }m > 0
+        \end{cases} \\
+    \text{example}_{n, m} & \defeq \forall v,\ \text{make}(n, m, v) = []
+  \end{aligned}$}%
+  \caption{\label{fig:micro:LiftLetsMap:code}Initial code for binders and recursive functions}
+\end{figure}
+
+\begin{figure}
+  \adjustbox{valign=t}{\resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
+      \einput{perf-UnderLetsPlus0.csv.md5}
+  }
+    \pgfplotsset{every axis legend/.append style={
+        at={(0.5,-0.2)},
+        anchor=north}}
+    \begin{axis}[xlabel=$n\cdot m$,
+        ylabel=time (s),
+        scaled x ticks=false,
+        ymax=27,
+        xmax=12000,
+        table/col sep=comma,
+        table/x=param-0-nm]
+      \addplot[mark=o,color=red]         table[y=rewrite-strat(bottomup-bottomup)-regression-exponential-user]{perf-LiftLetsMap.csv};
+      \addplot[mark=triangle,color=red]  table[y=rewrite-strat(topdown-bottomup)-regression-exponential-user]{perf-LiftLetsMap.csv};
+      \addplot[mark=square,color=red]    table[y=setoid-rewrite-regression-cubic-user]{perf-LiftLetsMap.csv};
+      \addplot[mark=+,color=blue]        table[y=Rewrite-for-gen-user]{perf-LiftLetsMap.csv};
+      \addplot[mark=x,color=ForestGreen] table[y=rewriting-user]{perf-LiftLetsMap.csv};
+      \addplot[mark=*,color=purple]      table[y=cps+vm-compute-regression-quadratic-user]{perf-LiftLetsMap.csv};
+      \legend{rewrite\_strat bottomup,rewrite\_strat topdown,repeat setoid\_rewrite,{Our approach including reification, cbv, etc.},Our approach (only rewriting),cps+vm\_compute}
+    \end{axis}
+  \end{tikzpicture}}}
+  \caption{\label{fig:timing-LiftLetsMap}Benchmark with recursive functions}
+\end{figure}
+
+The next experiment uses the code in \autoref{fig:micro:LiftLetsMap:code}.
+%Note that the ${}=[]$ at the end is just a dummy placeholder to allow us to check reduction and rewriting of this term.
+Note that the \letin{} binding blocks further reduction of map\_dbl when we iterate it $m$ times in make, and so we need to take care to preserve sharing when reducing here.
+
+\autoref{fig:timing-LiftLetsMap} compares performance between our approach, \texttt{repeat setoid\_rewrite}, and two variants of \texttt{rewrite\_strat}.
+Additionally, we consider another option, which was adopted by Fiat Cryptography at a larger scale: rewrite our functions to improve reduction behavior.
+% POPL 2022 review TODO:
+% line 958 (and elsewhere), "continuation-passing style [...] allows standard VM-based reduction to achieve good performance".
+% Is it folklore knowledge why writing in CPS style helps?
+% It might be worth briefly recalling why this is so.
+Specifically, both functions are rewritten in continuation-passing style, which makes them harder to read and reason about but allows standard VM-based reduction to achieve good performance.
+The figure shows that \texttt{rewrite\_strat} variants are essentially unusable for this example, with \texttt{setoid\_rewrite} performing only marginally better, while our approach applied to the original, more readable definitions loses ground steadily to VM-based reduction on CPS'd code.
+On the largest terms ($n \cdot m > 20,000$), the gap is 6s vs.\ 0.1s of compilation time, which should often be acceptable in return for simplified coding and proofs, plus the ability to mix proved rewrite rules with built-in reductions.
+Note that about 99\% of the difference between the full time of our method and just the rewriting is spent in the final \taccbv{} at the end, used to denote our output term from reified syntax.
+We blame this performance on the unfortunate fact that reduction in Coq is quadratic in the number of nested binders present; see Coq bug \coqbug{11151}.
+This bug has since been fixed, as of Coq 8.14; see Coq PR~\coqpr{13537}.
+%traded for simplified coding and proofs, plus the ability to mix proved rewrite rules with built-in reductions.
+
+We can perform this rewriting in four ways.
 
@@
 Throughout most of this paper, anywhere that \letin{} appears, we have actually used \mintinline{coq}{Let_In} in the code.
-It would alternatively be possible to extend the reification preprocessor to automatically convert \letin{} to \mintinline{coq}{Let_In}, but this may cause problems when converting the interpretation of the reified term with the pre-reified term, as Coq's conversion does not allow fine-tuning of when to inline or unfold \mintinline{coq}{let}s.
+It would alternatively be possible to extend the reification preprocessor to automatically convert \letin{} to \mintinline{coq}{Let_In}, but this may cause problems when converting the interpretation of the reified term with the prereified term, as Coq's conversion does not allow fine-tuning of when to inline or unfold \mintinline{coq}{let}s.
 
@@
 
-Then we can just run \texttt{vm\_compute}.
+Then we can just run \tacvmcompute{}.
 Note that this strategy, while quite fast, results in a stack overflow when $n \cdot m$ is larger than approximately $2.5\cdot 10^4$.
@@
 From this example, we conclude that \texttt{rewrite\_strat} is unsuitable for computations involving large terms with many binders, especially in cases where reduction and rewriting need to be interwoven, and that the many bugs in \texttt{rewrite\_strat} result in confusing gymnastics required for success.
-The prior state of the art---writing code in CPS---suitably tweaked by using module pacity to allow \texttt{vm\_compute}, remains the best performer here, though the cost of rewriting everything is CPS may be prohibitive.
+The prior state of the art---writing code in CPS---suitably tweaked by using module opacity to allow \tacvmcompute{}, remains the best performer here, though the cost of rewriting everything is CPS may be prohibitive.
 Our method soundly beats \texttt{rewrite\_strat}.
-We are additionally bottlenecked on \texttt{cbv}, which is used to unfold the goal post-rewriting and costs about a minute on the largest of terms; see Coq bug \coqbug{11151} for a discussion on what is wrong with Coq's reduction here.
-
-\subsection{SieveOfEratosthenes} \label{sec:Eratosthenes}
+We are additionally bottlenecked on \taccbv{}, which is used to unfold the goal post-rewriting and costs about a minute on the largest of terms; see Coq bug \coqbug{11151} for a discussion on what is wrong with Coq's reduction here.
+
+\subsection{SieveOfEratosthenes}\label{sec:Eratosthenes}\label{sec:micro:Eratosthenes}
+
+The final experiment involves full reduction in computing the Sieve of Eratosthenes, taking inspiration on benchmark choice from \citet{Aehlig}.
+We find in \autoref{fig:timing-SieveOfEratosthenes} that we are slower than \tacvmcompute{}, \tacnativecompute{}, and \taccbv{}, but faster than \mintinline{coq}{lazy}, and of course much faster than \tacsimpl{} and \taccbn{}, which are quite slow.
+
+\begin{figure}
+  \resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
+      \einput{perf-SieveOfEratosthenes.csv.md5}
+  }
+    \pgfplotsset{every axis legend/.append style={
+        at={(0.5,-0.2)},
+        anchor=north}}
+    \begin{axis}[xlabel=upper bound,
+        ylabel=time (s),
+        ymax=65,
+        xmax=5000,
+        table/col sep=comma,
+        table/x=param-n]
+      \addplot[mark=square,color=red]       table[y=simpl-regression-cubic-user]                {perf-SieveOfEratosthenes.csv};
+      \addplot[mark=triangle,color=red]     table[y=cbn-regression-cubic-user]                  {perf-SieveOfEratosthenes.csv};
+      \addplot[mark=square*,color=red]      table[y=lazy-regression-quadratic-user]             {perf-SieveOfEratosthenes.csv};
+      \addplot[mark=+,color=blue]           table[y=Rewrite-for-gen-user]                       {perf-SieveOfEratosthenes.csv};
+      \addplot[mark=x,color=ForestGreen]    table[y=rewriting-user]                             {perf-SieveOfEratosthenes.csv};
+      \addplot[mark=triangle*,color=orange] table[y=cbv-regression-quadratic-user]              {perf-SieveOfEratosthenes.csv};
+      \addplot[mark=*,color=purple]         table[y=vm-compute-regression-quadratic-user]       {perf-SieveOfEratosthenes.csv};
+      \addplot[mark=-,color=red]            table[y=native-compute(2)-regression-quadratic-real]{perf-SieveOfEratosthenes.csv};
+      \legend{simpl,cbn,lazy,{Our approach including reification, cbv, etc.},Our approach (only rewriting),cbv,vm\_compute,native\_compute}
+    \end{axis}
+  \end{tikzpicture}}
+  \caption{\label{fig:timing-SieveOfEratosthenes}Full evaluation, Sieve of Eratosthenes}
+\end{figure}
 
@@
 
-To get \texttt{cbn} and \texttt{simpl} to unfold our term fully, we emit
+To get \taccbn{} and \tacsimpl{} to unfold our term fully, we emit
 \begin{minted}[fontsize=\small]{coq}
@@
 
+\FloatBarrier\section{Fusing Compiler Passes}\label{sec:fusing-compiler-passes}
+
+When we moved the
+%aforementioned constant-folding rules
+%constant-folding rules from \autoref{sec:iteration:carries}
+constant-folding rules
+from before abstract interpretation to after it, the performance of our compiler on Word-by-Word Montgomery code synthesis decreased significantly.
+(The generated code did not change.)
+We discovered that the number of variable assignments in our intermediate code was quartic in the number of bits in the prime, while the number of variable assignments in the generated code is only quadratic.
+The performance numbers we measured supported this theory: the overall running time of synthesizing code for a prime near $2^k$ jumped from $\Theta(k^2)$ to $\Theta(k^4)$ when we made this change.
+We believe that fusing abstract interpretation with rewriting and partial evaluation would allow us to fix this asymptotic-complexity issue.
+
+To make this situation more concrete, consider the following example:
+Fiat Cryptography uses abstract interpretation to perform bounds analysis; each expression is associated with a range that describes the lower and upper bounds of values that expression might take on.
+Abstract interpretation on addition works as follows: if we have that $x_\ell \le x \le x_u$ and $y_\ell \le y \le y_u$, then we have that $x_\ell + y_\ell \le x + y \le x_u + y_u$.
+Performing bounds analysis on $+$ requires two additions.
+We might have an arithmetic simplification that says that $x + y = x$ whenever we know that $0 \le y \le 0$.
+If we perform the abstract interpretation and then the arithmetic simplification, we perform two additions (for the bounds analysis) and then two comparisons (to test the lower and upper bounds of $y$ for equality with 0).
+We cannot perform the arithmetic simplification before abstract interpretation, because we will not know the bounds of $y$.
+However, if we perform the arithmetic simplification for each expression after performing bounds analysis on its \emph{subexpressions} and only after this perform abstract interpretation on the resulting expression, then we need not use any additions to compute the bounds of $x + y$ when $0 \le y \le 0$, since the expression will just become $x$.
+
+Another essential pass to fuse with rewriting and partial evaluation is let-lifting.
+Unless all of the code is CPS-converted ahead of time, attempting to do let-lifting via rewriting, as must be done when using \mintinline{coq}{setoid_rewrite}, \mintinline{coq}{rewrite_strat}, or \Rtac, results in slower asymptotics.
+This pattern is already apparent in the \texttt{LiftLetsMap} / ``Binders and Recursive Functions'' example in \autoref{sec:micro:LiftLetsMap}.
+% POPL 2022 review TODO:
+% line 852, the forward pointer to section 5.2.4 is somewhat unfortunate; the text cannot be fully understood without looking ahead, it seems.
+% We do not know what the parameters `n` and `m` represents.
+% On line 853, it is not clear what "cubic" means (cubic in what parameter?).
+% On line 858, what is `make`?
+% What is `map_dbl`? The whole paragraph (848-861) is extremely technical.
+% It does not clearly explain why it is the fundamentally difficult to perform let-lifting in an efficient way.
+% It also does not explain its opening sentence, "Another essential pass to fuse with rewriting and partial evaluation is let-lifting".
+% Why is it essential?
+% Is it more efficient if fused than if performed separately?
+% Why is that?
+We achieve linear performance in $n\cdot m$ when ignoring the final \taccbv, while \mintinline{coq}{setoid_rewrite} and \mintinline{coq}{rewrite_strat} are both cubic.
+The rewriter in \Rtac\space cannot possibly achieve better than $\mathcal{O}\left(n\cdot m^2\right)$ unless it can be sublinear in the number of rewrites, because our rewriter gets away with a constant number of rewrites (four), plus evaluating recursion principles for a total amount of work $\mathcal{O}(n\cdot m)$.
+But without primitive support for let-lifting, it is instead necessary to lift the lets by rewrite rules, which requires
+$\mathcal{O}\left(n\cdot m^2\right)$ rewrites just to lift the lets.
+The analysis is thus: running \texttt{make} simply gives us $m$ nested applications of \texttt{map\_dbl} to a length-$n$ list.
+To reduce a given call to \texttt{map\_dbl}, all existing let-binders must first be lifted (there are $n\cdot k$ of them on the $k$-innermost-call) across \texttt{map\_dbl}, one-at-a-time.
+Then the \texttt{map\_dbl} adds another $n$ let binders, so we end up doing $\sum_{k=0}^{m} n\cdot k$ lifts, i.e., $n\cdot m(m+1)/2$ rewrites just to lift the lets.
+
 \section{Experience vs.\ Lean and \texorpdfstring{\texttt{setoid\_rewrite}}{setoid\_rewrite}\label{sec:lean}}
@@
 Although all of our toy examples work with \texttt{setoid\_rewrite} or \texttt{rewrite\_strat} (until the terms get too big), even the smallest of examples in Fiat Cryptography fell over using these tactics.
-When attempting to use \texttt{rewrite\_strat} for partial evaluation and rewriting on unsaturated Solinas with 1 limb on small primes (such as 29), we were able to get \texttt{rewrite\_strat} to finish after about 90 seconds.
+When attempting to use \texttt{setoid\_rewrite} for partial evaluation and rewriting on unsaturated Solinas with 1 limb on small primes (such as $2^{61}-1$), we were able to get \texttt{setoid\_rewrite} to finish after about 100 seconds.
+Trying to synthesize code for two limbs on slightly larger primes (such as $2^{107}-1$, which needs two limbs on a 64-bit machine) took about 10 minutes;
+three limbs took just under 3.5 hours, and four limbs failed to synthesize with an out-of-memory error after using over 60 GB of RAM.
+The widely used primes tend to have around five to ten limbs.
+See \coqbug{13576} for more details and for updates.
+
+The \texttt{rewrite\_strat} tactic, which does not require duplicating the entire goal at each rewriting step, fared a bit better.
+Small primes with 1 limb took about 90 seconds, but further performance tuning of the typeclass instances dropped this time down to 11 seconds.
 The bugs in \texttt{rewrite\_strat} made finding the right magic invocation quite painful, nonetheless; the invocation we settled on involved \emph{sixteen} consecutive calls to \texttt{rewrite\_strat} with varying arguments and strategies.
-Trying to synthesize code for two limbs on slightly larger primes (such as 113, which needs two limbs on a 64-bit machine) took about three hours.
-The widely used primes tend to have around five to ten limbs; we leave extrapolating this slowdown to the reader.
-
-We have attached this experiment using \verb|rewrite_strat| as \verb|fiat_crypto_via_rewrite_strat.v|, which is meant to be run in emacs/PG from inside the \verb|fiat-crypto| directory, or in \verb|coqc| by setting \verb|COQPATH| to the value emitted by \texttt{make printenv} in \verb|fiat-crypto| and then invoking the command \texttt{coqc -q -R /path/to/fiat-crypto/src Crypto /path/to/fiat\_crypto\_via\_rewrite\_strat.v}.
-To test with the two-limb prime 113, change \verb|of_string "2^5-3" 8| in the definition of \verb|p| to \verb|of_string "2^7-15" 64|.
+Two limbs took about 90 seconds, three limbs took a bit under 10 minutes, and four limbs took about 70 minutes and about 17 GB of RAM.
+Extrapolating out the exponential asymptotics of the fastest-growing subcall to \texttt{rewrite\_strat} indicates that 5 limbs would take 11--12 hours, 6 limbs would take 10--11 days, 7 limbs would take 31--32 weeks, 8 limbs would take 13--14 years, 9 limbs would take 2--3 centuries, 10 limbs would take 6--7 millennia, and 15 limbs would take 2--3 times the age of the universe, and 17 limbs, the largest example we might find at present in the real world, would take over $1000\times$ the age of the universe!
+See \coqbug{13708} for more details and updates.
+
+This experiment using \verb|rewrite_strat| can be found online in the Coq source file at \texttt{src/fiat\_crypto\_via\_setoid\_rewrite\_standalone.v} on GitHub at \githublink{coq-community}{coq-performance-tests}.
+To test with the two-limb prime $2^{107}-1$, change \verb|Goal goal| to \verb|Goal goal_of_size 2%nat| near the bottom of the file.
 
 We also tried Lean, in the hopes that rewriting in Lean, specifically optimized for performance, would be up to the challenge.
-Although Lean performed about 30\% better than Coq on the 1-limb example, taking a bit under a minute, it did not complete on the two-limb example even after four hours (after which we stopped trying), and a five-limb example was still going after 40 hours.
-
-We have attached our experiments with running \texttt{rewrite} in Lean on the Fiat Cryptography code as a supplement as well.
+Although Lean performed about 30\% better than Coq's \texttt{setoid\_rewrite} on the 1-limb example, taking a bit under a minute, it did not complete on the two-limb example even after four hours (after which we stopped trying), and a five-limb example was still going after 40 hours.
+
+Our experiments with running \texttt{rewrite} in Lean on the Fiat Cryptography code can be found in the file \texttt{fiat-crypto-lean/src/fiat\_crypto.lean} on GitHub at \githublink{mit-plv}{fiat-crypto}[lean].
 We used Lean version 3.4.2, commit cbd2b6686ddb, Release.
@@
 
-\section{Reading the Code Supplement} \label{sec:CodeSupplement-more}
+\section{Limitations and Preprocessing}\label{sec:implementation-and-usage}
+
+%\todo{It's not clear to Jason what in this section belongs here, and what isn't interesting (yet?) to readers}
+%\todo{I (Jason) have no idea what the story for this section should be}
+%\todo{Insofar as text is actually kept in this section, make sure any duplication between here and the out-of-paper appendix is deliberate}
+%\todo{If this section is moved or removed, be sure to restructure \autoref{sec:code-from-implementation-and-usage} accordingly}
+
+We now note some details of the rewriting framework that were previously glossed over, which are useful for using the code or implementing something similar, but which do not add fundamental capabilities to the approach.
+Although the rewriting framework does not support dependently typed constants, we can automatically preprocess uses of eliminators like \mintinline{coq}{nat_rect} and \mintinline{coq}{list_rect} into nondependent versions.
+The tactic that does this preprocessing is extensible via \Ltac{}'s reassignment feature.
+Since pattern-matching compilation mixed with NbE requires knowing how many arguments a constant can be applied to, internally we must use a version of the recursion principle whose type arguments do not contain arrows; current preprocessing can handle recursion principles with either no arrows or one arrow in motives.
+
+Recall from \autoref{sec:explain-eval-rect} that \mintinline{coq}{eval_rect} is a definition provided by our framework for eagerly evaluating recursion associated with certain types.
+It functions by triggering typeclass resolution for the lemmas reducing the recursion principle associated to the given type.
+We provide instances for \texttt{nat}, \texttt{prod}, \texttt{list}, \texttt{option}, and \texttt{bool}.
+Users may add more instances if they desire.
+
+Recall again from \autoref{sec:explain-ident.eagerly} that we use \mintinline{coq}{ident.eagerly} to ask the reducer to simplify a case of primitive recursion by complete traversal of the designated argument's constructor tree.
+Our current version only allows a limited, hard-coded set of eliminators with \mintinline{coq}{ident.eagerly} (\texttt{nat\_rect} on return types with either zero or one arrows, \texttt{list\_rect} on return types with either zero or one arrows, and \texttt{List.nth\_default}), but nothing in principle prevents automatic generation of the necessary code.
+
+We define a constant \mintinline{coq}{Let_In} which we use for writing \letin{} expressions that do not reduce under $\zeta$ (Coq's reduction rule for \mintinline{coq}{let}-inlining).
+Throughout most of this paper, anywhere that \letin{} appears, we have actually used \mintinline{coq}{Let_In} in the code.
+It would alternatively be possible to extend the reification preprocessor to automatically convert \letin{} to \mintinline{coq}{Let_In}, but this strategy may cause problems when converting the interpretation of the reified term with the prereified term, as Coq's conversion does not allow fine-tuning of when to inline or unfold \mintinline{coq}{let}s.
+
+\section{Reading the Code Supplement}\label{sec:CodeSupplement-more}
 
@@
 We have attached both the code for implementing the rewriter, as well as a copy of Fiat Cryptography adapted to use the rewriting framework.
-Both code supplements build with Coq 8.9 and Coq 8.10, and they require that whichever OCaml was used to build Coq be installed on the system to permit building plugins.
+Both code supplements build with Coq versions 8.9--8.13, and they require that whichever OCaml was used to build Coq be installed on the system to permit building plugins.
 (If Coq was installed via opam, then the correct version of OCaml will automatically be available.)
@@
 
-\subsubsection{Code from \autocommanameref{sec:motivating-example}}
-
-The \texttt{prefixSums} example appears in the Coq source file \texttt{rewriter/src/Rewriter/Rewriter/Examples/PrefixSums.v}.
-Note that we use \texttt{dlet} rather than \mintinline{coq}{let} in binding \texttt{acc'} so that we can preserve the \mintinline{coq}{let} binder even under $\iota$ reduction, which much of Coq's infrastructure performs eagerly.
-Because we attempt to isolate the dependency on the axiom of functional extensionality as much as possible, we also in practice require \texttt{Proper} instances for each higher-order identifier saying that each constant respects function extensionality.
-We hope to remove the dependency on function extensionality altogether in the future.
-Although we glossed over this detail in the body of this paper, we also prove
-\begin{minted}[fontsize=\small]{coq}
-Global Instance: forall A B,
- Proper ((eq ==> eq ==> eq) ==> eq ==> eq ==> eq)
-        (@fold_left A B).
-\end{minted}
-
-The \mintinline{coq}{Make} command is exposed in \texttt{rewriter/src/Rewriter/Util/plugins/RewriterBuild.v} and defined in \texttt{rewriter/src/Rewriter/Util/plugins/rewriter\_build\_plugin.mlg}.
-Note that one must run \texttt{make} to create this latter file; it is copied over from a version-specific file at the beginning of the build.
-
-\label{sec:code:eval-rect}%
-\label{sec:code:ident.eagerly}%
-The \verb|do_again|, \verb|eval_rect|, and \verb|ident.eagerly| constants are defined at the bottom of module \verb|RewriteRuleNotations| in \texttt{rewriter/src/Rewriter/Language/Pre.v}.
-
-\subsubsection{Code from \autocommanameref{sec:trusted-code-base-size}}
+The P-384 curve is mentioned.
+This is the curve with modulus $2^{384} - 2^{128} - 2^{96} + 2^{32} - 1$; its benchmarks can be found in files matching the glob \texttt{fiat-crypto/src/Rewriter/PerfTesting/Specific/generated/p2384m2128m296p232m1\_\_*\_\_word\_by\_word\_montgomery\_*}.
+The output \texttt{.log} files are included in the tarball; the \texttt{.v} and \texttt{.sh} files are automatically generated in the course of running \texttt{make perf -k}.
+
+\subsubsection{Code from \autocommanameref{sec:related-one}}
 
@@
   \item
-    We claimed that we \textbf{did not grow the trusted base} (excepting the axiom of functional extensionality).
+    We claimed that we \textbf{did not grow the trusted code base}.
     In any example file (of which a couple can be found in \texttt{rewriter/src/Rewriter/Rewriter/Examples/}), the \mintinline{coq}{Make} command creates a rewriter package.
-    Running \texttt{Print Assumptions} on this new constant (often named \texttt{rewriter} or \texttt{myrew}) should demonstrate a lack of axioms other than functional extensionality.
+    Running \texttt{Print Assumptions} on this new constant (often named \texttt{rewriter} or \texttt{myrew}) should demonstrate a lack of axioms.
     \texttt{Print Assumptions} may also be run on the proof that results from using the rewriter.
@@
     We claimed to allow reduction that \textbf{mixes} \emph{rules of the definitional equality} with \emph{equalities proven explicitly as theorems}; the ``rules of the definitional equality'' are, for example, $\beta$ reduction, and we assert that it should be self-evident that our rewriter supports this.
+  \item
+    We claimed to allow \textbf{rapid iteration} on rewrite rules with \emph{minimal verification overhead}.
+    We invite the reader to alter the list of constants in any of the \mintinline{coq}{Make ...
+:= Rewriter For ...} invocations in \texttt{rewriter/src/Rewriter/Rewriter/Examples/} or to alter the list of rewrite rules in \texttt{fiat-crypto/src/Rewriter/Rules.v} to experience iteration on rewrite rules.
   \item
@@
 
-The P-384 curve is mentioned.
-This is the curve with modulus $2^{384} - 2^{128} - 2^{96} + 2^{32} - 1$; its benchmarks can be found in files matching the glob \texttt{fiat-crypto/src/Rewriter/PerfTesting/Specific/generated/p2384m2128m296p232m1\_\_*\_\_word\_by\_word\_montgomery\_*}.
-The output \texttt{.log} files are included in the tarball; the \texttt{.v} and \texttt{.sh} files are automatically generated in the course of running \texttt{make perf -k}.
-
-We mention integration with abstract interpretation; the abstract-interpretation pass is implemented in \texttt{fiat-crypto/src/AbstractInterpretation/}.
-
-\subsection{Code from \autocommanameref{sec:trust}}
-
-The individual rewritings mentioned are implemented via the \texttt{Rewrite\_*} tactics exported at the top of \texttt{rewriter/src/Rewriter/Util/plugins/RewriterBuild.v}.
-These tactics bottom out in tactics defined at the bottom of \texttt{rewriter/src/Rewriter/Rewriter/AllTactics.v}.
-
-\subsubsection{Code from \autocommanameref{sec:nine-steps}} \label{sec:code:nine-steps}
+We mention encoding pattern matching explicitly by adopting the performance-tuned approach of \citet{maranget2008compiling}; the code for this is in \texttt{rewriter/src/Rewriter/Rewriter/Rewriter.v} starting from the comment above \mintinline{coq}{Inductive decision_tree} and including the Gallina definitions \mintinline{coq}{eval_decision_tree} and \mintinline{coq}{compile_rewrites}.
+
+We mention integration with abstract interpretation; the abstract-interpretation pass is implemented in \texttt{fiat-crypto/src/AbstractInterpretation/}; integration is achieved in rewrite rules in \texttt{fiat-crypto/src/Rewriter/Rules.v} making use of the various \mintinline{coq}{Local Notation}s defined in that file for \mintinline{coq}{ident.cast}.
+
+We mention parametric higher-order abstract syntax (PHOAS); the definition of our datatype is \mintinline{coq}{Inductive expr} in module \mintinline{coq}{Compilers.expr} in \texttt{rewriter/src/Rewriter/Language/Language.v}.
+We mention a let-lifting transformation threaded throughout reduction; this is \mintinline{coq}{Inductive UnderLets}, a monad defined in module \mintinline{coq}{Compilers.UnderLets} in the file \texttt{rewriter/src/Rewriter/Language/UnderLets.v}.
+
+\subsection{Code from \autocommanameref{sec:motivating-example}}
+
+The \texttt{prefixSums} example appears in the Coq source file \texttt{rewriter/src/Rewriter/Rewriter/Examples/PrefixSums.v}.
+Note that we use \texttt{dlet} rather than \mintinline{coq}{let} in binding \texttt{acc'} so that we can preserve the \mintinline{coq}{let} binder even under $\iota$ reduction, which much of Coq's infrastructure performs eagerly.
+Because we do not depend on the axiom of functional extensionality, we also in practice require \texttt{Proper} instances for each higher-order identifier saying that each constant respects function extensionality.
+Although we glossed over this detail in the body of this paper, we also prove
+\begin{minted}[fontsize=\small]{coq}
+Global Instance: forall A B,
+ Proper ((eq ==> eq ==> eq) ==> eq ==> eq ==> eq)
+        (@fold_left A B).
+\end{minted}
+
+The \mintinline{coq}{Make} command is exposed in \texttt{rewriter/src/Rewriter/Util/plugins/RewriterBuild.v} and defined in \texttt{rewriter/src/Rewriter/Util/plugins/rewriter\_build\_plugin.mlg}.
+Note that one must run \texttt{make} to create this latter file; it is copied over from a version-specific file at the beginning of the build.
+
+\label{sec:code:eval-rect}%
+\label{sec:code:ident.eagerly}%
+The \verb|do_again|, \verb|eval_rect|, and \verb|ident.eagerly| constants are defined at the bottom of module \verb|RewriteRuleNotations| in \texttt{rewriter/src/Rewriter/Language/Pre.v}.
+
+\subsection{Code from \autocommanameref{sec:structure}}
+
+%The individual rewritings mentioned are implemented via the \texttt{Rewrite\_*} tactics exported at the top of \texttt{rewriter/src/Rewriter/Util/plugins/RewriterBuild.v}.
+%These tactics bottom out in tactics defined at the bottom of \texttt{rewriter/src/Rewriter/Rewriter/AllTactics.v}.
+
+\subsubsection{Code from \autocommanameref{sec:nine-steps}}\label{sec:code:nine-steps}
 
@@
     We use a theorem stating that rewriting preserves denotations of well-formed terms to replace the denotation subterm with the denotation of the rewriter applied to the same reified term.
-    We use Coq's built-in full reduction (\texttt{vm\_compute}) to reduce the application of the rewriter to the reified term.
+    We use Coq's built-in full reduction (\tacvmcompute{}) to reduce the application of the rewriter to the reified term.
     This step is performed by the tactic \texttt{do\_rewrite\_with} in \texttt{rewriter/src/Rewriter/Rewriter/AllTactics.v}.
   \item
-    Finally, we run \texttt{cbv} (a standard call-by-value reducer) to simplify away the invocation of the denotation function on the concrete syntax tree from rewriting.
+    Finally, we run \taccbv{} (a standard call-by-value reducer) to simplify away the invocation of the denotation function on the concrete syntax tree from rewriting.
     This step is performed by the tactic \texttt{do\_final\_cbv} in \texttt{rewriter/src/Rewriter/Rewriter/AllTactics.v}.
@@
 
-\subsubsection{Our Approach in More Than Nine Steps}
+The expression language $e$ corresponds to the inductive \texttt{expr} type defined in the module \texttt{Compilers.expr} in \texttt{rewriter/src/Rewriter/Language/Language.v}.
+
+\subsubsection*{Our Approach in More Than Nine Steps}
 
@@
     We use a lemma stating that rewriting preserves denotations of well-formed terms to replace the goal with the rewriter applied to our reified term.
-    We use \texttt{vm\_compute} to prove the well-formedness side condition reflectively.
-    We use \texttt{vm\_compute} again to reduce the application of the rewriter to the reified term.
-    \item
-    Finally, we run \texttt{cbv} to unfold the denotation function, and we instantiate the evar with the resulting rewritten term.
+    We use \tacvmcompute{} to prove the well-formedness side condition reflectively.
+    We use \tacvmcompute{} again to reduce the application of the rewriter to the reified term.
+    \item
+    Finally, we run \taccbv{} to unfold the denotation function, and we instantiate the evar with the resulting rewritten term.
 \end{enumerate}
 
-There are a couple of steps that contribute to the trusted base.
+There are a couple of steps that contribute to the trusted code base.
 We must trust that the rewriter package we generate from the rewrite rules in fact matches the rewrite rules we want to rewrite with.
@@
 Otherwise, everything is checked by Coq.
-We do, however, depend on the axiom of function extensionality in one place in the rewriter proof; after spending a couple of hours trying to remove this axiom, we temporarily gave up.
-
-\subsection{Code from \autocommanameref{sec:structure}}
-
-The expression language $e$ corresponds to the inductive \texttt{expr} type defined in module \texttt{Compilers.expr} in \texttt{rewriter/src/Rewriter/Language/Language.v}.
 
@@
 
-\subsubsection{Code from \autocommanameref{sec:implementation-and-usage}} \label{sec:code-from-implementation-and-usage}
+\subsection{Code from \autocommanameref{sec:evaluation}}
+
+\subsubsection{Code from \autocommanameref{sec:iteration}}
+The old continuation-passing-style versions of verified arithmetic functions can be found in the folder \texttt{fiat-crypto/src/ArithmeticCPS/}, while the new versions can be found in the folder \texttt{fiat-crypto/src/Arithmetic/}.
+
+The rewrite rules
+%from \autocommanameref{sec:iteration:reassoc}
+for reassociating arithmetic
+can be found in \mintinline{coq}{arith_rewrite_rulesT} starting at the comment ``We reassociate some multiplication of small constants'' in \texttt{fiat-crypto/src/Rewriter/Rules.v}.
+
+The
+following
+frontend constructs
+%from \autocommanameref{sec:iteration:new-frontend}
+are in \verb|all_ident_named_interped| defined in \texttt{fiat-crypto/src/Language/IdentifierParameters.v}.
+\begin{itemize}
+\item
+  The multiplication primitives are \mintinline{coq}{with_name ident_Z_mul_split Z.mul_split} as well as \mintinline{coq}{with_name ident_Z_mul_high Z.mul_high}, as well as the various Coq expressions \mintinline{coq}{with_name ident_fancy_mulXX ident.fancy.mulXX} for each \mintinline{coq}{X} being either \texttt{l} or \texttt{h}.
+\item
+  The ``comment'' function is both \mintinline{coq}{with_name ident_comment (@ident.comment)} as well as \mintinline{coq}{with_name ident_comment_no_keep (@ident.comment_no_keep)}.
+\item
+  The bitwise exclusive-or is \mintinline{coq}{with_name ident_Z_lxor Z.lxor}.
+\item
+  The special identity function which prints in the backend as a call to some inline assembly is \mintinline{coq}{with_name ident_value_barrier (@Z.value_barrier)}.
+\end{itemize}
+
+The rules about bitmasking operations
+%from \autocommanameref{sec:iteration:new-middle-end}
+can be found in \mintinline{coq}{arith_with_casts_rewrite_rulesT} in \texttt{fiat-crypto/src/Rewriter/Rules.v} and involve \mintinline{coq}{Z.land} and \mintinline{coq}{Z.lor}.
+
+The compiler configuration about conditional-move instructions is the flag \texttt{--cmovznz-by-mul} defined in \texttt{fiat-crypto/src/CLI.v}.
+The \texttt{if}-statement using the thus-defined \mintinline{coq}{use_mul_for_cmovznz} is in \texttt{src/PushButtonSynthesis/Primitives.v}.
+
+The rewrite rules for the new backends
+%mentioned in \autocommanameref{sec:iteration:new-backend}
+are defined by \mintinline{coq}{fancy_with_casts_rewrite_rulesT} and \mintinline{coq}{mul_split_rewrite_rulesT} as well as \mintinline{coq}{multiret_split_rewrite_rulesT} as well as \mintinline{coq}{noselect_rewrite_rulesT} in \texttt{fiat-crypto/src/Rewriter/Rules.v}.
+The special function \mintinline{coq}{Z.combine_at_bitwidth} is defined in \texttt{fiat-crypto/src/Util/ZUtil/Definitions.v}.
+The designation of \mintinline{coq}{Z.combine_at_bitwidth} as an identifier that should be inlined occurs by listing it in the definition \mintinline{coq}{var_like_idents} in the source file \texttt{fiat-crypto/src/Language/IdentifierParameters.v}.
+
+The rules involving carries mentioned in
+%\autocommanameref{sec:iteration:carries} and
+\autocommanameref{sec:fusing-compiler-passes} are in \mintinline{coq}{arith_with_casts_rewrite_rulesT} in \texttt{fiat-crypto/src/Rewriter/Rules.v}.
+
+\subsubsection{Code from \autocommanameref{sec:micro}}
+
+This code is found in the files in \texttt{rewriter/src/Rewriter/Rewriter/Examples/}.
+We ran the microbenchmarks using the code in \texttt{rewriter/src/Rewriter/Rewriter/Examples/PerfTesting/Harness.v} together with some \texttt{Makefile} cleverness.
+
+The code for \autoref{fig:timing-Plus0Tree} from \autocommanameref{sec:micro:Plus0Tree} can be found in \texttt{Plus0Tree.v}.
+
+The code for \autoref{fig:timing-UnderLetsPlus0} from \autocommanameref{sec:micro:UnderLetsPlus0} can be found in \texttt{UnderLetsPlus0.v}.
+
+%The code used for the performance investigation mentioned in \autocommanameref{sec:micro:setoid-rewrite-bottlenecks-lite} and detailed in \autoref{sec:setoid-rewrite-bottlenecks} is not part of the framework we are presenting, and thus not in the supplement.
+
+The code for \autoref{fig:timing-LiftLetsMap} from \autocommanameref{sec:micro:LiftLetsMap} can be found in \texttt{LiftLetsMap.v}.
+
+The code for \autoref{fig:timing-SieveOfEratosthenes} from \autocommanameref{sec:micro:Eratosthenes} can be found in \texttt{SieveOfEratosthenes.v}.
+
+\subsubsection{Code from \autocommanameref{sec:macro}}
+
+The rewrite rules are defined in \texttt{fiat-crypto/src/Rewriter/Rules.v} and proven in the file \texttt{fiat-crypto/src/Rewriter/RulesProofs.v}.
+They are turned into rewriters in the various files in \texttt{fiat-crypto/src/Rewriter/Passes/}.
+The shared inductives and definitions are defined in the Coq source file \texttt{fiat-crypto/src/Language/IdentifiersBasicGENERATED.v}, the Coq source file \texttt{fiat-crypto/src/Language/IdentifiersGENERATED.v}, and finally also the Coq source file \texttt{fiat-crypto/src/Language/IdentifiersGENERATEDProofs.v}.
+Note that we invoke the subtactics of the \mintinline{coq}{Make} command manually to increase parallelism in the build and to allow a shared language across multiple rewriter packages.
+
+\subsection{Code from \autocommanameref{sec:implementation-and-usage}}\label{sec:code-from-implementation-and-usage}
 
@@
 
-The hard-coding of the eliminators for use with \mintinline{coq}{ident.eagerly} ({c.f.} \autoref{sec:code:ident.eagerly}) is done in the tactics \mintinline{coq}{reify_ident_preprocess} and \mintinline{coq}{rewrite_interp_eager} in \texttt{rewriter/src/Rewriter/Language/Language.v}, in the inductive type \mintinline{coq}{restricted_ident} and the typeclass \mintinline{coq}{BuildEagerIdentT} in \texttt{rewriter/src/Rewriter/Language/Language.v}, and in the \Ltac{} tactic \mintinline{coq}{handle_reified_rewrite_rules_interp} defined in the file \texttt{rewriter/src/Rewriter/Rewriter/ProofsCommonTactics.v}.
+The hard-coding of the eliminators for use with \mintinline{coq}{ident.eagerly} ({c.f.} \autoref{sec:code:ident.eagerly}) is done in the tactics \mintinline{coq}{reify_ident_preprocess} and \mintinline{coq}{rewrite_interp_eager} in \texttt{rewriter/src/Rewriter/Language/Language.v}, in the inductive type \mintinline{coq}{restricted_ident} and the typeclass \mintinline{coq}{BuildEagerIdentT} in \texttt{rewriter/src/Rewriter/Language/Language.v}, and in the \Ltac{} tactic with the name of \mintinline{coq}{handle_reified_rewrite_rules_interp} defined in the file \texttt{rewriter/src/Rewriter/Rewriter/ProofsCommonTactics.v}.
 
@@
 
-\subsection{Code from \autocommanameref{sec:evaluation}}
-
-\subsubsection{Code from \autocommanameref{sec:micro}}
-
-This code is found in the files in \texttt{rewriter/src/Rewriter/Rewriter/Examples/}.
-We ran the microbenchmarks using the code in \texttt{rewriter/src/Rewriter/Rewriter/Examples/PerfTesting/Harness.v} together with some \texttt{Makefile} cleverness.
-
-The code from \autocommanameref{sec:micro:Plus0Tree} can be found in \texttt{Plus0Tree.v}.
-
-The code from \autocommanameref{sec:micro:UnderLetsPlus0} can be found in \texttt{UnderLetsPlus0.v}.
-
-The code used for the performance investigation mentioned in \autocommanameref{sec:micro:setoid-rewrite-bottlenecks-lite} and detailed in \autoref{sec:setoid-rewrite-bottlenecks} is not part of the framework we are presenting, and thus not in the supplement.
-
-The code from \autocommanameref{sec:micro:LiftLetsMap} can be found in \texttt{LiftLetsMap.v}.
-
-The code from \autocommanameref{sec:micro:Eratosthenes} can be found in \texttt{SieveOfEratosthenes.v}.
-
-\subsubsection{Code from \autocommanameref{sec:macro}}
-
-The rewrite rules are defined in \texttt{fiat-crypto/src/Rewriter/Rules.v} and proven in the file \texttt{fiat-crypto/src/Rewriter/RulesProofs.v}.
-They are turned into rewriters in the various files in \texttt{fiat-crypto/src/Rewriter/Passes/}.
-The shared inductives and definitions are defined in the Coq source file \texttt{fiat-crypto/src/Language/IdentifiersBasicGENERATED.v}, the Coq source file \texttt{fiat-crypto/src/Language/IdentifiersGENERATED.v}, and finally also the Coq source file \texttt{fiat-crypto/src/Language/IdentifiersGENERATEDProofs.v}.
-Note that we invoke the subtactics of the \mintinline{coq}{Make} command manually to increase parallelism in the build and to allow a shared language across multiple rewriter packages.
-
 \end{document}

### 2022-05-03 papers/source/rewriting/popl2022/rewriting.tex
Add more final versions of 2022 ITP papers

--- 2022-05-03 papers/source/rewriting/arxiv/rewriting.tex
@@
-\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
 \RequirePackage{luatex85}
 \PassOptionsToPackage{dvipsnames}{xcolor}
+%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
 \RequirePackage{amssymb}
@@
 \RequirePackage{varioref}
-\PassOptionsToPackage{nameinlink}{cleverref}%[2012/02/15]% v0.18.4;
-\documentclass[a4paper,USenglish,cleveref,autoref,thm-restate]{lipics-v2021}
+\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 \usepackage{hyperref}
-%This is a template for producing LIPIcs articles.
-%See lipics-v2021-authors-guidelines.pdf for further information.
-%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
-%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
-%for section-numbered lemmas etc., use "numberwithinsect"
-%for enabling cleveref support, use "cleveref"
-%for enabling autoref support, use "autoref"
-%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
-%for enabling thm-restate support, use "thm-restate"
-%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
-%for producing a PDF according the PDF/A standard, add "pdfa"
-
-%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
-%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
-
-%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
-
-\bibliographystyle{plainurl}% the mandatory bibstyle
-
-%\usepackage{wrapfig}% forbidden in ITP
+\usepackage[nameinlink]{cleveref}[2012/02/15]% v0.18.4;
+%% For double-blind review submission, w/ CCS and ACM Reference
+%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
+%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
+%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%% For single-blind review submission, w/ CCS and ACM Reference
+%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
+%% For final camera-ready submission, w/ required CCS and ACM Reference
+%\documentclass[acmsmall]{acmart}\settopmatter{}
+
+%% Journal information
+%% Supplied to authors by publisher for camera-ready submission;
+%% use defaults for review submission.
+\acmJournal{PACMPL}
+\acmVolume{1}
+\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
+\acmArticle{1}
+\acmYear{2018}
+\acmMonth{1}
+\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+\startPage{1}
+
+%% Copyright information
+%% Supplied to authors (based on authors' rights management selection;
+%% see authors.acm.org) by publisher for camera-ready submission;
+%% use 'none' for review submission.
+\setcopyright{none}
+%\setcopyright{acmcopyright}
+%\setcopyright{acmlicensed}
+%\setcopyright{rightsretained}
+%\copyrightyear{2018}           %% If different from \acmYear
+
+%% Bibliography style
+\bibliographystyle{ACM-Reference-Format}
+%% Citation style
+%% Note: author/year citations are required for papers published as an
+%% issue of PACMPL.
+\citestyle{acmauthoryear}   %% For author/year citations
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Note: Authors migrating a paper from PACMPL format to traditional
+%% SIGPLAN proceedings format must update the '\documentclass' and
+%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Some recommended packages.
+%\usepackage{booktabs}   %% For formal tables:
+%                        %% http://ctan.org/pkg/booktabs
+%\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
+%                        %% http://ctan.org/pkg/subcaption
+
+\usepackage{wrapfig}
 \setlength{\intextsep}{0pt} % don't add space above nor below
-%\usepackage{subfig}
-%\newcommand{\subfigureautorefname}{\figureautorefname}
-
-\nolinenumbers% needef for arXiv
-\usepackage[frozencache]{minted}
+\usepackage{subfig}
+\newcommand{\subfigureautorefname}{\figureautorefname}
+%% http://ctan.org/pkg/subcaption
+
+\usepackage{minted}
 \usemintedstyle{bw}
@@
 \usepackage{apptools}
-\usepackage{placeins}
 \usepackage{adjustbox}
-\usepackage{xparse}
 \usepgfplotslibrary{units}
@@
 \makeatother
-
-\provideboolean{tobesplit}
-
-\ifthenelse{\boolean{tobesplit}}{%
-  \AtBeginDocument{\renewcommand{\appendixautorefname}{Online Appendix}}%
-}{}
-%https://tex.stackexchange.com/a/149897/2066
-\makeatletter
-\patchcmd{\hyper@makecurrent}{%
-    \ifx\Hy@param\Hy@chapterstring
-        \let\Hy@param\Hy@chapapp
-    \fi
-}{%
-    \iftoggle{inappendix}{%true-branch
-        % list the names of all sectioning counters here
-        \@checkappendixparam{chapter}%
-        \@checkappendixparam{section}%
-        \@checkappendixparam{subsection}%
-        \@checkappendixparam{subsubsection}%
-        \@checkappendixparam{paragraph}%
-        \@checkappendixparam{subparagraph}%
-    }{}%
-}{}{\show\hyper@makecurrent\errmessage{failed to patch}}
-
-\newcommand*{\@checkappendixparam}[1]{%
-    \def\@checkappendixparamtmp{#1}%
-    \ifx\Hy@param\@checkappendixparamtmp
-        \let\Hy@param\Hy@appendixstring
-    \fi
-}
-\makeatletter
-
-\newtoggle{inappendix}
-\togglefalse{inappendix}
-
-\apptocmd{\appendix}{\toggletrue{inappendix}\renewcommand{\appendixautorefname}{Appendix}}{}{\show\appendix\errmessage{failed to patch}}
-%\apptocmd{\subappendices}{\toggletrue{inappendix}}{}{\show\subappendices\errmessage{failed to patch}}
-% end appendix autoref patch
 
@@
 
+% https://tex.stackexchange.com/a/183682/2066
+\makeatletter
+
+% define a macro \Autoref to allow multiple references to be passed to \autoref
+\newcommand\Autoref[1]{\@first@ref#1,@}
+\def\@throw@dot#1.#2@{#1}% discard everything after the dot
+\def\@set@refname#1{%    % set \@refname to autoefname+s using \getrefbykeydefault
+    \edef\@tmp{\getrefbykeydefault{#1}{anchor}{}}%
+    \xdef\@tmp{\expandafter\@throw@dot\@tmp.@}%
+    \ltx@IfUndefined{\@tmp autorefnameplural}%
+         {\def\@refname{\@nameuse{\@tmp autorefname}s}}%
+         {\def\@refname{\@nameuse{\@tmp autorefnameplural}}}%
+}
+\def\@first@ref#1,#2{%
+  \ifx#2@\autoref{#1}\let\@nextref\@gobble% only one ref, revert to normal \autoref
+  \else%
+    \@set@refname{#1}%  set \@refname to autoref name
+    \@refname~\ref{#1}% add autoefname and first reference
+    \let\@nextref\@next@ref% push processing to \@next@ref
+  \fi%
+  \@nextref#2%
+}
+\def\@next@ref#1,#2{%
+   \ifx#2@ and~\ref{#1}\let\@nextref\@gobble% at end: print and+\ref and stop
+   \else, \ref{#1}% print  ,+\ref and continue
+   \fi%
+   \@nextref#2%
+}
+
+\makeatother
+
 \newbibfield{author}
@@
 \newcommand{\citetitle}[1]{\emph{\usebibentry{#1}{title}}\nocite{#1}}
-\newcommand{\citet}[1]{\usebibentry{#1}{citet}~\cite{#1}}
-\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO: #1}}}
+\newcommand{\customcitet}[1]{\usebibentry{#1}{citet}~\cite{#1}}
+\newcommand{\todo}[1]{\textbf{TODO: #1}}
 \newcommand{\coqbug}[1]{\href{https://github.com/coq/coq/issues/#1}{\##1}}
@@
 \newcommand{\coqissue}[1]{\href{https://github.com/coq/coq/issues/#1}{\##1}}
-\NewDocumentCommand{\githublink}{m m o}{%
-  \IfNoValueTF{#3}%
-    {\href{https://github.com/#1/#2}{\texttt{#1/#2}}}%
-    {\href{https://github.com/#1/#2/tree/#3}{\texttt{#1/#2@#3}}}%
-}
-\NewDocumentCommand{\githublinkurl}{m m o}{%
-  \IfNoValueTF{#3}%
-    {\url{https://github.com/#1/#2}}%
-    {\url{https://github.com/#1/#2/tree/#3}}%
-}
 \newcommand{\Rtac}{\ensuremath{\mathcal{R}_\text{\emph{tac}}}}
@@
 }
-\newcommand{\startsupplement}{%
+\newcommand{\supplement}{%
   \clearpage
@@
 
+\begin{document}
+%% Title information
+%\title[Short Title]{Full Title}         %% [Short Title] is optional;
+%                                        %% when present, will be used in
+%                                        %% header instead of Full Title.
+%\titlenote{with title note}             %% \titlenote is optional;
+%                                        %% can be repeated if necessary;
+%                                        %% contents suppressed with 'anonymous'
+%\subtitle{Subtitle}                     %% \subtitle is optional
+%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
+%                                        %% can be repeated if necessary;
+%                                        %% contents suppressed with 'anonymous'
 \title{Accelerating Verified-Compiler Development with a Verified Rewriting Engine}
 
-%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line
-
-\author{Jason Gross}{CSAIL, Massachusetts Institute of Technology, 77 Massachusetts Ave., Cambridge, MA 02139, USA \and MIRI, USA \and \url{https://jasongross.github.io/} }{jgross@mit.edu}{https://orcid.org/0000-0002-9427-4891}{}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty.
-Please provide at least the name of the affiliation and the country.
-The full address is optional.
-Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.
-
-\author{Andres Erbsen}{CSAIL, Massachusetts Institute of Technology, 77 Massachusetts Ave., Cambridge, MA 02139, USA \and \url{https://andres.systems/} }{andreser@mit.edu}{}{}
-
-\author{Jade Philipoom%\footnote{now with Google}
-}{CSAIL, Massachusetts Institute of Technology, 77 Massachusetts Ave., Cambridge, MA 02139, USA \and Google}{jade.philipoom@gmail.com}{}{}
-% should we \footnote{now with Google}?
-
-\author{Miraya Poddar-Agrawal}{Reed College, 3203 SE Woodstock Blvd, Portland, OR 97202, USA% \and \url{https://github.com/miraya-pa}
-}{ragrawal@reed.edu}{https://orcid.org/0000-0001-7617-9180}{}
-
-\author{Adam Chlipala}{CSAIL, Massachusetts Institute of Technology, 77 Massachusetts Ave., Cambridge, MA 02139, USA \and \url{http://adam.chlipala.net/} }{adamc@csail.mit.edu}{https://orcid.org/0000-0001-7085-9417}{%[funding]
+%% Author information
+%% Contents and number of authors suppressed with 'anonymous'.
+%% Each author should be introduced by \author, followed by
+%% \authornote (optional), \orcid (optional), \affiliation, and
+%% \email.
+%% An author may have multiple affiliations and/or emails; repeat the
+%% appropriate command.
+%% Many elements are not rendered, but should be provided for metadata
+%% extraction tools.
+
+%% Author with single affiliation.
+\author{Jason Gross}
+%\authornote{with author1 note}          %% \authornote is optional;
+%                                        %% can be repeated if necessary
+\orcid{0000-0002-9427-4891}
+\affiliation{
+%  \position{Position1}
+  \department{CSAIL}              %% \department is recommended
+  \institution{Massachusetts Institute of Technology}            %% \institution is required
+  \streetaddress{77 Massachusetts Avenue}
+  \city{Cambridge}
+  \state{MA}
+  \postcode{02139}
+  \country{USA}                    %% \country is recommended
 }
-
-\authorrunning{J.
-Gross and A.
-Erbsen and J.
-Philipoom and M.
-Poddar-Agrawal and A.
-Chlipala} %TODO mandatory.
-First: Use abbreviated first/middle names.
-Second (only in severe cases): Use first author plus 'et al.'
-
-\Copyright{Jason Gross \& Andres Erbsen \& Jade Philipoom \& Miraya Poddar-Agrawal \& Adam Chlipala} %TODO mandatory, please use full first names.
-LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
-
-\begin{CCSXML}
-<ccs2012>
-   <concept>
-       <concept_id>10003752.10003790.10002990</concept_id>
-       <concept_desc>Theory of computation~Logic and verification</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003790.10003798</concept_id>
-       <concept_desc>Theory of computation~Equational logic and rewriting</concept_desc>
-       <concept_significance>500</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10011007.10011006.10011041</concept_id>
-       <concept_desc>Software and its engineering~Compilers</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10011007.10011006.10011041.10011046</concept_id>
-       <concept_desc>Software and its engineering~Translator writing systems and compiler generators</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
- </ccs2012>
-\end{CCSXML}
-
-\ccsdesc[300]{Theory of computation~Logic and verification}
-\ccsdesc[500]{Theory of computation~Equational logic and rewriting}
-\ccsdesc[300]{Software and its engineering~Compilers}
-\ccsdesc[300]{Software and its engineering~Translator writing systems and compiler generators}
-
-\keywords{compiler verification, rewriting engines, cryptography} %TODO mandatory; please add comma-separated list of keywords
-
-%\category{} %optional, e.g. invited paper
-
-\ifthenelse{\boolean{tobesplit}}{%
-%  \relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
-%  \relatedversiondetails%[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]
-%      {Full Version}{https://arxiv.org/abs/2205.00862} %linktext and cite are optional
-}{}
-
-\supplement{%
-%  \url{https://jasongross.github.io/papers/2022-rewriting-itp-supplement.pdf},
-  \githublinkurl{mit-plv}{rewriter}[ITP-2022-perf-data],
-  \githublinkurl{mit-plv}{fiat-crypto}[perf-testing-data-ITP-2022-rewriting]%
-}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
-%\supplementdetails[linktext={opt.
-text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g.
-Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional
-
-\funding{%
-  This work was supported in part by a Google Research Award, National Science Foundation grants CCF-1253229, CCF-1512611, and CCF-1521584, and the National Science Foundation Graduate Research Fellowship under Grant Nos.\ 1122374 and 1745302.
-  Any opinion, findings, and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the National Science Foundation.%
+\email{jgross@mit.edu}          %% \email is recommended
+
+\author{Andres Erbsen}
+%\authornote{with author1 note}          %% \authornote is optional;
+%                                        %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+%  \position{Position1}
+  \department{CSAIL}              %% \department is recommended
+  \institution{Massachusetts Institute of Technology}            %% \institution is required
+  \streetaddress{77 Massachusetts Avenue}
+  \city{Cambridge}
+  \state{MA}
+  \postcode{02139}
+  \country{USA}                    %% \country is recommended
 }
-%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors.
-Please enter author specific funding statements as fifth argument of the \author macro.
-
-%\acknowledgements{I want to thank \dots}%optional
-
-%\nolinenumbers %uncomment to disable line numbering
-
-%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\EventEditors{June Andronick and Leonardo de Moura}
-\EventNoEds{2}
-\EventLongTitle{13th International Conference on Interactive Theorem Proving (ITP 2022)}
-\EventShortTitle{ITP 2022}
-\EventAcronym{ITP}
-\EventYear{2022}
-\EventDate{August 7--10, 2022}
-\EventLocation{Haifa, Israel}
-\EventLogo{}
-\SeriesVolume{237}
-\ArticleNo{5}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-\maketitle
-
-%TODO mandatory: add short abstract of the document
+\email{andreser@mit.edu}          %% \email is recommended
+
+\author{Rajashree Agrawal}
+%\authornote{with author1 note}          %% \authornote is optional;
+%                                        %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+%  \position{Position1}
+%  \department{}              %% \department is recommended
+  \institution{Reed College}            %% \institution is required
+  \streetaddress{3203 SE Woodstock Blvd}
+  \city{Portland}
+  \state{OR}
+  \postcode{97202}
+  \country{USA}                    %% \country is recommended
+}
+\email{ragrawal@reed.edu}          %% \email is recommended
+
+\author{Jade Philipoom}
+\authornote{now with Google}          %% \authornote is optional;
+                                        %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+%  \position{Position1}
+  \department{CSAIL}              %% \department is recommended
+  \institution{Massachusetts Institute of Technology}            %% \institution is required
+  \streetaddress{77 Massachusetts Avenue}
+  \city{Cambridge}
+  \state{MA}
+  \postcode{02139}
+  \country{USA}                    %% \country is recommended
+}
+\email{jadep@mit.edu}          %% \email is recommended
+
+\author{Adam Chlipala}
+%\authornote{with author1 note}          %% \authornote is optional;
+%                                        %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+%  \position{Position1}
+  \department{CSAIL}              %% \department is recommended
+  \institution{Massachusetts Institute of Technology}            %% \institution is required
+  \streetaddress{77 Massachusetts Avenue}
+  \city{Cambridge}
+  \state{MA}
+  \postcode{02139}
+  \country{USA}                    %% \country is recommended
+}
+\email{adamc@csail.mit.edu}          %% \email is recommended
+
+%
+%\authorrunning{J.\ Gross et al.\@}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
+
+%% Abstract
+%% Note: \begin{abstract}...\end{abstract} environment must come
+%% before \maketitle command
 \begin{abstract}
@@
 
+%\keywords{First keyword  \and Second keyword \and Another keyword.}
+
+%% 2012 ACM Computing Classification System (CSS) concepts
+%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10011007.10011006.10011008</concept_id>
+<concept_desc>Software and its engineering~General programming languages</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10003456.10003457.10003521.10003525</concept_id>
+<concept_desc>Social and professional topics~History of programming languages</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Software and its engineering~General programming languages}
+\ccsdesc[300]{Social and professional topics~History of programming languages}
+%% End of generated code
+
+%% Keywords
+%% comma separated list
+%\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
+
+%% \maketitle
+%% Note: \maketitle command must come after title commands, author
+%% commands, abstract environment, Computing Classification System
+%% environment and commands, and keywords command.
+\maketitle
+
 \section{Introduction}\label{sec:intro}
@@
 \citet{xcert} demonstrated a streamlined way to extend CompCert with new verified optimizations driven by dataflow analysis, but we are not aware of past work that supports easy extension for compilers from functional languages to C code.
-We present our work targeting that style.
+We present our new work targeting that sort of compilation.
 
@@
 In this paper, we present \textbf{the first proven-correct compiler-builder toolkit parameterized on rewrite rules}.
-Arbitrary sets of Coq theorems (quantified equalities) can be assembled by a single new Coq command into an extraction-ready verified compiler.
-We did not need to extend the trusted code base, so our compiler compiler need not be trusted.
-We achieve both good performance of compiler runs and good performance of generated code, via addressing a number of scale-up challenges vs.\ past work.
-%including efficient handling of intermediate code with thousands of nested binders and critical subterm sharing.
-
-We evaluate our toolkit by replacing a key component of Fiat Cryptography~\cite{FiatCryptoSP19}, a Coq library that generates code for big-integer modular arithmetic at the heart of elliptic-curve-cryptography algorithms.
-Routines generated (with proof) with Fiat Cryptography now ship with all major Web browsers and all major mobile operating systems.
-With our improved compiler architecture, it became easy to add two new backends and a variety of new supported source-code features, and we were easily able to try out new optimizations.
-%and later replace unsound optimizations with equivalent sound optimizations.
+Arbitrary sets of Coq theorems, proving quantified equalities, can be assembled by a single new Coq command into an extraction-ready verified compiler.
+We did not need to extend the trusted code base, so our compiler compiler need not be trusted in any way.
+We achieve both good performance of compiler runs and good performance of generated code, via addressing a number of scale-up challenges vs.\ past work, including efficient handling of intermediate code with thousands of nested variable binders and critical subterm sharing.
+
+We evaluate our builder toolkit by replacing a key component of Fiat Cryptography, a Coq library that generates code for big-integer modular arithmetic at the heart of elliptic-curve-cryptography algorithms.
+This domain-specific compiler has been adopted, for instance, in the Chrome Web browser, such that the majority of all HTTPS connections from browsers are now initiated using code generated (with proof) by Fiat Cryptography.
+With our improved compiler architecture, it became easy to add two new backends and a variety of new supported source-code features, and we were easily able to try out new optimizations and later replace unsound optimizations with equivalent sound optimizations.
 
 Replacing Fiat Cryptography's original compiler with the compiler generated by our toolkit has two additional benefits.
-Fiat Cryptography was previously only used successfully to build C code for two of the three most widely used curves (P-256 and Curve25519).
-Our prior version's execution timed out trying to compile code for the third most widely used curve (P-384).
-Using our new toolkit has made it possible to generate compiler-synthesized code for P-384 while generating completely identical code for the primes handled by the previous version, about 1000$\times$ more quickly.
-Additionally, Fiat Cryptography previously required source code to be written in continuation-passing style, and our compiler has enabled a direct-style approach, which pays off in simplifying theorem statements and proofs.
+Fiat Cryptography was previously only used successfully to build C code for the two most widely used curves (P-256 and Curve25519).
+Their execution timed out trying to compile code for the third most widely used curve (P-384).
+Using our toolkit has made it possible to generate compiler-synthesized code for P-384 while generating completely identical code for the primes handled by the previous version, about 1000$\times$ more quickly.
+Additionally, Fiat Cryptography previously required source code to be written in continuation-passing style, and our compiler has enabled a direct-style approach, which pays off in simplifying the theorem statement and proof for every arithmetic combinator.
 
@@
 
-\subsection{Related Work}\label{sec:related-one}
+\subsection{Related Work, Take One}\label{sec:related-one}
 % POPL 2022 review TODO:
@@
 Assume our mission is to take libraries of purely functional combinators, apply them to compile-time parameters, and compile the results down to lean C code.
-Furthermore, we ask for machine-checked proofs that the C programs preserve the behavior of the higher-order functional programs we started with.
+Furthermore, we ask for machine-checked proofs that the C programs preserve the input-output behavior of the higher-order functional programs we started with.
 What good ideas from the literature can we build on?
@@
 Their rewrite-engine driver is unproven OCaml code, while we will produce custom drivers with Coq proofs.
-%This suggests also that they cannot use rewrite rules which are only sound in the presence of side conditions, as such side-condition checking could not happen in the engine driving the rewrites.
-%Overall, their work complements ours; we focus on how to build a scalable engine for building a proven-correct compiler around shallowly embedded rewrite rules in a proof assistant like Coq; they focus on what rewrite rules are necessary for implementing a general-purpose compiler.
+This suggests also that they cannot use rewrite rules which are only sound in the presence of side conditions, as such side-condition checking could not happen in the engine driving the rewrites.
+Overall, their work complements ours; we focus on how to build a scalable engine for building a proven-correct compiler around shallowly embedded rewrite rules in a proof assistant like Coq; they focus on what rewrite rules are necessary for implementing a general-purpose compiler.
 
@@
 \Rtac{} is also more heavyweight to use, for instance requiring that theorems be restated manually in a deep embedding to bring them into automation procedures.
-Furthermore, we are not aware of any past experiments driving verified compilers with \Rtac{}.
-%, and we expect there would be other bottlenecks working with large, deeply nested terms.
+Furthermore, we are not aware of any past experiments driving verified compilers with \Rtac{}, and we expect there would be other bottlenecks working with large, deeply nested terms.
 
@@
 
-Our implementation builds on fast full reduction in Coq's kernel, via a virtual machine~\cite{vmcompute} or compilation to native code~\cite{nativecompute} (neither verified).
-Especially the latter is similar in adopting NbE for full reduction, simplifying even under $\lambda$s, on top of a more traditional implementation of OCaml that never executes preemptively under $\lambda$s.
-Neither approach unifies support for rewriting with proved rules, and partial evaluation only applies in very limited cases.
-
-A variety of forms of pragmatic partial evaluation have been demonstrated, with Lightweight Modular Staging~\cite{LMS} in Scala as one of the best-known current examples.
-The LMS-Verify system~\cite{LMSVerify} can be used for formal verification of generated code after-the-fact.
-Typically LMS-Verify has been used with relatively shallow properties (though potentially applied to larger and more sophisticated code bases than we tackle), not scaling to the kinds of functional-correctness properties that concern us here.
-%justifying investment in verified partial evaluators.
-
 So, overall, to our knowledge, no past compiler as a set of rewrite rules has come with a full proof of correctness as a standalone functional program.
 Related prior work with mechanized proofs suffered from both performance bottlenecks and usability problems, the latter in requiring that eligible rewrite rules be stated in special deep embeddings.
-%We will demonstrate all of these pieces working together well enough to be adopted by several high-profile open-source projects.
+We will demonstrate all of these pieces working together well enough to be adopted by several high-profile open-source projects.
 
@@
   Naive representation of terms without proper \emph{sharing of common subterms} can lead to fatal term-size blow-up.
-  %Fiat Cryptography's arithmetic routines rely on significant sharing of this kind.
+  Fiat Cryptography's arithmetic routines rely on significant sharing of this kind.
 \item
   Unconditional rewrite rules are in general insufficient, and we need \emph{rules with side conditions}.
-  E.g., Fiat Cryptography depends on checking lack-of-overflow conditions.
+  For instance, in Fiat Cryptography, some rules for simplifying modular arithmetic depend on proofs that operations in subterms do not overflow.
 \item
   However, it is also not reasonable to expect a general engine to discharge all side conditions on the spot.
-  We need integration with \emph{abstract interpretation}.
-  %that can analyze whole programs to support reduction.
+  We need integration with \emph{abstract interpretation} that can analyze whole programs to support reduction.
 \end{itemize}
@@
 
-We pause to give a motivating example before presenting the core structure of our engine (\autoref{sec:structure}), the additional scaling challenges we faced (\autoref{sec:scaling}), experiments (\autoref{sec:evaluation}), and conclusions.
-Our implementation is attached.
+We pause to give a motivating example before presenting the core structure of our engine (\autoref{sec:structure}), the additional scaling challenges we faced (\autoref{sec:scaling}), experiments (\autoref{sec:evaluation}), and related work (\autoref{sec:related}) and conclusions.
+Our implementation is included as an anonymous supplement.
 
@@
 
-Our compilation style involves source programs that mix higher-order functions and inductive types.
-We want to compile to C code, reducing away uses of fancier features while seizing opportunities for arithmetic simplification.
-Here is a small but illustrative example.
-
-\begin{minted}[fontsize=\small]{coq}
-Definition prefixSums (ls : list nat) : list nat :=
+Our style of compilation involves source programs that mix higher-order functions and inductive datatypes.
+We want to compile them to C code, reducing away all uses of fancier features while seizing opportunities for arithmetic simplification.
+Let us walk through a small example that nonetheless illustrates several key challenges.
+
+\begin{minted}[fontsize=\small]{coq}
+Definition prefixSums (ls:list nat) : list nat :=
   let ls' := combine ls (seq 0 (length ls)) in
@@
   let '(_, ls''') := fold_left (λ '(acc, ls''') n,
-    let acc' := acc + n in (acc', acc' :: ls''')) ls'' (0, []) in ls'''.
+    let acc' := acc + n in (acc', acc' :: ls''')) ls'' (0, []) in
+  ls'''.
 \end{minted}
@@
 Then we map over the list of pairs, multiplying the components at each position.
-Finally, we compute all prefix sums.
+Finally, we traverse that list, building up a list of all prefix sums.
 
@@
 \begin{minted}[fontsize=\small]{coq}
-let acc := b + c * 2 in let acc' := acc + d * 3 in [acc'; acc; b; 0]
+let acc  := b + c * 2   in
+let acc' := acc + d * 3 in
+[acc'; acc; b; 0]
 \end{minted}
@@
 We do not quite have C code yet, but, composing this code with another routine to consume the output list, we easily arrive at a form that looks almost like three-address code and is quite easy to translate to C and many other languages.
-%That actual translation we leave to other small compiler phases outside the scope of this paper.
+That actual translation we leave to other small compiler phases outside the scope of this paper.
 
@@
 Our plugin also provides new tactics like \mintinline{coq}{Rewrite_rhs_for}, which applies a rewriter to the right-hand side of an equality goal.
-That last tactic is just what we need to synthesize a specialized \mintinline{coq}{prefixSums} for list length four, along with its correctness proof.
+That last tactic is just what we need to synthesize a specialized \mintinline{coq}{prefixSums} for list length four, along with a proof of its equivalence to the original function.
 \begin{minted}[fontsize=\small]{coq}
 Definition prefixSums4 :
-{f:nat→nat→nat→nat→list nat | ∀ a b c d, f a b c d = prefixSums [a;b;c;d]}
+{f:nat→nat→nat→nat→list nat|∀ a b c d,f a b c d = prefixSums [a;b;c;d]}
   := ltac:(eexists; Rewrite_rhs_for rewriter; reflexivity).
@@
 That compiler execution ran inside of Coq, but an even more pragmatic approach is to \emph{extract} the compiler as a standalone program in OCaml or Haskell.
-Such a translation is possible because the \mintinline{coq}{Make} command produces a proved program in Gallina, Coq's logic.
+Such a translation is possible because the \mintinline{coq}{Make} command produces a proved program in Gallina, Coq's logic that doubles as a dependently typed functional language.
 As a result, our reworking of Fiat Cryptography compilation culminated in extraction of a command-line OCaml program that developers in industry have been able to run without our help, where Fiat Cryptography previously required installing and running Coq, with an elaborate build process to capture its output.
-It is also true that the standalone program is about 10$\times$ as fast as execution within Coq, though the trusted code base is larger.
-% for the extracted version, since extraction itself is not proved.
+It is also true that the standalone program is about 10$\times$ as fast as execution within Coq, though the trusted code base is larger for the extracted version, since extraction itself is not proved.
 
@@
 \item
-  Within the logic of the proof assistant (Isabelle/HOL, in their case), a type of syntax trees for ML programs is defined, with an associated (trusted) operational semantics.
-\item
-  They also wrote a reduction function in (deeply embedded) ML, parameterized on a function to choose the next rewrite, and proved it sound once-and-for-all.
+  Within the logic of the proof assistant (Isabelle/HOL, in their case), a type of syntax trees for ML programs is defined, with an associated operational semantics.
+  This operational semantics is trusted, much as Coq extraction is trusted.
+\item
+  They also wrote a reduction function in (deeply embedded) ML, parameterized on a function to choose the next rewrite, and proved it sound once-and-for-all, against the ML operational semantics.
 \end{enumerate}
@@
 \item
-  \emph{Generate a (deeply embedded) ML program that decides which rewrite rule, if any, to apply at the top node of a syntax tree}, along with a proof of its soundness.
+  \emph{Generate a (deeply embedded) ML program that decides which rewrite rule, if any, to apply at the top node of a syntax tree}, along with a proof of its soundness with respect to ML semantics.
 \item
@@
 \item
-  Combining the general proof of the rewrite engine with proofs generated by reification (the prior two steps), conclude that an application of the reduction function to the reified rules and term is indeed an ML term that generates correct answers.
+  Combine the general proof of the rewrite engine with proofs generated by reification (the prior two steps), conclude that an application of the reduction function to the reified rules and term is indeed an ML term that generates correct answers.
 \item
   ``Throw the ML term over the wall,'' using a general code-generation framework for Isabelle/HOL~\cite{CodeGen}.
-  Trusted code compiles the ML code into the concrete syntax of Standard ML, and compiles it, and runs it, asserting an axiom about the outcome.
+  Trusted code compiles the ML code into the concrete syntax of a mainstream ML language, Standard ML in their case, and compiles it with an off-the-shelf compiler.
+  The output of that compiled program is then passed back over to the tactic, in terms of an axiomatic assertion that the ML semantics really yields that answer.
 \end{enumerate}
@@
   Our reduction engine is written \emph{as a normal Gallina functional program}, rather than within a deeply embedded language.
-  As a result, we are able to prove its type-correctness and termination, and we are able to run it within Coq's kernel.
+  As a result, we are able to prove its type-correctness and termination, and we are able to run it within Coq's kernel, rather than through a plugin.
 \item
@@
 
-\subsection{Our Approach in Ten Steps}\label{sec:nine-steps}
+\subsection{Our Approach in Nine Steps}\label{sec:nine-steps}
 
 Here is a bit more detail on the steps that go into applying our Coq plugin, many of which we expand on in the following sections.
-For \mintinline{coq}{Make} to precompute a rewriter:
+In order to build a precomputed rewriter with the \mintinline{coq}{Make} command, the following actions are performed:
 \begin{enumerate}
 \item
-  The given lemma statements are scraped for which named identifiers to encode.
+  The given lemma statements are scraped for which named functions and types the rewriter package will support.
 \item
   Inductive types enumerating all available primitive types and functions are emitted.
-  This allows us to achieve the performance gains attributed in \citet{Boespflug2009} to having native metalanguage constructors for all constants, without manual coding.
+  This allows us to achieve the performance gains attributed in \citet{Boespflug2009} to having native metalanguage constructors for each constant without needing to manually code an enumeration.
 \item
   Tactics generate all of the necessary definitions and prove all of the necessary lemmas for dealing with this particular set of inductive codes.
-  Definitions include operations like Boolean equality on type codes and lemmas like ``all types have decidable equality.''
+  Definitions include operations like Boolean equality on type codes and lemmas like ``all representable primitive types have decidable equality.''
 \item
   The statements of rewrite rules are reified and soundness and syntactic-well-formedness lemmas are proven about each of them.
-  %Each instance of the former involves wrapping the user-provided proof with the right adapter to apply to the reified version.
-  %Automating this step allows rewrite rules to be proven in terms of their shallow embedding, which drastically accelerates iteration on the set of rewrite rules.
-\item
-  Definitions and lemmas needed to prove correctness are assembled into a single package.
+  Each instance of the former involves wrapping the user-provided proof with the right adapter to apply to the reified version.
+  Automating this step allows rewrite rules to be proven in terms of their shallow embedding, which drastically accelerates iteration on the set of rewrite rules.
+\item
+  The definitions needed to perform reification and rewriting and the lemmas needed to prove correctness are assembled into a single package that can be passed by name to the rewriting tactic.
 \end{enumerate}
 
-Then, to rewrite in a goal, the following steps are performed:
+When we want to rewrite with a rewriter package in a goal, the following steps are performed:
 \begin{enumerate}
 \item
-  Rearrange the goal into a single quantifier-free logical formula.%: free-variable quantification is replaced by equality between functions (taking free variables as inputs).
+  We rearrange the goal into a single logical formula: all free-variable quantification in the proof context is replaced by changing the equality goal into an equality between two functions (taking the free variables as inputs).
   %\todo{Pedantic Note From Jason: it's not actually an equality between two functions, it's an \mintinline{coq}{equiv} between two functions where \mintinline{coq}{equiv} is a custom relation we define indexed over type codes that is equality up to function extensionality}
 \item
-  Reify a selected subterm and replace it with a call to our denotation function.
-  %We reify the goal side we want to simplify, using inductive codes from the specified package.
-  %That goal side is replaced with a denotation-function call on the reified version.
-\item
-  Rewrite with a theorem, into a form calling our rewriter.
-\item Call Coq's built-in full reduction (\tacvmcompute{}) to reduce this application.
-\item
-  Run standard call-by-value reduction to simplify away use of the denotation function. %on the concrete syntax tree from rewriting.
+  We reify the side of the goal we want to simplify, using the inductive codes in the specified package.
+  That side of the goal is then replaced with a call to a denotation function on the reified version.
+\item
+  We use a theorem stating that rewriting preserves denotations of well-formed terms to replace the denotation subterm with the denotation of the rewriter applied to the same reified term.
+  We use Coq's built-in full reduction (\tacvmcompute{}) to reduce the application of the rewriter to the reified term.
+\item
+  Finally, we run \taccbv{} (a standard call-by-value reducer) to simplify away the invocation of the denotation function on the concrete syntax tree from rewriting.
   %\todo{Is it worth mentioning that we'd instead use \tacvmcompute{} if it supported whitelists (and also said whitelists were recorded in the casts)?}
@@
 % POPL 2022 review TODO: Line 312: I'm not sure what "nearly simply typed" means.
-The object language of our rewriter is nearly simply typed. %with limited support for calling polymorphic functions.
+The object language of our rewriter is nearly simply typed, with limited support for calling polymorphic functions.
 %However, we still support identifiers whose definitions use dependent types, since our reducer does not need to look into definitions.
@@
 As we instead build our rewriter in Coq's logic, we have no such option to defer to ML.
-%Indeed, Coq's logic only includes primitive pattern-matching constructs to match one constructor at a time.
+Indeed, Coq's logic only includes primitive pattern-matching constructs to match one constructor at a time.
 %The frontend desugars nested pattern matching into trees of matches, but we do not have access to that desugaring within the language.
@@
 Instead, we adopted the approach of \citet{maranget2008compiling}, who describes compilation of pattern matches in OCaml to decision trees that eliminate needless repeated work (for example, decomposing an expression into $x + y + z$ only once even if two different rules match on that pattern).
-%We have not yet implemented his optimizations for finding \emph{minimal} decision trees.
+We have not yet implemented any of the optimizations described therein for finding \emph{minimal} decision trees.
 
@@
 Here the correctness condition is that we must not change the semantics of the expression.
-%Said another way, any rewrite-rule replacement expression must match the semantics of the rewrite-rule pattern.
+Said another way, any rewrite-rule replacement expression must match the semantics of the rewrite-rule pattern.
 
@@
   \item \texttt{Switch icases app\_case default}:
-    With the first element of the vector, match on its kind; if it is an identifier matching something in \texttt{icases}, which is a list of pairs of identifiers and decision trees, remove the first element of the vector and run that decision tree; if it is an application and \texttt{app\_case} is not \texttt{None}, try the \texttt{app\_case} decision tree, replacing the first element of each vector with the two elements of the function and the argument it is applied to; otherwise, do not modify the vectors and use the \texttt{default}.
+    With the first element of the vector, match on its kind; if it is an identifier matching something in \texttt{icases}, which is a list of pairs of identifiers and decision trees, remove the first element of the vector and run that decision tree; if it is an application and \texttt{app\_case} is not \texttt{None}, try the \texttt{app\_case} decision tree, replacing the first element of each vector with the two elements of the function and the argument it is applied to; otherwise, do not modify the vectors and use the \texttt{default} decision tree.
   \item \texttt{Swap i cont}: Swap the first element of the vector with the $i^\texttt{th}$ element (0-indexed) and keep going with \texttt{cont}.
@@
 We begin by reviewing NbE's most classic variant, for performing full $\beta$-reduction in a simply typed term in a guaranteed-terminating way.
-Our simply typed $\lambda$-calculus syntax is:
+The simply typed $\lambda$-calculus syntax we use is:
 \begin{align*}
@@
 We can now define normalization by evaluation.
-First, we choose a ``semantic'' representation for each syntactic type, which serves as an interpreter's result type.
+First, we choose a ``semantic'' representation for each syntactic type, which serves as the result type of an intermediate interpreter.
 \begin{align*}
-  \text{NbE}_t(t_1 \to t_2) & \defeq \text{NbE}_t(t_1) \to \text{NbE}_t(t_2)
-  & \text{NbE}_t(b) & \defeq \texttt{expr}(b)
+  \text{NbE}_t(t_1 \to t_2) & \defeq \text{NbE}_t(t_1) \to \text{NbE}_t(t_2) \\
+  \text{NbE}_t(b) & \defeq \texttt{expr}(b)
 \end{align*}
@@
 \begin{figure}%[b]
-  \begin{subfigure}[b]{0.5\textwidth}
 \begin{align*}
@@
   \text{reflect}_{b}(e) & \defeq e \\ \noalign{\vskip7pt}
-    \end{align*}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.5\textwidth}
-    \begin{align*}
   \text{reduce} & : \text{expr}(t) \to \text{NbE}_t(t) \\
@@
 \end{align*}
-  \end{subfigure}
 \caption{\label{fig:nbe}Implementation of normalization by evaluation}
@@
 We split out analysis of term syntax in a separate function reduce, defined by primitive recursion on term syntax, when usually this functionality would be mixed in with reflect.
-The reason for this choice will become clear when we extend NbE.
+The reason for this choice will become clear when we extend NbE to handle our full problem domain.
 
@@
 % POPL 2022 review TODO: line 433, "First, [...]" but there is no "Second, [...]"?
-First, we bind our specialized decision-tree engine (which rewrites \emph{at the root of an AST only}) under the name rewrite-head.
+First, we bind our specialized decision-tree engine under the name rewrite-head.
+Recall that this function only tries to apply rewrite rules at the top level of its input.
 
@@
 It is important to note that a constant of function type will be $\eta$-expanded only once for each syntactic occurrence in the starting term, though the expanded function is effectively a thunk, waiting to perform rewriting again each time it is called.
-From first principles, it is not clear why such a strategy terminates on all possible input terms.
-%, though we work up to convincing Coq of that fact.
+From first principles, it is not clear why such a strategy terminates on all possible input terms, though we work up to convincing Coq of that fact.
 
@@
 Recall that their rewriter was implemented in a deeply embedded ML, while ours is implemented in Coq's logic, which enforces termination of all functions.
-Aehlig et al.\ did not prove termination, which indeed does not hold for their rewriter in general, which works with untyped terms, not to mention the possibility of divergent rule-specific ML functions.
+Aehlig et al.\ did not prove termination, which indeed does not hold for their rewriter in general, which works with untyped terms, not to mention the possibility of rule-specific ML functions that diverge themselves.
 In contrast, we need to convince Coq up-front that our interleaved $\lambda$-term normalization and algebraic simplification always terminate.
-Additionally, we must prove that rewriting preserves term denotations, which can easily devolve into tedious binder bookkeeping.
-%, depending on encoding.
+Additionally, we need to prove that our rewriter preserves denotations of terms, which can easily devolve into tedious binder bookkeeping, depending on encoding.
 
@@
 \begin{minted}[fontsize=\small]{coq}
-Fixpoint denoteT (t : type) : Type := match t with
+Fixpoint denoteT (t : type) : Type
+  := match t with
      | arrow s d => denoteT s -> denoteT d
-  | base b    => denote_base_type b  end.
+     | base b    => denote_base_type b
+     end.
 \end{minted}
@@
 Next we see the convenience of being able to \emph{use} an expression by choosing how it should represent variables.
-Specifically, it is natural to choose \emph{the type-denotation function itself} as variable representation.
+Specifically, it is natural to choose \emph{the type-denotation function itself} as the variable representation.
 % POPL 2022 review TODO:
@@
 % I believe I would prefer to see correct code up front.
-Especially note how this choice makes rigorous last section's convention (e.g., in the suspicious function-abstraction clause of reduce), where a recursive function enforces that values have always been substituted for variables early enough.
-\begin{minted}[fontsize=\small]{coq}
-Fixpoint denoteE {t} (e : expr denoteT t) : denoteT t := match e with
+Especially note how this choice makes rigorous the convention we followed in the prior section (e.g., in the suspicious function-abstraction clause of function reduce), where a recursive function enforces that values have always been substituted for variables early enough.
+\begin{minted}[fontsize=\small]{coq}
+Fixpoint denoteE {t} (e : expr denoteT t) : denoteT t
+  := match e with
      | Var v     => v
@@
      | LetIn x f => let xv := denoteE x in denoteE f xv
-  | Ident c   => denoteI c  end.
-Definition DenoteE {t} (E : Expr t) : denoteT t := denoteE (E denoteT).
-\end{minted}
-
-It is now easy to follow the same script in making our rewriting-enabled NbE fully formal, in \autoref{fig:nbe2}.
+     | Ident c   => denoteI c
+     end.
+Definition DenoteE {t} (E : Expr t) : denoteT t
+  := denoteE (E denoteT).
+\end{minted}
+
+It is now easy to follow the same script in making our rewriting-enabled NbE fully formal.
 Note especially the first clause of \texttt{reduce}, where we avoid variable substitution precisely because we have chosen to represent variables with normalized semantic values.
@@
 The final definition $\lambda$-quantifies over that choice.
-
-\begin{figure}
-  \begin{subfigure}[b]{0.5\textwidth}
-    \begin{minted}[fontsize=\small]{coq}
-Fixpoint nbeT var (t : type) : Type :=
-match t with
+\begin{minted}[fontsize=\small]{coq}
+Fixpoint nbeT var (t : type) : Type
+  := match t with
      | arrow s d => nbeT var s -> nbeT var d
@@
      end.
-
-Fixpoint reify {var t}
-  : nbeT var t -> expr var t :=
-match t with
-| arrow s d => λ f, Abs (λ x,
-    reify (f (reflect (Var x))))
-| base b    => λ e, e            end
-    \end{minted}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.5\textwidth}
-    \begin{minted}[fontsize=\small]{coq}
-with reflect{var t}:expr var t->nbeT var t
+Fixpoint reify {var t} : nbeT var t -> expr var t
   := match t with
-| arrow s d => λ e, λ x,
-  reflect (App e (reify x))
-| base b    => rewrite_head      end.
-Fixpoint reduce{var t}(e:expr (nbeT var) t)
-  : nbeT var t := match e with
+     | arrow s d => λ f, Abs (λ x, reify (f (reflect (Var x))))
+     | base b    => λ e, e
+     end
+with reflect {var t} : expr var t -> nbeT var t
+  := match t with
+     | arrow s d => λ e, λ x, reflect (App e (reify x))
+     | base b    => rewrite_head
+     end.
+Fixpoint reduce {var t} (e : expr (nbeT var) t) : nbeT var t
+  := match e with
      | Abs e     => λ x, reduce (e (Var x))
@@
      | Var x     => x
-| Ident c   => reflect (Ident c) end.
-Definition Rewrite {t} (E:Expr t) : Expr t
+     | Ident c   => reflect (Ident c)
+     end.
+Definition Rewrite {t} (E : Expr t) : Expr t
   := λ var, reify (reduce (E (nbeT var t))).
 \end{minted}
-  \end{subfigure}
-  \caption{\label{fig:nbe2}PHOAS implementation of normalization by evaluation}
-\end{figure}
-
-One subtlety hidden in \autoref{fig:nbe2} in implicit arguments is in the final clause of \texttt{reduce}, where the two applications of the \texttt{Ident} constructor use different variable representations.
+
+One subtlety hidden above in implicit arguments is in the final clause of \texttt{reduce}, where the two applications of the \texttt{Ident} constructor use different variable representations.
 With all those details hashed out, we can prove a pleasingly simple correctness theorem, with a lemma for each main definition, with inductive structure mirroring recursive structure of the definition, also appealing to correctness of last section's pattern-compilation operations.
@@
 % The phrasing is somewhat odd.
-%Even before getting to the correctness theorem, we needed to convince Coq that the function terminates.
-%While for \citet{Aehlig}, a termination proof would have been a whole separate enterprise, it turns out that PHOAS and NbE line up so well that Coq accepts the above code with no additional termination proof; each key function is obviously structurally recursive on either a type or an expression.
+Even before getting to the correctness theorem, we needed to convince Coq that the function terminates.
+While for \citet{Aehlig}, a termination proof would have been a whole separate enterprise, it turns out that PHOAS and NbE line up so well that Coq accepts the above code with no additional termination proof; each key function is obviously structurally recursive on either a type or an expression.
 % POPL 2022 review TODO:
@@
 % If you mean that the Coq type-checker performs reduction during type-checking, that is a well-known fact which perhaps does not need to be recalled here.
-%As a result, the Coq kernel is ready to run our \mintinline{coq}{Rewrite} procedure during checking.
+As a result, the Coq kernel is ready to run our \mintinline{coq}{Rewrite} procedure during checking.
 
@@
 Now we use the \mintinline{coq}{Rewrite} correctness theorem to replace $\llbracket E \rrbracket$ with $\llbracket \mintinline{coq}{Rewrite}(E) \rrbracket$.
-Next we ask the Coq kernel to simplify $\mintinline{coq}{Rewrite}(E)$ by \emph{full reduction via native compilation}.
-%since we carefully designed $\mintinline{coq}{Rewrite}(E)$ and its dependencies to produce closed syntax trees, so that reduction will not get stuck pattern-matching on free variables.
-Finally, where $E'$ is the result of that reduction, we simplify $\llbracket E' \rrbracket$ with standard reduction.
+Next we may ask the Coq kernel to simplify $\mintinline{coq}{Rewrite}(E)$ by \emph{full reduction via compilation to native code}, since we carefully designed $\mintinline{coq}{Rewrite}(E)$ and its dependencies to produce closed syntax trees, so that reduction will not get stuck pattern-matching on free variables.
+Finally, where $E'$ is the result of that reduction, we simplify $\llbracket E' \rrbracket$ with standard reduction, producing a normal-looking Coq term.
 
@@
 % POPL 2022 review(ish) TODO: actually it's no longer explained further in \autoref{sec:macro}
-We ran that old code generator (actually just a subset of its functionality, but optimized by us a bit further, as explained in \autoref{sec:macro}) on the P-256 example and found it took about 15 seconds to finish.
-Then we modified reduction to inline \mintinline{coq}{let} binders instead of preserving them, at which point the job terminated with an out-of-memory error, on a machine with 64 GB of RAM.
-%(The successful run uses under 2 GB.)
+We ran their code generator (actually just a subset of its functionality, but optimized by us a bit further, as explained in \autoref{sec:macro}) on the P-256 example and found it took about 15 seconds to finish.
+Then we modified reduction to inline \mintinline{coq}{let} binders instead of preserving them, at which point the reduction job terminated with an out-of-memory error, on a machine with 64 GB of RAM.
+(The successful run uses under 2 GB.)
 % POPL 2022 review TODO (maybe just include this point):
@@
 We see a tension here between performance and niceness of library implementation.
-When we built the original Fiat Cryptography, we found it necessary to CPS-convert the code to coax Coq into adequate reduction performance.
-Then all of our correctness theorems were complicated by reasoning about continuations.
-In fact, the CPS reasoning was so painful that at one point most algorithms in the template library were defined twice, once in continuation-passing style and once in direct-style code, because it was easier to prove the two equivalent and work with the non-CPS version than to reason about the CPS version directly.
+The Fiat Cryptography authors found it necessary to CPS-convert their code to coax Coq into adequate reduction performance.
+Then all of their correctness theorems were complicated by reasoning about continuations.
+In fact, the CPS reasoning was so painful that at one point most algorithms in their template library were defined twice, once in continuation-passing style and once in direct-style code, because it was easier to prove the two equivalent and work with the non-CPS version than to reason about the CPS version directly.
 It feels like a slippery slope on the path to implementing a domain-specific compiler, rather than taking advantage of the pleasing simplicity of partial evaluation on natural functional programs.
@@
 \begin{align*}
-  \texttt{map}\ (\lambda x.\, y+x)\ (\letin[{z:=e}{[0;1;z+1]}]) & %\\
-  \; \leadsto \;
-  \letin[{z:=e}{[y;y+1;y+(z+1)]}]
+  \texttt{map}\ (\lambda x.\, y+x)\ (\letin[{z:=e}{[0;1;2;z;z+1]}]) \\
+  \leadsto \; \letin[{z:=e}{[y;y+1;y+2;y+z;y+(z+1)]}]
 \end{align*}
@@
 \begin{minted}[fontsize=\small]{coq}
-Inductive UnderLets {var} (T : Type) := Base (v : T)
-| UnderLet {A} (e : @expr var A) (f : var A -> UnderLets T).
+Inductive UnderLets {var} (T : Type) :=
+| Base (v : T)
+| UnderLet {A}(e : @expr var A)(f : var A -> UnderLets T).
 \end{minted}
 A value of type \mintinline{coq}{UnderLets T} is a series of \texttt{let} binders (where each expression \mintinline{coq}{e} may mention earlier-bound variables) ending in a value of type \mintinline{coq}{T}.
-%It is easy to build various ``smart constructors'' working with this type, for instance to construct a function application by lifting the \texttt{let}s of both function and argument to a common top level.
-
-%Such constructors are used to implement an NbE strategy that outputs \mintinline{coq}{UnderLets} telescopes.
+It is easy to build various ``smart constructors'' working with this type, for instance to construct a function application by lifting the \texttt{let}s of both function and argument to a common top level.
+
+Such constructors are used to implement an NbE strategy that outputs \mintinline{coq}{UnderLets} telescopes.
 Recall that the NbE type interpretation mapped base types to expression syntax trees.
-We add flexibility, parameterizing by a Boolean declaring whether to introduce telescopes.
-
-\begin{minted}[fontsize=\small]{coq}
-Fixpoint nbeT' {var} (with_lets : bool) (t : type) := match t with
+We now parameterize that type interpretation by a Boolean declaring whether we want to introduce telescopes.
+
+\begin{minted}[fontsize=\small]{coq}
+Fixpoint nbeT' {var} (with_lets : bool) (t : type)
+  := match t with
      | base t => if with_lets then @UnderLets var (@expr var t) else @expr var t
-  | arrow s d => nbeT' false s -> nbeT' true d  end.
-Definition nbeT := nbeT' false.      Definition nbeT_with_lets := nbeT' true.
+     | arrow s d => nbeT' false s -> nbeT' true d
+     end.
+Definition nbeT := nbeT' false.
+Definition nbeT_with_lets := nbeT' true.
 \end{minted}
@@
 \begin{align*}
-  {?n} / {?m} & \to n \gg \log_2 m\text{\quad if\quad }2^{\lfloor \log_2 m \rfloor} = m
+  {?n} / {?m} & \to n \gg \log_2 m\text{\ \ if\ \ }2^{\lfloor \log_2 m \rfloor} = m
 \end{align*}
@@
 We added a variant of pattern variable that only matches constants.
-Semantically, this variable style has no additional meaning, and in fact we implement it as a special identity function (notated as an apostrophe) that should be called in the right places within Coq lemma statements.
+Semantically, this variable style has no additional meaning, and in fact we implement it as a special identity function that should be called in the right places within Coq lemma statements.
 Rather, use of this identity function triggers the right behavior in our tactic code that reifies lemma statements.
-%We introduce a notation where a prefixed apostrophe signals a call to the ``constants only'' function.
-\label{sec:explain-'}
+We introduce a notation where a prefixed apostrophe signals a call to the ``constants only'' function.\label{sec:explain-'}
 %We use a special apostrophe marker to indicate a quantified variable that may only match with \emph{compile-time constants}.
@@
 The \mintinline{coq}{Make} command fails if any such expression contains pattern variables not marked as constants.
-%Therefore, matching of rules can safely run side conditions, knowing that Coq's full-reduction engine can determine their truth efficiently.
-
-Hence, we encode the above rule as $\forall n, m. \; 2^{\lfloor \log_2(\texttt{'}m)\rfloor} = \texttt{'}m \to n / \texttt{'}m = n \gg \texttt{'}(\log_2 m)$.
+Therefore, matching of rules can safely run side conditions, knowing that Coq's full-reduction engine can determine their truth efficiently.
+
+Hence, we encode the above rule as
+\begin{align*}
+  \forall n, m. \; 2^{\lfloor \log_2(\texttt{'}m)\rfloor} = \texttt{'}m \to n / \texttt{'}m = n \gg \texttt{'}(\log_2 m)
+\end{align*}
 
@@
 
-See \autoref{sec:implementation-and-usage}\ifthenelse{\boolean{tobesplit}}{%
-  \footnote{%
-  The appendix of this paper is available online at \url{https://jasongross.github.io/papers/2022-rewriting-itp-supplement.pdf}.%
-  }%
-}{}
-for discussion of some further twists in the implementation.
+\subsection{Limitations and Preprocessing}\label{sec:implementation-and-usage}
+
+%\todo{It's not clear to Jason what in this section belongs here, and what isn't interesting (yet?) to readers}
+%\todo{I (Jason) have no idea what the story for this section should be}
+%\todo{Insofar as text is actually kept in this section, make sure any duplication between here and the out-of-paper appendix is deliberate}
+%\todo{If this section is moved or removed, be sure to restructure \autoref{sec:code-from-implementation-and-usage} accordingly}
+
+We now note some details of the rewriting framework that were previously glossed over, which are useful for using the code or implementing something similar, but which do not add fundamental capabilities to the approach.
+Although the rewriting framework does not support dependently typed constants, we can automatically preprocess uses of eliminators like \mintinline{coq}{nat_rect} and \mintinline{coq}{list_rect} into nondependent versions.
+The tactic that does this preprocessing is extensible via \Ltac{}'s reassignment feature.
+Since pattern-matching compilation mixed with NbE requires knowing how many arguments a constant can be applied to, internally we must use a version of the recursion principle whose type arguments do not contain arrows; current preprocessing can handle recursion principles with either no arrows or one arrow in motives.
+
+Recall from \autoref{sec:explain-eval-rect} that \mintinline{coq}{eval_rect} is a definition provided by our framework for eagerly evaluating recursion associated with certain types.
+It functions by triggering typeclass resolution for the lemmas reducing the recursion principle associated to the given type.
+We provide instances for \texttt{nat}, \texttt{prod}, \texttt{list}, \texttt{option}, and \texttt{bool}.
+Users may add more instances if they desire.
+
+Recall again from \autoref{sec:explain-ident.eagerly} that we use \mintinline{coq}{ident.eagerly} to ask the reducer to simplify a case of primitive recursion by complete traversal of the designated argument's constructor tree.
+Our current version only allows a limited, hard-coded set of eliminators with \mintinline{coq}{ident.eagerly} (\texttt{nat\_rect} on return types with either zero or one arrows, \texttt{list\_rect} on return types with either zero or one arrows, and \texttt{List.nth\_default}), but nothing in principle prevents automatic generation of the necessary code.
+
+We define a constant \mintinline{coq}{Let_In} which we use for writing \letin{} expressions that do not reduce under $\zeta$ (Coq's reduction rule for \mintinline{coq}{let}-inlining).
+Throughout most of this paper, anywhere that \letin{} appears, we have actually used \mintinline{coq}{Let_In} in the code.
+It would alternatively be possible to extend the reification preprocessor to automatically convert \letin{} to \mintinline{coq}{Let_In}, but this strategy may cause problems when converting the interpretation of the reified term with the prereified term, as Coq's conversion does not allow fine-tuning of when to inline or unfold \mintinline{coq}{let}s.
 
@@
 % TODO: Presumably ``attached to this submission as an anonymized supplement'' will change for the final submission
-Our implementation, available on GitHub at \githublink{mit-plv}{rewriter}[ITP-2022-perf-data] and with a roadmap in \autoref{sec:CodeSupplement-more}, includes a mix of Coq code for the proved core of rewriting, tactic code for setting up proper use of that core, and OCaml plugin code for the manipulations beyond the tactic language's current capabilities.
+Our implementation, attached to this submission as an anonymized supplement with a roadmap in \autoref{sec:CodeSupplement-more}, includes a mix of Coq code for the proved core of rewriting, tactic code for setting up proper use of that core, and OCaml plugin code for the manipulations beyond the current capabilities of the tactic language.
 We report here on evidence that the tool is effective, first in terms of productivity by users and then in terms of compile-time performance.
@@
 \subsection{Iteration on the Fiat Cryptography Compiler}\label{sec:iteration}
-
 We ported Fiat Cryptography's core compiler functionality to use our framework.
 The result is now used in production by a number of open-source projects.
-We were glad to retire the CPS versions of verified arithmetic functions, which had been present only to support predictable reduction with subterm sharing.
-More importantly, it became easy to experiment with new transformations via proving new rewrite theorems, directly in normal Coq syntax, including the following, all justified by demand from real users:
+The Fiat Cryptography authors were glad to retire the continuation-passing-style versions of verified arithmetic functions, which had been present only to support predictable reduction with subterm sharing.
+More importantly, it became easy to experiment with new transformations via proving new rewrite theorems, directly in normal Coq syntax and proof style.
+
+\subsubsection{Reassociation of Arithmetic}\label{sec:iteration:reassoc}
+One of the simplest and most minor examples of iterating on the Fiat Cryptography compiler is that of reassociating multiplication under certain conditions.
+In particular, if $y \cdot z$ fits in 64 bits, but $x \cdot y$ and $x \cdot y \cdot z$ both take 128 bits, then we want to emit $x \cdot (y \cdot z)$ rather than $(x \cdot y) \cdot z$.
+This way, the C compiler will emit one single-width (64-bit) multiplication and one double-width (128-bit) multiplication, rather than two double-width multiplications.
+The rule we use for this is roughly
+\[
+\forall x, y, c. \; |\texttt{'}c| \le 2^8 \to (x \cdot y) \cdot {\texttt{'}c} = x \cdot (y \cdot {\texttt{'}c})
+\]
+Recall from \autoref{sec:explain-'} that the prefix \texttt{'} is used to indicate that some pattern must be a compile-time constant.
+Since this is just a notation for a specially named identity function, we can prove this rule just by applying associativity of multiplication in the integers.
+
+\subsubsection{New Constructs in the Frontend}\label{sec:iteration:new-frontend}
+Adding support for new constructs in the code being compiled is quite simple: just add the types and constants to the list used to generate the rewriter package.
+The original Fiat Cryptography compiler proof required making changes in several places and modifying several proofs per new identifier.
+We added support for the following constructs in a purely local, modular way:
 \begin{itemize}
-\item Reassociating arithmetic to minimize the bitwidths of intermediate results
-\item Multiplication primitives that separately return high halves and low halves
-\item Strings and a ``comment'' function of type $\forall A. \; \texttt{string} \to A \to A$
-\item Support for bitwise exclusive-or
-\item A special marker to block C compilers from introducing conditional jumps in code that should be constant-time
-\item Eliding bitmask-with-constant operations that can be proved as no-ops
-\item Rules to introduce conditional moves (on supported platforms)
-\item New hardware backend, via rules that invoke special instructions of a cryptographic accelerator
-\item New hardware backend, with a requirement that all intermediate integers have the same bitwidth, via rules to break wider operations down into several narrower operations
+\item multiplication primitives that separately return the high half and low half of a double-width multiplication
+\item strings and a ``comment'' function of type $\forall A. \; \texttt{string} \to A \to A$
+\item support for bitwise exclusive-or
+\item an identity function which prints in the backend as a call to some inline assembly defined in the header which prevents C compilers from introducing conditional jumps into code that ought to be constant-time
 \end{itemize}
 
-% This is an odd one to include in my summarized list.
-% It sounds like more of an observation on the software-engineering experience,
-% rather than extending fiat-crypto.  Maybe BRING BACK if we get more space.
-%% \subsubsection{Moving Rules Involving Carries}\label{sec:iteration:carries}
-%% We originally had rules like ``adding 0 to $x$ produces just $x$, with no carry.''
-%% This rule is not true in general, because we must encode carrying as happening at a particular bitwidth, while we have no guarantee that $x$ fits within that bitwidth.
-%% The correct rule has a precondition: that $x$ is between 0 and $2^{64}$.
-
-%% We discovered this issue when trying to prove our rewrite rules correct.
-%% As a result, we had to move this sort of rule from happening before abstract interpretation to happening after abstract interpretation.
-%% Since the passes are just defined as lists of rewrite rules, moving the rules was quite simple.
-%(Rephrasing them to talk about the outputs of abstract interpretation was somewhat painful, though, because the proof assistant did not enforce the conventions we were using for where to store abstract-interpretation information.
-%We hypothesize that a more uniform approach to integrating abstract interpretation with rewriting and partial evaluation would solve this problem.)
-
-%% BRING BACK \label{sec:fusing-compiler-passes} subsubsection from appendix when we have plenty of space.
+\subsubsection{New Middle-End Variations}\label{sec:iteration:new-middle-end}
+
+Performance-debugging runs of our generated compilers led us to various ideas for improvement, which were quite easy to add locally.
+For instance, some bitmasking operations (applying bitwise ``and'' with constants) can be proved to be no-ops, based on properties of the terms being masked.
+This realization led us to add several new rules.
+
+On our way to code generation for different target languages, we often make different decisions based on which broad families those targets belong to.
+For instance, only some platforms support conditional-move instructions.
+We added easy compiler configuration of whether to try to use conditional moves.
+Based on such a flag, we choose different generated rewriters, which have been described with largely the same rewrite rules, just including or omitting conditional-move-oriented rules as appropriate.
+
+\subsubsection{New Backends}\label{sec:iteration:new-backend}
+Two new backends were added after this framework was integrated into Fiat Cryptography.
+A backend is one set of rewrite rules, motivated by a specific target machine, that can be composed with the frontend rules that are independent of target.
+Each such combination can be used to generate a separate (proved) command-line tool.
+
+The first new backend involved a separate set of identifiers for an unusual hardware platform with special instructions to accelerate cryptographic arithmetic.
+The rewrite rules we added merely translated the constructs in our intermediate representation to the corresponding identifiers in this output language.
+
+The other backend we added required all arithmetic operations to operate on a single bitwidth, rather than mixing, say, 64-bit and 128-bit integers in the same function.
+To perform this translation, we add rewrite rules for all operations that produce integers of different bitwidths than they consume.
+These rewrite rules replace such an operation with a couple of variable assignments in terms of operations on the smaller bitwidth, together with a special function \mintinline{coq}{Z.combine_at_bitwidth} which is defined to combine two smaller-bitwidth integers into a larger-bitwidth integer.
+We also added rewrite rules about how to push various operations on the larger bitwidth through this \mintinline{coq}{Z.combine_at_bitwidth} construct (for example, a rule about how to turn a multiplication of two outputs of \mintinline{coq}{Z.combine_at_bitwidth} into \mintinline{coq}{Z.combine_at_bitwidth} applied to multiplications of the high and low bits of the inputs).
+By designating \mintinline{coq}{Z.combine_at_bitwidth} as an identifier that should be inlined rather than let-bound during the rewriting, we automatically get a complete compilation of bitwidth-splitting without having to run the rewrite rules repeatedly.%\todo{I think this sentence is very unclear, but I'm not sure how to make the point that we don't need the ``do again'' marker here, because the ordering of rewriting works out well, and that there's some coolness in the subtlty of how it works out well.
+Maybe it doesn't need to be mentioned?}
+
+\subsubsection{Moving Rules Involving Carries}\label{sec:iteration:carries}
+We originally had rules like ``adding 0 to $x$ produces just $x$, with no carry.''
+This rule is not true in general, because we must encode carrying as happening at a particular bitwidth, while we have no guarantee that $x$ fits within that bitwidth.
+The correct rule has a precondition: that $x$ is between 0 and $2^{64}$.
+
+We discovered this issue when trying to prove our rewrite rules correct.
+As a result, we had to move this sort of rule from happening before abstract interpretation to happening after abstract interpretation.
+Since the passes are just defined as lists of rewrite rules, moving the rules was quite simple.
+(Rephrasing them to talk about the outputs of abstract interpretation was somewhat painful, though, because the proof assistant did not enforce the conventions we were using for where to store abstract-interpretation information.
+We hypothesize that a more uniform approach to integrating abstract interpretation with rewriting and partial evaluation would solve this problem.)
+
+\subsubsection{Fusing Compiler Passes}\label{sec:fusing-compiler-passes}
+
+When we moved the aforementioned constant-folding rules from before abstract interpretation to after it, the performance of our compiler on Word-by-Word Montgomery code synthesis decreased significantly.
+(The generated code did not change.)
+We discovered that the number of variable assignments in our intermediate code was quartic in the number of bits in the prime, while the number of variable assignments in the generated code is only quadratic.
+The performance numbers we measured supported this theory: the overall running time of synthesizing code for a prime near $2^k$ jumped from $\Theta(k^2)$ to $\Theta(k^4)$ when we made this change.
+We believe that fusing abstract interpretation with rewriting and partial evaluation would allow us to fix this asymptotic-complexity issue.
+
+To make this situation more concrete, consider the following example:
+Fiat Cryptography uses abstract interpretation to perform bounds analysis; each expression is associated with a range that describes the lower and upper bounds of values that expression might take on.
+Abstract interpretation on addition works as follows: if we have that $x_\ell \le x \le x_u$ and $y_\ell \le y \le y_u$, then we have that $x_\ell + y_\ell \le x + y \le x_u + y_u$.
+Performing bounds analysis on $+$ requires two additions.
+We might have an arithmetic simplification that says that $x + y = x$ whenever we know that $0 \le y \le 0$.
+If we perform the abstract interpretation and then the arithmetic simplification, we perform two additions (for the bounds analysis) and then two comparisons (to test the lower and upper bounds of $y$ for equality with 0).
+We cannot perform the arithmetic simplification before abstract interpretation, because we will not know the bounds of $y$.
+However, if we perform the arithmetic simplification for each expression after performing bounds analysis on its \emph{subexpressions} and only after this perform abstract interpretation on the resulting expression, then we need not use any additions to compute the bounds of $x + y$ when $0 \le y \le 0$, since the expression will just become $x$.
+
+Another essential pass to fuse with rewriting and partial evaluation is let-lifting.
+Unless all of the code is CPS-converted ahead of time, attempting to do let-lifting via rewriting, as must be done when using \mintinline{coq}{setoid_rewrite}, \mintinline{coq}{rewrite_strat}, or \Rtac, results in slower asymptotics.
+This pattern is already apparent in the \texttt{LiftLetsMap} / ``Binders and Recursive Functions'' example in \autoref{sec:micro:LiftLetsMap}.
+% POPL 2022 review TODO:
+% line 852, the forward pointer to section 5.2.4 is somewhat unfortunate; the text cannot be fully understood without looking ahead, it seems.
+% We do not know what the parameters `n` and `m` represents.
+% On line 853, it is not clear what "cubic" means (cubic in what parameter?).
+% On line 858, what is `make`?
+% What is `map_dbl`? The whole paragraph (848-861) is extremely technical.
+% It does not clearly explain why it is the fundamentally difficult to perform let-lifting in an efficient way.
+% It also does not explain its opening sentence, "Another essential pass to fuse with rewriting and partial evaluation is let-lifting".
+% Why is it essential?
+% Is it more efficient if fused than if performed separately?
+% Why is that?
+We achieve linear performance in $n\cdot m$ when ignoring the final \taccbv, while \mintinline{coq}{setoid_rewrite} and \mintinline{coq}{rewrite_strat} are both cubic.
+The rewriter in \Rtac\space cannot possibly achieve better than $\mathcal{O}\left(n\cdot m^2\right)$ unless it can be sublinear in the number of rewrites, because our rewriter gets away with a constant number of rewrites (four), plus evaluating recursion principles for a total amount of work $\mathcal{O}(n\cdot m)$.
+But without primitive support for let-lifting, it is instead necessary to lift the lets by rewrite rules, which requires
+$\mathcal{O}\left(n\cdot m^2\right)$ rewrites just to lift the lets.
+The analysis is thus: running \texttt{make} simply gives us $m$ nested applications of \texttt{map\_dbl} to a length-$n$ list.
+To reduce a given call to \texttt{map\_dbl}, all existing let-binders must first be lifted (there are $n\cdot k$ of them on the $k$-innermost-call) across \texttt{map\_dbl}, one-at-a-time.
+Then the \texttt{map\_dbl} adds another $n$ let binders, so we end up doing $\sum_{k=0}^{m} n\cdot k$ lifts, i.e., $n\cdot m(m+1)/2$ rewrites just to lift the lets.
 
@@
 
-\begin{figure*}
+Now we turn to evaluating performance of generated compilers.
+We start with microbenchmarks focusing attention on particular aspects of reduction and rewriting, with \autoref{sec:additionalMicro} going into more detail.
+
+\subsubsection{Rewriting Without Binders}\label{sec:micro:Plus0Tree}
+
+Consider the code defined by the expression $\text{tree}_{n,m}(v)$ in \autoref{fig:micro:Plus0Tree:code}.
+%\todo{make sure the figure spacing is okay}
+%RSolve[{t[0]==m, t[n]==2*t[n-1] + m},t[n],n]
+%t[n]=m(2^{n+1}-1)
+We want to remove all of the${}+0$s.
+There are $\Theta(m \cdot 2^n)$ such rewriting locations.
+We can start from this expression directly, in which case reification alone takes as much time as Coq's \texttt{rewrite}.
+\begin{wrapfigure}[7]{r}{0pt}
+{\small $\begin{aligned}
+\text{iter}_m(v) & = v + \underbrace{0 + 0 + \cdots + 0}_m \\
+\text{tree}_{0,m}(v) &= \text{iter}_m(v + v) \\
+\text{tree}_{n+1,m}(v) &= \text{iter}_m(\text{tree}_{n,m}(v) + \text{tree}_{n,m}(v))
+\end{aligned}$}%
+\caption{\label{fig:micro:Plus0Tree:code}Expressions computing initial code}
+\end{wrapfigure}
+As the reification method was not especially optimized, and there exist fast reification methods~\cite{ReificationITP18}, we instead start from a call to a recursive function that generates such an expression.
+
+% POPL 2022 review TODO:
+% The graphs in Figure 3 are pretty ugly because they have so many data points packed together.
+% It is nonetheless possible to distinguish the various series involved though, with a bit of geometrical deduction.
+% And I suppose you'd lose some information if you thinned out the data points or plotted lines only.
+% So I don't really have a concrete suggestion for improving the graphs; I just don't find them very pleasing to look at.
+\begin{figure*}[b]
   \newsavebox{\NestedBindersSubfloat}%
   \sbox{\NestedBindersSubfloat}{%
-    \adjustbox{valign=t}{\resizebox{0.32\textwidth}{!}{\beginTikzpictureStamped[only marks]{
+    \adjustbox{valign=t}{\resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
       \einput{perf-UnderLetsPlus0.csv.md5}
@@
   \sbox{\BindersAndRecursiveFunctionsSubfloat}{%
-    \adjustbox{valign=t}{\resizebox{0.32\textwidth}{!}{\beginTikzpictureStamped[only marks]{
+    \adjustbox{valign=t}{\resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
         \einput{perf-UnderLetsPlus0.csv.md5}
@@
     \edef\nval{\NoBindersSubfloatNval}%
-    \adjustbox{valign=t}{\resizebox{0.32\textwidth}{!}{\beginTikzpictureStamped[only marks]{
+    \adjustbox{valign=t}{\resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
         \einput{perf-Plus0Tree.csv.md5}
@@
           xmax=15000,
-          xtick distance=3000,
           table/col sep=comma,
@@
   \sbox{\FiatCryptoSubfloat}{%
-    \adjustbox{valign=t}{\resizebox{0.32\textwidth}{!}{\beginTikzpictureStamped[only marks]{
+    \adjustbox{valign=t}{\resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
       \einput{perf-fiat-crypto.csv.md5}
@@
   \subfloat[Nested binders]{\usebox{\NestedBindersSubfloat}\vphantomSubfloatOne\label{fig:timing-UnderLetsPlus0}}%
-  %% \quad
-  %% \subfloat[Binders and recursive functions]{\usebox{\BindersAndRecursiveFunctionsSubfloat}\vphantomSubfloatTwo\label{fig:timing-LiftLetsMap}}%
+  \quad
+  \subfloat[Binders and recursive functions]{\usebox{\BindersAndRecursiveFunctionsSubfloat}\vphantomSubfloatTwo\label{fig:timing-LiftLetsMap}}%
   \quad
@@
 
-% POPL 2022 review TODO:
-% The graphs in Figure 3 are pretty ugly because they have so many data points packed together.
-% It is nonetheless possible to distinguish the various series involved though, with a bit of geometrical deduction.
-% And I suppose you'd lose some information if you thinned out the data points or plotted lines only.
-% So I don't really have a concrete suggestion for improving the graphs; I just don't find them very pleasing to look at.
-
-Now we turn to evaluating performance of generated compilers.
-We start with microbenchmarks focusing attention on particular aspects of reduction and rewriting, with \autoref{sec:additionalMicro} going into more detail, including on a few more benchmarks.
-
-Our first example family, \emph{nested binders}, has two integer parameters $n$ and $m$.
-An expression tree is built with $2^n$ copies of an expression, which is itself a free variable with $m$ ``useless'' additions of zero.
-We want to see all copies of this expression reduced to just the variable.
-\autoref{fig:timing-Plus0Tree} shows the results for $n = \NoBindersSubfloatNval$ as we scale $m$.
+%RSolve[{t[0]==m, t[n]==2*t[n-1] + m},t[n],n]
+%t[n]=m(2^{n+1}-1)
+\Vref{fig:timing-Plus0Tree} shows the results for $n = \NoBindersSubfloatNval$ as we scale $m$.
 The comparison points are Coq's \texttt{rewrite!}, \texttt{setoid\_rewrite}, and \texttt{rewrite\_strat}.
@@
 For our own approach, we list both the total time and the time taken for core execution of a verified rewrite engine, without counting reification (converting goals to ASTs) or its inverse (interpreting results back to normal-looking goals).
+
 The comparison here is very favorable for our approach so long as $m > 2$.
-%% The competing tactics spike upward toward timeouts at just around a thousand rewrite locations, while our engine is still under two seconds for examples with tens of thousands of rewrite locations.
-%% When $m < 2$, Coq's \texttt{rewrite!} tactic does a little bit better than our engine, corresponding roughly to the overhead incurred by our term representation (which, for example, stores the types at every application node) when most of the term is in fact unchanged by rewriting.
-(See \autoref{sec:additionalPlots:Plus0Tree} %\footnote{Like several forward references in this section, this one goes to an appendix included within the main submission page limit, to avoid interrupting the flow in presenting the most important results.}
-for more detailed plots.)
-
-Now consider what happens when we use \mintinline{coq}{let} binders to share subterms within repeated addition of zero, incorporating exponentially many additions with linearly sized terms.
+The competing tactics spike upward toward timeouts at just around a thousand rewrite locations, while our engine is still under two seconds for examples with tens of thousands of rewrite locations.
+When $m < 2$, Coq's \texttt{rewrite!} tactic does a little bit better than our engine, corresponding roughly to the overhead incurred by our term representation (which, for example, stores the types at every application node) when most of the term is in fact unchanged by rewriting.
+See \autoref{sec:additionalPlots:Plus0Tree}%\footnote{Like several forward references in this section, this one goes to an appendix included within the main submission page limit, to avoid interrupting the flow in presenting the most important results.}
+ for more detailed plots.
+
+\subsubsection{Rewriting Under Binders}\label{sec:micro:UnderLetsPlus0}
+
+\begin{wrapfigure}[7]{r}{0pt}
+{\small $\begin{aligned}
+  & \letin[{v_1 := v_0 + v_0 + 0}{}] \\ \noalign{\vskip-7pt}
+  & \vdots \\ \noalign{\vskip-3pt}
+  & \letin[{v_n := v_{n-1} + v_{n-1} + 0}{}] \\
+  & v_n + v_n + 0
+\end{aligned}$}%
+%  & \letin[{v_2 := v_1 + v_1 + 0}{}] \\
+\caption{\label{fig:micro:UnderLetsPlus0:code}Initial code}
+\end{wrapfigure}
+
+Consider now the code in \autoref{fig:micro:UnderLetsPlus0:code}, which is a version of the code above where redundant expressions are shared via \mintinline{coq}{let} bindings.
+
 \Vref{fig:timing-UnderLetsPlus0} shows the results.
@@
 
+\subsubsection{Performance Bottlenecks of Proof-Producing Rewriting}\label{sec:micro:setoid-rewrite-bottlenecks-lite}
+
 Although we have made our comparison against the built-in tactics \mintinline{coq}{setoid_rewrite} and \mintinline{coq}{rewrite_strat}, by analyzing the performance in detail, we can argue that these performance bottlenecks are likely to hold for any proof assistant designed like Coq.
-Detailed debugging reveals five performance bottlenecks in the existing tactics, discussed in \autoref{sec:setoid-rewrite-bottlenecks}.
+Detailed debugging reveals five performance bottlenecks in the existing rewriting tactics, which we discuss in \autoref{sec:setoid-rewrite-bottlenecks}.\footnote{This forward reference goes to an appendix included within the main submission page limit, to avoid interrupting the flow in presenting the most important results.}%Also included within the submission page limit, though of interest mostly to proof-assistant-implementation experts.}
+
+\subsubsection{Binders and Recursive Functions}\label{sec:micro:LiftLetsMap}
+
+\begin{wrapfigure}[11]{r}{0pt}
+{\small %\allowdisplaybreaks
+$\begin{aligned}
+  \text{map\_dbl}(\ell) & \defeq \begin{cases} [] & \text{if }\ell = [] \\
+      \letin[{y := h + h}{}] & \text{if }\ell = h::t \\
+      y :: \text{map\_dbl}(t) &
+      \end{cases} \\
+  \text{make}(n, m, v) & \defeq \begin{cases} [\underbrace{v, \ldots, v}_n] & \text{if }m = 0 \\
+      \text{map\_dbl}(\text{make}(n, m-1, v)) & \text{if }m > 0
+      \end{cases} \\
+  \text{example}_{n, m} & \defeq \forall v,\ \text{make}(n, m, v) = []
+\end{aligned}$}%
+\caption{\label{fig:micro:LiftLetsMap:code}Initial code for binders and recursive functions}
+\end{wrapfigure}
+
+The next experiment uses the code in \autoref{fig:micro:LiftLetsMap:code}.
+%Note that the ${}=[]$ at the end is just a dummy placeholder to allow us to check reduction and rewriting of this term.
+Note that the \letin{} binding blocks further reduction of map\_dbl when we iterate it $m$ times in make, and so we need to take care to preserve sharing when reducing here.
+
+\autoref{fig:timing-LiftLetsMap} compares performance between our approach, \texttt{repeat setoid\_rewrite}, and two variants of \texttt{rewrite\_strat}.
+Additionally, we consider another option, which was adopted by Fiat Cryptography at a larger scale: rewrite our functions to improve reduction behavior.
+% POPL 2022 review TODO:
+% line 958 (and elsewhere), "continuation-passing style [...] allows standard VM-based reduction to achieve good performance".
+% Is it folklore knowledge why writing in CPS style helps?
+% It might be worth briefly recalling why this is so.
+Specifically, both functions are rewritten in continuation-passing style, which makes them harder to read and reason about but allows standard VM-based reduction to achieve good performance.
+The figure shows that \texttt{rewrite\_strat} variants are essentially unusable for this example, with \texttt{setoid\_rewrite} performing only marginally better, while our approach applied to the original, more readable definitions loses ground steadily to VM-based reduction on CPS'd code.
+On the largest terms ($n \cdot m > 20,000$), the gap is 6s vs.\ 0.1s of compilation time, which should often be acceptable in return for simplified coding and proofs, plus the ability to mix proved rewrite rules with built-in reductions.
+Note that about 99\% of the difference between the full time of our method and just the rewriting is spent in the final \taccbv{} at the end, used to denote our output term from reified syntax.
+We blame this performance on the unfortunate fact that reduction in Coq is quadratic in the number of nested binders present; see Coq bug \coqbug{11151}.
+This bug has since been fixed, as of Coq 8.14; see Coq PR~\coqpr{13537}.
+%traded for simplified coding and proofs, plus the ability to mix proved rewrite rules with built-in reductions.
+See \autoref{sec:LiftLetsMap-more} for more on this microbenchmark.
+
+\subsubsection{Full Reduction}\label{sec:micro:Eratosthenes}
+
+The final experiment involves full reduction in computing the Sieve of Eratosthenes, taking inspiration on benchmark choice from \citet{Aehlig}.
+We find in \autoref{fig:timing-SieveOfEratosthenes} that we are slower than \tacvmcompute{}, \tacnativecompute{}, and \taccbv{}, but faster than \mintinline{coq}{lazy}, and of course much faster than \tacsimpl{} and \taccbn{}, which are quite slow.
 
 \subsection{Macrobenchmark: Fiat Cryptography}\label{sec:macro}
+
+\begin{wrapfigure}[19]{r}{0pt}
+  \resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
+      \einput{perf-SieveOfEratosthenes.csv.md5}
+  }
+    \pgfplotsset{every axis legend/.append style={
+        at={(0.5,-0.2)},
+        anchor=north}}
+    \begin{axis}[xlabel=upper bound,
+        ylabel=time (s),
+        ymax=65,
+        xmax=5000,
+        table/col sep=comma,
+        table/x=param-n]
+      \addplot[mark=square,color=red]       table[y=simpl-regression-cubic-user]                {perf-SieveOfEratosthenes.csv};
+      \addplot[mark=triangle,color=red]     table[y=cbn-regression-cubic-user]                  {perf-SieveOfEratosthenes.csv};
+      \addplot[mark=square*,color=red]      table[y=lazy-regression-quadratic-user]             {perf-SieveOfEratosthenes.csv};
+      \addplot[mark=+,color=blue]           table[y=Rewrite-for-gen-user]                       {perf-SieveOfEratosthenes.csv};
+      \addplot[mark=x,color=ForestGreen]    table[y=rewriting-user]                             {perf-SieveOfEratosthenes.csv};
+      \addplot[mark=triangle*,color=orange] table[y=cbv-regression-quadratic-user]              {perf-SieveOfEratosthenes.csv};
+      \addplot[mark=*,color=purple]         table[y=vm-compute-regression-quadratic-user]       {perf-SieveOfEratosthenes.csv};
+      \addplot[mark=-,color=red]            table[y=native-compute(2)-regression-quadratic-real]{perf-SieveOfEratosthenes.csv};
+      \legend{simpl,cbn,lazy,{Our approach including reification, cbv, etc.},Our approach (only rewriting),cbv,vm\_compute,native\_compute}
+    \end{axis}
+  \end{tikzpicture}}
+  \caption{\label{fig:timing-SieveOfEratosthenes}Full evaluation, Sieve of Eratosthenes}
+\end{wrapfigure}
 
@@
 As the figure shows, our approach gives about a 10$\times$--1000$\times$ speed-up over the original Fiat Cryptography pipeline.
-%(We used the same set of prime moduli as in the experiments run by \citet{FiatCryptoSP19}, which were chosen based on searching the archives of an elliptic-curves mailing list for all prime numbers.)
+(We used the same set of prime moduli as in the experiments run by \citet{FiatCryptoSP19}, which were chosen based on searching the archives of an elliptic-curves mailing list for all prime numbers.)
 Inspection of the timing profiles of the original pipeline reveals that reification dominates the timing profile; since partial evaluation is performed by Coq's kernel, reification must happen \emph{after} partial evaluation, and hence the size of the term being reified grows with the size of the output code.
 Also recall that the old approach required rewriting Fiat Cryptography's library of arithmetic functions in continuation-passing style, enduring this complexity in library correctness proofs, while our new approach applies to a direct-style library.
-Finally, the old approach included a custom reflection-based arithmetic simplifier for term syntax, run after traditional reduction, whereas now we are able to apply a generic engine that combines both, without requiring more than proving traditional rewrites.
+Finally, the old approach included a custom reflection-based arithmetic simplifier for term syntax, run after traditional reduction, whereas now we are able to apply a generic engine that combines both, without requiring more than proving traditional rewrite rules.
 
@@
 
+\section{Related Work, Take Two}\label{sec:related}
+
+We have already discussed the work of \citet{Aehlig}, which introduced the basic structure that our engine shares, but which required a substantially larger trusted code base, did not tackle certain challenges in scaling to large partial-evaluation problems, and did not report any performance experiments in partial evaluation.
+
+We have also mentioned \Rtac{}~\cite{rtac}, which provides tactics as proved functions.
+Its rewrite tactic does not support preservation of subterm sharing, which is performance-prohibitive for the kind of compilation that interested us.
+There is also no support (in \Rtac{} or the standard Coq tactics) for rewriting in normalization-by-evaluation order, which turns out to be perfect for good performance scaling in these compilers.
+
+Our implementation builds on fast full reduction in Coq's kernel, via a virtual machine~\cite{vmcompute} or compilation to native code~\cite{nativecompute}.
+Especially the latter is similar in adopting an NbE style for full reduction, simplifying even under $\lambda$s, on top of a more traditional implementation of OCaml that never executes preemptively under $\lambda$s.
+Neither approach unifies support for rewriting with proved rules, and partial evaluation only applies in very limited cases, where functions that should not be evaluated at compile time must have properly opaque definitions that the evaluator will not consult.
+Neither implementation involved a machine-checked proof.
+
+\citet{Boespflug2009} discusses optimizations on (untyped) NbE for use in proof assistants.
+They mention the reuse of metalanguage pattern-matching facilities on datatypes as an optimization, wherein using metalanguage constructors for constants breaks modularity of the reduction engine.
+We achieve this optimization while avoiding this issue by automatically creating on-the-fly a datatype of constants as a sort of early compilation phase.
+This provides a way to have constructors in the object language be represented as constructors in the metalanguage, in a typed NbE setting, without breaking modularity nor incurring extra work for the user, for a minor upfront time cost.
+
+% A more limited form of code generation for cryptography was already widespread through libraries like OpenSSL, which specifically uses Perl scripts to generate assembly code.
+% The Vale tool suite~\cite{Vale} formalizes these practices with a more principled language and associated verification tools.
+% However, code generation done in this style is significantly simpler than what we treat here, amounting mostly to loop unrolling, macro substitution, and computation of compile-time constants.
+% Also, Vale involves a significantly larger trusted code base than with our approach, with no reduction to some kernel proof checker, instead placing trust in language-specific tooling and an SMT solver.
+
+A variety of forms of pragmatic partial evaluation have been demonstrated, with Lightweight Modular Staging~\cite{LMS} in Scala as one of the best-known current examples.
+A kind of type-based overloading for staging annotations is used to smooth the rough edges in writing code that manipulates syntax trees.
+The LMS-Verify system~\cite{LMSVerify} can be used for formal verification of generated code after-the-fact.
+Typically LMS-Verify has been used with relatively shallow properties (though potentially applied to larger and more sophisticated code bases than we tackle), not scaling to the kinds of functional-correctness properties that concern us here, justifying investment in verified partial evaluators.
+
 \section{Future Work}
@@
 By far the biggest next step for our engine is to integrate abstract interpretation with rewriting and partial evaluation.
-We expect this would net us asymptotic performance gains
-as described in \autoref{sec:fusing-compiler-passes}.
-Additionally, it would allow us to simplify the phrasing of many of our post-abstract-interpretation rewrite rules, by relegating bounds information to side conditions rather than requiring that they appear in the syntactic form of the rule.
+We expect this would net us asymptotic performance gains as described in \autoref{sec:fusing-compiler-passes}.
+Additionally, it would allow us to simplify the phrasing of many of our post-abstract-interpretation rewrite rules, by relegating bounds information to side-conditions rather than requiring that they appear in the syntactic form of the rule.
 
 There are also a number of natural extensions to our engine.
-For instance, we do not yet allow pattern variables marked as ``constants only'' to apply to container datatypes; we limit the mixing of higher-order and polymorphic types, as well as limiting use of first-class polymorphism; we do not support rewriting with equalities of nonfully-applied functions; we only support decidable predicates as rule side conditions, and the predicates may only mention pattern variables restricted to matching constants; we have hardcoded support for a small set of container types and their eliminators; we support rewriting with equality and no other relations% (e.g., subset inclusion)
-; and we require decidable equality for all types mentioned in rules.
-%It may be helpful to design an engine that lifts some or all of these limitations, building on the basic structure that we present here.
+For instance, we do not yet allow pattern variables marked as ``constants only'' to apply to container datatypes; we limit the mixing of higher-order and polymorphic types, as well as limiting use of first-class polymorphism; we do not support rewriting with equalities of nonfully-applied functions; we only support decidable predicates as rule side conditions, and the predicates may only mention pattern variables restricted to matching constants; we have hardcoded support for a small set of container types and their eliminators; we support rewriting with equality and no other relations (e.g., subset inclusion); and we require decidable equality for all types mentioned in rules.
+It may be helpful to design an engine that lifts some or all of these limitations, building on the basic structure that we present here.
+
+%% Acknowledgments
+\begin{acks}                            %% acks environment is optional
+                                        %% contents suppressed with 'anonymous'
+  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
+  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
+  %% acknowledge financial support and will be used by metadata
+  %% extraction tools.
+  This material is based upon work supported by the
+  \grantsponsor{GS100000001}{National Science
+    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
+  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
+  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
+  conclusions or recommendations expressed in this material are those
+  of the author and do not necessarily reflect the views of the
+  National Science Foundation.
+\end{acks}
 
@@
 
-\startsupplement
 %% Appendix
@@
 
-\FloatBarrier\section{Performance Bottlenecks of Proof-Producing Rewriting}\label{sec:setoid-rewrite-bottlenecks}
+\section{Performance Bottlenecks of Proof-Producing Rewriting}\label{sec:setoid-rewrite-bottlenecks}
 
@@
 \clearpage
+\supplement
 
@@
 
-The data for this benchmark can be found on GitHub at \githublink{mit-plv}{fiat-crypto}[perf-testing-data-ITP-2022-rewriting].
-
 \begin{figure*}[b]
@@
 
-\clearpage
-\begin{figure}[t]
-  \subfloat[Expressions computing initial code for Rewriting Without Binders]{%
-  {\small $\begin{aligned}
-  \text{iter}_m(v) & = v + \underbrace{0 + 0 + \cdots + 0}_m \\
-  \text{tree}_{0,m}(v) &= \text{iter}_m(v + v) \\
-  \text{tree}_{n+1,m}(v) &= \text{iter}_m(\text{tree}_{n,m}(v) + \text{tree}_{n,m}(v))
-    \end{aligned}$}%
-  \label{fig:micro:Plus0Tree:code}}
-  \qquad
-  \subfloat[Initial code for Rewriting Under Binders]{%
-  {\small $\begin{aligned}
-  & \letin[{v_1 := v_0 + v_0 + 0}{}] \\ \noalign{\vskip-7pt}
-  & \vdots \\ \noalign{\vskip-3pt}
+\section{Additional Information on Microbenchmarks}\label{sec:additionalMicro}
+
+We performed all benchmarks on a 3.5 GHz Intel Haswell running Linux and Coq 8.11.1.
+We name the subsections here with the names that show up in the code supplement.
+
+\subsection{UnderLetsPlus0}\label{sec:UnderLetsPlus0-more}
+
+We provide more detail on the ``nested binders'' microbenchmark of \autoref{sec:micro:UnderLetsPlus0} displayed in \autoref{fig:timing-UnderLetsPlus0}.
+
+Recall that we are removing all of the${}+0$s from
+{\small \begin{align*}
+  & \letin[{v_1 := v_0 + v_0 + 0}{}] \\
+%  & \letin[{v_2 := v_1 + v_1 + 0}{}] \\
+  & \vdots \\
   & \letin[{v_n := v_{n-1} + v_{n-1} + 0}{}] \\
   & v_n + v_n + 0
-\end{aligned}$}%
-%  & \letin[{v_2 := v_1 + v_1 + 0}{}] \\
-  \label{fig:micro:UnderLetsPlus0:code}%
-  }
-  \caption{\label{fig:micro:Plus0Tree+UnderLetsPlus0:code}Code for rewriting without and under binders}
-\end{figure}
-
-\FloatBarrier\section{Additional Information on Microbenchmarks}\label{sec:additionalMicro}
-
-We performed all benchmarks on a 3.5 GHz Intel Haswell running Linux and Coq 8.11.1.
-We name the subsections here with the names that show up in the code supplement.
-
-\subsection{Rewriting Without Binders: Plus0Tree}\label{sec:Plus0Tree-more}\label{sec:micro:Plus0Tree}
-
-Consider the code defined by the expression $\text{tree}_{n,m}(v)$ in \autoref{fig:micro:Plus0Tree:code}.
-%\todo{make sure the figure spacing is okay}
-%RSolve[{t[0]==m, t[n]==2*t[n-1] + m},t[n],n]
-%t[n]=m(2^{n+1}-1)
-We want to remove all of the${}+0$s.
-There are $\Theta(m \cdot 2^n)$ such rewriting locations.
-We can start from this expression directly, in which case reification alone takes as much time as Coq's \texttt{rewrite}.
-As the reification method was not especially optimized, and there exist fast reification methods~\cite{ReificationITP18}, we instead start from a call to a recursive function that generates such an expression.
-
-%This is a version of \autoref{sec:UnderLetsPlus0-more} without any let binders, discussed in \autoref{sec:micro:Plus0Tree} but not displayed in \autoref{fig:multi-timing}.
+\end{align*}}%
+
+The code used to define this microbenchmark is
+\begin{minted}[fontsize=\small]{coq}
+Definition make_lets_def (n:nat) (v acc : Z) :=
+ @nat_rect (fun _ => Z * Z -> Z)
+   (fun '(v, acc) => acc + acc + v)
+   (fun _ rec '(v, acc) =>
+     dlet acc := acc + acc + v in rec (v, acc))
+   n
+   (v, acc).
+\end{minted}
+We note some details of the rewriting framework that were glossed over in the main body of the paper, which are useful for using the code:
+Although the rewriting framework does not support dependently typed constants, we can automatically preprocess uses of eliminators like \mintinline{coq}{nat_rect} and \mintinline{coq}{list_rect} into nondependent versions.
+The tactic that does this preprocessing is extensible via \Ltac{}'s reassignment feature.
+Since pattern-matching compilation mixed with NbE requires knowing how many arguments a constant can be applied to, we must internally use a version of the recursion principle whose type arguments do not contain arrows; current preprocessing can handle recursion principles with either no arrows or one arrow in the motive.
+Even though we will eventually plug in 0 for $v$, we jump through some extra hoops to ensure that our rewriter cannot cheat by rewriting away the ${}+0$ before reducing the recursion on $n$.
+
+We can reduce this expression in three ways.
+
+\subsubsection{Our Rewriter}
+One lemma is required for rewriting with our rewriter:
+\begin{minted}[fontsize=\small]{coq}
+Lemma Z.add_0_r : forall z, z + 0 = z.
+\end{minted}
+
+Creating the rewriter takes about 12 seconds on the machine we used for running the performance experiments:
+\begin{minted}[fontsize=\small]{coq}
+Make myrew := Rewriter For (Z.add_0_r, eval_rect nat, eval_rect prod).
+\end{minted}
+Recall from \autoref{sec:explain-eval-rect} that \mintinline{coq}{eval_rect} is a definition provided by our framework for eagerly evaluating recursion associated with certain types.
+It functions by triggering typeclass resolution for the lemmas reducing the recursion principle associated to the given type.
+We provide instances for \texttt{nat}, \texttt{prod}, \texttt{list}, \texttt{option}, and \texttt{bool}.
+Users may add more instances if they desire.
+
+\subsubsection{\texorpdfstring{\texttt{setoid\_rewrite}}{setoid\_rewrite} and \texorpdfstring{\texttt{rewrite\_strat}}{rewrite\_strat}}
+To give as many advantages as we can to the preexisting work on rewriting, we pre-reduce the recursion on \mintinline{coq}{nat}s using \taccbv{} before performing \texttt{setoid\_rewrite}.
+(Note that \texttt{setoid\_rewrite} cannot itself perform reduction without generating large proof terms, and \texttt{rewrite\_strat} is not currently capable of sequencing reduction with rewriting internally due to bugs such as \coqbug{10923}.)
+Rewriting itself is easy; we may use any of \texttt{repeat setoid\_rewrite Z.add\_0\_r}, \texttt{rewrite\_strat topdown Z.add\_0\_r}, or \texttt{rewrite\_strat bottomup Z.add\_0\_r}.
+
+\subsection{Plus0Tree}\label{sec:Plus0Tree-more}
+
+This is a version of \autoref{sec:UnderLetsPlus0-more} without any let binders, discussed in \autoref{sec:micro:Plus0Tree} but not displayed in \autoref{fig:multi-timing}.
 
@@
 
-\subsection{Rewriting Under Binders: UnderLetsPlus0}\label{sec:UnderLetsPlus0-more}\label{sec:micro:UnderLetsPlus0}
-
-Consider now the code in \autoref{fig:micro:UnderLetsPlus0:code}, which is a version of the code above where redundant expressions are shared via \mintinline{coq}{let} bindings.
-
-The code used to define this microbenchmark is
-\begin{minted}[fontsize=\small]{coq}
-Definition make_lets_def (n:nat) (v acc : Z) :=
- @nat_rect (fun _ => Z * Z -> Z)
-   (fun '(v, acc) => acc + acc + v)
-   (fun _ rec '(v, acc) =>
-     dlet acc := acc + acc + v in rec (v, acc))
-   n
-   (v, acc).
-\end{minted}
-We note some details of the rewriting framework that were glossed over in the main body of the paper, which are useful for using the code:
-Although the rewriting framework does not support dependently typed constants, we can automatically preprocess uses of eliminators like \mintinline{coq}{nat_rect} and \mintinline{coq}{list_rect} into nondependent versions.
-The tactic that does this preprocessing is extensible via \Ltac{}'s reassignment feature.
-Since pattern-matching compilation mixed with NbE requires knowing how many arguments a constant can be applied to, we must internally use a version of the recursion principle whose type arguments do not contain arrows; current preprocessing can handle recursion principles with either no arrows or one arrow in the motive.
-Even though we will eventually plug in 0 for $v$, we jump through some extra hoops to ensure that our rewriter cannot cheat by rewriting away the ${}+0$ before reducing the recursion on $n$.
-
-We can reduce this expression in three ways.
-
-\subsubsection{Our Rewriter}
-One lemma is required for rewriting with our rewriter:
-\begin{minted}[fontsize=\small]{coq}
-Lemma Z.add_0_r : forall z, z + 0 = z.
-\end{minted}
-
-Creating the rewriter takes about 12 seconds on the machine we used for running the performance experiments:
-\begin{minted}[fontsize=\small]{coq}
-Make myrew := Rewriter For (Z.add_0_r, eval_rect nat, eval_rect prod).
-\end{minted}
-Recall from \autoref{sec:explain-eval-rect} that \mintinline{coq}{eval_rect} is a definition provided by our framework for eagerly evaluating recursion associated with certain types.
-It functions by triggering typeclass resolution for the lemmas reducing the recursion principle associated to the given type.
-We provide instances for \texttt{nat}, \texttt{prod}, \texttt{list}, \texttt{option}, and \texttt{bool}.
-Users may add more instances if they desire.
-
-\subsubsection{\texorpdfstring{\texttt{setoid\_rewrite}}{setoid\_rewrite} and \texorpdfstring{\texttt{rewrite\_strat}}{rewrite\_strat}}
-To give as many advantages as we can to the preexisting work on rewriting, we pre-reduce the recursion on \mintinline{coq}{nat}s using \taccbv{} before performing \texttt{setoid\_rewrite}.
-(Note that \texttt{setoid\_rewrite} cannot itself perform reduction without generating large proof terms, and \texttt{rewrite\_strat} is not currently capable of sequencing reduction with rewriting internally due to bugs such as \coqbug{10923}.)
-Rewriting itself is easy; we may use any of \texttt{repeat setoid\_rewrite Z.add\_0\_r}, \texttt{rewrite\_strat topdown Z.add\_0\_r}, or \texttt{rewrite\_strat bottomup Z.add\_0\_r}.
-
-\subsection{Binders and Recursive Functions: LiftLetsMap}\label{sec:LiftLetsMap-more}\label{sec:micro:LiftLetsMap}
-
-\begin{figure}
-  {\small %\allowdisplaybreaks
-  $\begin{aligned}
-    \text{map\_dbl}(\ell) & \defeq \begin{cases} [] & \text{if }\ell = [] \\
-        \letin[{y := h + h}{}] & \text{if }\ell = h::t \\
-        y :: \text{map\_dbl}(t) &
-        \end{cases} \\
-    \text{make}(n, m, v) & \defeq \begin{cases} [\underbrace{v, \ldots, v}_n] & \text{if }m = 0 \\
-        \text{map\_dbl}(\text{make}(n, m-1, v)) & \text{if }m > 0
-        \end{cases} \\
-    \text{example}_{n, m} & \defeq \forall v,\ \text{make}(n, m, v) = []
-  \end{aligned}$}%
-  \caption{\label{fig:micro:LiftLetsMap:code}Initial code for binders and recursive functions}
-\end{figure}
-
-\begin{figure}
-  \adjustbox{valign=t}{\resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
-      \einput{perf-UnderLetsPlus0.csv.md5}
-  }
-    \pgfplotsset{every axis legend/.append style={
-        at={(0.5,-0.2)},
-        anchor=north}}
-    \begin{axis}[xlabel=$n\cdot m$,
-        ylabel=time (s),
-        scaled x ticks=false,
-        ymax=27,
-        xmax=12000,
-        table/col sep=comma,
-        table/x=param-0-nm]
-      \addplot[mark=o,color=red]         table[y=rewrite-strat(bottomup-bottomup)-regression-exponential-user]{perf-LiftLetsMap.csv};
-      \addplot[mark=triangle,color=red]  table[y=rewrite-strat(topdown-bottomup)-regression-exponential-user]{perf-LiftLetsMap.csv};
-      \addplot[mark=square,color=red]    table[y=setoid-rewrite-regression-cubic-user]{perf-LiftLetsMap.csv};
-      \addplot[mark=+,color=blue]        table[y=Rewrite-for-gen-user]{perf-LiftLetsMap.csv};
-      \addplot[mark=x,color=ForestGreen] table[y=rewriting-user]{perf-LiftLetsMap.csv};
-      \addplot[mark=*,color=purple]      table[y=cps+vm-compute-regression-quadratic-user]{perf-LiftLetsMap.csv};
-      \legend{rewrite\_strat bottomup,rewrite\_strat topdown,repeat setoid\_rewrite,{Our approach including reification, cbv, etc.},Our approach (only rewriting),cps+vm\_compute}
-    \end{axis}
-  \end{tikzpicture}}}
-  \caption{\label{fig:timing-LiftLetsMap}Benchmark with recursive functions}
-\end{figure}
-
-The next experiment uses the code in \autoref{fig:micro:LiftLetsMap:code}.
-%Note that the ${}=[]$ at the end is just a dummy placeholder to allow us to check reduction and rewriting of this term.
-Note that the \letin{} binding blocks further reduction of map\_dbl when we iterate it $m$ times in make, and so we need to take care to preserve sharing when reducing here.
-
-\autoref{fig:timing-LiftLetsMap} compares performance between our approach, \texttt{repeat setoid\_rewrite}, and two variants of \texttt{rewrite\_strat}.
-Additionally, we consider another option, which was adopted by Fiat Cryptography at a larger scale: rewrite our functions to improve reduction behavior.
-% POPL 2022 review TODO:
-% line 958 (and elsewhere), "continuation-passing style [...] allows standard VM-based reduction to achieve good performance".
-% Is it folklore knowledge why writing in CPS style helps?
-% It might be worth briefly recalling why this is so.
-Specifically, both functions are rewritten in continuation-passing style, which makes them harder to read and reason about but allows standard VM-based reduction to achieve good performance.
-The figure shows that \texttt{rewrite\_strat} variants are essentially unusable for this example, with \texttt{setoid\_rewrite} performing only marginally better, while our approach applied to the original, more readable definitions loses ground steadily to VM-based reduction on CPS'd code.
-On the largest terms ($n \cdot m > 20,000$), the gap is 6s vs.\ 0.1s of compilation time, which should often be acceptable in return for simplified coding and proofs, plus the ability to mix proved rewrite rules with built-in reductions.
-Note that about 99\% of the difference between the full time of our method and just the rewriting is spent in the final \taccbv{} at the end, used to denote our output term from reified syntax.
-We blame this performance on the unfortunate fact that reduction in Coq is quadratic in the number of nested binders present; see Coq bug \coqbug{11151}.
-This bug has since been fixed, as of Coq 8.14; see Coq PR~\coqpr{13537}.
-%traded for simplified coding and proofs, plus the ability to mix proved rewrite rules with built-in reductions.
-
-We can perform this rewriting in four ways.
+\subsection{LiftLetsMap}\label{sec:LiftLetsMap-more}
+
+We now discuss in more detail the ``binders and recursive functions'' example from \autoref{sec:micro:LiftLetsMap}.
+
+The expression we want to get out at the end looks like:
+\begin{align*}
+    & \letin[{v_{1,1} := v + v}{}] \\
+    & \vdots \\
+    & \letin[{v_{1,n} := v + v}{}] \\
+    & \letin[{v_{2,1} := v_{1,1} + v_{1,1}}{}] \\
+    & \vdots \\
+    & \letin[{v_{2,n} := v_{1,n} + v_{1,n}}{}] \\
+    & \vdots \\
+    & [v_{m,1}, \ldots, v_{m,n}]
+\end{align*}
+
+Recall that we make this example with the code
+\begin{minted}[fontsize=\small]{coq}
+Definition map_double (ls : list Z) :=
+  list_rect _ [] (λ x xs rec, let y := x + x in y :: rec) ls.
+Definition make (n : nat) (m : nat) (v : Z) :=
+  nat_rect _ (List.repeat v n) (λ _ rec, map_double rec) m.
+\end{minted}
+
+We can perform this rewriting in four ways; see \autoref{fig:timing-LiftLetsMap}.
 
@@
 
-\subsection{SieveOfEratosthenes}\label{sec:Eratosthenes}\label{sec:micro:Eratosthenes}
-
-The final experiment involves full reduction in computing the Sieve of Eratosthenes, taking inspiration on benchmark choice from \citet{Aehlig}.
-We find in \autoref{fig:timing-SieveOfEratosthenes} that we are slower than \tacvmcompute{}, \tacnativecompute{}, and \taccbv{}, but faster than \mintinline{coq}{lazy}, and of course much faster than \tacsimpl{} and \taccbn{}, which are quite slow.
-
-\begin{figure}
-  \resizebox{0.4\textwidth}{!}{\beginTikzpictureStamped[only marks]{
-      \einput{perf-SieveOfEratosthenes.csv.md5}
-  }
-    \pgfplotsset{every axis legend/.append style={
-        at={(0.5,-0.2)},
-        anchor=north}}
-    \begin{axis}[xlabel=upper bound,
-        ylabel=time (s),
-        ymax=65,
-        xmax=5000,
-        table/col sep=comma,
-        table/x=param-n]
-      \addplot[mark=square,color=red]       table[y=simpl-regression-cubic-user]                {perf-SieveOfEratosthenes.csv};
-      \addplot[mark=triangle,color=red]     table[y=cbn-regression-cubic-user]                  {perf-SieveOfEratosthenes.csv};
-      \addplot[mark=square*,color=red]      table[y=lazy-regression-quadratic-user]             {perf-SieveOfEratosthenes.csv};
-      \addplot[mark=+,color=blue]           table[y=Rewrite-for-gen-user]                       {perf-SieveOfEratosthenes.csv};
-      \addplot[mark=x,color=ForestGreen]    table[y=rewriting-user]                             {perf-SieveOfEratosthenes.csv};
-      \addplot[mark=triangle*,color=orange] table[y=cbv-regression-quadratic-user]              {perf-SieveOfEratosthenes.csv};
-      \addplot[mark=*,color=purple]         table[y=vm-compute-regression-quadratic-user]       {perf-SieveOfEratosthenes.csv};
-      \addplot[mark=-,color=red]            table[y=native-compute(2)-regression-quadratic-real]{perf-SieveOfEratosthenes.csv};
-      \legend{simpl,cbn,lazy,{Our approach including reification, cbv, etc.},Our approach (only rewriting),cbv,vm\_compute,native\_compute}
-    \end{axis}
-  \end{tikzpicture}}
-  \caption{\label{fig:timing-SieveOfEratosthenes}Full evaluation, Sieve of Eratosthenes}
-\end{figure}
+\subsection{SieveOfEratosthenes}\label{sec:Eratosthenes}
 
@@
 
-\FloatBarrier\section{Fusing Compiler Passes}\label{sec:fusing-compiler-passes}
-
-When we moved the
-%aforementioned constant-folding rules
-%constant-folding rules from \autoref{sec:iteration:carries}
-constant-folding rules
-from before abstract interpretation to after it, the performance of our compiler on Word-by-Word Montgomery code synthesis decreased significantly.
-(The generated code did not change.)
-We discovered that the number of variable assignments in our intermediate code was quartic in the number of bits in the prime, while the number of variable assignments in the generated code is only quadratic.
-The performance numbers we measured supported this theory: the overall running time of synthesizing code for a prime near $2^k$ jumped from $\Theta(k^2)$ to $\Theta(k^4)$ when we made this change.
-We believe that fusing abstract interpretation with rewriting and partial evaluation would allow us to fix this asymptotic-complexity issue.
-
-To make this situation more concrete, consider the following example:
-Fiat Cryptography uses abstract interpretation to perform bounds analysis; each expression is associated with a range that describes the lower and upper bounds of values that expression might take on.
-Abstract interpretation on addition works as follows: if we have that $x_\ell \le x \le x_u$ and $y_\ell \le y \le y_u$, then we have that $x_\ell + y_\ell \le x + y \le x_u + y_u$.
-Performing bounds analysis on $+$ requires two additions.
-We might have an arithmetic simplification that says that $x + y = x$ whenever we know that $0 \le y \le 0$.
-If we perform the abstract interpretation and then the arithmetic simplification, we perform two additions (for the bounds analysis) and then two comparisons (to test the lower and upper bounds of $y$ for equality with 0).
-We cannot perform the arithmetic simplification before abstract interpretation, because we will not know the bounds of $y$.
-However, if we perform the arithmetic simplification for each expression after performing bounds analysis on its \emph{subexpressions} and only after this perform abstract interpretation on the resulting expression, then we need not use any additions to compute the bounds of $x + y$ when $0 \le y \le 0$, since the expression will just become $x$.
-
-Another essential pass to fuse with rewriting and partial evaluation is let-lifting.
-Unless all of the code is CPS-converted ahead of time, attempting to do let-lifting via rewriting, as must be done when using \mintinline{coq}{setoid_rewrite}, \mintinline{coq}{rewrite_strat}, or \Rtac, results in slower asymptotics.
-This pattern is already apparent in the \texttt{LiftLetsMap} / ``Binders and Recursive Functions'' example in \autoref{sec:micro:LiftLetsMap}.
-% POPL 2022 review TODO:
-% line 852, the forward pointer to section 5.2.4 is somewhat unfortunate; the text cannot be fully understood without looking ahead, it seems.
-% We do not know what the parameters `n` and `m` represents.
-% On line 853, it is not clear what "cubic" means (cubic in what parameter?).
-% On line 858, what is `make`?
-% What is `map_dbl`? The whole paragraph (848-861) is extremely technical.
-% It does not clearly explain why it is the fundamentally difficult to perform let-lifting in an efficient way.
-% It also does not explain its opening sentence, "Another essential pass to fuse with rewriting and partial evaluation is let-lifting".
-% Why is it essential?
-% Is it more efficient if fused than if performed separately?
-% Why is that?
-We achieve linear performance in $n\cdot m$ when ignoring the final \taccbv, while \mintinline{coq}{setoid_rewrite} and \mintinline{coq}{rewrite_strat} are both cubic.
-The rewriter in \Rtac\space cannot possibly achieve better than $\mathcal{O}\left(n\cdot m^2\right)$ unless it can be sublinear in the number of rewrites, because our rewriter gets away with a constant number of rewrites (four), plus evaluating recursion principles for a total amount of work $\mathcal{O}(n\cdot m)$.
-But without primitive support for let-lifting, it is instead necessary to lift the lets by rewrite rules, which requires
-$\mathcal{O}\left(n\cdot m^2\right)$ rewrites just to lift the lets.
-The analysis is thus: running \texttt{make} simply gives us $m$ nested applications of \texttt{map\_dbl} to a length-$n$ list.
-To reduce a given call to \texttt{map\_dbl}, all existing let-binders must first be lifted (there are $n\cdot k$ of them on the $k$-innermost-call) across \texttt{map\_dbl}, one-at-a-time.
-Then the \texttt{map\_dbl} adds another $n$ let binders, so we end up doing $\sum_{k=0}^{m} n\cdot k$ lifts, i.e., $n\cdot m(m+1)/2$ rewrites just to lift the lets.
-
 \section{Experience vs.\ Lean and \texorpdfstring{\texttt{setoid\_rewrite}}{setoid\_rewrite}\label{sec:lean}}
@@
 The widely used primes tend to have around five to ten limbs.
-See \coqbug{13576} for more details and for updates.
+%See \coqbug{13576} for more details and for updates.
 
@@
 Extrapolating out the exponential asymptotics of the fastest-growing subcall to \texttt{rewrite\_strat} indicates that 5 limbs would take 11--12 hours, 6 limbs would take 10--11 days, 7 limbs would take 31--32 weeks, 8 limbs would take 13--14 years, 9 limbs would take 2--3 centuries, 10 limbs would take 6--7 millennia, and 15 limbs would take 2--3 times the age of the universe, and 17 limbs, the largest example we might find at present in the real world, would take over $1000\times$ the age of the universe!
-See \coqbug{13708} for more details and updates.
-
-This experiment using \verb|rewrite_strat| can be found online in the Coq source file at \texttt{src/fiat\_crypto\_via\_setoid\_rewrite\_standalone.v} on GitHub at \githublink{coq-community}{coq-performance-tests}.
-To test with the two-limb prime $2^{107}-1$, change \verb|Goal goal| to \verb|Goal goal_of_size 2%nat| near the bottom of the file.
+%See \coqbug{13708} for more details and updates.
+
+We have attached this experiment using \verb|rewrite_strat| as \verb|fiat_crypto_via_rewrite_strat.v|, which is meant to be run in emacs/PG from inside the \verb|fiat-crypto| directory, or in \verb|coqc| by setting \verb|COQPATH| to the value emitted by \texttt{make printenv} in \verb|fiat-crypto| and then invoking the command \texttt{coqc -q -R /path/to/fiat-crypto/src Crypto /path/to/fiat\_crypto\_via\_rewrite\_strat.v}.
+To test with the two-limb prime 113, change \verb|of_string "2^5-3" 8| in the definition of \verb|p| to \verb|of_string "2^7-15" 64|.
 
@@
 
-Our experiments with running \texttt{rewrite} in Lean on the Fiat Cryptography code can be found in the file \texttt{fiat-crypto-lean/src/fiat\_crypto.lean} on GitHub at \githublink{mit-plv}{fiat-crypto}[lean].
+We have attached our experiments with running \texttt{rewrite} in Lean on the Fiat Cryptography code as a supplement as well.
 We used Lean version 3.4.2, commit cbd2b6686ddb, Release.
@@
 change \texttt{open ex} to \texttt{open ex2} to try the two-limb example, or to \texttt{open ex5} to try the five-limb example.
-
-\section{Limitations and Preprocessing}\label{sec:implementation-and-usage}
-
-%\todo{It's not clear to Jason what in this section belongs here, and what isn't interesting (yet?) to readers}
-%\todo{I (Jason) have no idea what the story for this section should be}
-%\todo{Insofar as text is actually kept in this section, make sure any duplication between here and the out-of-paper appendix is deliberate}
-%\todo{If this section is moved or removed, be sure to restructure \autoref{sec:code-from-implementation-and-usage} accordingly}
-
-We now note some details of the rewriting framework that were previously glossed over, which are useful for using the code or implementing something similar, but which do not add fundamental capabilities to the approach.
-Although the rewriting framework does not support dependently typed constants, we can automatically preprocess uses of eliminators like \mintinline{coq}{nat_rect} and \mintinline{coq}{list_rect} into nondependent versions.
-The tactic that does this preprocessing is extensible via \Ltac{}'s reassignment feature.
-Since pattern-matching compilation mixed with NbE requires knowing how many arguments a constant can be applied to, internally we must use a version of the recursion principle whose type arguments do not contain arrows; current preprocessing can handle recursion principles with either no arrows or one arrow in motives.
-
-Recall from \autoref{sec:explain-eval-rect} that \mintinline{coq}{eval_rect} is a definition provided by our framework for eagerly evaluating recursion associated with certain types.
-It functions by triggering typeclass resolution for the lemmas reducing the recursion principle associated to the given type.
-We provide instances for \texttt{nat}, \texttt{prod}, \texttt{list}, \texttt{option}, and \texttt{bool}.
-Users may add more instances if they desire.
-
-Recall again from \autoref{sec:explain-ident.eagerly} that we use \mintinline{coq}{ident.eagerly} to ask the reducer to simplify a case of primitive recursion by complete traversal of the designated argument's constructor tree.
-Our current version only allows a limited, hard-coded set of eliminators with \mintinline{coq}{ident.eagerly} (\texttt{nat\_rect} on return types with either zero or one arrows, \texttt{list\_rect} on return types with either zero or one arrows, and \texttt{List.nth\_default}), but nothing in principle prevents automatic generation of the necessary code.
-
-We define a constant \mintinline{coq}{Let_In} which we use for writing \letin{} expressions that do not reduce under $\zeta$ (Coq's reduction rule for \mintinline{coq}{let}-inlining).
-Throughout most of this paper, anywhere that \letin{} appears, we have actually used \mintinline{coq}{Let_In} in the code.
-It would alternatively be possible to extend the reification preprocessor to automatically convert \letin{} to \mintinline{coq}{Let_In}, but this strategy may cause problems when converting the interpretation of the reified term with the prereified term, as Coq's conversion does not allow fine-tuning of when to inline or unfold \mintinline{coq}{let}s.
 
@@
 
+\subsubsection{Code from \autocommanameref{sec:implementation-and-usage}}\label{sec:code-from-implementation-and-usage}
+
+The \Ltac{} hooks for extending the preprocessing of eliminators are \mintinline{coq}{reify_preprocess_extra} and \mintinline{coq}{reify_ident_preprocess_extra} in a submodule of \texttt{rewriter/src/Rewriter/Language/PreCommon.v}.
+These hooks are called by \mintinline{coq}{reify_preprocess} and \mintinline{coq}{reify_ident_preprocess} in a submodule of \texttt{rewriter/src/Rewriter/Language/Language.v}.
+Some recursion lemmas for use with these tactics are defined in the \verb|Thunked| module in \texttt{fiat-crypto/src/Language/PreExtra.v}.
+These tactics are overridden in the file \texttt{fiat-crypto/src/Language/IdentifierParameters.v}.
+
+The typeclass associated to \mintinline{coq}{eval_rect} ({c.f.} \autoref{sec:code:eval-rect}) is \mintinline{coq}{rules_proofs_for_eager_type} defined in \texttt{rewriter/src/Rewriter/Language/Pre.v}.
+The instances we provide by default are defined in a submodule of \texttt{src/Rewriter/Language/PreLemmas.v}.
+
+The hard-coding of the eliminators for use with \mintinline{coq}{ident.eagerly} ({c.f.} \autoref{sec:code:ident.eagerly}) is done in the tactics \mintinline{coq}{reify_ident_preprocess} and \mintinline{coq}{rewrite_interp_eager} in \texttt{rewriter/src/Rewriter/Language/Language.v}, in the inductive type \mintinline{coq}{restricted_ident} and the typeclass \mintinline{coq}{BuildEagerIdentT} in \texttt{rewriter/src/Rewriter/Language/Language.v}, and in the \Ltac{} tactic with the name of \mintinline{coq}{handle_reified_rewrite_rules_interp} defined in the file \texttt{rewriter/src/Rewriter/Rewriter/ProofsCommonTactics.v}.
+
+The \mintinline{coq}{Let_In} constant is defined in \texttt{rewriter/src/Rewriter/Util/LetIn.v}.
+
 \subsection{Code from \autocommanameref{sec:evaluation}}
@@
 
-The rewrite rules
-%from \autocommanameref{sec:iteration:reassoc}
-for reassociating arithmetic
-can be found in \mintinline{coq}{arith_rewrite_rulesT} starting at the comment ``We reassociate some multiplication of small constants'' in \texttt{fiat-crypto/src/Rewriter/Rules.v}.
-
-The
-following
-frontend constructs
-%from \autocommanameref{sec:iteration:new-frontend}
-are in \verb|all_ident_named_interped| defined in \texttt{fiat-crypto/src/Language/IdentifierParameters.v}.
+The rewrite rules from \autocommanameref{sec:iteration:reassoc} can be found in the definition \mintinline{coq}{arith_rewrite_rulesT} starting at the comment ``We reassociate some multiplication of small constants'' in \texttt{fiat-crypto/src/Rewriter/Rules.v}.
+
+The frontend constructs from \autocommanameref{sec:iteration:new-frontend} are in \verb|all_ident_named_interped| defined in \texttt{fiat-crypto/src/Language/IdentifierParameters.v}.
 \begin{itemize}
@@
 
-The rules about bitmasking operations
-%from \autocommanameref{sec:iteration:new-middle-end}
-can be found in \mintinline{coq}{arith_with_casts_rewrite_rulesT} in \texttt{fiat-crypto/src/Rewriter/Rules.v} and involve \mintinline{coq}{Z.land} and \mintinline{coq}{Z.lor}.
+The rules about bitmasking operations from \autocommanameref{sec:iteration:new-middle-end} can be found in \mintinline{coq}{arith_with_casts_rewrite_rulesT} in \texttt{fiat-crypto/src/Rewriter/Rules.v} and involve \mintinline{coq}{Z.land} and \mintinline{coq}{Z.lor}.
 
@@
 
-The rewrite rules for the new backends
-%mentioned in \autocommanameref{sec:iteration:new-backend}
-are defined by \mintinline{coq}{fancy_with_casts_rewrite_rulesT} and \mintinline{coq}{mul_split_rewrite_rulesT} as well as \mintinline{coq}{multiret_split_rewrite_rulesT} as well as \mintinline{coq}{noselect_rewrite_rulesT} in \texttt{fiat-crypto/src/Rewriter/Rules.v}.
+The rewrite rules for the new backends mentioned in \autocommanameref{sec:iteration:new-backend} are defined by \mintinline{coq}{fancy_with_casts_rewrite_rulesT} and \mintinline{coq}{mul_split_rewrite_rulesT} as well as \mintinline{coq}{multiret_split_rewrite_rulesT} and \mintinline{coq}{noselect_rewrite_rulesT} in \texttt{fiat-crypto/src/Rewriter/Rules.v}.
 The special function \mintinline{coq}{Z.combine_at_bitwidth} is defined in \texttt{fiat-crypto/src/Util/ZUtil/Definitions.v}.
-The designation of \mintinline{coq}{Z.combine_at_bitwidth} as an identifier that should be inlined occurs by listing it in the definition \mintinline{coq}{var_like_idents} in the source file \texttt{fiat-crypto/src/Language/IdentifierParameters.v}.
-
-The rules involving carries mentioned in
-%\autocommanameref{sec:iteration:carries} and
-\autocommanameref{sec:fusing-compiler-passes} are in \mintinline{coq}{arith_with_casts_rewrite_rulesT} in \texttt{fiat-crypto/src/Rewriter/Rules.v}.
+The designation of \mintinline{coq}{Z.combine_at_bitwidth} as an identifier that should be inlined occurs by listing it in the definition \mintinline{coq}{var_like_idents} in \texttt{fiat-crypto/src/Language/IdentifierParameters.v}.
+
+The rules involving carries mentioned in \autocommanameref{sec:iteration:carries} and \autocommanameref{sec:fusing-compiler-passes} are in \mintinline{coq}{arith_with_casts_rewrite_rulesT} in \texttt{fiat-crypto/src/Rewriter/Rules.v}.
 
@@
 
-The code for \autoref{fig:timing-Plus0Tree} from \autocommanameref{sec:micro:Plus0Tree} can be found in \texttt{Plus0Tree.v}.
-
-The code for \autoref{fig:timing-UnderLetsPlus0} from \autocommanameref{sec:micro:UnderLetsPlus0} can be found in \texttt{UnderLetsPlus0.v}.
-
-%The code used for the performance investigation mentioned in \autocommanameref{sec:micro:setoid-rewrite-bottlenecks-lite} and detailed in \autoref{sec:setoid-rewrite-bottlenecks} is not part of the framework we are presenting, and thus not in the supplement.
-
-The code for \autoref{fig:timing-LiftLetsMap} from \autocommanameref{sec:micro:LiftLetsMap} can be found in \texttt{LiftLetsMap.v}.
-
-The code for \autoref{fig:timing-SieveOfEratosthenes} from \autocommanameref{sec:micro:Eratosthenes} can be found in \texttt{SieveOfEratosthenes.v}.
+The code from \autocommanameref{sec:micro:Plus0Tree} can be found in \texttt{Plus0Tree.v}.
+
+The code from \autocommanameref{sec:micro:UnderLetsPlus0} can be found in \texttt{UnderLetsPlus0.v}.
+
+The code used for the performance investigation mentioned in \autocommanameref{sec:micro:setoid-rewrite-bottlenecks-lite} and detailed in \autoref{sec:setoid-rewrite-bottlenecks} is not part of the framework we are presenting, and thus not in the supplement.
+
+The code from \autocommanameref{sec:micro:LiftLetsMap} can be found in \texttt{LiftLetsMap.v}.
+
+The code from \autocommanameref{sec:micro:Eratosthenes} can be found in \texttt{SieveOfEratosthenes.v}.
 
@@
 
-\subsection{Code from \autocommanameref{sec:implementation-and-usage}}\label{sec:code-from-implementation-and-usage}
-
-The \Ltac{} hooks for extending the preprocessing of eliminators are \mintinline{coq}{reify_preprocess_extra} and \mintinline{coq}{reify_ident_preprocess_extra} in a submodule of \texttt{rewriter/src/Rewriter/Language/PreCommon.v}.
-These hooks are called by \mintinline{coq}{reify_preprocess} and \mintinline{coq}{reify_ident_preprocess} in a submodule of \texttt{rewriter/src/Rewriter/Language/Language.v}.
-Some recursion lemmas for use with these tactics are defined in the \verb|Thunked| module in \texttt{fiat-crypto/src/Language/PreExtra.v}.
-These tactics are overridden in the file \texttt{fiat-crypto/src/Language/IdentifierParameters.v}.
-
-The typeclass associated to \mintinline{coq}{eval_rect} ({c.f.} \autoref{sec:code:eval-rect}) is \mintinline{coq}{rules_proofs_for_eager_type} defined in \texttt{rewriter/src/Rewriter/Language/Pre.v}.
-The instances we provide by default are defined in a submodule of \texttt{src/Rewriter/Language/PreLemmas.v}.
-
-The hard-coding of the eliminators for use with \mintinline{coq}{ident.eagerly} ({c.f.} \autoref{sec:code:ident.eagerly}) is done in the tactics \mintinline{coq}{reify_ident_preprocess} and \mintinline{coq}{rewrite_interp_eager} in \texttt{rewriter/src/Rewriter/Language/Language.v}, in the inductive type \mintinline{coq}{restricted_ident} and the typeclass \mintinline{coq}{BuildEagerIdentT} in \texttt{rewriter/src/Rewriter/Language/Language.v}, and in the \Ltac{} tactic with the name of \mintinline{coq}{handle_reified_rewrite_rules_interp} defined in the file \texttt{rewriter/src/Rewriter/Rewriter/ProofsCommonTactics.v}.
-
-The \mintinline{coq}{Let_In} constant is defined in \texttt{rewriter/src/Rewriter/Util/LetIn.v}.
-
 \end{document}

### 2022-05-03 papers/source/rewriting/process-old-synthesis.py

@@ -1,0 +1,176 @@
+#!/usr/bin/env python3
+import re, math, csv, sys
+from functools import reduce
+
+def pretty_prime(prime):
+    return prime.replace('-', '+-').replace('p', '+').replace('m', '+-').replace('e', '^').replace('x', '*')
+
+def latex_prime(prime):
+    def latex_exps(expr):
+        return expr if '^' not in expr else '%s^{%s}' % tuple(expr.split('^'))
+    def latex_prods(expr): return r' \cdot '.join(latex_exps(v) for v in expr.split('*'))
+    def latex_sums(expr): return ' + '.join(latex_prods(v) for v in expr.split('+')).replace('+ -', '- ')
+    return latex_sums(pretty_prime(prime))
+
+def eval_prime(prime):
+    def eval_exps(expr): return int(expr) if '^' not in expr else pow(*map(int, expr.split('^')))
+    def eval_prods(expr): return math.prod(eval_exps(v) for v in expr.split('*'))
+    def eval_sums(expr): return sum(eval_prods(v) for v in expr.split('+'))
+    return eval_sums(pretty_prime(prime))
+
+LINE_REG = re.compile(r'^[^ ]*?(montgomery|solinas)([0-9]+)_([^_]+)_([0-9]+)limbs/([^\./]+)\.vo\s+\(real: ([0-9\.]+), user: ([0-9\.]+), sys: ([0-9\.]+), mem: ([0-9]+) ko\)')
+def parse_prime_line(line):
+    for kind, bitwidth, prime, nlimbs, operation, realt, usert, syst, memko in LINE_REG.findall(line):
+        return {'kind': {'montgomery':'WordByWordMontgomery','solinas':'UnsaturatedSolinas'}[kind],
+                'operation': operation,
+                'bitwidth': bitwidth,
+                'prime': eval_prime(prime),
+                'prime-str': prime,
+                'prime-tex': latex_prime(prime),
+                'nlimbs': nlimbs,
+                'real': realt,
+                'user': usert,
+                'sys': syst,
+                'mem': memko
+                }
+
+def rows_to_dict(keys, rows, descr=''):
+    rtn = {}
+    for i, row in enumerate(rows):
+        if row is None: continue
+        try:
+            key = tuple(f(row[k]) for k, f in keys)
+        except:
+            print('Invalid row %d: %s' % (i, str(row)))
+            raise
+        if key not in rtn.keys(): rtn[key] = []
[…]

### 2022-05-03 papers/source/rewriting/rewriting.tex
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@f661247ac1396f9fdc105addf8b50a96e4262555 🚀

@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
+%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
 \RequirePackage{luatex85}
 \PassOptionsToPackage{dvipsnames}{xcolor}
-\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2021}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-%\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-\usepackage{wrapfig}
+\RequirePackage{amssymb}
+% order matters, see https://tex.stackexchange.com/a/83051/2066
+\RequirePackage{varioref}
+\PassOptionsToPackage{nameinlink}{cleverref}%[2012/02/15]% v0.18.4;
+\documentclass[a4paper,USenglish,cleveref,autoref,thm-restate]{lipics-v2021}
+\usepackage{hyperref}
+%This is a template for producing LIPIcs articles.
+%See lipics-v2021-authors-guidelines.pdf for further information.
+%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
+%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
+%for section-numbered lemmas etc., use "numberwithinsect"
+%for enabling cleveref support, use "cleveref"
+%for enabling autoref support, use "autoref"
+%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
+%for enabling thm-restate support, use "thm-restate"
+%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
+%for producing a PDF according the PDF/A standard, add "pdfa"
+
+%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
+%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
+
+%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
+
+\bibliographystyle{plainurl}% the mandatory bibstyle
+
+%\usepackage{wrapfig}% forbidden in ITP
 \setlength{\intextsep}{0pt} % don't add space above nor below
-\usepackage{subfig}
-\newcommand{\subfigureautorefname}{\figureautorefname}
-%% http://ctan.org/pkg/subcaption
-
-\usepackage{hyperref}
+%\usepackage{subfig}
+%\newcommand{\subfigureautorefname}{\figureautorefname}
+
 \usepackage{minted}
 \usemintedstyle{bw}
-\usepackage{amsthm}
-%\usepackage{amssymb}
+\usepackage{amsmath}
+%\usepackage{amsthm}
+\usepackage{amssymb}
+\usepackage{comment}
 \usepackage[all]{xy}
@@
   \newwrite\@appendixpagefile
-  \immediate\openout\@appendixpagefile=\jobname.appendix.startpage
-  \immediate\write\@appendixpagefile{\detokenize{\def\appendixstartpage}{\thepage}}
+  \immediate\openout\@appendixpagefile=\jobname.appendix.startpage\relax
+  \immediate\write\@appendixpagefile{\detokenize{\def\appendixstartpage}{\thepage}}%
 }
-\newcommand{\supplement}{%
+\newcommand{\startsupplement}{%
   \clearpage
   \newwrite\@supplementpagefile
-  \immediate\openout\@supplementpagefile=\jobname.supplement.startpage
-  \immediate\write\@supplementpagefile{\detokenize{\def\supplementstartpage}{\thepage}}
+  \immediate\openout\@supplementpagefile=\jobname.supplement.startpage\relax
+  \immediate\write\@supplementpagefile{\detokenize{\def\supplementstartpage}{\thepage}}%
+  \newwrite\@supplementonlypagefile
+  \immediate\openout\@supplementonlypagefile=\jobname.supplement.onlystartpage\relax
+  \immediate\write\@supplementonlypagefile{\thepage}%
 }%
+\AtEndDocument{%
+  \newwrite\@endsupplementonlypagefile
+  \immediate\openout\@endsupplementonlypagefile=\jobname.supplement.onlyendpage\relax
+  \immediate\write\@endsupplementonlypagefile{\thepage}%
+}
 \makeatother
 
+\title{Accelerating Verified-Compiler Development with a Verified Rewriting Engine}
+
+%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line
+
+\author{Jason Gross}{CSAIL, Massachusetts Institute of Technology, 77 Massachusetts Ave., Cambridge, MA 02139, USA \and MIRI, USA \and \url{https://jasongross.github.io/} }{jgross@mit.edu}{https://orcid.org/0000-0002-9427-4891}{}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty.
+Please provide at least the name of the affiliation and the country.
+The full address is optional.
+Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.
+
+\author{Andres Erbsen}{CSAIL, Massachusetts Institute of Technology, 77 Massachusetts Ave., Cambridge, MA 02139, USA \and \url{https://andres.systems/} }{andreser@mit.edu}{}{}
+
+\author{Jade Philipoom%\footnote{now with Google}
+}{CSAIL, Massachusetts Institute of Technology, 77 Massachusetts Ave., Cambridge, MA 02139, USA \and Google}{jade.philipoom@gmail.com}{}{}
+% should we \footnote{now with Google}?
+
+\author{Miraya Poddar-Agrawal}{Reed College, 3203 SE Woodstock Blvd, Portland, OR 97202, USA% \and \url{https://github.com/miraya-pa}
+}{ragrawal@reed.edu}{https://orcid.org/0000-0001-7617-9180}{}
+
+\author{Adam Chlipala}{CSAIL, Massachusetts Institute of Technology, 77 Massachusetts Ave., Cambridge, MA 02139, USA \and \url{http://adam.chlipala.net/} }{adamc@csail.mit.edu}{https://orcid.org/0000-0001-7085-9417}{%[funding]
+}
+
+\authorrunning{J.
+Gross and A.
+Erbsen and J.
+Philipoom and M.
+Poddar-Agrawal and A.
+Chlipala} %TODO mandatory.
+First: Use abbreviated first/middle names.
+Second (only in severe cases): Use first author plus 'et al.'
+
+\Copyright{Jason Gross \& Andres Erbsen \& Jade Philipoom \& Miraya Poddar-Agrawal \& Adam Chlipala} %TODO mandatory, please use full first names.
+LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
+
+\begin{CCSXML}
+<ccs2012>
+   <concept>
+       <concept_id>10003752.10003790.10002990</concept_id>
+       <concept_desc>Theory of computation~Logic and verification</concept_desc>
+       <concept_significance>300</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10003752.10003790.10003798</concept_id>
+       <concept_desc>Theory of computation~Equational logic and rewriting</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10011007.10011006.10011041</concept_id>
+       <concept_desc>Software and its engineering~Compilers</concept_desc>
+       <concept_significance>300</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10011007.10011006.10011041.10011046</concept_id>
+       <concept_desc>Software and its engineering~Translator writing systems and compiler generators</concept_desc>
+       <concept_significance>300</concept_significance>
+       </concept>
+ </ccs2012>
+\end{CCSXML}
+
+\ccsdesc[300]{Theory of computation~Logic and verification}
+\ccsdesc[500]{Theory of computation~Equational logic and rewriting}
+\ccsdesc[300]{Software and its engineering~Compilers}
+\ccsdesc[300]{Software and its engineering~Translator writing systems and compiler generators}
+
+\keywords{compiler verification, rewriting engines, cryptography} %TODO mandatory; please add comma-separated list of keywords
+
+%\category{} %optional, e.g. invited paper
+
+\ifthenelse{\boolean{tobesplit}}{%
+%  \relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
+\relatedversiondetails%[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]
+    {Full Version}{https://arxiv.org/abs/2205.00862} %linktext and cite are optional
+}{}
+
+\supplement{%
+  \url{https://jasongross.github.io/papers/2022-rewriting-itp-supplement.pdf},
+  \githublinkurl{mit-plv}{rewriter}[ITP-2022-perf-data],
+  \githublinkurl{mit-plv}{fiat-crypto}[perf-testing-data-ITP-2022-rewriting]%
+}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
+%\supplementdetails[linktext={opt.
+text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g.
+Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional
+
+\funding{%
+  This work was supported in part by a Google Research Award, National Science Foundation grants CCF-1253229, CCF-1512611, and CCF-1521584, and the National Science Foundation Graduate Research Fellowship under Grant Nos.\ 1122374 and 1745302.
+  Any opinion, findings, and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the National Science Foundation.%
+}
+%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.
+
+%\acknowledgements{I want to thank \dots}%optional
+
+%\nolinenumbers %uncomment to disable line numbering
+
+%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\EventEditors{June Andronick and Leonardo de Moura}
+\EventNoEds{2}
+\EventLongTitle{13th International Conference on Interactive Theorem Proving (ITP 2022)}
+\EventShortTitle{ITP 2022}
+\EventAcronym{ITP}
+\EventYear{2022}
+\EventDate{August 7--10, 2022}
+\EventLocation{Haifa, Israel}
+\EventLogo{}
+\SeriesVolume{237}
+\ArticleNo{5}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{document}
 
-%% Title information
-\title{A Framework for Building Verified Partial Evaluators}         %% [Short Title] is optional;
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-%\titlenote{with title note}             %% \titlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-%\subtitle{Subtitle}                     %% \subtitle is optional
-%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{Jason Gross}
-%\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{0000-0002-9427-4891}             %% \orcid is optional
-\affiliation{
-  %\position{Position1}
-  \department{CSAIL}              %% \department is recommended
-  \institution{MIT}            %% \institution is required
-  %\streetaddress{Street1 Address1}
-  \city{Cambridge}
-  \state{MA}
-  \postcode{02139}
-  \country{USA}                    %% \country is recommended
-}
-\email{jgross@mit.edu}          %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{Andres Erbsen}
-%\authornote{with author2 note}          %% \authornote is optional;
-%% can be repeated if necessary
-%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-    %\position{Position2a}
-    \department{CSAIL}             %% \department is recommended
-    \institution{MIT}           %% \institution is required
-    %\streetaddress{Street2a Address2a}
-    \city{Cambridge}
-    \state{MA}
-    \postcode{02139}
-    \country{USA}                    %% \country is recommended
-}
-\email{andreser@mit.edu}         %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{Adam Chlipala}
-%\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  %\position{Position2a}
-  \department{CSAIL}             %% \department is recommended
-  \institution{MIT}           %% \institution is required
-  %\streetaddress{Street2a Address2a}
-  \city{Cambridge}
-  \state{MA}
-  \postcode{02139}
-  \country{USA}                    %% \country is recommended
-}
-\email{adamc@csail.mit.edu}         %% \email is recommended
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
+\maketitle
+
+%TODO mandatory: add short abstract of the document
 \begin{abstract}
-  \emph{Partial evaluation} is a classic technique for generating lean, customized code from libraries that start with more bells and whistles.
-  It is also an attractive approach to creation of \emph{formally verified} systems, where theorems can be proved about libraries, yielding correctness of all specializations ``for free.''
-  However, it can be challenging to make library specialization both performant (at compile time and runtime) and trustworthy.
-  We present a new approach, prototyped in the Coq proof assistant, which supports specialization at the speed of native-code execution, without adding to the trusted code base.
-  Our extensible engine, which combines the traditional concepts of tailored term reduction and automatic rewriting from hint databases, is also of interest to replace these ingredients in proof assistants' proof checkers and tactic engines, at the same time as it supports extraction to standalone compilers from library parameters to specialized code.
+  Compilers are a prime target for formal verification, since compiler bugs invalidate higher-level correctness guarantees, but compiler changes may become more labor-intensive to implement, if they must come with proof patches.
+  One appealing approach is to present compilers as sets of algebraic rewrite rules, which a generic engine can apply efficiently.
+  Now each rewrite rule can be proved separately, with no need to revisit past proofs for other parts of the compiler.
+  We present the first realization of this idea, in the form of a framework for the Coq proof assistant.
+  Our new Coq command takes normal proved theorems and combines them automatically into fast compilers with proofs.
+  We applied our framework to improve the Fiat Cryptography toolchain for generating cryptographic arithmetic, producing an extracted command-line compiler that is about 1000$\times$ faster while actually featuring simpler compiler-specific proofs.
 \end{abstract}
 
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-%% \begin{CCSXML}
-%% <ccs2012>
-%% <concept>
-%% <concept_id>10011007.10011006.10011008</concept_id>
-%% <concept_desc>Software and its engineering~General programming languages</concept_desc>
-%% <concept_significance>500</concept_significance>
-%% </concept>
-%% <concept>
-%% <concept_id>10003456.10003457.10003521.10003525</concept_id>
-%% <concept_desc>Social and professional topics~History of programming languages</concept_desc>
-%% <concept_significance>300</concept_significance>
-%% </concept>
-%% </ccs2012>
-%% \end{CCSXML}
-
-%% \ccsdesc[500]{Software and its engineering~General programming languages}
-%% \ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-%% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction} \label{sec:intro}
-
-Mechanized proof is gaining in importance for development of critical software infrastructure.
-Oft-cited examples include the CompCert verified C compiler~\cite{Compcert} and the seL4 verified operating-system microkernel~\cite{seL4SOSP09}.
-Here we have very flexible systems that are ready to adapt to varieties of workloads, be they C source programs for CompCert or application binaries for seL4.
-For a verified operating system, such adaptation takes place at \emph{runtime}, when we launch the application.
-However, some important bits of software infrastructure commonly do adaptation at \emph{compile time}, such that the fully general infrastructure software is not even installed in a deployed system.
-
-Of course, compilers are a natural example of that pattern, as we would not expect CompCert itself to be installed on an embedded system whose application code was compiled with it.
-The problem is that writing a compiler is rather labor-intensive, with its crafting of syntax-tree types for source, target, and intermediate languages, its fine-tuning of code for transformation passes that manipulate syntax trees explicitly, and so on.
-An appealing alternative is \emph{partial evaluation}~\cite{PartialEvaluation}, which relies on reusable compiler facilities to specialize library code to parameters, with no need to write that library code in terms of syntax-tree manipulations.
-Cutting-edge tools in this tradition even make it possible to use high-level functional languages to generate performance-competitive low-level code, as in Scala's Lightweight Modular Staging~\cite{LMS}.
-
-It is natural to try to port this approach to construction of systems with mechanized proofs.
-On one hand, the typed functional languages in popular proof assistants' logics make excellent hosts for flexible libraries, which can often be specialized through means as simple as partial application of curried functions.
-Term-reduction systems built into the proof assistants can then generate the lean residual programs.
-On the other hand, it is surprisingly difficult to realize the last sentence with good performance.
-The challenge is that we are not just implementing algorithms; we also want a proof to be checked by a small proof checker, and there is tension in designing such a checker, as fancier reduction strategies grow the trusted code base.
-It would seem like an abandonment of the spirit of proof assistants to bake in a reduction strategy per library, yet effective partial evaluation tends to be rather fine-tuned in this way.
-Performance tuning matters when generated code is thousands of lines long.
-
-In this paper, we present an approach to verified partial evaluation in proof assistants, which requires no changes to proof checkers.
-To make the relevance concrete, we use the example of Fiat Cryptography~\cite{FiatCryptoSP19}, a Coq library that generates code for big-integer modular arithmetic at the heart of elliptic-curve-cryptography algorithms.
-This domain-specific compiler has been adopted, for instance, in the Chrome Web browser, such that about half of all HTTPS connections from browsers are now initiated using code generated (with proof) by Fiat Cryptography.
-However, Fiat Cryptography was only used successfully to build C code for the two most widely used curves (P-256 and Curve25519).
-Their method of partial evaluation timed out trying to compile code for the third most widely used curve (P-384).
-Additionally, to achieve acceptable reduction performance, the library code had to be written manually in continuation-passing style.
-We will demonstrate a new Coq library that corrects both weaknesses, while maintaining the generality afforded by allowing rewrite rules to be mixed with partial evaluation.
-
-\subsection{A Motivating Example} \label{sec:motivating-example} \label{sec:explain-ident.eagerly} \label{sec:explain-eval-rect} \label{sec:explain-'}
-
-We are interested in partial-evaluation examples that mix higher-order functions, inductive datatypes, and arithmetic simplification.
-For instance, consider the following Coq code.
-
-\begin{minted}[fontsize=\small]{coq}
-Definition prefixSums (ls:list nat) : list nat :=
+\section{Introduction}\label{sec:intro}
+% TODO(Jason): Add section/info somewhere about NbE-order being the right order
+
+Formally verified compilers like CompCert~\cite{Compcert} and CakeML~\cite{CakeML} are success stories for proof assistants, helping close a trust gap for one of the most important categories of software infrastructure.
+A popular compiler cannot afford to stay still; developers will add new backends, new language features, and better optimizations.
+Proofs must be adjusted as these improvements arrive.
+It makes sense that the author of a new piece of compiler code must prove its correctness, but ideally there would be no need to revisit old proofs.
+There has been limited work, though, on avoiding that kind of coupling.
+\citet{xcert} demonstrated a streamlined way to extend CompCert with new verified optimizations driven by dataflow analysis, but we are not aware of past work that supports easy extension for compilers from functional languages to C code.
+We present our work targeting that style.
+
+One strategy for writing compilers modularly is to exercise foresight in designing a core that will change very rarely, such that feature iteration happens outside the core.
+Specifically, phrasing the compiler in terms of rewrite rules allows clean abstractions and conceptual boundaries~\cite{Hickey2006}.
+Then, most desired iteration on the compiler can be achieved through iteration on the rewrite rules.
+
+It is surprisingly difficult to realize this modular approach with good performance.
+Verified compilers can either be proof-producing (certifying) or proven-correct (certified).
+Proof-producing compilers usually operate on the functional languages of the proof assistants that they are written in, and variable assignments are encoded as let binders.
+All existing proof-producing rewriting strategies scale at least quadratically in the number of binders.
+This performance scaling is inadequate for applications like Fiat Cryptography~\cite{FiatCryptoSP19} where the generated code has 1000s of variables in a single function.
+Proven-correct compilers do not suffer from this asymptotic blowup in the number of binders.
+
+In this paper, we present \textbf{the first proven-correct compiler-builder toolkit parameterized on rewrite rules}.
+Arbitrary sets of Coq theorems (quantified equalities) can be assembled by a single new Coq command into an extraction-ready verified compiler.
+We did not need to extend the trusted code base, so our compiler compiler need not be trusted.
+We achieve both good performance of compiler runs and good performance of generated code, via addressing a number of scale-up challenges vs.\ past work.
+%including efficient handling of intermediate code with thousands of nested binders and critical subterm sharing.
+
+We evaluate our toolkit by replacing a key component of Fiat Cryptography~\cite{FiatCryptoSP19}, a Coq library that generates code for big-integer modular arithmetic at the heart of elliptic-curve-cryptography algorithms.
+Routines generated (with proof) with Fiat Cryptography now ship with all major Web browsers and all major mobile operating systems.
+With our improved compiler architecture, it became easy to add two new backends and a variety of new supported source-code features, and we were easily able to try out new optimizations.
+%and later replace unsound optimizations with equivalent sound optimizations.
+
+Replacing Fiat Cryptography's original compiler with the compiler generated by our toolkit has two additional benefits.
+Fiat Cryptography was previously only used successfully to build C code for two of the three most widely used curves (P-256 and Curve25519).
+Our prior version's execution timed out trying to compile code for the third most widely used curve (P-384).
+Using our new toolkit has made it possible to generate compiler-synthesized code for P-384 while generating completely identical code for the primes handled by the previous version, about 1000$\times$ more quickly.
+Additionally, Fiat Cryptography previously required source code to be written in continuation-passing style, and our compiler has enabled a direct-style approach, which pays off in simplifying theorem statements and proofs.
+
+%There are a number of subtleties that arise in building a verified compiler around rewrite rules.
+%There are two important performance criteria: performance of the compiler in generating code, and the performance of the generated code itself.
+
+%The typed functional languages in popular proof assistants' logics make excellent hosts for flexible libraries, which can often be specialized through means as simple as partial application of curried functions.
+%Term-reduction systems built into the proof assistants can then generate the lean residual programs.
+%It is surprisingly difficult to realize the last sentence with good performance.
+%The challenge is that we are not just implementing algorithms; we also want a proof to be checked by a small proof checker, and there is tension in designing such a checker, as fancier reduction strategies grow the trusted code base.
+%It would seem like an abandonment of the spirit of proof assistants to bake in a reduction strategy per library, yet effective partial evaluation tends to be rather fine-tuned in this way.
+%Performance tuning matters when generated code is thousands of lines long.
+
+% While there are nascent efforts in proof repair to mitigate the cost of iteration in a verified setting~\cite{Ringer2021}, we
+
+%Mechanized proof is gaining in importance for development of critical software infrastructure.
+%Oft-cited examples include the CompCert verified C compiler~\cite{Compcert} and the seL4 verified operating-system microkernel~\cite{seL4SOSP09}.
+%Here we have very flexible systems that are ready to adapt to varieties of workloads, be they C source programs for CompCert or application binaries for seL4.
+%For a verified operating system, such adaptation takes place at \emph{runtime}, when we launch the application.
+%However, some important bits of software infrastructure commonly do adaptation at \emph{compile time}, such that the fully general infrastructure software is not even installed in a deployed system.
+
+%Of course, compilers are a natural example of that pattern, as we would not expect CompCert itself to be installed on an embedded system whose application code was compiled with it.
+%The problem is that writing a compiler is rather labor-intensive, with its crafting of syntax-tree types for source, target, and intermediate languages, its fine-tuning of code for transformation passes that manipulate syntax trees explicitly, and so on.
+%An appealing alternative is \emph{partial evaluation}~\cite{PartialEvaluation}, which relies on reusable compiler facilities to specialize library code to parameters, with no need to write that library code in terms of syntax-tree manipulations.
+%Cutting-edge tools in this tradition even make it possible to use high-level functional languages to generate performance-competitive low-level code, as in Scala's Lightweight Modular Staging~\cite{LMS}.
+
+%It is natural to try to port this approach to construction of systems with mechanized proofs.
+%On one hand, the typed functional languages in popular proof assistants' logics make excellent hosts for flexible libraries, which can often be specialized through means as simple as partial application of curried functions.
+%Term-reduction systems built into the proof assistants can then generate the lean residual programs.
+%On the other hand, it is surprisingly difficult to realize the last sentence with good performance.
+%The challenge is that we are not just implementing algorithms; we also want a proof to be checked by a small proof checker, and there is tension in designing such a checker, as fancier reduction strategies grow the trusted code base.
+%It would seem like an abandonment of the spirit of proof assistants to bake in a reduction strategy per library, yet effective partial evaluation tends to be rather fine-tuned in this way.
+%Performance tuning matters when generated code is thousands of lines long.
+
+\subsection{Related Work}\label{sec:related-one}
+% POPL 2022 review TODO:
+% Section 1.1, it may be a little early at this point to discuss the related work, as I don't think it is quite clear yet exactly what problem this paper solves.
+% From the first 1.5 pages, I understand that the Fiat Cryptography compiler has been improved, but it is not clear yet how this is achieved and what are the merits or demerits of the approach presented in this paper.
+
+Assume our mission is to take libraries of purely functional combinators, apply them to compile-time parameters, and compile the results down to lean C code.
+Furthermore, we ask for machine-checked proofs that the C programs preserve the behavior of the higher-order functional programs we started with.
+What good ideas from the literature can we build on?
+
+\citet{Hickey2006} discuss at length how to build compilers around rewrite rules.
+``All program transformations, from parsing to code generation, are cleanly isolated and specified as term rewrites.''
+While they note that the correctness of the compiler is thus reduced to the correctness of the rewrite rules, they did not prove correctness mechanically.
+Furthermore, it is not clear that they manage to avoid the asymptotic blow-up associated with proof-producing rewriting of deeply nested let-binders.
+They give no performance numbers, so it is hard to say whether or not their compiler performs at the scale necessary for Fiat Cryptography.
+Their rewrite-engine driver is unproven OCaml code, while we will produce custom drivers with Coq proofs.
+%This suggests also that they cannot use rewrite rules which are only sound in the presence of side conditions, as such side-condition checking could not happen in the engine driving the rewrites.
+%Overall, their work complements ours; we focus on how to build a scalable engine for building a proven-correct compiler around shallowly embedded rewrite rules in a proof assistant like Coq; they focus on what rewrite rules are necessary for implementing a general-purpose compiler.
+
+\Rtac{}~\cite{rtac} is a more general framework for verified proof tactics in Coq, including an experimental reflective version of \texttt{rewrite\_strat} supporting arbitrary setoid relations, unification variables, and arbitrary semidecidable side conditions solvable by other verified tactics, using de Bruijn indexing to manage binders.
+We found that \Rtac{} misses a critical feature for compiling large programs: preserving subterm sharing.
+As a result, our experiments with compiler construction yielded clear asymptotic slowdown vs.\ what we eventually accomplished.
+\Rtac{} is also more heavyweight to use, for instance requiring that theorems be restated manually in a deep embedding to bring them into automation procedures.
+Furthermore, we are not aware of any past experiments driving verified compilers with \Rtac{}.
+%, and we expect there would be other bottlenecks working with large, deeply nested terms.
+
+\citet{Aehlig} came closest to a fitting approach, using \emph{normalization by evaluation (NbE)}~\cite{NbE} to bootstrap reduction of open terms on top of full reduction, as built into a proof assistant.
+% POPL 2022 review TODO: Line 99: I'd drop the "it was simultaneously true that".
+However, it was simultaneously true that they expanded the proof-assistant trusted code base in ways specific to their technique, and that they did not report any experiments actually using the tool for partial evaluation (just traditional full reduction), potentially hiding performance-scaling challenges or other practical issues.
+For instance, they also do not preserve subterm sharing explicitly, and they represent variable references as unary natural numbers (de Bruijn-style).
+They also require that rewrite rules be embodied in ML code, rather than stated as natural ``native'' lemmas of the proof assistant.
+We will follow their basic outline with important modifications.
+
+Our implementation builds on fast full reduction in Coq's kernel, via a virtual machine~\cite{vmcompute} or compilation to native code~\cite{nativecompute} (neither verified).
+Especially the latter is similar in adopting NbE for full reduction, simplifying even under $\lambda$s, on top of a more traditional implementation of OCaml that never executes preemptively under $\lambda$s.
+Neither approach unifies support for rewriting with proved rules, and partial evaluation only applies in very limited cases.
+
+A variety of forms of pragmatic partial evaluation have been demonstrated, with Lightweight Modular Staging~\cite{LMS} in Scala as one of the best-known current examples.
+The LMS-Verify system~\cite{LMSVerify} can be used for formal verification of generated code after-the-fact.
+Typically LMS-Verify has been used with relatively shallow properties (though potentially applied to larger and more sophisticated code bases than we tackle), not scaling to the kinds of functional-correctness properties that concern us here.
+%justifying investment in verified partial evaluators.
+
+So, overall, to our knowledge, no past compiler as a set of rewrite rules has come with a full proof of correctness as a standalone functional program.
+Related prior work with mechanized proofs suffered from both performance bottlenecks and usability problems, the latter in requiring that eligible rewrite rules be stated in special deep embeddings.
+%We will demonstrate all of these pieces working together well enough to be adopted by several high-profile open-source projects.
+
+\subsection{Our Solution}\label{sec:our-solution}
+
+% POPL 2022 review TODO:
+% line 114, "Our variant on the technique of Aehlig et al. [...]".
+% The authors have not yet recalled what is Aehlig et al.'s technique, nor explained how this work differs from it, so it is difficult to appreciate the list of claimed advantages at this point.
+Our variant on the technique of \citet{Aehlig} has these advantages:
+
+\begin{itemize}
+\item It integrates with a general-purpose, foundational proof assistant, \textbf{without growing the trusted code base}.
+\item For a wide variety of initial functional programs, it provides \textbf{fast} partial evaluation with reasonable memory use.
+\item It allows reduction that \textbf{mixes} \emph{rules of the definitional equality} with \emph{equalities proven explicitly as theorems}.
+\item It allows \textbf{rapid iteration} on rewrite rules with \emph{minimal verification overhead}.
+\item It \textbf{preserves sharing} of common subterms.
+\item It also allows \textbf{extraction of standalone compilers}.
+\end{itemize}
+
+Our contributions include answers to a number of challenges that arise in scaling NbE-based partial evaluation in a proof assistant.
+First, we rework the approach of \citet{Aehlig} to function \emph{without extending a proof assistant's trusted code base}, which, among other challenges, requires us to prove termination of reduction and encode pattern matching explicitly (leading us to adopt the performance-tuned approach of \citet{maranget2008compiling}).
+We also improve on Coq-specific related work (e.g., of \citet{rtac}) by allowing rewrites to be written in natural Coq form (not special embedded syntax-tree types), while supporting optimizations associated with past unverified engines (e.g., \citet{Boespflug2009}).
+
+Second, using partial evaluation to generate residual terms thousands of lines long raises \emph{new scaling challenges}:
+\begin{itemize}
+\item
+  Output terms may contain so \emph{many nested variable binders} that we expect it to be performance-prohibitive to perform bookkeeping operations on first-order-encoded terms (e.g., with de Bruijn indices, as is done in \Rtac{} by \citet{rtac}).
+  For instance, while the reported performance experiments of \citet{Aehlig} generate only closed terms with no binders, Fiat Cryptography may generate a single routine (e.g., multiplication for curve P-384) with nearly a thousand nested binders.
+%\todo{Consider being more specific here about expected cost of first-order representation, maybe using some experiment.}
+%  \todo{Note from Jason: We haven't actually checked that the bookkeeping is performance-prohibitive.
+I suspect there will be a cost, but that if when we made fiat-crypto there was an existing system that used de Bruijn indices and did this, (along with built-in let-lifting, etc), I suspect we might have been a bit more sad, but not prohibitively more sad.
+I think we should be much more tentative with this bullet point.}
+\item
+  Naive representation of terms without proper \emph{sharing of common subterms} can lead to fatal term-size blow-up.
+  %Fiat Cryptography's arithmetic routines rely on significant sharing of this kind.
+\item
+  Unconditional rewrite rules are in general insufficient, and we need \emph{rules with side conditions}.
+  E.g., Fiat Cryptography depends on checking lack-of-overflow conditions.
+\item
+  However, it is also not reasonable to expect a general engine to discharge all side conditions on the spot.
+  We need integration with \emph{abstract interpretation}.
+  %that can analyze whole programs to support reduction.
+\end{itemize}
+
+Briefly, our respective solutions to these problems are the \emph{parametric higher-order abstract syntax (PHOAS)}~\cite{PhoasICFP08} term encoding, a \emph{let-lifting} transformation threaded throughout reduction, extension of rewrite rules with executable Boolean side conditions, and a design pattern that uses decorator function calls to include analysis results in a program.
+
+Finally, we carry out the \emph{first large-scale performance-scaling evaluation} of a verified rewrite-rule-based compiler, covering all elliptic curves from the published Fiat Cryptography experiments, along with microbenchmarks.
+
+We pause to give a motivating example before presenting the core structure of our engine (\autoref{sec:structure}), the additional scaling challenges we faced (\autoref{sec:scaling}), experiments (\autoref{sec:evaluation}), and conclusions.
+Our implementation is attached.
+
+\section{A Motivating Example}\label{sec:motivating-example}\label{sec:explain-ident.eagerly}\label{sec:explain-eval-rect}
+
+Our compilation style involves source programs that mix higher-order functions and inductive types.
+We want to compile to C code, reducing away uses of fancier features while seizing opportunities for arithmetic simplification.
+Here is a small but illustrative example.
+
+\begin{minted}[fontsize=\small]{coq}
+Definition prefixSums (ls : list nat) : list nat :=
   let ls' := combine ls (seq 0 (length ls)) in

### 2023-03-31 resume/Resume.tex
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@4a20b836590007b68b441c3fc567c012fde26482 🚀
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@de630c64253fb9a226076168f6ef29b037bff769 🚀

@@
 \DeclareFile{curriculum-vitae}
-\DeclareFile{mathematics}
-\DeclareFile{computer-science}
-\DeclareFile{curriculum-vitae-long}
-\DeclareFile{curriculum-vitae-mathematics}
-\DeclareFile{curriculum-vitae-computer-science}
+%\DeclareFile{mathematics}
+%\DeclareFile{computer-science}
+%\DeclareFile{curriculum-vitae-long}
+%\DeclareFile{curriculum-vitae-mathematics}
+%\DeclareFile{curriculum-vitae-computer-science}
 \DeclareDocumentClass{\documentclass[11pt]{res}}
@@
 \end{multitex}
-\begin{multitex*}[\cvlongfiles]
-  \name{\LARGE \bf \hspace{\stretch{1}} Jason Gross \hspace{\stretch{1}} \\
-    \hspace{\stretch{1}}\large \href{mailto:jgross@mit.edu}{jgross@mit.edu} --- (631) 790-8962 --- \url{https://people.csail.mit.edu/jgross/}\hspace{\stretch{1}} \\}
 
-  \address{\textsc{Present Address}\\258 Prospect Street, Apt \# 1L\\Cambridge, MA 02139}
-  \address{\textsc{Permanent Address}\\126 Hayrick Lane \\  Commack, NY 11725}
-\end{multitex*}
-\begin{multitex}[\allexceptcvlongfiles]
+\begin{multitex}
   \name{\LARGE \bf \parbox{\widthof{github.com/JasonGross}}{%
@@
   \resume
-\end{multitex}
 
-\begin{multitex*}[\allexceptcvandcvlongfiles]
   \section{\textsc{Research Interests}}
-  Programming Languages, Formal Verification, Cryptography, Performance of Automation in Interactive Proof Assistants, Homotopy Type Theory, Category Theory, Program Synthesis
-\end{multitex*}
-\begin{multitex}[curriculum-vitae]
-  \section{\textsc{Research Interests}}
-  Programming Languages, Formal Verification, Cryptography, Performance Engineering
-\end{multitex}
+  \begin{itemize}
+    \item Programming Languages and AI: Löb's Theorem, Type Theory, Compact Proof Generation
+    \item Trust and Security Engineering: Formal Verification, Cryptography, Performance Engineering
+  \end{itemize}
 
-\begin{multitex}
-  %\section{JOB OBJECTIVE}
-    %A summer position that will use my accounting and computer
-    %skills.
-    %
+  \section{\textsc{Experience}}
+
+  \employer{Machine Intelligence Research Institute (MIRI)}
+  \dates{February 2021--Present}
+  \title{Research Staff}
+  \location{Berkeley, CA (remote)}
+  \begin{position}
+    \begin{itemize}
+      \item Performing self-directed research into topics in fundamental programming language theory and mathematics%, bringing insights and understanding back to my supervisor
+    \end{itemize}
+  \end{position}
+
+  \employer{Coq Development Team, INRIA}
+  \dates{June 2021--Present}
+  \title{Core Team Member}
+  \location{Nantes, France (remote)}
+  \begin{position}
+    \begin{itemize}
+      \item Reported the plurality of all-time bugs in Coq (since 2012)
+      \item Designed and engineered a bug minimizer for automatically producing minimized stand-alone test-cases from bug reports with automatic support for minimizing regressions in external projects tested by Coq CI
+      \item Researching performance issues that impact scalability of automated verification
+    \end{itemize}
+  \end{position}
+
+  \employer{MIT CSAIL}
+  \dates{September 2013--February 2021}
+  \title{PhD Researcher}
+  \location{Cambridge, MA}
+  \begin{position}
+    \begin{itemize}
+      \item Main Project: Fiat Cryptography (\href{https://github.com/mit-plv/fiat-crypto}{github.com/mit-plv/fiat-crypto})
+      \item Fiat Cryptography is a developer tool to generate proven-correct cryptographic code.
+The project has wide industry adoption, including powering secure connections in Google Chrome and Firefox
+      \item Lead development of one of the world's first algorithm-level-optimizing compilers
+      \item Collaboratively implemented the tool; wrote backends to C, Go, Java, and JSON; managed development of backends to Rust and Zig
+    \end{itemize}
+  \end{position}
+
   \section{\textsc{Education}}
-\end{multitex}
-
-\begin{multitex}[\allexceptcvlongfiles]
   \employer{Massachusetts Institute of Technology}
@@
   \end{position}
-\end{multitex}
 
-\begin{multitex*}[\cvlongfiles]
-    \textbf{Massachusetts Institute of Technology}, Cambridge, MA  \\
-    Began Doctorate of Philosophy in Computer Science in September 2013 \\
-    \onlymultitex*{\cvfiles}{Research Interests: Programming Languages, Performance of Automation in Interactive Proof Assistants, Formal Verification, Homotopy Type Theory, Category Theory, Program Synthesis, Type Theory \\}
-    G.P.A. 4.9/5.0 \\
-    \\
-    Bachelor of Science in Mathematics and Physics, June 2013   \\
-    G.P.A. 4.6/5.0 \\
-\end{multitex*}
-\begin{multitex*}[\resumefiles]%
-    \\
-    Relevant Coursework: %Computer Systems
-        Security, %Foundations of
-        Program Analysis, Performance Engineering of Software% Systems
-        , Statistical Physics, Quantum% Physics
-        , Topology, Analysis, Waves \& Vibrations, Special Relativity% and the Physics of Spacetime%
-        , \mbox{Algebra}% force no hyphenation
-\end{multitex*}
-\begin{multitex*}[\cvlongfiles]%
-    \\
-		Coursework:
+\clearpage
+
+  \section{\textsc{Internships}}
+
   \begin{itemize}
-			\item Computer Science: Inference and Information, Computer Systems Security, Geometric Computing, Foundations of Program Analysis, Performance Engineering of Software Systems, Structure and Interpretation of Computer Programs
-			\item Mathematics: Paradox and Infinity, Category Theory for Scientists, Algebraic Topology I, Seminar in Topology, Introduction to Topology, Real Analysis, Abstract Algebra I \& II, Differential Equations, Calculus I \& II, Linear Algebra
-			\item Physics: Modern Astrophysics, General Relativity (Graduate), Statistical Physics I \& II, Quantum Physics I, II, \& III, Classical Mechanics II, Waves and Vibrations, Special Relativity and the Physics of Spacetime
-			\item Other Sciences: Introductory Biology, Introductory Chemistry, Design a Concentrated Solar Power Water Heater (seminar)
-			\item Other: The Art and Science of Happiness, Metaphysics, Introduction to Musical Composition, Capitalism and Its Critics, Moral Psychology, Philosophy of Quantum Mechanics, Moral Problems and the Good Life, Philosophy of Love, Kaufman Teaching Certificate Program
+  \item MIRI, summer 2019: Formalized type theories, and proved properties of programs that reason about themselves
+  \item Google, summer 2018: Worked on integration of Fiat Cryptography with BoringSSL in Chrome
+  \item Google, summer 2016: Extended Fiat Cryptography with ECC primiatives for integration with Open Titan
+  \item Microsoft Research, summer 2014: Collaboratively created a language for specifying input/output behavior of x86 assembly programs, verfied the input/output behavior of a number of simple programs, and improved performance of the x86proved project
+  \item MIT CSAIL - PLV, 2012 - 2014: Entered a significant amount of category theory into the automated proof assistant Coq, and worked on building an interface for databases and database migration on top of category theory
+  \item MIT CSAIL - CoCoSci, 2009 - 2011: Designed and managed the data collection webpage for research in categorical learning and transfer learning
+  \item Commack High School, 2006 - 2009: Researched circuits over sets of natural numbers, winning 4th (2009) and 3rd (2008) place awards in mathematics at ISEF
   \end{itemize}
-\end{multitex*}
 
-% \item Presented ``Computational Higher Inductive Types: Computing with Custom Equalities'', October 6, 2013, The CSAIL Student Workshop
+  \section{\textsc{Professional Activities}}
 
-\begin{multitex*}[\cvlongfiles]
-\section{\textsc{Experience}}
-\end{multitex*}
-\begin{multitex*}[\cvlongfiles]
-  \employer{MIT CSAIL}
-  \dates{September 2013--February 2021}
-  \title{Researcher}
-  \location{Cambridge, MA}
-  \begin{position}
   \begin{itemize}
-      \item Researching program synthesis and verification with Adam Chlipala
-      \item Collaboratively working on \onlymultitex*{\cvfiles}{implementing }one of the world's first algorithm-level-optimizing compilers
-      \item Collaboratively implemented proven-correct crypto code now used by Google Chrome
+  \item Co-maintainer of the Fiat Cryptography project 
+  \item Co-maintainer of the homotopy type theory Coq repository (\href{https://github.com/HoTT/HoTT}{\texttt{HoTT/HoTT} on GitHub})
+  \item Program Committee Member of ITP 2023 and CoqPL 2022
+  \item Supervising research in formalizing correspondence of affine logic to two-player games
+  \item Supervising research in anti-inductive utility functions
+  \item Supervising research in performative power of predidiction markets 
+  \item Circling Facilitator at The Relateful Company
+  \item Member of SIPB (Student Information and Processing Board)
   \end{itemize}
-  \end{position}
-\end{multitex*}
 
-\begin{multitex}
-\section{\textsc{Experience}}
-\end{multitex}
+  \section{\textsc{Select Past Activities}}
 
-\begin{multitex*}[\cvfiles]
-  \employer{Coq Development Team, INRIA}
-  \dates{June 2021--Present}
-  \title{Core Team Member}
-  \location{Nantes, France (remote)}
-  \begin{position}
   \begin{itemize}
-      \item Stress Testing, Bug Reporting, Bug Minimizer, Compatibility Assurance
-      \item Engineering and maintaining a bug minimizer for automatically producing minimized stand-alone test-cases from buggy code
-      \item Researching performance issues that impact scalability of automated verification
-      \item Have reported the plurality of all-time bugs in Coq (since 2012)
+  \item Particpant in MIRI Decision Theory Workshops
+  \item Volunteer at CFAR workshops
+  \item Instructor at MIT ESP Programs
+  \item Instructor at Monsoon Math Camp
+  \item President of MIT Tech Squares
+  \item Contributor to the SIPB BarnOwn project
+  \item Project leader for MITeX, an online interface for composing \LaTeX
+  \item TA for 6.172: Performance Engineering
+  \item TA for 8.012: Physics I and 8.022: Physics II at the Experimental Study Group
+  \item Participant at Cananda/USA Mathcamp
   \end{itemize}
-  \end{position}
 
-  \employer{Machine Intelligence Research Institute}
-  \dates{February 2021--Present}
-  \title{Research Staff}
-  \location{Berkeley, CA (remote)}
-  \begin{position}
-    \begin{itemize}
-      \item Performing self-directed research into topics in fundamental programming language theory and mathematics%, bringing insights and understanding back to my supervisor
-    \end{itemize}
-  \end{position}
-
-  \employer{MIT CSAIL}
-  \dates{September 2013--February 2021}
-  \title{PhD Researcher}
-  \location{Cambridge, MA}
-  \begin{position}
-    \begin{itemize}
-      \item Main Research Project: Fiat Cryptography (\href{https://github.com/mit-plv/fiat-crypto}{github.com/mit-plv/fiat-crypto})
-      \item Collaboratively implemented proven-correct cryptographic code now used by Google Chrome, and in the majority of secure connects from web browsers
-      \item Led development of one of the world's first algorithm-level-optimizing compilers
-      \item Wrote backends to C, Go, Java, and JSON; managed development of backends to Rust and Zig
-    \end{itemize}
-  \end{position}
-\end{multitex*}
-
-\begin{multitex}
-\section{\textsc{Internships}}
-\end{multitex}
-
-\begin{multitex*}[\cvfiles]
-  \employer{Machine Intelligence Research Institute}
-  \dates{June 2019--August 2019}
-  \title{Type Theory Intern}
-  \location{Berkeley, CA}
-  \begin{position}
-    \begin{itemize}
-      \item Worked on formalizing type theories and on proving things within proof assistants
-    \end{itemize}
-  \end{position}
-\end{multitex*}
-\begin{multitex}
-  \bgroup
-  \interlinepenalty 10000
-  \employer{Google}
-  \dates{June 2018--August 2018}
-  \title{Software Engineering Intern}
-  \location{Cambridge, MA}
-  \begin{position}
-    \begin{itemize}
-      \item Worked with BoringSSL on integration of proven-correct low-level ECC primitives into Chrome
-    \end{itemize}
-  \end{position}
-  \egroup
-\end{multitex}
-\begin{multitex*}[\cvfiles]
-  \bgroup
-  \interlinepenalty 10000
-  \employer{Google}
-  \dates{June 2016--September 2016}
-  \title{Software Engineering Intern}
-  \location{Mountain View, CA}
-  \begin{position}
-    \begin{itemize}
-      \item Formalized low-level ECC primitives with proofs of correctness
-    \end{itemize}
-  \end{position}
-  \egroup
-\end{multitex*}
-\begin{multitex}[\cvlongandmathfiles]
-  \employer{MIT}
-  \dates{Fall 2009--Present}
-  \title{Teacher}
-  \location{Cambridge, MA}
-  \begin{position}
-    \begin{itemize}
-      \item Taught classes on \LaTeX, philosophy, linear algebra, and quantum mechanics for MIT Educational Studies Program's Splash, Spark, and Summer HSSP (High School Studies Program)
-      \onlymultitex{\cvandmathfiles}{\item Teaching Assistant for 8.012 (Physics I) and 8.022 (Physics II) in Experimental Study Group}
-    \end{itemize}
-  \end{position}
-\end{multitex}
-\begin{multitex*}[\cvlongandcomputersciencefiles]
-  \employer{MIT}
-  \dates{September 2015--December 2015}
-  \title{TA for 6.172 (Performance Engineering% of Software% Systems
-  )}
-  \location{Cambridge, MA}
-  \begin{position}
-    \begin{itemize}
-      \item Created and led recitations, taught students, and helped run class
-      \item Analyzed and explained assembly output of \texttt{gcc -O3} to teach vectorization
-    \end{itemize}
-  \end{position}
-\end{multitex*}
-\begin{multitex*}[\cvlongfiles]
-    \employer{MIRI}
-    \dates{June 12--14, 2015}
-    \title{Decision Theory Workshop Attendee}
-    \location{Berkeley, CA}
-    \begin{position}
-        \begin{itemize}
-            \item Formalized various versions of L\"ob's theorem in Agda and Coq
-            \item Learned about and worked on problems on the frontiers of decision theory
-        \end{itemize}
-    \end{position}
-\end{multitex*}
-\begin{multitex}
-  \employer{Microsoft Research}
-  \dates{June 2014--August 2014}
-  \title{Intern}
-  \location{Cambridge, United Kingdom}
-  \begin{position}
-    \begin{itemize}
-      \item \onlymultitex*{\allexceptcvlongfiles}{Collaboratively c}\onlymultitex*{\cvlongfiles}{C}reated a language for specifying input/output behavior of x86 assembly programs\onlymultitex*{\cvlongfiles}{with Andrew Kennedy and Nick Benton}; Verified the I/O behavior of a number of simple programs
-      \onlymultitex*{\cvfiles}{\item Improved automation of the x86proved library}
-    \end{itemize}
-  \end{position}
-\end{multitex}
-\begin{multitex*}[\cvfiles]
-    \employer{MIT CSAIL}
-    \dates{April 2012--June 2014}
-    \title{Researcher}
-    \location{Cambridge, MA}
-    \begin{position}
-        \begin{itemize}
-            \item Entered a significant amount of category theory into the automated proof assistant Coq \onlymultitex*{\cvfiles}{(\url{https://github.com/HoTT/HoTT/tree/master/theories/Categories})}
-            \onlymultitex{\cvfiles}{\item Made progress towards an interface for databases and database migration on top of category theory in Coq\onlymultitex*{\cvlongfiles}{ with David Spivak and Adam Chlipala}}
-            \onlymultitex*{\cvlongfiles}{\item Presented ``Building Database Management on top of Category Theory in Coq'', January 25, 2013, POPL 2013: 40th ACM SIGPLAN-SIGACT Symposium on Principles of Programming Languages}
-            \onlymultitex*{\cvlongfiles}{\item Presented	 ``Experience Implementing a Performant Category-Theory Library in Coq'', July 14, 2014, Interactive Theorem Proving 2014 Conference}
-        \end{itemize}
-    \end{position}
-\end{multitex*}
-\begin{multitex*}[\cvfiles]
-  \employer{MIT CSAIL}
-  \dates{November 2009--September 2011}
-  \title{Researcher}
-  \location{Cambridge, MA}
-  \begin{position}
-    \begin{itemize}
-      \onlymultitex*{\cvfiles}{\item Designed from scratch a data collection webpage, collected data for, and helped with research \onlymultitex*{\cvlongfiles}{ of Brenden Lake, Ruslan Salakhutdinov, and Josh Tenenbaum, }on categorical and transfer learning\onlymultitex*{\cvlongfiles}{ (\url{https://jgross.scripts.mit.edu/alphabets/})}}
-      \onlymultitex*{\cvlongfiles,\resumefiles}{%
-      \item Co-author of ``One shot learning of simple visual concepts''\onlymultitex*{\cvfiles}{ published in \emph{Proceedings of the 33rd Annual Meeting of the Cognitive Science Society}}.
-        }
-    \end{itemize}
-  \end{position}
-\end{multitex*}
-\begin{multitex*}[\cvlongfiles]
-  \employer{MIT OCW}
-    \dates{May--August 2010}
-    \title{}
-    \location{Cambridge, MA}
-    \begin{position}
-        \begin{itemize}
-            \item Evaluated and categorized videos of Walter Lewin solving physics problems for updating the 8.01 (Physics I) OpenCourseWare (OCW) website.
-        \end{itemize}
-    \end{position}
-\end{multitex*}
-\begin{multitex*}[\cvfiles]
-  \employer{Commack High School}
-  \dates{Fall 2006--Summer 2009}
-  \title{Independent Researcher}
-  \location{Commack, NY}
-  \begin{position}
-    \begin{itemize}
-      \item Independently researched circuits over sets of natural numbers for three years.
-      \item Won fourth place award in mathematics in ISEF (Intel International Science and Engineering Fair) in 2009, third place award in ISEF 2008.
-    \end{itemize}
-  \end{position}
-\end{multitex*}
-
-\begin{multitex*}[\cvlongfiles]
-  \employer{Turnpike Total Appliance}
-  \dates{Fall 2006--Summer 2009}
-  \title{Web Page Designer}
-  \location{Commack, NY}
-  \begin{position}
-    \begin{itemize}
-      \item Designed and improved main company website, took initiative to make more improvements.
-    \end{itemize}
-  \end{position}
-
-\end{multitex*}
-
-\begin{multitex}[curriculum-vitae]
   \section{\textsc{Programming Languages}}
-\end{multitex}
-\begin{multitex*}[\allexceptcv]
-\section{\textsc{Computer Skills}}
-\end{multitex*}
-\begin{multitex}
   \begin{itemize}
   \item
-        Proficient: Coq, \onlymultitex*{\cvlongfiles}{\TeX{} macro language, }Mathematica, git, Python, JavaScript, BASIC
+    Proficient: Coq, \onlymultitex{\cvlongfiles}{\TeX{} macro language, }Mathematica, git, Python, JavaScript, BASIC
   \item
-        Working knowledge: \onlymultitex*{\cvlongfiles}{\LaTeX{}, }C, C++, Agda, OCaml, Haskell, Scheme, HTML, CSS, Perl, Java
+    Working knowledge: \onlymultitex{\cvlongfiles}{\LaTeX{}, }C, C++, Agda, OCaml, Haskell, Scheme, HTML, CSS, Perl, Java
   \item
@@
   \end{itemize}
-\end{multitex}
 
-\begin{multitex}
-  \section{\textsc{Teaching}}
-  \begin{itemize}
-  \item Instructor at Monsoon Math Camp: category theory, linear logic, L\"ob's theorem
-  \item TA for 6.172 (Performance Engineering% of Software% Systems
-    ): Led recitations, analyzed and explained assembly output of \texttt{gcc -O3} to teach vectorization
-  \item TA for 8.012 (Physics I) and 8.022 (Physics II) in Experimental Study Group
-  \item Teacher at MIT ESP Programs: \LaTeX, philosophy, linear algebra, quantum mechanics
-  \end{itemize}
-\end{multitex}
+  \nocite{*}
 
-\begin{multitex*}[\cvlongfiles]
-\section{\textsc{Honors and Awards}}
-\end{multitex*}
-\begin{multitex*}[\cvlongfiles]
-    \begin{itemize}
-      \onlymultitex*{\highschoolfiles,\cvfiles}{%
-      \item
-        Mathematics Honor Society (Commack High School)
-      }
-      \item
-        Collection of 12 original K'NEX synagogue models exhibited in various museums (2004--2008)
-      \onlymultitex*{\highschoolfiles}{%
-        \begin{itemize}
-          \item
-            Collection of 12 original K'NEX synagogue models
-          \item
-            Volunteered to give speeches and conduct K'NEX workshops for children at opening of exhibits, as well as in many locations in NY area
-          \item
-            April 2008 --- Gave interview for April edition of BabagaNewz Magazine and audio interview for BabagaNewz.com
-          \item
-            October 2006 --- Displayed in Gurwin Nursing Home on Long Island
-          \item
-            February 2006 --- Exhibited at Lisa Ann Watson Children's Discovery Museum in Miami, Florida
-          \item
-            October 2005 --- Displayed at Shomrim (Society for New York City Jewish Police Officers)
-          \item
-            February--April 2005 --- Exhibited in Yeshiva University Museum in Manhattan
-          \item
-            December 2004 --- Displayed at Suffolk Y-JCC in Commack
-        \end{itemize}
-        }
-      \onlymultitex*{\cvfiles}{
-      \item
-        Graduated Cum Laude from Commack High School (June 2009)
-      }
-      \onlymultitex*{\highschoolfiles}{%
-      \item
-        Commack High School Outstanding Academic Performance (June 2009)
-      \item
-        Florence Burns Scholarship for Achievement of Excellence in Computer Science (June 2009)
-      \item
-        Theodore R. Spedalle Award for Achievement of Excellence in Mathematics (June 2009)
-      \item
-        Commack High School Mathematics Research Award (June 2009)
-      \item
-        President's Award for Educational Excellence (June 2009)
-      \item
-        AP Scholar with Distinction (2009)
-      \item
-        Robert C. Byrd Honors Scholarship (June 2009)
-      \item
-        ISEF (Intel International Science \& Engineering Fair)
-        \begin{itemize}
-          \item
-            Math research paper, ``Circuits over Sets of Natural Numbers'' won $3^\text{rd}$ Place in Math in Atlanta, May 2008
-          \item
-            Math research paper, ``An Investigation of the Closure of the Set of Singleton Sets of Natural Numbers under $\cup$, $\cap$, $\overline{\phantom{\emptyset}}$, $+$, $\times$'' won $4^\text{th}$ place in Math in Reno, May 2009
-       \end{itemize}
-      \item
-        LISEF (Long Island Science \& Engineering Fair) --- Won First Place Grand Award in Math in March 2009; selected to participate in ISEF 2009
-      \item
-        2009 National Merit Scholarship Finalist (February 2009)
-      \item
-        JSHS (Junior Science \& Humanities Symposium) --- Invited to present math research paper, February 2009 and February 2008
-      \item
-        Rensselaer Medal --- Awarded the Rensselaer Polytechnic Institute's Rensselaer Medal, June 2008
-      \item
-        Outstanding Junior Award --- sole recipient, awarded by Suffolk County Math Teacher's Association (June 2008)
-      \item
-        County All-Star Interscholastic Math League --- competed at NYSML (New York State Mathematics League) and ARML (American Regions Mathematics League) during junior, sophomore and freshmen years, 2006--2009; Achieved individual high score for team in 2009
-      \item
-        Suffolk County Senior High Interscholastic Math League --- $2^\text{nd}$ in county in 2009, $1^\text{st}$ in county in 2008, $3^\text{rd}$ in county in 2007
-      \item
-        NYSSEF (New York State Science and Engineering Fair)
-        \begin{itemize}
-          \item
-            2008 ($11^\text{th}$ grade) --- $1^\text{st}$ in math in Intel division, one of four first place individual winners in fair, Mu Alpha Theta award, Army award, Ohio Wesleyan University Scholarship
-          \item
-            2007 ($10^\text{th}$ grade) --- $1^\text{st}$ in math in Andromeda Division, earned Mu Alpha Theta Award from National Mathematics Honor Society
-        \end{itemize}
-      \item
-        Long Island Math Fair
-        \begin{itemize}
-          \item
-            2008 ($11^\text{th}$ grade) --- Earned gold medal for ``Group Theory'' project
-          \item
-            2007 ($10^\text{th}$ grade) --- Earned gold medal for ``Circuits over Sets of Natural Numbers'' project
-          \item
-            2006 ($9^\text{th}$ grade) --- Earned gold medal for ``Prime Numbers'' project
-          \item
-            2005 ($8^\text{th}$ grade) --- Earned gold medal for ``The Fourth Spatial Dimension'' project
-          \item
-            2004 ($7^\text{th}$ grade) --- Earned gold medal for ``The Seven Bridges of K\"onigsberg'' project
-        \end{itemize}
-      \item
-        AIME (American Invitational Mathematics Examination)
-        \begin{itemize}
-          \item
-            2009 ($12^\text{th}$ grade) - ???
-          \item
-            2008 ($11^\text{th}$ grade) - score of 8
-          \item
-            2007 ($10^\text{th}$ grade) - score of 4
-          \item
-            2006 ($9^\text{th}$ grade) - score of 1
-        \end{itemize}
-      \item
-        American Mathematics Contest 12 (AMC-12)
-        \begin{itemize}
-          \item
-            2009 ($12^\text{th}$ grade) - score of ???
-          \item
-            2008 ($11^\text{th}$ grade) - score of 120, top 1\% nationwide, earned title of ``School Winner''
-          \item
-            2007 ($10^\text{th}$ grade) - score of 103.5, top 1\% nationwide, earned title of ``School Winner''
-        \end{itemize}
-      \item
-        American Mathematics Contest 10 (AMC-10)
-        \begin{itemize}
-          \item
-            2006 ($9^\text{th}$ grade) - score of 127.5, top 1\% nationwide, earned title of ``School Winner''
-          \item
-            2005 ($8^\text{th}$ grade) - score of 118.5, earned title of ``School Winner'' in high school as an $8^\text{th}$ grader in the $9^\text{th}$ grade honors math class
-        \end{itemize}
-      \item
-        Suffolk County Mathematics Contest
-        \begin{itemize}
-          \item
-            2009 ($12^\text{th}$ grade) - ???
-          \item
-            2008 ($11^\text{th}$ grade) - $1^\text{st}$ place in county competing with $12^\text{th}$ graders
-          \item
-            2007 ($10^\text{th}$ grade) - $4^\text{th}$ place in county competing with $11^\text{th}$ graders
-          \item
-            2006 ($9^\text{th}$ grade) - $4^\text{th}$ place in county competing with $10^\text{th}$ graders
-          \item
-            2005 ($8^\text{th}$ grade) - $5^\text{th}$ place in county competing with $9^\text{th}$ graders
-          \item
-            2004 ($7^\text{th}$ grade) - $1^\text{st}$ place in county
-        \end{itemize}
-      \item
-        Suffolk County Mathematics Teachers Association (SCMTA) Tournament
-        \begin{itemize}
-          \item
-            2009 ($12^\text{th}$ grade) - ??????th in county
-          \item
-            2008 ($11^\text{th}$ grade) - $4^\text{th}$ in county
-          \item
-            2007 ($10^\text{th}$ grade) - $2^\text{nd}$ in county
-          \item
-            2006 ($9^\text{th}$ grade) - $1^\text{st}$ in county
-        \end{itemize}
-      \item
-        Commack High School Mathematics Department Awards --- June 2008, June 2007, and June 2006
-      \item
-        Commack Middle School Mathematics Departmental Award --- was only student of 605 students to receive Commack Middle School mathematics departmental award, 2005
-      \item
-        Suffolk County Mathcounts --- placed first in county as individual, 2005
-      \item
-        USA Mathematics Talent Search (USAMTS), 2007--2008; Placed on National List of Students with 6 Commended Solutions in 2008, silver medal in 2008
-      \item
-        Harvard-MIT Mathematics Tournament (HMMT), February 2009, 2008, and 2007
-      \item
-        New York State Science Olympiads
-        \begin{itemize}
-          \item
-            February 2008 --- $2^\text{nd}$ place in Environmental Chemistry, $4^\text{th}$ place in Circuit Lab, $4^\text{th}$ place in Five Star Science
-          \item
-            February 2007 --- $4^\text{th}$ place in Circuit Lab
-        \end{itemize}
-      \item
-        Mandelbrot Competition, team that competes in national competitions, 2006--2009; Placed on National Leaderboard (Top Tier) in 2008; co-captain in 2007--2008, captain in 2008--2009
-      \item
-        American Computer Science League (ACSL) Club, 2005--present; President in 2008--2009
-      \item
-        United States of America Computer Olympiad (USACO), 2005--present
-      \item
-        Suffolk County Senior High Interscholastic Math League, 2005--present
-      \item
-        Article was published in Teen Ink Magazine (November 2007)
-      }
-    \end{itemize}
-
-\end{multitex*}
-
-\begin{multitex*}[\highschoolfiles]
-\section{\textsc{Community Service and Involvement}}
-  \begin{itemize}
-    \item
-      Member of National Math Honor Society; Tutor students in math both in and out of school (2007--2009)
-    \item
-      Member of National Technology Honor Society (2007--2009)
-  \end{itemize}
-\end{multitex*}
-
-\begin{multitex}
-\section{\textsc{Other Activities}}
-\end{multitex}
-\begin{multitex}
-    \begin{itemize}
-      \onlymultitex*{\cvfiles,computer-science}{%
-      \item Co-maintainer of the Fiat Cryptography project (\href{https://github.com/mit-plv/fiat-crypto}{\texttt{mit-plv/fiat-crypto}} on GitHub)
-      \item Co-maintainer of the homotopy type theory Coq repository (\href{https://github.com/HoTT/HoTT}{\texttt{HoTT/HoTT} on GitHub})
-      \item Program Committee Member of CoqPL 2022
-      \onlymultitex*{\cvlongfiles}{%
-        \item MIRI Decision Theory Workshop Attendee: Formalized various versions of L\"ob's theorem in Agda and Coq
-        }
-      \item Commiter to the SIPB BarnOwl project (\url{https://barnowl.mit.edu})
-      }
-      \onlymultitex*{\cvfiles}{%
-      \item SIPB (Student Information and Processing Board) Member
-      }
-      \onlymultitex*{\cvlongfiles}{
-      \item Was project leader for MITeX, an online interface for composing \LaTeX
-      }
-      \onlymultitex*{\cvfiles}{%
-      \item President of Tech Squares, MIT's Square Dancing Club (May 2013--October 2014)
-      }
-      \onlymultitex*{\cvlongfiles}{%
-      \item HMMT Solutions Editor (2010)
-      }
-      \item Canada/USA Mathcamp (Summers 2006--2009)
-    \end{itemize}
-
-\end{multitex}
-
-\begin{multitex*}[\cvlongfiles]
-% this is a hack
-%\clearpage
-\section{\textsc{Interests}}
-    \begin{itemize}
-      \item Philosophy
-      \onlymultitex*{\cvlongfiles}{%
-      \item Programming
-      \item Dancing, especially tango, contra, and square dancing
-        }
-      \item Neuroscience
-      \onlymultitex*{\cvlongfiles}{%
-      \item Hiking
-      %\onlymultitex*{\highschoolfiles}{
-      \item MIT Mystery Hunt, Participated on Manic Sages Team, January 2008--2012
-        %  }
-        }
-    \end{itemize}
-\end{multitex*}
-
-\begin{multitex}
-  \nocite{*}
-\end{multitex}
-
-\begin{multitex*}[\cvshortfiles]
   \clearpage
@@
   \printbibliography[title={$\left.\right.$},notcategory=exclude]
-\end{multitex*}
 
-\begin{multitex*}[\cvlongfiles]
-  \section{\textsc{Presentations and Publications}}
-  \printbibliography[title={$\left.\right.$}]
-\end{multitex*}
-
-\begin{multitex}
   \endresume

### 2023-03-31 resume/ResumeOld.tex
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@de630c64253fb9a226076168f6ef29b037bff769 🚀

% LaTeX file for resume
% This file uses the resume document class (res.cls)

\documentclass[11pt]{res}
\usepackage[T1]{fontenc}
\usepackage{xifthen,verbatim}
\usepackage{hyperref}
\usepackage{etoolbox}[2011/01/03]

\usepackage[backend=biber,bibencoding=utf-8,maxbibnames=99,sorting=ymdnt,noerroretextools=true,style=alphabetic]{biblatex}
% http://tex.stackexchange.com/a/140641
\DeclareSortingTemplate{ymdnt}{
    \sort{
        \field{presort}
    }
    \sort[final]{
        \field{sortkey}
    }
    \sort[direction=descending]{
        \field{sortyear}
        \field{year}
        \literal{9999}
    }
    \sort[direction=descending]{
        \field[padside=left,padwidth=2,padchar=0]{month}
        \literal{99}
    }
    \sort[direction=descending]{
        \field[padside=left,padwidth=2,padchar=0]{day}
        \literal{99}
    }
    \sort{
        \field{sortname}
        \field{author}
        \field{editor}
        \field{translator}
        \field{sorttitle}
        \field{title}
    }
    \sort{
        \field{sorttitle}
        \field{title}
    }
}

\makeatletter

\let\origifempty\ifempty
\undef\ifempty
\let\origforlistloop\forlistloop
\usepackage{etextools}
\let\eifempty\ifempty
\let\ifempty\origifempty
\let\forlistloop\origforlistloop
\def\preetextools{\let\ifempty\eifempty}

\newif\ifmultitexcompile
\multitexcompiletrue

\newcommand\multitex@separator{-}
\newwrite\multitex@temp
% \immediate\openout\multitex@temp{\jobname.mtx}
\newtoks\multitex@names
\multitex@names={}
\newcommand{\multitex@postprocess}[1]{%
  \immediate\expandafter\closeout\csname @multitex@#1@file\endcsname
  \ifmultitexcompile
    \immediate\write18{pdflatex \csname @multitex@#1@name\endcsname}%
    \immediate\write18{biber \csname @multitex@#1@barename\endcsname}%
    \immediate\write18{pdflatex \csname @multitex@#1@name\endcsname}%
  \fi
}
\newcommand{\onlymultitex}[1]{\FE@ifstar{#1}{\only@other@multitex}{\only@multitex}}
\newcommand{\only@multitex}[2]{#2}
\newcommand{\only@other@multitex}[2]{}
{\catcode`\#=12
  \global\def\@pound{#}
}
\newcommand{\DeclareFile}[2][]{
  \global\multitex@names=\expandafter{\the\multitex@names,#2}
  \expandafter\xdef\csname @multitex@#2\endcsname{\xifempty{#1}{#2}{#1}}
  \expandafter\xdef\csname @multitex@#2@name\endcsname{"\jobname\multitex@separator\csname @multitex@#2\endcsname.tex"}
  \expandafter\xdef\csname @multitex@#2@auxname\endcsname{"\jobname\multitex@separator\csname @multitex@#2\endcsname.aux"}
  \expandafter\xdef\csname @multitex@#2@barename\endcsname{"\jobname\multitex@separator\csname @multitex@#2\endcsname"}
  \expandafter\newwrite\csname @multitex@#2@file\endcsname
  \immediate\expandafter\openout\csname @multitex@#2@file\expandafter\endcsname\csname @multitex@#2@name\endcsname
  \AtBeginDocument{%
    \multitex@write{#2}{%
      \detokenize{\newcommand{\onlymultitex}[1]}%
        {\detokenize{\csname FE@ifstar\endcsname}%
          {\@pound1}%
          \detokenize{%
            {\csname only@multitex\endcsname}{\csname only@multitex\endcsname}%
          }%
        }%
      \detokenize{%
        \expandafter\newcommand\csname only@multitex\endcsname[2]}%
        {%
          \noexpand\renewcommand*{\noexpand\do}[1]{%
            \noexpand\ifstrequal{\@pound\@pound1}{#2}%
              {\@pound2\noexpand\listbreak}%
              {}%
            }%
          {\noexpand\edef\noexpand\foo{\@pound1}\noexpand\expandafter}%
          \noexpand\expandafter\noexpand\docsvlist\noexpand\expandafter{\noexpand\foo}%
        }
      \detokenize{%
        \expandafter\let\csname only@other@multitex\expandafter\endcsname\csname only@multitex\endcsname
      }%
    }%
  }
  \AtEndDocument{\multitex@postprocess{#2}}
}
\newcommand{\multitex@write}[2]{\xifblank{#1}{}{\immediate\expandafter\write\csname @multitex@#1@file\endcsname{#2}}}
\edef\multitex@tempfilename{\jobname.mtx}
\newcommand{\multitex@pre@writetothis}{\immediate\openout\multitex@temp\multitex@tempfilename}%\long\gdef\multitex@thiscode{}}%
\newcommand{\multitex@writetothis}[1]{\immediate\write\multitex@temp{#1}}%\ExpandNext{\long\gappto\multitex@thiscode}{\expandnext{\relax\scantokens}{#1}}}%
\newcommand{\multitex@post@writetothis}{\immediate\closeout\multitex@temp\relax\input{\multitex@tempfilename}}%\show\multitex@thiscode\multitex@thiscode\relax}%
\newcommand{\multitex@writetoothers}[2][\the\multitex@names]{{\preetextools\ExpandNext{\forcsvloop}{#1}\do{\multitex@write{##1}{#2}}}}
\newcommand{\DeclareDocumentClass}[1]{\multitex@writetoothers{\detokenize{#1}}}
\AtBeginDocument{\multitex@writetoothers{\detokenize{\begin{document}}}}
\AtEndDocument{\multitex@writetoothers{\detokenize{\end{document}}}}
\newcommand{\includein}[2][]{\multitex@pre@writetothis\multitex@writetothis{\detokenize{#2}}\multitex@post@writetothis\multitex@writetoothers[#1]{\detokenize{#2}}}
\AfterEndEnvironment{multitex}{\multitex@post@writetothis}
\newenvironment{multitex}[1][\the\multitex@names]{%
  \@bsphack
  \begingroup% Lets Keep the Changes Local
    \multitex@pre@writetothis
    \let\do\@makeother\dospecials\catcode`\^^M\active
    \def\verbatim@processline{\multitex@writetothis{\the\verbatim@line}\multitex@writetoothers[#1]{\the\verbatim@line}}%
    \verbatim@start\space % I'm not sure why I need the space, but I seem to need it.
}{\endgroup
  \@esphack
}
% only others
\newenvironment{multitex*}[1][\the\multitex@names]{%
  \@bsphack
  \begingroup% Lets Keep the Changes Local
    \let\do\@makeother\dospecials\catcode`\^^M\active
    \def\verbatim@processline{\multitex@writetoothers[#1]{\the\verbatim@line}}%
    \verbatim@start\space % I'm not sure why I need the space, but I seem to need it.
}{\endgroup
  \@esphack
}
\makeatother

\DeclareFile{curriculum-vitae}
\DeclareFile{mathematics}
\DeclareFile{computer-science}
\DeclareFile{curriculum-vitae-long}
\DeclareFile{curriculum-vitae-mathematics}
\DeclareFile{curriculum-vitae-computer-science}
\DeclareDocumentClass{\documentclass[11pt]{res}}

\begin{multitex}
  \edef\resumefiles{mathematics,computer-science}
  \edef\cvshortspecialfiles{curriculum-vitae-mathematics,curriculum-vitae-computer-science}
  \edef\cvshortfiles{curriculum-vitae,\cvshortspecialfiles}
  \edef\cvlongfiles{curriculum-vitae-long}
  \edef\cvfiles{\cvshortfiles,\cvlongfiles}
  \edef\highschoolfiles{curriculum-vitae-long}
  \edef\mathfiles{mathematics,curriculum-vitae-mathematics}
  \edef\cvlongandmathfiles{\mathfiles,curriculum-vitae-long}
  \edef\cvandmathfiles{\cvlongandmathfiles,curriculum-vitae,curriculum-vitae-computer-science}
  \edef\computersciencefiles{computer-science,curriculum-vitae-computer-science}
  \edef\cvlongandcomputersciencefiles{\computersciencefiles,curriculum-vitae-long}
  \edef\cvandcomputersciencefiles{curriculum-vitae,curriculum-vitae-mathematics,\cvlongandcomputersciencefiles}
  \edef\allexceptcv{\resumefiles,\cvshortspecialfiles}
  \edef\allexceptcvlongfiles{\resumefiles,\cvshortfiles}
  \edef\allexceptcvandcvlongfiles{\resumefiles,\cvshortspecialfiles}
  %\usepackage[utf8x]{inputenc}
  %\usepackage[T1]{fontenc}
  %\usepackage{helvetica} % uses helvetica postscript font (download helvetica.sty)
  %\usepackage{newcent}   % uses new century schoolbook postscript font
  \usepackage[T1]{fontenc}
  \usepackage[margin=1in,top=0.5in,bottom=0.5in]{geometry}
  \usepackage{hyperref}
  \usepackage{calc}
  \usepackage{amsmath}
  \usepackage{etoolbox}[2011/01/03]
  %\bibliographystyle{plainyr-rev}
  \usepackage[backend=biber,maxbibnames=99,sorting=ymdnt,noerroretextools=true,style=alphabetic]{biblatex}
% http://tex.stackexchange.com/a/140641
\DeclareSortingTemplate{ymdnt}{
    \sort{
        \field{presort}
    }
    \sort[final]{
        \field{sortkey}
    }
    \sort[direction=descending]{
        \field{sortyear}
        \field{year}
        \literal{9999}
    }
    \sort[direction=descending]{
        \field[padside=left,padwidth=2,padchar=0]{month}
        \literal{99}
    }
    \sort[direction=descending]{
        \field[padside=left,padwidth=2,padchar=0]{day}
        \literal{99}
    }
    \sort{
        \field{sortname}
        \field{author}
        \field{editor}
        \field{translator}
        \field{sorttitle}
        \field{title}
    }
    \sort{
        \field{sorttitle}
        \field{title}
    }
}
  \addbibresource{jason-gross.bib}
  \DeclareBibliographyCategory{exclude}
  \makeatletter
  \@ifpackageloaded{etextools}{%
  }{%
    \let\origifempty\ifempty
    \let\origforlistloop\forlistloop
    \undef\ifempty
    \usepackage{etextools}
    \let\eifempty\ifempty
    \let\ifempty\origifempty
    \let\forlistloop\origforlistloop
    \def\preetextools{\let\ifempty\eifempty}
  }%
  \makeatother
  \begin{format}
    \employer{l}\dates{r}\\
    \title{l}\location{r} \\
    \body
  \end{format}
  \let\oldemployer=\employer
  \let\oldtitle=\title
  \renewcommand{\employer}[1]{\oldemployer{\textbf{#1}}}
  \renewcommand{\title}[1]{\oldtitle{\textit{#1}}}

  % \let\oldposition=\position
  % \let\oldendposition=\endposition
  % \renewenvironment{position}{\oldposition\vspace*{-2\baselineskip}}{\endoldposition}

  \makeatletter
  \let\old@itemize=\itemize
  \def\itemize{\old@itemize\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}\setlength{\leftskip}{-1em}}
  \renewenvironment{position}
    {%
    \begingroup
      \par
        \the\tabular@head
  %     \addpenalty{-\@secpenalty}% bad place for a page break
      \penalty -\@secpenalty % bad place for a page break
      \penalty 10000
      \ignorespaces
      \vspace*{-\baselineskip}
    }{%
        \the\tabular@tail
  %     \addpenalty{\@secpenalty}% good place for a page break
      \penalty \@secpenalty % good place for a page break
      \endgroup
  }
  \makeatother

  \newsectionwidth{1em}
  \resumewidth=7.5in

  \newcommand{\aswidthof}[2]{\rlap{#1}\hphantom{#2}}
\end{multitex}
\begin{multitex*}[\cvlongfiles]
  \name{\LARGE \bf \hspace{\stretch{1}} Jason Gross \hspace{\stretch{1}} \\
    \hspace{\stretch{1}}\large \href{mailto:jgross@mit.edu}{jgross@mit.edu} --- (631) 790-8962 --- \url{https://people.csail.mit.edu/jgross/}\hspace{\stretch{1}} \\}

  \address{\textsc{Present Address}\\258 Prospect Street, Apt \# 1L\\Cambridge, MA 02139}
  \address{\textsc{Permanent Address}\\126 Hayrick Lane \\  Commack, NY 11725}
\end{multitex*}
\begin{multitex}[\allexceptcvlongfiles]
  \name{\LARGE \bf \parbox{\widthof{github.com/JasonGross}}{%
      \centering Jason Gross \\
      \vspace{0.2cm}
      \large \href{https://github.com/JasonGross/}{github.com/JasonGross} \\
      \href{https://people.csail.mit.edu/jgross/}{people.csail.mit.edu/jgross} \\
      $\left.\right.$ \\
      $\left.\right.$ \\%
  }}

  \vspace{5cm}

  %\address{}%\textsc{Address}\\258 Prospect Street, Apt \# 1L\\Cambridge, MA 02139}
  %\address{\textsc{Contact}\\\href{mailto:jgross@mit.edu}{jgross@mit.edu}\\(631) 790-8962}
  \newcommand{\addressaswidth}[1]{\makebox[\widthof{\textsc{jgross@mit.edu}}][c]{#1}}
  \address{\addressaswidth{\textsc{Contact}}\\\addressaswidth{\href{mailto:jgross@mit.edu}{jgross@mit.edu}}\\\addressaswidth{(631) 790-8962}}
\end{multitex}

\begin{document}

\begin{multitex}
  \resume
\end{multitex}

\begin{multitex*}[\allexceptcvandcvlongfiles]
  \section{\textsc{Research Interests}}
  Programming Languages, Formal Verification, Cryptography, Performance of Automation in Interactive Proof Assistants, Homotopy Type Theory, Category Theory, Program Synthesis
\end{multitex*}
\begin{multitex}[\cvlongfiles]
  \section{\textsc{Research Interests}}
  Programming Languages, Formal Verification, Cryptography, Performance Engineering
\end{multitex}

\begin{multitex}
  %\section{JOB OBJECTIVE}
    %A summer position that will use my accounting and computer
    %skills.
    %
  \section{\textsc{Education}}
\end{multitex}

\begin{multitex}[\allexceptcvlongfiles]
  \employer{Massachusetts Institute of Technology}
  \dates{2013--2021}
  \title{{\rm PhD in Computer Science}}
  \location{Cambridge, MA}
  \begin{position}\\
      Advisor: Adam Chlipala\\
    \textit{Thesis: Performance Engineering of Proof-Based Software Systems at Scale}\\
    \textit{SM Thesis: An Extensible Framework for Synthesizing Efficient, Verified Parsers}
  \end{position}

  \employer{Massachusetts Institute of Technology}
  \dates{2009--2013}
  \title{{\rm BS in Mathematics and Physics}}
  \location{Cambridge, MA}
  \begin{position}\\
      GPA: 4.6/5
  \end{position}
\end{multitex}

\begin{multitex*}[\cvlongfiles]
    \textbf{Massachusetts Institute of Technology}, Cambridge, MA  \\
    Began Doctorate of Philosophy in Computer Science in September 2013 \\
    \onlymultitex*{\cvfiles}{Research Interests: Programming Languages, Performance of Automation in Interactive Proof Assistants, Formal Verification, Homotopy Type Theory, Category Theory, Program Synthesis, Type Theory \\}
    G.P.A. 4.9/5.0 \\
    \\
    Bachelor of Science in Mathematics and Physics, June 2013   \\
    G.P.A. 4.6/5.0 \\
\end{multitex*}
\begin{multitex*}[\resumefiles]%
    \\
    Relevant Coursework: %Computer Systems
        Security, %Foundations of
        Program Analysis, Performance Engineering of Software% Systems
        , Statistical Physics, Quantum% Physics
        , Topology, Analysis, Waves \& Vibrations, Special Relativity% and the Physics of Spacetime%
        , \mbox{Algebra}% force no hyphenation
\end{multitex*}
\begin{multitex*}[\cvlongfiles]%
    \\
		Coursework:
		\begin{itemize}
			\item Computer Science: Inference and Information, Computer Systems Security, Geometric Computing, Foundations of Program Analysis, Performance Engineering of Software Systems, Structure and Interpretation of Computer Programs
			\item Mathematics: Paradox and Infinity, Category Theory for Scientists, Algebraic Topology I, Seminar in Topology, Introduction to Topology, Real Analysis, Abstract Algebra I \& II, Differential Equations, Calculus I \& II, Linear Algebra
			\item Physics: Modern Astrophysics, General Relativity (Graduate), Statistical Physics I \& II, Quantum Physics I, II, \& III, Classical Mechanics II, Waves and Vibrations, Special Relativity and the Physics of Spacetime
			\item Other Sciences: Introductory Biology, Introductory Chemistry, Design a Concentrated Solar Power Water Heater (seminar)
			\item Other: The Art and Science of Happiness, Metaphysics, Introduction to Musical Composition, Capitalism and Its Critics, Moral Psychology, Philosophy of Quantum Mechanics, Moral Problems and the Good Life, Philosophy of Love, Kaufman Teaching Certificate Program
		\end{itemize}
\end{multitex*}

% \item Presented ``Computational Higher Inductive Types: Computing with Custom Equalities'', October 6, 2013, The CSAIL Student Workshop

\begin{multitex*}[\cvlongfiles]
\section{\textsc{Experience}}
\end{multitex*}
\begin{multitex*}[\cvlongfiles]
  \employer{MIT CSAIL}
  \dates{September 2013--February 2021}
  \title{Researcher}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \item Researching program synthesis and verification with Adam Chlipala
      \item Collaboratively working on \onlymultitex*{\cvfiles}{implementing }one of the world's first algorithm-level-optimizing compilers
      \item Collaboratively implemented proven-correct crypto code now used by Google Chrome
    \end{itemize}
  \end{position}
\end{multitex*}

\begin{multitex}
\section{\textsc{Experience}}
\end{multitex}

\begin{multitex*}[\cvfiles]
  \employer{Coq Development Team, INRIA}
  \dates{June 2021--Present}
  \title{Core Team Member}
  \location{Nantes, France (remote)}
  \begin{position}
    \begin{itemize}
      \item Stress Testing, Bug Reporting, Bug Minimizer, Compatibility Assurance
      \item Engineering and maintaining a bug minimizer for automatically producing minimized stand-alone test-cases from buggy code
      \item Researching performance issues that impact scalability of automated verification
      \item Have reported the plurality of all-time bugs in Coq (since 2012)
    \end{itemize}
  \end{position}

  \employer{Machine Intelligence Research Institute}
  \dates{February 2021--Present}
  \title{Research Staff}
  \location{Berkeley, CA (remote)}
  \begin{position}
    \begin{itemize}
      \item Performing self-directed research into topics in fundamental programming language theory and mathematics%, bringing insights and understanding back to my supervisor
    \end{itemize}
  \end{position}

  \employer{MIT CSAIL}
  \dates{September 2013--February 2021}
  \title{PhD Researcher}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \item Main Research Project: Fiat Cryptography (\href{https://github.com/mit-plv/fiat-crypto}{github.com/mit-plv/fiat-crypto})
      \item Collaboratively implemented proven-correct cryptographic code now used by Google Chrome, and in the majority of secure connects from web browsers
      \item Led development of one of the world's first algorithm-level-optimizing compilers
      \item Wrote backends to C, Go, Java, and JSON; managed development of backends to Rust and Zig
    \end{itemize}
  \end{position}
\end{multitex*}

\begin{multitex}
\section{\textsc{Internships}}
\end{multitex}

\begin{multitex*}[\cvfiles]
  \employer{Machine Intelligence Research Institute}
  \dates{June 2019--August 2019}
  \title{Type Theory Intern}
  \location{Berkeley, CA}
  \begin{position}
    \begin{itemize}
      \item Worked on formalizing type theories and on proving things within proof assistants
    \end{itemize}
  \end{position}
\end{multitex*}
\begin{multitex}
  \bgroup
  \interlinepenalty 10000
  \employer{Google}
  \dates{June 2018--August 2018}
  \title{Software Engineering Intern}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \item Worked with BoringSSL on integration of proven-correct low-level ECC primitives into Chrome
    \end{itemize}
  \end{position}
  \egroup
\end{multitex}
\begin{multitex*}[\cvfiles]
  \bgroup
  \interlinepenalty 10000
  \employer{Google}
  \dates{June 2016--September 2016}
  \title{Software Engineering Intern}
  \location{Mountain View, CA}
  \begin{position}
    \begin{itemize}
      \item Formalized low-level ECC primitives with proofs of correctness
    \end{itemize}
  \end{position}
  \egroup
\end{multitex*}
\begin{multitex}[\cvlongandmathfiles]
  \employer{MIT}
  \dates{Fall 2009--Present}
  \title{Teacher}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \item Taught classes on \LaTeX, philosophy, linear algebra, and quantum mechanics for MIT Educational Studies Program's Splash, Spark, and Summer HSSP (High School Studies Program)
      \onlymultitex{\cvandmathfiles}{\item Teaching Assistant for 8.012 (Physics I) and 8.022 (Physics II) in Experimental Study Group}
    \end{itemize}
  \end{position}
\end{multitex}
\begin{multitex*}[\cvlongandcomputersciencefiles]
  \employer{MIT}
  \dates{September 2015--December 2015}
  \title{TA for 6.172 (Performance Engineering% of Software% Systems
  )}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \item Created and led recitations, taught students, and helped run class
      \item Analyzed and explained assembly output of \texttt{gcc -O3} to teach vectorization
    \end{itemize}
  \end{position}
\end{multitex*}
\begin{multitex*}[\cvlongfiles]
    \employer{MIRI}
    \dates{June 12--14, 2015}
    \title{Decision Theory Workshop Attendee}
    \location{Berkeley, CA}
    \begin{position}
        \begin{itemize}
            \item Formalized various versions of L\"ob's theorem in Agda and Coq
            \item Learned about and worked on problems on the frontiers of decision theory
        \end{itemize}
    \end{position}
\end{multitex*}
\begin{multitex}
  \employer{Microsoft Research}
  \dates{June 2014--August 2014}
  \title{Intern}
  \location{Cambridge, United Kingdom}
  \begin{position}
    \begin{itemize}
      \item \onlymultitex*{\allexceptcvlongfiles}{Collaboratively c}\onlymultitex*{\cvlongfiles}{C}reated a language for specifying input/output behavior of x86 assembly programs\onlymultitex*{\cvlongfiles}{with Andrew Kennedy and Nick Benton}; Verified the I/O behavior of a number of simple programs
      \onlymultitex*{\cvfiles}{\item Improved automation of the x86proved library}
    \end{itemize}
  \end{position}
\end{multitex}
\begin{multitex*}[\cvfiles]
    \employer{MIT CSAIL}
    \dates{April 2012--June 2014}
    \title{Researcher}
    \location{Cambridge, MA}
    \begin{position}
        \begin{itemize}
            \item Entered a significant amount of category theory into the automated proof assistant Coq \onlymultitex*{\cvfiles}{(\url{https://github.com/HoTT/HoTT/tree/master/theories/Categories})}
            \onlymultitex{\cvfiles}{\item Made progress towards an interface for databases and database migration on top of category theory in Coq\onlymultitex*{\cvlongfiles}{ with David Spivak and Adam Chlipala}}
            \onlymultitex*{\cvlongfiles}{\item Presented ``Building Database Management on top of Category Theory in Coq'', January 25, 2013, POPL 2013: 40th ACM SIGPLAN-SIGACT Symposium on Principles of Programming Languages}
            \onlymultitex*{\cvlongfiles}{\item Presented	 ``Experience Implementing a Performant Category-Theory Library in Coq'', July 14, 2014, Interactive Theorem Proving 2014 Conference}
        \end{itemize}
    \end{position}
\end{multitex*}
\begin{multitex*}[\cvfiles]
  \employer{MIT CSAIL}
  \dates{November 2009--September 2011}
  \title{Researcher}
  \location{Cambridge, MA}
  \begin{position}
    \begin{itemize}
      \onlymultitex*{\cvfiles}{\item Designed from scratch a data collection webpage, collected data for, and helped with research \onlymultitex*{\cvlongfiles}{ of Brenden Lake, Ruslan Salakhutdinov, and Josh Tenenbaum, }on categorical and transfer learning\onlymultitex*{\cvlongfiles}{ (\url{https://jgross.scripts.mit.edu/alphabets/})}}
      \onlymultitex*{\cvlongfiles,\resumefiles}{%
      \item Co-author of ``One shot learning of simple visual concepts''\onlymultitex*{\cvfiles}{ published in \emph{Proceedings of the 33rd Annual Meeting of the Cognitive Science Society}}.
        }
    \end{itemize}
  \end{position}
\end{multitex*}
\begin{multitex*}[\cvlongfiles]
  \employer{MIT OCW}
    \dates{May--August 2010}
    \title{}
    \location{Cambridge, MA}
    \begin{position}
        \begin{itemize}
            \item Evaluated and categorized videos of Walter Lewin solving physics problems for updating the 8.01 (Physics I) OpenCourseWare (OCW) website.
        \end{itemize}
    \end{position}
\end{multitex*}
\begin{multitex*}[\cvfiles]
  \employer{Commack High School}
  \dates{Fall 2006--Summer 2009}
  \title{Independent Researcher}
  \location{Commack, NY}
  \begin{position}
    \begin{itemize}
      \item Independently researched circuits over sets of natural numbers for three years.
      \item Won fourth place award in mathematics in ISEF (Intel International Science and Engineering Fair) in 2009, third place award in ISEF 2008.
    \end{itemize}
  \end{position}
\end{multitex*}

\begin{multitex*}[\cvlongfiles]
  \employer{Turnpike Total Appliance}
  \dates{Fall 2006--Summer 2009}
  \title{Web Page Designer}
  \location{Commack, NY}
  \begin{position}
    \begin{itemize}
      \item Designed and improved main company website, took initiative to make more improvements.
    \end{itemize}
  \end{position}

\end{multitex*}

\begin{multitex}[curriculum-vitae]
\section{\textsc{Programming Languages}}
\end{multitex}
\begin{multitex*}[\allexceptcv]
\section{\textsc{Computer Skills}}
\end{multitex*}
\begin{multitex}
    \begin{itemize}
      \item
        Proficient: Coq, \onlymultitex*{\cvlongfiles}{\TeX{} macro language, }Mathematica, git, Python, JavaScript, BASIC
      \item
        Working knowledge: \onlymultitex*{\cvlongfiles}{\LaTeX{}, }C, C++, Agda, OCaml, Haskell, Scheme, HTML, CSS, Perl, Java
      \item
        Basic knowledge: Matlab, Lean, Idris, Ruby, Go, Ur/Web, x86 Assembly
    \end{itemize}
\end{multitex}

\begin{multitex}
  \section{\textsc{Teaching}}
  \begin{itemize}
  \item Instructor at Monsoon Math Camp: category theory, linear logic, L\"ob's theorem
  \item TA for 6.172 (Performance Engineering% of Software% Systems
    ): Led recitations, analyzed and explained assembly output of \texttt{gcc -O3} to teach vectorization
  \item TA for 8.012 (Physics I) and 8.022 (Physics II) in Experimental Study Group
  \item Teacher at MIT ESP Programs: \LaTeX, philosophy, linear algebra, quantum mechanics
  \end{itemize}
\end{multitex}

\begin{multitex*}[\cvlongfiles]
\section{\textsc{Honors and Awards}}
\end{multitex*}
\begin{multitex*}[\cvlongfiles]
    \begin{itemize}
      \onlymultitex*{\highschoolfiles,\cvfiles}{%
      \item
        Mathematics Honor Society (Commack High School)
      }
      \item
        Collection of 12 original K'NEX synagogue models exhibited in various museums (2004--2008)
      \onlymultitex*{\highschoolfiles}{%
        \begin{itemize}
          \item
            Collection of 12 original K'NEX synagogue models
          \item
            Volunteered to give speeches and conduct K'NEX workshops for children at opening of exhibits, as well as in many locations in NY area
          \item
            April 2008 --- Gave interview for April edition of BabagaNewz Magazine and audio interview for BabagaNewz.com
          \item
            October 2006 --- Displayed in Gurwin Nursing Home on Long Island
          \item
            February 2006 --- Exhibited at Lisa Ann Watson Children's Discovery Museum in Miami, Florida
          \item
            October 2005 --- Displayed at Shomrim (Society for New York City Jewish Police Officers)
          \item
            February--April 2005 --- Exhibited in Yeshiva University Museum in Manhattan
          \item
            December 2004 --- Displayed at Suffolk Y-JCC in Commack
        \end{itemize}
        }
      \onlymultitex*{\cvfiles}{
      \item
        Graduated Cum Laude from Commack High School (June 2009)
      }
      \onlymultitex*{\highschoolfiles}{%
      \item
        Commack High School Outstanding Academic Performance (June 2009)
      \item
        Florence Burns Scholarship for Achievement of Excellence in Computer Science (June 2009)
      \item
        Theodore R. Spedalle Award for Achievement of Excellence in Mathematics (June 2009)
      \item
        Commack High School Mathematics Research Award (June 2009)
      \item
        President's Award for Educational Excellence (June 2009)
      \item
        AP Scholar with Distinction (2009)
      \item
        Robert C. Byrd Honors Scholarship (June 2009)
      \item
        ISEF (Intel International Science \& Engineering Fair)
        \begin{itemize}
          \item
            Math research paper, ``Circuits over Sets of Natural Numbers'' won $3^\text{rd}$ Place in Math in Atlanta, May 2008
          \item
            Math research paper, ``An Investigation of the Closure of the Set of Singleton Sets of Natural Numbers under $\cup$, $\cap$, $\overline{\phantom{\emptyset}}$, $+$, $\times$'' won $4^\text{th}$ place in Math in Reno, May 2009
       \end{itemize}
      \item
        LISEF (Long Island Science \& Engineering Fair) --- Won First Place Grand Award in Math in March 2009; selected to participate in ISEF 2009
      \item
        2009 National Merit Scholarship Finalist (February 2009)
      \item
        JSHS (Junior Science \& Humanities Symposium) --- Invited to present math research paper, February 2009 and February 2008
      \item
        Rensselaer Medal --- Awarded the Rensselaer Polytechnic Institute's Rensselaer Medal, June 2008
      \item
        Outstanding Junior Award --- sole recipient, awarded by Suffolk County Math Teacher's Association (June 2008)
      \item
        County All-Star Interscholastic Math League --- competed at NYSML (New York State Mathematics League) and ARML (American Regions Mathematics League) during junior, sophomore and freshmen years, 2006--2009; Achieved individual high score for team in 2009
      \item
        Suffolk County Senior High Interscholastic Math League --- $2^\text{nd}$ in county in 2009, $1^\text{st}$ in county in 2008, $3^\text{rd}$ in county in 2007
      \item
        NYSSEF (New York State Science and Engineering Fair)
        \begin{itemize}
          \item
            2008 ($11^\text{th}$ grade) --- $1^\text{st}$ in math in Intel division, one of four first place individual winners in fair, Mu Alpha Theta award, Army award, Ohio Wesleyan University Scholarship
          \item
            2007 ($10^\text{th}$ grade) --- $1^\text{st}$ in math in Andromeda Division, earned Mu Alpha Theta Award from National Mathematics Honor Society
        \end{itemize}
      \item
        Long Island Math Fair
        \begin{itemize}
          \item
            2008 ($11^\text{th}$ grade) --- Earned gold medal for ``Group Theory'' project
          \item
            2007 ($10^\text{th}$ grade) --- Earned gold medal for ``Circuits over Sets of Natural Numbers'' project
          \item
            2006 ($9^\text{th}$ grade) --- Earned gold medal for ``Prime Numbers'' project
          \item
            2005 ($8^\text{th}$ grade) --- Earned gold medal for ``The Fourth Spatial Dimension'' project
          \item
            2004 ($7^\text{th}$ grade) --- Earned gold medal for ``The Seven Bridges of K\"onigsberg'' project
        \end{itemize}
      \item
        AIME (American Invitational Mathematics Examination)
        \begin{itemize}
          \item
            2009 ($12^\text{th}$ grade) - ???
          \item
            2008 ($11^\text{th}$ grade) - score of 8
          \item
            2007 ($10^\text{th}$ grade) - score of 4
          \item
            2006 ($9^\text{th}$ grade) - score of 1
        \end{itemize}
      \item
        American Mathematics Contest 12 (AMC-12)
        \begin{itemize}
          \item
            2009 ($12^\text{th}$ grade) - score of ???
          \item
            2008 ($11^\text{th}$ grade) - score of 120, top 1\% nationwide, earned title of ``School Winner''
          \item
            2007 ($10^\text{th}$ grade) - score of 103.5, top 1\% nationwide, earned title of ``School Winner''
        \end{itemize}
      \item
        American Mathematics Contest 10 (AMC-10)
        \begin{itemize}
          \item
            2006 ($9^\text{th}$ grade) - score of 127.5, top 1\% nationwide, earned title of ``School Winner''
          \item
            2005 ($8^\text{th}$ grade) - score of 118.5, earned title of ``School Winner'' in high school as an $8^\text{th}$ grader in the $9^\text{th}$ grade honors math class
        \end{itemize}
      \item
        Suffolk County Mathematics Contest
        \begin{itemize}
          \item
            2009 ($12^\text{th}$ grade) - ???
          \item
            2008 ($11^\text{th}$ grade) - $1^\text{st}$ place in county competing with $12^\text{th}$ graders
          \item
            2007 ($10^\text{th}$ grade) - $4^\text{th}$ place in county competing with $11^\text{th}$ graders
          \item
            2006 ($9^\text{th}$ grade) - $4^\text{th}$ place in county competing with $10^\text{th}$ graders
          \item
            2005 ($8^\text{th}$ grade) - $5^\text{th}$ place in county competing with $9^\text{th}$ graders
          \item
            2004 ($7^\text{th}$ grade) - $1^\text{st}$ place in county
        \end{itemize}
      \item
        Suffolk County Mathematics Teachers Association (SCMTA) Tournament
        \begin{itemize}
          \item
            2009 ($12^\text{th}$ grade) - ??????th in county
          \item
            2008 ($11^\text{th}$ grade) - $4^\text{th}$ in county
          \item
            2007 ($10^\text{th}$ grade) - $2^\text{nd}$ in county
          \item
            2006 ($9^\text{th}$ grade) - $1^\text{st}$ in county
        \end{itemize}
      \item
        Commack High School Mathematics Department Awards --- June 2008, June 2007, and June 2006
      \item
        Commack Middle School Mathematics Departmental Award --- was only student of 605 students to receive Commack Middle School mathematics departmental award, 2005
      \item
        Suffolk County Mathcounts --- placed first in county as individual, 2005
      \item
        USA Mathematics Talent Search (USAMTS), 2007--2008; Placed on National List of Students with 6 Commended Solutions in 2008, silver medal in 2008
      \item
        Harvard-MIT Mathematics Tournament (HMMT), February 2009, 2008, and 2007
      \item
        New York State Science Olympiads
        \begin{itemize}
          \item
            February 2008 --- $2^\text{nd}$ place in Environmental Chemistry, $4^\text{th}$ place in Circuit Lab, $4^\text{th}$ place in Five Star Science
          \item
            February 2007 --- $4^\text{th}$ place in Circuit Lab
        \end{itemize}
      \item
        Mandelbrot Competition, team that competes in national competitions, 2006--2009; Placed on National Leaderboard (Top Tier) in 2008; co-captain in 2007--2008, captain in 2008--2009
      \item
        American Computer Science League (ACSL) Club, 2005--present; President in 2008--2009
      \item
        United States of America Computer Olympiad (USACO), 2005--present
      \item
        Suffolk County Senior High Interscholastic Math League, 2005--present
      \item
        Article was published in Teen Ink Magazine (November 2007)
      }
    \end{itemize}

\end{multitex*}

\begin{multitex*}[\highschoolfiles]
\section{\textsc{Community Service and Involvement}}
  \begin{itemize}
    \item
      Member of National Math Honor Society; Tutor students in math both in and out of school (2007--2009)
    \item
      Member of National Technology Honor Society (2007--2009)
  \end{itemize}
\end{multitex*}

\begin{multitex}
\section{\textsc{Other Activities}}
\end{multitex}
\begin{multitex}
    \begin{itemize}
      \onlymultitex*{\cvfiles,computer-science}{%
      \item Co-maintainer of the Fiat Cryptography project (\href{https://github.com/mit-plv/fiat-crypto}{\texttt{mit-plv/fiat-crypto}} on GitHub)
      \item Co-maintainer of the homotopy type theory Coq repository (\href{https://github.com/HoTT/HoTT}{\texttt{HoTT/HoTT} on GitHub})
      \item Program Committee Member of ITP 2023 and CoqPL 2022
      \onlymultitex*{\cvlongfiles}{%
        \item MIRI Decision Theory Workshop Attendee: Formalized various versions of L\"ob's theorem in Agda and Coq
        }
      \item Commiter to the SIPB BarnOwl project (\url{https://barnowl.mit.edu})
      }
      \onlymultitex*{\cvfiles}{%
      \item SIPB (Student Information and Processing Board) Member
      }
      \onlymultitex*{\cvlongfiles}{
      \item Was project leader for MITeX, an online interface for composing \LaTeX
      }
      \onlymultitex*{\cvfiles}{%
      \item President of Tech Squares, MIT's Square Dancing Club (May 2013--October 2014)
      }
      \onlymultitex*{\cvlongfiles}{%
      \item HMMT Solutions Editor (2010)
      }
      \item Canada/USA Mathcamp (Summers 2006--2009)
    \end{itemize}

\end{multitex}

\begin{multitex*}[\cvlongfiles]
% this is a hack
%\clearpage
\section{\textsc{Interests}}
    \begin{itemize}
      \item Philosophy
      \onlymultitex*{\cvlongfiles}{%
      \item Programming
      \item Dancing, especially tango, contra, and square dancing
        }
      \item Neuroscience
      \onlymultitex*{\cvlongfiles}{%
      \item Hiking
      %\onlymultitex*{\highschoolfiles}{
      \item MIT Mystery Hunt, Participated on Manic Sages Team, January 2008--2012
        %  }
        }
    \end{itemize}
\end{multitex*}

\begin{multitex}
  \nocite{*}
\end{multitex}

\begin{multitex*}[\cvshortfiles]
  \clearpage
  \section{\textsc{Selected Presentations and Publications}}
  \addtocategory{exclude}{Evaluation2021Huhmann}
  \printbibliography[title={$\left.\right.$},notcategory=exclude]
\end{multitex*}

\begin{multitex*}[\cvlongfiles]
  \section{\textsc{Presentations and Publications}}
  \printbibliography[title={$\left.\right.$}]
\end{multitex*}

\begin{multitex}
\endresume
\end{multitex}
\end{document}

### 2023-03-31 papers/lob-paper/tipa/mf/gentipx.sh

@@ -1,90 +1,0 @@
-#!/bin/sh
-
-# gentipx.sh --- a shell script for producing TIPX parameter files.
-# 
-# 2001/11/21 first version
-# 2001/12/21 tt8, tt9, tt10, tt12, ts10 (=SLTT10) added.
-
-AWK=gawk
-
-DATE="2003/01/01"
-VERSION="1.2"
-
-PARAMFILE="tipaprm.def"
-
-FILES="tipx8 tipx9 tipx10 tipx12 tipx17 tipxsl8 tipxsl9 tipxsl10 tipxsl12 \
-	tipxbx8 tipxbx9 tipxbx10 tipxbx12 tipxss8 tipxss9 tipxss10 tipxss12 \
-	tipxss17 tipxb10 tipxbs10 tipxsb10 tipxsi10 \
-	tipxtt8 tipxtt9 tipxtt10 tipxtt12 tipxts10"
-
-for FNAME in $FILES
-do
-  case "$FNAME" in
-    tipx8)    NUM=3; FID="TIPX";  FSIZE="8"; FDESC="Roman 8" ;;
-    tipx9)    NUM=4; FID="TIPX";  FSIZE="9"; FDESC="Roman 9" ;;
-    tipx10)   NUM=5; FID="TIPX";  FSIZE="10";FDESC="Roman 10" ;;
-    tipx12)   NUM=6; FID="TIPX";  FSIZE="12";FDESC="Roman 12" ;;
-    tipx17)   NUM=7; FID="TIPX";  FSIZE="17.28";FDESC="Roman 17.28" ;;
-    tipxsl8)  NUM=8; FID="TIPXSL";FSIZE="8"; FDESC="Slanted Roman 8" ;;
-    tipxsl9)  NUM=9; FID="TIPXSL";FSIZE="9"; FDESC="Slanted Roman 9" ;;
-    tipxsl10) NUM=10;FID="TIPXSL";FSIZE="10";FDESC="Slanted Roman 10" ;;
-    tipxsl12) NUM=11;FID="TIPXSL";FSIZE="12";FDESC="Slanted Roman 12" ;;
-    tipxbx8)  NUM=12;FID="TIPXBX";FSIZE="8"; FDESC="Bold Extended Roman 8" ;;
-    tipxbx9)  NUM=13;FID="TIPXBX";FSIZE="9"; FDESC="Bold Extended Roman 9" ;;
[…]

### 2023-03-31 resume/Resume.tex
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@40bb8bd01b90aaf91b9564c15f587273acc75411 🚀

@@
-% LaTeX file for resume
-% This file uses the resume document class (res.cls)
+\documentclass[11pt]{res}
 
-\documentclass[11pt]{res}
-\usepackage[T1]{fontenc}
-\usepackage{xifthen,verbatim}
-\usepackage{hyperref}
-\usepackage{etoolbox}[2011/01/03]
-
-\usepackage[backend=biber,bibencoding=utf-8,maxbibnames=99,sorting=ymdnt,noerroretextools=true,style=alphabetic]{biblatex}
-% http://tex.stackexchange.com/a/140641
-\DeclareSortingTemplate{ymdnt}{
-    \sort{
-        \field{presort}
-    }
-    \sort[final]{
-        \field{sortkey}
-    }
-    \sort[direction=descending]{
-        \field{sortyear}
-        \field{year}
-        \literal{9999}
-    }
-    \sort[direction=descending]{
-        \field[padside=left,padwidth=2,padchar=0]{month}
-        \literal{99}
-    }
-    \sort[direction=descending]{
-        \field[padside=left,padwidth=2,padchar=0]{day}
-        \literal{99}
-    }
-    \sort{
-        \field{sortname}
-        \field{author}
-        \field{editor}
-        \field{translator}
-        \field{sorttitle}
-        \field{title}
-    }
-    \sort{
-        \field{sorttitle}
-        \field{title}
-    }
-}
-
-\makeatletter
-
-\let\origifempty\ifempty
-\undef\ifempty
-\let\origforlistloop\forlistloop
-\usepackage{etextools}
-\let\eifempty\ifempty
-\let\ifempty\origifempty
-\let\forlistloop\origforlistloop
-\def\preetextools{\let\ifempty\eifempty}
-
-\newif\ifmultitexcompile
-\multitexcompiletrue
-
-\newcommand\multitex@separator{-}
-\newwrite\multitex@temp
-% \immediate\openout\multitex@temp{\jobname.mtx}
-\newtoks\multitex@names
-\multitex@names={}
-\newcommand{\multitex@postprocess}[1]{%
-  \immediate\expandafter\closeout\csname @multitex@#1@file\endcsname
-  \ifmultitexcompile
-    \immediate\write18{pdflatex \csname @multitex@#1@name\endcsname}%
-    \immediate\write18{biber \csname @multitex@#1@barename\endcsname}%
-    \immediate\write18{pdflatex \csname @multitex@#1@name\endcsname}%
-  \fi
-}
-\newcommand{\onlymultitex}[1]{\FE@ifstar{#1}{\only@other@multitex}{\only@multitex}}
-\newcommand{\only@multitex}[2]{#2}
-\newcommand{\only@other@multitex}[2]{}
-{\catcode`\#=12
-  \global\def\@pound{#}
-}
-\newcommand{\DeclareFile}[2][]{
-  \global\multitex@names=\expandafter{\the\multitex@names,#2}
-  \expandafter\xdef\csname @multitex@#2\endcsname{\xifempty{#1}{#2}{#1}}
-  \expandafter\xdef\csname @multitex@#2@name\endcsname{"\jobname\multitex@separator\csname @multitex@#2\endcsname.tex"}
-  \expandafter\xdef\csname @multitex@#2@auxname\endcsname{"\jobname\multitex@separator\csname @multitex@#2\endcsname.aux"}
-  \expandafter\xdef\csname @multitex@#2@barename\endcsname{"\jobname\multitex@separator\csname @multitex@#2\endcsname"}
-  \expandafter\newwrite\csname @multitex@#2@file\endcsname
-  \immediate\expandafter\openout\csname @multitex@#2@file\expandafter\endcsname\csname @multitex@#2@name\endcsname
-  \AtBeginDocument{%
-    \multitex@write{#2}{%
-      \detokenize{\newcommand{\onlymultitex}[1]}%
-        {\detokenize{\csname FE@ifstar\endcsname}%
-          {\@pound1}%
-          \detokenize{%
-            {\csname only@multitex\endcsname}{\csname only@multitex\endcsname}%
-          }%
-        }%
-      \detokenize{%
-        \expandafter\newcommand\csname only@multitex\endcsname[2]}%
-        {%
-          \noexpand\renewcommand*{\noexpand\do}[1]{%
-            \noexpand\ifstrequal{\@pound\@pound1}{#2}%
-              {\@pound2\noexpand\listbreak}%
-              {}%
-            }%
-          {\noexpand\edef\noexpand\foo{\@pound1}\noexpand\expandafter}%
-          \noexpand\expandafter\noexpand\docsvlist\noexpand\expandafter{\noexpand\foo}%
-        }
-      \detokenize{%
-        \expandafter\let\csname only@other@multitex\expandafter\endcsname\csname only@multitex\endcsname
-      }%
-    }%
-  }
-  \AtEndDocument{\multitex@postprocess{#2}}
-}
-\newcommand{\multitex@write}[2]{\xifblank{#1}{}{\immediate\expandafter\write\csname @multitex@#1@file\endcsname{#2}}}
-\edef\multitex@tempfilename{\jobname.mtx}
-\newcommand{\multitex@pre@writetothis}{\immediate\openout\multitex@temp\multitex@tempfilename}%\long\gdef\multitex@thiscode{}}%
-\newcommand{\multitex@writetothis}[1]{\immediate\write\multitex@temp{#1}}%\ExpandNext{\long\gappto\multitex@thiscode}{\expandnext{\relax\scantokens}{#1}}}%
-\newcommand{\multitex@post@writetothis}{\immediate\closeout\multitex@temp\relax\input{\multitex@tempfilename}}%\show\multitex@thiscode\multitex@thiscode\relax}%
-\newcommand{\multitex@writetoothers}[2][\the\multitex@names]{{\preetextools\ExpandNext{\forcsvloop}{#1}\do{\multitex@write{##1}{#2}}}}
-\newcommand{\DeclareDocumentClass}[1]{\multitex@writetoothers{\detokenize{#1}}}
-\AtBeginDocument{\multitex@writetoothers{\detokenize{\begin{document}}}}
-\AtEndDocument{\multitex@writetoothers{\detokenize{\end{document}}}}
-\newcommand{\includein}[2][]{\multitex@pre@writetothis\multitex@writetothis{\detokenize{#2}}\multitex@post@writetothis\multitex@writetoothers[#1]{\detokenize{#2}}}
-\AfterEndEnvironment{multitex}{\multitex@post@writetothis}
-\newenvironment{multitex}[1][\the\multitex@names]{%
-  \@bsphack
-  \begingroup% Lets Keep the Changes Local
-    \multitex@pre@writetothis
-    \let\do\@makeother\dospecials\catcode`\^^M\active
-    \def\verbatim@processline{\multitex@writetothis{\the\verbatim@line}\multitex@writetoothers[#1]{\the\verbatim@line}}%
-    \verbatim@start\space % I'm not sure why I need the space, but I seem to need it.
-}{\endgroup
-  \@esphack
-}
-% only others
-\newenvironment{multitex*}[1][\the\multitex@names]{%
-  \@bsphack
-  \begingroup% Lets Keep the Changes Local
-    \let\do\@makeother\dospecials\catcode`\^^M\active
-    \def\verbatim@processline{\multitex@writetoothers[#1]{\the\verbatim@line}}%
-    \verbatim@start\space % I'm not sure why I need the space, but I seem to need it.
-}{\endgroup
-  \@esphack
-}
-\makeatother
-
-\DeclareFile{curriculum-vitae}
-%\DeclareFile{mathematics}
-%\DeclareFile{computer-science}
-%\DeclareFile{curriculum-vitae-long}
-%\DeclareFile{curriculum-vitae-mathematics}
-%\DeclareFile{curriculum-vitae-computer-science}
-\DeclareDocumentClass{\documentclass[11pt]{res}}
-
-\begin{multitex}
-  \edef\resumefiles{mathematics,computer-science}
-  \edef\cvshortspecialfiles{curriculum-vitae-mathematics,curriculum-vitae-computer-science}
-  \edef\cvshortfiles{curriculum-vitae,\cvshortspecialfiles}
-  \edef\cvlongfiles{curriculum-vitae-long}
-  \edef\cvfiles{\cvshortfiles,\cvlongfiles}
-  \edef\highschoolfiles{curriculum-vitae-long}
-  \edef\mathfiles{mathematics,curriculum-vitae-mathematics}
-  \edef\cvlongandmathfiles{\mathfiles,curriculum-vitae-long}
-  \edef\cvandmathfiles{\cvlongandmathfiles,curriculum-vitae,curriculum-vitae-computer-science}
-  \edef\computersciencefiles{computer-science,curriculum-vitae-computer-science}
-  \edef\cvlongandcomputersciencefiles{\computersciencefiles,curriculum-vitae-long}
-  \edef\cvandcomputersciencefiles{curriculum-vitae,curriculum-vitae-mathematics,\cvlongandcomputersciencefiles}
-  \edef\allexceptcv{\resumefiles,\cvshortspecialfiles}
-  \edef\allexceptcvlongfiles{\resumefiles,\cvshortfiles}
-  \edef\allexceptcvandcvlongfiles{\resumefiles,\cvshortspecialfiles}
-  %\usepackage[utf8x]{inputenc}
-  %\usepackage[T1]{fontenc}
-  %\usepackage{helvetica} % uses helvetica postscript font (download helvetica.sty)
-  %\usepackage{newcent}   % uses new century schoolbook postscript font
 \usepackage[T1]{fontenc}
@@
 \usepackage{etoolbox}[2011/01/03]
-  %\bibliographystyle{plainyr-rev}
 \usepackage[backend=biber,maxbibnames=99,sorting=ymdnt,noerroretextools=true,style=alphabetic]{biblatex}
@@
 \newsectionwidth{1em}
+\makeatletter
+\pretocmd{\employer}{\vskip-\baselineskip}{}{}
+\apptocmd{\@@section}{\vskip1.5\baselineskip\relax}{}{}
+\makeatother
 \resumewidth=7.5in
@@
 \newcommand{\aswidthof}[2]{\rlap{#1}\hphantom{#2}}
-\end{multitex}
 
-\begin{multitex}
 \name{\LARGE \bf \parbox{\widthof{github.com/JasonGross}}{%
@@
 \address{\addressaswidth{\textsc{Contact}}\\\addressaswidth{\href{mailto:jgross@mit.edu}{jgross@mit.edu}}\\\addressaswidth{(631) 790-8962}}
-\end{multitex}
 
@@
 
-\begin{multitex}
 \resume
@@
   \item Main Project: Fiat Cryptography (\href{https://github.com/mit-plv/fiat-crypto}{github.com/mit-plv/fiat-crypto})
-      \item Fiat Cryptography is a developer tool to generate proven-correct cryptographic code.
-The project has wide industry adoption, including powering secure connections in Google Chrome and Firefox
+  \item Fiat Cryptography is a developer tool to generate proven-correct cryptographic code. The project has wide industry adoption, including powering secure connections in Google Chrome and Firefox
   \item Lead development of one of the world's first algorithm-level-optimizing compilers
@@
 \item Microsoft Research, summer 2014: Collaboratively created a language for specifying input/output behavior of x86 assembly programs, verfied the input/output behavior of a number of simple programs, and improved performance of the x86proved project
-  \item MIT CSAIL - PLV, 2012 - 2014: Entered a significant amount of category theory into the automated proof assistant Coq, and worked on building an interface for databases and database migration on top of category theory
-  \item MIT CSAIL - CoCoSci, 2009 - 2011: Designed and managed the data collection webpage for research in categorical learning and transfer learning
-  \item Commack High School, 2006 - 2009: Researched circuits over sets of natural numbers, winning 4th (2009) and 3rd (2008) place awards in mathematics at ISEF
+\item MIT CSAIL --- PLV, 2012--2014: Entered a significant amount of category theory into the automated proof assistant Coq, and worked on building an interface for databases and database migration on top of category theory
+\item MIT CSAIL --- CoCoSci, 2009--2011: Designed and managed the data collection webpage for research in categorical learning and transfer learning
+\item Commack High School, 2006--2009: Researched circuits over sets of natural numbers, winning 4th (2009) and 3rd (2008) place awards in mathematics at ISEF
 \end{itemize}
@@
 \item
-    Proficient: Coq, \onlymultitex{\cvlongfiles}{\TeX{} macro language, }Mathematica, git, Python, JavaScript, BASIC
+  Proficient: Coq, %\onlymultitex{\cvlongfiles}{\TeX{} macro language, }
+  Mathematica, git, Python, JavaScript, BASIC
 \item
-    Working knowledge: \onlymultitex{\cvlongfiles}{\LaTeX{}, }C, C++, Agda, OCaml, Haskell, Scheme, HTML, CSS, Perl, Java
+  Working knowledge: %\onlymultitex{\cvlongfiles}{\LaTeX{}, }
+  C, C++, Agda, OCaml, Haskell, Scheme, HTML, CSS, Perl, Java
 \item
@@
 \endresume
-\end{multitex}
 \end{document}

### 2023-03-31 papers/lob-paper/tipa/mf/gentipa.sh
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@a4bd5f2eca34f14478f0fbf045e7c3a26639987b 🚀

@@ -1,91 +1,0 @@
-#!/bin/sh
-
-# gentipa.sh --- a shell script for producing TIPA parameter files.
-# 
-# 1996/06/24 first version
-# 2001/11/07 bs10 (= BXSL10), sb10 (=SSBX10), si10 (=SSSL10) added.
-# 2001/12/21 tt8, tt9, tt10, tt12, ts10 (=SLTT10) added.
-
-AWK=gawk
-
-DATE="2003/01/01"
-VERSION="1.2"
-
-PARAMFILE="tipaprm.def"
-
-FILES="tipa8 tipa9 tipa10 tipa12 tipa17 tipasl8 tipasl9 tipasl10 tipasl12 \
-	tipabx8 tipabx9 tipabx10 tipabx12 tipass8 tipass9 tipass10 tipass12 \
-	tipass17 tipab10 tipabs10 tipasb10 tipasi10 \
-	tipatt8 tipatt9 tipatt10 tipatt12 tipats10"
-
-for FNAME in $FILES
-do
-  case "$FNAME" in
-    tipa8)    NUM=3; FID="TIPA";  FSIZE="8"; FDESC="Roman 8" ;;
-    tipa9)    NUM=4; FID="TIPA";  FSIZE="9"; FDESC="Roman 9" ;;
-    tipa10)   NUM=5; FID="TIPA";  FSIZE="10";FDESC="Roman 10" ;;
-    tipa12)   NUM=6; FID="TIPA";  FSIZE="12";FDESC="Roman 12" ;;
-    tipa17)   NUM=7; FID="TIPA";  FSIZE="17.28";FDESC="Roman 17.28" ;;
-    tipasl8)  NUM=8; FID="TIPASL";FSIZE="8"; FDESC="Slanted Roman 8" ;;
-    tipasl9)  NUM=9; FID="TIPASL";FSIZE="9"; FDESC="Slanted Roman 9" ;;
-    tipasl10) NUM=10;FID="TIPASL";FSIZE="10";FDESC="Slanted Roman 10" ;;
-    tipasl12) NUM=11;FID="TIPASL";FSIZE="12";FDESC="Slanted Roman 12" ;;
-    tipabx8)  NUM=12;FID="TIPABX";FSIZE="8"; FDESC="Bold Extended Roman 8" ;;
-    tipabx9)  NUM=13;FID="TIPABX";FSIZE="9"; FDESC="Bold Extended Roman 9" ;;
[…]

### 2023-03-31 papers/lob-paper/tipa/mf/genxipa.sh

@@ -1,67 +1,0 @@
-#!/bin/sh
-
-# genxipa.sh --- a shell script for producing TIPA parameter files.
-# 
-# 1996/06/24 first version
-#
-
-AWK=gawk
-
-DATE="2003/01/01"
-VERSION="1.2"
-
-PARAMFILE="xipaprm.def"
-
-FILES="xipa10 xipasl10 xipab10 xipabs10 xipass10 xipasi10 xipasb10"
-
-for FNAME in $FILES
-do
-  case "$FNAME" in
-    xipa10)   NUM=3;FID="XIPA";  FSIZE="10";FDESC="Roman 10" ;;
-    xipasl10) NUM=4;FID="XIPASL";FSIZE="10";FDESC="Slanted Roman 10" ;;
-    xipab10)  NUM=5;FID="XIPAB"; FSIZE="10";FDESC="Bold Roman 10" ;;
-    xipabs10) NUM=6;FID="XIPABS";FSIZE="10";FDESC="Bold Slanted Roman 10" ;;
-    xipass10) NUM=7;FID="XIPASS";FSIZE="10";FDESC="Sans Serif 10" ;;
-    xipasi10) NUM=8;FID="XIPASI";FSIZE="10";FDESC="Sans Serif Slanted 10" ;;
-    xipasb10) NUM=9;FID="XIPASB";FSIZE="10";FDESC="Sans Serif Bold 10" ;;
-  esac
-
-  echo "% $FNAME.mf: XIPA $FDESC point parameter file" > $FNAME.mf
-  echo "% Copyright 1996-2003 FUKUI Rei" >> $FNAME.mf
-  echo "%" >> $FNAME.mf
-  echo "% This program may be distributed and/or modified under the" >> $FNAME.mf
-  echo "% conditions of the LaTeX Project Public License, either version 1.2" >> $FNAME.mf
-  echo "% of this license or (at your option) any later version." >> $FNAME.mf
-  echo "% The latest version of this license is in" >> $FNAME.mf
-  echo "%   http://www.latex-project.org/lppl.txt" >> $FNAME.mf
-  echo "% and version 1.2 or later is part of all distributions of LaTeX " >> $FNAME.mf
-  echo "% version 1999/12/01 or later." >> $FNAME.mf
-  echo "%" >> $FNAME.mf
[…]

### 2023-09-15 resume/Resume.tex
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@7403e26c3bbc263873a1f2225d67d0b3e7be99bf 🚀
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@ecc266e08e824c5124385cbb8bc7f31f866fa001 🚀
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@7d29e351f5536af50d1566b3d2295e629d222916 🚀

@@
 \usepackage{etoolbox}[2011/01/03]
-\usepackage[backend=biber,maxbibnames=99,sorting=ymdnt,noerroretextools=true,style=alphabetic]{biblatex}
+\expandafter\def\csname blx@noerroretextools\endcsname{}
+\usepackage[backend=biber,maxbibnames=99,sorting=ymdnt,style=alphabetic]{biblatex}
 % http://tex.stackexchange.com/a/140641
@@
 \makeatletter
-\pretocmd{\employer}{\vskip-\baselineskip}{}{}
-\apptocmd{\@@section}{\vskip1.5\baselineskip\relax}{}{}
+\def\firstemployerskip{\vskip-\baselineskip\def\employerskip{}}
+\pretocmd{\employer}{\employerskip}{}{}
+\apptocmd{\@@section}{\vskip1.5\baselineskip\relax\def\employerskip{\firstemployerskip}}{}{}
 \makeatother
@@
 
-\name{\LARGE \bf \parbox{\widthof{github.com/JasonGross}}{%
-    \centering Jason Gross \\
-    \vspace{0.2cm}
-    \large \href{https://github.com/JasonGross/}{github.com/JasonGross} \\
-    \href{https://people.csail.mit.edu/jgross/}{people.csail.mit.edu/jgross} \\
-    $\left.\right.$ \\
-    $\left.\right.$ \\%
-}}
+\name{\LARGE Jason Gross\vspace{0.75cm}}
+%\parbox{\textwidth}{%\widthof{github.com/JasonGross}}{%
+%    {\LARGE \centering \textbf{Jason Gross}} \\
+%    \vspace{0.2cm}
+%    \large
+%    \href{https://github.com/JasonGross/}{github.com/JasonGross} \\
+%    \href{https://people.csail.mit.edu/jgross/}{people.csail.mit.edu/jgross} \\
+%    $\left.\right.$ \\
+%    $\left.\right.$ \\%
+%}}
 
-\vspace{5cm}
-
+\address{\parbox{\textwidth}{%
+  \href{mailto:jgross@mit.edu}{\large Email: \normalsize jgross@mit.edu}\hfill
+  \href{https://people.csail.mit.edu/jgross/}{\large Website: \normalsize people.csail.mit.edu/jgross} \\
+  \href{https://github.com/JasonGross/}{\large GitHub: \normalsize JasonGross}\hfill
+  \href{https://scholar.google.com/citations?user=QouPlrMAAAAJ&hl=en}{\large Google Scholar: \normalsize QouPlrMAAAAJ}%
+}%
+}
 %\address{}%\textsc{Address}\\258 Prospect Street, Apt \# 1L\\Cambridge, MA 02139}
 %\address{\textsc{Contact}\\\href{mailto:jgross@mit.edu}{jgross@mit.edu}\\(631) 790-8962}
-\newcommand{\addressaswidth}[1]{\makebox[\widthof{\textsc{jgross@mit.edu}}][c]{#1}}
-\address{\addressaswidth{\textsc{Contact}}\\\addressaswidth{\href{mailto:jgross@mit.edu}{jgross@mit.edu}}\\\addressaswidth{(631) 790-8962}}
+%\newcommand{\addressaswidth}[1]{\makebox[\widthof{\textsc{jgross@mit.edu}}][c]{#1}}
+%\address{\addressaswidth{\textsc{Contact}}\\\addressaswidth{\href{mailto:jgross@mit.edu}{jgross@mit.edu}}\\\addressaswidth{(631) 790-8962}}
 
@@
 \begin{itemize}
-\item Programming Languages and AI: Löb's Theorem, Type Theory, Compact Proof Generation
-\item Trust and Security Engineering: Formal Verification, Cryptography, Performance Engineering
+\item AI: Mechanistic Interpretability, Heuristic Arguments, Compact Proofs, Löb's Theorem
+\item Security \& Engineering: Formal Verification, Cryptography, Bugs Infrastructure
+\item Programming Languages: Type Theory, Performance, Proof Engines \& Automation
+%
+%\item Mechanistic Interpretability: Heuristic Arguments, Compact Proofs, Type Theory
+%\item Security Engineering: Formal Verification, Cryptography, Performance Engineering
+%\item Programming Languages and AI: Löb's Theorem, Type Theory, Compact Proof Generation
+%%\item Trust and Security Engineering: Formal Verification, Cryptography, Performance Engineering
 \end{itemize}
@@
 
-\employer{Machine Intelligence Research Institute (MIRI)}
+\employer{Special Project of ARC Theory}
+\dates{August 2023--Present}
+\title{Project Lead}
+\location{Berkeley, CA}
+\begin{position}
+  \begin{itemize}
+  \item Building the first machine-checked functional-correctness proofs of mechanistic interpretability arguments about transformers in Coq (\href{https://github.com/JasonGross/neural-net-coq-interp}{\texttt{github.com/JasonGross/neural-net-coq-interp}})
+  \item Raised \$150k and leading an interdisciplinary team of ML researchers and mathematicians
+  \end{itemize}
+\end{position}
+
+\employer{\mbox{Machine Intelligence Research Institute (MIRI)}}
 \dates{February 2021--Present}
@@
   \begin{itemize}
-  \item Performing self-directed research into topics in fundamental programming language theory and mathematics%, bringing insights and understanding back to my supervisor
+  \item Performing self-directed research into fundamental programming language theory and math%ematics%, bringing insights and understanding back to my supervisor
   \end{itemize}
@@
 \dates{June 2021--Present}
-\title{Core Team Member}
+\title{Member of Core Team}
 \location{Nantes, France (remote)}
@@
   \item Reported the plurality of all-time bugs in Coq (since 2012)
-  \item Designed and engineered a bug minimizer for automatically producing minimized stand-alone test-cases from bug reports with automatic support for minimizing regressions in external projects tested by Coq CI
+  \item Designed and engineered a bug report minimizer for automatically producing minimized standalone test-cases and minimizing regressions in external projects tested by Coq's CI
   \item Researching performance issues that impact scalability of automated verification
@@
   \begin{itemize}
-  \item Main Project: Fiat Cryptography (\href{https://github.com/mit-plv/fiat-crypto}{github.com/mit-plv/fiat-crypto})
-  \item Fiat Cryptography is a developer tool to generate proven-correct cryptographic code. The project has wide industry adoption, including powering secure connections in Google Chrome and Firefox
+  \item Main Project: Fiat Cryptography (\href{https://github.com/mit-plv/fiat-crypto}{\texttt{github.com/mit-plv/fiat-crypto}})
+  \item Fiat Cryptography is a developer tool to generate proven-correct cryptographic code, with wide industry adoption, powering the plurality of secure connections in Chrome and Firefox
   \item Lead development of one of the world's first algorithm-level-optimizing compilers
@@
 \item
-  Proficient: Coq, %\onlymultitex{\cvlongfiles}{\TeX{} macro language, }
-  Mathematica, git, Python, JavaScript, BASIC
+  Proficient: Coq, Agda, Python, %\onlymultitex{\cvlongfiles}{\TeX{} macro language, }
+  Mathematica, BASIC, \TeX{} macro language, git, JavaScript
 \item
   Working knowledge: %\onlymultitex{\cvlongfiles}{\LaTeX{}, }
-  C, C++, Agda, OCaml, Haskell, Scheme, HTML, CSS, Perl, Java
+  C, C++, OCaml, Haskell, Scheme%/Racket
+  , HTML, CSS, Perl, Bash, Java
 \item
-  Basic knowledge: Matlab, Lean, Idris, Ruby, Go, Ur/Web, x86 Assembly
+  Basic knowledge: MATLAB, Lean, Idris, Ruby, Go, Ur/Web, x86 Assembly, SQL, Batch
 \end{itemize}

### 2023-10-03 resume/Resume.tex
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@fbdaea4438b1db099a79d1cdf1c0c4e92f1b7fdd 🚀

@@
 
-\section{\textsc{Research Interests}}
-\begin{itemize}
-\item AI: Mechanistic Interpretability, Heuristic Arguments, Compact Proofs, Löb's Theorem
-\item Security \& Engineering: Formal Verification, Cryptography, Bugs Infrastructure
-\item Programming Languages: Type Theory, Performance, Proof Engines \& Automation
-%
-%\item Mechanistic Interpretability: Heuristic Arguments, Compact Proofs, Type Theory
-%\item Security Engineering: Formal Verification, Cryptography, Performance Engineering
-%\item Programming Languages and AI: Löb's Theorem, Type Theory, Compact Proof Generation
-%%\item Trust and Security Engineering: Formal Verification, Cryptography, Performance Engineering
-\end{itemize}
+%\section{\textsc{Research Interests}}
+%\begin{itemize}
+%\item AI: Mechanistic Interpretability, Heuristic Arguments, Compact Proofs, Löb's Theorem
+%\item Security \& Engineering: Formal Verification, Cryptography, Bugs \& CI Infrastructure
+%\item Programming Languages: Type Theory, Performance, Proof Engines \& Automation
+%%
+%%\item Mechanistic Interpretability: Heuristic Arguments, Compact Proofs, Type Theory
+%%\item Security Engineering: Formal Verification, Cryptography, Performance Engineering
+%%\item Programming Languages and AI: Löb's Theorem, Type Theory, Compact Proof Generation
+%%%\item Trust and Security Engineering: Formal Verification, Cryptography, Performance Engineering
+%\end{itemize}
+\section{\textsc{About Me}}
+\vspace*{-0.25cm}
+I'm a programming languages research scientist transitioning into ML and alignment.
+I have a working knowledge of around two dozen programming languages, and expertise in a handful (Coq ($\approx$1M+ loc), Python ($\approx$50k loc), Agda ($\approx$50k loc), others).
+I spent my PhD on low-level cryptographic code generation, proof automation, performance engineering, and infrastructure around debugging and CI.
+My current project is in mechanistic interpretability, heuristic arguments, and compact proofs.
+I'm excited about what scalable performant automation can make possible.
 
+% TODO: https://github.com/avgupta456/github-trends
 \section{\textsc{Experience}}

### 2023-10-03 resume/Resume.tex
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@ce1015d2bef32d302bda5e7484df4ff634635750 🚀

@@
 \vspace*{-0.25cm}
-I'm a programming languages research scientist transitioning into ML and alignment.
-I have a working knowledge of around two dozen programming languages, and expertise in a handful (Coq ($\approx$1M+ loc), Python ($\approx$50k loc), Agda ($\approx$50k loc), others).
-I spent my PhD on low-level cryptographic code generation, proof automation, performance engineering, and infrastructure around debugging and CI.
-My current project is in mechanistic interpretability, heuristic arguments, and compact proofs.
-I'm excited about what scalable performant automation can make possible.
+Research scientist with extensive experience in proofs and formal verification transitioning to AI \& ML.
+Currently researching mechanistic interpretability, heuristic arguments, and compact proofs.
+Programming language polyglot, with working knowledge of around two dozen programming languages, and expertise in a handful (Coq (1M+ loc), Python (40k+ loc), Agda (40k+ loc), others).
+I spent the years of my PhD on low-level cryptographic code generation, automation, performance engineering, and infrastructure around debugging and CI.
 

### 2024-08-13 resume/Resume.tex
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@3f47a114b1f4a4215d649df0dfdf1074921ed186 🚀
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@b0a221edd069d45e659ded594349a8e2a9a4df17 🚀
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@f514aa8cdd283c872edc741b5d54be969118930c 🚀
Deploying to gh-pages from @ JasonGross/JasonGross.github.io@f133da88691b0bb5f0833d3abdde9d071cc038aa 🚀

--- 2023-10-03 resume/Resume.tex
@@
 \usepackage{etoolbox}[2011/01/03]
+\let\origifempty\ifempty
+\let\origforlistloop\forlistloop
+\undef\ifempty
+\usepackage{etextools}
+\let\eifempty\ifempty
+\let\ifempty\origifempty
+\let\forlistloop\origforlistloop
 \expandafter\def\csname blx@noerroretextools\endcsname{}
@@
 \DeclareBibliographyCategory{exclude}
-\makeatletter
-\@ifpackageloaded{etextools}{%
-}{%
-  \let\origifempty\ifempty
-  \let\origforlistloop\forlistloop
-  \undef\ifempty
-  \usepackage{etextools}
-  \let\eifempty\ifempty
-  \let\ifempty\origifempty
-  \let\forlistloop\origforlistloop
-  \def\preetextools{\let\ifempty\eifempty}
-}%
-\makeatother
 \begin{format}
@@
 I'm a programming languages research scientist transitioning into ML and alignment.
-I have a working knowledge of around two dozen programming languages, and expertise in a handful (Coq ($\approx$1M+ loc), Python ($\approx$50k loc), Agda ($\approx$50k loc), others).
+I have a working knowledge of around two dozen programming languages, and expertise in a handful (Coq ($\approx$1M+ loc), Python ($\approx$80k loc), Agda ($\approx$50k loc), others).
 I spent my PhD on low-level cryptographic code generation, proof automation, performance engineering, and infrastructure around debugging and CI.
@@
   \item Building the first machine-checked functional-correctness proofs of mechanistic interpretability arguments about transformers in Coq (\href{https://github.com/JasonGross/neural-net-coq-interp}{\texttt{github.com/JasonGross/neural-net-coq-interp}})
-  \item Raised \$150k and leading an interdisciplinary team of ML researchers and mathematicians
+  \item Raised \$150k \& leading an interdisciplinary team of eight ML researchers and mathematicians
   \end{itemize}
@@
 \employer{\mbox{Machine Intelligence Research Institute (MIRI)}}
-\dates{February 2021--Present}
+\dates{February 2021--September 2023}
 \title{Research Staff}
@@
 \addtocategory{exclude}{Evaluation2021Huhmann}
-\printbibliography[title={$\left.\right.$},notcategory=exclude]
+\printbibliography[title={\ensuremath{\left.\right.}},notcategory=exclude]
 

### 2024-11-16 misc/trash.tex
Add some generation files

\documentclass{article}
\usepackage[margin=0.5in]{geometry}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{nopageno}
\usepackage{svg}

\begin{document}
\pagestyle{empty}
% https://qci.qr-code.click/uploads/qr_codes/673826e31ca41.svg?1731733480
% https://api.qrserver.com/v1/create-qr-code/?size=150x150&data=https://jasongross.github.io/misc/trash.html

\begin{multicols}{2}
    \centering
    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash.svg}
    \vspace{0.5cm}

    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash.svg}
    \vspace{0.5cm}

    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash.svg}
    \vspace{0.5cm}

    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash.svg}
    \vspace{0.5cm}

    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash.svg}
    \vspace{0.5cm}

    \includesvg[width=\linewidth, height=.3\textheight, keepaspectratio]{qr-trash.svg}
\end{multicols}

\end{document}
